### 2020-03-26 paper/ms.tex
init paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt,twocolumn,tighten]{aastex62}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}

\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{\today}
\revised{---}
\accepted{---}
\submitjournal{AAS journals.}
\shorttitle{Against PTFO$\,$8-8695b}

\begin{document}

\defcitealias{bouma_wasp4b_2019}{B19}

\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
% ", Explaining its Decreasing Orbital Period"
% ": A Cautionary Tale in the Search for Tidal Orbital Decay"
% "An Explanation for WASP-4's Decreasing Orbital Period".
% "WASP-4's Decreasing Orbital Period: Solved."

\correspondingauthor{L. G. Bouma}
\email{luke@astro.princeton.edu}

%
% key authors:
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}

%
% contributing authors: alphabetical
%
% \author[0000-0001-8638-0320]{A. W. Howard}
% \affiliation{Cahill Center for Astrophysics, California Institute of
% Technology, Pasadena, CA 91125, USA}
% %
% \author[0000-0002-2532-2853]{S. B. Howell}
% \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
% %
% \author[0000-0002-0531-1073]{H. Isaacson}
% \affiliation{Astronomy Department, University of California, Berkeley,
% CA 94720, USA}
% %
% \author{H. Knutson}
% \affiliation{Division of Geological and Planetary Sciences, California
% Institute of Technology, Pasadena, CA 91125, USA}
% %
% \author[0000-0001-7233-7508]{R. A. Matson}
% \affiliation{U.S. Naval Observatory, Washington, DC 20392, USA}
% %

\begin{abstract}
  PTFO$\,$8-8695b could be the youngest, shortest-period 
  hot Jupiter known.  However it has not been shown to be a planet.
  TESS recently observed PTFO$\,$8-8695 for one month.
  The TESS light-curve shows that the dominant variability in this
  system is a sinusoidal modulation with period 11.96 hours, likely
  caused by stellar rotation.
  Also present is a complex signal, previously identified as the
  planet candidate, that repeats every 10.74 hours.
  The ``rotation'' and ``orbit'' have different frequencies,
  and show the expected beat every 4.48 days.
  There is a dip in the complex, shorter-period signal.
  However the orbital phase of the dip has changed over the past
  decade, in agreement with recent observations by Tanimoto et al.
  Planets do not change their orbital phase.
  PTFO$\,$8-8695 therefore seems broadly consistent with the
  ``transient dipping'' phenomenology observed in many young M dwarfs,
  and seems unlikely to be a planet.
\end{abstract}

%TODO
\keywords{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
If it were a planet, PTFO$\,$8-8695b would be exceptional.
A transiting hot Jupiter, orbiting a $\approx$3$\,$Myr old M dwarf
in ORION would make it the youngest hot Jupiter known.
Its orbital period of only 12 HOURS would also make it the shortest
period hot Jupiter known.

% Section~\ref{sec:observations} of this paper presents all of the
% available transit data as well as the new radial velocity and speckle
% imaging observations.  Section~\ref{sec:analysis} describes our
% analysis of the data, and our interpretation that WASP-4 is being
% pulled around by a brown dwarf or low-mass star.
% Section~\ref{sec:discussion} places this result within the context of
% orbital decay searches, and points out that line-of-sight
% accelerations will be a relatively common type of ``false positive.''
% Section~\ref{sec:conclusions} offers concluding remarks.

\section{Observations}
\label{sec:observations}

% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=0.7\textwidth]{f1.pdf}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
% 		The vertical axis shows the observed transit times minus the
% 		calculated times, assuming a constant orbital period.  More opaque
% 		points correspond to more precise data.  The $\pm1\sigma$
% 		uncertainties of the quadratic ephemeris are shown in blue.  The
% 		yellow star represents the weighted average of 18 data points
% 		obtained with TESS. The TESS data were averaged here for display
% 		purposes only; our analysis used the 18 individual transit times.
% 		\label{fig:times}
% 	}
% \end{figure*}

\section{Analysis}
\label{sec:analysis}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:lc_map}
	}
\end{figure*}

\begin{figure*}[t]
	% \begin{center}
	% 	\leavevmode
	% 	\includegraphics[width=0.95\textwidth]{f1.png}
	% \end{center}
  \gridline{
    \fig{f2a.png}{0.8\textwidth}{}
    \fig{f2b.png}{0.2\textwidth}{}
  }
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:splitsignal}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f3.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:phasefold}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf foo.}
    bar
		\label{fig:corner}
	}
\end{figure*}

\subsection{Visual Inspection}

% TODO:
% Figure: CDIPS LC + SPOC LC (left two panels). Right panel:
% periodogram.
%

\subsection{Model Fitting}

% I fitted some models of the form "transit + N harmonics of sines and
% cosines at the orbital frequency + M harmonics of sines and cosines at
% the rotation frequency". For example, if N=2 and M=1, then I fit for
% [logBrot0, logArot0, phirot, omegarot, logBorb1, logAorb1, logBorb0,
% logAorb0, b, r, u, logP, t0, meanflux]. A_N is the amplitude of the
% Nth harmonic sine function A_N * sin(N(ωt+φ)), B_N of the cosine. I
% went with quadratic limb darkening.
% 
% On fake data, I got the right parameters provided I didn't turn up the
% noise too high.
% 
% I gave it a pass on the PTFO data, binning from 2->10 minute sampling
% to speed up the execution time. Results from two models are below.
% Case 1 is N=2,M=1. Case 2 is N=2,M=2. Both fits are fine, and I get
% decent convergence with a few minutes each. The rotation parameters
% (phase + amplitudes) are quite correlated, so I set a reasonably
% strong prior on the rotation phase.
% 
% The main result that both models agree on: the extra power at the
% orbital frequency seems to not be ellipsoidal. Both of these models
% prefer to put the power at 1x the orbital frequency, not 2x the
% orbital frequency! This gives both Aorb0 and Borb0 non-zero, while
% Aorb1 and Borb1 seem to be more consistent with zero.

\section{Discussion}
\label{sec:discussion}

\section{Conclusions}
\label{sec:conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \acknowledgements
% %
% This paper includes data collected by the TESS mission, which are
% publicly available from the Mikulski Archive for Space Telescopes
% (MAST).
% %
% Funding for the TESS mission is provided by NASA's Science Mission
% directorate.
% %
% This work made use of NASA's Astrophysics Data System Bibliographic
% Services.
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  % \texttt{astroplan} \citep{astroplan2018},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  % \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{radvel} \citep{fulton_radvel_2018},
  % \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}

% \facilities{
% 	{\it Astrometry}:
% 	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}.
% 	{\it Imaging}:
% 	Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
% 	{\it Spectroscopy}:
% 	Keck:I~(HIRES; \citealt{vogt_hires_1994}),
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
% 	{\it Photometry}:
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
% 	El Sauce:0.356m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
% 	NTT,
% 	SOAR (SOI),
% 	TESS \citep{ricker_transiting_2015},
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
% }

%
% The following are entries from Table 1 that are not otherwise cited
% in the text
%
% \nocite{wilson_wasp-4b_2008}
% \nocite{gillon_improved_2009}
% \nocite{winn_transit_2009}
% \nocite{hoyer_tramos_2013}
% \nocite{dragomir_terms_2011}
% \nocite{sanchis-ojeda_starspots_2011}
% \nocite{nikolov_wasp-4b_2012}
% \nocite{ranjan_atmospheric_2014}
% \nocite{huitson_gemini_2017}

% \input{WASP-4b_transit_time_table.tex}
% \input{WASP-4b_rv_table.tex}
% \input{model_fit_table.tex}
% \input{rv_model_posterior_table.tex}
% \input{pdot_table.tex}

%\clearpage
\bibliographystyle{yahapj}                            
\bibliography{bibliography}

\listofchanges

\end{document}

### 2020-03-26 paper/collect.sh

@@ -1,0 +1,10 @@
+# Usage: copy in figures for the paper
+
+pdir=../results/PTFO_8-8695_results/20200320_linear_amplitude_prior/
+
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_sampleplot.png f1.png
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap.png f2a.png
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_periodogram.png f2b.png
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.png f3.png
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f4.png
+

### 2020-03-26 billy/plotting.py
add figures and paper pdf
... and compressed the corner plot

@@ -460,7 +460,7 @@ def plot_cornerplot(true_d, m, outpath):
     fig = corner.corner(trace_df, quantiles=[0.16, 0.5, 0.84],
                         show_titles=True, title_kwargs={"fontsize": 12},
                         truths=truths, title_fmt='.2g')
-    savefig(fig, outpath, writepdf=0)
+    savefig(fig, outpath, writepdf=0, dpi=100)
 
 
 def savefig(fig, figpath, writepdf=True, dpi=450):

### 2020-03-26 drivers/real.py

@@ -16,8 +16,8 @@ def main(modelid):
 
     traceplot = 0
     sampleplot = 0
-    cornerplot = 0
-    splitsignalplot = 1 if 'Porb' in modelid and 'Prot' in modelid else 0
+    cornerplot = 1
+    splitsignalplot = 0 if 'Porb' in modelid and 'Prot' in modelid else 0
 
     REALID = 'PTFO_8-8695'
     RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')

### 2020-03-26 paper/collect.sh

@@ -7,4 +7,4 @@ cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap.png f2a.png
 cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_periodogram.png f2b.png
 cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.png f3.png
 cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f4.png
-
+pngquant f4.png --output f4_comp.png

### 2020-03-28 billy/plotting.py
add sick plot_scene

@@ -1,5 +1,6 @@
 """
 Plots:
+
     plot_periodogram
     plot_test_data
     plot_MAP_data
@@ -11,6 +12,8 @@ Plots:
     plot_traceplot
     plot_cornerplot
 
+    plot_scene
+
 Convenience:
     savefig
     format_ax
@@ -19,6 +22,7 @@ import os, corner
 import numpy as np, matplotlib.pyplot as plt
 from datetime import datetime
 from pymc3.backends.tracetab import trace_to_dataframe
+from itertools import product
 
 from billy.convenience import flatten as bflatten
 from billy.convenience import get_clean_ptfo_data
@@ -484,3 +488,183 @@ def format_ax(ax):
         tick.label.set_fontsize('small')
     for tick in ax.yaxis.get_major_ticks():
         tick.label.set_fontsize('small')
+
+
+def plot_scene(c_obj, img_wcs, img, outpath, Tmag_cutoff=17, showcolorbar=0,
+               ap_mask=0, bkgd_mask=0):
+
+    from astrobase.plotbase import skyview_stamp
+    from astropy import units as u, constants as const
+    from astropy.wcs import WCS
+    from astroquery.mast import Catalogs
+    import astropy.visualization as vis
+    import matplotlib as mpl
+    from matplotlib import patches
+
+    # standard tick formatting fails for these images.
+    mpl.rcParams['xtick.direction'] = 'in'
+    mpl.rcParams['ytick.direction'] = 'in'
+
+    #
+    # wcs information parsing
+    # follow Clara Brasseur's https://github.com/ceb8/tessworkshop_wcs_hack
+    # (this is from the CDIPS vetting reports...)
+    #
+    radius = 5.0*u.arcminute
+
+    nbhr_stars = Catalogs.query_region(
+        "{} {}".format(float(c_obj.ra.value), float(c_obj.dec.value)),
+        catalog="TIC",
+        radius=radius
+    )
+
+    try:
+        px,py = img_wcs.all_world2pix(
+            nbhr_stars[nbhr_stars['Tmag'] < Tmag_cutoff]['ra'],
+            nbhr_stars[nbhr_stars['Tmag'] < Tmag_cutoff]['dec'],
+            0
+        )
+    except Exception as e:
+        print('ERR! wcs all_world2pix got {}'.format(repr(e)))
+        raise(e)
+
+    ticids = nbhr_stars[nbhr_stars['Tmag'] < Tmag_cutoff]['ID']
[…]

### 2020-03-28 drivers/plot_scene.py

@@ -1,0 +1,43 @@
+"""
+TESS image (x's numbered)
+DSS2 Red image.
+"""
+
+import numpy as np
+
+from astropy.io import fits
+from astropy import wcs
+from astropy.wcs import WCS
+from astropy.coordinates import SkyCoord
+import astropy.units as u
+
+from billy.plotting import plot_scene
+
+def main():
+    # J2015.5 gaia
+    ra = 81.28149227428
+    dec = 1.57343091124
+
+    c_obj = SkyCoord(ra, dec, unit=(u.deg), frame='icrs')
+
+    hdul = fits.open(
+        '../data/PTFO_8-8695/'
+        'tess2018349182459-s0006-0000000264461976-0126-s/'
+        'tess2018349182459-s0006-0000000264461976-0126-s_tp.fits'
+    )
+
+    img_wcs = wcs.WCS(hdul[2].header)
+
+    d = hdul[1].data
+    mean_img = np.nansum(d["FLUX"], axis=0) / d["FLUX"].shape[0]
+
+    bkgd_mask = (hdul[2].data == 37)
+    ap_mask = (hdul[2].data == 43)
+
+    outpath = '../results/PTFO_8-8695_results/scene.png'
+    plot_scene(c_obj, img_wcs, mean_img, outpath, ap_mask=ap_mask,
+               bkgd_mask=bkgd_mask)
+
+
+if __name__ == "__main__":
+    main()

### 2020-03-28 paper/ms.tex
add figures and paper pdf
... and compressed the corner plot
add works for blend analysis

@@
 \title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
-% ", Explaining its Decreasing Orbital Period"
-% ": A Cautionary Tale in the Search for Tidal Orbital Decay"
-% "An Explanation for WASP-4's Decreasing Orbital Period".
-% "WASP-4's Decreasing Orbital Period: Solved."
 
@@
   The TESS light-curve shows that the dominant variability in this
-  system is a sinusoidal modulation with period 11.96 hours, likely
-  caused by stellar rotation.
+  system is a sinusoidal modulation with a ``long'' period $P_{\rm
+  \ell}$
+  of 11.96 hours, likely caused by stellar rotation.
   Also present is a complex signal, previously identified as the
-  planet candidate, that repeats every 10.74 hours.
-  The ``rotation'' and ``orbit'' have different frequencies,
-  and show the expected beat every 4.48 days.
-  There is a dip in the complex, shorter-period signal.
-  However the orbital phase of the dip has changed over the past
-  decade, in agreement with recent observations by Tanimoto et al.
-  Planets do not change their orbital phase.
+  planet candidate, that repeats with a ``short'' period $P_{\rm s}$ of
+  10.74 hours.
+  The ``long'' and ``short'' signals show the expected beat every 4.48
+  days.
+  There is a dip in the complex, short-period signal.
+  However ground-based photometry from the past decade shows that the
+  orbital phase of the dip seems to have instantaneously jumped, at
+  least twice.
+  The TESS epoch of the dip is consistent with recent observations by
+  Tanimoto et al., and differs from the discovery epoch by 5.14 hours.
+  Planets do not ``jump'' in orbital phase.
   PTFO$\,$8-8695 therefore seems broadly consistent with the
@@
 
-% \begin{figure*}[t]
-% 	\begin{center}
-% 		\leavevmode
-% 		\includegraphics[width=0.7\textwidth]{f1.pdf}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-% 	\caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
-% 		The vertical axis shows the observed transit times minus the
-% 		calculated times, assuming a constant orbital period.  More opaque
-% 		points correspond to more precise data.  The $\pm1\sigma$
-% 		uncertainties of the quadratic ephemeris are shown in blue.  The
-% 		yellow star represents the weighted average of 18 data points
-% 		obtained with TESS. The TESS data were averaged here for display
-% 		purposes only; our analysis used the 18 individual transit times.
-% 		\label{fig:times}
-% 	}
-% \end{figure*}
+\subsection{Blend considerations}
+\label{subsec:blend}
+
+The TESS pixels are $\approx21$'' per side, and so we need to consider
+whether light from neighboring stars could affect the photometry.  The
+scene is shown in Figure~\ref{fig:scene}.  
+The pixels used to
+measure the background level are indicated with an `\texttt{X}' hatch,
+and the pixels used for the final light-curve are shown with the
+`\texttt{/}' hatch.
+
+The target star, PTFO
+8-8605 (TIC 264461976), has a $T$-band magnitude of 14.0, and its position is shown with a
+star.  
+The other (unlabeled) star inside the target aperture, TIC 264461979, has $T=16.8$ and so cannot
+contribute a signal with relative amplitude 10\%.
+The only neighbor that is sufficiently close and bright that
+its light might contaminate the target star is TIC 264461980, with
+$T=14.8$, which we denote ``Star A''.  Star A is 23.6'' NW of our
+target, and based on the magnitude difference could contribute up to
+48\% the flux of our target star, PTFO 8-8695.
+
+Because PTFO 8-8695 has previously been identified to have periodicity
+consistent with our measurement of $P_{\rm s}$, our main concern
+regarding blending is the degree to which we can be certain that the
+long-period signal at $P_{\rm \ell}$ also originates from PTFO 8-8695.
+We took two approaches towards determining the source of the long-period signal.
+
+First, we examined the CDIPS full frame image light-curves of the
+target, which are available on MAST \citep{bouma_cluster_2019}.
+The maximal peak-to-peak beat amplitude is consistently $\approx$10\%
+across apertures of radii 1, 1.5, and 2.25 pixels.
+If Star A were the source of the long-period variability, we would expect the
+peak variability amplitude to be smallest in the 1 pixel aperture, based on the
+separation of the sources (Figure~\ref{fig:scene}, bottom).
+From this test alone, it seems unlikely that Star A is the source of
+the long-period signal.
+
+Second, we examined the light-curve of each pixel in the scene
+individually.  We opted to use the
+interactive tools implemented in
+\texttt{lightkurve} \citep{lightkurve_2018}.  If Star A were the
+source of the long-period variability, we would expect the pixels
+nearest to Star A to show a sinusoidal signal with
+amplitude exceeding $10\%$.  We find no evidence for
+this being the case.  The pixel directly below Star A does not
+clearly show the sinusoidal variability, and the peak-to-peak 
+variability in that pixel is $\lesssim 8\%$.  In contrast, the
+south-easternmost pixel within PTFO 8-8695's aperture (the pixel 
+furthest from Star A that was used in the optimal aperture) shows the $P_{\rm \ell}$ sinusoidal
+variability signal at $\approx 10\%$ amplitude.
+
+As there is no evidence in favor of a blend scenario, we
+conclude that both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from PTFO 8-8695.
 
@@
 
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f1.png}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf foo.}
-    bar
-		\label{fig:lc_map}
-	}
-\end{figure*}
+%\begin{figure*}[t]
+%	\begin{center}
+%		\leavevmode
+%		\includegraphics[width=0.95\textwidth]{f2.png}
+%	\end{center}
+%	\vspace{-0.7cm}
+%	\caption{ {\bf foo.}
+%    bar
+%		\label{fig:lc_map}
+%	}
+%\end{figure*}
 
 \begin{figure*}[t]
-	% \begin{center}
-	% 	\leavevmode
-	% 	\includegraphics[width=0.95\textwidth]{f1.png}
-	% \end{center}
   \gridline{
-    \fig{f2a.png}{0.8\textwidth}{}
-    \fig{f2b.png}{0.2\textwidth}{}
+    \fig{f3a.png}{0.8\textwidth}{}
+    \fig{f3b.png}{0.2\textwidth}{}
   }
@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f3.png}
+		\includegraphics[width=0.8\textwidth]{f4.png}
 	\end{center}
@@
 
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.48\textwidth]{f1.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Scene used for blend analysis.}
+		{\it Top:} Mean TESS image of PTFO 8-8695 over Sector~6, with a log-stretch.
+		The position of PTFO 8-8695 is shown with a yellow star.
+		Neighbors with $T<17$ are shown with red crosses.
+		The apertures used to measure the background and target star flux
+		are shown with \texttt{X} and \texttt{/} hatches, respectively.
+		{\it Bottom:} DDS2 $R$-band image of the same field, with a linear stretch. The circles show
+		apertures of radii 1, 1.5, and 2.25 pixels used in part of our blend
+		analysis.
+		The pixel level TESS data show that ``Star A''  does not contribute variability at either of the two observed periods (see Section~\ref{subsec:blend}).
+		\label{fig:scene}
+	}
+\end{figure}
+
 \begin{figure*}[t]
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f4.png}
+		\includegraphics[width=0.9\textwidth]{o_minus_c.png}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf foo.}
+		bar
+		\label{fig:o_minus_c}
+	}
+\end{figure*}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f5_comp.png}
 	\end{center}
@@
   \texttt{IPython} \citep{perez_2007},
+	\texttt{lightkurve} \citep{lightkurve_2018},
   \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 

### 2020-03-28 paper/collect.sh
update paper

@@ -2,9 +2,15 @@
 
 pdir=../results/PTFO_8-8695_results/20200320_linear_amplitude_prior/
 
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_sampleplot.png f1.png
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap.png f2a.png
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_periodogram.png f2b.png
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.png f3.png
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f4.png
-pngquant f4.png --output f4_comp.png
+#cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_joined.pdf f1.pdf
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap.png f1.png
+
+#cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.pdf f2.pdf
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.png f2.png
+
+cp ../results/PTFO_8-8695_results/scene.pdf f3.pdf
+
+cp ../results/PTFO_8-8695_literature_and_TESS_times_O-C_vs_epoch_all_tanimoto2020_ephem.png f4.png
+
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f5.png
+pngquant f5.png --output f5_comp.png

### 2020-03-28 billy/plotting.py
made plots even cooler

@@ -126,12 +126,14 @@ def plot_splitsignal_map(m, outpath):
     things at rotation frequency
     things at orbital frequency
     """
-    fig, axs = plt.subplots(nrows=4, figsize=(14, 12), sharex=True)
 
-    axs[0].set_ylabel('flux')
-    axs[0].plot(m.x_obs, m.y_obs, ".k", ms=4, label="data")
+    # 8.5x11 is letter paper. x10 allows space for caption.
+    fig, axs = plt.subplots(nrows=4, figsize=(8.5, 10), sharex=True)
+
+    axs[0].set_ylabel('f', fontsize='x-large')
+    axs[0].plot(m.x_obs, m.y_obs, ".k", ms=4, label="data", zorder=2, rasterized=True)
     axs[0].plot(m.x_obs, m.map_estimate['mu_model'], lw=0.5, label='MAP',
-                color='C0', alpha=1, zorder=5)
+                color='C0', alpha=1, zorder=1)
 
     y_tra = m.map_estimate['mu_transit']
     for ix, f in enumerate(['rot', 'orb']):
@@ -144,35 +146,40 @@ def plot_splitsignal_map(m, outpath):
             yval += m.map_estimate[k1]
         if f == 'rot':
             y_rot = yval
-            axs[0].plot(m.x_obs, y_rot, lw=0.5, label='model '+f,
-                        color='C{}'.format(ix+1), alpha=1, zorder=ix+3)
+            # axs[0].plot(m.x_obs, y_rot, lw=0.5, label='model '+f,
+            #             color='C{}'.format(ix+1), alpha=1, zorder=ix+3)
         if f == 'orb':
             y_orb = yval + y_tra
-            axs[0].plot(m.x_obs, y_orb, lw=0.5, label='model '+f,
-                        color='C{}'.format(ix+1), alpha=1, zorder=ix+3)
+            # axs[0].plot(m.x_obs, y_orb, lw=0.5, label='model '+f,
+            #             color='C{}'.format(ix+1), alpha=1, zorder=ix+3)
 
-    axs[1].set_ylabel('flux-orb (rot)')
-    axs[1].plot(m.x_obs, m.y_obs-y_orb, ".k", ms=4, label="data-orb")
+    #axs[1].set_ylabel('flux-orb (rot)')
[…]

### 2020-03-28 drivers/real.py

@@ -16,19 +16,22 @@ def main(modelid):
 
     traceplot = 0
     sampleplot = 0
-    cornerplot = 1
-    splitsignalplot = 0 if 'Porb' in modelid and 'Prot' in modelid else 0
+    cornerplot = 0
+    splitsignalplot = 1 if 'Porb' in modelid and 'Prot' in modelid else 0
 
     REALID = 'PTFO_8-8695'
     RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
-    PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID))
+    PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID),
+                           '20200320_linear_amplitude_prior')
+
+    ##########
+
     if not os.path.exists(PLOTDIR):
         os.mkdir(PLOTDIR)
     pklpath = os.path.join(
         os.path.expanduser('~'), 'local', 'billy',
         '{}_model_{}.pkl'.format(REALID, modelid)
     )
-
     np.random.seed(42)
 
     x_obs, y_obs, y_err = get_clean_ptfo_data()

### 2020-03-28 paper/ms.tex
update paper
paper updates

@@
 
-%\begin{figure*}[t]
-%	\begin{center}
-%		\leavevmode
-%		\includegraphics[width=0.95\textwidth]{f2.png}
-%	\end{center}
-%	\vspace{-0.7cm}
-%	\caption{ {\bf foo.}
-%    bar
-%		\label{fig:lc_map}
-%	}
-%\end{figure*}
-
 \begin{figure*}[t]
-  \gridline{
-    \fig{f3a.png}{0.8\textwidth}{}
-    \fig{f3b.png}{0.2\textwidth}{}
-  }
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.98\textwidth]{f1.png}
+	\end{center}
 	\vspace{-0.7cm}
-	\caption{ {\bf foo.}
-    bar
+	\caption{ {\bf TESS light-curve of PTFO 8-8695 (Sector 6, Orbit 1).}
+	{\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux versus time. The beat period of 4.48 days is visible by eye.
+	The model plotted underneath the data includes 2 harmonics at the long
+	period $P_{\rm \ell}$, plus 2 harmonics and a transit at the short period $P_{\rm s}$.
+	{\it Upper middle}: Long-period signal, equal to the raw signal minus the short-period signal.
+	{\it Lower middle}: Short-period signal, equal to the raw signal minus the long-period signal.
+	{\it Bottom}: residual.
+	The data are binned from 2 to 10 minute cadence as a convenience for plotting and fitting.
 		\label{fig:splitsignal}
@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f4.png}
+		\includegraphics[width=0.95\textwidth]{f2.pdf}
 	\end{center}
 	\vspace{-0.7cm}
-	\caption{ {\bf foo.}
-    bar
+	\caption{ {\bf Phase-folded long and short-period signals.}
+	{\it Top}: Long-period signal, as in Figure~\ref{fig:splitsignal}.
+	{\it Bottom}: Short-period signal. The reference phase is set to the most ``planetary'' dip.
+	Gray points are the 10 minute cadence \texttt{PDCSAP} flux.
+	Black points are binned to 100 points per period.
 	\label{fig:phasefold}
@@
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f1.pdf}
+		\includegraphics[width=0.48\textwidth]{f3.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{o_minus_c.png}
+		\includegraphics[width=0.9\textwidth]{f4.png}
 	\end{center}

### 2020-03-28 paper/collect.sh

@@ -5,8 +5,7 @@ pdir=../results/PTFO_8-8695_results/20200320_linear_amplitude_prior/
 #cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_joined.pdf f1.pdf
 cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap.png f1.png
 
-#cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.pdf f2.pdf
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.png f2.png
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.pdf f2.pdf
 
 cp ../results/PTFO_8-8695_results/scene.pdf f3.pdf
 

### 2020-04-02 billy/models.py
add plot_O_minus_C

@@ -32,3 +32,37 @@ def transit_model(params, t, texp=30/(60*24), mstar=1, rstar=1):
         .get_light_curve(orbit=orbit, r=r, t=t, texp=texp)
         .eval().flatten()
     )
+
+
+def linear_model(params, x, x_occ=None):
+    """
+    Linear model. Parameters (t0, P).
+    Must pass transit times.
+
+    If x_occ is none, returns model t_tra array.
+    If x_occ is a numpy array, returns tuple of model t_tra and t_occ arrays.
+    """
+    t0, period = params
+    if not isinstance(x_occ,np.ndarray):
+        return t0 + period*x
+    else:
+        return t0 + period*x, t0 + period/2 + period*x_occ
+
+
+def quadratic_model(params, x, x_occ=None):
+    """
+    Quadratic model. Parameters (t0, P, 0.5dP/dE).
+    Must pass transit times.
+
+    If x_occ is none, returns model t_tra array.
+    If x_occ is a numpy array, returns tuple of model t_tra and t_occ arrays.
+    """
+    t0, period, half_dP_dE = params
+    if not isinstance(x_occ,np.ndarray):
+        return t0 + period*x + half_dP_dE*x**2
+    else:
+        return (t0 + period*x + half_dP_dE*x**2,
+                t0 + period/2 + period*x_occ + half_dP_dE*x_occ**2
+               )
+
+

### 2020-04-02 billy/plotting.py

@@ -14,6 +14,8 @@ Plots:
 
     plot_scene
 
+    plot_O_minus_C
+
 Convenience:
     savefig
     format_ax
@@ -26,9 +28,14 @@ from itertools import product
 
 from billy.convenience import flatten as bflatten
 from billy.convenience import get_clean_ptfo_data
+from billy.models import linear_model
 
 from astrobase.lcmath import phase_magseries, phase_bin_magseries
 from astropy.stats import LombScargle
+from astropy import units as u, constants as const
+
+from numpy import array as nparr
+from astropy.time import Time
 
 def plot_periodogram(outdir, islinear=True):
 
@@ -251,6 +258,7 @@ def plot_splitsignal_map_periodogram(ydict, outpath):
             ax.set_xlim([period_min, period_max])
         else:
             ax.set_xlim([_period_min, _period_max])
+            ax.set_xscale('log')
 
     axs[-1].set_xlabel('Period [days]')
 
@@ -529,7 +537,6 @@ def plot_scene(c_obj, img_wcs, img, outpath, Tmag_cutoff=17, showcolorbar=0,
                ap_mask=0, bkgd_mask=0):
 
     from astrobase.plotbase import skyview_stamp
-    from astropy import units as u, constants as const
     from astropy.wcs import WCS
     from astroquery.mast import Catalogs
     import astropy.visualization as vis
@@ -703,3 +710,137 @@ def plot_scene(c_obj, img_wcs, img, outpath, Tmag_cutoff=17, showcolorbar=0,
         fig.tight_layout(h_pad=1, w_pad=1)
 
     savefig(fig, outpath, dpi=300)
+
+
+
+def plot_O_minus_C(
+    x, y, sigma_y, theta_linear, refs, savpath=None, xlabel='Epoch',
+    ylabel='Deviation from constant period [min]', xlim=None, ylim=None,
+    ylim1=None, include_all_points=False, onlytransits=False):
+
+    xfit = np.linspace(10*np.min(x), 10*np.max(x), 10000)
+
+    if not onlytransits:
+        raise NotImplementedError
+
+    fig, a0 = plt.subplots(nrows=1, ncols=1, figsize=(4*1.3,3*1.3))
+
+    refs = refs.astype(str)
+    istess = ( np.core.defchararray.find(refs, 'tess')!=-1 )
+
[…]

### 2020-04-02 drivers/plot_O_minus_C.py

@@ -1,0 +1,76 @@
+'''
+make O-C model figure
+'''
+import pandas as pd, numpy as np
+import os
+from numpy import array as nparr
+from astrobase.timeutils import get_epochs_given_midtimes_and_period
+
+import billy.plotting as bp
+
+EPOCH = 2455543.94300
+PERIOD = 0.448399
+
+def get_data(
+    datacsv='../data/WASP-18b_transits_and_TESS_times_O-C_vs_epoch_selected.csv',
+    is_occultation=False
+    ):
+
+    df = pd.read_csv(datacsv, sep=';')
+
+    # 'period', 'period_err', 't0_HJD', 'err_t0', 't0_BJD_TDB', 'reference',
+    #    'where_I_got_time', 'comment'
+
+    y = nparr(df.t0_BJD_TDB)
+    sigma_y = nparr(df.err_t0)
+    refs = nparr(df.reference)
+
+    sel = np.isfinite(y) & np.isfinite(sigma_y)
+
+    y, sigma_y, refs = y[sel], sigma_y[sel], refs[sel]
+
+    x, _ = get_epochs_given_midtimes_and_period(
+        y, PERIOD, t0_fixed=EPOCH, verbose=True
+    )
+
+    return x, y, sigma_y, refs
+
+
+def main(xlim=None, ylim=None, include_all_points=False, ylim1=None):
+
+    homedir = os.path.expanduser('~')
+    basedir = os.path.join(homedir, 'Dropbox/proj/billy/')
+    transitpath = (
+        basedir+'data/ephemeris/PTFO_8-8695_manual_all.csv'
+    )
+    occpath = 'foo.csv'
+
+    print('getting data from {:s}'.format(transitpath))
+    x, y, sigma_y, refs = get_data(datacsv=transitpath)
+    print('getting data from {:s}'.format(occpath))
+
+    if os.path.exists(occpath):
+        raise NotImplementedError
+
+    linear_params = nparr([EPOCH,  PERIOD])
+
+    savpath = '../results/ephemeris/O_minus_C.png'
+
+    onlytransits = True
+    bp.plot_O_minus_C(
+        x, y, sigma_y, linear_params, refs, savpath=savpath,
+        xlabel='Cycle number', ylabel=None, xlim=xlim, ylim=ylim,
+        include_all_points=include_all_points, ylim1=ylim1,
+        onlytransits=onlytransits)
+
+    if onlytransits:
[…]

### 2020-04-02 paper/ms.tex
minor paper edits

@@
   orbital phase of the dip seems to have instantaneously jumped, at
-  least twice.
+  least once, and maybe twice.
   The TESS epoch of the dip is consistent with recent observations by
@@
 
-\subsection{Blend considerations}
-\label{subsec:blend}
-
-The TESS pixels are $\approx21$'' per side, and so we need to consider
-whether light from neighboring stars could affect the photometry.  The
-scene is shown in Figure~\ref{fig:scene}.  
-The pixels used to
-measure the background level are indicated with an `\texttt{X}' hatch,
-and the pixels used for the final light-curve are shown with the
-`\texttt{/}' hatch.
-
-The target star, PTFO
-8-8605 (TIC 264461976), has a $T$-band magnitude of 14.0, and its position is shown with a
-star.  
-The other (unlabeled) star inside the target aperture, TIC 264461979, has $T=16.8$ and so cannot
-contribute a signal with relative amplitude 10\%.
-The only neighbor that is sufficiently close and bright that
-its light might contaminate the target star is TIC 264461980, with
-$T=14.8$, which we denote ``Star A''.  Star A is 23.6'' NW of our
-target, and based on the magnitude difference could contribute up to
-48\% the flux of our target star, PTFO 8-8695.
-
-Because PTFO 8-8695 has previously been identified to have periodicity
-consistent with our measurement of $P_{\rm s}$, our main concern
-regarding blending is the degree to which we can be certain that the
-long-period signal at $P_{\rm \ell}$ also originates from PTFO 8-8695.
-We took two approaches towards determining the source of the long-period signal.
-
-First, we examined the CDIPS full frame image light-curves of the
-target, which are available on MAST \citep{bouma_cluster_2019}.
-The maximal peak-to-peak beat amplitude is consistently $\approx$10\%
-across apertures of radii 1, 1.5, and 2.25 pixels.
-If Star A were the source of the long-period variability, we would expect the
-peak variability amplitude to be smallest in the 1 pixel aperture, based on the
-separation of the sources (Figure~\ref{fig:scene}, bottom).
-From this test alone, it seems unlikely that Star A is the source of
-the long-period signal.
-
-Second, we examined the light-curve of each pixel in the scene
-individually.  We opted to use the
-interactive tools implemented in
-\texttt{lightkurve} \citep{lightkurve_2018}.  If Star A were the
-source of the long-period variability, we would expect the pixels
-nearest to Star A to show a sinusoidal signal with
-amplitude exceeding $10\%$.  We find no evidence for
-this being the case.  The pixel directly below Star A does not
-clearly show the sinusoidal variability, and the peak-to-peak 
-variability in that pixel is $\lesssim 8\%$.  In contrast, the
-south-easternmost pixel within PTFO 8-8695's aperture (the pixel 
-furthest from Star A that was used in the optimal aperture) shows the $P_{\rm \ell}$ sinusoidal
-variability signal at $\approx 10\%$ amplitude.
-
-As there is no evidence in favor of a blend scenario, we
-conclude that both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from PTFO 8-8695.
-
 \section{Analysis}
@@
 	{\it Top}: Long-period signal, as in Figure~\ref{fig:splitsignal}.
-	{\it Bottom}: Short-period signal. The reference phase is set to the most ``planetary'' dip.
+	{\it Bottom}: Short-period signal. The reference phase is set to the ``planetary'' dip.
 	Gray points are the 10 minute cadence \texttt{PDCSAP} flux.
@@
 		The position of PTFO 8-8695 is shown with a yellow star.
-		Neighbors with $T<17$ are shown with red crosses.
+		Neighbors with $T<17$ are shown with orange crosses.
 		The apertures used to measure the background and target star flux
 		are shown with \texttt{X} and \texttt{/} hatches, respectively.
-		{\it Bottom:} DDS2 $R$-band image of the same field, with a linear stretch. The circles show
+		{\it Bottom:} Digitized Sky Survey $R$-band image of the same field, with a linear stretch. The circles show
 		apertures of radii 1, 1.5, and 2.25 pixels used in part of our blend
@@
 
+\paragraph{Fourier analysis}
+Figure~\ref{fig:splitsignal} could also be given in the frequency
+domain, rather than the time domain.
+The largest peaks in 
+the Lomb-Scargle periodogram (CITE) of the raw light-curve are
+at $P_\rm {s}$ and $P_{\rm \ell}$.
+Both peaks of course have aliases.
+The $P_\rm {\ell}$ peak has larger ampitude.
+Subtracting out the long-period signal, the short-period signal
+dominates the periodogram, and vice-versa.
+The periodogram of the final residual (Figure~\ref{fig:splitsignal}
+bottom row) shows a weakly significant, poorly resolved peak at
+$\approx$8 days, consistent with the visual impression in the time
+domain that there could be a weak long-period signal present.
+
+\subsection{Blend considerations}
+\label{subsec:blend}
+
+The TESS pixels are $\approx21$'' per side, and so we need to consider
+whether light from neighboring stars could affect the photometry.  The
+scene is shown in Figure~\ref{fig:scene}.  
+The pixels used to
+measure the background level are indicated with an `\texttt{X}' hatch,
+and the pixels used for the final light-curve are shown with the
+`\texttt{/}' hatch.
+
+The target star, PTFO
+8-8605 (TIC 264461976), has a $T$-band magnitude of 14.0, and its position is shown with a
+star.  
+The other (unlabeled) star inside the target aperture, TIC 264461979, has $T=16.8$ and so cannot
+contribute a signal with relative amplitude 10\%.
+The only neighbor that is sufficiently close and bright that
+its light might contaminate the target star is TIC 264461980, with
+$T=14.8$, which we denote ``Star A''.  Star A is 23.6'' NW of our
+target, and based on the magnitude difference could contribute up to
+48\% the flux of our target star, PTFO 8-8695.
+
+Because PTFO 8-8695 has previously been identified to have periodicity
+consistent with our measurement of $P_{\rm s}$, our main concern
+regarding blending is the degree to which we can be certain that the
+long-period signal at $P_{\rm \ell}$ also originates from PTFO 8-8695.
+We took two approaches towards determining the source of the long-period signal.
+
+First, we examined the CDIPS full frame image light-curves of the
+target, which are available on MAST \citep{bouma_cluster_2019}.
+The maximal peak-to-peak beat amplitude is consistently $\approx$10\%
+across apertures of radii 1, 1.5, and 2.25 pixels.
+If Star A were the source of the long-period variability, we would expect the
+peak variability amplitude to be smallest in the 1 pixel aperture, based on the
+separation of the sources (Figure~\ref{fig:scene}, bottom).
+From this test alone, it seems unlikely that Star A is the source of
+the long-period signal.
+
+Second, we examined the light-curve of each pixel in the scene
+individually.  We opted to use the
+interactive tools implemented in
+\texttt{lightkurve} \citep{lightkurve_2018}.  If Star A were the
+source of the long-period variability, we would expect the pixels
+nearest to Star A to show a sinusoidal signal with
+amplitude exceeding $10\%$.  We find no evidence for
+this being the case.  The pixel directly below Star A does not
+clearly show the sinusoidal variability, and the peak-to-peak 
+variability in that pixel is $\lesssim 8\%$.  In contrast, the
+south-easternmost pixel within PTFO 8-8695's aperture (the pixel 
+furthest from Star A that was used in the optimal aperture) shows the $P_{\rm \ell}$ sinusoidal
+variability signal at $\approx 10\%$ amplitude.
+
+As there is no evidence in favor of a blend scenario, we
+conclude that both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from PTFO 8-8695.
+
 \section{Discussion}
 \label{sec:discussion}
+
+The TESS dip does not phase up where it is supposed to...
+\label{fig:o_minus_c}
 

### 2020-04-02 paper/collect.sh

@@ -9,7 +9,7 @@ cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.pdf f2.pdf
 
 cp ../results/PTFO_8-8695_results/scene.pdf f3.pdf
 
-cp ../results/PTFO_8-8695_literature_and_TESS_times_O-C_vs_epoch_all_tanimoto2020_ephem.png f4.png
+cp ../results/ephemeris/O_minus_C.pdf f4.pdf
 
 cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f5.png
 pngquant f5.png --output f5_comp.png

### 2020-04-02 paper/ms.tex
update paper

@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f4.png}
+		\includegraphics[width=0.9\textwidth]{f4.pdf}
 	\end{center}
 	\vspace{-0.7cm}
-	\caption{ {\bf foo.}
-		bar
+	\caption{
+    {\bf Timing residuals for PTFO 8-8695b from a decade of monitoring.}
+		Black points are times of ``dips'', minus the indicated linear ephemeris.
+    The $y$-axis is given in units of phase for the
+    short-period signal.
+    The star shows the binned TESS ephemeris.
+		``Dips'' have been observed by
+		\citet{van_eyken_ptf_2012}, \citet{ciardi_follow-up_2015},
+		\citet{yu_tests_2015}, \citet{raetz_yeti_2016},
+		\citet{onitsuka_multi-color_2017}, and \citet{tanimoto_evidence_2020}.
+    Certain dips ({\it e.g.}, the one at phase 0 in mid-2019) are
+    consistent with noise, and were likely reported because
+    something was {\it expected}, rather than convincingly {\it
+    observed}.
+    Horizontal dashed lines are drawn at $\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$,
+    highlighting what could be either a numerical coincidence or an observational
+    bias.
+    The orbital phase observed by TESS is consistent with that of \citet{tanimoto_evidence_2020}, and quite different from the original phase.
 		\label{fig:o_minus_c}
@@
 the Lomb-Scargle periodogram (CITE) of the raw light-curve are
-at $P_\rm {s}$ and $P_{\rm \ell}$.
+at $P_{\rm s}$ and $P_{\rm \ell}$.
 Both peaks of course have aliases.
-The $P_\rm {\ell}$ peak has larger ampitude.
+The $P_{\rm \ell}$ peak has larger ampitude.
 Subtracting out the long-period signal, the short-period signal

### 2020-04-02 billy/plotting.py

@@ -737,12 +737,15 @@ def plot_O_minus_C(
             zorder=8, label='binned TESS', markerfacecolor='yellow',
             markersize=9, marker='*', color='black', lw=0)
 
+    P_short = 0.44854
+    P_long = 0.49914
+    phase_diff = np.abs(P_long - P_short) / P_short
     a0.hlines(
-        1/9, min(xlim), max(xlim), colors='gray', alpha=0.5,
+        phase_diff, min(xlim), max(xlim), colors='gray', alpha=0.5,
         linestyles='--', zorder=-2, linewidths=0.5
     )
     a0.hlines(
-        -1/9, min(xlim), max(xlim), colors='gray', alpha=0.5,
+        -phase_diff, min(xlim), max(xlim), colors='gray', alpha=0.5,
         linestyles='--', zorder=-2, linewidths=0.5
     )
     a0.hlines(

### 2020-04-02 paper/ms.tex
paper edits

@@
 
+% \subsection{Observations}
+
+PTFO 8-8695 was observed by TESS with Camera 1, CCD 1, from December
+15, 2018 to January 6, 2019, during the sixth sector of science
+operations.  The star is designated TIC 264461976 in the TESS Input
+Catalog \citep{stassun_TIC_2018}.  The pixel data for an $11\times11$
+array surrounding PTFO 8-8695 were averaged into 2-minute stacks by
+the onboard computer. 
+Each 2048$\times$2048 image was also averaged into 30-minute stacks,
+and saved as a ``full frame image'' (FFI).
+
+%FIXME these paragraphs are directly copied from WASP-4 paperI
+%FIXME these paragraphs are directly copied from WASP-4 paperI
+The data were downlinked via the Deep Space
+Network\footnote{\url{deepspace.jpl.nasa.gov}}, and the
+spacecraft timestamps were calibrated against the ground-station
+clocks.  The spacecraft clock times were then transformed by the
+Payload Operations Center into the {\it Temps Dynamique Barycentrique}
+(TDB) reference system.  The images were then reduced to lightcurves
+by the Science Processing Operations Center (SPOC) at NASA
+Ames~\citep{jenkins_tess_2016}.
+
+%FIXME these paragraphs are directly copied from WASP-4 paperI
+%FIXME these paragraphs are directly copied from WASP-4 paperI
+We began our analysis with the Presearch Data Conditioning (PDC)
+lightcurve, which has had non-astrophysical variability removed
+through the methods discussed by \citet{smith_kepler_apertures_2017}
+and \citet{smith_kepler_PDC_2017}.  We then processed the lightcurve
+as follows.  First, we removed all points with non-zero quality flags.
+This removed data that might have been adversely affected by
+``momentum dumps,'' the firing of thrusters and resetting of reaction
+wheels\footnote{The spacecraft pointing and momentum dumps are
+described in the data release notes:
+\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_07_drn09_v03.pdf}}
+that took place every 2.5 days during sector 2.  The data during these
+events were assigned quality flags corresponding to ``Reaction Wheel
+Desaturation Event'' and ``Manual Exclude'' \citep[][Table
+28]{tess_data_product_description_2018}.  For WASP-4, these flags were
+simultaneously set for 54 distinct cadences, and there were 10
+momentum dumps, averaging about 10 minutes of flagged data per dump.
+Out of caution, we clipped out an additional 10 minutes before and
+after every momentum dump.  We also removed the data within the first
+and last hours of both orbits, because of correlated red noise that
+appears during those time ranges.
+
+%FIXME these paragraphs are directly copied from WASP-4 paperI
+%FIXME these paragraphs are directly copied from WASP-4 paperI
+All told, we removed 8\% of the original data points, and were left
+with 18{,}165 measurements of the relative flux of WASP-4.  We
+normalized the data by dividing out the median flux.  We converted the
+timestamps from BTJD$_{\rm TDB}$ into BJD$_{\rm TDB}$ by adding the
+appropriate 2{,}457{,}000 day offset
+\citep{tess_data_product_description_2018}.  Many of these and
+subsequent processing steps were performed using
+\texttt{astrobase}~\citep{bhatti_astrobase_2018}. We did not
+``flatten'' the lightcurves, as is often done with splines,
+polynomials, or Gaussian processes.  Instead, we modeled the
+out-of-transit flux variations simultaneously with the transit
+parameters, as described below.
+
+% \subsection{Star and planet parameters}
+% \label{sec:system_parameters}
+% 
+% \begin{deluxetable}{lccc}
+% \tabletypesize{\scriptsize}
+% \tablecaption{Selected system parameters of WASP-4b\label{tbl:params}}
+% \tablenum{1}
+% 
+% \tablehead{
+% \colhead{Parameter} & \colhead{Value} & \colhead{68\% Confidence Interval} & \colhead{Comment}
+% }
+% 
+% \startdata
+% {\it Transit/RV parameters:} & & & \\
+%   $R_{\rm p}/R_\star$                        & $0.15201$              & $+0.00040$, $-0.00033$      & A \\
+%   $i$~[deg]                                  & $89.06$                & $+0.65$, $-0.84$            & A \\
+%   $a/R_\star$                                & $5.451$                & $+0.023$, $-0.052$          & A \\
+%   $u_{\rm linear}$                           & $0.382$                & ---                         & A \\
+%   $u_{\rm quad}$                             & $0.210$                & ---                         & A \\
+%   $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & B \\
+% {\it Stellar parameters:} & & & \\
+%   $T_{\rm eff}$~[K]                          & $5400$                 & $\pm 90$                    & C \\
+%   $\log g_\star$~[cgs]                       & $4.47$                 & $\pm 0.11$                  & C \\
+%   $[{\rm Fe/H}]$                             & $-0.07$                & $\pm 0.19$                  & C \\
+%   $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & D \\
+%   $A_V$~[mag]                                & $0.03$                 & $+0.02, -0.01$              & D \\
+%   $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & F \\
+%   $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & E \\
+%   $\rho_\star$~[g~cm$^{-3}$]                 & $1.711$                & $+0.022$, $-0.048$          & E \\
+%   $M_\star$~[M$_{\odot}$]                    & $0.864$                & $+0.084$, $-0.090$          & E \\
+%   $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
+% {\it Planetary parameters:} & & & \\
+%   $a$~[AU]                                   & $0.0226$               & $+0.0007$, $-0.0008$        & E \\
+%   $M_p$~[M$_{\rm Jup}$]                      & $1.186$                & $+0.090$, $-0.098$          & E \\
+%   $R_p$~[R$_{\rm Jup}$]                      & $1.321$                & $\pm 0.039$                 & E \\
+% \enddata
+% 
+% \tablecomments{
+%   (A) From phase-folded TESS lightcurve (\S~\ref{sec:measurement}).
+%   Orbital periods are in Table~4. The limb darkening parameters were
+%   allowed to float around the \citet{claret_limb_2017} prediction, but
+%   were unconstrained.
+%   (B) \citet{triaud_spin-orbit_2010}.
+%   (C) From HARPS spectra \citep{doyle_accurate_2013}.
+%   (D) \citet{stassun_accurate_2017}.
+%   (E) This work, see \S~\ref{sec:system_parameters}.
+%   (F) \citet{gaia_collaboration_gaia_2018}.
+%   (G) \citet{stassun_TIC_2018}.
+% }
+% 
+% \vspace{-1cm}
+% \end{deluxetable}
+% 
+% We calculated the stellar and planetary parameters in the following
+% way.  We computed the star's spectral energy distribution based on the
+% {\it Gaia} DR2 parallax (after making the small correction advocated
+% by \citealt{stassun_evidence_2018}) and the broadband magnitudes from
+% the available all-sky catalogs: $G$ from {\it Gaia\/} DR2, $B_T$ and
+% $V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from {\it
+% 2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
+% wavelength range 0.4--22~$\mu$m.
+% We adopted the effective temperature
+% from the work by \citet{doyle_accurate_2013}, who determined the
+% spectroscopic parameters of WASP-4 using high-signal-to-noise
+% observations with the High Accuracy Radial-velocity Planet Searcher
+% (HARPS).  Then, we determined the stellar radius through the
+% combination of the bolometric luminosity and the effective
+% temperature, using the Stefan-Boltzmann law.  To determine the stellar
+% mass, we first computed the mean stellar density based on the value of
+% $a/R_\star$ that gave the best fit to the phase-folded TESS lightcurve
+% (for the relevant equation, see \citealt{seager_unique_2003} or
+% \citealt{winn_exoplanet_2010}).  The mass was calculated from the
+% radius and density, and the orbital distance was also calculated from
+% the radius and $a/R_\star$.  The planetary radius was calculated as
+% the product of $R_\star$ and $R_{\rm p}/R_\star$.  Finally, the planet
+% mass was calculated based on the stellar mass, the radial-velocity
+% amplitude observed by \citet{triaud_spin-orbit_2010}, and the orbital
+% inclination.
+% 
+% Table~1 gives the resulting parameters, which we adopted for the
+% remaining analysis.  \added{The uncertainties in our derived stellar
+% and planetary parameters are propagated according to standard analytic
+% formulae, under the assumption that the variables are uncorrelated and
+% normally distributed.} Our \added{system} parameters are in agreement
+% with those of previous investigators, but have the benefit of
+% incorporating the {\it Gaia} parallax
+% \citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013,huitson_gemini_2017}.
+% By comparing the star's luminosity and spectroscopic parameters with
+% the outputs of the Yonsei-Yale stellar-evolutionary models, we found
+% that WASP-4 is a main-sequence star with an age of approximately
+% $7\,{\rm Gyr}$ \added{\citep{demarque_y2_2004}}.
+
 \section{Analysis}
@@
 
-% TODO:
-% Figure: CDIPS LC + SPOC LC (left two panels). Right panel:
-% periodogram.
-%
+Our initial inspection of the light-curve, in both its 2-minute PDCSAP
+and 30-minute FFI forms, revealed an obvious beat signal
+(Figure~\ref{fig:splitsignal}, top panel).
+
+As a precursor to more detailed analysis, we calculated generalized
+Lomb-Scargle periodograms using \texttt{astrobase}
+\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}. 
+The two largest peaks in the Lomb-Scargle periodogram of the
+light-curve were clearly separated at $P_{\rm s} \approx 0.448\,{\rm
+days}$ and $P_{\rm \ell} \approx 0.499\,{\rm days}$.
+The $P_{\rm \ell}$ peak had the larger ampitude of the two.
+Smaller harmonics from each of these dominant two signals were also
+present.
+
+Initial fitting experiments with splines showed that subtracting out
+the long-period signal, the short-period signal dominated the
+periodogram, and vice-versa.
+However it quickly became clear that a mroe careful modelling approach
+than ``flattening'' the light-curve with splines or Gaussian proceses
+would be preferable, in order to preserve the power at each frequency.
 
 \subsection{Model Fitting}
+
+We opted to model the light-curve as a linear combination of Fourier
+harmonics at the short and long periods, plus a transit at the
+short period.
+Symbolically,
+\begin{equation}
+  f = f_{\rm s} + f_{\rm \ell} = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
+\end{equation}
+where $f_{\rm s}$ is the relative flux at the short period, and
+$f_{\rm \ell}$ is the flux at the long period.
+Decomposing each term,
+\begin{align}
+  f = &f_{\rm transit,s} + \sum_{n=1}^{n=N} A_n \sin(n\omega_{\rm s}t)
+  + \sum_{n=1}^{n=N} B_n \cos(n\omega_{\rm s}t)\\
+  &+ \sum_{m=1}^{m=M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
+  + \sum_{m=1}^{m=M} B_m \cos([m\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
+\end{align}
+for $N$ and $M$ the total number of harmonics at the short and long period, respective,
+$A_i$ and $B_i$ for each harmonic term, and $\omega_i = 2\pi / P_i$ the angular
+frequency for $i$ the short or long period index. 
+We fixed the ``phase-offset'' for the short period signal to be zero,
+and let the reference time of the long period signal float by introducing
+$\phi_{\rm \ell}$. 
+Since we did not a priori know how many harmonics would be
+appropriate, we let the number of harmonics vary, and
+used the Bayesian information criterion to choose the
+appropriate model.
+
+For example, one possible model could be a transit, plus $N=2$ harmonics
+of sines and cosines at the short period, plus $M=1$  harmonics at the long
+period.
+In this case, the free parameters would be as follows.
+For the transit, we would fit for
+the impact parameter, the planet-to-star radius ratio, two quadratic limb darkening parameters,
+the planet orbital period (equal to the short period), the reference time
+for the transit, and the mean flux.
+There would be 4 additional Fourier amplitudes at the short period,
+plus 2 Fourier amplitudes at the long period, and well as the long period itself
+and its phase.
+For this case, we therefore fitted 14 free parameters.
+
+% 
+% I gave it a pass on the PTFO data, binning from 2->10 minute sampling
+% to speed up the execution time. Results from two models are below.
+% Case 1 is N=2,M=1. Case 2 is N=2,M=2. Both fits are fine, and I get
+% decent convergence with a few minutes each. The rotation parameters
+% (phase + amplitudes) are quite correlated, so I set a reasonably
+% strong prior on the rotation phase.
+% 
+% The main result that both models agree on: the extra power at the
+% orbital frequency seems to not be ellipsoidal. Both of these models
+% prefer to put the power at 1x the orbital frequency, not 2x the
+% orbital frequency! This gives both Aorb0 and Borb0 non-zero, while
+% Aorb1 and Borb1 seem to be more consistent with zero.
+
+% Using the cleaned PDC lightcurve, we applied the Box Least Squares
+% algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
+% period, transit duration, and a reference epoch using the TESS data
+% alone.  Based on the results, we isolated the data within 4 transit
+% durations of each transit midpoint.  To find the transit parameters
+% that best fit the data, we first fitted a line to the out-of-transit
+% flux measurements surrounding each transit, and divided it out.  We
+% then created a phase-folded lightcurve \added{from all 18 transits}
+% and fitted a standard transit model using the analytic formulae given
+% by \citet{mandel_analytic_2002} and implemented by
+% \replaced{\citet{kreidberg_batman_2015}}{\citet[][\texttt{BATMAN}]{kreidberg_batman_2015}}
+% We assumed the orbit to be circular, consistent with the limits from
+% radial velocities and occultation timing
+% \citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
+% The free parameters were the reference epoch, the planet to star
+% radius ratio $R_{\rm p}/R_\star$, the orbital distance to stellar
+% radius ratio $a/R_\star$, the inclination $i$, two quadratic
+% limb-darkening coefficients $(u_{\rm linear}, u_{\rm quad})$, and the
+% orbital period $P$.
+% 
+% We sampled the posterior probability distribution for all the
+% parameters using the algorithm proposed by
+% \citet{goodman_ensemble_2010} and implemented by
+% \replaced{\citet{foreman-mackey_emcee_2013}}{\citet[][\texttt{emcee}]{foreman-mackey_emcee_2013}}.
+% Table~1 gives the results, which are in reasonable agreement with the
+% parameters reported by {\it e.g.},
+% \citet{southworth_high-precision_2009} and
+% \citet{huitson_gemini_2017}.  Figure~\ref{fig:phasefold} shows the
+% phase-folded lightcurve.
+% 
+% To measure the transit times, we returned to the `cleaned' PDC time
+% series and fitted the data within four transit durations of each
+% transit separately. We used four free parameters: the time of
+% mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
+% slope and intercept of a linear trend to account for any slow
+% variations unrelated to the transit.  We fixed the remaining
+% parameters at the values that had been determined from the
+% phase-folded TESS lightcurve.  The uncertainty in each
+% \added{photometric} data point was set equal to the root-mean-square
+% (rms) level of the out-of-transit data.
+% 
+% To verify that the measured uncertainties are estimated accurately, we
+% computed the \deleted{reduced }$\chi^2$\added{ value} for a linear
+% ephemeris fit to the measured TESS mid-transit times.  We found that
+% $\chi^2 = 9.2$, with $n=16$ degrees of freedom.  The variance of the
+% $\chi^2$ distribution is $2n$, so we would expect $\chi^2 = 16 \pm
+% 5.7$.  Visually inspecting the residuals showed that the error
+% variance had been overestimated, so we multiplied the measured TESS
+% errors by a factor $f=0.76$, forcing a reduced $\chi^2$ of unity.
+% This lowered the mean uncertainty of the transit midtimes from $29.8$
+% to $22.6$ seconds.  We verified that omitting this step did not
+% appreciably alter any of our conclusions.
+% 
+% Figure~\ref{fig:lightcurves} shows the lightcurve of each individual
+% transit, the best-fit models, and the residuals.  Table~2 reports the
+% mid-transit times and their uncertainties.  After binning the
+% residuals to 1-hour windows, the lightcurves have an rms scatter of
+% 586\,{\rm ppm}.  The pre-launch TESS noise
+% model\footnote{\url{github.com/lgbouma/tnm}, commit \texttt{be06f09}}
+% (\citealt{winn_photonflux_2013}, \citealt{Sullivan_2015} Section 6.4)
+% would have predicted an error budget consisting of the following terms
+% added in quadrature: 410\,{\rm ppm} from photon-counting noise,
+% 202\,{\rm ppm} from detector read noise, and 673\,{\rm ppm} from the
+% zodiacal background light.  The level of background light appears to
+% have been overestimated in the model.
 
@@
 
-\paragraph{Fourier analysis}
-Figure~\ref{fig:splitsignal} could also be given in the frequency
-domain, rather than the time domain.
-The largest peaks in 
-the Lomb-Scargle periodogram (CITE) of the raw light-curve are
-at $P_{\rm s}$ and $P_{\rm \ell}$.
-Both peaks of course have aliases.
-The $P_{\rm \ell}$ peak has larger ampitude.
-Subtracting out the long-period signal, the short-period signal
-dominates the periodogram, and vice-versa.
 The periodogram of the final residual (Figure~\ref{fig:splitsignal}

### 2020-04-03 paper/ms.tex
little bit of extra text

@@
 
-\section{Observations}
+\section{The Data}
 \label{sec:observations}
 
-% \subsection{Observations}
+\subsection{Observations}
 
@@
 operations.  The star is designated TIC 264461976 in the TESS Input
-Catalog \citep{stassun_TIC_2018}.  The pixel data for an $11\times11$
+Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.  The pixel data for an $11\times11$
 array surrounding PTFO 8-8695 were averaged into 2-minute stacks by
 the onboard computer. 
-Each 2048$\times$2048 image was also averaged into 30-minute stacks,
-and saved as a ``full frame image'' (FFI).
+Each 2048$\times$2048 image from the CCD was also averaged into
+30-minute stacks, and saved as a ``full frame image'' (FFI).
 
-%FIXME these paragraphs are directly copied from WASP-4 paperI
-%FIXME these paragraphs are directly copied from WASP-4 paperI
-The data were downlinked via the Deep Space
-Network\footnote{\url{deepspace.jpl.nasa.gov}}, and the
-spacecraft timestamps were calibrated against the ground-station
-clocks.  The spacecraft clock times were then transformed by the
-Payload Operations Center into the {\it Temps Dynamique Barycentrique}
-(TDB) reference system.  The images were then reduced to lightcurves
+The 2-minute stacks for PTFO 8-8695 were then reduced to lightcurves
 by the Science Processing Operations Center (SPOC) at NASA
-Ames~\citep{jenkins_tess_2016}.
+Ames~\citep{jenkins_tess_2016}.  Our main analysis used the resulting
+Presearch Data Conditioning (PDC) lightcurve, which had
+non-astrophysical variability removed through the methods discussed by
+\citet{smith_kepler_apertures_2017} and \citet{smith_kepler_PDC_2017}.
 
-%FIXME these paragraphs are directly copied from WASP-4 paperI
-%FIXME these paragraphs are directly copied from WASP-4 paperI
-We began our analysis with the Presearch Data Conditioning (PDC)
-lightcurve, which has had non-astrophysical variability removed
-through the methods discussed by \citet{smith_kepler_apertures_2017}
-and \citet{smith_kepler_PDC_2017}.  We then processed the lightcurve
-as follows.  First, we removed all points with non-zero quality flags.
-This removed data that might have been adversely affected by
-``momentum dumps,'' the firing of thrusters and resetting of reaction
-wheels\footnote{The spacecraft pointing and momentum dumps are
-described in the data release notes:
-\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_07_drn09_v03.pdf}}
-that took place every 2.5 days during sector 2.  The data during these
-events were assigned quality flags corresponding to ``Reaction Wheel
-Desaturation Event'' and ``Manual Exclude'' \citep[][Table
-28]{tess_data_product_description_2018}.  For WASP-4, these flags were
-simultaneously set for 54 distinct cadences, and there were 10
-momentum dumps, averaging about 10 minutes of flagged data per dump.
-Out of caution, we clipped out an additional 10 minutes before and
-after every momentum dump.  We also removed the data within the first
-and last hours of both orbits, because of correlated red noise that
-appears during those time ranges.
+We also processed the 30-minute image stacks separately, as part of
+the Cluster Difference Imaging
+Photometric Survey (CDIPS; \citep{bouma_cluster_2019}).
+The raw CDIPS light-curve shows identical features as the 2-minute
+light-curve, though at 30-minute cadence.
 
-%FIXME these paragraphs are directly copied from WASP-4 paperI
-%FIXME these paragraphs are directly copied from WASP-4 paperI
-All told, we removed 8\% of the original data points, and were left
-with 18{,}165 measurements of the relative flux of WASP-4.  We
-normalized the data by dividing out the median flux.  We converted the
-timestamps from BTJD$_{\rm TDB}$ into BJD$_{\rm TDB}$ by adding the
-appropriate 2{,}457{,}000 day offset
-\citep{tess_data_product_description_2018}.  Many of these and
+%FIXME: these numbers
+%FIXME: gotta mask orbit edges too...
+Both light-curves showed a clear discontinuous ``jump'' in the last
+few days, which seemed likely to be an instrumental systematic.
+We masked out HOW MUCH DATA(?).
+As an added precaution, we masked out the first and last N hours
+of each orbit, since there is often
+systematic red noise during those times.
+Finally, we removed all points with non-zero quality flags \citep[see][]{tess_data_product_description_2018}.
+
+%FIXME: these numbers
+All told, we removed N\% of the original data points, and were left
+with NN{,}NNN measurements of the relative flux of PTFO 8-8695.  We
+normalized the data by dividing out the median flux, and then
+subtracted out the mean.  Many of these and
 subsequent processing steps were performed using
-\texttt{astrobase}~\citep{bhatti_astrobase_2018}. We did not
-``flatten'' the lightcurves, as is often done with splines,
-polynomials, or Gaussian processes.  Instead, we modeled the
-out-of-transit flux variations simultaneously with the transit
-parameters, as described below.
+\texttt{astrobase}~\citep{bhatti_astrobase_2018}.
+
+\subsection{Visual Inspection}
+
+Our initial inspection of the light-curve, in both its 2-minute PDCSAP
+and 30-minute FFI forms, revealed an obvious beat signal
+(Figure~\ref{fig:splitsignal}, top panel).
+
+As a precursor to more detailed analysis, we calculated generalized
+Lomb-Scargle periodograms using \texttt{astrobase}
+\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
+The two largest peaks in the Lomb-Scargle periodogram of the
+light-curve were clearly separated at $P_{\rm s} \approx 0.448\,{\rm
+days}$ and $P_{\rm \ell} \approx 0.499\,{\rm days}$.  The $P_{\rm
+\ell}$ peak had the greater power of the two.  Smaller harmonics from
+each of these dominant two signals were also present.
+
+% 0.14 = A1+A2
+% 0.06 = A1-A2
+% 2A1 = 0.20
+% -> A1 = 0.1
+% 2A2 = 0.08
+% -> A2 = 0.04
+
+The beat peak-to-peak amplitude at maximum, when the two signals
+constructively interfere, is about 14\%.  At minimum, the peak-to-peak
+amplitude is about 6\%.  Assuming the signals are just two sinusoids,
+algebra tells us that the peak-to-peak amplitudes should therefore be
+10\% for the long-period signal, and 4\% for the short-period signal.
+These numbers will turn out to be over-estimates, but only slightly.
+
+Initial signal-processing experiments fitting out splines or sinusoids
+showed that after subtracting out the long-period signal, the short-period
+signal dominated the periodogram, and vice-versa.
+However it quickly became clear that it would be beneficial to simultaneously
+model the signals separately, in order to
+preserve the power at each frequency.
 
@@
 
-\section{Analysis}
+\section{The Model}
 \label{sec:analysis}
@@
 
-\subsection{Visual Inspection}
-
-Our initial inspection of the light-curve, in both its 2-minute PDCSAP
-and 30-minute FFI forms, revealed an obvious beat signal
-(Figure~\ref{fig:splitsignal}, top panel).
-
-As a precursor to more detailed analysis, we calculated generalized
-Lomb-Scargle periodograms using \texttt{astrobase}
-\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}. 
-The two largest peaks in the Lomb-Scargle periodogram of the
-light-curve were clearly separated at $P_{\rm s} \approx 0.448\,{\rm
-days}$ and $P_{\rm \ell} \approx 0.499\,{\rm days}$.
-The $P_{\rm \ell}$ peak had the larger ampitude of the two.
-Smaller harmonics from each of these dominant two signals were also
-present.
-
-Initial fitting experiments with splines showed that subtracting out
-the long-period signal, the short-period signal dominated the
-periodogram, and vice-versa.
-However it quickly became clear that a mroe careful modelling approach
-than ``flattening'' the light-curve with splines or Gaussian proceses
-would be preferable, in order to preserve the power at each frequency.
-
 \subsection{Model Fitting}
@@
 We opted to model the light-curve as a linear combination of Fourier
-harmonics at the short and long periods, plus a transit at the
-short period.
-Symbolically,
+harmonics at the short and long periods, plus a transit at the short
+period.  Symbolically,
 \begin{equation}
-  f = f_{\rm s} + f_{\rm \ell} = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
+  f = f_{\rm s} + f_{\rm \ell}
+  = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
 \end{equation}
 where $f_{\rm s}$ is the relative flux at the short period, and
-$f_{\rm \ell}$ is the flux at the long period.
-Decomposing each term,
+$f_{\rm \ell}$ is the flux at the long period.  Writing out the
+Fourier terms,
 \begin{align}
@@
 \end{align}
-for $N$ and $M$ the total number of harmonics at the short and long period, respective,
-$A_i$ and $B_i$ for each harmonic term, and $\omega_i = 2\pi / P_i$ the angular
-frequency for $i$ the short or long period index. 
-We fixed the ``phase-offset'' for the short period signal to be zero,
-and let the reference time of the long period signal float by introducing
-$\phi_{\rm \ell}$. 
-Since we did not a priori know how many harmonics would be
-appropriate, we let the number of harmonics vary, and
-used the Bayesian information criterion to choose the
-appropriate model.
+for $N$ and $M$ the total number of harmonics at the short and long
+period, respective, $A_i$ and $B_i$ for each harmonic term, and
+$\omega_i = 2\pi / P_i$ the angular frequency for $i$ the short or
+long period index.  We fixed the ``phase-offset'' for the short period
+signal to be zero, and let the reference time for the long period
+signal float by introducing $\phi_{\rm \ell}$.  Since we did not a
+priori know how many harmonics would be appropriate, we considered a
+number of different choices for $N$ and $M$, and used the Bayesian
+information criterion to choose the appropriate model.
 
-For example, one possible model could be a transit, plus $N=2$ harmonics
-of sines and cosines at the short period, plus $M=1$  harmonics at the long
-period.
-In this case, the free parameters would be as follows.
-For the transit, we would fit for
-the impact parameter, the planet-to-star radius ratio, two quadratic limb darkening parameters,
-the planet orbital period (equal to the short period), the reference time
-for the transit, and the mean flux.
-There would be 4 additional Fourier amplitudes at the short period,
-plus 2 Fourier amplitudes at the long period, and well as the long period itself
-and its phase.
-For this case, we therefore fitted 14 free parameters.
+As an example, one possible model could be a transit, plus $N=2$
+harmonics of sines and cosines at the short period, plus $M=1$
+harmonics at the long period.  In this case, the free parameters would
+be as follows.  For the transit, we would fit for the impact
+parameter, the planet-to-star radius ratio, two quadratic limb
+darkening parameters, the planet orbital period (equal to the short
+period), the reference time for the transit, and the mean flux.  There
+would be 4 additional Fourier amplitudes at the short period, plus 2
+Fourier amplitudes at the long period, and well as the long period
+itself and its phase.  For this case, we therefore fitted 14 free
+parameters.
+
+We implemented the models using \texttt{PyMC3}, which is built on \texttt{theano} \citep{salvatier_2016_PyMC3,exoplanet:theano}.
+For the Fourier terms, we used the default math operators.
+For the exoplanet transit, we used the model and derivatives implemented in
+\texttt{exoplanet} \citep{exoplanet:exoplanet}.
+
+Our priors were as follows.
+%FIXME
+
+We initialized to the maximum a posteriori (MAP) solution.
+We then sampled using \texttt{PyMC3}'s gradient-based No-U-Turns Sampler, 
+and used R-hat as our convergence diagnostic.
 
@@
 
-\section{Discussion}
+\section{Interpretation}
 \label{sec:discussion}
@@
   \texttt{corner} \citep{corner_2016},
-  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
+  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
+  \texttt{exoplanet} \citep{exoplanet:agol19}
+  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
+  dependencies \citep{exoplanet:agol19, exoplanet:kipping13, exoplanet:luger18,
+  	exoplanet:theano}.
   \texttt{IPython} \citep{perez_2007},
@@
   \texttt{pandas} \citep{mckinney-proc-scipy-2010},
+  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
   \texttt{radvel} \citep{fulton_radvel_2018},

### 2020-04-13 billy/convenience.py
billy.convenience: better data cleaning

@@ -4,7 +4,7 @@ from collections import OrderedDict
 from astropy.io import fits
 
 from astrobase.lcmath import time_bin_magseries_with_errs
-
+from cdips.lcproc.mask_orbit_edges import mask_orbit_start_and_end
 
 def chisq(y_mod, y_obs, y_err):
     return np.sum( (y_mod - y_obs )**2 / y_err**2 )
@@ -89,17 +89,53 @@ def get_ptfo_data(cdips=1, spoc=0):
 
 def get_clean_ptfo_data(binsize=120*5):
     """
-    get data. quality cut and remove weird end points. bin to 10 minutes, to
+    get data. mask orbit edges... quality cut and remove weird end points. bin to 10 minutes, to
     speed fitting (which is linear in time).
     """
 
     d = get_ptfo_data(cdips=0, spoc=1)[0]
+
+    time = d.TIME
+    flux = d.PDCSAP_FLUX
+    flux_err = d.PDCSAP_FLUX_ERR
+
+    N_i = len(time) # initial
+
     quality = d.QUALITY
-    x_obs = d.TIME - 1468.2
-    sel = (quality == 0) & (x_obs < 20.1)
-    x_obs = d.TIME[sel] - 1468.2
-    y_obs = (d.PDCSAP_FLUX[sel] / np.nanmedian(d.PDCSAP_FLUX[sel])) - 1
-    y_err = d.PDCSAP_FLUX_ERR[sel] / np.nanmedian(d.PDCSAP_FLUX[sel])
+    time, flux, flux_err = (
+        time[quality == 0], flux[quality == 0], flux_err[quality == 0]
+    )
+
+    N_ii = len(time) # after quality cut
+
+    time, flux, sel = mask_orbit_start_and_end(time, flux, orbitgap=0.5,
+                                               expected_norbits=2,
+                                               orbitpadding=6/(24),
+                                               raise_expectation_error=True,
+                                               return_inds=True)
+    flux_err = flux_err[sel]
+
+    N_iii = len(time) # after orbit edge masking
+
+    # 2457000 + 1488.3 = 2458488.3
+    sel = (x_obs < 1488.3)
+
+    x_obs = x_obs[sel]
+
+    time_offset = 1468.2
+    x_obs -= time_offset
+
+    N_iv = len(x_obs) # after dropping end of orbit 20
+
+    y_obs = (flux[sel] / np.nanmedian(flux[sel])) - 1
[…]

### 2020-04-13 paper/ms.tex
update paper to describe data cleaning

@@
   TESS recently observed PTFO$\,$8-8695 for one month.
-  The TESS light-curve shows that the dominant variability in this
+  The TESS lightcurve shows that the dominant variability in this
   system is a sinusoidal modulation with a ``long'' period $P_{\rm
@@
   Planets do not ``jump'' in orbital phase.
-  PTFO$\,$8-8695 therefore seems broadly consistent with the
-  ``transient dipping'' phenomenology observed in many young M dwarfs,
-  and seems unlikely to be a planet.
+  Given the available evidence,
+  PTFO$\,$8-8695 seems consistent with the
+  ``transient dipping'' phenomenology observed in many young M dwarfs.
+  It seems rather unlikely to be a planet.
 \end{abstract}
@@
 \section{Introduction}
-If it were a planet, PTFO$\,$8-8695b would be exceptional.
+If PTFO$\,$8-8695b were a planet, it would be exceptional.
 A transiting hot Jupiter, orbiting a $\approx$3$\,$Myr old M dwarf
@@
 operations.  The star is designated TIC 264461976 in the TESS Input
-Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.  The pixel data for an $11\times11$
-array surrounding PTFO 8-8695 were averaged into 2-minute stacks by
-the onboard computer. 
-Each 2048$\times$2048 image from the CCD was also averaged into
-30-minute stacks, and saved as a ``full frame image'' (FFI).
+Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.  The pixel data
+for an $11\times11$ array surrounding PTFO 8-8695 were averaged into
+2-minute stacks by the onboard computer.  Each 2048$\times$2048 image
+from the CCD was also averaged into 30-minute stacks, and saved as a
+``full frame image'' (FFI).
 
@@
 Ames~\citep{jenkins_tess_2016}.  Our main analysis used the resulting
-Presearch Data Conditioning (PDC) lightcurve, which had
-non-astrophysical variability removed through the methods discussed by
-\citet{smith_kepler_apertures_2017} and \citet{smith_kepler_PDC_2017}.
+Presearch Data Conditioning (PDC) lightcurve.  The PDC lightcurve
+aperture used pixels chosen to maximize the SNR of the total flux of
+the target \citep{smith_kepler_apertures_2017}.  Non-astrophysical
+variability was removed through the methods discussed by
+\citet{smith_kepler_PDC_2017}.
 
-We also processed the 30-minute image stacks separately, as part of
-the Cluster Difference Imaging
-Photometric Survey (CDIPS; \citep{bouma_cluster_2019}).
-The raw CDIPS light-curve shows identical features as the 2-minute
-light-curve, though at 30-minute cadence.
+As an independent check on the shorter cadence SPOC light-curve, we
+separately processed the 30-minute image stacks as part of the Cluster
+Difference Imaging Photometric Survey (CDIPS;
+\citep{bouma_cluster_2019}).  The CDIPS lightcurve used a circular
+aperture with radius 1 pixel.
 
-%FIXME: these numbers
-%FIXME: gotta mask orbit edges too...
-Both light-curves showed a clear discontinuous ``jump'' in the last
-few days, which seemed likely to be an instrumental systematic.
-We masked out HOW MUCH DATA(?).
-As an added precaution, we masked out the first and last N hours
-of each orbit, since there is often
-systematic red noise during those times.
-Finally, we removed all points with non-zero quality flags \citep[see][]{tess_data_product_description_2018}.
+For data cleaning, we removed all points with non-zero quality flags
+\citep[{\it e.g.},][]{tess_data_product_description_2018}.  We also
+masked out the first and last 6 hours of each orbit, since there is
+often systematic red noise during those times.  Both the CDIPS and PDC
+lightcurves showed a clear discontinuous ``jump'' in the last few days
+of orbit 20, which seemed likely to be an instrumental systematic.  We
+correspondingly masked out times from BJD 2458488.3 until the end of
+the orbit.  The PDC lightcurve initially had 15{,}678 points.  The
+quality cut removed 854, masking the orbit edges removed 716, and
+removing the final few days of orbit 20 removed 1079.  83\% of the
+highest-quality data remained.
 
-%FIXME: these numbers
-All told, we removed N\% of the original data points, and were left
-with NN{,}NNN measurements of the relative flux of PTFO 8-8695.  We
-normalized the data by dividing out the median flux, and then
-subtracted out the mean.  Many of these and
-subsequent processing steps were performed using
+We normalized these points by dividing out the median flux. We then
+subtracted by unity to simplify subsequent analysis.  Many of these
+and subsequent processing steps were performed using
 \texttt{astrobase}~\citep{bhatti_astrobase_2018}.
@@
 
-Our initial inspection of the light-curve, in both its 2-minute PDCSAP
+Our initial inspection of the lightcurve, in both its 2-minute PDCSAP
 and 30-minute FFI forms, revealed an obvious beat signal
@@
 The two largest peaks in the Lomb-Scargle periodogram of the
-light-curve were clearly separated at $P_{\rm s} \approx 0.448\,{\rm
+lightcurve were clearly separated at $P_{\rm s} \approx 0.448\,{\rm
 days}$ and $P_{\rm \ell} \approx 0.499\,{\rm days}$.  The $P_{\rm
@@
 10\% for the long-period signal, and 4\% for the short-period signal.
-These numbers will turn out to be over-estimates, but only slightly.
+These order-of-magnitude numbers will turn out to be roughly correct.
 
 Initial signal-processing experiments fitting out splines or sinusoids
-showed that after subtracting out the long-period signal, the short-period
-signal dominated the periodogram, and vice-versa.
-However it quickly became clear that it would be beneficial to simultaneously
-model the signals separately, in order to
-preserve the power at each frequency.
+showed that after subtracting out the long-period signal, the
+short-period signal dominated the periodogram, and vice-versa.
+However it quickly became clear that it would be beneficial to
+simultaneously model the signals separately, in order to preserve the
+power at each frequency.
 
@@
 	\vspace{-0.7cm}
-	\caption{ {\bf TESS light-curve of PTFO 8-8695 (Sector 6, Orbit 1).}
+	\caption{ {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 1).}
 	{\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux versus time. The beat period of 4.48 days is visible by eye.
@@
 
-We opted to model the light-curve as a linear combination of Fourier
+We opted to model the lightcurve as a linear combination of Fourier
 harmonics at the short and long periods, plus a transit at the short
@@
   &+ \sum_{m=1}^{m=M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
-  + \sum_{m=1}^{m=M} B_m \cos([m\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
+  + \sum_{m=1}^{m=M} B_m \cos(m[\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
 \end{align}
@@
 measure the background level are indicated with an `\texttt{X}' hatch,
-and the pixels used for the final light-curve are shown with the
+and the pixels used for the final lightcurve are shown with the
 `\texttt{/}' hatch.
@@
 
-First, we examined the CDIPS full frame image light-curves of the
+First, we examined the CDIPS full frame image lightcurves of the
 target, which are available on MAST \citep{bouma_cluster_2019}.
@@
 
-Second, we examined the light-curve of each pixel in the scene
+Second, we examined the lightcurve of each pixel in the scene
 individually.  We opted to use the

### 2020-04-13 billy/convenience.py
fix da bug

@@ -118,9 +118,9 @@ def get_clean_ptfo_data(binsize=120*5):
     N_iii = len(time) # after orbit edge masking
 
     # 2457000 + 1488.3 = 2458488.3
-    sel = (x_obs < 1488.3)
+    sel = (time < 1488.3)
 
-    x_obs = x_obs[sel]
+    x_obs = time[sel]
 
     time_offset = 1468.2
     x_obs -= time_offset

### 2020-04-13 billy/convenience.py
splitsignal plot: add part i vs part ii

@@ -124,6 +124,7 @@ def get_clean_ptfo_data(binsize=120*5):
 
     time_offset = 1468.2
     x_obs -= time_offset
+    # reverse offset: 2457000 + 1468.2 = 2458468.2
 
     N_iv = len(x_obs) # after dropping end of orbit 20
 
@@ -143,6 +144,8 @@ def get_clean_ptfo_data(binsize=120*5):
         x_obs = bd['binnedtimes']
         y_obs = bd['binnedmags']
         y_err = bd['binnederrs']
+
+    assert len(x_obs) == len(y_obs) == len(y_err)
 
     return (
         x_obs.astype(np.float64),

### 2020-04-13 billy/plotting.py

@@ -127,7 +127,7 @@ def plot_sampleplot(m, outpath, N_samples=100):
     savefig(fig, outpath, writepdf=0, dpi=300)
 
 
-def plot_splitsignal_map(m, outpath):
+def plot_splitsignal_map(m, outpath, part='i'):
     """
     y_obs + y_MAP + y_rot + y_orb
     things at rotation frequency
@@ -183,7 +183,21 @@ def plot_splitsignal_map(m, outpath):
         #a.legend()
         format_ax(a)
         a.set_ylim((-.075, .075))
-        a.set_xlim((0, 9))
+        if part == 'i':
+            a.set_xlim((0, 9))
+        else:
+            a.set_xlim((10, 20.3))
+
+    props = dict(boxstyle='square', facecolor='white', alpha=0.9, pad=0.15,
+                 linewidth=0)
+    if part == 'i':
+        axs[3].text(0.97, 0.03, 'Orbit 19', ha='right', va='bottom',
+                    transform=axs[3].transAxes, bbox=props, zorder=3,
+                    fontsize='x-large')
+    else:
+        axs[3].text(0.97, 0.03, 'Orbit 20', ha='right', va='bottom',
+                    transform=axs[3].transAxes, bbox=props, zorder=3,
+                    fontsize='x-large')
 
     fig.tight_layout(h_pad=0., w_pad=0.)
     savefig(fig, outpath, writepdf=1, dpi=300)

### 2020-04-13 drivers/real.py
ditto

@@ -4,6 +4,7 @@ Fit data for "transit_NsincosPorb_NsincosProt" model.
 import os
 import numpy as np, pandas as pd, matplotlib.pyplot as plt, pymc3 as pm
 from os.path import join
+from itertools import product
 
 from billy.modelfitter import ModelFitter, ModelParser
 import billy.plotting as bp
@@ -15,14 +16,14 @@ from billy import __path__
 def main(modelid):
 
     traceplot = 0
-    sampleplot = 0
-    cornerplot = 0
+    sampleplot = 1
+    cornerplot = 1
     splitsignalplot = 1 if 'Porb' in modelid and 'Prot' in modelid else 0
 
     REALID = 'PTFO_8-8695'
     RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
     PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID),
-                           '20200320_linear_amplitude_prior')
+                           '20200413_v0')
 
     ##########
 
@@ -60,8 +61,10 @@ def main(modelid):
             outpath = join(PLOTDIR, '{}_{}_phasefoldpost.png'.format(REALID, modelid))
             bp.plot_phasefold_post(m, ydict, outpath)
         if do_map:
-            outpath = join(PLOTDIR, '{}_{}_splitsignalmap.png'.format(REALID, modelid))
-            ydict = bp.plot_splitsignal_map(m, outpath)
+            outpath = join(PLOTDIR, '{}_{}_splitsignalmap_i.png'.format(REALID, modelid))
+            ydict = bp.plot_splitsignal_map(m, outpath, part='i')
+            outpath = join(PLOTDIR, '{}_{}_splitsignalmap_ii.png'.format(REALID, modelid))
+            ydict = bp.plot_splitsignal_map(m, outpath, part='ii')
             outpath = join(PLOTDIR, '{}_{}_splitsignalmap_periodogram.png'.format(REALID, modelid))
             bp.plot_splitsignal_map_periodogram(ydict, outpath)
             outpath = join(PLOTDIR, '{}_{}_phasefoldmap.png'.format(REALID, modelid))
@@ -76,7 +79,13 @@ def main(modelid):
 
 
 if __name__ == "__main__":
-    main('transit_2sincosPorb_2sincosProt')
     # main('transit_1sincosPorb_1sincosProt')
+
+    # TODO:
[…]

### 2020-04-14 billy/convenience.py
update billy.convenience for N,M model

@@ -27,16 +27,13 @@ def get_bic(m, ydict):
     χ2 = chisq(y_mod, y_obs, y_err)
 
     k = None
-    if m.modelid == 'transit_1sincosPorb_1sincosProt':
-        k = 13
-    if m.modelid == 'transit_2sincosPorb_1sincosProt':
-        k = 15
-    if m.modelid == 'transit_1sincosPorb_2sincosProt':
-        k = 15
-    if m.modelid == 'transit_2sincosPorb_2sincosProt':
-        k = 17
-    if k is None:
-        raise NotImplementedError
+    _N = int(m.modelid.split('_')[1][0])
+    _M = int(m.modelid.split('_')[2][0])
+
+    k_tra = 7
+    k_Porb = 2*_N # amplitudes
+    k_Prot = 2*_M + 2 # amplitudes, plus period and phase
+    k = k_tra + k_Porb + k_Prot
 
     n = len(y_obs)
     BIC = bic( χ2, k, n )
@@ -194,6 +191,9 @@ def initialize_ptfo_prior_d(x_obs, modelcomponents):
                     elif ix == 1:
                         prior_d['A{}{}'.format(k,ix)] = 1e-2
                         prior_d['B{}{}'.format(k,ix)] = 1e-2
+                    elif ix == 2:
+                        prior_d['A{}{}'.format(k,ix)] = 1e-2
+                        prior_d['B{}{}'.format(k,ix)] = 1e-2
                     else:
                         raise NotImplementedError
 
@@ -204,6 +204,9 @@ def initialize_ptfo_prior_d(x_obs, modelcomponents):
                     elif ix == 1:
                         prior_d['A{}{}'.format(k,ix)] = 1e-2
                         prior_d['B{}{}'.format(k,ix)] = -1e-2
+                    elif ix == 2:
+                        prior_d['A{}{}'.format(k,ix)] = 1e-2
+                        prior_d['B{}{}'.format(k,ix)] = 1e-2
                     else:
                         raise NotImplementedError
 

### 2020-04-14 billy/modelfitter.py
add overwrite option to ModelFitter

@@ -60,12 +60,13 @@ class ModelFitter(ModelParser):
 
     def __init__(self, modelid, x_obs, y_obs, y_err, prior_d,
                  mstar=1, rstar=1, N_samples=1000, N_cores=16, N_chains=4,
-                 plotdir=None, pklpath=None):
+                 plotdir=None, pklpath=None, overwrite=1):
 
         self.N_samples = N_samples
         self.N_cores = N_cores
         self.N_chains = N_chains
         self.PLOTDIR = plotdir
+        self.OVERWRITE = overwrite
         self.x_obs = x_obs
         self.y_obs = y_obs
         self.y_err = y_err

### 2020-04-14 billy/plotting.py
add overwrite option to plots

@@ -49,6 +49,7 @@ def plot_periodogram(outdir, islinear=True):
 
     P_rot, P_orb = 0.49914, 0.4485
 
+    plt.close('all')
     f, ax = plt.subplots(figsize=(4,3))
     ax.plot(
         period, power, lw=0.5, c='k'
@@ -78,6 +79,7 @@ def plot_periodogram(outdir, islinear=True):
 
 
 def plot_test_data(x_obs, y_obs, y_mod, modelid, outdir):
+    plt.close('all')
     fig = plt.figure(figsize=(14, 4))
     ax = fig.add_subplot(111, xlabel='x_obs', ylabel='y_obs',
                          title='Generated data and underlying model')
@@ -90,6 +92,7 @@ def plot_test_data(x_obs, y_obs, y_mod, modelid, outdir):
 
 
 def plot_MAP_data(x_obs, y_obs, y_MAP, outpath):
+    plt.close('all')
     plt.figure(figsize=(14, 4))
     plt.plot(x_obs, y_obs, ".k", ms=4, label="data")
     plt.plot(x_obs, y_MAP, lw=1)
@@ -101,6 +104,11 @@ def plot_MAP_data(x_obs, y_obs, y_MAP, outpath):
 
 
 def plot_sampleplot(m, outpath, N_samples=100):
+
+    if os.path.exists(outpath) and not m.OVERWRITE:
+        return
+
+    plt.close('all')
     fig, ax = plt.subplots(figsize=(14, 4))
     ax.plot(m.x_obs, m.y_obs, ".k", ms=4, label="data", zorder=N_samples+1)
     ax.plot(m.x_obs, m.map_estimate['mu_model'], lw=0.5, label='MAP',
@@ -134,6 +142,7 @@ def plot_splitsignal_map(m, outpath, part='i'):
     things at orbital frequency
     """
 
+    plt.close('all')
     # 8.5x11 is letter paper. x10 allows space for caption.
     fig, axs = plt.subplots(nrows=4, figsize=(8.5, 10), sharex=True)
 
@@ -200,7 +209,9 @@ def plot_splitsignal_map(m, outpath, part='i'):
                     fontsize='x-large')
 
     fig.tight_layout(h_pad=0., w_pad=0.)
-    savefig(fig, outpath, writepdf=1, dpi=300)
+
+    if not os.path.exists(outpath) or m.OVERWRITE:
+        savefig(fig, outpath, writepdf=1, dpi=300)
 
     ydict = {
         'x_obs': m.x_obs,
[…]

### 2020-04-14 drivers/real.py
add overwrite option to main fitter

@@ -20,6 +20,7 @@ def main(modelid):
     cornerplot = 1
     splitsignalplot = 1 if 'Porb' in modelid and 'Prot' in modelid else 0
 
+    OVERWRITE = 0
     REALID = 'PTFO_8-8695'
     RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
     PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID),
@@ -40,7 +41,7 @@ def main(modelid):
     mp = ModelParser(modelid)
     prior_d = initialize_ptfo_prior_d(x_obs, mp.modelcomponents)
     m = ModelFitter(modelid, x_obs, y_obs, y_err, prior_d, plotdir=PLOTDIR,
-                    pklpath=pklpath)
+                    pklpath=pklpath, overwrite=OVERWRITE)
 
     print(pm.summary(m.trace, varnames=list(prior_d.keys())))
 
@@ -66,7 +67,8 @@ def main(modelid):
             outpath = join(PLOTDIR, '{}_{}_splitsignalmap_ii.png'.format(REALID, modelid))
             ydict = bp.plot_splitsignal_map(m, outpath, part='ii')
             outpath = join(PLOTDIR, '{}_{}_splitsignalmap_periodogram.png'.format(REALID, modelid))
-            bp.plot_splitsignal_map_periodogram(ydict, outpath)
+            if not os.path.exists(outpath) or m.OVERWRITE:
+                bp.plot_splitsignal_map_periodogram(ydict, outpath)
             outpath = join(PLOTDIR, '{}_{}_phasefoldmap.png'.format(REALID, modelid))
             bp.plot_phasefold_map(m, ydict, outpath)
             get_bic(m, ydict)
@@ -79,11 +81,12 @@ def main(modelid):
 
 
 if __name__ == "__main__":
-    # main('transit_1sincosPorb_1sincosProt')
 
-    # TODO:
     for N, M in product(range(1,4), range(1,4)):
         main('transit_{}sincosPorb_{}sincosProt'.format(N,M))
+
+    # NOTE: DEBUG
+    # main('transit_1sincosPorb_1sincosProt')
 
     # NOTE: DEPRECATED
     # main('transit_2sincosPorb_2sincosProt')

### 2020-04-14 paper/ms.tex
paper: many updates

@@
 15, 2018 to January 6, 2019, during the sixth sector of science
-operations.  The star is designated TIC 264461976 in the TESS Input
+operations.  The star was designated TIC 264461976 in the TESS Input
 Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.  The pixel data
@@
 
-For data cleaning, we removed all points with non-zero quality flags
+To clean the data, we removed all points with non-zero quality flags
 \citep[{\it e.g.},][]{tess_data_product_description_2018}.  We also
@@
 the orbit.  The PDC lightcurve initially had 15{,}678 points.  The
-quality cut removed 854, masking the orbit edges removed 716, and
-removing the final few days of orbit 20 removed 1079.  83\% of the
-highest-quality data remained.
+quality cut removed 854 points, masking the orbit edges removed an
+additional 716, and removing the final few days of orbit 20 removed an
+additional 1079.  After cleaning, 83\% of the initial flux
+measurements remained.
 
@@
 Our initial inspection of the lightcurve, in both its 2-minute PDCSAP
-and 30-minute FFI forms, revealed an obvious beat signal
+and 30-minute FFI forms, showed a strong sinusoidal beat signal
 (Figure~\ref{fig:splitsignal}, top panel).
@@
 The two largest peaks in the Lomb-Scargle periodogram of the
-lightcurve were clearly separated at $P_{\rm s} \approx 0.448\,{\rm
-days}$ and $P_{\rm \ell} \approx 0.499\,{\rm days}$.  The $P_{\rm
-\ell}$ peak had the greater power of the two.  Smaller harmonics from
-each of these dominant two signals were also present.
+lightcurve were clearly separated at a ``short'' period $P_{\rm s}
+\approx 0.448\,{\rm days}$ and a ``long'' period $P_{\rm \ell} \approx
+0.499\,{\rm days}$.  The $P_{\rm \ell}$ peak had the greater power of
+the two.  Smaller harmonics from each of these two dominants peaks
+were also present.
 
@@
 
-The beat peak-to-peak amplitude at maximum, when the two signals
+The peak-to-peak amplitude at maximum, when the two signals
 constructively interfere, is about 14\%.  At minimum, the peak-to-peak
@@
 
-% \subsection{Star and planet parameters}
-% \label{sec:system_parameters}
-% 
-% \begin{deluxetable}{lccc}
-% \tabletypesize{\scriptsize}
-% \tablecaption{Selected system parameters of WASP-4b\label{tbl:params}}
-% \tablenum{1}
-% 
-% \tablehead{
-% \colhead{Parameter} & \colhead{Value} & \colhead{68\% Confidence Interval} & \colhead{Comment}
-% }
-% 
-% \startdata
-% {\it Transit/RV parameters:} & & & \\
-%   $R_{\rm p}/R_\star$                        & $0.15201$              & $+0.00040$, $-0.00033$      & A \\
-%   $i$~[deg]                                  & $89.06$                & $+0.65$, $-0.84$            & A \\
-%   $a/R_\star$                                & $5.451$                & $+0.023$, $-0.052$          & A \\
-%   $u_{\rm linear}$                           & $0.382$                & ---                         & A \\
-%   $u_{\rm quad}$                             & $0.210$                & ---                         & A \\
-%   $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & B \\
-% {\it Stellar parameters:} & & & \\
-%   $T_{\rm eff}$~[K]                          & $5400$                 & $\pm 90$                    & C \\
-%   $\log g_\star$~[cgs]                       & $4.47$                 & $\pm 0.11$                  & C \\
-%   $[{\rm Fe/H}]$                             & $-0.07$                & $\pm 0.19$                  & C \\
-%   $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & D \\
-%   $A_V$~[mag]                                & $0.03$                 & $+0.02, -0.01$              & D \\
-%   $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & F \\
-%   $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & E \\
-%   $\rho_\star$~[g~cm$^{-3}$]                 & $1.711$                & $+0.022$, $-0.048$          & E \\
-%   $M_\star$~[M$_{\odot}$]                    & $0.864$                & $+0.084$, $-0.090$          & E \\
-%   $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
-% {\it Planetary parameters:} & & & \\
-%   $a$~[AU]                                   & $0.0226$               & $+0.0007$, $-0.0008$        & E \\
-%   $M_p$~[M$_{\rm Jup}$]                      & $1.186$                & $+0.090$, $-0.098$          & E \\
-%   $R_p$~[R$_{\rm Jup}$]                      & $1.321$                & $\pm 0.039$                 & E \\
-% \enddata
-% 
-% \tablecomments{
-%   (A) From phase-folded TESS lightcurve (\S~\ref{sec:measurement}).
-%   Orbital periods are in Table~4. The limb darkening parameters were
-%   allowed to float around the \citet{claret_limb_2017} prediction, but
-%   were unconstrained.
-%   (B) \citet{triaud_spin-orbit_2010}.
-%   (C) From HARPS spectra \citep{doyle_accurate_2013}.
-%   (D) \citet{stassun_accurate_2017}.
-%   (E) This work, see \S~\ref{sec:system_parameters}.
-%   (F) \citet{gaia_collaboration_gaia_2018}.
-%   (G) \citet{stassun_TIC_2018}.
-% }
-% 
-% \vspace{-1cm}
-% \end{deluxetable}
-% 
-% We calculated the stellar and planetary parameters in the following
-% way.  We computed the star's spectral energy distribution based on the
-% {\it Gaia} DR2 parallax (after making the small correction advocated
-% by \citealt{stassun_evidence_2018}) and the broadband magnitudes from
-% the available all-sky catalogs: $G$ from {\it Gaia\/} DR2, $B_T$ and
-% $V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from {\it
-% 2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
-% wavelength range 0.4--22~$\mu$m.
-% We adopted the effective temperature
-% from the work by \citet{doyle_accurate_2013}, who determined the
-% spectroscopic parameters of WASP-4 using high-signal-to-noise
-% observations with the High Accuracy Radial-velocity Planet Searcher
-% (HARPS).  Then, we determined the stellar radius through the
-% combination of the bolometric luminosity and the effective
-% temperature, using the Stefan-Boltzmann law.  To determine the stellar
-% mass, we first computed the mean stellar density based on the value of
-% $a/R_\star$ that gave the best fit to the phase-folded TESS lightcurve
-% (for the relevant equation, see \citealt{seager_unique_2003} or
-% \citealt{winn_exoplanet_2010}).  The mass was calculated from the
-% radius and density, and the orbital distance was also calculated from
-% the radius and $a/R_\star$.  The planetary radius was calculated as
-% the product of $R_\star$ and $R_{\rm p}/R_\star$.  Finally, the planet
-% mass was calculated based on the stellar mass, the radial-velocity
-% amplitude observed by \citet{triaud_spin-orbit_2010}, and the orbital
-% inclination.
-% 
-% Table~1 gives the resulting parameters, which we adopted for the
-% remaining analysis.  \added{The uncertainties in our derived stellar
-% and planetary parameters are propagated according to standard analytic
-% formulae, under the assumption that the variables are uncorrelated and
-% normally distributed.} Our \added{system} parameters are in agreement
-% with those of previous investigators, but have the benefit of
-% incorporating the {\it Gaia} parallax
-% \citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013,huitson_gemini_2017}.
-% By comparing the star's luminosity and spectroscopic parameters with
-% the outputs of the Yonsei-Yale stellar-evolutionary models, we found
-% that WASP-4 is a main-sequence star with an age of approximately
-% $7\,{\rm Gyr}$ \added{\citep{demarque_y2_2004}}.
-
 \section{The Model}
@@
 		\leavevmode
-		\includegraphics[width=0.98\textwidth]{f1.png}
+		\includegraphics[width=0.98\textwidth]{f1.pdf}
 	\end{center}
 	\vspace{-0.7cm}
-	\caption{ {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 1).}
-	{\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux versus time. The beat period of 4.48 days is visible by eye.
-	The model plotted underneath the data includes 2 harmonics at the long
-	period $P_{\rm \ell}$, plus 2 harmonics and a transit at the short period $P_{\rm s}$.
-	{\it Upper middle}: Long-period signal, equal to the raw signal minus the short-period signal.
-	{\it Lower middle}: Short-period signal, equal to the raw signal minus the long-period signal.
-	{\it Bottom}: residual.
-	The data are binned from 2 to 10 minute cadence as a convenience for plotting and fitting.
+	\caption{
+  {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 19).}
+  {\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux
+  versus time. The beat period of 4.48 days is visible by eye.  The
+  model plotted underneath the data includes 2 harmonics at the long
+  period $P_{\rm \ell}$, plus 2 harmonics and a transit at the short
+  period $P_{\rm s}$.
+  {\it Upper middle}: Long-period signal, equal to the raw signal
+  minus the short-period signal.
+  {\it Lower middle}: Short-period signal, equal to the raw signal
+  minus the long-period signal.
+  {\it Bottom}: residual.  The data are binned from 2 to 10 minute
+  cadence as a convenience for plotting and fitting.
   \label{fig:splitsignal}
@@
 		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f2.pdf}
+		\includegraphics[width=0.98\textwidth]{f2.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 20).}
+  Panels are as in Figure~\ref{fig:splitsignal}.
+  \label{fig:splitsignalii}
+	}
+\end{figure*}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.95\textwidth]{f3.pdf}
 	\end{center}
@@
 	{\it Top}: Long-period signal, as in Figure~\ref{fig:splitsignal}.
-	{\it Bottom}: Short-period signal. The reference phase is set to the ``planetary'' dip.
-	Gray points are the 10 minute cadence \texttt{PDCSAP} flux.
-	Black points are binned to 100 points per period.
+  {\it Bottom}: Short-period signal. The reference phase is set to the
+  ``planetary'' dip.  Gray points are the 10 minute cadence
+  \texttt{PDCSAP} flux.  Black points are binned to 100 points per
+  period.
 	\label{fig:phasefold}
@@
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f3.pdf}
+		\includegraphics[width=0.48\textwidth]{f4.pdf}
 	\end{center}
@@
 	\caption{ {\bf Scene used for blend analysis.}
-		{\it Top:} Mean TESS image of PTFO 8-8695 over Sector~6, with a log-stretch.
-		The position of PTFO 8-8695 is shown with a yellow star.
-		Neighbors with $T<17$ are shown with orange crosses.
-		The apertures used to measure the background and target star flux
-		are shown with \texttt{X} and \texttt{/} hatches, respectively.
-		{\it Bottom:} Digitized Sky Survey $R$-band image of the same field, with a linear stretch. The circles show
-		apertures of radii 1, 1.5, and 2.25 pixels used in part of our blend
-		analysis.
-		The pixel level TESS data show that ``Star A''  does not contribute variability at either of the two observed periods (see Section~\ref{subsec:blend}).
+    {\it Top:} Mean TESS image of PTFO 8-8695 over Sector~6, with a
+    log-stretch.  The position of PTFO 8-8695 is shown with a yellow
+    star.  Neighbors with $T<17$ are shown with orange crosses.  The
+    apertures used to measure the background and target star flux are
+    shown with \texttt{X} and \texttt{/} hatches, respectively.
+    {\it Bottom:} Digitized Sky Survey $R$-band image of the same
+    field, with a linear stretch. The circles show apertures of radii
+    1, 1.5, and 2.25 pixels used in part of our blend analysis.  The
+    pixel level TESS data show that ``Star A''  does not contribute
+    variability at either of the two observed periods (see
+    Section~\ref{subsec:blend}).
 		\label{fig:scene}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f4.pdf}
+		\includegraphics[width=0.9\textwidth]{f5.pdf}
 	\end{center}
@@
     {\bf Timing residuals for PTFO 8-8695b from a decade of monitoring.}
-		Black points are times of ``dips'', minus the indicated linear ephemeris.
-    The $y$-axis is given in units of phase for the
-    short-period signal.
-    The star shows the binned TESS ephemeris.
-		``Dips'' have been observed by
-		\citet{van_eyken_ptf_2012}, \citet{ciardi_follow-up_2015},
-		\citet{yu_tests_2015}, \citet{raetz_yeti_2016},
-		\citet{onitsuka_multi-color_2017}, and \citet{tanimoto_evidence_2020}.
-    Certain dips ({\it e.g.}, the one at phase 0 in mid-2019) are
-    consistent with noise, and were likely reported because
-    something was {\it expected}, rather than convincingly {\it
-    observed}.
-    Horizontal dashed lines are drawn at $\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$,
-    highlighting what could be either a numerical coincidence or an observational
-    bias.
-    The orbital phase observed by TESS is consistent with that of \citet{tanimoto_evidence_2020}, and quite different from the original phase.
+    Black points are times of ``dips'', minus the indicated linear
+    ephemeris.  The $y$-axis is given in units of phase for the
+    short-period signal.  The star shows the binned TESS ephemeris.
+    ``Dips'' have been observed by \citet{van_eyken_ptf_2012},
+    \citet{ciardi_follow-up_2015}, \citet{yu_tests_2015},
+    \citet{raetz_yeti_2016}, \citet{onitsuka_multi-color_2017}, and
+    \citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
+    at phase 0 in mid-2019) are consistent with noise, and were likely
+    reported because something was {\it expected}, rather than
+    convincingly {\it observed}.  Horizontal dashed lines are drawn at
+    $\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting what
+    could be either a numerical coincidence or an observational bias.
+    The orbital phase observed by TESS is consistent with that of
+    \citet{tanimoto_evidence_2020}, and quite different from the
+    original phase.
 		\label{fig:o_minus_c}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f5_comp.png}
+		\includegraphics[width=0.9\textwidth]{f6_comp.png}
 	\end{center}
@@
 
-\subsection{Model Fitting}
+\subsection{Model Description}
 
@@
 harmonics at the short and long periods, plus a transit at the short
-period.  Symbolically,
+period.  Symbolically, the total flux $f$ is given as
 \begin{equation}
@@
 \begin{align}
-  f = &f_{\rm transit,s} + \sum_{n=1}^{n=N} A_n \sin(n\omega_{\rm s}t)
-  + \sum_{n=1}^{n=N} B_n \cos(n\omega_{\rm s}t)\\
-  &+ \sum_{m=1}^{m=M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
-  + \sum_{m=1}^{m=M} B_m \cos(m[\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
+  f = &f_{\rm transit,s} + \sum_{n=1}^{N} A_n \sin(n\omega_{\rm s}t)
+  + \sum_{n=1}^{N} B_n \cos(n\omega_{\rm s}t)\\
+  &+ \sum_{m=1}^{M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
+  + \sum_{m=1}^{M} B_m \cos(m[\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
 \end{align}
 for $N$ and $M$ the total number of harmonics at the short and long
-period, respective, $A_i$ and $B_i$ for each harmonic term, and
-$\omega_i = 2\pi / P_i$ the angular frequency for $i$ the short or
-long period index.  We fixed the ``phase-offset'' for the short period
-signal to be zero, and let the reference time for the long period
-signal float by introducing $\phi_{\rm \ell}$.  Since we did not a
-priori know how many harmonics would be appropriate, we considered a
-number of different choices for $N$ and $M$, and used the Bayesian
-information criterion to choose the appropriate model.
+periods, respectively, $A_i$ and $B_i$ the amplitudes for each
+harmonic term (potentially negative), and $\omega_i = 2\pi / P_i$ the
+angular frequency for $i$ the short or long period index.  We fixed
+the ``phase-offset'' for the short period signal to be zero, and let
+the reference time for the long period signal float by introducing
+$\phi_{\rm \ell}$.  Since we did not a priori know how many harmonics
+would be appropriate, we considered a number of different choices for
+$N$ and $M$, and used the Bayesian information criterion to choose the
+appropriate model (Table~\ref{tab:modelcompare}).
 
@@
 period), the reference time for the transit, and the mean flux.  There
-would be 4 additional Fourier amplitudes at the short period, plus 2
-Fourier amplitudes at the long period, and well as the long period
-itself and its phase.  For this case, we therefore fitted 14 free
-parameters.
+would be $2N=4$ additional Fourier amplitudes at the short period,
+plus $2M=2$ Fourier amplitudes at the long period, and well as the
+long period itself and its phase.  For this case, we therefore fitted
+14 free parameters.
 
-We implemented the models using \texttt{PyMC3}, which is built on \texttt{theano} \citep{salvatier_2016_PyMC3,exoplanet:theano}.
-For the Fourier terms, we used the default math operators.
-For the exoplanet transit, we used the model and derivatives implemented in
-\texttt{exoplanet} \citep{exoplanet:exoplanet}.
+We implemented and fitted the models using \texttt{PyMC3}, which is
+built on \texttt{theano}
+\citep{salvatier_2016_PyMC3,exoplanet:theano}.  For the Fourier terms,
+we used the default math operators.  For the exoplanet transit, we
+used the model and derivatives implemented in \texttt{exoplanet}
+\citep{exoplanet:exoplanet}.  Our priors are listed in
+Table~\ref{tab:posterior}.  To speed up the fitting, we binned the
+cleaned 2 minute lightcurves to 10 minute bins.  We correspondingly
+scaled the uncertainties in the flux measurements by a factor of
+$\sqrt{5}$.  Before sampling, we initialized each model to the maximum
+a posteriori (MAP) solution.  We then sampled using \texttt{PyMC3}'s
+gradient-based No-U-Turn Sampler \citep{hoffman_no-u-turn_2014}, and
+used $\hat{R}$ as our convergence diagnostic
+\citep{gelman_inference_1992}.
+We tested our ability to successfully recover injected parameters
+using synthetic data, before switching to the actual PTFO 8-8695
+lightcurves.
 
-Our priors were as follows.
-%FIXME
+\subsection{Fitting Results}
 
-We initialized to the maximum a posteriori (MAP) solution.
-We then sampled using \texttt{PyMC3}'s gradient-based No-U-Turns Sampler, 
-and used R-hat as our convergence diagnostic.
+\input{model_comparison_table.tex}
+\input{model_posterior_table.tex}
 
-% 
-% I gave it a pass on the PTFO data, binning from 2->10 minute sampling
-% to speed up the execution time. Results from two models are below.
-% Case 1 is N=2,M=1. Case 2 is N=2,M=2. Both fits are fine, and I get
-% decent convergence with a few minutes each. The rotation parameters
-% (phase + amplitudes) are quite correlated, so I set a reasonably
-% strong prior on the rotation phase.
-% 
+We considered nine models, with the number of harmonics per frequency
+$N$ and $M$ ranging from one to three.  To select our preferred model,
+we used the Bayesian information criterion
+(Table~\ref{tab:modelcompare}).  The model with the lowest BIC had
+$N=2$ harmonics at the short 11.74$\,$hr period, and $M=2$ harmonics
+at the long 11.96$\,$hr period.  All of the models have reduced
+$\chi^2$ ranging between 1.37 and 1.51, which suggests a plausible,
+though not perfect agreement between the data and models.
+
+To explore where each model succeeded and failed, we split the raw
+signal into its respective components (Figures~\ref{fig:splitsignal}
+and~\ref{fig:splitsignalii}).  We also examined the phase-folded
+signals (Figure~\ref{fig:phasefold}).
+
+In every model, the variability at the long period is a simple
+sinusoid with peak-to-peak amplitude $\approx$10\%
+(Figure~\ref{fig:phasefold}, top).  The variability at the short
+period is always more complex.  A dip of depth $\approx$1.2\%, fit in
+our model as a transit, lasts $\approx$0.75 hours.  Superposed on the
+dip is a complex signal with peak-to-peak amplitude of about 4\%,
+which peaks near phase 0.25, and reaches minimum brightness between
+phases -0.5 and -0.25.
+
+Outside of the primary dip, the short-period signal is relatively
+smooth, at least from phases 0 to 0.5.  However the short-period
+signal is asymmetric.  The flux from phases -0.5 to 0 shows what could
+be a discontinuous jump, shortly after reaching minimum.  This jump
+was visible in each of the nine models we considered, with different
+choices for the number of harmonics.
+
 % The main result that both models agree on: the extra power at the
@@
 
-% Using the cleaned PDC lightcurve, we applied the Box Least Squares
-% algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
-% period, transit duration, and a reference epoch using the TESS data
-% alone.  Based on the results, we isolated the data within 4 transit
-% durations of each transit midpoint.  To find the transit parameters
-% that best fit the data, we first fitted a line to the out-of-transit
-% flux measurements surrounding each transit, and divided it out.  We
-% then created a phase-folded lightcurve \added{from all 18 transits}
-% and fitted a standard transit model using the analytic formulae given
-% by \citet{mandel_analytic_2002} and implemented by
-% \replaced{\citet{kreidberg_batman_2015}}{\citet[][\texttt{BATMAN}]{kreidberg_batman_2015}}
-% We assumed the orbit to be circular, consistent with the limits from
-% radial velocities and occultation timing
-% \citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
-% The free parameters were the reference epoch, the planet to star
-% radius ratio $R_{\rm p}/R_\star$, the orbital distance to stellar
-% radius ratio $a/R_\star$, the inclination $i$, two quadratic
-% limb-darkening coefficients $(u_{\rm linear}, u_{\rm quad})$, and the
-% orbital period $P$.
-% 
-% We sampled the posterior probability distribution for all the
-% parameters using the algorithm proposed by
-% \citet{goodman_ensemble_2010} and implemented by
-% \replaced{\citet{foreman-mackey_emcee_2013}}{\citet[][\texttt{emcee}]{foreman-mackey_emcee_2013}}.
-% Table~1 gives the results, which are in reasonable agreement with the
-% parameters reported by {\it e.g.},
-% \citet{southworth_high-precision_2009} and
-% \citet{huitson_gemini_2017}.  Figure~\ref{fig:phasefold} shows the
-% phase-folded lightcurve.
-% 
-% To measure the transit times, we returned to the `cleaned' PDC time
-% series and fitted the data within four transit durations of each
-% transit separately. We used four free parameters: the time of
-% mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
-% slope and intercept of a linear trend to account for any slow
-% variations unrelated to the transit.  We fixed the remaining
-% parameters at the values that had been determined from the
-% phase-folded TESS lightcurve.  The uncertainty in each
-% \added{photometric} data point was set equal to the root-mean-square
-% (rms) level of the out-of-transit data.
-% 
-% To verify that the measured uncertainties are estimated accurately, we
-% computed the \deleted{reduced }$\chi^2$\added{ value} for a linear
-% ephemeris fit to the measured TESS mid-transit times.  We found that
-% $\chi^2 = 9.2$, with $n=16$ degrees of freedom.  The variance of the
-% $\chi^2$ distribution is $2n$, so we would expect $\chi^2 = 16 \pm
-% 5.7$.  Visually inspecting the residuals showed that the error
-% variance had been overestimated, so we multiplied the measured TESS
-% errors by a factor $f=0.76$, forcing a reduced $\chi^2$ of unity.
-% This lowered the mean uncertainty of the transit midtimes from $29.8$
-% to $22.6$ seconds.  We verified that omitting this step did not
-% appreciably alter any of our conclusions.
-% 
-% Figure~\ref{fig:lightcurves} shows the lightcurve of each individual
-% transit, the best-fit models, and the residuals.  Table~2 reports the
-% mid-transit times and their uncertainties.  After binning the
-% residuals to 1-hour windows, the lightcurves have an rms scatter of
-% 586\,{\rm ppm}.  The pre-launch TESS noise
-% model\footnote{\url{github.com/lgbouma/tnm}, commit \texttt{be06f09}}
-% (\citealt{winn_photonflux_2013}, \citealt{Sullivan_2015} Section 6.4)
-% would have predicted an error budget consisting of the following terms
-% added in quadrature: 410\,{\rm ppm} from photon-counting noise,
-% 202\,{\rm ppm} from detector read noise, and 673\,{\rm ppm} from the
-% zodiacal background light.  The level of background light appears to
-% have been overestimated in the model.
-
-% I fitted some models of the form "transit + N harmonics of sines and
-% cosines at the orbital frequency + M harmonics of sines and cosines at
-% the rotation frequency". For example, if N=2 and M=1, then I fit for
-% [logBrot0, logArot0, phirot, omegarot, logBorb1, logAorb1, logBorb0,
-% logAorb0, b, r, u, logP, t0, meanflux]. A_N is the amplitude of the
-% Nth harmonic sine function A_N * sin(N(ωt+φ)), B_N of the cosine. I
-% went with quadratic limb darkening.
-% 
-% On fake data, I got the right parameters provided I didn't turn up the
-% noise too high.
-% 
-% I gave it a pass on the PTFO data, binning from 2->10 minute sampling
-% to speed up the execution time. Results from two models are below.
-% Case 1 is N=2,M=1. Case 2 is N=2,M=2. Both fits are fine, and I get
-% decent convergence with a few minutes each. The rotation parameters
-% (phase + amplitudes) are quite correlated, so I set a reasonably
-% strong prior on the rotation phase.
-% 
-% The main result that both models agree on: the extra power at the
-% orbital frequency seems to not be ellipsoidal. Both of these models
-% prefer to put the power at 1x the orbital frequency, not 2x the
-% orbital frequency! This gives both Aorb0 and Borb0 non-zero, while
-% Aorb1 and Borb1 seem to be more consistent with zero.
-
-The periodogram of the final residual (Figure~\ref{fig:splitsignal}
-bottom row) shows a weakly significant, poorly resolved peak at
-$\approx$8 days, consistent with the visual impression in the time
-domain that there could be a weak long-period signal present.
+% The periodogram of the final residual (Figure~\ref{fig:splitsignal}
+% bottom row) shows a weakly significant, poorly resolved peak at
+% $\approx$8 days, consistent with the visual impression in the time
+% domain that there could be a weak long-period signal present.
 

### 2020-04-14 paper/collect.sh

@@ -1,15 +1,15 @@
 # Usage: copy in figures for the paper
 
-pdir=../results/PTFO_8-8695_results/20200320_linear_amplitude_prior/
+pdir=../results/PTFO_8-8695_results/20200413_v0/
 
-#cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_joined.pdf f1.pdf
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap.png f1.png
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_i.pdf f1.pdf
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_ii.pdf f2.pdf
 
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.pdf f2.pdf
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.pdf f3.pdf
 
-cp ../results/PTFO_8-8695_results/scene.pdf f3.pdf
+cp ../results/PTFO_8-8695_results/scene.pdf f4.pdf
 
-cp ../results/ephemeris/O_minus_C.pdf f4.pdf
+cp ../results/ephemeris/O_minus_C.pdf f5.pdf
 
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f5.png
-pngquant f5.png --output f5_comp.png
+cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f6.png
+pngquant f6.png --output f6_comp.png

### 2020-04-14 drivers/make_model_comparison_table.py
add drivers.make_model_comparison_table

@@ -1,0 +1,28 @@
+import pickle, os
+import numpy as np, pandas as pd
+from glob import glob
+
+pklpaths = glob(
+    '/Users/luke/Dropbox/proj/billy/results/PTFO_8-8695_results/20200413_v0/*bicdict*'
+)
+ds = []
+for p in pklpaths:
+    d = pickle.load(open(p,'rb'))
+    ds.append(d)
+
+df = pd.DataFrame(ds)
+
+df = df.sort_values(by='BIC')
+
+df['D_BIC'] = df.BIC - df.BIC.min()
+
+df = df.drop('modelid', axis=1)
+
+df['chisq'] = np.round(df.chisq, 1)
+df['redchisq'] = np.round(df.redchisq, 3)
+df['BIC'] = np.round(df.BIC, 1)
+df['D_BIC'] = np.round(df.D_BIC, 1)
+
+outpath = '../results/PTFO_8-8695_results/20200413_v0/bic_table_data.tex'
+df.to_latex(outpath, index=False)
+print('wrote {}'.format(outpath))

### 2020-04-14 billy/convenience.py
add BIC pickling

@@ -1,5 +1,5 @@
 import numpy as np, pandas as pd
-import collections
+import collections, pickle, os
 from collections import OrderedDict
 from astropy.io import fits
 
@@ -18,7 +18,7 @@ def bic(chisq, k, n):
     return chisq + k*np.log(n)
 
 
-def get_bic(m, ydict):
+def get_bic(m, ydict, outdir):
 
     y_obs = ydict['y_obs']
     y_err = m.y_err
@@ -41,12 +41,27 @@ def get_bic(m, ydict):
     dof = n-k
 
     msg = (
-        '{}: χ2 = {:.1f}, redχ2 = {:.2f}, BIC = {:.1f}'.
-        format(m.modelid, χ2, χ2/dof, BIC)
+        '{}: Ndata = {:d}, Nparam = {:d}, χ2 = {:.1f}, redχ2 = {:.2f}, BIC = {:.1f}'.
+        format(m.modelid, n, k, χ2, χ2/dof, BIC)
     )
     print(42*'=')
     print(msg)
     print(42*'=')
+
+    bicdict = {
+        'modelid': m.modelid,
+        'N': _N,
+        'M': _M,
+        'Ndata': n,
+        'Nparam': k,
+        'chisq': χ2,
+        'redchisq': χ2/dof,
+        'BIC': BIC
+    }
+    pklpath = os.path.join(outdir, m.modelid+'_bicdict.pkl')
+    with open(pklpath, 'wb') as buff:
+        pickle.dump(bicdict, buff)
+    print('Wrote {}'.format(pklpath))
 
 
 def flatten(l):
@@ -140,7 +155,10 @@ def get_clean_ptfo_data(binsize=120*5):
                                           minbinelems=5)
         x_obs = bd['binnedtimes']
         y_obs = bd['binnedmags']
-        y_err = bd['binnederrs']
+
+        # assume errors scale as sqrt(N)
+        original_cadence = 120
+        y_err = bd['binnederrs'] / (binsize/original_cadence)**(1/2)
 
     assert len(x_obs) == len(y_obs) == len(y_err)
 

### 2020-04-14 drivers/real.py
minor

@@ -71,7 +71,7 @@ def main(modelid):
                 bp.plot_splitsignal_map_periodogram(ydict, outpath)
             outpath = join(PLOTDIR, '{}_{}_phasefoldmap.png'.format(REALID, modelid))
             bp.plot_phasefold_map(m, ydict, outpath)
-            get_bic(m, ydict)
+            get_bic(m, ydict, PLOTDIR)
 
     if cornerplot:
         prior_d.pop('omegaorb', None) # not sampled; only used in data generation
@@ -82,13 +82,15 @@ def main(modelid):
 
 if __name__ == "__main__":
 
-    for N, M in product(range(1,4), range(1,4)):
-        main('transit_{}sincosPorb_{}sincosProt'.format(N,M))
+    DEBUG = 0
 
-    # NOTE: DEBUG
-    # main('transit_1sincosPorb_1sincosProt')
+    if DEBUG:
+        main('transit_1sincosPorb_1sincosProt')
 
-    # NOTE: DEPRECATED
-    # main('transit_2sincosPorb_2sincosProt')
-    # main('transit_1sincosPorb_2sincosProt')
-    # main('transit_2sincosPorb_1sincosProt')
+    else:
+        for N, M in product(range(1,4), range(1,4)):
+            main('transit_{}sincosPorb_{}sincosProt'.format(N,M))
+        # DEPRECATED
+        # main('transit_2sincosPorb_2sincosProt')
+        # main('transit_1sincosPorb_2sincosProt')
+        # main('transit_2sincosPorb_1sincosProt')

### 2020-04-14 paper/ms.tex
compile da paper
update paper

@@
 
+\begin{figure*}[t!]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f1.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 19).}
+		{\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux
+		versus time. The beat period of 4.48 days is visible by eye.  The
+		model plotted underneath the data includes 2 harmonics at the long
+		period $P_{\rm \ell}$, plus 2 harmonics and a transit at the short
+		period $P_{\rm s}$.
+		{\it Upper middle}: Long-period signal, equal to the raw signal
+		minus the short-period signal.
+		{\it Lower middle}: Short-period signal, equal to the raw signal
+		minus the long-period signal.
+		{\it Bottom}: residual.  The data are binned from 2 to 10 minute
+		cadence as a convenience for plotting and fitting.
+		\label{fig:splitsignal}
+	}
+\end{figure*}
+
+%\newpage
+\begin{figure*}[hbtp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f2.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 20).}
+		Panels are as in Figure~\ref{fig:splitsignal}.
+		\label{fig:splitsignalii}
+	}
+\end{figure*}
+%\newpage
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.95\textwidth]{f3.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Phase-folded long and short-period signals.}
+		{\it Top}: Long-period signal, as in Figure~\ref{fig:splitsignal}.
+		{\it Bottom}: Short-period signal. The reference phase is set to the
+		``planetary'' dip.  Gray points are the 10 minute cadence
+		\texttt{PDCSAP} flux.  Black points are binned to 100 points per
+		period.
+		\label{fig:phasefold}
+	}
+\end{figure*}
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.48\textwidth]{f4.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Scene used for blend analysis.}
+		{\it Top:} Mean TESS image of PTFO 8-8695 over Sector~6, with a
+		log-stretch.  The position of PTFO 8-8695 is shown with a yellow
+		star.  Neighbors with $T<17$ are shown with orange crosses.  The
+		apertures used to measure the background and target star flux are
+		shown with \texttt{X} and \texttt{/} hatches, respectively.
+		{\it Bottom:} Digitized Sky Survey $R$-band image of the same
+		field, with a linear stretch. The circles show apertures of radii
+		1, 1.5, and 2.25 pixels used in part of our blend analysis.  The
+		pixel level TESS data show that ``Star A''  does not contribute
+		variability at either of the two observed periods (see
+		Section~\ref{subsec:blend}).
+		\label{fig:scene}
+	}
+\end{figure}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f5.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Timing residuals for PTFO 8-8695b from a decade of monitoring.}
+		Black points are times of ``dips'', minus the indicated linear
+		ephemeris.  The $y$-axis is given in units of phase for the
+		short-period signal.  The star shows the binned TESS ephemeris.
+		``Dips'' have been observed by \citet{van_eyken_ptf_2012},
+		\citet{ciardi_follow-up_2015}, \citet{yu_tests_2015},
+		\citet{raetz_yeti_2016}, \citet{onitsuka_multi-color_2017}, and
+		\citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
+		at phase 0 in mid-2019) are consistent with noise, and were likely
+		reported because something was {\it expected}, rather than
+		convincingly {\it observed}.  Horizontal dashed lines are drawn at
+		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting what
+		could be either a numerical coincidence or an observational bias.
+		The orbital phase observed by TESS is consistent with that of
+		\citet{tanimoto_evidence_2020}, and quite different from the
+		original phase.
+		\label{fig:o_minus_c}
+	}
+\end{figure*}
+
+\begin{figure*}[t!]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f6_comp.png}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf foo.}
+		bar
+		\label{fig:corner}
+	}
+\end{figure*}
+
 \subsection{Observations}
@@
 
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.98\textwidth]{f1.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-  {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 19).}
-  {\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux
-  versus time. The beat period of 4.48 days is visible by eye.  The
-  model plotted underneath the data includes 2 harmonics at the long
-  period $P_{\rm \ell}$, plus 2 harmonics and a transit at the short
-  period $P_{\rm s}$.
-  {\it Upper middle}: Long-period signal, equal to the raw signal
-  minus the short-period signal.
-  {\it Lower middle}: Short-period signal, equal to the raw signal
-  minus the long-period signal.
-  {\it Bottom}: residual.  The data are binned from 2 to 10 minute
-  cadence as a convenience for plotting and fitting.
-  \label{fig:splitsignal}
-	}
-\end{figure*}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.98\textwidth]{f2.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 20).}
-  Panels are as in Figure~\ref{fig:splitsignal}.
-  \label{fig:splitsignalii}
-	}
-\end{figure*}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f3.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf Phase-folded long and short-period signals.}
-	{\it Top}: Long-period signal, as in Figure~\ref{fig:splitsignal}.
-  {\it Bottom}: Short-period signal. The reference phase is set to the
-  ``planetary'' dip.  Gray points are the 10 minute cadence
-  \texttt{PDCSAP} flux.  Black points are binned to 100 points per
-  period.
-	\label{fig:phasefold}
-	}
-\end{figure*}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f4.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf Scene used for blend analysis.}
-    {\it Top:} Mean TESS image of PTFO 8-8695 over Sector~6, with a
-    log-stretch.  The position of PTFO 8-8695 is shown with a yellow
-    star.  Neighbors with $T<17$ are shown with orange crosses.  The
-    apertures used to measure the background and target star flux are
-    shown with \texttt{X} and \texttt{/} hatches, respectively.
-    {\it Bottom:} Digitized Sky Survey $R$-band image of the same
-    field, with a linear stretch. The circles show apertures of radii
-    1, 1.5, and 2.25 pixels used in part of our blend analysis.  The
-    pixel level TESS data show that ``Star A''  does not contribute
-    variability at either of the two observed periods (see
-    Section~\ref{subsec:blend}).
-		\label{fig:scene}
-	}
-\end{figure}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f5.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Timing residuals for PTFO 8-8695b from a decade of monitoring.}
-    Black points are times of ``dips'', minus the indicated linear
-    ephemeris.  The $y$-axis is given in units of phase for the
-    short-period signal.  The star shows the binned TESS ephemeris.
-    ``Dips'' have been observed by \citet{van_eyken_ptf_2012},
-    \citet{ciardi_follow-up_2015}, \citet{yu_tests_2015},
-    \citet{raetz_yeti_2016}, \citet{onitsuka_multi-color_2017}, and
-    \citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
-    at phase 0 in mid-2019) are consistent with noise, and were likely
-    reported because something was {\it expected}, rather than
-    convincingly {\it observed}.  Horizontal dashed lines are drawn at
-    $\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting what
-    could be either a numerical coincidence or an observational bias.
-    The orbital phase observed by TESS is consistent with that of
-    \citet{tanimoto_evidence_2020}, and quite different from the
-    original phase.
-		\label{fig:o_minus_c}
-	}
-\end{figure*}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f6_comp.png}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf foo.}
-    bar
-		\label{fig:corner}
-	}
-\end{figure*}
-
 \subsection{Model Description}
@@
 \input{model_comparison_table.tex}
-\input{model_posterior_table.tex}
+%\input{model_posterior_table.tex}
 
@@
 (Table~\ref{tab:modelcompare}).  The model with the lowest BIC had
-$N=2$ harmonics at the short 11.74$\,$hr period, and $M=2$ harmonics
+two harmonics at the short 11.74$\,$hr period, and two harmonics
 at the long 11.96$\,$hr period.  All of the models have reduced
-$\chi^2$ ranging between 1.37 and 1.51, which suggests a plausible,
-though not perfect agreement between the data and models.
+$\chi^2$ ranging between 1.37 and 1.51, which suggests a plausible
+though imperfect agreement between the data and models.
 
@@
 In every model, the variability at the long period is a simple
-sinusoid with peak-to-peak amplitude $\approx$10\%
-(Figure~\ref{fig:phasefold}, top).  The variability at the short
-period is always more complex.  A dip of depth $\approx$1.2\%, fit in
-our model as a transit, lasts $\approx$0.75 hours.  Superposed on the
-dip is a complex signal with peak-to-peak amplitude of about 4\%,
-which peaks near phase 0.25, and reaches minimum brightness between
-phases -0.5 and -0.25.
+sinusoid with peak-to-peak amplitude $\approx$10\%.  The variability
+at the short period is always more complex.  A dip of depth
+$\approx$1.2\%, fit in our model as a transit, lasts $\approx$0.75
+hours.  Superposed on the dip is a complex signal with peak-to-peak
+amplitude of about 4\%, which peaks near phase 0.25, and reaches
+minimum brightness between phases -0.5 and -0.25.
 
@@
 be a discontinuous jump, shortly after reaching minimum.  This jump
-was visible in each of the nine models we considered, with different
-choices for the number of harmonics.
+was visible in each of the nine models we considered.
+
+The periodogram of the final residual (Figure~\ref{fig:splitsignal}
+bottom row) shows a weakly significant, poorly resolved peak at
+$\approx$8 days, consistent with the visual impression in the time
+domain that there could be a weak long-period signal present.
+
+\subsection{Blend considerations}
+\label{subsec:blend}
+
+The TESS pixels are $\approx21$'' per side, and so before making an
+interptations, we need to consider whether light from neighboring
+stars could have affected the photometry.  The scene is shown in
+Figure~\ref{fig:scene}.  The pixels used to measure the background
+level in the SPOC lightcuirve are indicated with an `\texttt{X}'
+hatch, and the pixels used in the final lightcurve aperture are shown
+with the `\texttt{/}' hatch.
+
+The target star, PTFO 8-8605 (TIC 264461976), has a $T$-band magnitude
+of 14.0, and its position is shown with a star.  The other (unlabeled)
+star inside the target aperture, TIC 264461979, has $T=16.8$ and so
+cannot contribute a signal with relative amplitude 10\%.  The only
+neighbor that is sufficiently close and bright that its light might
+contaminate the target star is TIC 264461980, with $T=14.8$, which we
+denote ``Star A''.  Star A is 23.6'' NW of our target, and based on
+the magnitude difference could contribute up to 48\% the flux of our
+target star, PTFO 8-8695.
+
+Because PTFO 8-8695 has previously been identified to have periodicity
+consistent with our measurement of $P_{\rm s}$, our main concern
+regarding blending is the degree to which we can be certain that the
+long-period signal at $P_{\rm \ell}$ also originates from PTFO 8-8695.
+We took two approaches towards determining the source of the
+long-period signal.
+
+First, we examined the CDIPS full frame image lightcurves of the
+target, which are available on MAST \citep{bouma_cluster_2019}.  The
+maximal peak-to-peak beat amplitude is consistently $\approx$10\%
+across apertures of radii 1, 1.5, and 2.25 pixels.  If Star A were the
+source of the long-period variability, we would expect the peak
+variability amplitude to be smallest in the 1 pixel aperture, based on
+the separation of the sources (Figure~\ref{fig:scene}, bottom).  From
+this test alone, it seems unlikely that Star A is the source of the
+long-period signal.
+
+Second, we examined the lightcurve of each pixel in the scene
+individually.  We opted to use the interactive tools implemented in
+\texttt{lightkurve} \citep{lightkurve_2018}.  If Star A were the
+source of the long-period variability, we would expect the pixels
+nearest to Star A to show a sinusoidal signal with amplitude exceeding
+$10\%$.  We find no evidence for this being the case.  The pixel
+directly below Star A does not clearly show the sinusoidal
+variability, and the peak-to-peak variability in that pixel is
+$\lesssim 8\%$.  In contrast, the south-easternmost pixel within PTFO
+8-8695's aperture (the pixel furthest from Star A that was used in the
+optimal aperture) shows the $P_{\rm \ell}$ sinusoidal variability
+signal at $\approx 10\%$ amplitude.
+
+As there is no evidence in favor of a blend scenario, we conclude that
+both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from PTFO
+8-8695.
+
+\section{Discussion}
+\label{sec:discussion}
+
+\subsection{Long period sinusoid}
+The standard interpretation of sinusoidal modulations for a
+pre-main-sequence M dwarf is that we are observing a rotation period.
+There is no evidence to suggest that we are seeing anything different.
+
+\subsection{Short period dip}
+The dip lasts about 45 minutes, and seems to re-occur every 11.74
+hours
+(Figures~\ref{fig:splitsignal},~\ref{fig:splitsignalii},~\ref{fig:phasefold}).
+The dip duration is roughly the same as that observed by previous
+investigators \citep{van_eyken_ptf_2012,yu_tests_2015,}
+The dip depth is comparable to what has been observed in red visual
+bands...
+%TODO: and how does it compare to IR? bigger or smaller? cite relevant
+%studies.
+
+\paragraph{Epoch}
+The TESS dip does not phase up where it is supposed to...
+Figure~\ref{fig:o_minus_c}
+
+\subsection{Short period out-of-dip modulation}
+If there were a giant planet transiting PTFO 8-8695, it would tidally
+distort the host star, and cause ellipsoidal photometric modulations.
+The amplitude of the ellipsoidal distortion for a $1\,M_{\rm Jup}$
+companion would be about 1400$\,$ppm
+\citep{shporer_astrophysics_2017}.  This is significantly larger than
+the typical ellipsopidal modulation induced by close-in giant planets
+because the host star is puffy, and still on the pre-main-sequence.
+For our estimate, we assumed $R_\star = 1.39 R_\odot$, and $M_\star =
+0.39 M_\odot$ \citep{van_eyken_ptf_2012}.
+
+Our preferred model does detect a significant ellipsoidal signal,
+parametrized as the ``$B_1$'' component.  The amplitude of the signal
+is $0.53 \pm 0.06\%$.  Interpreted as being caused by a planet, it
+would imply a minimum planet mass $M_{\rm p} \sin i$ of $3.8\,M_{\rm
+Jup}$.
 
@@
 
-% The periodogram of the final residual (Figure~\ref{fig:splitsignal}
-% bottom row) shows a weakly significant, poorly resolved peak at
-% $\approx$8 days, consistent with the visual impression in the time
-% domain that there could be a weak long-period signal present.
-
-\subsection{Blend considerations}
-\label{subsec:blend}
-
-The TESS pixels are $\approx21$'' per side, and so we need to consider
-whether light from neighboring stars could affect the photometry.  The
-scene is shown in Figure~\ref{fig:scene}.  
-The pixels used to
-measure the background level are indicated with an `\texttt{X}' hatch,
-and the pixels used for the final lightcurve are shown with the
-`\texttt{/}' hatch.
-
-The target star, PTFO
-8-8605 (TIC 264461976), has a $T$-band magnitude of 14.0, and its position is shown with a
-star.  
-The other (unlabeled) star inside the target aperture, TIC 264461979, has $T=16.8$ and so cannot
-contribute a signal with relative amplitude 10\%.
-The only neighbor that is sufficiently close and bright that
-its light might contaminate the target star is TIC 264461980, with
-$T=14.8$, which we denote ``Star A''.  Star A is 23.6'' NW of our
-target, and based on the magnitude difference could contribute up to
-48\% the flux of our target star, PTFO 8-8695.
-
-Because PTFO 8-8695 has previously been identified to have periodicity
-consistent with our measurement of $P_{\rm s}$, our main concern
-regarding blending is the degree to which we can be certain that the
-long-period signal at $P_{\rm \ell}$ also originates from PTFO 8-8695.
-We took two approaches towards determining the source of the long-period signal.
-
-First, we examined the CDIPS full frame image lightcurves of the
-target, which are available on MAST \citep{bouma_cluster_2019}.
-The maximal peak-to-peak beat amplitude is consistently $\approx$10\%
-across apertures of radii 1, 1.5, and 2.25 pixels.
-If Star A were the source of the long-period variability, we would expect the
-peak variability amplitude to be smallest in the 1 pixel aperture, based on the
-separation of the sources (Figure~\ref{fig:scene}, bottom).
-From this test alone, it seems unlikely that Star A is the source of
-the long-period signal.
-
-Second, we examined the lightcurve of each pixel in the scene
-individually.  We opted to use the
-interactive tools implemented in
-\texttt{lightkurve} \citep{lightkurve_2018}.  If Star A were the
-source of the long-period variability, we would expect the pixels
-nearest to Star A to show a sinusoidal signal with
-amplitude exceeding $10\%$.  We find no evidence for
-this being the case.  The pixel directly below Star A does not
-clearly show the sinusoidal variability, and the peak-to-peak 
-variability in that pixel is $\lesssim 8\%$.  In contrast, the
-south-easternmost pixel within PTFO 8-8695's aperture (the pixel 
-furthest from Star A that was used in the optimal aperture) shows the $P_{\rm \ell}$ sinusoidal
-variability signal at $\approx 10\%$ amplitude.
-
-As there is no evidence in favor of a blend scenario, we
-conclude that both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from PTFO 8-8695.
-
-\section{Interpretation}
-\label{sec:discussion}
-
-The TESS dip does not phase up where it is supposed to...
-\label{fig:o_minus_c}
+\section{Plausible Physical Interpretations}
 
@@
 \label{sec:conclusions}
+
+PTFO 8-8695 was previously thought to potentially host a hot Jupiter.
+The TESS lightcurve of PTFO 8-8695 showed a number of new features,
+many of which seem to disfavor the hot Jupiter interpretatation.
+The TESS data showed two key pieces of evidence.
+\begin{enumerate}
+  \item {\it Two periodic signals.} The ``long'' signal is a 10\%
+      peak-to-peak sinusoidal modulation repeating every 11.96 hours.
+      The ``short'' signal is a 4\% peak-to-peak complex modulation
+      repeating every 10.74 hours. It is composed of a dip, plus at
+      least two harmonics. The signals beat, and therefore cannot be
+      an artifact linked to data processing.
+  \item {\it A dip at the wrong orbital phase.} The clearest dip in
+    the ``short'' signal was consistent with recent observations by
+    \citet{tanimoto_evidence_2020}, and differed from the discovery
+    epoch by 5.14 hours.
+\end{enumerate}
+
+The physical mechanism responsible for all these features remains a
+matter of speculation.
+With that said, the TESS data support new arguments against the planetary
+interpretation of PTFO 8-8695.
+First, if the long signal is caused by starspot modulation, and the
+short signal by a transiting planet, what causes the additional
+complex modulations seen at the short, ``orbital'', period?
+
+Similarly, if the planet truly orbits every 10.74 hours, while the
+star's equator spins every 11.96 hours, the situation is clearly
+Darwin unstable.
+
+  Given the available evidence,
+  PTFO$\,$8-8695 seems consistent with the
+  ``transient dipping'' phenomenology observed in many young M dwarfs.
+  It seems rather unlikely to be a planet.
 

### 2020-04-15 paper/ms.tex
update paper. it a photometric binary.
this will need another section!

@@
 \renewcommand*{\subsectionautorefname}{Section} %for \autoref
+
+\newcommand{\ptfo}{PTFO$\,$8-8695}
+\newcommand{\ptfob}{PTFO$\,$8-8695b}
 
@@
 \begin{abstract}
-  PTFO$\,$8-8695b could be the youngest, shortest-period 
-  hot Jupiter known.  However it has not been shown to be a planet.
-  TESS recently observed PTFO$\,$8-8695 for one month.
+  PTFO$\,$8-8695b, a candidate hot Jupiter in the 7--10$,$Myr old
+  Orion-OB1a cluster, might be the youngest hot Jupiter known.
+  We inspected data from TESS and Gaia to clarify whether it is
+  actually a planet.
   The TESS lightcurve shows that the dominant variability in this
   system is a sinusoidal modulation with a ``long'' period $P_{\rm
-  \ell}$
-  of 11.96 hours, likely caused by stellar rotation.
-  Also present is a complex signal, previously identified as the
-  planet candidate, that repeats with a ``short'' period $P_{\rm s}$ of
-  10.74 hours.
-  The ``long'' and ``short'' signals show the expected beat every 4.48
-  days.
-  There is a dip in the complex, short-period signal.
-  However ground-based photometry from the past decade shows that the
-  orbital phase of the dip seems to have instantaneously jumped, at
-  least once, and maybe twice.
-  The TESS epoch of the dip is consistent with recent observations by
-  Tanimoto et al., and differs from the discovery epoch by 5.14 hours.
+  \ell}=11.96\,$hr, likely caused by stellar rotation.  Also present
+  is a complex signal, previously identified as the planet candidate,
+  that repeats with a ``short'' period $P_{\rm s}= 10.74\,$hr.
+  The ``long'' and ``short'' signals beat every 4.48 days.
+  Although there is a dip in the short-period signal, ground-based
+  photometry from the past decade shows that the orbital phase of the
+  dip seems to have instantaneously jumped, at least once, and perhaps
+  twice.
   Planets do not ``jump'' in orbital phase.
-  Given the available evidence,
-  PTFO$\,$8-8695 seems consistent with the
-  ``transient dipping'' phenomenology observed in many young M dwarfs.
-  It seems rather unlikely to be a planet.
+  Furthermore, the Gaia data show that PTFO$\,$8-8695 is probably
+  a photometric binary (the comparison is made relative to the
+  kinematic subgroup of 25-Ori for which PTFO$\,$8-8695 is a member).
+  Given the available evidence, we believe that PTFO$\,$8-8695 is a
+  binary M dwarf system in which one star is showing the ``long''
+  rotation signal, and the other star is showing ``transient dipping''
+  that has been observed in many young M dwarfs.
+  The planetary interpretation seems highly unlikely.
 \end{abstract}
@@
 \section{Introduction}
-If PTFO$\,$8-8695b were a planet, it would be exceptional.
+If \ptfob\ were a planet, it would be exceptional.
 A transiting hot Jupiter, orbiting a $\approx$3$\,$Myr old M dwarf
@@
 	\caption{
-		{\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 19).}
+		{\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 19).}
 		{\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux
@@
 
-%\newpage
 \begin{figure*}[hbtp]
@@
 	\vspace{-0.7cm}
-	\caption{ {\bf TESS lightcurve of PTFO 8-8695 (Sector 6, Orbit 20).}
+  \caption{ {\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 20).}
 		Panels are as in Figure~\ref{fig:splitsignal}.
@@
 \end{figure*}
-%\newpage
 
@@
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f4.pdf}
+		\includegraphics[width=0.45\textwidth]{f4.pdf}
 	\end{center}
@@
 	\caption{ {\bf Scene used for blend analysis.}
-		{\it Top:} Mean TESS image of PTFO 8-8695 over Sector~6, with a
-		log-stretch.  The position of PTFO 8-8695 is shown with a yellow
+		{\it Top:} Mean TESS image of \ptfo\ over Sector~6, with a
+		log-stretch.  The position of \ptfo\ is shown with a yellow
 		star.  Neighbors with $T<17$ are shown with orange crosses.  The
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f5.pdf}
+		\includegraphics[width=0.5\textwidth]{f5.pdf}
 	\end{center}
@@
 	\caption{
-		{\bf Timing residuals for PTFO 8-8695b from a decade of monitoring.}
+		{\bf Timing residuals for \ptfob\ from a decade of monitoring.}
 		Black points are times of ``dips'', minus the indicated linear
@@
 	}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[t!]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f6_comp.png}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf foo.}
-		bar
-		\label{fig:corner}
-	}
-\end{figure*}
+%\begin{figure*}[t!]
+%	\begin{center}
+%		\leavevmode
+%		\includegraphics[width=0.9\textwidth]{f6_comp.png}
+%	\end{center}
+%	\vspace{-0.7cm}
+%	\caption{ {\bf foo.}
+%	bar
+%		\label{fig:corner}
+%	}
+%\end{figure*}
 
-\subsection{Observations}
+\subsection{TESS Observations}
 
-PTFO 8-8695 was observed by TESS with Camera 1, CCD 1, from December
+\ptfo\ was observed by TESS with Camera 1, CCD 1, from December
 15, 2018 to January 6, 2019, during the sixth sector of science
-operations.  The star was designated TIC 264461976 in the TESS Input
-Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.  The pixel data
-for an $11\times11$ array surrounding PTFO 8-8695 were averaged into
-2-minute stacks by the onboard computer.  Each 2048$\times$2048 image
-from the CCD was also averaged into 30-minute stacks, and saved as a
-``full frame image'' (FFI).
+operations \citep{ricker_transiting_2015}.  The star was designated
+TIC 264461976 in the TESS Input Catalog
+\citep{stassun_TIC_2018,stassun_TIC8_2019}.  The pixel data for an
+$11\times11$ array surrounding \ptfo\ were averaged into 2-minute
+stacks by the onboard computer.  Each 2048$\times$2048 image from the
+CCD was also averaged into 30-minute stacks, and saved as a ``full
+frame image'' (FFI).
 
-The 2-minute stacks for PTFO 8-8695 were then reduced to lightcurves
+The 2-minute stacks for \ptfo\ were then reduced to lightcurves
 by the Science Processing Operations Center (SPOC) at NASA
@@
 Difference Imaging Photometric Survey (CDIPS;
-\citep{bouma_cluster_2019}).  The CDIPS lightcurve used a circular
+\citealt{bouma_cluster_2019}).  The CDIPS lightcurve used a circular
 aperture with radius 1 pixel.
@@
 
-\subsection{Visual Inspection}
+\subsection{Gaia Observations}
+
+Between July 25, 2014 and May 23, 2016, Gaia acquired 121 ``good''
+observations of \ptfo, {\it i.e.}, observations that were not strongly
+downweighted in the astrometric solution of the source (CITE).
+%TODO: give an original explanation of how Gaia works
+The Gaia data processing team at XXXINSTITUTION combined the XXX and
+YYY into a global astrometric solution, from which the parallax of 
+\ptfo\ was derived (CITE).
+The star was assigned a Gaia DR2 identifier of 3222255959210123904
+(CITE).
+In addition, measurements of the stellar proper motion between
+observation epochs were made.
+
+One point of note:
+Gaia DR2 measured a significant ``astrometric excess'', at a level of
+10.3$\sigma$.
+This astrometric excess indicates the degree
+to which a single-source model fails to explain the observed astrometric
+measurements (CITE).
+
+\subsection{Cluster Membership}
+The Orion molecular cloud complex has numerous subgroups, with ages
+spanning 1 to 15$\,$Myr (CITE).
+\ptfo\ has been known to be a member of the Orion OB1a association
+since at least CITE (Briceno 2005).
+Its relation to the broader Orion complex has been further explored by
+CITEX, CITEY, and CITEZ (Briceno 05, 08, 18, whomever Van Eyken cites,
+and Kounkel 18).
+
+In describing the cluster membership of \ptfo, we
+follow the notation and results of \citet{kounkel_apogee2_2018}.
+\citet{kounkel_apogee2_2018} combined astrometric data from Gaia DR2
+with spectroscopic data from APOGEE-2 (CITE).
+They then performed a hierarchical clustering on the six dimensional
+position and velocity information to identify subgroups within the
+Orion complex.
+From smallest to largest groups, \ptfo\ was identified as being a
+member of the following subgroups:
+\begin{equation}
+  {\rm PTFO\,8\text{-}8695}
+  \in {\rm 25\,Ori\text{-}1}
+  \subset {\rm 25\,Ori}
+  \subset {\rm Orion\ OB1a}
+  \subset {\rm Orion\ D},
+\end{equation}
+where from set-notation, `$\subset$' denotes ``is a proper subset of''.
+
+While all members of the Orion complex are young relative to the field, the internal
+age dispersion between different subgroups is measurable.
+The Orion Nebula Cluster (M42) is a site of ongoing star-formation, and its
+stars are 1-3$\,$Myr old (CITE).
+25$\,$Ori\-1, by contrast, is XX-XX$\,$Myr old (CITE).
+These details are essential when assessing any evidence for photometric binarity
+in \ptfo, because there is a degeneracy between stellar luminosity
+and age for stars on the pre-main-sequence.
+Having a clean sample of tightly spatially and kinematically associated stars
+is essential to minimize contamination not just from the field, but from older and younger
+members of the Orion complex itself.
+
+\section{TESS Analysis}
+\label{sec:analysis}
+
+\subsection{Inspection}
 
@@
 
-\section{The Model}
-\label{sec:analysis}
-
-\subsection{Model Description}
+\subsection{ Model}
 
@@
 We tested our ability to successfully recover injected parameters
-using synthetic data, before switching to the actual PTFO 8-8695
+using synthetic data, before switching to the actual \ptfo\
 lightcurves.
 
-\subsection{Fitting Results}
+\subsection{ Fitting Results}
 
@@
 
-The target star, PTFO 8-8605 (TIC 264461976), has a $T$-band magnitude
+The target star, \ptfo\ (TIC 264461976), has a $T$-band magnitude
 of 14.0, and its position is shown with a star.  The other (unlabeled)
@@
 the magnitude difference could contribute up to 48\% the flux of our
-target star, PTFO 8-8695.
+target star, \ptfo.
 
-Because PTFO 8-8695 has previously been identified to have periodicity
+Because \ptfo\ has previously been identified to have periodicity
 consistent with our measurement of $P_{\rm s}$, our main concern
 regarding blending is the degree to which we can be certain that the
-long-period signal at $P_{\rm \ell}$ also originates from PTFO 8-8695.
+long-period signal at $P_{\rm \ell}$ also originates from \ptfo.
 We took two approaches towards determining the source of the
@@
 variability, and the peak-to-peak variability in that pixel is
-$\lesssim 8\%$.  In contrast, the south-easternmost pixel within PTFO
-8-8695's aperture (the pixel furthest from Star A that was used in the
+$\lesssim 8\%$.  In contrast, the south-easternmost pixel within \ptfo's aperture (the pixel furthest from Star A that was used in the
 optimal aperture) shows the $P_{\rm \ell}$ sinusoidal variability
@@
 As there is no evidence in favor of a blend scenario, we conclude that
-both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from PTFO
-8-8695.
+both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from \ptfo.
+
+\section{Gaia Analysis}
+\label{sec:gaia}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f6.png}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Positions, kinematics, and HR diagram of \ptfo.}
+	Members of the 25$\,$Ori-1 subgroup are shown with black circles, and were identified by \citet{kounkel_apogee2_2018} through clustering on Gaia-DR2 and APOGEE data.
+	The ``neighborhood'' (gray circles) is defined as the group of at most $10^4$ randomly selected non-member stars within 5 standard deviations of the mean right ascension, declination, and parallax.
+	\label{fig:gaia}
+	}
+\end{figure*}
+
+We did an analysis.
+The result was
+Figure~\ref{fig:gaia}.
 
@@
 \subsection{Short period out-of-dip modulation}
-If there were a giant planet transiting PTFO 8-8695, it would tidally
+If there were a giant planet transiting \ptfo, it would tidally
 distort the host star, and cause ellipsoidal photometric modulations.
@@
 
-PTFO 8-8695 was previously thought to potentially host a hot Jupiter.
-The TESS lightcurve of PTFO 8-8695 showed a number of new features,
+\ptfo\ was previously thought to potentially host a hot Jupiter.
+The TESS lightcurve of \ptfo\ showed a number of new features,
 many of which seem to disfavor the hot Jupiter interpretatation.
@@
 With that said, the TESS data support new arguments against the planetary
-interpretation of PTFO 8-8695.
+interpretation of \ptfo.
 First, if the long signal is caused by starspot modulation, and the

### 2020-04-15 paper/collect.sh

@@ -11,5 +11,7 @@ cp ../results/PTFO_8-8695_results/scene.pdf f4.pdf
 
 cp ../results/ephemeris/O_minus_C.pdf f5.pdf
 
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f6.png
-pngquant f6.png --output f6_comp.png
+cp ../results/cluster_membership/3222255959210123904_nbhd_plot.png f6.png
+
+# cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f6.png
+# pngquant f6.png --output f6_comp.png

### 2020-04-15 billy/plotting.py
add plot_hr

@@ -14,13 +14,15 @@ Plots:
 
     plot_scene
 
+    plot_hr
+
     plot_O_minus_C
 
 Convenience:
     savefig
     format_ax
 """
-import os, corner
+import os, corner, pickle
 import numpy as np, matplotlib.pyplot as plt
 from datetime import datetime
 from pymc3.backends.tracetab import trace_to_dataframe
@@ -747,6 +749,66 @@ def plot_scene(c_obj, img_wcs, img, outpath, Tmag_cutoff=17, showcolorbar=0,
         fig.tight_layout(h_pad=1, w_pad=1)
 
     savefig(fig, outpath, dpi=300)
+
+
+def plot_hr(outdir):
+
+    pklpath = '/Users/luke/Dropbox/proj/billy/results/cluster_membership/nbhd_info_3222255959210123904.pkl'
+    info = pickle.load(open(pklpath, 'rb'))
+    (targetname, groupname, group_df_dr2, target_df, nbhd_df,
+     cutoff_probability, pmdec_min, pmdec_max, pmra_min, pmra_max,
+     group_in_k13, group_in_cg18, group_in_kc19, group_in_k18
+    ) = info
+
+    ##########
+
+    plt.close('all')
+
+    f, ax = plt.subplots(figsize=(4,3))
+
+    nbhd_yval = np.array([nbhd_df['phot_g_mean_mag'] +
+                          5*np.log10(nbhd_df['parallax']/1e3) + 5])
+    ax.scatter(
+        nbhd_df['phot_bp_mean_mag']-nbhd_df['phot_rp_mean_mag'], nbhd_yval,
+        c='gray', alpha=1., zorder=2, s=7, rasterized=True, linewidths=0,
+        label='Neighborhood'
+    )
+
+    yval = group_df_dr2['phot_g_mean_mag'] + 5*np.log10(group_df_dr2['parallax']/1e3) + 5
+    ax.scatter(
+        group_df_dr2['phot_bp_mean_mag']-group_df_dr2['phot_rp_mean_mag'],
+        yval,
+        c='k', alpha=1., zorder=3, s=9, rasterized=True, linewidths=0,
+        label='K+18 members'
+    )
+
+    target_yval = np.array([target_df['phot_g_mean_mag'] +
+                            5*np.log10(target_df['parallax']/1e3) + 5])
+    ax.plot(
[…]

### 2020-04-15 drivers/plot_hr.py

@@ -1,0 +1,8 @@
+import os
+import billy.plotting as bp
+from billy import __path__
+
+RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
+PLOTDIR = os.path.join(RESULTSDIR, 'cluster_membership')
+
+bp.plot_hr(PLOTDIR)

### 2020-04-15 paper/ms.tex
update paper

@@
 \begin{abstract}
-  PTFO$\,$8-8695b, a candidate hot Jupiter in the 7--10$,$Myr old
-  Orion-OB1a cluster, might be the youngest hot Jupiter known.
+  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10$,$Myr old
+  Orion-OB1a cluster. It might be the youngest hot Jupiter known.
   We inspected data from TESS and Gaia to clarify whether it is
@@
   Planets do not ``jump'' in orbital phase.
-  Furthermore, the Gaia data show that PTFO$\,$8-8695 is probably
-  a photometric binary (the comparison is made relative to the
-  kinematic subgroup of 25-Ori for which PTFO$\,$8-8695 is a member).
+  Furthermore, the Gaia data show that PTFO$\,$8-8695 is probably a
+  photometric binary, relative to the subgroup of Orion for which
+  PTFO$\,$8-8695 is a member.
   Given the available evidence, we believe that PTFO$\,$8-8695 is a
@@
 observation epochs were made.
+%FIXME sentence is bad
 
@@
 
-In describing the cluster membership of \ptfo, we
-follow the notation and results of \citet{kounkel_apogee2_2018}.
+In describing the cluster membership of \ptfo, we follow the notation
+and results of \citet{kounkel_apogee2_2018}.
 \citet{kounkel_apogee2_2018} combined astrometric data from Gaia DR2
-with spectroscopic data from APOGEE-2 (CITE).
-They then performed a hierarchical clustering on the six dimensional
-position and velocity information to identify subgroups within the
-Orion complex.
-From smallest to largest groups, \ptfo\ was identified as being a
-member of the following subgroups:
+with spectroscopic data from APOGEE-2 (CITE).  They then performed a
+hierarchical clustering on the six dimensional position and velocity
+information to identify subgroups within the Orion complex.  From
+smallest to largest groups, \ptfo\ was identified as being a member of
+the following subgroups:
 \begin{equation}
@@
 
-While all members of the Orion complex are young relative to the field, the internal
-age dispersion between different subgroups is measurable.
-The Orion Nebula Cluster (M42) is a site of ongoing star-formation, and its
-stars are 1-3$\,$Myr old (CITE).
-25$\,$Ori\-1, by contrast, is XX-XX$\,$Myr old (CITE).
-These details are essential when assessing any evidence for photometric binarity
-in \ptfo, because there is a degeneracy between stellar luminosity
-and age for stars on the pre-main-sequence.
-Having a clean sample of tightly spatially and kinematically associated stars
-is essential to minimize contamination not just from the field, but from older and younger
-members of the Orion complex itself.
+While all members of the Orion complex are young relative to the
+field, the internal age dispersion between different subgroups is
+measurable.  The Orion Nebula Cluster (M42) is a site of ongoing
+star-formation, and its stars are 1-3$\,$Myr old (CITE).
+25$\,$Ori\-1, by contrast, is XX-XX$\,$Myr old (CITE).  These details
+are essential when assessing any evidence for photometric binarity in
+\ptfo, because there is a degeneracy between stellar luminosity and
+age for stars on the pre-main-sequence.  Having a clean sample of
+tightly spatially and kinematically associated stars is essential to
+minimize contamination not just from the field, but from older and
+younger members of the Orion complex itself.
 
@@
 
-\subsection{ Model}
+\subsection{Lightcurve Model}
 
@@
 
-\subsection{ Fitting Results}
+\subsection{Fitting Results}
 
@@
 variability, and the peak-to-peak variability in that pixel is
-$\lesssim 8\%$.  In contrast, the south-easternmost pixel within \ptfo's aperture (the pixel furthest from Star A that was used in the
+$\lesssim 8\%$.  In contrast, the south-easternmost pixel within
+\ptfo's aperture (the pixel furthest from Star A that was used in the
 optimal aperture) shows the $P_{\rm \ell}$ sinusoidal variability
@@
 As there is no evidence in favor of a blend scenario, we conclude that
-both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from \ptfo.
+both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from \ptfo,
+at least within the resolution of the Gaia-DR2 source catalog.
+However, as we shall see, \ptfo itself could still be a binary.
 
-\section{Gaia Analysis}
+\section{Gaia Analysis \& Photometric Binarity}
 \label{sec:gaia}
 
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\includegraphics[width=0.9\textwidth]{f6.png}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+% 	\caption{ {\bf Positions, kinematics, and HR diagram of \ptfo.}
+%   Members of the 25$\,$Ori-1 subgroup are shown with black circles,
+%   and were identified by \citet{kounkel_apogee2_2018} through
+%   clustering on Gaia-DR2 and APOGEE data.  The ``neighborhood'' (gray
+%   circles) is defined as the group of at most $10^4$ randomly selected
+%   non-member stars within 5 standard deviations of the mean right
+%   ascension, declination, and parallax.
+% 	\label{fig:gaia}
+% 	}
+% \end{figure*}
 \begin{figure*}[t]
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f6.png}
+		\includegraphics[width=0.9\textwidth]{f6.pdf}
 	\end{center}
 	\vspace{-0.7cm}
-	\caption{ {\bf Positions, kinematics, and HR diagram of \ptfo.}
-	Members of the 25$\,$Ori-1 subgroup are shown with black circles, and were identified by \citet{kounkel_apogee2_2018} through clustering on Gaia-DR2 and APOGEE data.
-	The ``neighborhood'' (gray circles) is defined as the group of at most $10^4$ randomly selected non-member stars within 5 standard deviations of the mean right ascension, declination, and parallax.
+	\caption{ {\bf HR diagram of \ptfo and members of 25$\,$Ori-1.}
+  Members of the 25$\,$Ori-1 group (black circles) were identified by
+  \citet{kounkel_apogee2_2018} by clustering on six-dimensional
+  Gaia-DR2 and APOGEE-2 data.  We defined the reference
+  ``neighborhood'' (gray circles) as the group of at most $10^4$
+  randomly selected non-member stars within 5 standard deviations of
+  the mean 25$\,$Ori-1 right ascension, declination, and parallax.
 	\label{fig:gaia}
@@
 
-We did an analysis.
-The result was
-Figure~\ref{fig:gaia}.
+To assess the cluster membership, stellar mass, age, and potential
+binarity of \ptfo, we need to identify stars with which it is and is
+not coeval.  The simplest way to do this---clustering based on
+six-dimensional position and kinematic information---has already been
+done by \citet{kounkel_apogee2_2018}.  For simplicity, we considered
+the members they identified brighter than $G_{\rm Rp}$ of 16.  This
+yielded 149 members of 25$\,$Ori-1.
+
+To define a set of non-associated stars with comparable selection
+functions, we defined a reference ``neighborhood'' as the group of at
+most $10^4$ randomly selected non-member stars within 5 standard
+deviations of the mean 25$\,$Ori-1 right ascension, declination, and
+parallax.  We queried these stars using the \texttt{astroquery}
+package, which provides a convenient interface to the Gaia archive
+(CITE, CITE).  This yielded 1{,}819 neighbor stars.  While some of
+these stars may indeed be members of the Orion complex, or even of
+25$\,$Ori-1, enforcing this cut on positions and parallaxes ensures
+that we are querying stars with comparable amounts of interstellar
+reddening.
+
+We examined the resulting five-dimensional right ascension,
+declination, proper motions, and parallaxes.  The first point we noted
+was that 25$\,$Ori-1 was a clearly defined over-density in each
+dimension---the cluster definitely exists, and is different from the
+neighborhood.  \ptfo\ was also clearly a member in each of these
+projected dimensions.
+
+Given our detection of two separate signals, the issue of whether
+\ptfo\ could be a photometric binary was of relatively high concern.
+Figure~\ref{fig:gaia} shows the HR diagram from which we can assess
+this question.
 
@@
 \subsection{Long period sinusoid}
+
 The standard interpretation of sinusoidal modulations for a
@@
 There is no evidence to suggest that we are seeing anything different.
+
+This is the dominant signal in the system with 10\% amplitude.
+Previous analyses, including the discovery study by
+\citet{van_eyken_ptf_2012}, saw the same signal ({\it e.g.}, their
+Figure~7).  However it was not previously thought to be astrophysical.
+\citet{van_eyken_ptf_2012} identified it as a periodogram peak at
+$0.9985 \pm 0.0061\,$days, and ascribed it to the observing cadence,
+because of its close correspondence to the sidereal day.  While the
+TESS data can show significant reflected light from the Earth
+\citep[{\it e.g.},][]{luger_tess_2019}, our pixel-level analysis
+showed that the signal is specific to only pixels near \ptfo, and no
+other pixels.  We therefore conclude that the signal is astrophysical.
 

### 2020-04-15 paper/collect.sh

@@ -11,7 +11,7 @@ cp ../results/PTFO_8-8695_results/scene.pdf f4.pdf
 
 cp ../results/ephemeris/O_minus_C.pdf f5.pdf
 
-cp ../results/cluster_membership/3222255959210123904_nbhd_plot.png f6.png
+cp ../results/cluster_membership/hr.pdf f6.pdf
 
 # cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f6.png
 # pngquant f6.png --output f6_comp.png

### 2020-04-15 paper/ms.tex
update paper

@@
 		versus time. The beat period of 4.48 days is visible by eye.  The
-		model plotted underneath the data includes 2 harmonics at the long
+		preferred model plotted underneath the data includes 2 harmonics at the long
 		period $P_{\rm \ell}$, plus 2 harmonics and a transit at the short
@@
 \end{figure*}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f3.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf Phase-folded long and short-period signals.}
-		{\it Top}: Long-period signal, as in Figure~\ref{fig:splitsignal}.
-		{\it Bottom}: Short-period signal. The reference phase is set to the
-		``planetary'' dip.  Gray points are the 10 minute cadence
-		\texttt{PDCSAP} flux.  Black points are binned to 100 points per
-		period.
-		\label{fig:phasefold}
-	}
-\end{figure*}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.45\textwidth]{f4.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf Scene used for blend analysis.}
-		{\it Top:} Mean TESS image of \ptfo\ over Sector~6, with a
-		log-stretch.  The position of \ptfo\ is shown with a yellow
-		star.  Neighbors with $T<17$ are shown with orange crosses.  The
-		apertures used to measure the background and target star flux are
-		shown with \texttt{X} and \texttt{/} hatches, respectively.
-		{\it Bottom:} Digitized Sky Survey $R$-band image of the same
-		field, with a linear stretch. The circles show apertures of radii
-		1, 1.5, and 2.25 pixels used in part of our blend analysis.  The
-		pixel level TESS data show that ``Star A''  does not contribute
-		variability at either of the two observed periods (see
-		Section~\ref{subsec:blend}).
-		\label{fig:scene}
-	}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.5\textwidth]{f5.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Timing residuals for \ptfob\ from a decade of monitoring.}
-		Black points are times of ``dips'', minus the indicated linear
-		ephemeris.  The $y$-axis is given in units of phase for the
-		short-period signal.  The star shows the binned TESS ephemeris.
-		``Dips'' have been observed by \citet{van_eyken_ptf_2012},
-		\citet{ciardi_follow-up_2015}, \citet{yu_tests_2015},
-		\citet{raetz_yeti_2016}, \citet{onitsuka_multi-color_2017}, and
-		\citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
-		at phase 0 in mid-2019) are consistent with noise, and were likely
-		reported because something was {\it expected}, rather than
-		convincingly {\it observed}.  Horizontal dashed lines are drawn at
-		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting what
-		could be either a numerical coincidence or an observational bias.
-		The orbital phase observed by TESS is consistent with that of
-		\citet{tanimoto_evidence_2020}, and quite different from the
-		original phase.
-		\label{fig:o_minus_c}
-	}
-\end{figure}
 
@@
 \label{sec:analysis}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.8\textwidth]{f3.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Phase-folded long and short-period signals.}
+		{\it Top}: Long-period signal, as in Figure~\ref{fig:splitsignal}.
+		{\it Bottom}: Short-period signal. The reference phase is set to the
+		``planetary'' dip.  Gray points are the 10 minute cadence
+		\texttt{PDCSAP} flux.  Black points are binned to 100 points per
+		period.
+		\label{fig:phasefold}
+	}
+\end{figure*}
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.45\textwidth]{f4.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Scene used for blend analysis.}
+		{\it Top:} Mean TESS image of \ptfo\ over Sector~6, with a
+		log-stretch.  The position of \ptfo\ is shown with a yellow
+		star.  Neighbors with $T<17$ are shown with orange crosses.  The
+		apertures used to measure the background and target star flux are
+		shown with \texttt{X} and \texttt{/} hatches, respectively.
+		{\it Bottom:} Digitized Sky Survey $R$-band image of the same
+		field, with a linear stretch. The circles show apertures of radii
+		1, 1.5, and 2.25 pixels used in part of our blend analysis.  The
+		pixel level TESS data show that ``Star A''  does not contribute
+		variability at either of the two observed periods (see
+		Section~\ref{subsec:blend}).
+		\label{fig:scene}
+	}
+\end{figure}
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.5\textwidth]{f5.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Timing residuals for \ptfob\ from a decade of monitoring.}
+		Black points are times of ``dips'', minus the indicated linear
+		ephemeris.  The $y$-axis is given in units of phase for the
+		short-period signal.  The star shows the binned TESS ephemeris.
+		``Dips'' have been observed by \citet{van_eyken_ptf_2012},
+		\citet{ciardi_follow-up_2015}, \citet{yu_tests_2015},
+		\citet{raetz_yeti_2016}, \citet{onitsuka_multi-color_2017}, and
+		\citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
+		at phase 0 in mid-2019) are consistent with noise, and were likely
+		reported because something was {\it expected}, rather than
+		convincingly {\it observed}.  Horizontal dashed lines are drawn at
+		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting what
+		could be either a numerical coincidence or an observational bias.
+		The orbital phase observed by TESS is consistent with that of
+		\citet{tanimoto_evidence_2020}, and quite different from the
+		original phase.
+		\label{fig:o_minus_c}
+	}
+\end{figure}
 
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f6.pdf}
+		\includegraphics[width=0.8\textwidth]{f6.pdf}
 	\end{center}
 	\vspace{-0.7cm}
-	\caption{ {\bf HR diagram of \ptfo and members of 25$\,$Ori-1.}
+	\caption{ {\bf HR diagram of \ptfo\ and members of 25$\,$Ori-1.}
   Members of the 25$\,$Ori-1 group (black circles) were identified by
-  \citet{kounkel_apogee2_2018} by clustering on six-dimensional
+  \citet{kounkel_apogee2_2018} through clustering on six-dimensional
   Gaia-DR2 and APOGEE-2 data.  We defined the reference
@@
   the mean 25$\,$Ori-1 right ascension, declination, and parallax.
+  \ptfo\ seems likely to be a photometric binary.
 	\label{fig:gaia}

### 2020-04-16 paper/ms.tex
update paper

@@
   \ptfo\ seems likely to be a photometric binary.
+  ``$G$'' is the Gaia broadband, $Bp$ is Gaia blue, and $Rp$ is Gaia red.
+  $\omega_{\rm as}$ is the parallax in arcseconds.
 	\label{fig:gaia}
@@
 
-To assess the cluster membership, stellar mass, age, and potential
-binarity of \ptfo, we need to identify stars with which it is and is
+To assess the cluster membership and potential
+binarity of \ptfo, we needed to identify stars with which it was and was
 not coeval.  The simplest way to do this---clustering based on
-six-dimensional position and kinematic information---has already been
+six-dimensional position and kinematic information---had already been
 done by \citet{kounkel_apogee2_2018}.  For simplicity, we considered
 the members they identified brighter than $G_{\rm Rp}$ of 16.  This
-yielded 149 members of 25$\,$Ori-1.
+yielded 149 stars in 25$\,$Ori-1, mostly M dwarfs.
+\citet{kounkel_apogee2_2018} identified seven other smaller
+groups in the Orion complex near the Be star 25$\,$Ori. These groups
+received higher numbers, {\it e.g.}, 25$\,$Ori-2.
 
-To define a set of non-associated stars with comparable selection
+To define a set of non-member stars that nonetheless had comparable selection
 functions, we defined a reference ``neighborhood'' as the group of at
@@
 package, which provides a convenient interface to the Gaia archive
-(CITE, CITE).  This yielded 1{,}819 neighbor stars.  While some of
+(CITE, CITE).  This yielded 1{,}819 neighbors.  While some of
 these stars may indeed be members of the Orion complex, or even of
@@
 was that 25$\,$Ori-1 was a clearly defined over-density in each
-dimension---the cluster definitely exists, and is different from the
+dimension, so the cluster exists, and is different from the
 neighborhood.  \ptfo\ was also clearly a member in each of these
@@
 
-Given our detection of two separate signals, the issue of whether
-\ptfo\ could be a photometric binary was of relatively high concern.
-Figure~\ref{fig:gaia} shows the HR diagram from which we can assess
-this question.
+Given our detection of two separate signals, whether
+\ptfo\ could be a photometric binary was of great interest.
+Figure~\ref{fig:gaia} shows the HR diagram from which we assessed
+this issue.
+The diagram shows that \ptfo\ is $\approx$0.8 magnitudes brighter than the average
+25$\,$Ori-1 star of the same color.
+In other words, it is about twice as bright.
+It also seems to be on the photometric binary track of the cluster, which has a few other stars.
+
+The implication is that either 
+{\it (i)} \ptfo\ is notably younger than the kinematically identical cluster members, or
+{\it (ii)} \ptfo\ is a photometric binary.
+Given the independent presence of two resolved signals,
+we favor the interpretation that \ptfo\ is a binary star system.
 
@@
 The dip duration is roughly the same as that observed by previous
-investigators \citep{van_eyken_ptf_2012,yu_tests_2015,}
+investigators \citep{van_eyken_ptf_2012,yu_tests_2015}
 The dip depth is comparable to what has been observed in red visual

### 2020-04-16 paper/ms.tex
update title

@@
 
-\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
+%\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
+% \title{Seeing Double: Still Against the Planetary Interpretation of PTFO$\,$8-8695b}
+% \title{Seeing Double: TESS and Gaia Show That PTFO$\,$8-8695b is Unlikely a
+% Planet}
+\title{Seeing Double: Indications From {\it TESS} and {\it Gaia} That
+PTFO$\,$8-8695b is Not a Planet}
 
@@
 \begin{abstract}
-  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10$,$Myr old
+  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10 million year old
   Orion-OB1a cluster. It might be the youngest hot Jupiter known.
@@
   The TESS lightcurve shows that the dominant variability in this
-  system is a sinusoidal modulation with a ``long'' period $P_{\rm
-  \ell}=11.96\,$hr, likely caused by stellar rotation.  Also present
+  system is a sinusoid with a ``long'' period $P_{\rm
+  \ell}=11.96\,$hr, previously thought to be a systematic, but in fact likely caused by stellar rotation.  Also present
   is a complex signal, previously identified as the planet candidate,
   that repeats with a ``short'' period $P_{\rm s}= 10.74\,$hr.
-  The ``long'' and ``short'' signals beat every 4.48 days.
+  The two signals beat every 4.48 days.
   Although there is a dip in the short-period signal, ground-based
@@
   Furthermore, the Gaia data show that PTFO$\,$8-8695 is probably a
-  photometric binary, relative to the subgroup of Orion for which
-  PTFO$\,$8-8695 is a member.
+  photometric binary.
   Given the available evidence, we believe that PTFO$\,$8-8695 is a
-  binary M dwarf system in which one star is showing the ``long''
+  binary M dwarf in which one star shows the ``long''
   rotation signal, and the other star is showing ``transient dipping''
-  that has been observed in many young M dwarfs.
-  The planetary interpretation seems highly unlikely.
+  that has also been observed in other young M dwarfs.
+  The planetary interpretation seems quite unlikely.
 \end{abstract}
@@
   \ptfo\ seems likely to be a photometric binary.
-  ``$G$'' is the Gaia broadband, $Bp$ is Gaia blue, and $Rp$ is Gaia red.
+  $G$ is the Gaia broadband, $Bp$ is Gaia blue, $Rp$ is Gaia red, and
   $\omega_{\rm as}$ is the parallax in arcseconds.

### 2020-04-16 paper/ms.tex
wrote a sassy introduction

@@
 \submitjournal{AAS journals.}
-\shorttitle{Against PTFO$\,$8-8695b}
+\shorttitle{Seeing Double: PTFO$\,$8-8695}
 
@@
 %\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
-% \title{Seeing Double: Still Against the Planetary Interpretation of PTFO$\,$8-8695b}
+% \title{Seeing Double: Still Against the Planetary Interpretation of
+% PTFO$\,$8-8695b}
 % \title{Seeing Double: TESS and Gaia Show That PTFO$\,$8-8695b is Unlikely a
@@
 
-%
-% contributing authors: alphabetical
-%
-% \author[0000-0001-8638-0320]{A. W. Howard}
-% \affiliation{Cahill Center for Astrophysics, California Institute of
-% Technology, Pasadena, CA 91125, USA}
-% %
-% \author[0000-0002-2532-2853]{S. B. Howell}
-% \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
-% %
-% \author[0000-0002-0531-1073]{H. Isaacson}
-% \affiliation{Astronomy Department, University of California, Berkeley,
-% CA 94720, USA}
-% %
-% \author{H. Knutson}
-% \affiliation{Division of Geological and Planetary Sciences, California
-% Institute of Technology, Pasadena, CA 91125, USA}
-% %
-% \author[0000-0001-7233-7508]{R. A. Matson}
-% \affiliation{U.S. Naval Observatory, Washington, DC 20392, USA}
-% %
-
 \begin{abstract}
   PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10 million year old
-  Orion-OB1a cluster. It might be the youngest hot Jupiter known.
-  We inspected data from TESS and Gaia to clarify whether it is
-  actually a planet.
-  The TESS lightcurve shows that the dominant variability in this
-  system is a sinusoid with a ``long'' period $P_{\rm
-  \ell}=11.96\,$hr, previously thought to be a systematic, but in fact likely caused by stellar rotation.  Also present
-  is a complex signal, previously identified as the planet candidate,
-  that repeats with a ``short'' period $P_{\rm s}= 10.74\,$hr.
-  The two signals beat every 4.48 days.
-  Although there is a dip in the short-period signal, ground-based
-  photometry from the past decade shows that the orbital phase of the
-  dip seems to have instantaneously jumped, at least once, and perhaps
-  twice.
-  Planets do not ``jump'' in orbital phase.
-  Furthermore, the Gaia data show that PTFO$\,$8-8695 is probably a
-  photometric binary.
-  Given the available evidence, we believe that PTFO$\,$8-8695 is a
-  binary M dwarf in which one star shows the ``long''
-  rotation signal, and the other star is showing ``transient dipping''
-  that has also been observed in other young M dwarfs.
-  The planetary interpretation seems quite unlikely.
+  Orion-OB1a cluster. We inspected data from TESS and Gaia to clarify whether
+  it is actually a planet.  The TESS lightcurve shows that the dominant
+  variability in this system is a sinusoid with a ``long'' period $P_{\rm
+  \ell}=11.96\,$hr, likely caused by stellar rotation.  Also present is a
+  complex signal, previously identified as the planet candidate, that repeats
+  with a ``short'' period $P_{\rm s}= 10.74\,$hr.  The two signals beat every
+  4.48 days.  Although there is a dip in the short-period signal, ground-based
+  photometry from the past decade shows that the orbital phase of the dip seems
+  to have instantaneously jumped, at least once, and perhaps twice.  Planets do
+  not ``jump'' in orbital phase.  Furthermore, the Gaia data show that
+  PTFO$\,$8-8695 is probably a photometric binary.  Given the evidence, we
+  believe that PTFO$\,$8-8695 is a binary M dwarf in which one star shows the
+  ``long'' rotation signal, and the other is showing ``transient dipping'' that
+  has also been observed in other young M dwarfs.  The planetary interpretation
+  seems unlikely.
 \end{abstract}
@@
 %TODO
-\keywords{}
+\keywords{
+	Exoplanet evolution (491),
+	Open star clusters (1160),
+	Stellar ages (1581),
+	Stellar rotation (1629),
+	Variable stars (1761)
+}
 
@@
 \section{Introduction}
-If \ptfob\ were a planet, it would be exceptional.
-A transiting hot Jupiter, orbiting a $\approx$3$\,$Myr old M dwarf
-in ORION would make it the youngest hot Jupiter known.
-Its orbital period of only 12 HOURS would also make it the shortest
-period hot Jupiter known.
+If \ptfob\ were a planet, it would be exceptional.  Transiting a sub-10$\,$Myr
+old weak-lined T Tauri M dwarf in Orion, it would be the youngest hot Jupiter
+known \citep{van_eyken_ptf_2012}.  Its orbital period of only 10.7 hours would
+also give it the shortest period of any known hot Jupiter.  With such a short
+period, it would almost certainly have filled its Roche lobe, and would be
+actively losing mass to its host star.  Not only that, but the rapidly rotating
+host star would also be oblate and gravity darkened, and so the planet's orbit
+would likely precess into and out of transitability
+\citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.
 
-% Section~\ref{sec:observations} of this paper presents all of the
-% available transit data as well as the new radial velocity and speckle
-% imaging observations.  Section~\ref{sec:analysis} describes our
-% analysis of the data, and our interpretation that WASP-4 is being
-% pulled around by a brown dwarf or low-mass star.
-% Section~\ref{sec:discussion} places this result within the context of
-% orbital decay searches, and points out that line-of-sight
-% accelerations will be a relatively common type of ``false positive.''
-% Section~\ref{sec:conclusions} offers concluding remarks.
+Additional lines of evidence could imply further fascinating ``firsts'' for
+this candidate planet.  One first would be that its transits are about three
+times deeper in optical bandpasses ({\it e.g.,} $g$) than in the near-infrared
+({\it e.g.}, $z$-band) \citep{onitsuka_multicolor_2017,tanimoto_evidence_2020}.
+A cloud-free hydrogen-dominated planetary atmosphere cannot explain such a
+wavelength dependence.  The proposed explanation is that the planet is
+surrounded by a dust cloud \citep{tanimoto_evidence_2020}.
+
+A second first could be the detection of direct H$\alpha$ emission from the
+planet itself \citep{johnskrull_h_2016}.  While the stellar chromosphere emits
+in H$\alpha$, it seems that there is an additional excess H$\alpha$ emission
+that could be in phase with the planetary orbit.  The average velocity width of
+the excess H$\alpha$ emission is 87$\,$km$\,$s$^{-1}$, and its equivalent width
+is 70-80\% that from the stellar chromosphere \citep{johnskrull_h_2016}.  The
+proposed explanation is that the outflowing mass from the planet may explain
+this excess emission as well \citep{johnskrull_h_2016}.
+
+There are perhaps a few challenges to the planetary interpretation.  They
+include that the planet does not seem to emit infrared radiation in occulation,
+at least anywhere near the expected amplitude \citep{yu_tests_2015}.  In
+addition, despite measurement attempts by multiple investigators, it does not
+seem to show the Rossiter effect at the amplitude expected given the rapid
+stellar rotation and large planet size
+\citep{yu_tests_2015,ciardi_followup_2015}.  Finally, detailed modelling of the
+``precession + gravity darkening'' transits has shown that the necessary degree
+of gravity darkening is too great, given the spectroscopically observed
+equatorial velocity \citep{howarth_reappraisal_2016}.  Additionally, the
+gravity-darkened star would need to show photometric variability from its
+precession rotation axis, and such variability has not been observed.
+
+While the planetary interpetation has challenges, so too do alternative
+explanations.  High-latitude accretion hotspots might produce the observed
+H$\alpha$ variability, but \ptfo\ does not have an infrared (IR) excess
+associated with the presence of an inner disk.  Low-latitude spots, hot or
+cold, struggle to produce the necessary dip durations.  A circumstellar cloud
+of dust or gas might be viable, but clumps must be invoked to explain the
+short-duration flux dips.  The material might be a remnant of the primordial
+gas disk, or it could be ejected and entrained in the stellar magnetosphere.
+
+Section~\ref{sec:observations} presents newly available observations from the
+Transiting Exoplanet Survey Satellite (TESS; \citealt{ricker_transiting_2015})
+and Gaia \citep{gaia_collaboration_gaia_2018}.  Analyzing the TESS data in
+Section~\ref{sec:tess}, we find two distinct signals.  A long sinusoid repeats
+every 12.0 hours, and is probably stellar rotation.  A short ``dip + complex
+modulation'' repeats every 10.7 hours.  Analyzing the Gaia data in
+Section~\ref{sec:gaia}, we show that \ptfo\ is quite likely a photometric
+binary.  In Section~\ref{sec:discussion}, we note that the orbital phase of the
+dip seems to have instantaneously changed, perhaps twice.  We argue that
+PTFO$\,$8-8695 is a binary M dwarf in which one star shows a rotation signal,
+and the other is showing ``transient dipping'' that has been observed in other
+young M dwarfs, and is likely caused by eclipses of warm particulate clouds.
+Section~\ref{sec:conclusions} summarizes our main points.
 
@@
 \section{TESS Analysis}
-\label{sec:analysis}
+\label{sec:tess}
 
@@
 		``Dips'' have been observed by \citet{van_eyken_ptf_2012},
-		\citet{ciardi_follow-up_2015}, \citet{yu_tests_2015},
+		\citet{ciardi_followup_2015}, \citet{yu_tests_2015},
 		\citet{raetz_yeti_2016}, \citet{onitsuka_multi-color_2017}, and
@@
 
+%FIXME: summarize existant AO constraints here too
+\citet{schmidt_direct_2016}
+\citet{lee_evidence_2018}
+% \citet{czesla_xray_2019}
+
 To assess the cluster membership and potential
@@
 
-The standard interpretation of sinusoidal modulations for a
-pre-main-sequence M dwarf is that we are observing a rotation period.
-There is no evidence to suggest that we are seeing anything different.
+The standard interpretation of sinusoidal modulations for a pre-main-sequence M
+dwarf is that we are observing a stellar rotation period of 11.96 hours.  This
+is the dominant signal in the system with 10\% amplitude, and there is no
+evidence to suggest that this signal has any other origin.
 
-This is the dominant signal in the system with 10\% amplitude.
-Previous analyses, including the discovery study by
-\citet{van_eyken_ptf_2012}, saw the same signal ({\it e.g.}, their
-Figure~7).  However it was not previously thought to be astrophysical.
-\citet{van_eyken_ptf_2012} identified it as a periodogram peak at
-$0.9985 \pm 0.0061\,$days, and ascribed it to the observing cadence,
-because of its close correspondence to the sidereal day.  While the
-TESS data can show significant reflected light from the Earth
-\citep[{\it e.g.},][]{luger_tess_2019}, our pixel-level analysis
-showed that the signal is specific to only pixels near \ptfo, and no
-other pixels.  We therefore conclude that the signal is astrophysical.
+The discovery study by \citet{van_eyken_ptf_2012} saw the same signal ({\it
+e.g.}, their Figure~7), but identified its alias as a periodogram peak at
+$0.9985 \pm 0.0061\,$days. They ascribed it to their observing cadence, because
+of its close correspondence to the sidereal day.  While the TESS data can show
+significant reflected light from the Earth \citep[{\it
+e.g.},][]{luger_tess_2019}, our pixel-level analysis showed that the signal is
+specific to only pixels near \ptfo, and no other pixels.  We therefore conclude
+that the signal is not a systematic.
+
+We are not the first to reach the conclusion that the long period sinusoidal
+modulation is astrophysical.  A follow-up study by
+\citet{koen_multicolour_2015} identified the same modes and aliases as
+\citet{van_eyken_ptf_2012}, but argued that the $0.50\,{\rm d}$ signal was
+astrophysical.  Using archival photometry and photometry from the YETI global
+telescope network, \citet{raetz_yeti_2016} eventually came to the conclusion
+that the that the $0.50\,{\rm d}$ signal was indeed from stellar rotation.  The
+TESS data strongly support this conclusion.
 

### 2020-04-16 billy/plotting.py
minor

@@ -797,7 +797,9 @@ def plot_hr(outdir):
 
     ylim = ax.get_ylim()
     ax.set_ylim((max(ylim),min(ylim)))
-    # ax.set_xlim((-0.5, 3.0))
+
+    ax.set_xlim((0.9, 3.1))
+    ax.set_ylim((9.5, 4.5))
 
     # # set M_omega y limits
     # min_y = np.nanmin(np.array([np.nanpercentile(nbhd_yval, 2), target_yval]))

### 2020-04-16 paper/ms.tex
paper smithing

@@
 
-Additional lines of evidence could imply further fascinating ``firsts'' for
-this candidate planet.  One first would be that its transits are about three
-times deeper in optical bandpasses ({\it e.g.,} $g$) than in the near-infrared
+Other lines of evidence would imply further planetary ``firsts'' for
+this planet candidate.  One first would be that its transits are about three
+times deeper in optical bandpasses ({\it e.g.,} $g$-band) than in the near-infrared
 ({\it e.g.}, $z$-band) \citep{onitsuka_multicolor_2017,tanimoto_evidence_2020}.
 A cloud-free hydrogen-dominated planetary atmosphere cannot explain such a
-wavelength dependence.  The proposed explanation is that the planet is
-surrounded by a dust cloud \citep{tanimoto_evidence_2020}.
+wavelength dependence.  The planet might therefore be surrounded by a
+dust cloud \citep{tanimoto_evidence_2020}.
 
-A second first could be the detection of direct H$\alpha$ emission from the
+Another first could be the direct detection of H$\alpha$ emission from the
 planet itself \citep{johnskrull_h_2016}.  While the stellar chromosphere emits
@@
 the excess H$\alpha$ emission is 87$\,$km$\,$s$^{-1}$, and its equivalent width
-is 70-80\% that from the stellar chromosphere \citep{johnskrull_h_2016}.  The
+is 70-80\% that of the stellar chromosphere \citep{johnskrull_h_2016}.  The
 proposed explanation is that the outflowing mass from the planet may explain
@@
 
-There are perhaps a few challenges to the planetary interpretation.  They
-include that the planet does not seem to emit infrared radiation in occulation,
-at least anywhere near the expected amplitude \citep{yu_tests_2015}.  In
-addition, despite measurement attempts by multiple investigators, it does not
-seem to show the Rossiter effect at the amplitude expected given the rapid
-stellar rotation and large planet size
-\citep{yu_tests_2015,ciardi_followup_2015}.  Finally, detailed modelling of the
-``precession + gravity darkening'' transits has shown that the necessary degree
-of gravity darkening is too great, given the spectroscopically observed
-equatorial velocity \citep{howarth_reappraisal_2016}.  Additionally, the
-gravity-darkened star would need to show photometric variability from its
-precession rotation axis, and such variability has not been observed.
+There are perhaps a few challenges to the planetary interpretation (if
+these ``features'' are not already seen as such).  They include that
+the planet does not seem to emit infrared radiation in occulation, at
+least anywhere near the expected amplitude \citep{yu_tests_2015}.  In
+addition, despite measurement attempts by multiple investigators, it
+does not seem to show the Rossiter effect at the amplitude expected
+given the rapid stellar rotation and large planet size
+\citep{yu_tests_2015,ciardi_followup_2015}.  Finally, detailed
+modelling of the ``precession + gravity darkening'' transits has shown
+that the necessary degree of gravity darkening is too great, given the
+spectroscopically observed equatorial velocity
+\citep{howarth_reappraisal_2016}.  Additionally, as the gravity-darkened
+star precessed about its rotation axis, it would need to show
+photometric variability that has not been observed.
 
-While the planetary interpetation has challenges, so too do alternative
-explanations.  High-latitude accretion hotspots might produce the observed
-H$\alpha$ variability, but \ptfo\ does not have an infrared (IR) excess
-associated with the presence of an inner disk.  Low-latitude spots, hot or
-cold, struggle to produce the necessary dip durations.  A circumstellar cloud
-of dust or gas might be viable, but clumps must be invoked to explain the
-short-duration flux dips.  The material might be a remnant of the primordial
-gas disk, or it could be ejected and entrained in the stellar magnetosphere.
+While the planetary interpetation has challenges, alternative
+explanations do as well.  High-latitude accretion hotspots might
+produce the observed H$\alpha$ variability, but \ptfo\ does not have
+an infrared (IR) excess associated with the presence of an inner disk.
+Low-latitude starspots, hot or cold, struggle to produce the necessary
+dip durations.  A circumstellar cloud of dust or gas might be
+entrained in the stellar magnetosphere, but clumps would need to be
+invoked to explain the short-duration flux dips.
 
-Section~\ref{sec:observations} presents newly available observations from the
-Transiting Exoplanet Survey Satellite (TESS; \citealt{ricker_transiting_2015})
-and Gaia \citep{gaia_collaboration_gaia_2018}.  Analyzing the TESS data in
-Section~\ref{sec:tess}, we find two distinct signals.  A long sinusoid repeats
-every 12.0 hours, and is probably stellar rotation.  A short ``dip + complex
-modulation'' repeats every 10.7 hours.  Analyzing the Gaia data in
-Section~\ref{sec:gaia}, we show that \ptfo\ is quite likely a photometric
-binary.  In Section~\ref{sec:discussion}, we note that the orbital phase of the
-dip seems to have instantaneously changed, perhaps twice.  We argue that
-PTFO$\,$8-8695 is a binary M dwarf in which one star shows a rotation signal,
-and the other is showing ``transient dipping'' that has been observed in other
-young M dwarfs, and is likely caused by eclipses of warm particulate clouds.
+In the following, Section~\ref{sec:observations} presents newly
+available observations from the Transiting Exoplanet Survey Satellite
+(TESS; \citealt{ricker_transiting_2015}) and Gaia
+\citep{gaia_collaboration_gaia_2018}.  Analyzing the TESS data in
+Section~\ref{sec:tess}, we find two distinct signals.  A long sinusoid
+repeats every 12.0 hours, and is probably stellar rotation.  A short
+``dip + complex modulation'' repeats every 10.7 hours.  Analyzing the
+Gaia data in Section~\ref{sec:gaia}, we show that \ptfo\ is quite
+likely a photometric binary.  In Section~\ref{sec:discussion}, we note
+that the orbital phase of the dip seems to have instantaneously
+changed, perhaps twice.  We argue that PTFO$\,$8-8695 is a binary M
+dwarf in which one star shows a rotation signal, and the other is
+showing ``transient dipping'' that has been observed in other young M
+dwarfs, and is likely caused by eclipses of warm particulate clouds.
 Section~\ref{sec:conclusions} summarizes our main points.
@@
 smallest to largest groups, \ptfo\ was identified as being a member of
-the following subgroups:
+the following nested subgroups:
 \begin{equation}
-  {\rm PTFO\,8\text{-}8695}
-  \in {\rm 25\,Ori\text{-}1}
+  %{\rm PTFO\,8\text{-}8695}
+  % \in
+  {\rm 25\,Ori\text{-}1}
   \subset {\rm 25\,Ori}
@@
 \end{figure*}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.45\textwidth]{f4.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{ {\bf Scene used for blend analysis.}
-		{\it Top:} Mean TESS image of \ptfo\ over Sector~6, with a
-		log-stretch.  The position of \ptfo\ is shown with a yellow
-		star.  Neighbors with $T<17$ are shown with orange crosses.  The
-		apertures used to measure the background and target star flux are
-		shown with \texttt{X} and \texttt{/} hatches, respectively.
-		{\it Bottom:} Digitized Sky Survey $R$-band image of the same
-		field, with a linear stretch. The circles show apertures of radii
-		1, 1.5, and 2.25 pixels used in part of our blend analysis.  The
-		pixel level TESS data show that ``Star A''  does not contribute
-		variability at either of the two observed periods (see
-		Section~\ref{subsec:blend}).
-		\label{fig:scene}
-	}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.5\textwidth]{f5.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Timing residuals for \ptfob\ from a decade of monitoring.}
-		Black points are times of ``dips'', minus the indicated linear
-		ephemeris.  The $y$-axis is given in units of phase for the
-		short-period signal.  The star shows the binned TESS ephemeris.
-		``Dips'' have been observed by \citet{van_eyken_ptf_2012},
-		\citet{ciardi_followup_2015}, \citet{yu_tests_2015},
-		\citet{raetz_yeti_2016}, \citet{onitsuka_multi-color_2017}, and
-		\citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
-		at phase 0 in mid-2019) are consistent with noise, and were likely
-		reported because something was {\it expected}, rather than
-		convincingly {\it observed}.  Horizontal dashed lines are drawn at
-		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting what
-		could be either a numerical coincidence or an observational bias.
-		The orbital phase observed by TESS is consistent with that of
-		\citet{tanimoto_evidence_2020}, and quite different from the
-		original phase.
-		\label{fig:o_minus_c}
-	}
-\end{figure}
 
@@
 
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.45\textwidth]{f4.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{ {\bf Scene used for blend analysis.}
+		{\it Top:} Mean TESS image of \ptfo\ over Sector~6, with a
+		log-stretch.  The position of \ptfo\ is shown with a yellow
+		star.  Neighbors with $T<17$ are shown with orange crosses.  The
+		apertures used to measure the background and target star flux are
+		shown with \texttt{X} and \texttt{/} hatches, respectively.
+		{\it Bottom:} Digitized Sky Survey $R$-band image of the same
+		field, with a linear stretch. The circles show apertures of radii
+		1, 1.5, and 2.25 pixels used in part of our blend analysis.  The
+		pixel level TESS data show that ``Star A''  does not contribute
+		variability at either of the two observed periods (see
+		Section~\ref{subsec:blend}).
+		\label{fig:scene}
+	}
+\end{figure}
+
 The TESS pixels are $\approx21$'' per side, and so before making an
@@
 	\vspace{-0.7cm}
-	\caption{ {\bf HR diagram of \ptfo\ and members of 25$\,$Ori-1.}
+	\caption{ {\bf HR diagram of \ptfo\ and late-type members of 25$\,$Ori-1.}
   Members of the 25$\,$Ori-1 group (black circles) were identified by
@@
   the mean 25$\,$Ori-1 right ascension, declination, and parallax.
-  \ptfo\ seems likely to be a photometric binary.
   $G$ is the Gaia broadband, $Bp$ is Gaia blue, $Rp$ is Gaia red, and
   $\omega_{\rm as}$ is the parallax in arcseconds.
+  The $x$-axis limits have been set to show only K and M dwarf
+  members, to accentuate \ptfo's separation from the
+  single-star sequence.
 	\label{fig:gaia}
@@
 %studies.
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.5\textwidth]{f5.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Timing residuals for \ptfob\ from a decade of monitoring.}
+		Black points are times of ``dips'', minus the indicated linear
+		ephemeris.  The $y$-axis is given in units of phase for the
+		short-period signal.  The star shows the binned TESS ephemeris.
+		``Dips'' have been observed by \citet{van_eyken_ptf_2012},
+		\citet{ciardi_followup_2015}, \citet{yu_tests_2015},
+		\citet{raetz_yeti_2016}, \citet{onitsuka_multi-color_2017}, and
+		\citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
+		at phase 0 in mid-2019) are consistent with noise, and were likely
+		reported because something was {\it expected}, rather than
+		convincingly {\it observed}.  Horizontal dashed lines are drawn at
+		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting what
+		could be either a numerical coincidence or an observational bias.
+		The orbital phase observed by TESS is consistent with that of
+		\citet{tanimoto_evidence_2020}, and quite different from the
+		original phase.
+		\label{fig:o_minus_c}
+	}
+\end{figure}
 

### 2020-04-17 paper/ms.tex
made sick posterior table

@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f3.pdf}
+		\includegraphics[width=0.99\textwidth]{f3.pdf}
 	\end{center}
@@
 
-\input{model_comparison_table.tex}
-%\input{model_posterior_table.tex}
-
 We considered nine models, with the number of harmonics per frequency
@@
 (Table~\ref{tab:modelcompare}).  The model with the lowest BIC had
-two harmonics at the short 11.74$\,$hr period, and two harmonics
+two harmonics at the short 10.74$\,$hr period, and two harmonics
 at the long 11.96$\,$hr period.  All of the models have reduced
@@
 \subsection{Short period dip}
-The dip lasts about 45 minutes, and seems to re-occur every 11.74
+The dip lasts about 45 minutes, and seems to re-occur every 10.74
 hours
@@
 
-%\clearpage
+\clearpage
+\input{model_comparison_table.tex}
+\input{model_posterior_table.tex}
+\clearpage
+
 \bibliographystyle{yahapj}                            

### 2020-04-17 drivers/make_posterior_table.py

@@ -1,0 +1,166 @@
+import os, re
+from copy import deepcopy
+import numpy as np, pandas as pd, matplotlib.pyplot as plt, pymc3 as pm
+from os.path import join
+from itertools import product
+
+from billy.modelfitter import ModelFitter, ModelParser
+import billy.plotting as bp
+from billy.convenience import (
+    get_clean_ptfo_data, get_ptfo_data, initialize_ptfo_prior_d, get_bic
+)
+from billy import __path__
+
+def normal_str(mu, sd, fmtstr=None):
+    # fmtstr: e.g., "({:.5f}; {:.2f})"
+    if fmtstr is None:
+        return '$\\mathcal{N}'+'({}; {})$'.format(mu, sd)
+    else:
+        return '$\\mathcal{N}'+'{}$'.format(fmtstr).format(mu, sd)
+
+
+def uniform_str(lower, upper, fmtstr=None):
+    # fmtstr: e.g., "({:.5f}; {:.2f})"
+    if fmtstr is None:
+        return '$\\mathcal{U}'+'({}; {})$'.format(lower, upper)
+    else:
+        return '$\\mathcal{U}'+'{}$'.format(fmtstr).format(lower, upper)
+
+
+def main(modelid):
+
+    OVERWRITE = 0
+    REALID = 'PTFO_8-8695'
+    RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
+    PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID),
+                           '20200413_v0')
+
+    summarypath = os.path.join(PLOTDIR, 'posterior_table_raw.csv')
+
+    pklpath = os.path.join(
+        os.path.expanduser('~'), 'local', 'billy',
+        '{}_model_{}.pkl'.format(REALID, modelid)
+    )
+    np.random.seed(42)
+
+    x_obs, y_obs, y_err = get_clean_ptfo_data()
+
+    mp = ModelParser(modelid)
+    prior_d = initialize_ptfo_prior_d(x_obs, mp.modelcomponents)
+
+    if not os.path.exists(summarypath):
+
+        m = ModelFitter(modelid, x_obs, y_obs, y_err, prior_d, plotdir=PLOTDIR,
+                        pklpath=pklpath, overwrite=OVERWRITE)
+
+        df = pm.summary(
+            m.trace, varnames=list(prior_d.keys()),
+            round_to=10, kind='stats'
+        )
+
+        df.to_csv(summarypath, index=True)
+
[…]

### 2020-04-17 paper/ms.tex
paper update

@@
 		\citet{ciardi_followup_2015}, \citet{yu_tests_2015},
-		\citet{raetz_yeti_2016}, \citet{onitsuka_multi-color_2017}, and
+		\citet{raetz_yeti_2016}, \citet{onitsuka_multicolor_2017}, and
 		\citet{tanimoto_evidence_2020}.  Certain dips ({\it e.g.}, the one
@@
 		convincingly {\it observed}.  Horizontal dashed lines are drawn at
-		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting what
-		could be either a numerical coincidence or an observational bias.
-		The orbital phase observed by TESS is consistent with that of
-		\citet{tanimoto_evidence_2020}, and quite different from the
-		original phase.
+		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting a
+    possible observational bias.  The orbital phase observed by TESS
+    is consistent with that of \citet{tanimoto_evidence_2020}, and
+    quite different from the original phase.
 		\label{fig:o_minus_c}
@@
 
-% The main result that both models agree on: the extra power at the
-% orbital frequency seems to not be ellipsoidal. Both of these models
-% prefer to put the power at 1x the orbital frequency, not 2x the
-% orbital frequency! This gives both Aorb0 and Borb0 non-zero, while
-% Aorb1 and Borb1 seem to be more consistent with zero.
+\section{Physical Interpretation}
 
-\section{Plausible Physical Interpretations}
+Given the evidence, we believe that PTFO$\,$8-8695 is a binary M dwarf
+in which one star shows the ``long'' rotation signal, and the other is
+showing ``transient dipping'' that has also been observed in other
+young M dwarfs.
+
+Many other young M-dwarf photometric binaries
+have been observed in {\it e.g.,} Upper Sco and the Pleiades to show
+multiple periods \citep{rebull_usco_2018,stauffer_rotevol_2018}.
+Typically, the periods are both rotational modulation.
+
+However, occasionally one or both periods can be ``scallop-shell''
+variability, {\it e.g.}, EPIC 203956650 in $\rho$Oph
+\citep{rebull_usco_2018}.
+
+There are
+
+The gas in the disks is gone (CITE).
+The upper limits on the SED from \citet{yu_tests_2015} imply X, Y, Z.
+The stars are therefore presumably no longer ``magnetically locked''
+to their disks.
+This is consistent with the $\approx$half-day periodicities of both
+signals.
+In the broader context of rapidly rotating young M-dwarfs, ``magnetic
+locking'' causes classical disked T-Tauri stars tend to rotate more
+{\it slowly}, and almost none have periods less than two days
+\citep[{\it e.g.},][]{rebull_rotation_2020}.
+
+The main physical question is what is causing the ``transient
+dipping''. This is an unsolved problem not only for \ptfo\ but also
+for an entire class of young rapidly rotating M-dwarfs
+
+The roughly half-day periods of both signals imply that
 

### 2020-04-17 paper/ms.tex
make HR diagram even better

@@
 	\vspace{-0.7cm}
-	\caption{ {\bf HR diagram of \ptfo\ and late-type members of 25$\,$Ori-1.}
+	\caption{ {\bf Hertzsprung-Russell diagram of \ptfo\ and late-type members of 25$\,$Ori-1.}
   Members of the 25$\,$Ori-1 group (black circles) were identified by
   \citet{kounkel_apogee2_2018} through clustering on six-dimensional
-  Gaia-DR2 and APOGEE-2 data.  We defined the reference
-  ``neighborhood'' (gray circles) as the group of at most $10^4$
+  Gaia-DR2 and APOGEE-2 data.  The reference
+  ``neighborhood'' (gray circles) is the group of at most $10^4$
   randomly selected non-member stars within 5 standard deviations of
   the mean 25$\,$Ori-1 right ascension, declination, and parallax.
+  It contains members of the full Orion complex with a wider spread of ages,
+  in addition to field interlopers.
   $G$ is the Gaia broadband, $Bp$ is Gaia blue, $Rp$ is Gaia red, and

### 2020-04-17 billy/plotting.py

@@ -771,7 +771,7 @@ def plot_hr(outdir):
     ax.scatter(
         nbhd_df['phot_bp_mean_mag']-nbhd_df['phot_rp_mean_mag'], nbhd_yval,
         c='gray', alpha=1., zorder=2, s=7, rasterized=True, linewidths=0,
-        label='Neighborhood'
+        label='Neighborhood', marker='.'
     )
 
     yval = group_df_dr2['phot_g_mean_mag'] + 5*np.log10(group_df_dr2['parallax']/1e3) + 5

### 2020-04-17 billy/plotting.py
phasefold plot: add point dumper

@@ -346,6 +346,20 @@ def plot_phasefold_map(m, d, outpath):
     axs[1].scatter(orb_bd['binnedphases'], orb_bd['binnedmags'], color='black',
                    s=8, alpha=1, zorder=5, linewidths=0)
 
+    out_d = {
+        'orb_d': orb_d,
+        'orb_bd': orb_bd,
+        'rot_d': rot_d,
+        'rot_bd': rot_bd
+    }
+    pklpath = os.path.join(
+        os.path.dirname(outpath),
+        os.path.basename(outpath).replace('.png','_points.pkl')
+    )
+    with open(pklpath, 'wb') as buff:
+        pickle.dump(out_d, buff)
+    print('made {}'.format(pklpath))
+
     txt1 = '$P_{{\mathrm{{s}}}}$ = {:.5f}$\,$d'.format(P_orb)
     axs[1].text(0.98, 0.98, txt1, ha='right', va='top',
                 transform=axs[1].transAxes, bbox=props, zorder=3)

### 2020-04-18 billy/plotting.py
add plot_brethren

@@ -18,11 +18,14 @@ Plots:
 
     plot_O_minus_C
 
+    plot_brethren
+
 Convenience:
     savefig
     format_ax
 """
 import os, corner, pickle
+from glob import glob
 import numpy as np, matplotlib.pyplot as plt
 from datetime import datetime
 from pymc3.backends.tracetab import trace_to_dataframe
@@ -32,12 +35,17 @@ from billy.convenience import flatten as bflatten
 from billy.convenience import get_clean_ptfo_data
 from billy.models import linear_model
 
-from astrobase.lcmath import phase_magseries, phase_bin_magseries
+from astrobase.lcmath import (
+    phase_magseries, phase_bin_magseries, sigclip_magseries
+)
+from astrobase import periodbase
+
 from astropy.stats import LombScargle
 from astropy import units as u, constants as const
+from astropy.io import fits
+from astropy.time import Time
 
 from numpy import array as nparr
-from astropy.time import Time
 
 def plot_periodogram(outdir, islinear=True):
 
@@ -962,3 +970,202 @@ def plot_O_minus_C(
     fig.tight_layout(h_pad=0, w_pad=0)
 
     savefig(fig, savpath, dpi=350)
+
+
+def plot_brethren(outdir):
+
+    from transitleastsquares import transitleastsquares
+
+    # get campaign 2 upper sco everest light-curves.
+    # NOTE: 204270520 has a C15 LC, but it's the only one.
+    epic_ids = [
+        "204143627",
+        "204270520",
+        "204321142",
+        # "205046529", # kind of messed up, b/c it's a binary
+        "205483258",
+        '204787516',
+    ]
+    lc_dict = {}
+
+    k2pkl = os.path.join(outdir, 'k2data.pkl')
+    if not os.path.exists(k2pkl):
+        for epic_id in epic_ids:
+
+            lc_dict[epic_id] = {}
+
+            lcpath = glob(
+                '/Users/luke/Dropbox/proj/billy/data/analogs/hlsp_everest_k2_llc_{}-c02_kepler_v2.0_lc/*fits'.format(epic_id)
+            )
+            assert len(lcpath)==1
+
+            hdul = fits.open(lcpath[0])
+            time = hdul[1].data['TIME']
+            flux = hdul[1].data['FLUX']
+            qual = hdul[1].data['QUALITY']
+            hdul.close()
+
+            plt.close('all')
+            fig = plt.figure(figsize=(16,4))
[…]

### 2020-04-18 drivers/plot_brethren.py

@@ -1,0 +1,8 @@
+import os
+import billy.plotting as bp
+from billy import __path__
+
+RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
+PLOTDIR = os.path.join(RESULTSDIR, 'brethren')
+
+bp.plot_brethren(PLOTDIR)

### 2020-04-18 drivers/real.py

@@ -20,7 +20,7 @@ def main(modelid):
     cornerplot = 1
     splitsignalplot = 1 if 'Porb' in modelid and 'Prot' in modelid else 0
 
-    OVERWRITE = 0
+    OVERWRITE = 1
     REALID = 'PTFO_8-8695'
     RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
     PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID),
@@ -82,10 +82,10 @@ def main(modelid):
 
 if __name__ == "__main__":
 
-    DEBUG = 0
+    DEBUG = 1
 
     if DEBUG:
-        main('transit_1sincosPorb_1sincosProt')
+        main('transit_2sincosPorb_2sincosProt')
 
     else:
         for N, M in product(range(1,4), range(1,4)):

### 2020-04-18 paper/ms.tex
update paper

@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f6.pdf}
+		\includegraphics[width=0.8\textwidth]{f5.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.5\textwidth]{f5.pdf}
+		\includegraphics[width=0.5\textwidth]{f6.pdf}
 	\end{center}
@@
 \section{Physical Interpretation}
+
+%FIXME: the PTFO lightcurve looks so much worse! Why is it less dense??
+\begin{figure*}[hbtp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f7.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf \ptfo\ and its brethren.}
+  	Five transient and persistent flux dip stars, selected based on their visual similarity to the short-period signal in \ptfo, are as follows.
+  	EPIC 204143627, 1.1250d, dips change depth too
+  	EPIC 204321142  Usco flux dip, 0.476d
+  	EPIC 204270520  USco flux dip, 0.512d
+  	% EPIC 205046529B, 1.8358d, indicates dips can be at weird phases too
+  	RIK-210 = EPIC 205483258
+ 	EPIC 204787516  USco flux dip or possible? EB, 0.487d
+ 	We found these objects through studies by \citet{stauffer_orbiting_2017}, \citet{david_transient_2017}, and \citet{rebull_usco_2018}.
+		\label{fig:brethren}
+	}
+\end{figure*}
+
+cp ../results/brethren/brethren.pdf f7.pdf
 

### 2020-04-18 paper/collect.sh

@@ -9,9 +9,11 @@ cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.pdf f3.pdf
 
 cp ../results/PTFO_8-8695_results/scene.pdf f4.pdf
 
-cp ../results/ephemeris/O_minus_C.pdf f5.pdf
+cp ../results/cluster_membership/hr.pdf f5.pdf
 
-cp ../results/cluster_membership/hr.pdf f6.pdf
+cp ../results/ephemeris/O_minus_C.pdf f6.pdf
+
+cp ../results/brethren/brethren.pdf f7.pdf
 
 # cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_cornerplot.png f6.png
 # pngquant f6.png --output f6_comp.png

### 2020-04-19 paper/ms.tex
cleanup tex

@@
 % Planet}
-\title{Seeing Double: Indications From {\it TESS} and {\it Gaia} That
-PTFO$\,$8-8695b is Not a Planet}
+%\title{Seeing Double: Indications From {\it TESS} and {\it Gaia} That
+%PTFO$\,$8-8695b is Not a Planet}
+\title{Seeing Double: PTFO$\,$8-8695 is a Photometric Binary With One
+Component Showing Transient Dips }
 
@@
 \section{Introduction}
-If \ptfob\ were a planet, it would be exceptional.  Transiting a sub-10$\,$Myr
-old weak-lined T Tauri M dwarf in Orion, it would be the youngest hot Jupiter
-known \citep{van_eyken_ptf_2012}.  Its orbital period of only 10.7 hours would
-also give it the shortest period of any known hot Jupiter.  With such a short
-period, it would almost certainly have filled its Roche lobe, and would be
-actively losing mass to its host star.  Not only that, but the rapidly rotating
-host star would also be oblate and gravity darkened, and so the planet's orbit
-would likely precess into and out of transitability
+If \ptfob\ were a planet, it would be exceptional.  Transiting a
+sub-10$\,$Myr old weak-lined T Tauri M dwarf in Orion, it would be the
+youngest hot Jupiter known \citep{van_eyken_ptf_2012}.  Its orbital
+period of only 10.7 hours would also give it the shortest period of
+any known hot Jupiter.  With such a short period, it would almost
+certainly have filled its Roche lobe, and would be actively losing
+mass to its host star.  Not only that, but the rapidly rotating host
+star would also be oblate and gravity darkened, and so the planet's
+orbit would likely precess into and out of transitability
 \citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.
@@
 Other lines of evidence would imply further planetary ``firsts'' for
-this planet candidate.  One first would be that its transits are about three
-times deeper in optical bandpasses ({\it e.g.,} $g$-band) than in the near-infrared
-({\it e.g.}, $z$-band) \citep{onitsuka_multicolor_2017,tanimoto_evidence_2020}.
-A cloud-free hydrogen-dominated planetary atmosphere cannot explain such a
+this planet candidate.  One first would be that its transits are about
+three times deeper in optical bandpasses ({\it e.g.,} $g$-band) than
+in the near-infrared ({\it e.g.}, $z$-band)
+\citep{onitsuka_multicolor_2017,tanimoto_evidence_2020}.  A cloud-free
+hydrogen-dominated planetary atmosphere cannot explain such a
 wavelength dependence.  The planet might therefore be surrounded by a
@@
 
-Another first could be the direct detection of H$\alpha$ emission from the
-planet itself \citep{johnskrull_h_2016}.  While the stellar chromosphere emits
-in H$\alpha$, it seems that there is an additional excess H$\alpha$ emission
-that could be in phase with the planetary orbit.  The average velocity width of
-the excess H$\alpha$ emission is 87$\,$km$\,$s$^{-1}$, and its equivalent width
-is 70-80\% that of the stellar chromosphere \citep{johnskrull_h_2016}.  The
-proposed explanation is that the outflowing mass from the planet may explain
+Another first could be the direct detection of H$\alpha$ emission from
+the planet itself \citep{johnskrull_h_2016}.  While the stellar
+chromosphere emits in H$\alpha$, it seems that there is an additional
+excess H$\alpha$ emission that could be in phase with the planetary
+orbit.  The average velocity width of the excess H$\alpha$ emission is
+87$\,$km$\,$s$^{-1}$, and its equivalent width is 70-80\% that of the
+stellar chromosphere \citep{johnskrull_h_2016}.  The proposed
+explanation is that the outflowing mass from the planet may explain
 this excess emission as well \citep{johnskrull_h_2016}.
@@
 spectroscopically observed equatorial velocity
-\citep{howarth_reappraisal_2016}.  Additionally, as the gravity-darkened
-star precessed about its rotation axis, it would need to show
-photometric variability that has not been observed.
+\citep{howarth_reappraisal_2016}.  Additionally, as the
+gravity-darkened star precessed about its rotation axis, it would need
+to show photometric variability that has not been observed.
 
@@
     versus time. The beat period of 4.48 days is visible by eye.  The
-		preferred model plotted underneath the data includes 2 harmonics at the long
-		period $P_{\rm \ell}$, plus 2 harmonics and a transit at the short
-		period $P_{\rm s}$.
+    preferred model plotted underneath the data includes 2 harmonics
+    at the long period $P_{\rm \ell}$, plus 2 harmonics and a transit
+    at the short period $P_{\rm s}$.
 		{\it Upper middle}: Long-period signal, equal to the raw signal
@@
 
-\ptfo\ was observed by TESS with Camera 1, CCD 1, from December
-15, 2018 to January 6, 2019, during the sixth sector of science
-operations \citep{ricker_transiting_2015}.  The star was designated
-TIC 264461976 in the TESS Input Catalog
-\citep{stassun_TIC_2018,stassun_TIC8_2019}.  The pixel data for an
-$11\times11$ array surrounding \ptfo\ were averaged into 2-minute
-stacks by the onboard computer.  Each 2048$\times$2048 image from the
-CCD was also averaged into 30-minute stacks, and saved as a ``full
-frame image'' (FFI).
+\ptfo\ was observed by TESS with Camera 1, CCD 1, from December 15,
+2018 to January 6, 2019, during the sixth sector of science operations
+\citep{ricker_transiting_2015}.  The star was designated TIC 264461976
+in the TESS Input Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.
+The pixel data for an $11\times11$ array surrounding \ptfo\ were
+averaged into 2-minute stacks by the onboard computer.  Each
+2048$\times$2048 image from the CCD was also averaged into 30-minute
+stacks, and saved as a ``full frame image'' (FFI).
 
-The 2-minute stacks for \ptfo\ were then reduced to lightcurves
-by the Science Processing Operations Center (SPOC) at NASA
+The 2-minute stacks for \ptfo\ were then reduced to lightcurves by the
+Science Processing Operations Center (SPOC) at NASA
 Ames~\citep{jenkins_tess_2016}.  Our main analysis used the resulting
@@
 	\vspace{-0.7cm}
-	\caption{ {\bf Hertzsprung-Russell diagram of \ptfo\ and late-type members of 25$\,$Ori-1.}
+  \caption{ {\bf Hertzsprung-Russell diagram of \ptfo\ and late-type
+  members of 25$\,$Ori-1.}
   Members of the 25$\,$Ori-1 group (black circles) were identified by
   \citet{kounkel_apogee2_2018} through clustering on six-dimensional
-  Gaia-DR2 and APOGEE-2 data.  The reference
-  ``neighborhood'' (gray circles) is the group of at most $10^4$
-  randomly selected non-member stars within 5 standard deviations of
-  the mean 25$\,$Ori-1 right ascension, declination, and parallax.
-  It contains members of the full Orion complex with a wider spread of ages,
-  in addition to field interlopers.
-  $G$ is the Gaia broadband, $Bp$ is Gaia blue, $Rp$ is Gaia red, and
-  $\omega_{\rm as}$ is the parallax in arcseconds.
-  The $x$-axis limits have been set to show only K and M dwarf
-  members, to accentuate \ptfo's separation from the
-  single-star sequence.
+  Gaia-DR2 and APOGEE-2 data.  The reference ``neighborhood'' (gray
+  circles) is the group of at most $10^4$ randomly selected non-member
+  stars within 5 standard deviations of the mean 25$\,$Ori-1 right
+  ascension, declination, and parallax.  It contains members of the
+  full Orion complex with a wider spread of ages, in addition to field
+  interlopers.  $G$ is the Gaia broadband, $Bp$ is Gaia blue, $Rp$ is
+  Gaia red, and $\omega_{\rm as}$ is the parallax in arcseconds.  The
+  $x$-axis limits have been set to show only K and M dwarf members, to
+  accentuate \ptfo's separation from the single-star sequence.
 	\label{fig:gaia}
@@
 
-To assess the cluster membership and potential
-binarity of \ptfo, we needed to identify stars with which it was and was
-not coeval.  The simplest way to do this---clustering based on
-six-dimensional position and kinematic information---had already been
-done by \citet{kounkel_apogee2_2018}.  For simplicity, we considered
-the members they identified brighter than $G_{\rm Rp}$ of 16.  This
+To assess the cluster membership and potential binarity of \ptfo, we
+needed to identify stars with which it was and was not coeval.  The
+simplest way to do this---clustering based on six-dimensional position
+and kinematic information---had already been done by
+\citet{kounkel_apogee2_2018}.  For simplicity, we considered the
+members they identified brighter than $G_{\rm Rp}$ of 16.  This
 yielded 149 stars in 25$\,$Ori-1, mostly M dwarfs.
-\citet{kounkel_apogee2_2018} identified seven other smaller
-groups in the Orion complex near the Be star 25$\,$Ori. These groups
-received higher numbers, {\it e.g.}, 25$\,$Ori-2.
+\citet{kounkel_apogee2_2018} identified seven other smaller groups in
+the Orion complex near the Be star 25$\,$Ori. These groups received
+higher numbers, {\it e.g.}, 25$\,$Ori-2.
 
-To define a set of non-member stars that nonetheless had comparable selection
-functions, we defined a reference ``neighborhood'' as the group of at
-most $10^4$ randomly selected non-member stars within 5 standard
-deviations of the mean 25$\,$Ori-1 right ascension, declination, and
-parallax.  We queried these stars using the \texttt{astroquery}
-package, which provides a convenient interface to the Gaia archive
-(CITE, CITE).  This yielded 1{,}819 neighbors.  While some of
-these stars may indeed be members of the Orion complex, or even of
-25$\,$Ori-1, enforcing this cut on positions and parallaxes ensures
-that we are querying stars with comparable amounts of interstellar
-reddening.
+To define a set of non-member stars that nonetheless had comparable
+selection functions, we defined a reference ``neighborhood'' as the
+group of at most $10^4$ randomly selected non-member stars within 5
+standard deviations of the mean 25$\,$Ori-1 right ascension,
+declination, and parallax.  We queried these stars using the
+\texttt{astroquery} package, which provides a convenient interface to
+the Gaia archive (CITE, CITE).  This yielded 1{,}819 neighbors.  While
+some of these stars may indeed be members of the Orion complex, or
+even of 25$\,$Ori-1, enforcing this cut on positions and parallaxes
+ensures that we are querying stars with comparable amounts of
+interstellar reddening.
 
@@
 
-Given our detection of two separate signals, whether
-\ptfo\ could be a photometric binary was of great interest.
-Figure~\ref{fig:gaia} shows the HR diagram from which we assessed
-this issue.
-The diagram shows that \ptfo\ is $\approx$0.8 magnitudes brighter than the average
-25$\,$Ori-1 star of the same color.
-In other words, it is about twice as bright.
-It also seems to be on the photometric binary track of the cluster, which has a few other stars.
+Given our detection of two separate signals, whether \ptfo\ could be a
+photometric binary was of great interest.  Figure~\ref{fig:gaia} shows
+the HR diagram from which we assessed this issue.  The diagram shows
+that \ptfo\ is $\approx$0.8 magnitudes brighter than the average
+25$\,$Ori-1 star of the same color.  In other words, it is about twice
+as bright.  It also seems to be on the photometric binary track of the
+cluster, which has a few other stars.
 
-The implication is that either 
-{\it (i)} \ptfo\ is notably younger than the kinematically identical cluster members, or
-{\it (ii)} \ptfo\ is a photometric binary.
-Given the independent presence of two resolved signals,
-we favor the interpretation that \ptfo\ is a binary star system.
+The implication is that either {\it (i)} \ptfo\ is notably younger
+than the kinematically identical cluster members, or {\it (ii)} \ptfo\
+is a photometric binary.  Given the independent presence of two
+resolved signals, we favor the interpretation that \ptfo\ is a binary
+star system.
 
@@
 
-The standard interpretation of sinusoidal modulations for a pre-main-sequence M
-dwarf is that we are observing a stellar rotation period of 11.96 hours.  This
-is the dominant signal in the system with 10\% amplitude, and there is no
-evidence to suggest that this signal has any other origin.
+The standard interpretation of sinusoidal modulations for a
+pre-main-sequence M dwarf is that we are observing a stellar rotation
+period of 11.96 hours.  This is the dominant signal in the system with
+10\% amplitude, and there is no evidence to suggest that this signal
+has any other origin.
 
-The discovery study by \citet{van_eyken_ptf_2012} saw the same signal ({\it
-e.g.}, their Figure~7), but identified its alias as a periodogram peak at
-$0.9985 \pm 0.0061\,$days. They ascribed it to their observing cadence, because
-of its close correspondence to the sidereal day.  While the TESS data can show
-significant reflected light from the Earth \citep[{\it
-e.g.},][]{luger_tess_2019}, our pixel-level analysis showed that the signal is
-specific to only pixels near \ptfo, and no other pixels.  We therefore conclude
-that the signal is not a systematic.
+The discovery study by \citet{van_eyken_ptf_2012} saw the same signal
+({\it e.g.}, their Figure~7), but identified its alias as a
+periodogram peak at $0.9985 \pm 0.0061\,$days. They ascribed it to
+their observing cadence, because of its close correspondence to the
+sidereal day.  While the TESS data can show significant reflected
+light from the Earth \citep[{\it e.g.},][]{luger_tess_2019}, our
+pixel-level analysis showed that the signal is specific to only pixels
+near \ptfo, and no other pixels.  We therefore conclude that the
+signal is not a systematic.
 
-We are not the first to reach the conclusion that the long period sinusoidal
-modulation is astrophysical.  A follow-up study by
+We are not the first to reach the conclusion that the long period
+sinusoidal modulation is astrophysical.  A follow-up study by
 \citet{koen_multicolour_2015} identified the same modes and aliases as
-\citet{van_eyken_ptf_2012}, but argued that the $0.50\,{\rm d}$ signal was
-astrophysical.  Using archival photometry and photometry from the YETI global
-telescope network, \citet{raetz_yeti_2016} eventually came to the conclusion
-that the that the $0.50\,{\rm d}$ signal was indeed from stellar rotation.  The
-TESS data strongly support this conclusion.
+\citet{van_eyken_ptf_2012}, but argued that the $0.50\,{\rm d}$ signal
+was astrophysical.  Using archival photometry and photometry from the
+YETI global telescope network, \citet{raetz_yeti_2016} eventually came
+to the conclusion that the that the $0.50\,{\rm d}$ signal was indeed
+from stellar rotation.  The TESS data strongly support this
+conclusion.
 
@@
   \caption{ {\bf \ptfo\ and its brethren.}
-  	Five transient and persistent flux dip stars, selected based on their visual similarity to the short-period signal in \ptfo, are as follows.
+    Five transient and persistent flux dip stars, selected based on
+    their visual similarity to the short-period signal in \ptfo, are
+    as follows.
   	EPIC 204143627, 1.1250d, dips change depth too
@@
 
-cp ../results/brethren/brethren.pdf f7.pdf
-
 Given the evidence, we believe that PTFO$\,$8-8695 is a binary M dwarf
@@
 
-Many other young M-dwarf photometric binaries
-have been observed in {\it e.g.,} Upper Sco and the Pleiades to show
-multiple periods \citep{rebull_usco_2018,stauffer_rotevol_2018}.
-Typically, the periods are both rotational modulation.
+Many other young M-dwarf photometric binaries have been observed in
+{\it e.g.,} Upper Sco and the Pleiades to show multiple periods
+\citep{rebull_usco_2018,stauffer_rotevol_2018}.  Typically, the
+periods are both rotational modulation.
 

### 2020-04-20 paper/ms.tex
update paper

@@
 %PTFO$\,$8-8695b is Not a Planet}
-\title{Seeing Double: PTFO$\,$8-8695 is a Photometric Binary With One
-Component Showing Transient Dips }
+\title{Seeing Double: Transient Dips and Photometric Binarity in
+PTFO$\,$8-8695}
 
@@
 \begin{abstract}
-  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10 million year old
-  Orion-OB1a cluster. We inspected data from TESS and Gaia to clarify whether
-  it is actually a planet.  The TESS lightcurve shows that the dominant
-  variability in this system is a sinusoid with a ``long'' period $P_{\rm
-  \ell}=11.96\,$hr, likely caused by stellar rotation.  Also present is a
-  complex signal, previously identified as the planet candidate, that repeats
-  with a ``short'' period $P_{\rm s}= 10.74\,$hr.  The two signals beat every
-  4.48 days.  Although there is a dip in the short-period signal, ground-based
-  photometry from the past decade shows that the orbital phase of the dip seems
-  to have instantaneously jumped, at least once, and perhaps twice.  Planets do
-  not ``jump'' in orbital phase.  Furthermore, the Gaia data show that
-  PTFO$\,$8-8695 is probably a photometric binary.  Given the evidence, we
-  believe that PTFO$\,$8-8695 is a binary M dwarf in which one star shows the
-  ``long'' rotation signal, and the other is showing ``transient dipping'' that
-  has also been observed in other young M dwarfs.  The planetary interpretation
-  seems unlikely.
+  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10 million year
+  old Orion-OB1a cluster. We inspected data from TESS and Gaia to
+  clarify whether it is actually a planet.  The Gaia data show that
+  PTFO$\,$8-8695 is a photometric binary with respect to members of
+  its kinematic group.  The TESS lightcurve shows that the dominant
+  variability in this system is a sinusoid with a ``long'' period
+  $P_{\rm \ell}=11.96\,$hr, presumably caused by stellar rotation.
+  Also present is a complex signal, previously identified as the
+  planet candidate, that repeats with a ``short'' period $P_{\rm s}=
+  10.74\,$hr.  The two signals beat every 4.48 days.  Although there
+  is a dip in the short-period signal, ground-based photometry from
+  the past decade shows that the orbital phase of the dip seems to
+  have instantaneously jumped, at least once, and perhaps twice.
+  Planets do not ``jump'' in orbital phase.  Given the evidence, we
+  believe that PTFO$\,$8-8695 is a binary M dwarf in which one star
+  shows the long rotation signal, and the other star is showing
+  ``transient dipping'' that has been seen in a few other young
+  weak-lined T Tauri stars.  The origin of these transient dips is
+  still undetermined, but our preferred explanation is eclipses of
+  clouds of gas or dust at the Keplerian co-rotation radius.
 \end{abstract}
 
-%TODO
+\object{PTFO 8-8695}
+\object{PTFO 8-8695b}
+\object{CVSO 30}
+
 \keywords{
 	Exoplanet evolution (491),
-	Open star clusters (1160),
+  Pre-main sequence stars (1290),
 	Stellar ages (1581),
 	Stellar rotation (1629),
-	Variable stars (1761)
+	Variable stars (1761),
+  Low mass stars (2050)
 }
@@
 stellar chromosphere \citep{johnskrull_h_2016}.  The proposed
-explanation is that the outflowing mass from the planet may explain
-this excess emission as well \citep{johnskrull_h_2016}.
+explanation is that outflowing mass from the planet may explain this
+excess emission as well \citep{johnskrull_h_2016}.
 
@@
 gravity-darkened star precessed about its rotation axis, it would need
-to show photometric variability that has not been observed.
+to show photometric variability that has not been observed.  While the
+planetary interpetation clearly faces challenges, alternative
+explanations do as well.  High-latitude accretion hotspots might
+produce the observed H$\alpha$ variability, and require fine-tuning
+produce dips of the approach duration. Furthermore, \ptfo\ does not
+have an infrared (IR) excess associated with the presence of an inner
+disk \citep[{\it e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude
+starspots, hot or cold, struggle to produce the necessary dip
+durations.
 
-While the planetary interpetation has challenges, alternative
-explanations do as well.  High-latitude accretion hotspots might
-produce the observed H$\alpha$ variability, but \ptfo\ does not have
-an infrared (IR) excess associated with the presence of an inner disk.
-Low-latitude starspots, hot or cold, struggle to produce the necessary
-dip durations.  A circumstellar cloud of dust or gas might be
-entrained in the stellar magnetosphere, but clumps would need to be
-invoked to explain the short-duration flux dips.
+Alternatively, between 0.1\% and 1\% of rapidly rotating low-mass
+stars in $\mathcal{O}$(10)$\,$Myr old associations show narrow dips in
+phase with strong stellar rotation signals \citep{rebull_usco_2018}.
+The dips can persist over months, but their depths often vary, and
+sometimes change immediately after stellar flares.  The explanation
+proposed by {\it e.g.}, \citet{stauffer_orbiting_2017} and
+\citet{david_transient_2017} is that a circumstellar cloud of dust or
+gas might be orbiting near the Keplerian co-rotation radius.  For this
+explanation to be viable for \ptfo, a clear determination of the
+stellar rotation period is necessary.  To date, this has not been
+possible
+\citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.
 
-In the following, Section~\ref{sec:observations} presents newly
+We begin in Section~\ref{sec:observations} by describing newly
 available observations from the Transiting Exoplanet Survey Satellite
 (TESS; \citealt{ricker_transiting_2015}) and Gaia
-\citep{gaia_collaboration_gaia_2018}.  Analyzing the TESS data in
-Section~\ref{sec:tess}, we find two distinct signals.  A long sinusoid
-repeats every 12.0 hours, and is probably stellar rotation.  A short
-``dip + complex modulation'' repeats every 10.7 hours.  Analyzing the
-Gaia data in Section~\ref{sec:gaia}, we show that \ptfo\ is quite
-likely a photometric binary.  In Section~\ref{sec:discussion}, we note
-that the orbital phase of the dip seems to have instantaneously
-changed, perhaps twice.  We argue that PTFO$\,$8-8695 is a binary M
+\citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows two
+distinct signals, which we extract and analyze in
+Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
+hours, and is probably stellar rotation.  A short-period ``dip +
+complex modulation'' repeats every 10.7 hours.  Analyzing the Gaia
+data in Section~\ref{sec:gaia}, we show that relative to its kinematic
+group, \ptfo\ is a photometric binary.  We collect and discuss the
+puzzle pieces in Section~\ref{sec:discussion}.  The orbital phase of
+the dip seems to have instantaneously changed over the past decade.
+In addition, a number of other young stars show similar lightcurve
+morphologies.  We therefore argue that PTFO$\,$8-8695 is a binary M
 dwarf in which one star shows a rotation signal, and the other is
-showing ``transient dipping'' that has been observed in other young M
-dwarfs, and is likely caused by eclipses of warm particulate clouds.
+showing ``transient dipping'' that is likely caused by eclipses of
+clouds of material at the Keplerian co-rotation radius.
 Section~\ref{sec:conclusions} summarizes our main points.
@@
 \end{figure*}
-
-%\begin{figure*}[t!]
-%	\begin{center}
-%		\leavevmode
-%		\includegraphics[width=0.9\textwidth]{f6_comp.png}
-%	\end{center}
-%	\vspace{-0.7cm}
-%	\caption{ {\bf foo.}
-%	bar
-%		\label{fig:corner}
-%	}
-%\end{figure*}
 
@@
 
-Between July 25, 2014 and May 23, 2016, Gaia acquired 121 ``good''
-observations of \ptfo, {\it i.e.}, observations that were not strongly
-downweighted in the astrometric solution of the source (CITE).
-%TODO: give an original explanation of how Gaia works
-The Gaia data processing team at XXXINSTITUTION combined the XXX and
-YYY into a global astrometric solution, from which the parallax of 
-\ptfo\ was derived (CITE).
-The star was assigned a Gaia DR2 identifier of 3222255959210123904
-(CITE).
-In addition, measurements of the stellar proper motion between
-observation epochs were made.
-%FIXME sentence is bad
+Between July 25, 2014 and May 23, 2016, Gaia measured about 300
+billion centroid positions of 1{.}6 billion stars
+\citep{gaia_collaboration_gaia_2016,lindegren_gaiasoln_2018,gaia_collaboration_gaia_2018}.
+These CCD observations were used to estimate positions, proper
+motions, and parallax for the brighest 1{.}3 billion stars, including
+\ptfo.  There were 121 ``good'' observations of \ptfo, {\it i.e.},
+observations that were not strongly downweighted in the astrometric
+solution of the source.  The star was assigned a Gaia DR2 identifier
+of 3222255959210123904.  It photometric brightness was measured using
+selected bands ($G$, $Rp$, and $Bp$) of the Gaia Radial Velocity
+Spectrometer \citep{cropper_gaia_2018,evans_gaia_2018}.
+
+%FIXME FROM HERE
 

### 2020-04-20 billy/plotting.py
add plot_astrometric_excess

@@ -15,6 +15,7 @@ Plots:
     plot_scene
 
     plot_hr
+    plot_astrometric_excess
 
     plot_O_minus_C
 
@@ -814,8 +815,8 @@ def plot_hr(outdir):
     )
 
     ax.legend(loc='best', handletextpad=0.1, fontsize='small')
-    ax.set_ylabel('$G + 5\log_{10}(\omega_{\mathrm{as}}) + 5$', fontsize='large')
-    ax.set_xlabel('$Bp-Rp$', fontsize='large')
+    ax.set_ylabel('G + $5\log_{10}(\omega_{\mathrm{as}}) + 5$', fontsize='large')
+    ax.set_xlabel('Bp - Rp', fontsize='large')
 
     ylim = ax.get_ylim()
     ax.set_ylim((max(ylim),min(ylim)))
@@ -834,6 +835,37 @@ def plot_hr(outdir):
     outpath = os.path.join(outdir, 'hr.png')
     savefig(f, outpath)
 
+
+def plot_astrometric_excess(outdir):
+
+    pklpath = '/Users/luke/Dropbox/proj/billy/results/cluster_membership/nbhd_info_3222255959210123904.pkl'
+    d = pickle.load(open(pklpath, 'rb'))
+    g_df = d[2]
+    chisq_red = g_df.astrometric_chi2_al / (g_df.astrometric_n_obs_al - 5)
+
+    f,ax = plt.subplots(figsize=(4,3))
+
+    ax.scatter(g_df.phot_rp_mean_mag, chisq_red, c='k', alpha=1., zorder=3, s=9,
+               rasterized=True, linewidths=0, label='K+18 members')
+
+    sel = (g_df.source_id.astype(str) == '3222255959210123904')
+
+    ax.plot(
+        g_df[sel].phot_rp_mean_mag,
+        chisq_red[sel],
+        alpha=1, mew=0.5, zorder=8, label='PTFO 8-8695', markerfacecolor='yellow',
+        markersize=12, marker='*', color='black', lw=0
+    )
+
+    ax.set_xlabel('Rp', fontsize='large')
+    ax.set_ylabel('Astrometric Reduced $\chi^2$', fontsize='large')
+    ax.set_xlim([7.5,16.5])
+
+    ax.legend(loc='best', fontsize='small')
+
+    format_ax(ax)
+    outpath = os.path.join(outdir, 'astrometric_excess.png')
+    savefig(f, outpath)
 
 
 def plot_O_minus_C(

### 2020-04-20 drivers/plot_astrometric_excess.py

@@ -1,0 +1,8 @@
+import os
+import billy.plotting as bp
+from billy import __path__
+
+RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
+PLOTDIR = os.path.join(RESULTSDIR, 'cluster_membership')
+
+bp.plot_astrometric_excess(PLOTDIR)

### 2020-04-20 paper/ms.tex
add words and figures on astrometric binarity

@@
 \usepackage{natbib}
+\usepackage[caption=false]{subfig} % for subfloat
 
@@
 \shorttitle{Seeing Double: PTFO$\,$8-8695}
+
+%\objectname{PTFO 8-8695}
+%\objectname{PTFO 8-8695b}
+%\objectname{CVSO 30}
 
@@
   old Orion-OB1a cluster. We inspected data from TESS and Gaia to
-  clarify whether it is actually a planet.  The Gaia data show that
+  clarify whether it is truly a planet.  The Gaia data show that
   PTFO$\,$8-8695 is a photometric binary with respect to members of
@@
   believe that PTFO$\,$8-8695 is a binary M dwarf in which one star
-  shows the long rotation signal, and the other star is showing
+  shows the long rotation signal, and the other star shows
   ``transient dipping'' that has been seen in a few other young
@@
 \end{abstract}
-
-\object{PTFO 8-8695}
-\object{PTFO 8-8695b}
-\object{CVSO 30}
 
@@
 sometimes change immediately after stellar flares.  The explanation
-proposed by {\it e.g.}, \citet{stauffer_orbiting_2017} and
+proposed by \citet{stauffer_orbiting_2017} and
 \citet{david_transient_2017} is that a circumstellar cloud of dust or
@@
 explanation to be viable for \ptfo, a clear determination of the
-stellar rotation period is necessary.  To date, this has not been
+stellar rotation period is necessary.  To date this has not been
 possible
@@
 We begin in Section~\ref{sec:observations} by describing newly
-available observations from the Transiting Exoplanet Survey Satellite
-(TESS; \citealt{ricker_transiting_2015}) and Gaia
+available observations from TESS
+\citep{ricker_transiting_2015} and Gaia
 \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows two
@@
 Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
-hours, and is probably stellar rotation.  A short-period ``dip +
-complex modulation'' repeats every 10.7 hours.  Analyzing the Gaia
+hours, and is probably stellar rotation.  A short-period dip with an additional
+complex modulation repeats every 10.7 hours.  Analyzing the Gaia
 data in Section~\ref{sec:gaia}, we show that relative to its kinematic
@@
 the dip seems to have instantaneously changed over the past decade.
-In addition, a number of other young stars show similar lightcurve
-morphologies.  We therefore argue that PTFO$\,$8-8695 is a binary M
-dwarf in which one star shows a rotation signal, and the other is
-showing ``transient dipping'' that is likely caused by eclipses of
-clouds of material at the Keplerian co-rotation radius.
+In addition, a number of other young stars show lightcurve
+morphologies similar to the short-period signal.  We therefore argue that PTFO$\,$8-8695 is a binary M
+dwarf in which one star shows a rotation signal, and the other 
+shows ``transient dipping'' caused by eclipses of
+material at the Keplerian co-rotation radius.
 Section~\ref{sec:conclusions} summarizes our main points.
@@
 
-The 2-minute stacks for \ptfo\ were then reduced to lightcurves by the
+The 2-minute stacks for \ptfo\ were reduced to lightcurves by the
 Science Processing Operations Center (SPOC) at NASA
@@
 Difference Imaging Photometric Survey (CDIPS;
-\citealt{bouma_cluster_2019}).  The CDIPS lightcurve used a circular
+\citealt{bouma_cluster_2019}).  Our CDIPS lightcurve  of choice used a circular
 aperture with radius 1 pixel.
@@
 the orbit.  The PDC lightcurve initially had 15{,}678 points.  The
-quality cut removed 854 points, masking the orbit edges removed an
-additional 716, and removing the final few days of orbit 20 removed an
+quality-flag cut removed 854 points; masking the orbit edges removed an
+additional 716; removing the final few days of orbit 20 removed an
 additional 1079.  After cleaning, 83\% of the initial flux
@@
 
-We normalized these points by dividing out the median flux. We then
-subtracted by unity to simplify subsequent analysis.  Many of these
-and subsequent processing steps were performed using
+We normalized these points by dividing out the median flux. We opted
+to then subtract by unity to simplify subsequent interpretation.  Many
+of these and subsequent processing steps were performed using
 \texttt{astrobase}~\citep{bhatti_astrobase_2018}.
@@
 
+\subsubsection{Astrometry of \ptfo}
+
 Between July 25, 2014 and May 23, 2016, Gaia measured about 300
@@
 \citep{gaia_collaboration_gaia_2016,lindegren_gaiasoln_2018,gaia_collaboration_gaia_2018}.
-These CCD observations were used to estimate positions, proper
-motions, and parallax for the brighest 1{.}3 billion stars, including
-\ptfo.  There were 121 ``good'' observations of \ptfo, {\it i.e.},
-observations that were not strongly downweighted in the astrometric
-solution of the source.  The star was assigned a Gaia DR2 identifier
+In the Gaia second data release (DR2), these CCD observations were
+then used to estimate positions, proper motions, and parallax for the
+brighest 1{.}3 billion stars, including \ptfo
+\citep{lindegren_gaiasoln_2018}.  There were 121 ``good'' observations
+of \ptfo, that is observations that were not strongly down-weighted in
+its astrometric solution.  \ptfo\ was assigned a Gaia DR2 identifier
 of 3222255959210123904.  It photometric brightness was measured using
 selected bands ($G$, $Rp$, and $Bp$) of the Gaia Radial Velocity
-Spectrometer \citep{cropper_gaia_2018,evans_gaia_2018}.
+Spectrometer \citep{cropper_gaia_2018,evans_gaia_2018}.  We accessed
+the pipeline parameters for \ptfo\ using the Gaia
+archive\footnote{\url{gea.esac.esa.int/archive/}}.
 
-%FIXME FROM HERE
+The majority of Gaia's derived parameters for \ptfo\ agreed with
+expectation from former studies
+\citep{briceno_cida_2005,van_eyken_ptf_2012}.  One novelty however was
+that Gaia DR2 detected a significant ``astrometric excess'', at a
+level of 10.3$\sigma$.  
+We comment on the significance and interpretation of this
+excess in Section~\ref{sec:gaia}.
 
-One point of note:
-Gaia DR2 measured a significant ``astrometric excess'', at a level of
-10.3$\sigma$.
-This astrometric excess indicates the degree
-to which a single-source model fails to explain the observed astrometric
-measurements (CITE).
+\subsubsection{Cluster Membership}
 
-\subsection{Cluster Membership}
 The Orion molecular cloud complex has numerous subgroups, with ages
@@
 
-\section{Gaia Analysis \& Photometric Binarity}
+\section{Gaia Analysis}
 \label{sec:gaia}
@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f5.pdf}
+		%\includegraphics[width=0.8\textwidth]{f5a.pdf}
+    \subfloat{
+        \includegraphics[width=0.7\textwidth]{f5a}
+    }
+
+	\vspace{-0.7cm}
+    \subfloat{
+        \includegraphics[width=0.7\textwidth]{f5b}
+    }
 	\end{center}
 	\vspace{-0.7cm}
-  \caption{ {\bf Hertzsprung-Russell diagram of \ptfo\ and late-type
+  \caption{ {\bf Evidence for binarity in \ptfo}.
+  {\it Top: Hertzsprung-Russell diagram of \ptfo\ and late-type
   members of 25$\,$Ori-1.}
@@
   \citet{kounkel_apogee2_2018} through clustering on six-dimensional
-  Gaia-DR2 and APOGEE-2 data.  The reference ``neighborhood'' (gray
-  circles) is the group of at most $10^4$ randomly selected non-member
+  Gaia-DR2 and APOGEE-2 data.  The ``neighborhood'' (gray
+  circles) is non-member
   stars within 5 standard deviations of the mean 25$\,$Ori-1 right
   ascension, declination, and parallax.  It contains members of the
-  full Orion complex with a wider spread of ages, in addition to field
-  interlopers.  $G$ is the Gaia broadband, $Bp$ is Gaia blue, $Rp$ is
-  Gaia red, and $\omega_{\rm as}$ is the parallax in arcseconds.  The
+  Orion complex with its full spread of ages, in addition to field
+  interlopers.  $G$ denotes Gaia broadband magnitudes, $Bp$ Gaia blue, $Rp$
+  Gaia red, and $\omega_{\rm as}$ the parallax in arcseconds.  The
   $x$-axis limits have been set to show only K and M dwarf members, to
   accentuate \ptfo's separation from the single-star sequence.
+  {\it Bottom: Astrometric goodness-of-fit versus $Rp$ magnitude
+  for 25$\,$Ori-1 members}.
+  The single-source astrometric model provides a poor fit, which
+  could be due to stellar variability or binarity.
 	\label{fig:gaia}
@@
 
+\subsection{Photometric Binarity}
 To assess the cluster membership and potential binarity of \ptfo, we
@@
 star system.
+
+\subsection{Astrometric Binarity}
+
+As noted in Section~\ref{sec:observations}, the Gaia DR2 solution for
+\ptfo\ shows a 10.3$\sigma$ astrometric excess.  This astrometric
+excess indicates the degree to which a single-source model fails to
+explain the observed astrometric measurements.  Specifically, the
+single-source astrometric model for yielded $\chi^2=325.2$.  There
+are 121 astrometric measurements, and 5 free parameters, and therefore
+116 degrees of freedom. The reduced $\chi^2$ is 2.80.  The
+majority of stars with comparable brightness in Gaia do not show such
+poor goodness-of-fit \citep[][Appendix A]{lindegren_gaiasoln_2018}.
+
+Potential explanations for the poor astrometric fit include photometric
+variability and unresolved stellar binarity \citep[{\it
+e.g.},][]{rizzuto_ZEIT8_2018,belokurov_unresolved_2020}.
+If photometric variability were the cause, we would expect comparably
+faint stars in the same kinematic group of Orion to
+show similar astrometric excesses, as the majority of
+young stars are highly variable.
+
+Using the same 149 members in the 25$\,$Ori-1 subgroup from
+\citet{kounkel_apogee2_2018}, we calculated the reduced $chi^2$ for
+each member.  The lower panel of Figure~\ref{fig:gaia} shows the
+resulting values, as a function of stellar brightness.  \ptfo\ is in
+the upper 90$^{\rm th}$ percentile of stars showing astrometric
+excesses within the 25$\,$Ori-1 group.  Although we will have to wait
+for the full release of the nominal Gaia mission to definitively
+determine whether the astrometric excess is caused by stellar binarity
+or photometric variability, this result suggests that stellar binarity
+is the root cause, because other member stars have identical variability
+characteristics but do not show such a large astrometric excess.
+%FIXME: you might want to color the points by their TESS amplitude, or
+%similar.
+% or even better... just do a CUT on TESS amplitude. Require it to be
+% like 10% or above -- so LARGER amplitudes than PTFO! That's the best
+% test.
 

### 2020-04-20 paper/collect.sh

@@ -9,7 +9,8 @@ cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.pdf f3.pdf
 
 cp ../results/PTFO_8-8695_results/scene.pdf f4.pdf
 
-cp ../results/cluster_membership/hr.pdf f5.pdf
+cp ../results/cluster_membership/hr.pdf f5a.pdf
+cp ../results/cluster_membership/astrometric_excess.pdf f5b.pdf
 
 cp ../results/ephemeris/O_minus_C.pdf f6.pdf
 

### 2020-04-21 paper/ms.tex
update paper

@@
 Difference Imaging Photometric Survey (CDIPS;
-\citealt{bouma_cluster_2019}).  Our CDIPS lightcurve  of choice used a circular
-aperture with radius 1 pixel.
+\citealt{bouma_cluster_2019}).  Our CDIPS lightcurve  of choice used a
+circular aperture with radius 1 pixel.
 
@@
 In the Gaia second data release (DR2), these CCD observations were
-then used to estimate positions, proper motions, and parallax for the
-brighest 1{.}3 billion stars, including \ptfo
+used to estimate positions, proper motions, and parallax for the
+brighest 1{.}3 billion stars, including \ptfo\ 
 \citep{lindegren_gaiasoln_2018}.  There were 121 ``good'' observations
@@
 that Gaia DR2 detected a significant ``astrometric excess'', at a
-level of 10.3$\sigma$.  
-We comment on the significance and interpretation of this
-excess in Section~\ref{sec:gaia}.
-
-\subsubsection{Cluster Membership}
-
-The Orion molecular cloud complex has numerous subgroups, with ages
-spanning 1 to 15$\,$Myr (CITE).
-\ptfo\ has been known to be a member of the Orion OB1a association
-since at least CITE (Briceno 2005).
-Its relation to the broader Orion complex has been further explored by
-CITEX, CITEY, and CITEZ (Briceno 05, 08, 18, whomever Van Eyken cites,
-and Kounkel 18).
-
-In describing the cluster membership of \ptfo, we follow the notation
-and results of \citet{kounkel_apogee2_2018}.
-\citet{kounkel_apogee2_2018} combined astrometric data from Gaia DR2
-with spectroscopic data from APOGEE-2 (CITE).  They then performed a
-hierarchical clustering on the six dimensional position and velocity
-information to identify subgroups within the Orion complex.  From
-smallest to largest groups, \ptfo\ was identified as being a member of
-the following nested subgroups:
+level of 10.3$\sigma$.  We comment on the significance and
+interpretation of this excess in Section~\ref{sec:gaia}.
+
+\subsubsection{Hierarchical Cluster Membership}
+\label{subsec:hierarchical}
+
+Gaia also provided astrometric parameters for tens of thousands of
+young stars in the Orion complex.  The stellar populations in giant
+molecular cloud complexes are not monolithic; substructured groups are
+the norm \citep{briceno_lowmassOB_2007}.  The Orion molecular cloud
+complex in particular has numerous subgroups, with ages spanning 0.5
+to 15$\,$Myr. For an incomplete sampling, see for instance
+\citet{briceno_cida_2005,jeffries_kinematic_2006,briceno_25_2007,kounkel_apogee2_2018}
+and \citet{briceno_cidaII_2019}.
+
+\ptfo\ was initially identified as a member of the Orion$\,$OB1a
+sub-association by \citet{briceno_cida_2005} through combined
+photometry and spectroscopy.  Later work by \citet{briceno_25_2007}
+clarified that \ptfo\ was in a kinematically distinct subgroup of
+Orion~OB1a, named the ``25$\,$Ori'' group after its brightest
+member. The 25$\,$Ori group had an isochrone age of 7--10$\,$Myr, and
+had a smaller disk fraction than younger nearby sub-associations
+\citep{hernandez_spitzer_2007}.
+
+With the Gaia astrometry, it has become clear that 25$\,$Ori itself
+has further subgroups
+\citep{kounkel_apogee2_2018,briceno_cidaII_2019}.  In describing the
+cluster membership of \ptfo, we follow the notation and results of
+\citet{kounkel_apogee2_2018}.  These authors combined astrometric data
+from Gaia DR2 with near-infrared spectra from APOGEE-2
+\citep{gunn_sdss_2006,majewski_apache_2017,blanton_sloan_2017,zasowski_target_2017,cottle_apogee2_2018}.
+They performed a hierarchical clustering on the six dimensional
+position and velocity information to identify subgroups within the
+Orion complex.  From smallest to largest, \ptfo\ was identified
+as being a member of the following hierarchical subgroups:
 \begin{equation}
@@
 where from set-notation, `$\subset$' denotes ``is a proper subset of''.
-
-While all members of the Orion complex are young relative to the
-field, the internal age dispersion between different subgroups is
-measurable.  The Orion Nebula Cluster (M42) is a site of ongoing
-star-formation, and its stars are 1-3$\,$Myr old (CITE).
-25$\,$Ori\-1, by contrast, is XX-XX$\,$Myr old (CITE).  These details
-are essential when assessing any evidence for photometric binarity in
-\ptfo, because there is a degeneracy between stellar luminosity and
-age for stars on the pre-main-sequence.  Having a clean sample of
-tightly spatially and kinematically associated stars is essential to
-minimize contamination not just from the field, but from older and
+25$\,$Ori-1 is the largest subgroup of 25$\,$Ori, with 149 identified
+members.
+Its mean age from its CMD was found to be 6.9$\,$Myr, and from its HR
+diagram 8.5$\,$Myr \citep{kounkel_apogee2_2018}.
+\citet{kounkel_apogee2_2018} identified seven other smaller groups in
+the Orion complex near the Be star 25$\,$Ori. These groups received
+higher numbers, {\it e.g.}, 25$\,$Ori-2 (${\rm Age}_{\rm CMD} =15.1\,$Myr; ${\rm Age}_{\rm CMD} =12.9\,$Myr; see also \citealt{briceno_cidaII_2019}).
+
+These details concerning the group membership for one object may seem
+cumbersome to those accustomed to comparing ``young cluster
+members'' with ``old field stars''.  Though all members of the Orion
+complex are indeed young relative to the field, these details are
+essential for assessing any evidence for photometric binarity in
+\ptfo, because of the degeneracy between stellar luminosity and
+age for pre-main-sequence stars.  Having a clean sample of tightly
+spatially and kinematically associated reference stars
+minimizes contamination not just from the field, but from older and
 younger members of the Orion complex itself.
@@
 
-Our initial inspection of the lightcurve, in both its 2-minute PDCSAP
+Our initial inspection of the TESS lightcurve, in both its 2-minute PDCSAP
 and 30-minute FFI forms, showed a strong sinusoidal beat signal
-(Figure~\ref{fig:splitsignal}, top panel).
+(Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}, top panel).
 
@@
 0.499\,{\rm days}$.  The $P_{\rm \ell}$ peak had the greater power of
-the two.  Smaller harmonics from each of these two dominants peaks
+the two.  Smaller harmonics surrounding each of these two dominants peaks
 were also present.
@@
 
-The peak-to-peak amplitude at maximum, when the two signals
+The peak-to-peak lightcurve amplitude at maximum, when the two signals
 constructively interfere, is about 14\%.  At minimum, the peak-to-peak
@@
 
-We opted to model the lightcurve as a linear combination of Fourier
+We fitted the lightcurve as a linear combination of Fourier
 harmonics at the short and long periods, plus a transit at the short
@@
 $N$ and $M$, and used the Bayesian information criterion to choose the
-appropriate model (Table~\ref{tab:modelcompare}).
+appropriate model (Table~1).
 
@@
 harmonics of sines and cosines at the short period, plus $M=1$
-harmonics at the long period.  In this case, the free parameters would
+harmonics at the long period.  For this case the free parameters would
 be as follows.  For the transit, we would fit for the impact
@@
 \citep{exoplanet:exoplanet}.  Our priors are listed in
-Table~\ref{tab:posterior}.  To speed up the fitting, we binned the
-cleaned 2 minute lightcurves to 10 minute bins.  We correspondingly
+Table~2.  To speed up the fitting, we re-sampled the
+cleaned 2 minute lightcurves to 10 minute binning.  We correspondingly
 scaled the uncertainties in the flux measurements by a factor of
@@
 We tested our ability to successfully recover injected parameters
-using synthetic data, before switching to the actual \ptfo\
+using synthetic data, before fitting the \ptfo\
 lightcurves.
@@
 we used the Bayesian information criterion
-(Table~\ref{tab:modelcompare}).  The model with the lowest BIC had
+(Table~1).  The model with the lowest BIC had
 two harmonics at the short 10.74$\,$hr period, and two harmonics
-at the long 11.96$\,$hr period.  All of the models have reduced
+at the long 11.96$\,$hr period.  
+The next-best model had an additional harmonic at the longer period ($M=3$),
+but was otherwise identical.
+All nine models have reduced
 $\chi^2$ ranging between 1.37 and 1.51, which suggests a plausible
-though imperfect agreement between the data and models.
+though imperfect agreement between the data and fitting results.
+The best-fit parameters for the lowest BIC model are given in Table~2.
 
@@
 
-In every model, the variability at the long period is a simple
-sinusoid with peak-to-peak amplitude $\approx$10\%.  The variability
-at the short period is always more complex.  A dip of depth
+In every model, the $11.96\,{\rm hr}$ variability is a simple
+sinusoid with peak-to-peak amplitude $\approx$10\%.  The $10.74\,{\rm hr}$ variability
+is always more complex.  A dip of depth
 $\approx$1.2\%, fit in our model as a transit, lasts $\approx$0.75
-hours.  Superposed on the dip is a complex signal with peak-to-peak
-amplitude of about 4\%, which peaks near phase 0.25, and reaches
+hours.  Superposed on the dip is a slightly asymmetric sinusoid with peak-to-peak
+amplitude of about 4\%. The asymmetric sinusoid peaks near phase 0.25, and reaches
 minimum brightness between phases -0.5 and -0.25.
-
-Outside of the primary dip, the short-period signal is relatively
-smooth, at least from phases 0 to 0.5.  However the short-period
-signal is asymmetric.  The flux from phases -0.5 to 0 shows what could
-be a discontinuous jump, shortly after reaching minimum.  This jump
+The flux at phase $\approx -0.33$ shows what could
+be a discontinuous jump, shortly after reaching minimum (Figure~\ref{fig:phasefold}).  This jump
 was visible in each of the nine models we considered.
@@
 
-\subsection{Blend considerations}
+\section{Binarity Analysis}
+\label{sec:gaia}
+
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\includegraphics[width=0.9\textwidth]{f6.png}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+% 	\caption{ {\bf Positions, kinematics, and HR diagram of \ptfo.}
+%   Members of the 25$\,$Ori-1 subgroup are shown with black circles,
+%   and were identified by \citet{kounkel_apogee2_2018} through
+%   clustering on Gaia DR2 and APOGEE data.  The ``neighborhood'' (gray
+%   circles) is defined as the group of at most $10^4$ randomly selected
+%   non-member stars within 5 standard deviations of the mean right
+%   ascension, declination, and parallax.
+% 	\label{fig:gaia}
+% 	}
+% \end{figure*}
+
+\subsection{Visual Binarity}
 \label{subsec:blend}
@@
 
-The TESS pixels are $\approx21$'' per side, and so before making an
-interptations, we need to consider whether light from neighboring
+The TESS pixels are $\approx21$'' per side. Before making any
+interpretations, we needed to consider whether light from known neighboring
 stars could have affected the photometry.  The scene is shown in
-Figure~\ref{fig:scene}.  The pixels used to measure the background
-level in the SPOC lightcuirve are indicated with an `\texttt{X}'
+Figure~\ref{fig:scene}.  
+In the upper panels, the pixels used to measure the background
+level in the SPOC lightcurve are indicated with an `\texttt{X}'
 hatch, and the pixels used in the final lightcurve aperture are shown
@@
 contaminate the target star is TIC 264461980, with $T=14.8$, which we
-denote ``Star A''.  Star A is 23.6'' NW of our target, and based on
+dub ``Star A''.  Star A is 23.6'' NW of our target, and based on
 the magnitude difference could contribute up to 48\% the flux of our
@@
 
-Because \ptfo\ has previously been identified to have periodicity
+Because \ptfo\ was previously identified to have periodicity
 consistent with our measurement of $P_{\rm s}$, our main concern
-regarding blending is the degree to which we can be certain that the
-long-period signal at $P_{\rm \ell}$ also originates from \ptfo.
-We took two approaches towards determining the source of the
+regarding blending was the degree to which we could be certain that the
+long-period signal at $P_{\rm \ell}$ also originated from \ptfo.
+We took two approaches to verifying the source of the
 long-period signal.
 
-First, we examined the CDIPS full frame image lightcurves of the
-target, which are available on MAST \citep{bouma_cluster_2019}.  The
-maximal peak-to-peak beat amplitude is consistently $\approx$10\%
-across apertures of radii 1, 1.5, and 2.25 pixels.  If Star A were the
+First, we examined the CDIPS FFI lightcurves of the
+target, which were available on MAST \citep{bouma_cluster_2019}.  The
+maximal peak-to-peak beat amplitude was constant
+across apertures of radii 1, 1.5, and 2.25 pixels to visual precision ($\lesssim 1\%$).
+If Star A were the
 source of the long-period variability, we would expect the peak
@@
 
-Second, we examined the lightcurve of each pixel in the scene
+Second, we examined the 2-minute lightcurve of each pixel in the scene
 individually.  We opted to use the interactive tools implemented in
@@
 nearest to Star A to show a sinusoidal signal with amplitude exceeding
-$10\%$.  We find no evidence for this being the case.  The pixel
+$10\%$.  The data do not support this possibility.  The pixel
 directly below Star A does not clearly show the sinusoidal
@@
 optimal aperture) shows the $P_{\rm \ell}$ sinusoidal variability
-signal at $\approx 10\%$ amplitude.
-
-As there is no evidence in favor of a blend scenario, we conclude that
-both the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from \ptfo,
-at least within the resolution of the Gaia-DR2 source catalog.
-However, as we shall see, \ptfo itself could still be a binary.
-
-\section{Gaia Analysis}
-\label{sec:gaia}
-
-% \begin{figure*}[t]
-% 	\begin{center}
-% 		\leavevmode
-% 		\includegraphics[width=0.9\textwidth]{f6.png}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-% 	\caption{ {\bf Positions, kinematics, and HR diagram of \ptfo.}
-%   Members of the 25$\,$Ori-1 subgroup are shown with black circles,
-%   and were identified by \citet{kounkel_apogee2_2018} through
-%   clustering on Gaia-DR2 and APOGEE data.  The ``neighborhood'' (gray
-%   circles) is defined as the group of at most $10^4$ randomly selected
-%   non-member stars within 5 standard deviations of the mean right
-%   ascension, declination, and parallax.
-% 	\label{fig:gaia}
-% 	}
-% \end{figure*}
+signal at $\approx 14\%$ amplitude.
+We conclude that
+within the resolution of the Gaia DR2 source catalog,
+the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from \ptfo.
+From \citet{ziegler_measuring_2018}, we can surmise that
+stellar companions
+wider than $\approx$1'' (349~AU) and within
+$\Delta G \approx 3$ magnitudes of \ptfo\ would have likely been
+detected through this approach.
+
+\citet{van_eyken_ptf_2012} obtained
+stronger constraints on possible stellar companions using the
+NIRC2 camera on Keck II.
+They reported 3$\sigma$ $H$-band magnitude difference limits
+of 4.3, 6.4, and 8.9 at angular separations of 0.25, 0.5, and 1.0
+arcseconds (87, 175, and 349~AU).
+They also detected a point-source, not present in Gaia DR2,
+7.0 magnitudes fainter than the target, 
+and 1.8$''$ north-east.
+Due to the brightness difference, this coincident star\footnote{This companion was claimed to be a potential planetary-mass
+	object \citep{schmidt_direct_2016}. Subsequent analysis of its colors showed that it is
+	a background star \citep{lee_evidence_2018}.} cannot be the source
+of our signals.
+
+%FIXME: would be better to actually be able to get them...
+While long-baseline radial velocity measurements could provide complimentary
+lower limits on companion masses and separations, the RV
+data for \ptfo\ are rather poor owing to the stellar faintness.
+The longest single-instrument baseline in the literature appears to be 5 Keck/HIRES
+measurements acquired over 10 days in April 2011 by \citet{van_eyken_ptf_2012}.
+The RV RMS over that 10 day span was $1.6\,{\rm km}\,{\rm s}^{-1}$, consistent with
+the measurement precision.
+Though \citet{yu_tests_2015} acquired 22 further Keck/HIRES observations over
+one night in December 2013, these points seem to not have been
+reduced to velocities.
+Further Keck/HIRES measurements of this target could confirm or
+refute the existence of possible binary companions.
+
+\subsection{Photometric Binarity}
+
 \begin{figure*}[t]
@@
 		\leavevmode
-		%\includegraphics[width=0.8\textwidth]{f5a.pdf}
 		\subfloat{
-        \includegraphics[width=0.7\textwidth]{f5a}
+			\includegraphics[width=0.7\textwidth]{f5a.pdf}
 		}
@@
 		\subfloat{
-        \includegraphics[width=0.7\textwidth]{f5b}
+			\includegraphics[width=0.7\textwidth]{f5b.pdf}
 		}
@@
 		\citet{kounkel_apogee2_2018} through clustering on six-dimensional
-  Gaia-DR2 and APOGEE-2 data.  The ``neighborhood'' (gray
+		Gaia DR2 and APOGEE-2 data.  The ``neighborhood'' (gray
 		circles) is non-member
@@
 
-%FIXME: summarize existant AO constraints here too
-\citet{schmidt_direct_2016}
-\citet{lee_evidence_2018}
-% \citet{czesla_xray_2019}
-
-\subsection{Photometric Binarity}
-To assess the cluster membership and potential binarity of \ptfo, we
-needed to identify stars with which it was and was not coeval.  The
-simplest way to do this---clustering based on six-dimensional position
-and kinematic information---had already been done by
-\citet{kounkel_apogee2_2018}.  For simplicity, we considered the
-members they identified brighter than $G_{\rm Rp}$ of 16.  This
-yielded 149 stars in 25$\,$Ori-1, mostly M dwarfs.
-\citet{kounkel_apogee2_2018} identified seven other smaller groups in
-the Orion complex near the Be star 25$\,$Ori. These groups received
-higher numbers, {\it e.g.}, 25$\,$Ori-2.
+Aside from visual binarity, we can also check the Gaia data
+for photometric binarity.
+To assemble a set of stars coeval with \ptfo,
+we used the 25$\,$Ori-1 members identified by
+\citet{kounkel_apogee2_2018}, and discussed in Section~\ref{subsec:hierarchical}.
 
@@
 standard deviations of the mean 25$\,$Ori-1 right ascension,
-declination, and parallax.  We queried these stars using the
-\texttt{astroquery} package, which provides a convenient interface to
-the Gaia archive (CITE, CITE).  This yielded 1{,}819 neighbors.  While
+declination, and parallax.  We queried Gaia DR2 for these stars using
+\texttt{astroquery} \citep{astroquery_2018}.  This yielded 1{,}819 neighbors.  While
 some of these stars may indeed be members of the Orion complex, or
@@
 was that 25$\,$Ori-1 was a clearly defined over-density in each
-dimension, so the cluster exists, and is different from the
-neighborhood.  \ptfo\ was also clearly a member in each of these
+dimension---the cluster exists, and is distinct from the
+neighborhood.  \ptfo\ was also within the cluster in each of these
 projected dimensions.
@@
 photometric binary was of great interest.  Figure~\ref{fig:gaia} shows
-the HR diagram from which we assessed this issue.  The diagram shows
-that \ptfo\ is $\approx$0.8 magnitudes brighter than the average
+the HR diagram we constructed to assess this issue.  The diagram shows
+that \ptfo\ is $\approx$0.75 magnitudes brighter than the average
 25$\,$Ori-1 star of the same color.  In other words, it is about twice
-as bright.  It also seems to be on the photometric binary track of the
+as bright.  It also seems to be coincident with the photometric binary track of the
 cluster, which has a few other stars.
@@
 The implication is that either {\it (i)} \ptfo\ is notably younger
-than the kinematically identical cluster members, or {\it (ii)} \ptfo\
-is a photometric binary.  Given the independent presence of two
-resolved signals, we favor the interpretation that \ptfo\ is a binary
-star system.
+than the kinematically identical 25$\,$Ori-1 members, or {\it (ii)} \ptfo\
+is a photometric binary.
+Since there is no a priori reason to suspect an age difference,
+but we have resolved two
+separate photometric signals, we favor the binary interpretation.
 
@@
 
+A separate possible line of evidence for binarity is the Gaia DR2 astrometry.
 As noted in Section~\ref{sec:observations}, the Gaia DR2 solution for
@@
 explain the observed astrometric measurements.  Specifically, the
-single-source astrometric model for yielded $\chi^2=325.2$.  There
+single-source astrometric model yielded $\chi^2=325.2$.  There
 are 121 astrometric measurements, and 5 free parameters, and therefore
@@
 Using the same 149 members in the 25$\,$Ori-1 subgroup from
-\citet{kounkel_apogee2_2018}, we calculated the reduced $chi^2$ for
+\citet{kounkel_apogee2_2018}, we calculated the reduced $\chi^2$ for
 each member.  The lower panel of Figure~\ref{fig:gaia} shows the
@@
 the upper 90$^{\rm th}$ percentile of stars showing astrometric
-excesses within the 25$\,$Ori-1 group.  Although we will have to wait
+excesses within the 25$\,$Ori-1 group.  
+Relative to other M-dwarf group members with comparable
+brightnesses and variability characteristics, \ptfo\ still stands out
+as behaving astrometrically poorly.
+Ultimately, we will have to wait
 for the full release of the nominal Gaia mission to definitively
 determine whether the astrometric excess is caused by stellar binarity
-or photometric variability, this result suggests that stellar binarity
-is the root cause, because other member stars have identical variability
+or photometric variability.
+Nonetheless this result suggests that stellar binarity
+is indeed the root cause, because other member stars have identical variability
 characteristics but do not show such a large astrometric excess.
-%FIXME: you might want to color the points by their TESS amplitude, or
-%similar.
-% or even better... just do a CUT on TESS amplitude. Require it to be
-% like 10% or above -- so LARGER amplitudes than PTFO! That's the best
-% test.
 
@@
 
-The standard interpretation of sinusoidal modulations for a
-pre-main-sequence M dwarf is that we are observing a stellar rotation
-period of 11.96 hours.  This is the dominant signal in the system with
+The standard interpretation for 11.96$\,$hour sinusoidal modulations of a
+pre-main-sequence M dwarf is stellar rotation.  This is the dominant signal in the system with
 10\% amplitude, and there is no evidence to suggest that this signal
@@
 
-The discovery study by \citet{van_eyken_ptf_2012} saw the same signal
-({\it e.g.}, their Figure~7), but identified its alias as a
+The discovery study by \citet{van_eyken_ptf_2012} saw an alias of the same signal
+({\it e.g.}, their Figure~7), and identified it as a
 periodogram peak at $0.9985 \pm 0.0061\,$days. They ascribed it to
@@
 We are not the first to reach the conclusion that the long period
-sinusoidal modulation is astrophysical.  A follow-up study by
+sinusoidal modulation is astrophysical.  A study by
 \citet{koen_multicolour_2015} identified the same modes and aliases as
-\citet{van_eyken_ptf_2012}, but argued that the $0.50\,{\rm d}$ signal
-was astrophysical.  Using archival photometry and photometry from the
+\citet{van_eyken_ptf_2012}, but argued that the signal
+was astrophysical (however they were still unsure of the exact period).
+Using photometry from the
 YETI global telescope network, \citet{raetz_yeti_2016} eventually came
-to the conclusion that the that the $0.50\,{\rm d}$ signal was indeed
+to the conclusion that the $0.50\,{\rm d}$ signal was indeed
 from stellar rotation.  The TESS data strongly support this
@@
 \subsection{Short period dip}
-The dip lasts about 45 minutes, and seems to re-occur every 10.74
+The TESS lightcurve shows a 
+dip that lasts about 45 minutes, and seems to re-occur every 10.74
 hours
@@
 The dip duration is roughly the same as that observed by previous
-investigators \citep{van_eyken_ptf_2012,yu_tests_2015}
-The dip depth is comparable to what has been observed in red visual
-bands...
-%TODO: and how does it compare to IR? bigger or smaller? cite relevant
-%studies.
+investigators \citep{van_eyken_ptf_2012,yu_tests_2015}.
+The 1.2\% depth is similar to what has been observed in the near-infrared
+\citep{onitsuka_multicolor_2017}.
+However the dip depth seems likely to have evolved over time between
+being not present at all, to a maximum of $\approx$5\%
+\citep[{\it e.g.},]{koen_multicolour_2015,yu_tests_2015,tanimoto_evidence_2020}.
+
+One particularly interesting feature of the dip is its epoch.
+Over the past decade, many investigators have observed \ptfo.
+Its dips do not always occur on a 
+perfectly linear ephemeris \citep{yu_tests_2015}.
+In fact, \citet{tanimoto_evidence_2020} recently provided stark
+evidence for different behavior altogether: over a time-span of
+years, the dip ``splits'' into distinct groups at particular repeating phases.
+See for instance their Figures~2 through~4.
+Fitting a decade of observations, they provided the following linear
+ephemeris.
+\begin{align}
+t_0\ {\rm BJD}_{\rm TDB} &= 2455543.943 \pm 0.002 \\
+P &= 0.4483993 \pm 0.0000006 d.
+\end{align}
+
+In Figure~\ref{fig:o_minus_c}, we show the phase of the dip we
+detect in the TESS data, relative to their ephemeris.
+It agrees with the independent December 2018 measurements by \citet{tanimoto_evidence_2020}:
+the dip has drastically shifted phase over the past decade.
+
+Figure~\ref{fig:o_minus_c} shows two additional strange features:
+{\it (i)}  multiple dips per cycle, and {\it (ii)} a set of dips numerically coincident
+with phase $(P_{\rm \ell} - P_{\rm s}) / P_{\rm s}$.
+The observation of multiple
+dips per cycle in 2015 seems to have been detected by both \citet{yu_tests_2015}
+and \citet{tanimoto_evidence_2020}.
+It therefore seems credible.
+Inspecting the \citet{tanimoto_evidence_2020} lightcurves,
+the claim of multiple dips per cycle in December 2018 at phase 0 and -0{.}47
+seems much less plausible---the phase -0{.}47 dips are strongly detected,
+while the suggested phase 0 dip is not clearly present.
+
+We are not sure what to make of the numerical coincidence.
+The ratio of long to short periods is roughly 10:9, which does produce a beat.
+However it is not clear that this would obviously translate into
+an observational bias unless by some fluke three season's worth 
+of observations managed to only observe every ninth dip.
+This is of course not the case, and we therefore leave this 
+curiosity as observation {\it sans} interpretation.
 
@@
 		{\bf Timing residuals for \ptfob\ from a decade of monitoring.}
-    Black points are times of ``dips'', minus the indicated linear
+    Black points are times of dips, minus the indicated linear
     ephemeris.  The $y$-axis is given in units of phase for the
     short-period signal.  The star shows the binned TESS ephemeris.
-    ``Dips'' have been observed by \citet{van_eyken_ptf_2012},
+    Dips have been observed by \citet{van_eyken_ptf_2012},
     \citet{ciardi_followup_2015}, \citet{yu_tests_2015},
@@
     convincingly {\it observed}.  Horizontal dashed lines are drawn at
-    $\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting a
-    possible observational bias.  The orbital phase observed by TESS
+    $\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting either a numerical
+    coincidence or an observational bias.  The orbital phase observed by TESS
     is consistent with that of \citet{tanimoto_evidence_2020}, and
@@
 \end{figure}
-
-\paragraph{Epoch}
-The TESS dip does not phase up where it is supposed to...
-Figure~\ref{fig:o_minus_c}
 
@@
 parametrized as the ``$B_1$'' component.  The amplitude of the signal
-is $0.53 \pm 0.06\%$.  Interpreted as being caused by a planet, it
+is $0.53 \pm 0.06\%$ (see Table~2).  Interpreted as being caused by a planet, it
 would imply a minimum planet mass $M_{\rm p} \sin i$ of $3.8\,M_{\rm
 Jup}$.
+
+%FIXME but argue against
 

### 2020-04-21 paper/ms.tex
fix TeX column width

@@
   its kinematic group.  The TESS lightcurve shows that the dominant
-  variability in this system is a sinusoid with a ``long'' period
+  variability in this system is a sinusoid with a long period
   $P_{\rm \ell}=11.96\,$hr, presumably caused by stellar rotation.
   Also present is a complex signal, previously identified as the
-  planet candidate, that repeats with a ``short'' period $P_{\rm s}=
+  planet candidate, that repeats with a short period $P_{\rm s}=
   10.74\,$hr.  The two signals beat every 4.48 days.  Although there
   is a dip in the short-period signal, ground-based photometry from
-  the past decade shows that the orbital phase of the dip seems to
+  the past decade indicates that the orbital phase of the dip seems to
   have instantaneously jumped, at least once, and perhaps twice.
-  Planets do not ``jump'' in orbital phase.  Given the evidence, we
-  believe that PTFO$\,$8-8695 is a binary M dwarf in which one star
-  shows the long rotation signal, and the other star shows
-  ``transient dipping'' that has been seen in a few other young
+  Planets do not ``jump'' in orbital phase.  Given the available
+  evidence, we believe that PTFO$\,$8-8695 is a binary M dwarf in
+  which one star shows the long rotation signal, and the other star
+  shows ``transient dipping'' that has been seen in a few other young
   weak-lined T Tauri stars.  The origin of these transient dips is
-  still undetermined, but our preferred explanation is eclipses of
+  still undetermined, but our preferred explanation is eclipsing
   clouds of gas or dust at the Keplerian co-rotation radius.
@@
 \section{Introduction}
-If \ptfob\ were a planet, it would be exceptional.  Transiting a
-sub-10$\,$Myr old weak-lined T Tauri M dwarf in Orion, it would be the
-youngest hot Jupiter known \citep{van_eyken_ptf_2012}.  Its orbital
-period of only 10.7 hours would also give it the shortest period of
-any known hot Jupiter.  With such a short period, it would almost
-certainly have filled its Roche lobe, and would be actively losing
-mass to its host star.  Not only that, but the rapidly rotating host
-star would also be oblate and gravity darkened, and so the planet's
-orbit would likely precess into and out of transitability
+If \ptfob\ were a planet, it would be exceptional.  At below 10
+million years of age, it could be the youngest hot Jupiter known
+\citep{van_eyken_ptf_2012}.  Its orbital period of only 10.7 hours
+around a weak-lined T Tauri M dwarf would also give it the shortest
+period of any known hot Jupiter.  With such a short period, it would
+almost certainly have filled its Roche lobe, and would be actively
+losing mass to its host star.  Not only that, but the rapidly rotating
+host star would also be oblate and gravity darkened, and so the
+planet's orbit would likely precess into and out of transitability
 \citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.
@@
 stellar chromosphere \citep{johnskrull_h_2016}.  The proposed
-explanation is that outflowing mass from the planet may explain this
+explanation is that outflowing mass from the planet could explain this
 excess emission as well \citep{johnskrull_h_2016}.
@@
 the planet does not seem to emit infrared radiation in occulation, at
-least anywhere near the expected amplitude \citep{yu_tests_2015}.  In
+least at the expected amplitude \citep{yu_tests_2015}.  In
 addition, despite measurement attempts by multiple investigators, it
@@
 gravity-darkened star precessed about its rotation axis, it would need
-to show photometric variability that has not been observed.  While the
-planetary interpetation clearly faces challenges, alternative
-explanations do as well.  High-latitude accretion hotspots might
-produce the observed H$\alpha$ variability, and require fine-tuning
-produce dips of the approach duration. Furthermore, \ptfo\ does not
-have an infrared (IR) excess associated with the presence of an inner
-disk \citep[{\it e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude
-starspots, hot or cold, struggle to produce the necessary dip
-durations.
-
-Alternatively, between 0.1\% and 1\% of rapidly rotating low-mass
-stars in $\mathcal{O}$(10)$\,$Myr old associations show narrow dips in
-phase with strong stellar rotation signals \citep{rebull_usco_2018}.
-The dips can persist over months, but their depths often vary, and
-sometimes change immediately after stellar flares.  The explanation
-proposed by \citet{stauffer_orbiting_2017} and
-\citet{david_transient_2017} is that a circumstellar cloud of dust or
-gas might be orbiting near the Keplerian co-rotation radius.  For this
-explanation to be viable for \ptfo, a clear determination of the
-stellar rotation period is necessary.  To date this has not been
-possible
+to show photometric variability that has not been observed.
+
+While the planetary interpetation clearly faces challenges,
+alternative explanations do as well.  High-latitude accretion hotspots
+might produce the observed H$\alpha$ variability, and require
+fine-tuning produce dips of the approach duration. Furthermore, \ptfo\
+does not have an infrared (IR) excess associated with the presence of
+an inner disk \citep[{\it e.g.},][Figure~18]{yu_tests_2015}.
+Low-latitude starspots, hot or cold, struggle to produce the necessary
+dip durations.
+
+Another potential explanation is that between 0.1\% and 1\% of rapidly
+rotating low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations
+show narrow dips in phase with strong stellar rotation signals
+\citep{rebull_usco_2018}.  The dips can persist over months, but their
+depths often vary, and sometimes change immediately after stellar
+flares.  The explanation proposed by \citet{stauffer_orbiting_2017}
+and \citet{david_transient_2017} to explain this novel class of
+variable stars is that a circumstellar cloud of dust or gas might be
+orbiting near the Keplerian co-rotation radius.  Given available
+ground-based data, this explanation is not clearly viable for \ptfo,
+because a clear determination of the stellar rotation period has not
+yet been possible
 \citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.
@@
 We begin in Section~\ref{sec:observations} by describing newly
-available observations from TESS
-\citep{ricker_transiting_2015} and Gaia
-\citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows two
-distinct signals, which we extract and analyze in
+available observations from TESS \citep{ricker_transiting_2015} and
+Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
+two distinct signals, which we extract and analyze in
 Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
-hours, and is probably stellar rotation.  A short-period dip with an additional
-complex modulation repeats every 10.7 hours.  Analyzing the Gaia
-data in Section~\ref{sec:gaia}, we show that relative to its kinematic
-group, \ptfo\ is a photometric binary.  We collect and discuss the
-puzzle pieces in Section~\ref{sec:discussion}.  The orbital phase of
-the dip seems to have instantaneously changed over the past decade.
-In addition, a number of other young stars show lightcurve
-morphologies similar to the short-period signal.  We therefore argue that PTFO$\,$8-8695 is a binary M
-dwarf in which one star shows a rotation signal, and the other 
-shows ``transient dipping'' caused by eclipses of
-material at the Keplerian co-rotation radius.
-Section~\ref{sec:conclusions} summarizes our main points.
+hours, and is probably stellar rotation.  A short-period dip with an
+additional complex modulation repeats every 10.7 hours.  Analyzing the
+Gaia data in Section~\ref{sec:gaia}, we show that relative to its
+kinematic group, \ptfo\ is a photometric binary.  We collect and
+discuss the puzzle pieces in Section~\ref{sec:discussion}.  The
+orbital phase of the dip seems to have instantaneously changed over
+the past decade.  In addition, a number of other young stars show
+lightcurve morphologies similar to the short-period signal.  We
+therefore argue that PTFO$\,$8-8695 is a binary M dwarf in which one
+star shows a rotation signal, and the other shows ``transient
+dipping'' caused by eclipses of material at the Keplerian co-rotation
+radius.  Section~\ref{sec:conclusions} summarizes our main points.
 
@@
 Gaia also provided astrometric parameters for tens of thousands of
-young stars in the Orion complex.  The stellar populations in giant
+young stars in the Orion complex.  Stellar populations in giant
 molecular cloud complexes are not monolithic; substructured groups are
@@
 Orion complex.  From smallest to largest, \ptfo\ was identified
-as being a member of the following hierarchical subgroups:
+as a member of the following hierarchical subgroups:
 \begin{equation}
@@
 \end{equation}
-where from set-notation, `$\subset$' denotes ``is a proper subset of''.
-25$\,$Ori-1 is the largest subgroup of 25$\,$Ori, with 149 identified
-members.
-Its mean age from its CMD was found to be 6.9$\,$Myr, and from its HR
-diagram 8.5$\,$Myr \citep{kounkel_apogee2_2018}.
-\citet{kounkel_apogee2_2018} identified seven other smaller groups in
-the Orion complex near the Be star 25$\,$Ori. These groups received
-higher numbers, {\it e.g.}, 25$\,$Ori-2 (${\rm Age}_{\rm CMD} =15.1\,$Myr; ${\rm Age}_{\rm CMD} =12.9\,$Myr; see also \citealt{briceno_cidaII_2019}).
+where from set-notation, `$\subset$' denotes ``is a proper subset
+of''.  25$\,$Ori-1 is the largest subgroup of 25$\,$Ori, with 149
+identified members.  Its mean age from its CMD was found to be
+6.9$\,$Myr, and from its HR diagram 8.5$\,$Myr
+\citep{kounkel_apogee2_2018}.  \citet{kounkel_apogee2_2018} identified
+seven other smaller groups in the Orion complex near the Be star
+25$\,$Ori. These groups received higher numbers, {\it e.g.},
+25$\,$Ori-2 (${\rm Age}_{\rm CMD} =15.1\,$Myr; ${\rm Age}_{\rm CMD}
+=12.9\,$Myr; see also \citealt{briceno_cidaII_2019}).
 
 These details concerning the group membership for one object may seem
-cumbersome to those accustomed to comparing ``young cluster
-members'' with ``old field stars''.  Though all members of the Orion
-complex are indeed young relative to the field, these details are
-essential for assessing any evidence for photometric binarity in
-\ptfo, because of the degeneracy between stellar luminosity and
-age for pre-main-sequence stars.  Having a clean sample of tightly
-spatially and kinematically associated reference stars
-minimizes contamination not just from the field, but from older and
-younger members of the Orion complex itself.
+cumbersome to those accustomed to comparing ``young cluster members''
+with ``old field stars''.  Though all members of the Orion complex are
+indeed young relative to the field, these details are essential for
+assessing any evidence for photometric binarity in \ptfo, because of
+the degeneracy between stellar luminosity and age for
+pre-main-sequence stars.  Having a clean sample of tightly spatially
+and kinematically associated reference stars minimizes contamination
+not just from the field, but from older and younger members of the
+Orion complex itself.
 
@@
     {\it Top}: Long-period signal, as in Figure~\ref{fig:splitsignal}.
-		{\it Bottom}: Short-period signal. The reference phase is set to the
-		``planetary'' dip.  Gray points are the 10 minute cadence
+    {\it Bottom}: Short-period signal. The reference phase is set to
+    the ``planetary'' dip.  Gray points are the 10 minute cadence
     \texttt{PDCSAP} flux.  Black points are binned to 100 points per
@@
 
-Our initial inspection of the TESS lightcurve, in both its 2-minute PDCSAP
-and 30-minute FFI forms, showed a strong sinusoidal beat signal
-(Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}, top panel).
+Our initial inspection of the TESS lightcurve, in both its 2-minute
+PDCSAP and 30-minute FFI forms, showed a strong sinusoidal beat signal
+(Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}, top
+panel).
 
@@
 0.499\,{\rm days}$.  The $P_{\rm \ell}$ peak had the greater power of
-the two.  Smaller harmonics surrounding each of these two dominants peaks
-were also present.
+the two.  Smaller harmonics surrounding each of these two dominants
+peaks were also present.
 
@@
 
-We fitted the lightcurve as a linear combination of Fourier
-harmonics at the short and long periods, plus a transit at the short
-period.  Symbolically, the total flux $f$ is given as
+We fitted the lightcurve as a linear combination of Fourier harmonics
+at the short and long periods, plus a transit at the short period.
+Symbolically, the total flux $f$ is given as
 \begin{equation}
@@
 used the model and derivatives implemented in \texttt{exoplanet}
-\citep{exoplanet:exoplanet}.  Our priors are listed in
-Table~2.  To speed up the fitting, we re-sampled the
-cleaned 2 minute lightcurves to 10 minute binning.  We correspondingly
-scaled the uncertainties in the flux measurements by a factor of
-$\sqrt{5}$.  Before sampling, we initialized each model to the maximum
-a posteriori (MAP) solution.  We then sampled using \texttt{PyMC3}'s
-gradient-based No-U-Turn Sampler \citep{hoffman_no-u-turn_2014}, and
-used $\hat{R}$ as our convergence diagnostic
-\citep{gelman_inference_1992}.
-We tested our ability to successfully recover injected parameters
-using synthetic data, before fitting the \ptfo\
-lightcurves.
+\citep{exoplanet:exoplanet}.  Our priors are listed in Table~2.  To
+speed up the fitting, we re-sampled the cleaned 2 minute lightcurves
+to 10 minute binning.  We correspondingly scaled the uncertainties in
+the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
+initialized each model to the maximum a posteriori (MAP) solution.  We
+then sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
+\citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
+diagnostic \citep{gelman_inference_1992}.  We tested our ability to
+successfully recover injected parameters using synthetic data, before
+fitting the \ptfo\ lightcurves.
 
@@
 $N$ and $M$ ranging from one to three.  To select our preferred model,
-we used the Bayesian information criterion
-(Table~1).  The model with the lowest BIC had
-two harmonics at the short 10.74$\,$hr period, and two harmonics
-at the long 11.96$\,$hr period.  
-The next-best model had an additional harmonic at the longer period ($M=3$),
-but was otherwise identical.
-All nine models have reduced
-$\chi^2$ ranging between 1.37 and 1.51, which suggests a plausible
-though imperfect agreement between the data and fitting results.
-The best-fit parameters for the lowest BIC model are given in Table~2.
+we used the Bayesian information criterion (Table~1).  The model with
+the lowest BIC had two harmonics at the short 10.74$\,$hr period, and
+two harmonics at the long 11.96$\,$hr period.  The next-best model had
+an additional harmonic at the longer period ($M=3$), but was otherwise
+identical.  All nine models have reduced $\chi^2$ ranging between 1.37
+and 1.51, which suggests a plausible though imperfect agreement
+between the data and fitting results.  The best-fit parameters for the
+lowest BIC model are given in Table~2.
 
@@
 
-In every model, the $11.96\,{\rm hr}$ variability is a simple
-sinusoid with peak-to-peak amplitude $\approx$10\%.  The $10.74\,{\rm hr}$ variability
-is always more complex.  A dip of depth
-$\approx$1.2\%, fit in our model as a transit, lasts $\approx$0.75
-hours.  Superposed on the dip is a slightly asymmetric sinusoid with peak-to-peak
-amplitude of about 4\%. The asymmetric sinusoid peaks near phase 0.25, and reaches
-minimum brightness between phases -0.5 and -0.25.
-The flux at phase $\approx -0.33$ shows what could
-be a discontinuous jump, shortly after reaching minimum (Figure~\ref{fig:phasefold}).  This jump
-was visible in each of the nine models we considered.
+In every model, the $11.96\,{\rm hr}$ variability is a simple sinusoid
+with peak-to-peak amplitude $\approx$10\%.  The $10.74\,{\rm hr}$
+variability is always more complex.  A dip of depth $\approx$1.2\%,
+fit in our model as a transit, lasts $\approx$0.75 hours.  Superposed
+on the dip is a slightly asymmetric sinusoid with peak-to-peak
+amplitude of about 4\%. The asymmetric sinusoid peaks near phase 0.25,
+and reaches minimum brightness between phases -0.5 and -0.25.  The
+flux at phase $\approx -0.33$ shows what could be a discontinuous
+jump, shortly after reaching minimum (Figure~\ref{fig:phasefold}).
+This jump was visible in each of the nine models we considered.
 
@@
 The TESS pixels are $\approx21$'' per side. Before making any
-interpretations, we needed to consider whether light from known neighboring
-stars could have affected the photometry.  The scene is shown in
-Figure~\ref{fig:scene}.  
-In the upper panels, the pixels used to measure the background
-level in the SPOC lightcurve are indicated with an `\texttt{X}'
-hatch, and the pixels used in the final lightcurve aperture are shown
-with the `\texttt{/}' hatch.
-
-The target star, \ptfo\ (TIC 264461976), has a $T$-band magnitude
-of 14.0, and its position is shown with a star.  The other (unlabeled)
+interpretations, we needed to consider whether light from known
+neighboring stars could have affected the photometry.  The scene is
+shown in Figure~\ref{fig:scene}.  In the upper panels, the pixels used
+to measure the background level in the SPOC lightcurve are indicated
+with an `\texttt{X}' hatch, and the pixels used in the final
+lightcurve aperture are shown with the `\texttt{/}' hatch.
+
+The target star, \ptfo\ (TIC 264461976), has a $T$-band magnitude of
+14.0, and its position is shown with a star.  The other (unlabeled)
 star inside the target aperture, TIC 264461979, has $T=16.8$ and so
@@
 contaminate the target star is TIC 264461980, with $T=14.8$, which we
-dub ``Star A''.  Star A is 23.6'' NW of our target, and based on
-the magnitude difference could contribute up to 48\% the flux of our
+dub ``Star A''.  Star A is 23.6'' NW of our target, and based on the
+magnitude difference could contribute up to 48\% the flux of our
 target star, \ptfo.
@@
 consistent with our measurement of $P_{\rm s}$, our main concern
-regarding blending was the degree to which we could be certain that the
-long-period signal at $P_{\rm \ell}$ also originated from \ptfo.
-We took two approaches to verifying the source of the
-long-period signal.
-
-First, we examined the CDIPS FFI lightcurves of the
-target, which were available on MAST \citep{bouma_cluster_2019}.  The
-maximal peak-to-peak beat amplitude was constant
-across apertures of radii 1, 1.5, and 2.25 pixels to visual precision ($\lesssim 1\%$).
-If Star A were the
-source of the long-period variability, we would expect the peak
-variability amplitude to be smallest in the 1 pixel aperture, based on
-the separation of the sources (Figure~\ref{fig:scene}, bottom).  From
-this test alone, it seems unlikely that Star A is the source of the
-long-period signal.
+regarding blending was the degree to which we could be certain that
+the long-period signal at $P_{\rm \ell}$ also originated from \ptfo.
+We took two approaches to verifying the source of the long-period
+signal.
+
+First, we examined the CDIPS FFI lightcurves of the target, which were
+available on MAST \citep{bouma_cluster_2019}.  The maximal
+peak-to-peak beat amplitude was constant across apertures of radii 1,
+1.5, and 2.25 pixels to visual precision ($\lesssim 1\%$).  If Star A
+were the source of the long-period variability, we would expect the
+peak variability amplitude to be smallest in the 1 pixel aperture,
+based on the separation of the sources (Figure~\ref{fig:scene},
+bottom).  From this test alone, it seems unlikely that Star A is the
+source of the long-period signal.
 
@@
 nearest to Star A to show a sinusoidal signal with amplitude exceeding
-$10\%$.  The data do not support this possibility.  The pixel
-directly below Star A does not clearly show the sinusoidal
-variability, and the peak-to-peak variability in that pixel is
-$\lesssim 8\%$.  In contrast, the south-easternmost pixel within
-\ptfo's aperture (the pixel furthest from Star A that was used in the
-optimal aperture) shows the $P_{\rm \ell}$ sinusoidal variability
-signal at $\approx 14\%$ amplitude.
-We conclude that
-within the resolution of the Gaia DR2 source catalog,
-the $P_{\rm s}$ and $P_{\rm \ell}$ signals originate from \ptfo.
-From \citet{ziegler_measuring_2018}, we can surmise that
-stellar companions
-wider than $\approx$1'' (349~AU) and within
-$\Delta G \approx 3$ magnitudes of \ptfo\ would have likely been
-detected through this approach.
-
-\citet{van_eyken_ptf_2012} obtained
-stronger constraints on possible stellar companions using the
-NIRC2 camera on Keck II.
-They reported 3$\sigma$ $H$-band magnitude difference limits
-of 4.3, 6.4, and 8.9 at angular separations of 0.25, 0.5, and 1.0
-arcseconds (87, 175, and 349~AU).
-They also detected a point-source, not present in Gaia DR2,
-7.0 magnitudes fainter than the target, 
-and 1.8$''$ north-east.
-Due to the brightness difference, this coincident star\footnote{This companion was claimed to be a potential planetary-mass
-	object \citep{schmidt_direct_2016}. Subsequent analysis of its colors showed that it is
-	a background star \citep{lee_evidence_2018}.} cannot be the source
-of our signals.
+$10\%$.  The data do not support this possibility.  The pixel directly
+below Star A does not clearly show the sinusoidal variability, and the
+peak-to-peak variability in that pixel is $\lesssim 8\%$.  In
+contrast, the south-easternmost pixel within \ptfo's aperture (the
+pixel furthest from Star A that was used in the optimal aperture)
+shows the $P_{\rm \ell}$ sinusoidal variability signal at $\approx
+14\%$ amplitude.  We conclude that within the resolution of the Gaia
+DR2 source catalog, the $P_{\rm s}$ and $P_{\rm \ell}$ signals
+originate from \ptfo.  From \citet{ziegler_measuring_2018}, we can
+surmise that stellar companions wider than $\approx$1'' (349~AU) and
+within $\Delta G \approx 3$ magnitudes of \ptfo\ would have likely
+been detected through this approach.
+
+\citet{van_eyken_ptf_2012} obtained stronger constraints on possible
+stellar companions using the NIRC2 camera on Keck II.  They reported
+3$\sigma$ $H$-band magnitude difference limits of 4.3, 6.4, and 8.9 at
+angular separations of 0.25, 0.5, and 1.0 arcseconds (87, 175, and
+349~AU).  They also detected a point-source, not present in Gaia DR2,
+7.0 magnitudes fainter than the target, and 1.8$''$ north-east.  Due
+to the brightness difference, this coincident star\footnote{This
+companion was claimed to be a potential planetary-mass object
+\citep{schmidt_direct_2016}. Subsequent analysis of its colors showed
+that it is a background star \citep{lee_evidence_2018}.} cannot be the
+source of our signals.
 
 %FIXME: would be better to actually be able to get them...
-While long-baseline radial velocity measurements could provide complimentary
-lower limits on companion masses and separations, the RV
-data for \ptfo\ are rather poor owing to the stellar faintness.
-The longest single-instrument baseline in the literature appears to be 5 Keck/HIRES
-measurements acquired over 10 days in April 2011 by \citet{van_eyken_ptf_2012}.
-The RV RMS over that 10 day span was $1.6\,{\rm km}\,{\rm s}^{-1}$, consistent with
-the measurement precision.
-Though \citet{yu_tests_2015} acquired 22 further Keck/HIRES observations over
-one night in December 2013, these points seem to not have been
-reduced to velocities.
-Further Keck/HIRES measurements of this target could confirm or
-refute the existence of possible binary companions.
+While long-baseline radial velocity measurements could provide
+complimentary lower limits on companion masses and separations, the RV
+data for \ptfo\ are rather poor owing to the stellar faintness.  The
+longest single-instrument baseline in the literature appears to be 5
+Keck/HIRES measurements acquired over 10 days in April 2011 by
+\citet{van_eyken_ptf_2012}.  The RV RMS over that 10 day span was
+$1.6\,{\rm km}\,{\rm s}^{-1}$, consistent with the measurement
+precision.  Though \citet{yu_tests_2015} acquired 22 further
+Keck/HIRES observations over one night in December 2013, these points
+seem to not have been reduced to velocities.  Further Keck/HIRES
+measurements of this target could confirm or refute the existence of
+possible binary companions.
 
@@
     {\it Top: Hertzsprung-Russell diagram of \ptfo\ and late-type
-			members of 25$\,$Ori-1.}
-		Members of the 25$\,$Ori-1 group (black circles) were identified by
-		\citet{kounkel_apogee2_2018} through clustering on six-dimensional
-		Gaia DR2 and APOGEE-2 data.  The ``neighborhood'' (gray
-		circles) is non-member
-		stars within 5 standard deviations of the mean 25$\,$Ori-1 right
-		ascension, declination, and parallax.  It contains members of the
-		Orion complex with its full spread of ages, in addition to field
-		interlopers.  $G$ denotes Gaia broadband magnitudes, $Bp$ Gaia blue, $Rp$
-		Gaia red, and $\omega_{\rm as}$ the parallax in arcseconds.  The
-		$x$-axis limits have been set to show only K and M dwarf members, to
-		accentuate \ptfo's separation from the single-star sequence.
-		{\it Bottom: Astrometric goodness-of-fit versus $Rp$ magnitude
-			for 25$\,$Ori-1 members}.
-		The single-source astrometric model provides a poor fit, which
-		could be due to stellar variability or binarity.
+    members of 25$\,$Ori-1.} Members of the 25$\,$Ori-1 group (black
+    circles) were identified by \citet{kounkel_apogee2_2018} through
+    clustering on six-dimensional Gaia DR2 and APOGEE-2 data.  The
+    ``neighborhood'' (gray circles) is non-member stars within 5
+    standard deviations of the mean 25$\,$Ori-1 right ascension,
+    declination, and parallax.  It contains members of the Orion
+    complex with its full spread of ages, in addition to field
+    interlopers.  $G$ denotes Gaia broadband magnitudes, $Bp$ Gaia
+    blue, $Rp$ Gaia red, and $\omega_{\rm as}$ the parallax in
+    arcseconds.  The $x$-axis limits have been set to show only K and
+    M dwarf members, to accentuate \ptfo's separation from the
+    single-star sequence.  {\it Bottom: Astrometric goodness-of-fit
+    versus $Rp$ magnitude for 25$\,$Ori-1 members}.  The single-source
+    astrometric model provides a poor fit, which could be due to
+    stellar variability or binarity.
 		\label{fig:gaia}
@@
 
-Aside from visual binarity, we can also check the Gaia data
-for photometric binarity.
-To assemble a set of stars coeval with \ptfo,
+Aside from visual binarity, we can also check the Gaia data for
+photometric binarity.  To assemble a set of stars coeval with \ptfo,
 we used the 25$\,$Ori-1 members identified by
-\citet{kounkel_apogee2_2018}, and discussed in Section~\ref{subsec:hierarchical}.
+\citet{kounkel_apogee2_2018}, and discussed in
+Section~\ref{subsec:hierarchical}.
 
@@
 declination, and parallax.  We queried Gaia DR2 for these stars using
-\texttt{astroquery} \citep{astroquery_2018}.  This yielded 1{,}819 neighbors.  While
-some of these stars may indeed be members of the Orion complex, or
-even of 25$\,$Ori-1, enforcing this cut on positions and parallaxes
-ensures that we are querying stars with comparable amounts of
-interstellar reddening.
+\texttt{astroquery} \citep{astroquery_2018}.  This yielded 1{,}819
+neighbors.  While some of these stars may indeed be members of the
+Orion complex, or even of 25$\,$Ori-1, enforcing this cut on positions
+and parallaxes ensures that we are querying stars with comparable
+amounts of interstellar reddening.
 
@@
 was that 25$\,$Ori-1 was a clearly defined over-density in each
-dimension---the cluster exists, and is distinct from the
-neighborhood.  \ptfo\ was also within the cluster in each of these
-projected dimensions.
+dimension---the cluster exists, and is distinct from the neighborhood.
+\ptfo\ was also within the cluster in each of these projected
+dimensions.
 
@@
 25$\,$Ori-1 star of the same color.  In other words, it is about twice
-as bright.  It also seems to be coincident with the photometric binary track of the
-cluster, which has a few other stars.
+as bright.  It also seems to be coincident with the photometric binary
+track of the cluster, which has a few other stars.
 
 The implication is that either {\it (i)} \ptfo\ is notably younger
-than the kinematically identical 25$\,$Ori-1 members, or {\it (ii)} \ptfo\
-is a photometric binary.
-Since there is no a priori reason to suspect an age difference,
-but we have resolved two
-separate photometric signals, we favor the binary interpretation.
+than the kinematically identical 25$\,$Ori-1 members, or {\it (ii)}
+\ptfo\ is a photometric binary.  Since there is no a priori reason to
+suspect an age difference, but we have resolved two separate
+photometric signals, the binary interpretation seems somewhat more
+likely.
 
@@
 
-A separate possible line of evidence for binarity is the Gaia DR2 astrometry.
-As noted in Section~\ref{sec:observations}, the Gaia DR2 solution for
-\ptfo\ shows a 10.3$\sigma$ astrometric excess.  This astrometric
-excess indicates the degree to which a single-source model fails to
-explain the observed astrometric measurements.  Specifically, the
-single-source astrometric model yielded $\chi^2=325.2$.  There
-are 121 astrometric measurements, and 5 free parameters, and therefore
-116 degrees of freedom. The reduced $\chi^2$ is 2.80.  The
-majority of stars with comparable brightness in Gaia do not show such
-poor goodness-of-fit \citep[][Appendix A]{lindegren_gaiasoln_2018}.
-
-Potential explanations for the poor astrometric fit include photometric
-variability and unresolved stellar binarity \citep[{\it
-e.g.},][]{rizzuto_ZEIT8_2018,belokurov_unresolved_2020}.
-If photometric variability were the cause, we would expect comparably
-faint stars in the same kinematic group of Orion to
-show similar astrometric excesses, as the majority of
-young stars are highly variable.
+A separate possible line of evidence for binarity is the Gaia DR2
+astrometry.  As noted in Section~\ref{sec:observations}, the Gaia DR2
+solution for \ptfo\ shows a 10.3$\sigma$ astrometric excess.  This
+astrometric excess indicates the degree to which a single-source model
+fails to explain the observed astrometric measurements.  Specifically,
+the single-source astrometric model yielded $\chi^2=325.2$.  There are
+121 astrometric measurements, and 5 free parameters, and therefore 116
+degrees of freedom. The reduced $\chi^2$ is 2.80.  The majority of
+stars with comparable brightness in Gaia do not show such poor
+goodness-of-fit \citep[][Appendix A]{lindegren_gaiasoln_2018}.
+
+Potential explanations for the poor astrometric fit include
+photometric variability and unresolved stellar binarity \citep[{\it
+e.g.},][]{rizzuto_ZEIT8_2018,belokurov_unresolved_2020}.  If
+photometric variability were the cause, we would expect comparably
+faint stars in the same kinematic group of Orion to show similar
+astrometric excesses, as the majority of young stars are highly
+variable.
 
@@
 the upper 90$^{\rm th}$ percentile of stars showing astrometric
-excesses within the 25$\,$Ori-1 group.  
-Relative to other M-dwarf group members with comparable
-brightnesses and variability characteristics, \ptfo\ still stands out
-as behaving astrometrically poorly.
-Ultimately, we will have to wait
-for the full release of the nominal Gaia mission to definitively
-determine whether the astrometric excess is caused by stellar binarity
-or photometric variability.
-Nonetheless this result suggests that stellar binarity
-is indeed the root cause, because other member stars have identical variability
+excesses within the 25$\,$Ori-1 group.  Relative to other M-dwarf
+group members with comparable brightnesses and variability
+characteristics, \ptfo\ still stands out as behaving astrometrically
+poorly.  Ultimately, we will have to wait for the full release of the
+nominal Gaia mission to definitively determine whether the astrometric
+excess is caused by stellar binarity or photometric variability.
+Nonetheless this result suggests that stellar binarity is indeed the
+root cause, because other member stars have identical variability
 characteristics but do not show such a large astrometric excess.
@@
 
-The standard interpretation for 11.96$\,$hour sinusoidal modulations of a
-pre-main-sequence M dwarf is stellar rotation.  This is the dominant signal in the system with
-10\% amplitude, and there is no evidence to suggest that this signal
-has any other origin.
-
-The discovery study by \citet{van_eyken_ptf_2012} saw an alias of the same signal
-({\it e.g.}, their Figure~7), and identified it as a
+The standard interpretation for 11.96$\,$hour sinusoidal modulations
+of a pre-main-sequence M dwarf is stellar rotation.  This is the
+dominant signal in the system with 10\% amplitude, and there is no
+evidence to suggest that this signal has any other origin.
+
+The discovery study by \citet{van_eyken_ptf_2012} saw an alias of the
+same signal ({\it e.g.}, their Figure~7), and identified it as a
 periodogram peak at $0.9985 \pm 0.0061\,$days. They ascribed it to
@@
 \citet{koen_multicolour_2015} identified the same modes and aliases as
-\citet{van_eyken_ptf_2012}, but argued that the signal
-was astrophysical (however they were still unsure of the exact period).
-Using photometry from the
-YETI global telescope network, \citet{raetz_yeti_2016} eventually came
-to the conclusion that the $0.50\,{\rm d}$ signal was indeed
-from stellar rotation.  The TESS data strongly support this
-conclusion.
+\citet{van_eyken_ptf_2012}, but argued that the signal was
+astrophysical (however they were still unsure of the exact period).
+Using photometry from the YETI global telescope network,
+\citet{raetz_yeti_2016} eventually came to the conclusion that the
+$0.50\,{\rm d}$ signal was indeed from stellar rotation.  The TESS
+data strongly support this conclusion.
 
 \subsection{Short period dip}
-The TESS lightcurve shows a 
-dip that lasts about 45 minutes, and seems to re-occur every 10.74
-hours
+The TESS lightcurve shows a dip that lasts about 45 minutes, and seems
+to re-occur every 10.74 hours
 (Figures~\ref{fig:splitsignal},~\ref{fig:splitsignalii},~\ref{fig:phasefold}).
 The dip duration is roughly the same as that observed by previous
-investigators \citep{van_eyken_ptf_2012,yu_tests_2015}.
-The 1.2\% depth is similar to what has been observed in the near-infrared
-\citep{onitsuka_multicolor_2017}.
-However the dip depth seems likely to have evolved over time between
-being not present at all, to a maximum of $\approx$5\%
-\citep[{\it e.g.},]{koen_multicolour_2015,yu_tests_2015,tanimoto_evidence_2020}.
-
-One particularly interesting feature of the dip is its epoch.
-Over the past decade, many investigators have observed \ptfo.
-Its dips do not always occur on a 
-perfectly linear ephemeris \citep{yu_tests_2015}.
-In fact, \citet{tanimoto_evidence_2020} recently provided stark
-evidence for different behavior altogether: over a time-span of
-years, the dip ``splits'' into distinct groups at particular repeating phases.
-See for instance their Figures~2 through~4.
-Fitting a decade of observations, they provided the following linear
-ephemeris.
+investigators \citep{van_eyken_ptf_2012,yu_tests_2015}.  The 1.2\%
+depth is similar to what has been observed in the near-infrared
+\citep{onitsuka_multicolor_2017}.  However the dip depth seems likely
+to have evolved over time between being not present at all, to a
+maximum of $\approx$5\% \citep[{\it
+e.g.},]{koen_multicolour_2015,yu_tests_2015,tanimoto_evidence_2020}.
+
+One particularly interesting feature of the dip is its epoch.  Over
+the past decade, many investigators have observed \ptfo.  Its dips do
+not always occur on a perfectly linear ephemeris
+\citep{yu_tests_2015}.  In fact, \citet{tanimoto_evidence_2020}
+recently provided stark evidence for different behavior altogether:
+over a time-span of years, the dip ``splits'' into distinct groups at
+particular repeating phases.  See for instance their Figures~2
+through~4.  Fitting a decade of observations, they provided the
+following linear ephemeris.
 \begin{align}
@@
 
-In Figure~\ref{fig:o_minus_c}, we show the phase of the dip we
-detect in the TESS data, relative to their ephemeris.
-It agrees with the independent December 2018 measurements by \citet{tanimoto_evidence_2020}:
-the dip has drastically shifted phase over the past decade.
-
-Figure~\ref{fig:o_minus_c} shows two additional strange features:
-{\it (i)}  multiple dips per cycle, and {\it (ii)} a set of dips numerically coincident
-with phase $(P_{\rm \ell} - P_{\rm s}) / P_{\rm s}$.
-The observation of multiple
-dips per cycle in 2015 seems to have been detected by both \citet{yu_tests_2015}
-and \citet{tanimoto_evidence_2020}.
-It therefore seems credible.
-Inspecting the \citet{tanimoto_evidence_2020} lightcurves,
-the claim of multiple dips per cycle in December 2018 at phase 0 and -0{.}47
-seems much less plausible---the phase -0{.}47 dips are strongly detected,
-while the suggested phase 0 dip is not clearly present.
-
-We are not sure what to make of the numerical coincidence.
-The ratio of long to short periods is roughly 10:9, which does produce a beat.
-However it is not clear that this would obviously translate into
-an observational bias unless by some fluke three season's worth 
-of observations managed to only observe every ninth dip.
-This is of course not the case, and we therefore leave this 
-curiosity as observation {\it sans} interpretation.
+In Figure~\ref{fig:o_minus_c}, we show the phase of the dip we detect
+in the TESS data, relative to their ephemeris.  It agrees with the
+independent December 2018 measurements by
+\citet{tanimoto_evidence_2020}: the dip has drastically shifted phase
+over the past decade.
+
+Figure~\ref{fig:o_minus_c} shows two additional strange features: {\it
+(i)}  multiple dips per cycle, and {\it (ii)} a set of dips
+numerically coincident with phase $(P_{\rm \ell} - P_{\rm s}) / P_{\rm
+s}$.  The observation of multiple dips per cycle in 2015 was seen
+independently by both \citet{yu_tests_2015} and
+\citet{tanimoto_evidence_2020}.  It therefore seems credible.
+Inspecting the \citet{tanimoto_evidence_2020} lightcurves, the claim
+of multiple dips per cycle in December 2018 at phase 0 and -0{.}47
+seems less plausible---the phase -0{.}47 dips are strongly detected,
+while the suggested phase 0 dip is not clearly present in the data.
+
+We are not sure what to make of the numerical coincidence.  The ratio
+of long to short periods is roughly 10:9.  It is not clear that this
+would obviously translate into an observational bias unless by some
+fluke three season's worth of observations managed to only observe
+every ninth dip.  This is of course not the case, and we therefore
+leave this curiosity as observation {\it sans} interpretation.
 
@@
     convincingly {\it observed}.  Horizontal dashed lines are drawn at
-    $\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting either a numerical
-    coincidence or an observational bias.  The orbital phase observed by TESS
-    is consistent with that of \citet{tanimoto_evidence_2020}, and
-    quite different from the original phase.
+    $\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting either a
+    numerical coincidence or an observational bias.  The orbital phase
+    observed by TESS is consistent with that of
+    \citet{tanimoto_evidence_2020}, and quite different from the
+    original phase.
 		\label{fig:o_minus_c}
@@
 \subsection{Short period out-of-dip modulation}
+
+%FIXME write it from here!
+
 If there were a giant planet transiting \ptfo, it would tidally
@@
 parametrized as the ``$B_1$'' component.  The amplitude of the signal
-is $0.53 \pm 0.06\%$ (see Table~2).  Interpreted as being caused by a planet, it
-would imply a minimum planet mass $M_{\rm p} \sin i$ of $3.8\,M_{\rm
-Jup}$.
+is $0.53 \pm 0.06\%$ (see Table~2).  Interpreted as being caused by a
+planet, it would imply a minimum planet mass $M_{\rm p} \sin i$ of
+$3.8\,M_{\rm Jup}$.
 
 %FIXME but argue against
+
+\subsection{Similar Lightcurves}
+
+Figure~\ref{fig:brethren}
 

### 2020-04-21 billy/plotting.py
update brethren plot

@@ -1008,16 +1008,31 @@ def plot_brethren(outdir):
 
     from transitleastsquares import transitleastsquares
 
-    # get campaign 2 upper sco everest light-curves.
-    # NOTE: 204270520 has a C15 LC, but it's the only one.
+    # IDs to plot
     epic_ids = [
-        "204143627",
-        "204270520",
-        "204321142",
-        # "205046529", # kind of messed up, b/c it's a binary
-        "205483258",
-        '204787516',
+        '204143627',
+        '204270520',
+        '204321142',
+        '246938594', # 'Taurus',
+        '205483258' # RIK-210
     ]
+
+    # see /Users/luke/Dropbox/proj/billy/doc/20200417_list_of_analogs.txt
+    id_dict = {
+        "204143627": 'USco',
+        "204270520": 'USco', # 204270520 has a C15 LC, but it's the only one.
+        "204321142": 'USco',
+        "205046529": 'USco', # kind of messed up, b/c it has two components
+        "205483258": 'USco',
+        '204787516': 'USco',
+        '246938594': 'Taurus',
+        '246969828': 'Taurus',
+        '247794636': 'Taurus',  # does some wild stuff
+        '246676629': 'Taurus',
+        '246682490': 'Taurus',
+        '247343526': 'Taurus'
+    }
+
     lc_dict = {}
 
     k2pkl = os.path.join(outdir, 'k2data.pkl')
@@ -1026,9 +1041,14 @@ def plot_brethren(outdir):
 
             lc_dict[epic_id] = {}
 
-            lcpath = glob(
-                '/Users/luke/Dropbox/proj/billy/data/analogs/hlsp_everest_k2_llc_{}-c02_kepler_v2.0_lc/*fits'.format(epic_id)
-            )
+            if id_dict[epic_id] == 'USco':
+                lcpath = glob(
+                    '/Users/luke/Dropbox/proj/billy/data/analogs/hlsp_everest_k2_llc_{}-c02_kepler_v2.0_lc/*fits'.format(epic_id)
+                )
+            elif id_dict[epic_id] == 'Taurus':
+                lcpath = glob(
+                    '/Users/luke/Dropbox/proj/billy/data/analogs/hlsp_everest_k2_llc_{}-c13_kepler_v2.0_lc/*fits'.format(epic_id)
+                )
             assert len(lcpath)==1
 
             hdul = fits.open(lcpath[0])
[…]

### 2020-04-21 drivers/plot_brethren.py

@@ -3,6 +3,12 @@ import billy.plotting as bp
 from billy import __path__
 
 RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
+
+# # exploration options
+# PLOTDIR = os.path.join(RESULTSDIR, 'brethren_taurus')
+# PLOTDIR = os.path.join(RESULTSDIR, 'brethren_usco')
+
+# final plot
 PLOTDIR = os.path.join(RESULTSDIR, 'brethren')
 
 bp.plot_brethren(PLOTDIR)

### 2020-04-21 paper/ms.tex
added words

@@
 through~4.  Fitting a decade of observations, they provided the
-following linear ephemeris.
+following linear ephemeris, which we did not find any need to update.
 \begin{align}
@@
 
-%FIXME write it from here!
-
+Visually, the out-of-dip modulation at the 10.74$\,$hour period
+resembles a slightly asymmetric sinusoid (Figure~\ref{fig:phasefold}).
+Non-zero contributions in both the first and second harmonics are
+detected (Table~2).  The first sine and cosine harmonic both have
+amplitudes of roughly $0.9\pm0.1\%$.  The second sine harmonic has
+amplitude $0.16 \pm 0.07\%$, so is non-zero at a significance of only
+2.3$\sigma$.  The second cosine harmonic has negative amplitude $0.53
+\pm 0.06\%$.  In our sign convention, the fact that it is negative
+means that this component peaks at phase 0.25 and 0.75.
+
+\subsubsection{Planetary interpretation}
 If there were a giant planet transiting \ptfo, it would tidally
-distort the host star, and cause ellipsoidal photometric modulations.
-The amplitude of the ellipsoidal distortion for a $1\,M_{\rm Jup}$
-companion would be about 1400$\,$ppm
-\citep{shporer_astrophysics_2017}.  This is significantly larger than
-the typical ellipsopidal modulation induced by close-in giant planets
-because the host star is puffy, and still on the pre-main-sequence.
-For our estimate, we assumed $R_\star = 1.39 R_\odot$, and $M_\star =
-0.39 M_\odot$ \citep{van_eyken_ptf_2012}.
-
-Our preferred model does detect a significant ellipsoidal signal,
-parametrized as the ``$B_1$'' component.  The amplitude of the signal
-is $0.53 \pm 0.06\%$ (see Table~2).  Interpreted as being caused by a
-planet, it would imply a minimum planet mass $M_{\rm p} \sin i$ of
-$3.8\,M_{\rm Jup}$.
-
-%FIXME but argue against
-
-\subsection{Similar Lightcurves}
-
-Figure~\ref{fig:brethren}
-
-\section{Physical Interpretation}
-
-%FIXME: the PTFO lightcurve looks so much worse! Why is it less dense??
+distort the host star, and cause ellipsoidal photometric modulations
+that also peak at quadrature \citep[see][]{shporer_astrophysics_2017}.
+Interpreting the second cosine harmonic as being caused by a planet,
+it would imply a minimum planet mass $M_{\rm p} \sin i$ of
+$3.8\,M_{\rm Jup}$.  For this estimate, we assumed $R_\star = 1.39
+R_\odot$, and $M_\star = 0.39 M_\odot$ \citep{van_eyken_ptf_2012}.
+This ellipsoidal amplitude is larger than the typical
+modulations induced by close-in giant planets because the host star is
+puffy, and still on the pre-main-sequence.
+
+The planetary interpretation however does not readily explain the
+first sine and cosine harmonics.  Interpreting the sine component as
+Doppler beaming would imply a secondary mass greater than the primary
+($0.86\,M_\odot$).  Interpreting the cosine component as reflected or
+emitted light from the planet's surface is nonsensical because the sign is wrong---the
+planet would need to be {\it absorbing} light.
+
+\subsubsection{Similar Lightcurves}
+
 \begin{figure*}[hbtp]
@@
   \caption{ {\bf \ptfo\ and its brethren.}
-    Five transient and persistent flux dip stars, selected based on
-    their visual similarity to the short-period signal in \ptfo, are
-    as follows.
-  	EPIC 204143627, 1.1250d, dips change depth too
-  	EPIC 204321142  Usco flux dip, 0.476d
-  	EPIC 204270520  USco flux dip, 0.512d
-  	% EPIC 205046529B, 1.8358d, indicates dips can be at weird phases too
-  	RIK-210 = EPIC 205483258
- 	EPIC 204787516  USco flux dip or possible? EB, 0.487d
- 	We found these objects through studies by \citet{stauffer_orbiting_2017}, \citet{david_transient_2017}, and \citet{rebull_usco_2018}.
+    Five transient and persistent flux dip stars selected based on
+    their visual similarity to the short-period signal in \ptfo are as
+    shown.  They include EPIC 204143627, EPIC 204321142, EPIC
+    204270520, EPIC 205483258 (RIK-210), and EPIC 204787516.  RIK-210
+    has the longest period of any of these objects.  EPIC 204787516
+    has a less angular dip, and could also be an eclipsing binary.
+    All stars have similar properties as \ptfo, but are in Upper Sco.
+    We found these objects through studies by
+    \citet{stauffer_orbiting_2017}, \citet{david_transient_2017}, and
+    \citet{rebull_usco_2018}.
 		\label{fig:brethren}
@@
 \end{figure*}
+
+When physical explanations are not immediately clear, taxonomy can be
+a useful exercise.  To our knowledge, of order 10 lightcurves with
+similar morphologies have been seen in surveys of low-mass weak-lined
+T Tauri stars in regions including $\rho$ Oph, Upper Sco, Taurus, and perhaps
+the Pleiades
+\citep{rebull_rotation_2016,david_transient_2017,stauffer_orbiting_2017,stauffer_rotevol_2018,rebull_usco_2018,rebull_rotation_2020}.
+These surveys were performed using K2 \citep{howell_k2_2014}.  We
+downloaded a few of these lightcurves from MAST, opting for the
+EVEREST reductions \citep{luger_everest_2016,luger_update_2018}.  They
+are plotted in Figure~\ref{fig:brethren}
+
+These lightcurves have been classified phenomenologically as
+``persistent flux dips'' or ``transient flux dips'', based on whether
+their depths and durations show variability over the 90-day K2
+campaigns \citep{stauffer_orbiting_2017}.
+This class of objects has the following properties.
+\begin{enumerate}
+  \item Weak-lined T Tauri stars.
+  \item Spectral type typically M2-M5 ({\it e.g.},
+    \citet{rebull_usco_2018}~Figure~20).
+  \item Ages $\lesssim$ 100$\,$Myr.
+  \item Stars typically lack spectroscopic accretion indicators 
+  \item Phased lightcurves show shallow, angular dips, usually
+    superposed on large-amplitude smooth variability (interpreted as
+    stellar rotation).
+  \item Rotation is usually rapid ($P<1\,{\rm day}$).
+  \item Rarely any infrared excess (never any W4 detection; only a few
+    cases with W3 detctions).
+  \item Can show multiple dips per cycle ({\it e.g.}, EPIC 203692610,
+    RIK-210).
+  \item Dip depths, durations, and phasescan vary over timescales of days
+    ({\it e.g.}, EPIC 204143627).
+  \item Dip depths can change after flares ({\it e.g.}, RIK-210.
+  \item Rare ($\lesssim 1\%$ of relevant stars) at a population level
+    \citep{rebull_usco_2018}.
+\end{enumerate}
+The 10.74 hour signal in \ptfo\ meets essentially all of these
+criteria.
+
+\section{Physical Interpretation}
 

### 2020-04-22 paper/ms.tex
update paper

@@
 \subsection{Short period dip}
-The TESS lightcurve shows a dip that lasts about 45 minutes, and seems
-to re-occur every 10.74 hours
-(Figures~\ref{fig:splitsignal},~\ref{fig:splitsignalii},~\ref{fig:phasefold}).
-The dip duration is roughly the same as that observed by previous
-investigators \citep{van_eyken_ptf_2012,yu_tests_2015}.  The 1.2\%
-depth is similar to what has been observed in the near-infrared
-\citep{onitsuka_multicolor_2017}.  However the dip depth seems likely
-to have evolved over time between being not present at all, to a
-maximum of $\approx$5\% \citep[{\it
-e.g.},]{koen_multicolour_2015,yu_tests_2015,tanimoto_evidence_2020}.
-
-One particularly interesting feature of the dip is its epoch.  Over
-the past decade, many investigators have observed \ptfo.  Its dips do
-not always occur on a perfectly linear ephemeris
-\citep{yu_tests_2015}.  In fact, \citet{tanimoto_evidence_2020}
-recently provided stark evidence for different behavior altogether:
-over a time-span of years, the dip ``splits'' into distinct groups at
-particular repeating phases.  See for instance their Figures~2
-through~4.  Fitting a decade of observations, they provided the
-following linear ephemeris, which we did not find any need to update.
-\begin{align}
-t_0\ {\rm BJD}_{\rm TDB} &= 2455543.943 \pm 0.002 \\
-P &= 0.4483993 \pm 0.0000006 d.
-\end{align}
-
-In Figure~\ref{fig:o_minus_c}, we show the phase of the dip we detect
-in the TESS data, relative to their ephemeris.  It agrees with the
-independent December 2018 measurements by
-\citet{tanimoto_evidence_2020}: the dip has drastically shifted phase
-over the past decade.
-
-Figure~\ref{fig:o_minus_c} shows two additional strange features: {\it
-(i)}  multiple dips per cycle, and {\it (ii)} a set of dips
-numerically coincident with phase $(P_{\rm \ell} - P_{\rm s}) / P_{\rm
-s}$.  The observation of multiple dips per cycle in 2015 was seen
-independently by both \citet{yu_tests_2015} and
-\citet{tanimoto_evidence_2020}.  It therefore seems credible.
-Inspecting the \citet{tanimoto_evidence_2020} lightcurves, the claim
-of multiple dips per cycle in December 2018 at phase 0 and -0{.}47
-seems less plausible---the phase -0{.}47 dips are strongly detected,
-while the suggested phase 0 dip is not clearly present in the data.
-
-We are not sure what to make of the numerical coincidence.  The ratio
-of long to short periods is roughly 10:9.  It is not clear that this
-would obviously translate into an observational bias unless by some
-fluke three season's worth of observations managed to only observe
-every ninth dip.  This is of course not the case, and we therefore
-leave this curiosity as observation {\it sans} interpretation.
 
@@
 
+The TESS lightcurve shows a dip that lasts about 45 minutes, and seems
+to re-occur every 10.74 hours
+(Figures~\ref{fig:splitsignal},~\ref{fig:splitsignalii},~\ref{fig:phasefold}).
+The dip duration is roughly the same as that observed by previous
+investigators \citep{van_eyken_ptf_2012,yu_tests_2015}.  The 1.2\%
+depth is similar to what has been observed in the near-infrared
+\citep{onitsuka_multicolor_2017}.  However the dip depth seems likely
+to have evolved over time between being not present at all, to a
+maximum of $\approx$5\% \citep[{\it
+e.g.},]{koen_multicolour_2015,yu_tests_2015,tanimoto_evidence_2020}.
+
+One particularly interesting feature of the dip is its epoch.  Over
+the past decade, many investigators have observed \ptfo.  Its dips do
+not always occur on a perfectly linear ephemeris
+\citep{yu_tests_2015}.  In fact, \citet{tanimoto_evidence_2020}
+recently provided stark evidence for different behavior altogether:
+over a time-span of years, the dip ``splits'' into distinct groups at
+particular repeating phases.  See for instance their Figures~2
+through~4.  Fitting a decade of observations, they provided the
+following linear ephemeris, which we did not find any need to update.
+\begin{align}
+t_0\ {\rm BJD}_{\rm TDB} &= 2455543.943 \pm 0.002 \\
+P &= 0.4483993 \pm 0.0000006 d.
+\end{align}
+
+In Figure~\ref{fig:o_minus_c}, we show the phase of the dip we detect
+in the TESS data, relative to their ephemeris.  It agrees with the
+independent December 2018 measurements by
+\citet{tanimoto_evidence_2020}: the dip has drastically shifted phase
+over the past decade.
+
+Figure~\ref{fig:o_minus_c} shows two additional strange features: {\it
+(i)}  multiple dips per cycle, and {\it (ii)} a set of dips
+numerically coincident with phase $(P_{\rm \ell} - P_{\rm s}) / P_{\rm
+s}$.  The observation of multiple dips per cycle in 2015 was seen
+independently by both \citet{yu_tests_2015} and
+\citet{tanimoto_evidence_2020}.  It therefore seems credible.
+Inspecting the \citet{tanimoto_evidence_2020} lightcurves, the claim
+of multiple dips per cycle in December 2018 at phase 0 and -0{.}47
+seems less plausible---the phase -0{.}47 dips are strongly detected,
+while the suggested phase 0 dip is not clearly present in the data.
+
+We are not sure what to make of the numerical coincidence.  The ratio
+of long to short periods is roughly 10:9.  It is not clear that this
+would obviously translate into an observational bias unless by some
+fluke three season's worth of observations managed to only observe
+every ninth dip.  This is of course not the case, and we therefore
+leave this curiosity as observation {\it sans} interpretation.
+
 \subsection{Short period out-of-dip modulation}
@@
 that also peak at quadrature \citep[see][]{shporer_astrophysics_2017}.
-Interpreting the second cosine harmonic as being caused by a planet,
+Interpreting the second cosine harmonic as planet-induced tidal distortion,
 it would imply a minimum planet mass $M_{\rm p} \sin i$ of
@@
 When physical explanations are not immediately clear, taxonomy can be
-a useful exercise.  To our knowledge, of order 10 lightcurves with
+a useful exercise.  To our knowledge, of order 10-20 lightcurves with
 similar morphologies have been seen in surveys of low-mass weak-lined
@@
 EVEREST reductions \citep{luger_everest_2016,luger_update_2018}.  They
-are plotted in Figure~\ref{fig:brethren}
-
-These lightcurves have been classified phenomenologically as
+are plotted in Figure~\ref{fig:brethren}.
+
+These lightcurves have been phenomenologically classified as
 ``persistent flux dips'' or ``transient flux dips'', based on whether
@@
 campaigns \citep{stauffer_orbiting_2017}.
-This class of objects has the following properties.
+As \citet{stauffer_orbiting_2017} explain,
+these  objects are morphologically distinct from ``scallop shell' lightcurves,
+and are present in stars at more advanced evolutionary disk stages than
+the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
+The persistent and transient flux dip stars have the following properties.
 \begin{enumerate}
@@
   \item Spectral type typically M2-M5 ({\it e.g.},
-    \citet{rebull_usco_2018}~Figure~20).
+    \citealt{rebull_usco_2018},~Figure~20).
   \item Ages $\lesssim$ 100$\,$Myr.
@@
   \item Phased lightcurves show shallow, angular dips, usually
-    superposed on large-amplitude smooth variability (interpreted as
-    stellar rotation).
-  \item Rotation is usually rapid ($P<1\,{\rm day}$).
+    superposed on large-amplitude smooth variability. The latter
+    is interpreted as stellar rotation.
+  \item Rotation is usually between 0.5 to 2.0 days.
   \item Rarely any infrared excess (never any W4 detection; only a few
@@
     RIK-210).
-  \item Dip depths, durations, and phasescan vary over timescales of days
+  \item Dip depths, durations, and phases can vary over just a few cycles
     ({\it e.g.}, EPIC 204143627).
-  \item Dip depths can change after flares ({\it e.g.}, RIK-210.
-  \item Rare ($\lesssim 1\%$ of relevant stars) at a population level
+  \item Dip depths can change after flares.
+  \item Rare at a population level: $\lesssim 1\%$ of relevant stars
     \citep{rebull_usco_2018}.
@@
 
+A crucial additional point concerning the transient flux dips is
+that their dip durations seem to scale linearly with
+their periods \citep[][Figure~26]{stauffer_orbiting_2017}.
+In the usual idealized limits, the transit duration of a point source
+across the stellar disk scales as $T_0 \propto R_\star (P/M_\star)^{1/3}$
+\citep{winn_exoplanet_2010}.
+While the $\approx$0.5-day transient flux dip stars have dip
+durations consistent with point sources, at longer periods of
+1 to 5 days, the dip durations become many hours, and far
+too long for a point-source origin.
+
 \section{Physical Interpretation}
 
-Given the evidence, we believe that PTFO$\,$8-8695 is a binary M dwarf
-in which one star shows the ``long'' rotation signal, and the other is
-showing ``transient dipping'' that has also been observed in other
-young M dwarfs.
-
-Many other young M-dwarf photometric binaries have been observed in
-{\it e.g.,} Upper Sco and the Pleiades to show multiple periods
-\citep{rebull_usco_2018,stauffer_rotevol_2018}.  Typically, the
-periods are both rotational modulation.
-
-However, occasionally one or both periods can be ``scallop-shell''
-variability, {\it e.g.}, EPIC 203956650 in $\rho$Oph
-\citep{rebull_usco_2018}.
-
-There are
-
-The gas in the disks is gone (CITE).
-The upper limits on the SED from \citet{yu_tests_2015} imply X, Y, Z.
-The stars are therefore presumably no longer ``magnetically locked''
-to their disks.
-This is consistent with the $\approx$half-day periodicities of both
-signals.
-In the broader context of rapidly rotating young M-dwarfs, ``magnetic
-locking'' causes classical disked T-Tauri stars tend to rotate more
-{\it slowly}, and almost none have periods less than two days
-\citep[{\it e.g.},][]{rebull_rotation_2020}.
-
-The main physical question is what is causing the ``transient
-dipping''. This is an unsolved problem not only for \ptfo\ but also
-for an entire class of young rapidly rotating M-dwarfs
+The evidence for binarity in \ptfo\ is as follows.  First, the star is
+photometrically twice as bright as stars of the same color in its
+kinematic group (Figure~\ref{fig:gaia}.  Second, it shows two distinct
+photometric signals.  These two points alone are typically sufficient
+to imply binarity \citep{stauffer_rotevol_2018}.  For the case of
+\ptfo, there is a third line of evidence: the Gaia astrometry shows a
+poor goodness-of-fit to a single-source model.  While this could be
+caused by an additional companion or stellar variability, cluster
+members that are equally if not more variable still show \ptfo\ to be
+an outlier.  Therefore the astrometric excess is a suggestive third
+line of evidence for binarity in \ptfo.  We take these three points
+combined to imply that \ptfo\ is probably a binary.
+
+% rumination: is the lack of an infrared excess even significant? In
+% USco, we know late-M dwarfs show them. What about Orion? Distance is
+% 4x, at least...
+% Hernandez et al 2007a and 2007b show that there are some infrared
+% excesses seen with Spitzer. But their completeness for fainter M
+% dwarfs may be worse -- they see a disk fraction peaking at M0,
+% instead of later as seen by e.g., Barenfeld+2016.
+Based on the lack of an infrared excess seen by \citet{yu_tests_2015},
+the primordial gas disks seem to be have been depleted around both
+stars, and so the stars are no longer magnetically locked to their
+disks.  This is consistent with the $\approx$half-day periodicities of
+both signals: young disked M dwarfs typically rotate with periods of
+two days or more due to magnetic locking \citep[{\it
+e.g.},][]{rebull_rotation_2020}.  If the two stars are within
+$\approx$50$\,$AU of each other, as required by the NIRC2 adaptive
+optics imaging, then it would also be expected that the stars would
+truncate the outer edges of their respective disks, in a manner seen
+at the population level in exoplanetary systems
+\citep{kraus_impact_2016,moe_impact_2019}.
+
+The main physical question remaining is clearly what causes the
+transient dipping. We have argued that this is an unsolved problem not
+only for \ptfo\ but also for an emerging class of similar young
+rapidly rotating M-dwarfs.
+
+The explanation offered by \citet{stauffer_orbiting_2017} for this
+variability class is that warm clouds of gas and dust orbit the star
+at the Keplerian corotation radius.
+
+Why???
+
+What are the forces at play?
+
+FIXME
+
+Given the evidence for binarity, we believe that PTFO$\,$8-8695 is a
+binary M dwarf in which one star shows the 11.96~hour rotation signal,
+and the other shows a combination of transient dipping and stellar
+rotation, similar to that observed in other young M dwarfs.  The
+combined transient dipping and stellar rotation at the 10.74~hour
+cycle was only detected with the TESS data.
 

### 2020-04-22 paper/ms.tex
added words

@@
 had a smaller disk fraction than younger nearby sub-associations
-\citep{hernandez_spitzer_2007}.
+\citep{hernandez_spitzer_ob1_2007}.
 
@@
     has a less angular dip, and could also be an eclipsing binary.
-    All stars have similar properties as \ptfo, but are in Upper Sco.
+    All stars have similar properties as \ptfo, but are in Upper~Sco.
     We found these objects through studies by
@@
 
-When physical explanations are not immediately clear, taxonomy can be
-a useful exercise.  To our knowledge, of order 10-20 lightcurves with
+When physical explanations are not forthcoming, taxonomy can be a
+useful exercise.  To our knowledge, of order 10-20 lightcurves with
 similar morphologies have been seen in surveys of low-mass weak-lined
-T Tauri stars in regions including $\rho$ Oph, Upper Sco, Taurus, and perhaps
-the Pleiades
+T Tauri stars in regions including $\rho$ Oph, Upper~Sco, Taurus, and
+perhaps the Pleiades
 \citep{rebull_rotation_2016,david_transient_2017,stauffer_orbiting_2017,stauffer_rotevol_2018,rebull_usco_2018,rebull_rotation_2020}.
@@
     is interpreted as stellar rotation.
-  \item Rotation is usually between 0.5 to 2.0 days.
+  \item Rotation is fast---usually between 0.5 and 2.0 days.
   \item Rarely any infrared excess (never any W4 detection; only a few
@@
 \end{enumerate}
-The 10.74 hour signal in \ptfo\ meets essentially all of these
-criteria.
-
-A crucial additional point concerning the transient flux dips is
-that their dip durations seem to scale linearly with
-their periods \citep[][Figure~26]{stauffer_orbiting_2017}.
-In the usual idealized limits, the transit duration of a point source
-across the stellar disk scales as $T_0 \propto R_\star (P/M_\star)^{1/3}$
-\citep{winn_exoplanet_2010}.
-While the $\approx$0.5-day transient flux dip stars have dip
-durations consistent with point sources, at longer periods of
-1 to 5 days, the dip durations become many hours, and far
-too long for a point-source origin.
+The 10.74 hour signal in \ptfo\ meets all of these criteria.
+This is the first connection of \ptfo\ with this class of objects
+because the TESS lightcurve was needed to resolve the different
+rotation signals.
+
+There are two crucial additional points concerning the transient flux
+dips.  First, the dip durations seem to scale linearly with their
+periods \citep[][Figure~26]{stauffer_orbiting_2017}.  In the usual
+idealized limits, the transit duration $T$ of a point source across
+the stellar disk scales as $T \propto R_\star (P/M_\star)^{1/3}$
+\citep{winn_exoplanet_2010}.  While the shortest period
+$\approx$0.5-day transient flux dip stars have dip durations
+consistent with point sources, at longer periods of 1 to 5 days the
+dip durations become many hours, and far too long for a point-source
+origin.
+
+Second, between 40-50\% of the transient flux dip stars discovered in
+$\rho$~Oph and Upper~Sco show two
+Lomb-Scargle periods, and so are apparently binaries
+\citep[][Table~1]{stauffer_orbiting_2017}.
+This is higher than the corresponding main-sequence companion
+fraction of ${\rm CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$
+\citep{henry_solar_2006,duchene_stellar_2013}.
+Low-mass pre-main-sequence stars however have been shown to 
+companion fractions up to twice as high in dispersed clusters such as
+Upper~Sco and Taurus \citep{kraus_mapping_2008,kraus_mapping_2011}.
+A detailed high-resolution imaging survey would be necessary to determine
+whether the transient flux dip stars truly have any distinct
+population-level binarity properties relative to other young low-mass
+stars.
 
@@
 kinematic group (Figure~\ref{fig:gaia}.  Second, it shows two distinct
-photometric signals.  These two points alone are typically sufficient
-to imply binarity \citep{stauffer_rotevol_2018}.  For the case of
-\ptfo, there is a third line of evidence: the Gaia astrometry shows a
-poor goodness-of-fit to a single-source model.  While this could be
-caused by an additional companion or stellar variability, cluster
-members that are equally if not more variable still show \ptfo\ to be
-an outlier.  Therefore the astrometric excess is a suggestive third
-line of evidence for binarity in \ptfo.  We take these three points
+photometric signals.  These points alone suggest binarity
+\citep{stauffer_rotevol_2018}.  For the case of \ptfo, there is a
+third line of evidence: the Gaia astrometry shows a poor
+goodness-of-fit to a single-source model.  While this could be caused
+by an additional companion or stellar variability, cluster members
+that are equally if not more variable still show \ptfo\ to be an
+outlier.  Therefore the astrometric excess is a suggestive third line
+of evidence for binarity in \ptfo.  We take these three points
 combined to imply that \ptfo\ is probably a binary.
 
-% rumination: is the lack of an infrared excess even significant? In
-% USco, we know late-M dwarfs show them. What about Orion? Distance is
-% 4x, at least...
-% Hernandez et al 2007a and 2007b show that there are some infrared
-% excesses seen with Spitzer. But their completeness for fainter M
-% dwarfs may be worse -- they see a disk fraction peaking at M0,
-% instead of later as seen by e.g., Barenfeld+2016.
 Based on the lack of an infrared excess seen by \citet{yu_tests_2015},
-the primordial gas disks seem to be have been depleted around both
-stars, and so the stars are no longer magnetically locked to their
-disks.  This is consistent with the $\approx$half-day periodicities of
-both signals: young disked M dwarfs typically rotate with periods of
-two days or more due to magnetic locking \citep[{\it
-e.g.},][]{rebull_rotation_2020}.  If the two stars are within
-$\approx$50$\,$AU of each other, as required by the NIRC2 adaptive
-optics imaging, then it would also be expected that the stars would
-truncate the outer edges of their respective disks, in a manner seen
-at the population level in exoplanetary systems
-\citep{kraus_impact_2016,moe_impact_2019}.
+the primordial gas disks seem to be have been depleted\footnote{A
+potential confounding factor is completeness. Is Spitzer sensitive to
+infrared excesses at the distance of Orion? The series of studies by
+\cite{hernandez_spitzer_2006,hernandez_spitzer_ob1_2007,hernandez_spitzer_sig_2007,hernandez_spitzer_2009}
+show that it is.} around both stars in \ptfo. The stars are therefore
+no longer magnetically locked to their disks.  This is consistent with
+the $\approx$half-day periodicities of both rotation signals: young
+disked M dwarfs typically rotate with periods of two days or more due
+to magnetic locking \citep[{\it e.g.},][]{rebull_rotation_2020}.  If
+the two stars are within $\approx$50$\,$AU of each other, as required
+by the NIRC2 adaptive optics imaging, then it would also be expected
+that the stars would truncate the outer edges of their respective
+disks, in a manner seen at the population level in exoplanetary
+systems \citep{kraus_impact_2016,moe_impact_2019}.  This hard outer
+boundary condition could propagate to the inner disk and affect its
+evolution.
 
 The main physical question remaining is clearly what causes the
-transient dipping. We have argued that this is an unsolved problem not
-only for \ptfo\ but also for an emerging class of similar young
-rapidly rotating M-dwarfs.
-
-The explanation offered by \citet{stauffer_orbiting_2017} for this
-variability class is that warm clouds of gas and dust orbit the star
-at the Keplerian corotation radius.
-
-Why???
-
-What are the forces at play?
-
-FIXME
-
-Given the evidence for binarity, we believe that PTFO$\,$8-8695 is a
-binary M dwarf in which one star shows the 11.96~hour rotation signal,
-and the other shows a combination of transient dipping and stellar
-rotation, similar to that observed in other young M dwarfs.  The
-combined transient dipping and stellar rotation at the 10.74~hour
-cycle was only detected with the TESS data.
-
-The roughly half-day periods of both signals imply that
+transient dipping. This is an unsolved problem not only for \ptfo\ but
+also for an emerging class of similar young rapidly rotating M-dwarfs.
+Some possible explanations are highly disfavored, for reasons that
+have been detailed by \citet{david_transient_2017},
+\citet{stauffer_orbiting_2017}, and \citet{zhan_complex_2019}.
+These disfavored explanations include that the dips are caused by
+{\it (i)} eclipsing binaries;
+{\it (ii)} ``dipper''-flavor Class-I or Class-II disks;
+{\it (iii)} eclipses of prominences;
+{\it (iv)} high-latitude accretion hotspots;
+{\it (v)} high-latitude starspots;
+and
+{\it (vi)} dust clouds of reasonable composition.
+We also view the possibility of {\it (vii)} tidally disrupted planetary
+or cometary material to be implausible, given the synchronicity
+between dip and rotation periods, and the stability of signals
+observed in some systems.
+
+The explanations that are not yet ruled out include
+{\it (i)} transiting clumps of gas at the Keplerian corotation radius;
+{\it (ii)} transits of enshrouded protoplanets;
+{\it (iii)} occultations of starspots by an optically thick disk.
+The first and last explanations offer added appeal because they are
+flexible enough to explain not only the transient and persistent-dip
+M-dwarfs, but also the ``scallop shell'' M-dwarfs
+\citep{stauffer_orbiting_2017}.
+Despite this appeal, the possibility of distinct mechanisms 
+explaining these
+distinct variability classes should remain open.
 
@@
 
-Similarly, if the planet truly orbits every 10.74 hours, while the
-star's equator spins every 11.96 hours, the situation is clearly
-Darwin unstable.
-
 Given the available evidence,
@@
 ``transient dipping'' phenomenology observed in many young M dwarfs.
-  It seems rather unlikely to be a planet.
+Though it seems rather unlikely to be a planet, understanding \ptfo\
+and its analogs will help us understand the natal environments 
+of the majority of Earth-sized planets in the Milky Way
+\citep{dressing_occurrence_2013}.
 

### 2020-04-23 billy/plotting.py
splitsignal: fix annoying gap line

@@ -37,7 +37,8 @@ from billy.convenience import get_clean_ptfo_data
 from billy.models import linear_model
 
 from astrobase.lcmath import (
-    phase_magseries, phase_bin_magseries, sigclip_magseries
+    phase_magseries, phase_bin_magseries, sigclip_magseries,
+    find_lc_timegroups
 )
 from astrobase import periodbase
 
@@ -152,14 +153,23 @@ def plot_splitsignal_map(m, outpath, part='i'):
     things at rotation frequency
     things at orbital frequency
     """
+    ngroups, groupinds = find_lc_timegroups(m.x_obs, mingap=1.0)
+    assert ngroups == 2
+    if part == 'i':
+        g = groupinds[0]
+    elif part == 'ii':
+        g = groupinds[1]
+    else:
+        raise NotImplementedError
 
     plt.close('all')
     # 8.5x11 is letter paper. x10 allows space for caption.
     fig, axs = plt.subplots(nrows=4, figsize=(8.5, 10), sharex=True)
 
     axs[0].set_ylabel('f', fontsize='x-large')
-    axs[0].plot(m.x_obs, m.y_obs, ".k", ms=4, label="data", zorder=2, rasterized=True)
-    axs[0].plot(m.x_obs, m.map_estimate['mu_model'], lw=0.5, label='MAP',
+    axs[0].plot(m.x_obs[g], m.y_obs[g], ".k", ms=4, label="data", zorder=2,
+                rasterized=True)
+    axs[0].plot(m.x_obs[g], m.map_estimate['mu_model'][g], lw=0.5, label='MAP',
                 color='C0', alpha=1, zorder=1)
 
     y_tra = m.map_estimate['mu_transit']
@@ -180,21 +190,26 @@ def plot_splitsignal_map(m, outpath, part='i'):
             # axs[0].plot(m.x_obs, y_orb, lw=0.5, label='model '+f,
             #             color='C{}'.format(ix+1), alpha=1, zorder=ix+3)
 
-    #axs[1].set_ylabel('flux-orb (rot)')
-    axs[1].set_ylabel('$f_{{\mathrm{{\ell}}}} = f - f_{{\mathrm{{s}}}}$', fontsize='x-large')
[…]

### 2020-04-23 paper/ms.tex
astrometric residual: added the variability check

@@
 
-% \begin{figure*}[t]
-% 	\begin{center}
-% 		\leavevmode
-% 		\includegraphics[width=0.9\textwidth]{f6.png}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-% 	\caption{ {\bf Positions, kinematics, and HR diagram of \ptfo.}
-%   Members of the 25$\,$Ori-1 subgroup are shown with black circles,
-%   and were identified by \citet{kounkel_apogee2_2018} through
-%   clustering on Gaia DR2 and APOGEE data.  The ``neighborhood'' (gray
-%   circles) is defined as the group of at most $10^4$ randomly selected
-%   non-member stars within 5 standard deviations of the mean right
-%   ascension, declination, and parallax.
-% 	\label{fig:gaia}
-% 	}
-% \end{figure*}
-
 \subsection{Visual Binarity}
@@
     versus $Rp$ magnitude for 25$\,$Ori-1 members}.  The single-source
-    astrometric model provides a poor fit, which could be due to
-    stellar variability or binarity.
+    astrometric model for \ptfo\ provides a poor fit, which could be due to
+    stellar variability or binarity.  Cluster members that
+    are at least as variable as \ptfo\ show lower astrometric excesses
+    (black squares), suggesting binarity as the root cause.
 		\label{fig:gaia}
@@
 stars with comparable brightness in Gaia do not show such poor
-goodness-of-fit \citep[][Appendix A]{lindegren_gaiasoln_2018}.
+goodness-of-fit \citep[see][Appendix A]{lindegren_gaiasoln_2018}.
 
@@
 
-Using the same 149 members in the 25$\,$Ori-1 subgroup from
-\citet{kounkel_apogee2_2018}, we calculated the reduced $\chi^2$ for
-each member.  The lower panel of Figure~\ref{fig:gaia} shows the
-resulting values, as a function of stellar brightness.  \ptfo\ is in
+% NB: CDIPS-I only made 131 light-curves for 25Ori-1, because I wasn't
+% aware of the Kounkel+18 member list at the time.
+Using the same 149 members in the 25$\,$Ori-1 subgroup, we calculated
+the astrometric reduced $\chi^2$ for each member.  We then queried the
+CDIPS lightcurve database at MAST \citep{bouma_cluster_2019} to find
+the subset of stars that were at least as variable as \ptfo.
+We measured the variability amplitude by taking the difference of the
+$95^{\rm th}$ and $5^{\rm th}$ percentiles of the flux.
+This yielded 30 stars of equal or greater variability.
+The lower panel of Figure~\ref{fig:gaia} shows the
+reduced $\chi^2$ as a function of stellar brightness.  \ptfo\ is in
 the upper 90$^{\rm th}$ percentile of stars showing astrometric
@@
 excess is caused by stellar binarity or photometric variability.
-Nonetheless this result suggests that stellar binarity is indeed the
-root cause, because other member stars have identical variability
-characteristics but do not show such a large astrometric excess.
+Nonetheless the fact that comparably variable stars do not show
+comparably large astrometric excesses suggests that stellar binarity
+is indeed the root cause.
 
@@
 		numerical coincidence or an observational bias.  The orbital phase
-		observed by TESS is consistent with that of
+		observed by TESS (lower-right) is consistent with that of
 		\citet{tanimoto_evidence_2020}, and quite different from the

### 2020-04-23 billy/plotting.py

@@ -27,7 +27,7 @@ Convenience:
 """
 import os, corner, pickle
 from glob import glob
-import numpy as np, matplotlib.pyplot as plt
+import numpy as np, matplotlib.pyplot as plt, pandas as pd
 from datetime import datetime
 from pymc3.backends.tracetab import trace_to_dataframe
 from itertools import product
@@ -853,21 +853,37 @@ def plot_hr(outdir):
 
 def plot_astrometric_excess(outdir):
 
+    varamppath = '/Users/luke/Dropbox/proj/billy/data/25ori-1/var_amps.csv'
+    va_df = pd.read_csv(varamppath)
+
     pklpath = '/Users/luke/Dropbox/proj/billy/results/cluster_membership/nbhd_info_3222255959210123904.pkl'
     d = pickle.load(open(pklpath, 'rb'))
     g_df = d[2]
+
+    g_df = g_df.merge(va_df, on='source_id', how='left')
+
     chisq_red = g_df.astrometric_chi2_al / (g_df.astrometric_n_obs_al - 5)
+
+    ######
 
     f,ax = plt.subplots(figsize=(4,3))
 
-    ax.scatter(g_df.phot_rp_mean_mag, chisq_red, c='k', alpha=1., zorder=3, s=9,
-               rasterized=True, linewidths=0, label='K+18 members')
+    ax.scatter(g_df.phot_rp_mean_mag, chisq_red, c='gray', zorder=3, s=3,
+               marker='s', rasterized=True, linewidths=0, label='All members',
+               alpha=1)
 
-    sel = (g_df.source_id.astype(str) == '3222255959210123904')
+    ptfosel = (g_df.source_id.astype(str) == '3222255959210123904')
+    ptfo_amp = float(g_df[ptfosel].var_amp)
+    sel = (g_df.var_amp > ptfo_amp)
+
+    ax.scatter(g_df[sel].phot_rp_mean_mag, chisq_red[sel], c='k', alpha=1.,
+               zorder=4, s=12,
+               rasterized=True, linewidths=0, marker='s',
+               label='Amplitude > {:.1f}%'.format(ptfo_amp*100))
 
     ax.plot(
-        g_df[sel].phot_rp_mean_mag,
-        chisq_red[sel],
+        g_df[ptfosel].phot_rp_mean_mag,
+        chisq_red[ptfosel],
[…]

### 2020-04-23 drivers/abbreviate_the_bibliography.py
add abbreviate_the_bibliography

@@ -1,0 +1,34 @@
+"""
+Zotero generates long journal names. Abbreviate them in bulk to conform to the
+AAS style guide.
+"""
+
+from collections import OrderedDict
+
+mdict = OrderedDict({})
+
+mdict['The Astrophysical Journal Letters'] = r'\apjl'
+mdict['The Astrophysical Journal Supplement Series'] = r'\apjs'
+mdict['The Astronomical Journal'] = r'\aj'
+mdict['Annual Review of Astronomy and Astrophysics'] = r'\araa'
+mdict['The Astrophysical Journal'] = r'\apj'
+mdict['Publications of the Astronomical Society of the Pacific'] = r'\pasp'
+mdict['Astronomy and Astrophysics'] = r'\aap'
+mdict['Monthly Notices of the Royal Astronomical Society: Letters'] = r'\mnras:l'
+mdict['Monthly Notices of the Royal Astronomical Society'] = r'\mnras'
+mdict['Nature'] = r'\nat'
+
+with open('../paper/bibliography.bib', 'r') as f:
+    lines = f.readlines()
+
+for k,v in mdict.items():
+    for ix, l in enumerate(lines):
+        if k in l:
+            _newline = l.replace(k, v)
+            lines[ix] = _newline
+
+outpath = 'abbrevbib.bib'
+with open(outpath, 'w') as f:
+    f.writelines(lines)
+
+print('made {}'.format(outpath))

### 2020-04-23 paper/ms.tex
better conclusion

@@
 \usepackage[caption=false]{subfig} % for subfloat
+\usepackage{enumitem} % for specific spacing of enumerate
 
@@
   PTFO$\,$8-8695 is a photometric binary with respect to members of
-  its kinematic group.  The TESS lightcurve shows that the dominant
-  variability in this system is a sinusoid with a long period
-  $P_{\rm \ell}=11.96\,$hr, presumably caused by stellar rotation.
-  Also present is a complex signal, previously identified as the
-  planet candidate, that repeats with a short period $P_{\rm s}=
-  10.74\,$hr.  The two signals beat every 4.48 days.  Although there
-  is a dip in the short-period signal, ground-based photometry from
-  the past decade indicates that the orbital phase of the dip seems to
-  have instantaneously jumped, at least once, and perhaps twice.
-  Planets do not ``jump'' in orbital phase.  Given the available
-  evidence, we believe that PTFO$\,$8-8695 is a binary M dwarf in
-  which one star shows the long rotation signal, and the other star
-  shows ``transient dipping'' that has been seen in a few other young
-  weak-lined T Tauri stars.  The origin of these transient dips is
-  still undetermined, but our preferred explanation is eclipsing
-  clouds of gas or dust at the Keplerian co-rotation radius.
+  its kinematic group, and also suggest that it is an astrometric
+  binary.  The TESS lightcurve shows that the dominant variability in
+  this system is a sinusoid with a long period $P_{\rm
+  \ell}=11.96\,$hr, presumably caused by stellar rotation.  Also
+  present is a combined dip and asymmetric sinusoid, previously
+  identified as the planet candidate, that repeats with a short period
+  $P_{\rm s}= 10.74\,$hr.  The two signals beat every 4.48 days.
+  Although there is a dip in the short-period signal, ground-based
+  photometry from the past decade indicates that the orbital phase of
+  the dip seems to have instantaneously jumped, at least once, and
+  perhaps twice.  Planets do not ``jump'' in orbital phase.  Given the
+  available evidence, we believe that PTFO$\,$8-8695 is a binary M
+  dwarf in which one star shows the long rotation signal, and the
+  other star shows ``transient dipping'' that has been seen in a few
+  other young low-mass stars.  The origin of these transient dips is
+  still undetermined.
 \end{abstract}
@@
 Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
-hours, and is probably stellar rotation.  A short-period dip with an
-additional complex modulation repeats every 10.7 hours.  Analyzing the
+hours, and is probably stellar rotation.  A short-period dip superposed on an
+asymmetric sinusoid repeats every 10.7 hours.  Analyzing the
 Gaia data in Section~\ref{sec:gaia}, we show that relative to its
-kinematic group, \ptfo\ is a photometric binary.  We collect and
-discuss the puzzle pieces in Section~\ref{sec:discussion}.  The
-orbital phase of the dip seems to have instantaneously changed over
-the past decade.  In addition, a number of other young stars show
-lightcurve morphologies similar to the short-period signal.  We
-therefore argue that PTFO$\,$8-8695 is a binary M dwarf in which one
-star shows a rotation signal, and the other shows ``transient
-dipping'' caused by eclipses of material at the Keplerian co-rotation
-radius.  Section~\ref{sec:conclusions} summarizes our main points.
+kinematic group, \ptfo\ is a photometric binary, and suggest it also an
+astrometric binary.  We collect and discuss the puzzle pieces in
+Section~\ref{sec:discussion}.  The orbital phase of the dip seems to
+have instantaneously changed over the past decade.  In addition, a
+number of other young stars show lightcurve morphologies similar to
+the short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
+binary M dwarf in which one star shows a rotation signal, and the
+other shows ``transient dipping'' potentially caused by eclipses of plasma at
+the Keplerian co-rotation radius.  Section~\ref{sec:conclusions}
+summarizes our main points.
 
@@
 The majority of Gaia's derived parameters for \ptfo\ agreed with
-expectation from former studies
+expectations from former studies
 \citep{briceno_cida_2005,van_eyken_ptf_2012}.  One novelty however was
-that Gaia DR2 detected a significant ``astrometric excess'', at a
-level of 10.3$\sigma$.  We comment on the significance and
+that Gaia DR2 detected a 10.3$\sigma$ ``astrometric excess''.
+We comment on the significance and
 interpretation of this excess in Section~\ref{sec:gaia}.
@@
 diagnostic \citep{gelman_inference_1992}.  We tested our ability to
-successfully recover injected parameters using synthetic data, before
+successfully recover injected parameters using synthetic data before
 fitting the \ptfo\ lightcurves.
@@
 fit in our model as a transit, lasts $\approx$0.75 hours.  Superposed
-on the dip is a slightly asymmetric sinusoid with peak-to-peak
+on the dip is an asymmetric sinusoid with peak-to-peak
 amplitude of about 4\%. The asymmetric sinusoid peaks near phase 0.25,
@@
 
-The periodogram of the final residual (Figure~\ref{fig:splitsignal}
+The periodogram of the final residual (Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}
 bottom row) shows a weakly significant, poorly resolved peak at
@@
 CDIPS lightcurve database at MAST \citep{bouma_cluster_2019} to find
-the subset of stars that were at least as variable as \ptfo.
+the subset of members that were at least as variable as \ptfo.
 We measured the variability amplitude by taking the difference of the
@@
 \subsubsection{Similar Lightcurves}
+\label{subsec:dipstars}
 
@@
 The persistent and transient flux dip stars have the following properties.
-\begin{enumerate}
+\begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
   \item Weak-lined T Tauri stars.
@@
     cases with W3 detctions).
-  \item Can show multiple dips per cycle ({\it e.g.}, EPIC 203692610,
-    RIK-210).
+  \item Can show multiple dips per cycle.
+  %({\it e.g.}, EPIC 203692610, RIK-210).
   \item Dip depths, durations, and phases can vary over just a few cycles
@@
 
-\section{Physical Interpretation}
+\subsection{Physical Interpretation}
+\label{subsec:physical}
 
@@
 photometrically twice as bright as stars of the same color in its
-kinematic group (Figure~\ref{fig:gaia}.  Second, it shows two distinct
+kinematic group (Figure~\ref{fig:gaia}).  Second, it shows two distinct
 photometric signals.  These points alone suggest binarity
@@
 goodness-of-fit to a single-source model.  While this could be caused
-by an additional companion or stellar variability, cluster members
-that are equally if not more variable still show \ptfo\ to be an
+by stellar variability, cluster members
+that are at least as variable still show \ptfo\ to be an
 outlier.  Therefore the astrometric excess is a suggestive third line
@@
 also for an emerging class of similar young rapidly rotating M-dwarfs.
-Some possible explanations are highly disfavored, for reasons that
-have been detailed by \citet{david_transient_2017},
+Many possible explanations been detailed by \citet{rebull_rotation_2016}, \citet{david_transient_2017},
 \citet{stauffer_orbiting_2017}, and \citet{zhan_complex_2019}.
-These disfavored explanations include that the dips are caused by
+Their disfavored explanations include that the dips are caused by
 {\it (i)} eclipsing binaries;
@@
 {\it (v)} high-latitude starspots;
-and
-{\it (vi)} dust clouds of reasonable composition.
+or
+{\it (vi)} dust clouds of plausible composition.
 We also view the possibility of {\it (vii)} tidally disrupted planetary
 or cometary material to be implausible, given the synchronicity
-between dip and rotation periods, and the stability of signals
-observed in some systems.
+between dip and rotation periods seen across many systems.
 
@@
 \citep{stauffer_orbiting_2017}.
-Despite this appeal, the possibility of distinct mechanisms 
-explaining these
-distinct variability classes should remain open.
+Despite this appeal, the possibility of distinct mechanisms explaining
+these distinct variability classes remains open.
+
+The evolution of \ptfo\
+over the past decade (Figure~\ref{fig:o_minus_c}) could offer important
+hints.  Specifically, \ptfo's transition between having none, one, and
+multiple dips per cycle seems important.  It strains the ``enshrouded
+protoplanet'' interpretation, because there are no known processes
+that cause a planet's orbital phase to jump.  The dips would then need
+to be caused by material that was somehow disrupted from the planet,
+but somehow remained co-orbital for an extended duration. This seems
+implausible.
 
@@
 
-\ptfo\ was previously thought to potentially host a hot Jupiter.
-The TESS lightcurve of \ptfo\ showed a number of new features,
-many of which seem to disfavor the hot Jupiter interpretatation.
-The TESS data showed two key pieces of evidence.
-\begin{enumerate}
-  \item {\it Two periodic signals.} The ``long'' signal is a 10\%
-      peak-to-peak sinusoidal modulation repeating every 11.96 hours.
-      The ``short'' signal is a 4\% peak-to-peak complex modulation
-      repeating every 10.74 hours. It is composed of a dip, plus at
-      least two harmonics. The signals beat, and therefore cannot be
-      an artifact linked to data processing.
-  \item {\it A dip at the wrong orbital phase.} The clearest dip in
-    the ``short'' signal was consistent with recent observations by
-    \citet{tanimoto_evidence_2020}, and differed from the discovery
-    epoch by 5.14 hours.
-\end{enumerate}
-
-The physical mechanism responsible for all these features remains a
-matter of speculation.
-With that said, the TESS data support new arguments against the planetary
-interpretation of \ptfo.
-First, if the long signal is caused by starspot modulation, and the
-short signal by a transiting planet, what causes the additional
-complex modulations seen at the short, ``orbital'', period?
-
-Given the available evidence,
-PTFO$\,$8-8695 seems consistent with the
-``transient dipping'' phenomenology observed in many young M dwarfs.
-Though it seems rather unlikely to be a planet, understanding \ptfo\
-and its analogs will help us understand the natal environments 
-of the majority of Earth-sized planets in the Milky Way
+The combination of TESS and Gaia observations have provided new
+clarity in the \ptfo\ system.  Our main results are as follows.
+\begin{itemize}
+  \item {\it The TESS light-curve shows two periodic signals.} The
+    ``long'' signal is a 10\% peak-to-peak sinusoid that repeats every
+    11.96 hours.  The ``short'' signal is a 4\% peak-to-peak ``dip +
+    asymmetric sinusoid'' that repeats every 10.74 hours. The signals
+    beat, and therefore cannot be an artifact linked to data
+    processing.
+    Within the angular resolution of the Gaia source
+    catalog, both signals need to originate from \ptfo.
+  \item {\it The Gaia data imply binarity.} Relative to stars in its
+    kinematic group, \ptfo\ is a photometric binary
+    (Figure~\ref{fig:gaia}, top).  Relative to stars in its group that
+    are at least as photometrically variable, \ptfo\ also shows
+    signs of astrometric binarity (Figure~\ref{fig:gaia}, bottom).
+  \item {\it The orbital phase of the dip has changed since the
+    discovery by \citet{van_eyken_ptf_2012}.} As shown in
+    Figure~\ref{fig:o_minus_c}, the phase seems to have jumped,
+    perhaps twice. This agrees with the recent study by
+    \citet{tanimoto_evidence_2020}.
+  \item {\it All properties of \ptfo\ are consistent with the emerging
+    class of transient and persistent flux dip stars.} Analogous
+    lightcurves are shown in Figure~\ref{fig:brethren}.  Properties of
+    this variability class are enumerated in
+    Section~\ref{subsec:dipstars}.
+\end{itemize}
+
+The physical mechanism that explains the transient and persistent flux
+dips is unresolved. Our preferred explanations include
+transiting clumps of gas at the Keplerian corotation radius, and
+occultations of starspots by an optically thick disk \citep[{\it
+e.g.},][]{stauffer_orbiting_2017,david_transient_2017,zhan_complex_2019}.
+The jumping orbital phase disfavors the enticing explanation of an
+enshrouded, transiting protoplanet.  Though \ptfob seems unlikely
+to be a planet, understanding it and its analogs is a worthy
+problem that promises to teach us about the natal environments of the
+majority of habitable-zone Earth-sized planets in the Milky Way
 \citep{dressing_occurrence_2013}.
@@
 
-% \acknowledgements
-% %
-% This paper includes data collected by the TESS mission, which are
-% publicly available from the Mikulski Archive for Space Telescopes
-% (MAST).
-% %
-% Funding for the TESS mission is provided by NASA's Science Mission
-% directorate.
-% %
-% This work made use of NASA's Astrophysics Data System Bibliographic
-% Services.
+\acknowledgements
+%
+%This paper includes data collected by the TESS mission, which are
+%publicly available from the Mikulski Archive for Space Telescopes
+%(MAST).
+%
+%Funding for the TESS mission is provided by NASA's Science Mission
+%directorate.
+%
+\ptfo\ was included on the TESS ``short-cadence'' target list thanks to the
+Guest Investigator programs of S.\ Czesla and C.\ Huang (G011128 and
+G011132 respectively).
+%
 % %
@@
   \texttt{astrobase} \citep{bhatti_astrobase_2018},
-  % \texttt{astroplan} \citep{astroplan2018},
+  %\texttt{astroplan} \citep{astroplan2018},
   \texttt{astropy} \citep{astropy_2018},
   \texttt{astroquery} \citep{astroquery_2018},
-  % \texttt{BATMAN} \citep{kreidberg_batman_2015},
+  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
+  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019}
   \texttt{corner} \citep{corner_2016},
@@
   \texttt{pandas} \citep{mckinney-proc-scipy-2010},
+  \texttt{pyGAM} \citep{serven_pygam_2018_1476122},
   \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
-  \texttt{radvel} \citep{fulton_radvel_2018},
-  % \texttt{scikit-learn} \citep{scikit-learn},
+  %\texttt{radvel} \citep{fulton_radvel_2018},
+  %\texttt{scikit-learn} \citep{scikit-learn},
   \texttt{scipy} \citep{jones_scipy_2001}.
+  \texttt{tesscut} \citep{brasseur_astrocut_2019},
+  \texttt{wotan} \citep{hippke_wotan_2019}.
 }
 
-% \facilities{
-% 	{\it Astrometry}:
-% 	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}.
-% 	{\it Imaging}:
-% 	Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
-% 	{\it Spectroscopy}:
-% 	Keck:I~(HIRES; \citealt{vogt_hires_1994}),
+\facilities{
+ 	{\it Astrometry}:
+ 	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}.
+ 	{\it Imaging}:
+ 	Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2})
+ 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
+ 	{\it Spectroscopy}:
+ 	Keck:I~(HIRES; \citealt{vogt_hires_1994}),
 % 	Euler1.2m~(CORALIE),
 % 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
-% 	{\it Photometry}:
+ 	{\it Photometry}:
 % 	CTIO:1.0m (Y4KCam),
@@
 % 	SOAR (SOI),
-% 	TESS \citep{ricker_transiting_2015},
+ 	TESS \citep{ricker_transiting_2015}.
 % 	TRAPPIST \citep{jehin_trappist_2011},
 % 	VLT:Antu (FORS2).
-% }
+}
 
@@
 
-\clearpage
+% \clearpage
 \input{model_comparison_table.tex}
 \input{model_posterior_table.tex}
-\clearpage
+% \clearpage
 

### 2020-04-23 paper/ms.tex
improve text

@@
   PTFO$\,$8-8695 is a photometric binary with respect to members of
-  its kinematic group, and also suggest that it is an astrometric
-  binary.  The TESS lightcurve shows that the dominant variability in
-  this system is a sinusoid with a long period $P_{\rm
+  its kinematic group, and independently suggest that it is an
+  astrometric binary.  The TESS lightcurve shows that the dominant
+  variability in this system is sinusoidal with a long period $P_{\rm
   \ell}=11.96\,$hr, presumably caused by stellar rotation.  Also
@@
   dwarf in which one star shows the long rotation signal, and the
-  other star shows ``transient dipping'' that has been seen in a few
-  other young low-mass stars.  The origin of these transient dips is
-  still undetermined.
+  other star shows transient dipping that has been seen in a few other
+  young low-mass stars.  The origin of these transient dips is still
+  undetermined.
 \end{abstract}
@@
 the planet does not seem to emit infrared radiation in occulation, at
-least at the expected amplitude \citep{yu_tests_2015}.  In
-addition, despite measurement attempts by multiple investigators, it
-does not seem to show the Rossiter effect at the amplitude expected
-given the rapid stellar rotation and large planet size
+least at the expected amplitude \citep{yu_tests_2015}.  In addition,
+despite measurement attempts by multiple investigators, \ptfob\ does
+not seem to show the Rossiter effect at the amplitude expected given
+the rapid stellar rotation and large planet size
 \citep{yu_tests_2015,ciardi_followup_2015}.  Finally, detailed
@@
 alternative explanations do as well.  High-latitude accretion hotspots
-might produce the observed H$\alpha$ variability, and require
-fine-tuning produce dips of the approach duration. Furthermore, \ptfo\
-does not have an infrared (IR) excess associated with the presence of
-an inner disk \citep[{\it e.g.},][Figure~18]{yu_tests_2015}.
-Low-latitude starspots, hot or cold, struggle to produce the necessary
-dip durations.
-
-Another potential explanation is that between 0.1\% and 1\% of rapidly
+might produce the observed H$\alpha$ variability, but require
+fine-tuning to produce dips of the appropriate duration. Furthermore,
+\ptfo\ does not have an infrared (IR) excess associated with the
+presence of dust at the inner edge of a primordial disk \citep[{\it
+e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude starspots, hot or
+cold, struggle to produce the necessary dip durations.
+
+One potentially relevant fact is that between 0.1\% and 1\% of rapidly
 rotating low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations
@@
 and \citet{david_transient_2017} to explain this novel class of
-variable stars is that a circumstellar cloud of dust or gas might be
-orbiting near the Keplerian co-rotation radius.  Given available
-ground-based data, this explanation is not clearly viable for \ptfo,
-because a clear determination of the stellar rotation period has not
-yet been possible
+variable stars is that a circumstellar cloud of gas might be orbiting
+near the Keplerian co-rotation radius.  To date, this explanation has
+not clearly been viable for \ptfo, because a clear determination of
+the stellar rotation period has not yet been possible
 \citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.
@@
 Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
-hours, and is probably stellar rotation.  A short-period dip superposed on an
-asymmetric sinusoid repeats every 10.7 hours.  Analyzing the
-Gaia data in Section~\ref{sec:gaia}, we show that relative to its
-kinematic group, \ptfo\ is a photometric binary, and suggest it also an
-astrometric binary.  We collect and discuss the puzzle pieces in
-Section~\ref{sec:discussion}.  The orbital phase of the dip seems to
-have instantaneously changed over the past decade.  In addition, a
-number of other young stars show lightcurve morphologies similar to
-the short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
+hours, and is probably stellar rotation.  A short-period dip
+superposed on an asymmetric sinusoid repeats every 10.7 hours.
+Analyzing the Gaia data in Section~\ref{sec:gaia}, we show that
+relative to its kinematic group, \ptfo\ is a photometric binary, and
+suggest it also an astrometric binary.  We discuss the puzzle pieces
+in Section~\ref{sec:discussion}.  The orbital phase of the dip seems
+to have instantaneously changed over the past decade.  In addition,
+other young stars show lightcurve morphologies similar to the
+short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
 binary M dwarf in which one star shows a rotation signal, and the
-other shows ``transient dipping'' potentially caused by eclipses of plasma at
-the Keplerian co-rotation radius.  Section~\ref{sec:conclusions}
-summarizes our main points.
+other shows transient dipping potentially caused by eclipses of
+plasma at the Keplerian co-rotation radius.
+Section~\ref{sec:conclusions} summarizes our main points.
 
@@
 the target \citep{smith_kepler_apertures_2017}.  Non-astrophysical
-variability was removed through the methods discussed by
-\citet{smith_kepler_PDC_2017}.
+variability was removed by fitting out trends common to many stars
+\citep{smith_kepler_PDC_2017}.
 
@@
 The two largest peaks in the Lomb-Scargle periodogram of the
-lightcurve were clearly separated at a ``short'' period $P_{\rm s}
+lightcurve were separated at a ``short'' period $P_{\rm s}
 \approx 0.448\,{\rm days}$ and a ``long'' period $P_{\rm \ell} \approx
@@
 However it quickly became clear that it would be beneficial to
-simultaneously model the signals separately, in order to preserve the
+simultaneously model each signal, in order to preserve the
 power at each frequency.
@@
 \end{equation}
-where $f_{\rm s}$ is the relative flux at the short period, and
+where $f_{\rm s}$ is the flux at the short period, and
 $f_{\rm \ell}$ is the flux at the long period.  Writing out the
@@
     members of 25$\,$Ori-1.} Members of the 25$\,$Ori-1 group (black
-    circles) were identified by \citet{kounkel_apogee2_2018} through
-    clustering on six-dimensional Gaia DR2 and APOGEE-2 data.  The
+    circles) were identified by \citet{kounkel_apogee2_2018}.  The
     ``neighborhood'' (gray circles) is non-member stars within 5
@@
     blue, $Rp$ Gaia red, and $\omega_{\rm as}$ the parallax in
-    arcseconds.  The $x$-axis limits have been set to show only K and
-    M dwarf members, to accentuate \ptfo's separation from the
+    arcseconds.  The $x$-axis limits are set to show only K and
+    M dwarfs, to accentuate \ptfo's separation from the
     single-star sequence.  {\it Bottom: Astrometric goodness-of-fit
@@
     are at least as variable as \ptfo\ show lower astrometric excesses
-    (black squares), suggesting binarity as the root cause.
+    (black squares), suggesting binarity as the probable root cause.
 		\label{fig:gaia}
@@
 
-\subsection{Long period sinusoid}
+\subsection{Long Period Sinusoid}
 
@@
 
-\subsection{Short period dip}
+\subsection{Short Period Dip}
 
@@
 		at phase 0 in mid-2019) are consistent with noise, and were likely
-		reported because something was {\it expected}, rather than
-		convincingly {\it observed}.  Horizontal dashed lines are drawn at
+		reported because dips were expected, rather than
+		convincingly observed.  Horizontal dashed lines are drawn at
 		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting either a
@@
 		observed by TESS (lower-right) is consistent with that of
-		\citet{tanimoto_evidence_2020}, and quite different from the
-		original phase.
+		\citet{tanimoto_evidence_2020}.
 		\label{fig:o_minus_c}
@@
 
-\subsection{Short period out-of-dip modulation}
+\subsection{Short Period Out-of-Dip Modulation}
 
@@
 
-\subsubsection{Planetary interpretation}
+\subsubsection{Ellipsoidal Variability?}
 If there were a giant planet transiting \ptfo, it would tidally
@@
     their visual similarity to the short-period signal in \ptfo are as
-    shown.  They include EPIC 204143627, EPIC 204321142, EPIC
-    204270520, EPIC 205483258 (RIK-210), and EPIC 204787516.  RIK-210
-    has the longest period of any of these objects.  EPIC 204787516
-    has a less angular dip, and could also be an eclipsing binary.
-    All stars have similar properties as \ptfo, but are in Upper~Sco.
+    shown.
+    They include
+    EPIC 204143627,
+    EPIC 204270520,
+    EPIC 204321142,
+    EPIC 246938594,
+    and
+    EPIC 205483258 (RIK-210).
+    %and EPIC 204787516.
+    RIK-210 has the longest period of any of these objects. 
+    All the analogs displayed are either in Taurus or Upper~Sco, and
+    meet the characteristics of Section~\ref{subsec:dipstars}.
+    % EPIC 204787516 has a less angular dip, and could also be an
+    % eclipsing binary.
+    % All stars have similar properties as \ptfo, but are in Upper~Sco.
     We found these objects through studies by
@@
 the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
-The persistent and transient flux dip stars have the following properties.
+The persistent and transient flux dip stars all show angular dips
+that are cannot be explained with starspots.
+They typically have the following properties.
 \begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
@@
 \citep[][Table~1]{stauffer_orbiting_2017}.
-This is higher than the corresponding main-sequence companion
+This is higher than the main-sequence companion
 fraction of ${\rm CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$
@@
 
-The main physical question remaining is clearly what causes the
+The main physical question is what causes the
 transient dipping. This is an unsolved problem not only for \ptfo\ but
@@
 
-The combination of TESS and Gaia observations have provided new
+The combination of TESS and Gaia has provided new
 clarity in the \ptfo\ system.  Our main results are as follows.
@@
 e.g.},][]{stauffer_orbiting_2017,david_transient_2017,zhan_complex_2019}.
-The jumping orbital phase disfavors the enticing explanation of an
-enshrouded, transiting protoplanet.  Though \ptfob seems unlikely
+The jumping orbital phase disfavors the explanation of an
+enshrouded, transiting protoplanet.  Though \ptfob\ seems unlikely
 to be a planet, understanding it and its analogs is a worthy
@@
 G011132 respectively).
+%
+The Digitized Sky Survey was produced at the Space Telescope Science
+Institute under U.S. Government grant NAG W-2166.
+Figure~\ref{fig:scene} is based on photographic data obtained using
+the Oschin Schmidt Telescope on Palomar Mountain.
 %
@@
  	{\it Imaging}:
- 	Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2})
+  Second Generation Digitized Sky Survey,
+ 	Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
  	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
  	{\it Spectroscopy}:
- 	Keck:I~(HIRES; \citealt{vogt_hires_1994}),
+ 	Keck:I~(HIRES; \citealt{vogt_hires_1994}).
 % 	Euler1.2m~(CORALIE),

### 2020-04-23 billy/plotting.py

@@ -1045,6 +1045,7 @@ def plot_brethren(outdir):
         '204143627',
         '204270520',
         '204321142',
+        # '246969828', # maybe? also Taurus, double dips
         '246938594', # 'Taurus',
         '205483258' # RIK-210
     ]

### 2020-04-23 billy/plotting.py
minor

@@ -742,7 +742,7 @@ def plot_scene(c_obj, img_wcs, img, outpath, Tmag_cutoff=17, showcolorbar=0,
 
     ax0.text(3.2, 5, 'A', fontsize=16, color='C1', zorder=6, style='italic')
 
-    ax0.set_title('TESS (log)', fontsize='xx-large')
+    ax0.set_title('TESS', fontsize='xx-large')
 
     if showcolorbar:
         cb0 = fig.colorbar(cset0, ax=ax0, extend='neither', fraction=0.046, pad=0.04)
@@ -753,7 +753,7 @@ def plot_scene(c_obj, img_wcs, img, outpath, Tmag_cutoff=17, showcolorbar=0,
     cset1 = ax1.imshow(dss, origin='lower', cmap=plt.cm.gray_r)
 
     ax1.grid(ls='--', alpha=0.5)
-    ax1.set_title('DSS2 Red (linear)', fontsize='xx-large')
+    ax1.set_title('DSS2 Red', fontsize='xx-large')
     if showcolorbar:
         cb1 = fig.colorbar(cset1, ax=ax1, extend='neither', fraction=0.046,
                            pad=0.04)

### 2020-04-23 paper/ms.tex
update
important commit

@@
 \usepackage{enumitem} % for specific spacing of enumerate
+\usepackage{epigraph}
 
@@
 %PTFO$\,$8-8695b is Not a Planet}
-\title{Seeing Double: Transient Dips and Photometric Binarity in
-PTFO$\,$8-8695}
+%\title{Double Trouble: Transient Dips and Photometric Binarity in
+%PTFO$\,$8-8695}
+\title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
+PTFO 8-8695}
 
@@
 
+\epigraph{
+	Nubble [n]: a small lump or protuberance.
+}
+
 \section{Introduction}
+
 If \ptfob\ were a planet, it would be exceptional.  At below 10
@@
     {\it Bottom}: Short-period signal. The reference phase is set to
-    the ``planetary'' dip.  Gray points are the 10 minute cadence
+    the dip.  Gray points are the 10 minute cadence
     \texttt{PDCSAP} flux.  Black points are binned to 100 points per
@@
 %
+The authors thank D.~Fabrycky for helpful tidal evolution calculations
+that were eventually omitted.
+%
 \ptfo\ was included on the TESS ``short-cadence'' target list thanks to the

### 2020-04-26 paper/ms.tex
Update ms.tex
Having fun editing
Fun, fun, fun

@@
 PTFO 8-8695}
+\title{PTFO\,8-8695: A binary star, not a hot Jupiter host}
 
@@
 %
-\author[0000-0002-4265-047X]{J. N. Winn}
+\author[0000-0002-4265-047X]{J.\,N.\,Winn}
 \affiliation{ Department of Astrophysical Sciences, Princeton
@@
 \begin{abstract}
-  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10 million year
-  old Orion-OB1a cluster. We inspected data from TESS and Gaia to
-  clarify whether it is truly a planet.  The Gaia data show that
+  PTFO$\,$8-8695 is a star in the 7--10 million year
+  old Orion-OB1a cluster that shows brightness dips that resemble planetary transits.
+  While strong evidence against the planet hypothesis has already been provided,
+  the possibility remains debated in the literature.
+  To obtain further clues, we inspected data from the
+  NASA {\it Transiting Exoplanet Survey Satellite} (TESS) and the ESA Gaia mission.
+  The Gaia data show that
   PTFO$\,$8-8695 is a photometric binary with respect to members of
   its kinematic group, and independently suggest that it is an
-  astrometric binary.  The TESS lightcurve shows that the dominant
-  variability in this system is sinusoidal with a long period $P_{\rm
-  \ell}=11.96\,$hr, presumably caused by stellar rotation.  Also
-  present is a combined dip and asymmetric sinusoid, previously
-  identified as the planet candidate, that repeats with a short period
-  $P_{\rm s}= 10.74\,$hr.  The two signals beat every 4.48 days.
-  Although there is a dip in the short-period signal, ground-based
-  photometry from the past decade indicates that the orbital phase of
-  the dip seems to have instantaneously jumped, at least once, and
-  perhaps twice.  Planets do not ``jump'' in orbital phase.  Given the
-  available evidence, we believe that PTFO$\,$8-8695 is a binary M
-  dwarf in which one star shows the long rotation signal, and the
-  other star shows transient dipping that has been seen in a few other
-  young low-mass stars.  The origin of these transient dips is still
-  undetermined.
+  astrometric binary.  The TESS lightcurve shows two different photometric
+  periods. The variability is dominated by a sinusoidal signal with a period
+  of 11.96\,$hr, presumably caused by stellar rotation.  Also
+  present is a 10.74-day signal consisting of a not-quite sinusoid interrupted
+  by 1.2\% hour-long dips, the type of signal previously
+  intepreted as planetary transits.
+  The phase of the dips is nearly 180$^\circ$ away from the phase of the originally reported dips,
+  which, as noted previously, makes them difficult to explain as planetary transits.
+  Instead, we believe that PTFO$\,$8-8695 is a pair of young
+  and rapidly rotating M dwarfs, one of which shows the same ``transient-dipper''
+  behavior that been seen in at least 5 other cases.
+  The origin of these transient dips is still unknown but likely involves circumstellar material.
 \end{abstract}

### 2020-04-26 paper/ms.tex
minor
call an ice-cream truck
we've been scooped! not the end of the world, though.

--- 2020-04-23 paper/ms.tex
@@
 %PTFO$\,$8-8695}
-\title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
-PTFO 8-8695}
+%\title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
+%PTFO 8-8695}
+\title{Two Stars and Two Signals in PTFO 8-8695}
 
@@
 \author[0000-0002-4265-047X]{J. N. Winn}
+\affiliation{ Department of Astrophysical Sciences, Princeton
+University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+%
+\author[0000-0002-4265-047X]{...others!}
 \affiliation{ Department of Astrophysical Sciences, Princeton
@@
 
-\epigraph{
-	Nubble [n]: a small lump or protuberance.
-}
+%\epigraph{
+%	Nubble [n]: a small lump or protuberance.
+%}
 
@@
   \item Weak-lined T Tauri stars.
+    (Few to no spectroscopic accretion indicators). 
   \item Spectral type typically M2-M5 ({\it e.g.},
@@
   \item Ages $\lesssim$ 100$\,$Myr.
-  \item Stars typically lack spectroscopic accretion indicators 
   \item Phased lightcurves show shallow, angular dips, usually
@@
 \acknowledgements
+When this manuscript was at an advanced stage, we learned of the
+complementary work by \citet{koen_2020}, which was in press at the
+{\it Monthly Notices} before submission of our manuscript.  Our
+studies independently reached the key conclusions that the TESS
+lightcurve shows two periodic signals, and that the properties of
+\ptfo\ are consistent with the emerging class of transient and
+persistent flux dip stars.  \citet{koen_2020} reached these
+conclusions by modeling the TESS lightcurve as a truncated sum of
+Fourier terms, and concluded that the two signals are most simply
+interpreted as coming from two stars.  Our analysis of the Gaia data
+provide independent support for the conclusion that \ptfo\ is a
+binary.
+
+Minor differences in our approaches are that \citet{koen_2020}
+interprets the ``out of dip'' signal to potentially show multiple
+dips, while our BIC-based model comparison did not indicate the need
+for additional harmonics after fitting for the largest dip and two
+harmonics at $P_{\rm s}$ (Table~1). Our analysis emphasized the
+agreement between the TESS dip ephemeris and that from
+\citet{tanimoto_evidence_2020}.  \citet{koen_2020} however made a
+clearer case for testing the binarity hypothesis using interferometric
+imaging and high-precision radial velocities.
+\\
+\\
 %

### 2020-04-27 paper/ms.tex
Update ms.tex

--- 2020-04-26 paper/ms.tex
@@
 \begin{abstract}
-  PTFO$\,$8-8695 is a star in the 7--10 million year
+  PTFO$\,$8-8695 (CVSO\,30) is a star in the 7--10 million year
   old Orion-OB1a cluster that shows brightness dips that resemble planetary transits.
@@
 
-If \ptfob\ were a planet, it would be exceptional.  At below 10
-million years of age, it could be the youngest hot Jupiter known
-\citep{van_eyken_ptf_2012}.  Its orbital period of only 10.7 hours
-around a weak-lined T Tauri M dwarf would also give it the shortest
-period of any known hot Jupiter.  With such a short period, it would
-almost certainly have filled its Roche lobe, and would be actively
-losing mass to its host star.  Not only that, but the rapidly rotating
-host star would also be oblate and gravity darkened, and so the
-planet's orbit would likely precess into and out of transitability
+We wish \ptfob\ were a planet. It would be quite exceptional. It would
+be the youngest known hot Jupiter
+\citep{van_eyken_ptf_2012}, orbiting a T Tauri star in the Orion-OB1a cluster.
+It would have the shortest orbital period (10.7 hours) of any hot Jupiter.
+With such a short period, it would
+probably be filling its Roche lobe, and actively
+losing mass to its host star.  Not only that, but the rapidly-rotating
+host star is probably oblate enough to torque the planet's orbit
+into and out of the transiting configuration on a time scale of years
 \citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.
 
-Other lines of evidence would imply further planetary ``firsts'' for
-this planet candidate.  One first would be that its transits are about
-three times deeper in optical bandpasses ({\it e.g.,} $g$-band) than
-in the near-infrared ({\it e.g.}, $z$-band)
-\citep{onitsuka_multicolor_2017,tanimoto_evidence_2020}.  A cloud-free
-hydrogen-dominated planetary atmosphere cannot explain such a
-wavelength dependence.  The planet might therefore be surrounded by a
-dust cloud \citep{tanimoto_evidence_2020}.
-
-Another first could be the direct detection of H$\alpha$ emission from
-the planet itself \citep{johnskrull_h_2016}.  While the stellar
-chromosphere emits in H$\alpha$, it seems that there is an additional
-excess H$\alpha$ emission that could be in phase with the planetary
+Another first would be the direct detection of H$\alpha$ emission from
+the planet itself \citep{johnskrull_h_2016}.  In addition to the
+chromospheric H$\alpha$ emisson, it seems that there is an additional
+H$\alpha$ emission with radial velocity variations in phase with the planetary
 orbit.  The average velocity width of the excess H$\alpha$ emission is
@@
 stellar chromosphere \citep{johnskrull_h_2016}.  The proposed
-explanation is that outflowing mass from the planet could explain this
-excess emission as well \citep{johnskrull_h_2016}.
-
-There are perhaps a few challenges to the planetary interpretation (if
-these ``features'' are not already seen as such).  They include that
-the planet does not seem to emit infrared radiation in occulation, at
-least at the expected amplitude \citep{yu_tests_2015}.  In addition,
-despite measurement attempts by multiple investigators, \ptfob\ does
+explanation is that the emission is from hot material flowing away
+from the planet \citep{johnskrull_h_2016}.
+
+However, the observed signals have some peculiarities that
+make the planet seem even more unusual, to the point that they cast into doubt the premise
+that \ptfob\ is real.
+First, the transit-like brightness dips are about
+three times deeper in optical bandpasses ({\it e.g.,} $g$-band) than
+in the near-infrared ({\it e.g.}, $z$-band)
+\citep{onitsuka_multicolor_2017,tanimoto_evidence_2020}.  An ordinary
+atmosphere expected for a Jovian planet would not lead to such a strong
+color-dependence of the transits.  
+Second, the planet does not seem to emit as much infrared radiation
+as would be expected for such a hot Jovian planet \citep{yu_tests_2015}.
+Third, despite measurement attempts by multiple investigators, \ptfob\ does
 not seem to show the Rossiter effect at the amplitude expected given
 the rapid stellar rotation and large planet size
-\citep{yu_tests_2015,ciardi_followup_2015}.  Finally, detailed
-modelling of the ``precession + gravity darkening'' transits has shown
-that the necessary degree of gravity darkening is too great, given the
-spectroscopically observed equatorial velocity
-\citep{howarth_reappraisal_2016}.  Additionally, as the
-gravity-darkened star precessed about its rotation axis, it would need
-to show photometric variability that has not been observed.
+\citep{yu_tests_2015,ciardi_followup_2015}.
+Fourth, the phase of the dips within the overall period of photometric variability
+has changed drastically over the years since their initial discovery.
+To counter these objections, it has been proposed 
+that the planet may be much smaller than Jupiter and that the dips
+are produced by dust clouds emitted from the planet \citep{tanimoto_evidence_2020}.
+
+A separate issue is that the brightness dips
+change shape over many orbital cycles. This was initially explained by
+\cite{barnes_measurement_2013} as the natural effects of gravity darkening
+However, \cite{howarth_reappraisal_2016} argued
+that the necessary amplitude of gravity darkening is too large to be realistic, given the
+spectroscopically-determined rotation velocity.
+Additionally, as the
+gravity-darkened star precessed about its rotation axis, it would 
+show photometric variability that has not been observed.
 
 While the planetary interpetation clearly faces challenges,
-alternative explanations do as well.  High-latitude accretion hotspots
+there is no completely satisfactory alternate explanation.  High-latitude accretion hotspots
 might produce the observed H$\alpha$ variability, but require
@@
 e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude starspots, hot or
-cold, struggle to produce the necessary dip durations.
-
-One potentially relevant fact is that between 0.1\% and 1\% of rapidly
+cold, struggle to produce photometric features as short as some
+of the observed dips.
+
+A relevant fact is that between 0.1\% and 1\% of rapidly
 rotating low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations
-show narrow dips in phase with strong stellar rotation signals
+show short-duration dips as part of their overall periodic variability
 \citep{rebull_usco_2018}.  The dips can persist over months, but their
@@
 and \citet{david_transient_2017} to explain this novel class of
-variable stars is that a circumstellar cloud of gas might be orbiting
-near the Keplerian co-rotation radius.  To date, this explanation has
-not clearly been viable for \ptfo, because a clear determination of
-the stellar rotation period has not yet been possible
+variable stars is that a circumstellar cloud of gas is orbiting
+near the co-rotation radius.  To this point, though, it has not
+been clear if this explanation applies to
+\ptfo, because the determination of
+the stellar rotation period has been somewhat ambiguous
 \citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.
@@
 Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
-two distinct signals, which we extract and analyze in
-Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
-hours, and is probably stellar rotation.  A short-period dip
-superposed on an asymmetric sinusoid repeats every 10.7 hours.
-Analyzing the Gaia data in Section~\ref{sec:gaia}, we show that
-relative to its kinematic group, \ptfo\ is a photometric binary, and
-suggest it also an astrometric binary.  We discuss the puzzle pieces
-in Section~\ref{sec:discussion}.  The orbital phase of the dip seems
-to have instantaneously changed over the past decade.  In addition,
-other young stars show lightcurve morphologies similar to the
-short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
-binary M dwarf in which one star shows a rotation signal, and the
-other shows transient dipping potentially caused by eclipses of
-plasma at the Keplerian co-rotation radius.
-Section~\ref{sec:conclusions} summarizes our main points.
+two different periodic signals, which we analyze in
+Section~\ref{sec:tess}.
+The Gaia data, analyzed in Section~\ref{sec:gaia}, show that
+\ptfo\ is a photometric binary, and suggest it also an astrometric binary.
+We discuss the pieces of the puzzle
+in Section~\ref{sec:discussion}, and summarize the situation in
+Section~\ref{sec:conclusions}.
+% JNW: this kind of stuff is for the abstract/conclusion, not intro.
+%The phase of the dip within the seems
+%to have instantaneously changed over the past decade.  In addition,
+%other young stars show lightcurve morphologies similar to the
+%short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
+%binary M dwarf in which one star shows a rotation signal, and the
+%other shows transient dipping potentially caused by eclipses of
+%lasma at the Keplerian co-rotation radius.
 
@@
 \label{sec:observations}
+
+% JNW: How about more descriptive labels? f --> Total flux
+% fl = Longer-period signal
+% fs = Shorter-period signal
+% last panel, Residuals
 
@@
 		{\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 19).}
-    {\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux
-    versus time. The beat period of 4.48 days is visible by eye.  The
-    preferred model plotted underneath the data includes 2 harmonics
-    at the long period $P_{\rm \ell}$, plus 2 harmonics and a transit
-    at the short period $P_{\rm s}$.
-		{\it Upper middle}: Long-period signal, equal to the raw signal
-		minus the short-period signal.
-		{\it Lower middle}: Short-period signal, equal to the raw signal
-		minus the long-period signal.
+    {\it Top}: The ``raw'' (\texttt{PDCSAP} mean-subtracted) relative flux
+    data. The beat period of 4.48 days is visible by eye.  The
+    blue curve is a model including 2 harmonics
+    at the longer period $P_{\rm \ell}$, plus 2 harmonics and a transit
+    at the shorter period $P_{\rm s}$.
+		{\it Upper middle}: Longer-period signal, equal to the raw signal
+		minus the shorter-period signal.
+		{\it Lower middle}: Shorter-period signal, equal to the raw signal
+		minus the longer-period signal.
 		{\it Bottom}: residual.  The data are binned from 2 to 10 minute
@@
   \caption{ {\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 20).}
-		Panels are as in Figure~\ref{fig:splitsignal}.
+		Same format as Figure~\ref{fig:splitsignal}.
 		\label{fig:splitsignalii}
@@
 
-\ptfo\ was observed by TESS with Camera 1, CCD 1, from December 15,
-2018 to January 6, 2019, during the sixth sector of science operations
-\citep{ricker_transiting_2015}.  The star was designated TIC 264461976
+\ptfo\ [also known as CVSO\,30; \cite{briceno_cida_2005}}]
+was observed by TESS with Camera 1, CCD 1, from December 15,
+2018 until January 6, 2019, during the sixth sector of science operations
+\citep{ricker_transiting_2015}.  The star was designated TIC\,264461976
 in the TESS Input Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.
@@
 Science Processing Operations Center (SPOC) at NASA
-Ames~\citep{jenkins_tess_2016}.  Our main analysis used the resulting
+Ames~\citep{jenkins_tess_2016}.  We mainly used the
 Presearch Data Conditioning (PDC) lightcurve.  The PDC lightcurve
@@
 
-As an independent check on the shorter cadence SPOC light-curve, we
-separately processed the 30-minute image stacks as part of the Cluster
+As an independent check on the 2-minute SPOC lightcurve, we
+examined the lightcurve based upon 30-minute image stacks which was
+produced as part of the Cluster
 Difference Imaging Photometric Survey (CDIPS;
@@
 circular aperture with radius 1 pixel.
-
-To clean the data, we removed all points with non-zero quality flags
+% JNW: What does that mean? You used fractional pixels in calculating the pixel sums?
+
+To clean the data, we removed all points with non-zero quality flags,
+which indicate known problems
 \citep[{\it e.g.},][]{tess_data_product_description_2018}.  We also
-masked out the first and last 6 hours of each orbit, since there is
-often systematic red noise during those times.  Both the CDIPS and PDC
-lightcurves showed a clear discontinuous ``jump'' in the last few days
-of orbit 20, which seemed likely to be an instrumental systematic.  We
-correspondingly masked out times from BJD 2458488.3 until the end of
-the orbit.  The PDC lightcurve initially had 15{,}678 points.  The
+masked out the data from the first and last 6 hours of each orbit, since there are
+often systematic red noise in the photometry during those times.  Both the CDIPS and PDC
+lightcurves showed a clear discontinuous jump in the last few days
+of orbit 20, which seemed likely to be an instrumental systematic effect.  We
+correspondingly masked out the data with timestamps ranging from BJD\,2458488.3 until
+the end of the orbit.  The PDC lightcurve initially had 15{,}678 points.  The
 quality-flag cut removed 854 points; masking the orbit edges removed an
-additional 716; removing the final few days of orbit 20 removed an
-additional 1079.  After cleaning, 83\% of the initial flux
+additional 716 points; and removing the data from the final few days of orbit 20 removed an
+additional 1079 points.  After cleaning, 83\% of the initial flux
 measurements remained.
 
-We normalized these points by dividing out the median flux. We opted
-to then subtract by unity to simplify subsequent interpretation.  Many
+We normalized the lightcurve by dividing out the median flux, and then opted
+to subtract 1.0 to set the median value to zero, which simplified
+subsequent interpretation.  Many
 of these and subsequent processing steps were performed using
@@
 
-\subsubsection{Astrometry of \ptfo}
+\subsubsection{Astrometric measurements}
 
@@
 In the Gaia second data release (DR2), these CCD observations were
-used to estimate positions, proper motions, and parallax for the
+used to determine positions, proper motions, and parallaxes of the
 brighest 1{.}3 billion stars, including \ptfo\
-\citep{lindegren_gaiasoln_2018}.  There were 121 ``good'' observations
-of \ptfo, that is observations that were not strongly down-weighted in
-its astrometric solution.  \ptfo\ was assigned a Gaia DR2 identifier
-of 3222255959210123904.  It photometric brightness was measured using
+\citep{lindegren_gaiasoln_2018}.  For \ptfo, there were 121 ``good'' observations, i.e., observations that were not strongly down-weighted in
+the astrometric solution.  \ptfo\ was assigned the Gaia DR2 identifier
+3222255959210123904.  It photometric brightness was measured using
 selected bands ($G$, $Rp$, and $Bp$) of the Gaia Radial Velocity
@@
 expectations from former studies
-\citep{briceno_cida_2005,van_eyken_ptf_2012}.  One novelty however was
-that Gaia DR2 detected a 10.3$\sigma$ ``astrometric excess''.
+\citep{briceno_cida_2005,van_eyken_ptf_2012}.  The main novelty is
+that Gaia DR2 reported a 10.3$\sigma$ ``astrometric excess'', indicating
+that the residuals to the best-fitting astrometric model were larger
+than expected based on the statistical uncertainties.
 We comment on the significance and
@@
 the norm \citep{briceno_lowmassOB_2007}.  The Orion molecular cloud
-complex in particular has numerous subgroups, with ages spanning 0.5
+complex in particular has numerous subgroups, with ages ranging from 0.5
 to 15$\,$Myr. For an incomplete sampling, see for instance
@@
 \ptfo\ was initially identified as a member of the Orion$\,$OB1a
-sub-association by \citet{briceno_cida_2005} through combined
+sub-association and designated CVSO\,30 by \citet{briceno_cida_2005} through combined
 photometry and spectroscopy.  Later work by \citet{briceno_25_2007}
@@
 Orion~OB1a, named the ``25$\,$Ori'' group after its brightest
-member. The 25$\,$Ori group had an isochrone age of 7--10$\,$Myr, and
-had a smaller disk fraction than younger nearby sub-associations
+member. They reported that the 25$\,$Ori group has an isochrone age of 7--10$\,$Myr, and
+a smaller fraction of stars with disks than younger nearby sub-associations
 \citep{hernandez_spitzer_ob1_2007}.
@@
 With the Gaia astrometry, it has become clear that 25$\,$Ori itself
-has further subgroups
+has distinct subgroups
 \citep{kounkel_apogee2_2018,briceno_cidaII_2019}.  In describing the
@@
 \citep{gunn_sdss_2006,majewski_apache_2017,blanton_sloan_2017,zasowski_target_2017,cottle_apogee2_2018}.
-They performed a hierarchical clustering on the six dimensional
+They performed a hierarchical clustering on the six-dimensional
 position and velocity information to identify subgroups within the
@@
 \end{equation}
-where from set-notation, `$\subset$' denotes ``is a proper subset
+where `$\subset$' means ``is a proper subset
 of''.  25$\,$Ori-1 is the largest subgroup of 25$\,$Ori, with 149
-identified members.  Its mean age from its CMD was found to be
-6.9$\,$Myr, and from its HR diagram 8.5$\,$Myr
-\citep{kounkel_apogee2_2018}.  \citet{kounkel_apogee2_2018} identified
+identified members.  Based on the CMD, the mean age was found to be
+6.9$\,$Myr, while the HR diagram led to an estimated age of 8.5$\,$Myr
+\citep{kounkel_apogee2_2018}. 
+% JNW: What's the difference between a CMD and an HR diagram?
+\citet{kounkel_apogee2_2018} identified
 seven other smaller groups in the Orion complex near the Be star
@@
 
-These details concerning the group membership for one object may seem
-cumbersome to those accustomed to comparing ``young cluster members''
-with ``old field stars''.  Though all members of the Orion complex are
+These details concerning the group membership for one object are
+cumbersome to those accustomed to the simple distinction between ``young cluster members''
+and ``old field stars''.  Although all members of the Orion complex are
 indeed young relative to the field, these details are essential for
-assessing any evidence for photometric binarity in \ptfo, because of
+assessing the photometric evidence for the binarity of \ptfo, because of
 the degeneracy between stellar luminosity and age for
-pre-main-sequence stars.  Having a clean sample of tightly spatially
-and kinematically associated reference stars minimizes contamination
-not just from the field, but from older and younger members of the
-Orion complex itself.
+pre-main-sequence stars.  Having a clean sample of reference stars that are
+tightly associated with \ptfo\ --- both spatially
+and kinematically --- minimizes contamination
+not only from field stars, but also from older and younger members of the
+Orion complex.
 
@@
 	\vspace{-0.7cm}
-	\caption{ {\bf Phase-folded long and short-period signals.}
-    {\it Top}: Long-period signal, as in Figure~\ref{fig:splitsignal}.
-    {\it Bottom}: Short-period signal. The reference phase is set to
+	\caption{ {\bf Phase-folded longer and shorter-period signals.}
+    {\it Top}: Longer-period signal, as in Figure~\ref{fig:splitsignal}.
+    {\it Bottom}: Shorter-period signal. The reference phase is set to
     the dip.  Gray points are the 10 minute cadence
@@
 (Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}, top
-panel).
-
-As a precursor to more detailed analysis, we calculated generalized
+panel). As a precursor to more detailed analysis, we calculated generalized
 Lomb-Scargle periodograms using \texttt{astrobase}
 \citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
-The two largest peaks in the Lomb-Scargle periodogram of the
-lightcurve were separated at a ``short'' period $P_{\rm s}
-\approx 0.448\,{\rm days}$ and a ``long'' period $P_{\rm \ell} \approx
-0.499\,{\rm days}$.  The $P_{\rm \ell}$ peak had the greater power of
-the two.  Smaller harmonics surrounding each of these two dominants
-peaks were also present.
+The two tallest peaks occur at 0.448\,days and 0.499\,days.
+We will refer to these two periods as the ``shorter period'' $P_{\rm s}$
+and the ``longer period'' $P_{\rm ell}$.
+The peak at $P_{\rm \ell}$ is the taller of the two peaks.
+Lower-power harmonics of both signals are also present.
 
@@
 The peak-to-peak lightcurve amplitude at maximum, when the two signals
-constructively interfere, is about 14\%.  At minimum, the peak-to-peak
-amplitude is about 6\%.  Assuming the signals are just two sinusoids,
-algebra tells us that the peak-to-peak amplitudes should therefore be
-10\% for the long-period signal, and 4\% for the short-period signal.
-These order-of-magnitude numbers will turn out to be roughly correct.
-
-Initial signal-processing experiments fitting out splines or sinusoids
-showed that after subtracting out the long-period signal, the
-short-period signal dominated the periodogram, and vice-versa.
-However it quickly became clear that it would be beneficial to
-simultaneously model each signal, in order to preserve the
-power at each frequency.
+interfere constructively, is about 14\%.  During the times of destructive interference, the peak-to-peak
+amplitude is about 6\%.  Assuming the signals are mainly sinusoidal,
+simple algebra tells us that the peak-to-peak amplitudes should be
+about 10\% for the longer-period signal, and 4\% for the shorter-period signal.
+To view the phase-folded light curves of the longer-period signal,
+we subtracted the best-fitting sinusoid at the shorter period; the resulting
+light curve appears smooth and nearly sinusoidal.
+But after subtracting the best-fitting sinusoid at the longer period,
+visual inspection of the phase-folded light curve of the shorter-period signal
+revealed substructure resembling the ``dips'' seen in previous
+observations. In particular, there was a $\approx$1\% dip lasting about an hour.
+These initial impressions turned out to be consistent with the results of our
+more complicated analysis, described below.
+
+% JNW This doesn't seem necessary to me.
+%Initial signal-processing experiments fitting out splines or sinusoids
+%showed that after subtracting out the long-period signal, the
+%short-period signal dominated the periodogram, and vice-versa.
+%However it quickly became clear that it would be beneficial to
+%simultaneously model each signal, in order to preserve the
+%power at each frequency.
 
@@
 
-We fitted the lightcurve as a linear combination of Fourier harmonics
-at the short and long periods, plus a transit at the short period.
+We fitted a model to the lightcurve consisting of a linear combination of Fourier modes
+with periods $P_{\rm s}$ and $P_{\rm ell}$, as well as a number of harmonics chosen as described below. To try accounting for the dips, we also added an
+analytic transit model with period $P_{\rm s}$.
 Symbolically, the total flux $f$ is given as
@@
 \end{equation}
-where $f_{\rm s}$ is the flux at the short period, and
-$f_{\rm \ell}$ is the flux at the long period.  Writing out the
-Fourier terms,
+where $f_{\rm s}$ is the flux at the shorter period, and
+$f_{\rm \ell}$ is the flux at the longer period.  Writing out the
+Fourier terms explicitly,
 \begin{align}
@@
 \end{align}
-for $N$ and $M$ the total number of harmonics at the short and long
-periods, respectively, $A_i$ and $B_i$ the amplitudes for each
-harmonic term (potentially negative), and $\omega_i = 2\pi / P_i$ the
-angular frequency for $i$ the short or long period index.  We fixed
-the ``phase-offset'' for the short period signal to be zero, and let
-the reference time for the long period signal float by introducing
-$\phi_{\rm \ell}$.  Since we did not a priori know how many harmonics
-would be appropriate, we considered a number of different choices for
-$N$ and $M$, and used the Bayesian information criterion to choose the
-appropriate model (Table~1).
-
-As an example, one possible model could be a transit, plus $N=2$
-harmonics of sines and cosines at the short period, plus $M=1$
-harmonics at the long period.  For this case the free parameters would
-be as follows.  For the transit, we would fit for the impact
+where $N$ and $M$ are the total number of modes at the shorter and longer
+periods, respectively,
+% JNW: maybe N_\ell and N_s?
+$A_i$ and $B_i$ the amplitudes of each
+mode (which can be positive or negative), and $\omega_\ell$
+and $\omega_{\rm s}$ are the angular frequencies of the longer-period
+and shorter-period signals. By not including a phase parameter
+in the shorter-period model, we have implicitly defined the zero point
+of the phase scale. The relative phase of the longer-period model
+is specified by the phase parameter $\phi_\ell$.
+Since we did not know in advance how many harmonics
+would be appropriate to include in the model, we considered a number of different choices for
+$N$ and $M$, and used the Bayesian information criterion to select
+the final model (Table~1).
+
+As an example, one possible model consists of a transit, $N=2$
+sines and cosines at the shorter period, plus $M=1$
+sine and cosine at the longer period.  In this case, the free parameters would
+be as follows.  The transit model parameters are the impact
 parameter, the planet-to-star radius ratio, two quadratic limb
-darkening parameters, the planet orbital period (equal to the short
-period), the reference time for the transit, and the mean flux.  There
-would be $2N=4$ additional Fourier amplitudes at the short period,
-plus $2M=2$ Fourier amplitudes at the long period, and well as the
-long period itself and its phase.  For this case, we therefore fitted
-14 free parameters.
+darkening parameters, the planet's orbital period (set equal to $P_{\rm s}$)
+the time of a particular transit, and the mean flux.  There
+would be $2N=4$ additional Fourier amplitudes at the shorter period,
+plus $2M=2$ Fourier amplitudes at the longer period, and well as $P_\ell$
+itself and the relative phase $\phi_\ell$.  The total number
+of parameters is 14.
 
@@
 built on \texttt{theano}
-\citep{salvatier_2016_PyMC3,exoplanet:theano}.  For the Fourier terms,
-we used the default math operators.  For the exoplanet transit, we
-used the model and derivatives implemented in \texttt{exoplanet}
+\citep{salvatier_2016_PyMC3,exoplanet:theano}.
+For the Fourier terms,
+we used the default math operators.
+% JNW what does that mean? If you just mean 'numpy', is it necessary to state?
+For the exoplanet transit, we
+used the model and derivatives implemented in the \texttt{exoplanet} code
 \citep{exoplanet:exoplanet}.  Our priors are listed in Table~2.  To
-speed up the fitting, we re-sampled the cleaned 2 minute lightcurves
-to 10 minute binning.  We correspondingly scaled the uncertainties in
+speed up the fitting process, we averaged the 2-minute lightcurve
+to 10-minute samples.  We correspondingly scaled the uncertainties in
 the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
-initialized each model to the maximum a posteriori (MAP) solution.  We
-then sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
+initialized each model with the parameters of the maximum a posteriori (MAP) model.
+We then sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
 \citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
@@
 
-We considered nine models, with the number of harmonics per frequency
+We considered nine models, with the number of modes per frequency
 $N$ and $M$ ranging from one to three.  To select our preferred model,
 we used the Bayesian information criterion (Table~1).  The model with
-the lowest BIC had two harmonics at the short 10.74$\,$hr period, and
-two harmonics at the long 11.96$\,$hr period.  The next-best model had
+the lowest BIC had two modes at the shorter 10.74$\,$hr period, and
+two modes at the longer 11.96$\,$hr period.  The next-best model had
 an additional harmonic at the longer period ($M=3$), but was otherwise
-identical.  All nine models have reduced $\chi^2$ ranging between 1.37
+identical.  All nine models have reduced $\chi^2$ values ranging between 1.37
 and 1.51, which suggests a plausible though imperfect agreement
-between the data and fitting results.  The best-fit parameters for the
-lowest BIC model are given in Table~2.
+between the data and the model to within the formal uncertainties.
+Table~2 gives the best-fitting parameters for the preferred model,
+which has the lowest BIC value.
 
@@
 with peak-to-peak amplitude $\approx$10\%.  The $10.74\,{\rm hr}$
-variability is always more complex.  A dip of depth $\approx$1.2\%,
-fit in our model as a transit, lasts $\approx$0.75 hours.  Superposed
-on the dip is an asymmetric sinusoid with peak-to-peak
-amplitude of about 4\%. The asymmetric sinusoid peaks near phase 0.25,
-and reaches minimum brightness between phases -0.5 and -0.25.  The
-flux at phase $\approx -0.33$ shows what could be a discontinuous
-jump, shortly after reaching minimum (Figure~\ref{fig:phasefold}).
-This jump was visible in each of the nine models we considered.
+variability is always more complex.  The overall impression is of
+a distorted sinusoidal function,
+with a peak-to-peak
+amplitude of about 4\%.
+The asymmetric sinusoid rises to a maximum near phase 0.25,
+and reaches minimum brightness between phases $-0.5$ and $-0.25$.
+Between phases $-0.5$ and $0.0$ there appears to be complex shorter-timescale variability, 
+including an abrupt jump at phase $\approx -0.33$,
+and ending with a ``dip'' 
+of depth $\approx$1.2\%, lasting $\approx$0.75 hours.
 
 The periodogram of the final residual (Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}
-bottom row) shows a weakly significant, poorly resolved peak at
-$\approx$8 days, consistent with the visual impression in the time
-domain that there could be a weak long-period signal present.
-
-\section{Binarity Analysis}
+bottom row) shows a barely significant and poorly-resolved peak at
+$\approx$8 days, consistent with the visual impression of some slower
+trends in the residuals after subtracting the shorter- and longer-period
+signals.
+
+\section{Tests for Binarity}
 \label{sec:gaia}
@@
 		{\it Top:} Mean TESS image of \ptfo\ over Sector~6, with a
-		log-stretch.  The position of \ptfo\ is shown with a yellow
-		star.  Neighbors with $T<17$ are shown with orange crosses.  The
-		apertures used to measure the background and target star flux are
-		shown with \texttt{X} and \texttt{/} hatches, respectively.
+		logarithnic grayscale. The yellow star is the position of \ptfo\.
+		Orange crosses are neighboring stars with $T<17$. The
+		\texttt{X} and \texttt{/} hatches show the apertures
+		used to measure the background and target star flux, respectively.
 		{\it Bottom:} Digitized Sky Survey $R$-band image of the same
-		field, with a linear stretch. The circles show apertures of radii
-		1, 1.5, and 2.25 pixels used in part of our blend analysis.  The
-		pixel level TESS data show that ``Star A''  does not contribute
-		variability at either of the two observed periods (see
+		field, with a linear grayscale. The circles show the apertures of radii
+		1, 1.5, and 2.25 pixels used in our blend analysis. To the northwest
+		of \ptfo\ and between the
+		blue and orange circles is ``Star A'',
+		the only star bright and close enough to be contributing
+		to the signal attributed to \ptfo\. However, the pixel-level TESS data
+		showed that any such contamination is negligible (see
 		Section~\ref{subsec:blend}).
+		% JNW Label Star A?
 		\label{fig:scene}
@@
 The TESS pixels are $\approx21$'' per side. Before making any
-interpretations, we needed to consider whether light from known
-neighboring stars could have affected the photometry.  The scene is
+interpretations, we needed to consider whether light from
+neighboring stars could have contributed to the photometric signal
+we are attributing to \ptfo. The scene is
 shown in Figure~\ref{fig:scene}.  In the upper panels, the pixels used
 to measure the background level in the SPOC lightcurve are indicated
-with an `\texttt{X}' hatch, and the pixels used in the final
-lightcurve aperture are shown with the `\texttt{/}' hatch.
+with `\texttt{X}' hatching, and the pixels used in the final
+lightcurve aperture are shown with `\texttt{/}' hatching.
 
@@
 star inside the target aperture, TIC 264461979, has $T=16.8$ and so
-cannot contribute a signal with relative amplitude 10\%.  The only
+cannot contribute more than about 10\% to the total signal.  The only
 neighbor that is sufficiently close and bright that its light might
 contaminate the target star is TIC 264461980, with $T=14.8$, which we
-dub ``Star A''.  Star A is 23.6'' NW of our target, and based on the
-magnitude difference could contribute up to 48\% the flux of our
+dub ``Star A''.  Star A is 23.6'' northwest of the target, and based on the
+magnitude difference could contribute flux variations as large as 48\% the flux of our
 target star, \ptfo.
 
-Because \ptfo\ was previously identified to have periodicity
-consistent with our measurement of $P_{\rm s}$, our main concern
-regarding blending was the degree to which we could be certain that
-the long-period signal at $P_{\rm \ell}$ also originated from \ptfo.
-We took two approaches to verifying the source of the long-period
+The variability of \ptfo\ with a period consistent with $P_{\rm s}$ had already been observed
+based on images with arcsecond resolution.
+Thus, our main concern regarding blending
+was whether 
+the longer-period signal with period $P_{\rm \ell}$ originated from \ptfo,
+or from Star A. We took two approaches to investigate the source of the long-period
 signal.
 
-First, we examined the CDIPS FFI lightcurves of the target, which were
-available on MAST \citep{bouma_cluster_2019}.  The maximal
-peak-to-peak beat amplitude was constant across apertures of radii 1,
-1.5, and 2.25 pixels to visual precision ($\lesssim 1\%$).  If Star A
+First, we examined the CDIPS FFI lightcurves of the target, which are
+available on MAST \citep{bouma_cluster_2019}. Three lightcurves are available,
+based on photometric apertures with a radius of 1, 1.5, or 2.5 pixels. The maximal
+peak-to-peak beat amplitude was the same to within a percent,
+regardless of the size of the photometric
+aperture that was used to create the lightcurve.  If Star A
 were the source of the long-period variability, we would expect the
@@
 
-Second, we examined the 2-minute lightcurve of each pixel in the scene
-individually.  We opted to use the interactive tools implemented in
+Second, we examined the 2-minute lightcurve of each individual pixel in the scene,
+using the interactive tools implemented in
 \texttt{lightkurve} \citep{lightkurve_2018}.  If Star A were the
@@
 nearest to Star A to show a sinusoidal signal with amplitude exceeding
-$10\%$.  The data do not support this possibility.  The pixel directly
-below Star A does not clearly show the sinusoidal variability, and the
+$10\%$.  The data do not show this pattern.  The pixel directly
+below Star A does not clearly show any sinusoidal variability; the
 peak-to-peak variability in that pixel is $\lesssim 8\%$.  In
-contrast, the south-easternmost pixel within \ptfo's aperture (the
+contrast, the south-easternmost pixel within the \ptfo\ aperture (the
 pixel furthest from Star A that was used in the optimal aperture)
-shows the $P_{\rm \ell}$ sinusoidal variability signal at $\approx
-14\%$ amplitude.  We conclude that within the resolution of the Gaia
+shows the longer-period sinusoidal variability signal with an amplitude of
+14\%.  We conclude that within the resolution of the Gaia
 DR2 source catalog, the $P_{\rm s}$ and $P_{\rm \ell}$ signals
-originate from \ptfo.  From \citet{ziegler_measuring_2018}, we can
-surmise that stellar companions wider than $\approx$1'' (349~AU) and
+originate from \ptfo.  Based on the work of \citet{ziegler_measuring_2018}, we can
+surmise that stellar companions with separations wider than $\approx$1'' (349~AU) and
 within $\Delta G \approx 3$ magnitudes of \ptfo\ would have likely
@@
 \citet{van_eyken_ptf_2012} obtained stronger constraints on possible
-stellar companions using the NIRC2 camera on Keck II.  They reported
-3$\sigma$ $H$-band magnitude difference limits of 4.3, 6.4, and 8.9 at
+stellar companions based on high-resolution imaging with the NIRC2 camera on Keck II.  They reported
+3-$\sigma$ $H$-band magnitude difference limits of 4.3, 6.4, and 8.9 at
 angular separations of 0.25, 0.5, and 1.0 arcseconds (87, 175, and
-349~AU).  They also detected a point-source, not present in Gaia DR2,
-7.0 magnitudes fainter than the target, and 1.8$''$ north-east.  Due
-to the brightness difference, this coincident star\footnote{This
-companion was claimed to be a potential planetary-mass object
+349~AU).  They also detected a point source, not present in Gaia DR2,
+7.0 magnitudes fainter than the target, and 1.8$''$ to the north-east.  Due
+to the brightness difference, this point source\footnote{This
+point source was claimed to be a potential planetary-mass object
 \citep{schmidt_direct_2016}. Subsequent analysis of its colors showed
@@
 %FIXME: would be better to actually be able to get them...
-While long-baseline radial velocity measurements could provide
-complimentary lower limits on companion masses and separations, the RV
-data for \ptfo\ are rather poor owing to the stellar faintness.  The
-longest single-instrument baseline in the literature appears to be 5
+While long-baseline radial velocity (RV) measurements could provide
+complementary lower limits on companion masses and separations, the available RV
+data for \ptfo\ have large uncertainties due to the faintness of the star.  The
+RV dataset with the longest timespan in the literature appears to be the 5
 Keck/HIRES measurements acquired over 10 days in April 2011 by
-\citet{van_eyken_ptf_2012}.  The RV RMS over that 10 day span was
+\citet{van_eyken_ptf_2012}.  The root-mean-squared RV over that 10 day span was
 $1.6\,{\rm km}\,{\rm s}^{-1}$, consistent with the measurement
-precision.  Though \citet{yu_tests_2015} acquired 22 further
-Keck/HIRES observations over one night in December 2013, these points
-seem to not have been reduced to velocities.  Further Keck/HIRES
+precision.  Although \citet{yu_tests_2015} acquired 22 further
+Keck/HIRES spectra on one night in December 2013, those points
+were not reduced to velocities. Further Keck/HIRES
 measurements of this target could confirm or refute the existence of
@@
 	\caption{ {\bf Evidence for binarity in \ptfo}.
-    {\it Top: Hertzsprung-Russell diagram of \ptfo\ and late-type
-    members of 25$\,$Ori-1.} Members of the 25$\,$Ori-1 group (black
+    {\it Top:} Hertzsprung-Russell diagram of \ptfo\ and late-type
+    members of 25$\,$Ori-1. Members of the 25$\,$Ori-1 group (black
     circles) were identified by \citet{kounkel_apogee2_2018}.  The
-    ``neighborhood'' (gray circles) is non-member stars within 5
+    gray circles are stars in the ``neighborhood'', i.e.,
+    non-member stars within 5
     standard deviations of the mean 25$\,$Ori-1 right ascension,
@@
     M dwarfs, to accentuate \ptfo's separation from the
-    single-star sequence.  {\it Bottom: Astrometric goodness-of-fit
-    versus $Rp$ magnitude for 25$\,$Ori-1 members}.  The single-source
+    single-star sequence.  {\it Bottom:} Astrometric goodness-of-fit
+    versus $Rp$ magnitude for 25$\,$Ori-1 members.  The single-source
     astrometric model for \ptfo\ provides a poor fit, which could be due to
@@
     are at least as variable as \ptfo\ show lower astrometric excesses
-    (black squares), suggesting binarity as the probable root cause.
+    (black squares), suggesting binarity as the root cause.
 		\label{fig:gaia}
@@
 
-Aside from visual binarity, we can also check the Gaia data for
-photometric binarity.  To assemble a set of stars coeval with \ptfo,
+We also used the Gaia data to see if the observed luminosity of \ptfo
+is too bright to be a single star, i.e., if it is a ``photometric
+binary.''
+To assemble a set of stars coeval with \ptfo,
 we used the 25$\,$Ori-1 members identified by
@@
 
-To define a set of non-member stars that nonetheless had comparable
-selection functions, we defined a reference ``neighborhood'' as the
+To define a set of non-member stars that nonetheless are subject
+to similar selection criteria,
+we defined the reference ``neighborhood'' as the
 group of at most $10^4$ randomly selected non-member stars within 5
@@
 Orion complex, or even of 25$\,$Ori-1, enforcing this cut on positions
-and parallaxes ensures that we are querying stars with comparable
+and parallaxes ensures that we are comparing stars with similar
 amounts of interstellar reddening.
 
-We examined the resulting five-dimensional right ascension,
-declination, proper motions, and parallaxes.  The first point we noted
-was that 25$\,$Ori-1 was a clearly defined over-density in each
-dimension---the cluster exists, and is distinct from the neighborhood.
-\ptfo\ was also within the cluster in each of these projected
+We examined the resulting five-dimensional distribution of right ascension,
+declination, proper motion in both directions, and parallax.  The first point we noted
+was that 25$\,$Ori-1 is a clearly defined over-density in each
+dimension: the cluster was confirmed to exist, and to be distinct from the neighborhood. \ptfo\ was also within the cluster in each of these projected
 dimensions.
 
-Given our detection of two separate signals, whether \ptfo\ could be a
-photometric binary was of great interest.  Figure~\ref{fig:gaia} shows
-the HR diagram we constructed to assess this issue.  The diagram shows
+%Given our detection of two separate signals, whether \ptfo\ could be a
+%photometric binary was of great interest.  
+Figure~\ref{fig:gaia} shows
+the HR diagram we constructed from the data.  The diagram shows
 that \ptfo\ is $\approx$0.75 magnitudes brighter than the average
 25$\,$Ori-1 star of the same color.  In other words, it is about twice
-as bright.  It also seems to be coincident with the photometric binary
-track of the cluster, which has a few other stars.
+as bright as expected for a single star in the cluster.
+It also seems to be part of a ``photometric binary''
+track that runs parallel to the main track.
 
@@
 than the kinematically identical 25$\,$Ori-1 members, or {\it (ii)}
-\ptfo\ is a photometric binary.  Since there is no a priori reason to
-suspect an age difference, but we have resolved two separate
-photometric signals, the binary interpretation seems somewhat more
-likely.
+\ptfo\ is a binary with two components of nearly equal brightness.
+Since there is no other reason to
+suspect an age difference, and because the source showed two separate
+photometric signals with similar but distinct periods,
+the binary interpretation seems more probable.
 
@@
 astrometry.  As noted in Section~\ref{sec:observations}, the Gaia DR2
-solution for \ptfo\ shows a 10.3$\sigma$ astrometric excess.  This
-astrometric excess indicates the degree to which a single-source model
-fails to explain the observed astrometric measurements.  Specifically,
+astrometric solution for \ptfo\ shows a 10.3$\sigma$ ``astrometric excess'',
+a parameter that quantifies the degree to which a single-star model
+fails to fit the astrometric measurements.  Specifically,
 the single-source astrometric model yielded $\chi^2=325.2$.  There are
@@
 the subset of members that were at least as variable as \ptfo.
-We measured the variability amplitude by taking the difference of the
-$95^{\rm th}$ and $5^{\rm th}$ percentiles of the flux.
+We measured the variability amplitude by taking the difference between the
+$95^{\rm th}$ and $5^{\rm th}$ percentiles of the flux measurements.
 This yielded 30 stars of equal or greater variability.
@@
 group members with comparable brightnesses and variability
-characteristics, \ptfo\ still stands out as behaving astrometrically
-poorly.  Ultimately, we will have to wait for the full release of the
-nominal Gaia mission to definitively determine whether the astrometric
+characteristics, \ptfo\ still stands out by virtue of its failure
+to conform to a single-star astrometric model. This supports
+the interpretation that \ptfo\ is a binary star.
+
+We will have to wait for the full release of the
+nominal Gaia mission to determine definitively whether the astrometric
 excess is caused by stellar binarity or photometric variability.
@@
 
-\subsection{Long Period Sinusoid}
-
-The standard interpretation for 11.96$\,$hour sinusoidal modulations
+\subsection{Longer-Period Signal}
+
+The standard interpretation for 11.96$\,$hour nearly sinusoidal modulations
 of a pre-main-sequence M dwarf is stellar rotation.  This is the
@@
 
-The discovery study by \citet{van_eyken_ptf_2012} saw an alias of the
-same signal ({\it e.g.}, their Figure~7), and identified it as a
+In their report on the discovery of the unusual
+photometric variability, \citet{van_eyken_ptf_2012} saw an alias of the
+same signal ({\it e.g.}, their Figure~7),
+% JNW: Don't understand this sentence. What is the 'same signal'?
+and identified it as a
 periodogram peak at $0.9985 \pm 0.0061\,$days. They ascribed it to
 their observing cadence, because of its close correspondence to the
-sidereal day.  While the TESS data can show significant reflected
-light from the Earth \citep[{\it e.g.},][]{luger_tess_2019}, our
+sidereal day.
+% JNW What is this next phrase doing here?
+%While the TESS data can show significant reflected
+%light from the Earth \citep[{\it e.g.},][]{luger_tess_2019}, 
+Our
 pixel-level analysis showed that the signal is specific to only pixels
 near \ptfo, and no other pixels.  We therefore conclude that the
-signal is not a systematic.
+signal is not an artifact of systematic errors.
 
@@
 
-\subsection{Short Period Dip}
+\subsection{Shorter-Period Signal, Including the ``Dip''}
 
@@
 	\caption{
-		{\bf Timing residuals for \ptfob\ from a decade of monitoring.}
+		{\bf Timing residuals for \ptfob\ based on a decade of monitoring.}
 		Black points are times of dips, minus the indicated linear
-		ephemeris.  The $y$-axis is given in units of phase for the
-		short-period signal.  The star shows the binned TESS ephemeris.
-		Dips have been observed by \citet{van_eyken_ptf_2012},
+		ephemeris.  The phase of the shorter-period signal
+		is plotted on the $y$-axis. The star symbol represents
+		the TESS ephemeris.
+		Dips were observed by \citet{van_eyken_ptf_2012},
 		\citet{ciardi_followup_2015}, \citet{yu_tests_2015},
@@
 The TESS lightcurve shows a dip that lasts about 45 minutes, and seems
-to re-occur every 10.74 hours
+to recur every 10.74 hours
 (Figures~\ref{fig:splitsignal},~\ref{fig:splitsignalii},~\ref{fig:phasefold}).
@@
 
-One particularly interesting feature of the dip is its epoch.  Over
-the past decade, many investigators have observed \ptfo.  Its dips do
-not always occur on a perfectly linear ephemeris
+One particularly interesting feature of the dip is its epoch.  The dips do
+not occur with strict periodicity
 \citep{yu_tests_2015}.  In fact, \citet{tanimoto_evidence_2020}
-recently provided stark evidence for different behavior altogether:
-over a time-span of years, the dip ``splits'' into distinct groups at
+provided stark evidence for different behavior altogether:
+over a time-span of years, the dip ``split'' into distinct groups at
 particular repeating phases.  See for instance their Figures~2
@@
 
-In Figure~\ref{fig:o_minus_c}, we show the phase of the dip we detect
+Figure~\ref{fig:o_minus_c} shows the phase of the dip we detected
 in the TESS data, relative to their ephemeris.  It agrees with the
 independent December 2018 measurements by
-\citet{tanimoto_evidence_2020}: the dip has drastically shifted phase
-over the past decade.
+\citet{tanimoto_evidence_2020}: either the dip abruptly shifted phase
+over the past decade or, more likely, there are multiple dips that
+have come and gone at different phases.
 
@@
 Inspecting the \citet{tanimoto_evidence_2020} lightcurves, the claim
-of multiple dips per cycle in December 2018 at phase 0 and -0{.}47
-seems less plausible---the phase -0{.}47 dips are strongly detected,
+of multiple dips per cycle in December 2018 at phase 0 and $-0{.}47$
+seems less plausible---the phase $-0{.}47$ dips are strongly detected,
 while the suggested phase 0 dip is not clearly present in the data.
@@
 leave this curiosity as observation {\it sans} interpretation.
+% JNW: This seems like an important clue. I don't have the mental
+% energy to ponder it now, but it deserves pondering.
 
@@
 
-When physical explanations are not forthcoming, taxonomy can be a
-useful exercise.  To our knowledge, of order 10-20 lightcurves with
-similar morphologies have been seen in surveys of low-mass weak-lined
+When physical explanations are not forthcoming, taxonomy is a
+useful exercise.  By searching the literature, we have found about a dozen lightcurves with
+similar morphologies to \ptfo, drawn from surveys of low-mass weak-lined
 T Tauri stars in regions including $\rho$ Oph, Upper~Sco, Taurus, and
@@
 These surveys were performed using K2 \citep{howell_k2_2014}.  We
-downloaded a few of these lightcurves from MAST, opting for the
-EVEREST reductions \citep{luger_everest_2016,luger_update_2018}.  They
-are plotted in Figure~\ref{fig:brethren}.
+downloaded some of these lightcurves from MAST, opting for the
+EVEREST reductions \citep{luger_everest_2016,luger_update_2018}.
+They are plotted in Figure~\ref{fig:brethren}.
 
@@
 campaigns \citep{stauffer_orbiting_2017}.
-As \citet{stauffer_orbiting_2017} explain,
-these  objects are morphologically distinct from ``scallop shell' lightcurves,
+In the terminology of \citet{stauffer_orbiting_2017},
+these objects are morphologically distinct from ``scallop shell'' lightcurves,
 and are present in stars at more advanced evolutionary disk stages than
@@
 The persistent and transient flux dip stars all show angular dips
-that are cannot be explained with starspots.
-They typically have the following properties.
+that are cannot be explained as the effects of starspots.
+These stars typically have the following things in common:
 \begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
-  \item Weak-lined T Tauri stars.
-  \item Spectral type typically M2-M5 ({\it e.g.},
+  \item They are weak-lined T Tauri stars.
+  \item The spectral type is M2 to M5 ({\it e.g.},
     \citealt{rebull_usco_2018},~Figure~20).
-  \item Ages $\lesssim$ 100$\,$Myr.
-  \item Stars typically lack spectroscopic accretion indicators 
-  \item Phased lightcurves show shallow, angular dips, usually
+  \item The age is $\lesssim$ 100$\,$Myr.
+%  \item They lack spectroscopic indicators of strong accretion.
+% JNW: Same as WTTS?
+  \item The lightcurves show shallow, angular dips, usually
     superposed on large-amplitude smooth variability. The latter
     is interpreted as stellar rotation.
-  \item Rotation is fast---usually between 0.5 and 2.0 days.
-  \item Rarely any infrared excess (never any W4 detection; only a few
-    cases with W3 detctions).
-  \item Can show multiple dips per cycle.
+  \item The rotation is rapid---usually between 0.5 and 2.0 days.
+  \item There is rarely any detectable infrared excess in WISE data
+  (never any W4 detection; only a few W3 detections).
+  \item They sometimes show multiple dips per cycle.
   %({\it e.g.}, EPIC 203692610, RIK-210).
-  \item Dip depths, durations, and phases can vary over just a few cycles
+  \item The dip depths, durations, and phases can vary over just a few cycles
     ({\it e.g.}, EPIC 204143627).
-  \item Dip depths can change after flares.
-  \item Rare at a population level: $\lesssim 1\%$ of relevant stars
+  \item The dip depths can change after flares.
+  \item They are rare at a population level: $\lesssim 1\%$ of relevant stars
     \citep{rebull_usco_2018}.
@@
 There are two crucial additional points concerning the transient flux
-dips.  First, the dip durations seem to scale linearly with their
-periods \citep[][Figure~26]{stauffer_orbiting_2017}.  In the usual
+dips.  First, the dip durations seem to scale linearly with
+the photometric periods \citep[][Figure~26]{stauffer_orbiting_2017}.  In the usual
 idealized limits, the transit duration $T$ of a point source across
@@
 consistent with point sources, at longer periods of 1 to 5 days the
-dip durations become many hours, and far too long for a point-source
-origin.
+dip durations become many hours, far too long to be caused by
+planetary transits.
 
@@
 \citep{stauffer_rotevol_2018}.  For the case of \ptfo, there is a
-third line of evidence: the Gaia astrometry shows a poor
-goodness-of-fit to a single-source model.  While this could be caused
-by stellar variability, cluster members
-that are at least as variable still show \ptfo\ to be an
-outlier.  Therefore the astrometric excess is a suggestive third line
-of evidence for binarity in \ptfo.  We take these three points
-combined to imply that \ptfo\ is probably a binary.
+third line of evidence: the Gaia DR2 entry for \ptfo\ reports a poor
+fit of the single-star model to the astrometric data.  While this could be caused
+by stellar variability, other cluster members
+that are just as variable do not typically show the same level of
+excess astrometric motion.
+Therefore the astrometric excess is a suggestive third line
+of evidence for binarity in \ptfo. To us, the evidence leads to
+the conclusion that \ptfo\ is a nearly equal-mass binary consisting
+of two rapidly rotating stars.
 
@@
 evolution.
+% JNW I wonder if there could be a circumbinary ring, a la KH 15D.
 
 The main physical question is what causes the
-transient dipping. This is an unsolved problem not only for \ptfo\ but
+transient dips. This is an unsolved problem not only for \ptfo\ but
 also for an emerging class of similar young rapidly rotating M-dwarfs.
@@
 {\it (iii)} occultations of starspots by an optically thick disk.
-The first and last explanations offer added appeal because they are
+The first and last explanations have added appeal because they are
 flexible enough to explain not only the transient and persistent-dip
@@
 
-The combination of TESS and Gaia has provided new
-clarity in the \ptfo\ system.  Our main results are as follows.
+The combination of TESS and Gaia data has clarified a few things about
+the \ptfo\ system.  Our main results are as follows.
 \begin{itemize}
@@
 The jumping orbital phase disfavors the explanation of an
-enshrouded, transiting protoplanet.  Though \ptfob\ seems unlikely
-to be a planet, understanding it and its analogs is a worthy
-problem that promises to teach us about the natal environments of the
+enshrouded, transiting protoplanet.  Though \ptfob\ may not be a planet,
+as we and others had hoped, understanding \ptfo\ and its analogs is a worthy
+problem that might even teach us about the natal environments of the
 majority of habitable-zone Earth-sized planets in the Milky Way
 \citep{dressing_occurrence_2013}.
+% JNW: seems like a stretch but I will not strike it down
 
@@
 that were eventually omitted.
+%
+We thank the Heising-Simons Foundation for their generous support of this work.
 %

### 2020-04-27 paper/ms.tex
JNW edits
Co-authored-by: joshwinn <jwinn@astro.princeton.edu>

This also merges https://github.com/lgbouma/billy/pull/2

@@
 %PTFO$\,$8-8695}
-\title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
-PTFO 8-8695}
-\title{PTFO\,8-8695: A binary star, not a hot Jupiter host}
-
-\correspondingauthor{L. G. Bouma}
+% \title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
+% PTFO 8-8695}
+% \title{PTFO\,8-8695: A binary star, not a hot Jupiter host}   
+%\title{Two Stars and Two Signals in PTFO 8-8695}
+\title{PTFO 8-8695: Two Stars, Two Signals, No Planet}
+
+\correspondingauthor{L.\,G.\,Bouma}
 \email{luke@astro.princeton.edu}
@@
 %
-\author[0000-0002-0514-5538]{L. G. Bouma}
+\author[0000-0002-0514-5538]{L.\,G.\,Bouma}
 \affiliation{ Department of Astrophysical Sciences, Princeton
@@
 \author[0000-0002-4265-047X]{J.\,N.\,Winn}
+\affiliation{ Department of Astrophysical Sciences, Princeton
+University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+%
+\author[0000-0002-4265-047X]{TESS collaborators}
 \affiliation{ Department of Astrophysical Sciences, Princeton
@@
   periods. The variability is dominated by a sinusoidal signal with a period
-  of 11.96\,$hr, presumably caused by stellar rotation.  Also
+  of 11.96$\,$hr, presumably caused by stellar rotation.  Also
   present is a 10.74-day signal consisting of a not-quite sinusoid interrupted
-  by 1.2\% hour-long dips, the type of signal previously
+  by hour-long dips, the type of signal previously
   intepreted as planetary transits.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\epigraph{
-	Nubble [n]: a small lump or protuberance.
-}
 
@@
 Section~\ref{sec:conclusions}.
-% JNW: this kind of stuff is for the abstract/conclusion, not intro.
-%The phase of the dip within the seems
-%to have instantaneously changed over the past decade.  In addition,
-%other young stars show lightcurve morphologies similar to the
-%short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
-%binary M dwarf in which one star shows a rotation signal, and the
-%other shows transient dipping potentially caused by eclipses of
-%lasma at the Keplerian co-rotation radius.
 
@@
 \label{sec:observations}
-
-% JNW: How about more descriptive labels? f --> Total flux
-% fl = Longer-period signal
-% fs = Shorter-period signal
-% last panel, Residuals
 
@@
 
-\ptfo\ [also known as CVSO\,30; \cite{briceno_cida_2005}}]
+\ptfo\ [also known as CVSO\,30; \citealt{briceno_cida_2005}]
 was observed by TESS with Camera 1, CCD 1, from December 15,
@@
 circular aperture with radius 1 pixel.
-% JNW: What does that mean? You used fractional pixels in calculating the pixel sums?
 
@@
 We normalized the lightcurve by dividing out the median flux, and then opted
-to subtract 1.0 to set the median value to zero, which simplified
+to subtract 1{.}0 to set the median value to zero, which simplified
 subsequent interpretation.  Many
@@
 brighest 1{.}3 billion stars, including \ptfo\
-\citep{lindegren_gaiasoln_2018}.  For \ptfo, there were 121 ``good'' observations, i.e., observations that were not strongly down-weighted in
+\citep{lindegren_gaiasoln_2018}.  For \ptfo, there were 121 ``good''
+observations, {\it i.e.}, observations that were not strongly down-weighted in
 the astrometric solution.  \ptfo\ was assigned the Gaia DR2 identifier
@@
 6.9$\,$Myr, while the HR diagram led to an estimated age of 8.5$\,$Myr
-\citep{kounkel_apogee2_2018}. 
-% JNW: What's the difference between a CMD and an HR diagram?
+\citep[see][Section 2.3]{kounkel_apogee2_2018}. 
 \citet{kounkel_apogee2_2018} identified
@@
 
-% JNW This doesn't seem necessary to me.
-%Initial signal-processing experiments fitting out splines or sinusoids
-%showed that after subtracting out the long-period signal, the
-%short-period signal dominated the periodogram, and vice-versa.
-%However it quickly became clear that it would be beneficial to
-%simultaneously model each signal, in order to preserve the
-%power at each frequency.
-
 \subsection{Lightcurve Model}
@@
 \begin{align}
-  f = &f_{\rm transit,s} + \sum_{n=1}^{N} A_n \sin(n\omega_{\rm s}t)
-  + \sum_{n=1}^{N} B_n \cos(n\omega_{\rm s}t)\\
-  &+ \sum_{m=1}^{M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
-  + \sum_{m=1}^{M} B_m \cos(m[\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
+  f = &f_{\rm transit,s} + \sum_{n=1}^{N_{\rm s}} A_n \sin(n\omega_{\rm s}t)
+  + \sum_{n=1}^{N_{\rm s}} B_n \cos(n\omega_{\rm s}t)\\
+  &+ \sum_{m=1}^{N_{\rm \ell}} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
+  + \sum_{m=1}^{N_{\rm \ell}} B_m \cos(m[\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
 \end{align}
-where $N$ and $M$ are the total number of modes at the shorter and longer
+where $N_{\rm s}$ and $N_{\rm \ell}$ are the total number of modes at the shorter and longer
 periods, respectively,
-% JNW: maybe N_\ell and N_s?
-$A_i$ and $B_i$ the amplitudes of each
+$A_i$ and $B_i$ are the amplitudes of each
 mode (which can be positive or negative), and $\omega_\ell$
@@
 would be appropriate to include in the model, we considered a number of different choices for
-$N$ and $M$, and used the Bayesian information criterion to select
+$N_{\rm s}$ and $N_{\rm \ell}$, and used the Bayesian information criterion to select
 the final model (Table~1).
 
-As an example, one possible model consists of a transit, $N=2$
-sines and cosines at the shorter period, plus $M=1$
+As an example, one possible model consists of a transit, $N_{\rm s}=2$
+sines and cosines at the shorter period, plus $N_{\rm \ell}=1$
 sine and cosine at the longer period.  In this case, the free parameters would
@@
 the time of a particular transit, and the mean flux.  There
-would be $2N=4$ additional Fourier amplitudes at the shorter period,
-plus $2M=2$ Fourier amplitudes at the longer period, and well as $P_\ell$
+would be $2N_{\rm s}=4$ additional Fourier amplitudes at the shorter period,
+plus $2N_{\rm \ell}=2$ Fourier amplitudes at the longer period, and well as $P_\ell$
 itself and the relative phase $\phi_\ell$.  The total number
@@
 we used the default math operators.
-% JNW what does that mean? If you just mean 'numpy', is it necessary to state?
 For the exoplanet transit, we
@@
 We considered nine models, with the number of modes per frequency
-$N$ and $M$ ranging from one to three.  To select our preferred model,
+$N_{\rm s}$ and $N_{\rm \ell}$ ranging from one to three.  To select our preferred model,
 we used the Bayesian information criterion (Table~1).  The model with
@@
 two modes at the longer 11.96$\,$hr period.  The next-best model had
-an additional harmonic at the longer period ($M=3$), but was otherwise
+an additional harmonic at the longer period ($N_{\rm \ell}=3$), but was otherwise
 identical.  All nine models have reduced $\chi^2$ values ranging between 1.37
@@
 		{\it Top:} Mean TESS image of \ptfo\ over Sector~6, with a
-		logarithnic grayscale. The yellow star is the position of \ptfo\.
+		logarithmic grayscale. The yellow star is the position of \ptfo.
 		Orange crosses are neighboring stars with $T<17$. The
@@
 		Section~\ref{subsec:blend}).
-		% JNW Label Star A?
 		\label{fig:scene}
@@
 
-%FIXME: would be better to actually be able to get them...
 While long-baseline radial velocity (RV) measurements could provide
@@
 
-%Given our detection of two separate signals, whether \ptfo\ could be a
-%photometric binary was of great interest.  
 Figure~\ref{fig:gaia} shows
@@
 photometric variability, \citet{van_eyken_ptf_2012} saw an alias of the
-same signal ({\it e.g.}, their Figure~7),
-% JNW: Don't understand this sentence. What is the 'same signal'?
+longer-period signal ({\it e.g.}, their Figure~7),
 and identified it as a
@@
 sidereal day.
-% JNW What is this next phrase doing here?
-%While the TESS data can show significant reflected
-%light from the Earth \citep[{\it e.g.},][]{luger_tess_2019}, 
 Our
@@
 of multiple dips per cycle in December 2018 at phase 0 and $-0{.}47$
-seems less plausible---the phase $-0{.}47$ dips are strongly detected,
+seems less plausible---the phase $\text{-}0{.}47$ dips are strongly detected,
 while the suggested phase 0 dip is not clearly present in the data.
@@
   \item The age is $\lesssim$ 100$\,$Myr.
-%  \item They lack spectroscopic indicators of strong accretion.
-% JNW: Same as WTTS?
   \item The lightcurves show shallow, angular dips, usually
@@
   \item They sometimes show multiple dips per cycle.
-  %({\it e.g.}, EPIC 203692610, RIK-210).
   \item The dip depths, durations, and phases can vary over just a few cycles
@@
 evolution.
-% JNW I wonder if there could be a circumbinary ring, a la KH 15D.
 
@@
 \citep{dressing_occurrence_2013}.
-% JNW: seems like a stretch but I will not strike it down
 
@@
 
-% \clearpage
+\clearpage
 \input{model_comparison_table.tex}
 \input{model_posterior_table.tex}
-% \clearpage
+\clearpage
 

### 2020-04-27 billy/plotting.py
ditto

@@ -166,7 +166,8 @@ def plot_splitsignal_map(m, outpath, part='i'):
     # 8.5x11 is letter paper. x10 allows space for caption.
     fig, axs = plt.subplots(nrows=4, figsize=(8.5, 10), sharex=True)
 
-    axs[0].set_ylabel('f', fontsize='x-large')
+    axs[0].set_ylabel('Raw flux', fontsize='x-large')
+    # axs[0].set_ylabel('f', fontsize='x-large')
     axs[0].plot(m.x_obs[g], m.y_obs[g], ".k", ms=4, label="data", zorder=2,
                 rasterized=True)
     axs[0].plot(m.x_obs[g], m.map_estimate['mu_model'][g], lw=0.5, label='MAP',
@@ -190,22 +191,28 @@ def plot_splitsignal_map(m, outpath, part='i'):
             # axs[0].plot(m.x_obs, y_orb, lw=0.5, label='model '+f,
             #             color='C{}'.format(ix+1), alpha=1, zorder=ix+3)
 
-    axs[1].set_ylabel('$f_{{\mathrm{{\ell}}}} = f - f_{{\mathrm{{s}}}}$',
+    axs[1].set_ylabel('Longer period',
                       fontsize='x-large')
+    # axs[1].set_ylabel('$f_{{\mathrm{{\ell}}}} = f - f_{{\mathrm{{s}}}}$',
+    #                   fontsize='x-large')
     axs[1].plot(m.x_obs[g], m.y_obs[g]-y_orb[g], ".k", ms=4, label="data-orb",
                 zorder=2, rasterized=True)
     axs[1].plot(m.x_obs[g], m.map_estimate['mu_model'][g]-y_orb[g], lw=0.5,
                 label='model-orb', color='C0', alpha=1, zorder=1)
 
-    axs[2].set_ylabel('$f_{{\mathrm{{s}}}} = f - f_{{\mathrm{{\ell}}}}$',
+    axs[2].set_ylabel('Shorter period',
                       fontsize='x-large')
+    # axs[2].set_ylabel('$f_{{\mathrm{{s}}}} = f - f_{{\mathrm{{\ell}}}}$',
+    #                   fontsize='x-large')
     axs[2].plot(m.x_obs[g], m.y_obs[g]-y_rot[g], ".k", ms=4, label="data-rot",
                 zorder=2, rasterized=True)
[…]

### 2020-04-27 paper/ms.tex
minor update

@@
 Section~\ref{sec:conclusions}.
+We comment on the
+concurrent study by \citet{koen_2020} in postscript.
 
@@
 for additional harmonics after fitting for the largest dip and two
-harmonics at $P_{\rm s}$ (Table~1). Our analysis emphasized the
+harmonics at the shorter period (Table~1). Our analysis emphasized the
 agreement between the TESS dip ephemeris and that from

### 2020-04-27 paper/ms.tex
update manuscript

@@
 We comment on the
-concurrent study by \citet{koen_2020} in postscript.
+study by \citet{koen_2020} in postscript.
 
@@
 	\caption{ {\bf Phase-folded longer and shorter-period signals.}
-    {\it Top}: Longer-period signal, as in Figure~\ref{fig:splitsignal}.
+    {\it Top}: Relative flux at the longer-period, as in Figure~\ref{fig:splitsignal}.
     {\it Bottom}: Shorter-period signal. The reference phase is set to

### 2020-04-27 paper/ms.tex
fix column width

--- 2020-04-27 paper/ms.tex
@@
 \begin{abstract}
-  PTFO$\,$8-8695 (CVSO\,30) is a star in the 7--10 million year
-  old Orion-OB1a cluster that shows brightness dips that resemble planetary transits.
-  While strong evidence against the planet hypothesis has already been provided,
-  the possibility remains debated in the literature.
-  To obtain further clues, we inspected data from the
-  NASA {\it Transiting Exoplanet Survey Satellite} (TESS) and the ESA Gaia mission.
-  The Gaia data show that
+  PTFO$\,$8-8695 (CVSO\,30) is a star in the 7--10 million year old
+  Orion-OB1a cluster that shows brightness dips that resemble
+  planetary transits.  While strong evidence against the planet
+  hypothesis has already been provided, the possibility remains
+  debated in the literature.  To obtain further clues, we inspected
+  data from the NASA {\it Transiting Exoplanet Survey Satellite}
+  (TESS) and the ESA Gaia mission.  The Gaia data show that
   PTFO$\,$8-8695 is a photometric binary with respect to members of
   its kinematic group, and independently suggest that it is an
-  astrometric binary.  The TESS lightcurve shows two different photometric
-  periods. The variability is dominated by a sinusoidal signal with a period
-  of 11.96$\,$hr, presumably caused by stellar rotation.  Also
-  present is a 10.74-day signal consisting of a not-quite sinusoid interrupted
-  by hour-long dips, the type of signal previously
-  intepreted as planetary transits.
-  The phase of the dips is nearly 180$^\circ$ away from the phase of the originally reported dips,
-  which, as noted previously, makes them difficult to explain as planetary transits.
-  Instead, we believe that PTFO$\,$8-8695 is a pair of young
-  and rapidly rotating M dwarfs, one of which shows the same ``transient-dipper''
-  behavior that been seen in at least 5 other cases.
-  The origin of these transient dips is still unknown but likely involves circumstellar material.
+  astrometric binary.  The TESS lightcurve shows two different
+  photometric periods. The variability is dominated by a sinusoidal
+  signal with a period of 11.96$\,$hr, presumably caused by stellar
+  rotation.  Also present is a 10.74-day signal consisting of a
+  not-quite sinusoid interrupted by hour-long dips, the type of signal
+  previously intepreted as planetary transits.  The phase of the dips
+  is nearly 180$^\circ$ away from the phase of the originally reported
+  dips, which, as noted previously, makes them difficult to explain as
+  planetary transits.  Instead, we believe that PTFO$\,$8-8695 is a
+  pair of young and rapidly rotating M dwarfs, one of which shows the
+  same ``transient-dipper'' behavior that been seen in at least 5
+  other cases.  The origin of these transient dips is still unknown
+  but likely involves circumstellar material.
 \end{abstract}
@@
 We wish \ptfob\ were a planet. It would be quite exceptional. It would
-be the youngest known hot Jupiter
-\citep{van_eyken_ptf_2012}, orbiting a T Tauri star in the Orion-OB1a cluster.
-It would have the shortest orbital period (10.7 hours) of any hot Jupiter.
-With such a short period, it would
-probably be filling its Roche lobe, and actively
+be the youngest known hot Jupiter \citep{van_eyken_ptf_2012}, orbiting
+a T Tauri star in the Orion-OB1a cluster.  It would have the shortest
+orbital period (10.7 hours) of any hot Jupiter.  With such a short
+period, it would probably be filling its Roche lobe, and actively
 losing mass to its host star.  Not only that, but the rapidly-rotating
-host star is probably oblate enough to torque the planet's orbit
-into and out of the transiting configuration on a time scale of years
+host star is probably oblate enough to torque the planet's orbit into
+and out of the transiting configuration on a time scale of years
 \citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.
@@
 chromospheric H$\alpha$ emisson, it seems that there is an additional
-H$\alpha$ emission with radial velocity variations in phase with the planetary
-orbit.  The average velocity width of the excess H$\alpha$ emission is
-87$\,$km$\,$s$^{-1}$, and its equivalent width is 70-80\% that of the
-stellar chromosphere \citep{johnskrull_h_2016}.  The proposed
-explanation is that the emission is from hot material flowing away
-from the planet \citep{johnskrull_h_2016}.
-
-However, the observed signals have some peculiarities that
-make the planet seem even more unusual, to the point that they cast into doubt the premise
-that \ptfob\ is real.
-First, the transit-like brightness dips are about
-three times deeper in optical bandpasses ({\it e.g.,} $g$-band) than
-in the near-infrared ({\it e.g.}, $z$-band)
+H$\alpha$ emission with radial velocity variations in phase with the
+planetary orbit.  The average velocity width of the excess H$\alpha$
+emission is 87$\,$km$\,$s$^{-1}$, and its equivalent width is 70-80\%
+that of the stellar chromosphere \citep{johnskrull_h_2016}.  The
+proposed explanation is that the emission is from hot material flowing
+away from the planet \citep{johnskrull_h_2016}.
+
+However, the observed signals have some peculiarities that make the
+planet seem even more unusual, to the point that they cast into doubt
+the premise that \ptfob\ is real.  First, the transit-like brightness
+dips are about three times deeper in optical bandpasses ({\it e.g.,}
+$g$-band) than in the near-infrared ({\it e.g.}, $z$-band)
 \citep{onitsuka_multicolor_2017,tanimoto_evidence_2020}.  An ordinary
-atmosphere expected for a Jovian planet would not lead to such a strong
-color-dependence of the transits.  
-Second, the planet does not seem to emit as much infrared radiation
-as would be expected for such a hot Jovian planet \citep{yu_tests_2015}.
-Third, despite measurement attempts by multiple investigators, \ptfob\ does
-not seem to show the Rossiter effect at the amplitude expected given
-the rapid stellar rotation and large planet size
-\citep{yu_tests_2015,ciardi_followup_2015}.
-Fourth, the phase of the dips within the overall period of photometric variability
-has changed drastically over the years since their initial discovery.
-To counter these objections, it has been proposed 
-that the planet may be much smaller than Jupiter and that the dips
-are produced by dust clouds emitted from the planet \citep{tanimoto_evidence_2020}.
-
-A separate issue is that the brightness dips
-change shape over many orbital cycles. This was initially explained by
-\cite{barnes_measurement_2013} as the natural effects of gravity darkening
-However, \cite{howarth_reappraisal_2016} argued
-that the necessary amplitude of gravity darkening is too large to be realistic, given the
-spectroscopically-determined rotation velocity.
-Additionally, as the
-gravity-darkened star precessed about its rotation axis, it would 
-show photometric variability that has not been observed.
-
-While the planetary interpetation clearly faces challenges,
-there is no completely satisfactory alternate explanation.  High-latitude accretion hotspots
-might produce the observed H$\alpha$ variability, but require
-fine-tuning to produce dips of the appropriate duration. Furthermore,
-\ptfo\ does not have an infrared (IR) excess associated with the
-presence of dust at the inner edge of a primordial disk \citep[{\it
-e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude starspots, hot or
-cold, struggle to produce photometric features as short as some
-of the observed dips.
-
-A relevant fact is that between 0.1\% and 1\% of rapidly
-rotating low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations
-show short-duration dips as part of their overall periodic variability
+atmosphere expected for a Jovian planet would not lead to such a
+strong color-dependence of the transits.  Second, the planet does not
+seem to emit as much infrared radiation as would be expected for such
+a hot Jovian planet \citep{yu_tests_2015}.  Third, despite measurement
+attempts by multiple investigators, \ptfob\ does not seem to show the
+Rossiter effect at the amplitude expected given the rapid stellar
+rotation and large planet size
+\citep{yu_tests_2015,ciardi_followup_2015}.  Fourth, the phase of the
+dips within the overall period of photometric variability has changed
+drastically over the years since their initial discovery.  To counter
+these objections, it has been proposed that the planet may be much
+smaller than Jupiter and that the dips are produced by dust clouds
+emitted from the planet \citep{tanimoto_evidence_2020}.
+
+A separate issue is that the brightness dips change shape over many
+orbital cycles. This was initially explained by
+\cite{barnes_measurement_2013} as the natural effects of gravity
+darkening However, \cite{howarth_reappraisal_2016} argued that the
+necessary amplitude of gravity darkening is too large to be realistic,
+given the spectroscopically-determined rotation velocity.
+Additionally, as the gravity-darkened star precessed about its
+rotation axis, it would show photometric variability that has not been
+observed.
+
+While the planetary interpetation clearly faces challenges, there is
+no completely satisfactory alternate explanation.  High-latitude
+accretion hotspots might produce the observed H$\alpha$ variability,
+but require fine-tuning to produce dips of the appropriate duration.
+Furthermore, \ptfo\ does not have an infrared (IR) excess associated
+with the presence of dust at the inner edge of a primordial disk
+\citep[{\it e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude
+starspots, hot or cold, struggle to produce photometric features as
+short as some of the observed dips.
+
+A relevant fact is that between 0.1\% and 1\% of rapidly rotating
+low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations show
+short-duration dips as part of their overall periodic variability
 \citep{rebull_usco_2018}.  The dips can persist over months, but their
@@
 and \citet{david_transient_2017} to explain this novel class of
-variable stars is that a circumstellar cloud of gas is orbiting
-near the co-rotation radius.  To this point, though, it has not
-been clear if this explanation applies to
-\ptfo, because the determination of
-the stellar rotation period has been somewhat ambiguous
+variable stars is that a circumstellar cloud of gas is orbiting near
+the co-rotation radius.  To this point, though, it has not been clear
+if this explanation applies to \ptfo, because the determination of the
+stellar rotation period has been somewhat ambiguous
 \citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.
@@
 two different periodic signals, which we analyze in
-Section~\ref{sec:tess}.
-The Gaia data, analyzed in Section~\ref{sec:gaia}, show that
-\ptfo\ is a photometric binary, and suggest it also an astrometric binary.
-We discuss the pieces of the puzzle
-in Section~\ref{sec:discussion}, and summarize the situation in
-Section~\ref{sec:conclusions}.
-We comment on the
-concurrent study by \citet{koen_2020} in postscript.
+Section~\ref{sec:tess}.  The Gaia data, analyzed in
+Section~\ref{sec:gaia}, show that \ptfo\ is a photometric binary, and
+suggest it also an astrometric binary.  We discuss the pieces of the
+puzzle in Section~\ref{sec:discussion}, and summarize the situation in
+Section~\ref{sec:conclusions}.  We comment on the study by
+\citet{koen_2020} in postscript.
 
@@
 	\caption{
-		{\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 19).}
-    {\it Top}: The ``raw'' (\texttt{PDCSAP} mean-subtracted) relative flux
-    data. The beat period of 4.48 days is visible by eye.  The
-    blue curve is a model including 2 harmonics
-    at the longer period $P_{\rm \ell}$, plus 2 harmonics and a transit
-    at the shorter period $P_{\rm s}$.
+    {\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 19).} {\it Top}:
+    The ``raw'' (\texttt{PDCSAP} mean-subtracted) relative flux.
+    The beat period of 4.48 days is visible by eye.  The blue curve is
+    a model including 2 harmonics at the longer period $P_{\rm \ell}$,
+    plus 2 harmonics and a transit at the shorter period $P_{\rm s}$.
     {\it Upper middle}: Longer-period signal, equal to the raw signal
-		minus the shorter-period signal.
-		{\it Lower middle}: Shorter-period signal, equal to the raw signal
-		minus the longer-period signal.
-		{\it Bottom}: residual.  The data are binned from 2 to 10 minute
-		cadence as a convenience for plotting and fitting.
+    minus the shorter-period signal.  {\it Lower middle}:
+    Shorter-period signal, equal to the raw signal minus the
+    longer-period signal.  {\it Bottom}: residual relative flux.  The
+    data are binned from 2 to 10 minute cadence as a convenience for
+    plotting and fitting.
 		\label{fig:splitsignal}
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 20).}
-		Same format as Figure~\ref{fig:splitsignal}.
+  \caption{ {\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 20).} Same
+  format as Figure~\ref{fig:splitsignal}.
   \label{fig:splitsignalii}
@@
 
-\ptfo\ [also known as CVSO\,30; \citealt{briceno_cida_2005}]
-was observed by TESS with Camera 1, CCD 1, from December 15,
-2018 until January 6, 2019, during the sixth sector of science operations
-\citep{ricker_transiting_2015}.  The star was designated TIC\,264461976
-in the TESS Input Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.
-The pixel data for an $11\times11$ array surrounding \ptfo\ were
-averaged into 2-minute stacks by the onboard computer.  Each
-2048$\times$2048 image from the CCD was also averaged into 30-minute
-stacks, and saved as a ``full frame image'' (FFI).
+\ptfo\ [also known as CVSO\,30; \citealt{briceno_cida_2005}] was
+observed by TESS with Camera 1, CCD 1, from December 15, 2018 until
+January 6, 2019, during the sixth sector of science operations
+\citep{ricker_transiting_2015}.  The star was designated
+TIC\,264461976 in the TESS Input Catalog
+\citep{stassun_TIC_2018,stassun_TIC8_2019}.  The pixel data for an
+$11\times11$ array surrounding \ptfo\ were averaged into 2-minute
+stacks by the onboard computer.  Each 2048$\times$2048 image from the
+CCD was also averaged into 30-minute stacks, and saved as a ``full
+frame image'' (FFI).
 
@@
 Science Processing Operations Center (SPOC) at NASA
-Ames~\citep{jenkins_tess_2016}.  We mainly used the
-Presearch Data Conditioning (PDC) lightcurve.  The PDC lightcurve
-aperture used pixels chosen to maximize the SNR of the total flux of
-the target \citep{smith_kepler_apertures_2017}.  Non-astrophysical
-variability was removed by fitting out trends common to many stars
+Ames~\citep{jenkins_tess_2016}.  We mainly used the Presearch Data
+Conditioning (PDC) lightcurve.  The PDC lightcurve aperture used
+pixels chosen to maximize the SNR of the total flux of the target
+\citep{smith_kepler_apertures_2017}.  Non-astrophysical variability
+was removed by fitting out trends common to many stars
 \citep{smith_kepler_PDC_2017}.
 
-As an independent check on the 2-minute SPOC lightcurve, we
-examined the lightcurve based upon 30-minute image stacks which was
-produced as part of the Cluster
-Difference Imaging Photometric Survey (CDIPS;
+As an independent check on the 2-minute SPOC lightcurve, we examined
+the lightcurve based upon 30-minute image stacks which was produced as
+part of the Cluster Difference Imaging Photometric Survey (CDIPS;
 \citealt{bouma_cluster_2019}).  Our CDIPS lightcurve of choice used a
@@
 To clean the data, we removed all points with non-zero quality flags,
-which indicate known problems
-\citep[{\it e.g.},][]{tess_data_product_description_2018}.  We also
-masked out the data from the first and last 6 hours of each orbit, since there are
-often systematic red noise in the photometry during those times.  Both the CDIPS and PDC
-lightcurves showed a clear discontinuous jump in the last few days
-of orbit 20, which seemed likely to be an instrumental systematic effect.  We
-correspondingly masked out the data with timestamps ranging from BJD\,2458488.3 until
-the end of the orbit.  The PDC lightcurve initially had 15{,}678 points.  The
-quality-flag cut removed 854 points; masking the orbit edges removed an
-additional 716 points; and removing the data from the final few days of orbit 20 removed an
-additional 1079 points.  After cleaning, 83\% of the initial flux
-measurements remained.
-
-We normalized the lightcurve by dividing out the median flux, and then opted
-to subtract 1{.}0 to set the median value to zero, which simplified
-subsequent interpretation.  Many
-of these and subsequent processing steps were performed using
+which indicate known problems \citep[{\it
+e.g.},][]{tess_data_product_description_2018}.  We also masked out the
+data from the first and last 6 hours of each orbit, since there are
+often systematic red noise in the photometry during those times.  Both
+the CDIPS and PDC lightcurves showed a clear discontinuous jump in the
+last few days of orbit 20, which seemed likely to be an instrumental
+systematic effect.  We correspondingly masked out the data with
+timestamps ranging from BJD\,2458488.3 until the end of the orbit.
+The PDC lightcurve initially had 15{,}678 points.  The quality-flag
+cut removed 854 points; masking the orbit edges removed an additional
+716 points; and removing the data from the final few days of orbit 20
+removed an additional 1079 points.  After cleaning, 83\% of the
+initial flux measurements remained.
+
+We normalized the lightcurve by dividing out the median flux, and then
+opted to subtract 1{.}0 to set the median value to zero, which
+simplified subsequent interpretation.  Many of these and subsequent
+processing steps were performed using
 \texttt{astrobase}~\citep{bhatti_astrobase_2018}.
@@
 \citep{lindegren_gaiasoln_2018}.  For \ptfo, there were 121 ``good''
-observations, {\it i.e.}, observations that were not strongly down-weighted in
-the astrometric solution.  \ptfo\ was assigned the Gaia DR2 identifier
-3222255959210123904.  It photometric brightness was measured using
-selected bands ($G$, $Rp$, and $Bp$) of the Gaia Radial Velocity
-Spectrometer \citep{cropper_gaia_2018,evans_gaia_2018}.  We accessed
-the pipeline parameters for \ptfo\ using the Gaia
+observations, {\it i.e.}, observations that were not strongly
+down-weighted in the astrometric solution.  \ptfo\ was assigned the
+Gaia DR2 identifier 3222255959210123904.  It photometric brightness
+was measured using selected bands ($G$, $Rp$, and $Bp$) of the Gaia
+Radial Velocity Spectrometer
+\citep{cropper_gaia_2018,evans_gaia_2018}.  We accessed the pipeline
+parameters for \ptfo\ using the Gaia
 archive\footnote{\url{gea.esac.esa.int/archive/}}.
@@
 \citep{briceno_cida_2005,van_eyken_ptf_2012}.  The main novelty is
-that Gaia DR2 reported a 10.3$\sigma$ ``astrometric excess'', indicating
-that the residuals to the best-fitting astrometric model were larger
-than expected based on the statistical uncertainties.
-We comment on the significance and
-interpretation of this excess in Section~\ref{sec:gaia}.
+that Gaia DR2 reported a 10.3$\sigma$ ``astrometric excess'',
+indicating that the residuals to the best-fitting astrometric model
+were larger than expected based on the statistical uncertainties.  We
+comment on the significance and interpretation of this excess in
+Section~\ref{sec:gaia}.
 
@@
 the norm \citep{briceno_lowmassOB_2007}.  The Orion molecular cloud
-complex in particular has numerous subgroups, with ages ranging from 0.5
-to 15$\,$Myr. For an incomplete sampling, see for instance
+complex in particular has numerous subgroups, with ages ranging from
+0.5 to 15$\,$Myr. For an incomplete sampling, see for instance
 \citet{briceno_cida_2005,jeffries_kinematic_2006,briceno_25_2007,kounkel_apogee2_2018}
@@
 \ptfo\ was initially identified as a member of the Orion$\,$OB1a
-sub-association and designated CVSO\,30 by \citet{briceno_cida_2005} through combined
-photometry and spectroscopy.  Later work by \citet{briceno_25_2007}
-clarified that \ptfo\ was in a kinematically distinct subgroup of
-Orion~OB1a, named the ``25$\,$Ori'' group after its brightest
-member. They reported that the 25$\,$Ori group has an isochrone age of 7--10$\,$Myr, and
-a smaller fraction of stars with disks than younger nearby sub-associations
+sub-association and designated CVSO\,30 by \citet{briceno_cida_2005}
+through combined photometry and spectroscopy.  Later work by
+\citet{briceno_25_2007} clarified that \ptfo\ was in a kinematically
+distinct subgroup of Orion~OB1a, named the ``25$\,$Ori'' group after
+its brightest member. They reported that the 25$\,$Ori group has an
+isochrone age of 7--10$\,$Myr, and a smaller fraction of stars with
+disks than younger nearby sub-associations
 \citep{hernandez_spitzer_ob1_2007}.
@@
 position and velocity information to identify subgroups within the
-Orion complex.  From smallest to largest, \ptfo\ was identified
-as a member of the following hierarchical subgroups:
+Orion complex.  From smallest to largest, \ptfo\ was identified as a
+member of the following hierarchical subgroups:
 \begin{equation}
@@
 \end{equation}
-where `$\subset$' means ``is a proper subset
-of''.  25$\,$Ori-1 is the largest subgroup of 25$\,$Ori, with 149
-identified members.  Based on the CMD, the mean age was found to be
-6.9$\,$Myr, while the HR diagram led to an estimated age of 8.5$\,$Myr
-\citep[see][Section 2.3]{kounkel_apogee2_2018}. 
-\citet{kounkel_apogee2_2018} identified
+where `$\subset$' means ``is a proper subset of''.  25$\,$Ori-1 is the
+largest subgroup of 25$\,$Ori, with 149 identified members.  Based on
+the CMD, the mean age was found to be 6.9$\,$Myr, while the HR diagram
+led to an estimated age of 8.5$\,$Myr \citep[see][Section
+2.3]{kounkel_apogee2_2018}.  \citet{kounkel_apogee2_2018} identified
 seven other smaller groups in the Orion complex near the Be star
@@
 These details concerning the group membership for one object are
-cumbersome to those accustomed to the simple distinction between ``young cluster members''
-and ``old field stars''.  Although all members of the Orion complex are
-indeed young relative to the field, these details are essential for
-assessing the photometric evidence for the binarity of \ptfo, because of
-the degeneracy between stellar luminosity and age for
-pre-main-sequence stars.  Having a clean sample of reference stars that are
-tightly associated with \ptfo\ --- both spatially
-and kinematically --- minimizes contamination
-not only from field stars, but also from older and younger members of the
-Orion complex.
+cumbersome to those accustomed to the simple distinction between
+``young cluster members'' and ``old field stars''.  Although all
+members of the Orion complex are indeed young relative to the field,
+these details are essential for assessing the photometric evidence for
+the binarity of \ptfo, because of the degeneracy between stellar
+luminosity and age for pre-main-sequence stars.  Having a clean sample
+of reference stars that are tightly associated with \ptfo\ --- both
+spatially and kinematically --- minimizes contamination not only from
+field stars, but also from older and younger members of the Orion
+complex.
 
@@
 	\caption{ {\bf Phase-folded longer and shorter-period signals.}
-    {\it Top}: Longer-period signal, as in Figure~\ref{fig:splitsignal}.
-    {\it Bottom}: Shorter-period signal. The reference phase is set to
-    the dip.  Gray points are the 10 minute cadence
-    \texttt{PDCSAP} flux.  Black points are binned to 100 points per
-    period.
+    {\it Top}: Relative flux at the longer period, as in
+    Figure~\ref{fig:splitsignal}.  {\it Bottom}: Relative flux at the
+    shorter period. The reference phase is set to the dip.  Gray
+    points are the 10 minute cadence \texttt{PDCSAP} flux.  Black
+    points are binned to 100 points per period.
 		\label{fig:phasefold}
@@
 (Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}, top
-panel). As a precursor to more detailed analysis, we calculated generalized
-Lomb-Scargle periodograms using \texttt{astrobase}
+panel). As a precursor to more detailed analysis, we calculated
+generalized Lomb-Scargle periodograms using \texttt{astrobase}
 \citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
-The two tallest peaks occur at 0.448\,days and 0.499\,days.
-We will refer to these two periods as the ``shorter period'' $P_{\rm s}$
-and the ``longer period'' $P_{\rm ell}$.
-The peak at $P_{\rm \ell}$ is the taller of the two peaks.
-Lower-power harmonics of both signals are also present.
+The two tallest peaks occur at 0.448\,days and 0.499\,days.  We will
+refer to these two periods as the ``shorter period'' $P_{\rm s}$ and
+the ``longer period'' $P_{\rm ell}$.  The peak at $P_{\rm \ell}$ is
+the taller of the two peaks.  Lower-power harmonics of both signals
+are also present.
 
@@
 The peak-to-peak lightcurve amplitude at maximum, when the two signals
-interfere constructively, is about 14\%.  During the times of destructive interference, the peak-to-peak
-amplitude is about 6\%.  Assuming the signals are mainly sinusoidal,
-simple algebra tells us that the peak-to-peak amplitudes should be
-about 10\% for the longer-period signal, and 4\% for the shorter-period signal.
-To view the phase-folded light curves of the longer-period signal,
-we subtracted the best-fitting sinusoid at the shorter period; the resulting
-light curve appears smooth and nearly sinusoidal.
-But after subtracting the best-fitting sinusoid at the longer period,
-visual inspection of the phase-folded light curve of the shorter-period signal
-revealed substructure resembling the ``dips'' seen in previous
-observations. In particular, there was a $\approx$1\% dip lasting about an hour.
-These initial impressions turned out to be consistent with the results of our
-more complicated analysis, described below.
+interfere constructively, is about 14\%.  During the times of
+destructive interference, the peak-to-peak amplitude is about 6\%.
+Assuming the signals are mainly sinusoidal, simple algebra tells us
+that the peak-to-peak amplitudes should be about 10\% for the
+longer-period signal, and 4\% for the shorter-period signal.  To view
+the phase-folded light curves of the longer-period signal, we
+subtracted the best-fitting sinusoid at the shorter period; the
+resulting light curve appears smooth and nearly sinusoidal.  But after
+subtracting the best-fitting sinusoid at the longer period, visual
+inspection of the phase-folded light curve of the shorter-period
+signal revealed substructure resembling the ``dips'' seen in previous
+observations. In particular, there was a $\approx$1\% dip lasting
+about an hour.  These initial impressions turned out to be consistent
+with the results of our more complicated analysis, described below.
 
@@
 
-We fitted a model to the lightcurve consisting of a linear combination of Fourier modes
-with periods $P_{\rm s}$ and $P_{\rm ell}$, as well as a number of harmonics chosen as described below. To try accounting for the dips, we also added an
-analytic transit model with period $P_{\rm s}$.
-Symbolically, the total flux $f$ is given as
+We fitted a model to the lightcurve consisting of a linear combination
+of Fourier modes with periods $P_{\rm s}$ and $P_{\rm ell}$, as well
+as a number of harmonics chosen as described below. To try accounting
+for the dips, we also added an analytic transit model with period
+$P_{\rm s}$.  Symbolically, the total flux $f$ is given as
 \begin{equation}
@@
 \end{equation}
-where $f_{\rm s}$ is the flux at the shorter period, and
-$f_{\rm \ell}$ is the flux at the longer period.  Writing out the
-Fourier terms explicitly,
+where $f_{\rm s}$ is the flux at the shorter period, and $f_{\rm
+\ell}$ is the flux at the longer period.  Writing out the Fourier
+terms explicitly,
 \begin{align}
@@
 \end{align}
-where $N_{\rm s}$ and $N_{\rm \ell}$ are the total number of modes at the shorter and longer
-periods, respectively,
-$A_i$ and $B_i$ are the amplitudes of each
-mode (which can be positive or negative), and $\omega_\ell$
-and $\omega_{\rm s}$ are the angular frequencies of the longer-period
-and shorter-period signals. By not including a phase parameter
-in the shorter-period model, we have implicitly defined the zero point
-of the phase scale. The relative phase of the longer-period model
-is specified by the phase parameter $\phi_\ell$.
-Since we did not know in advance how many harmonics
-would be appropriate to include in the model, we considered a number of different choices for
-$N_{\rm s}$ and $N_{\rm \ell}$, and used the Bayesian information criterion to select
-the final model (Table~1).
+where $N_{\rm s}$ and $N_{\rm \ell}$ are the total number of modes at
+the shorter and longer periods, respectively, $A_i$ and $B_i$ are the
+amplitudes of each mode (which can be positive or negative), and
+$\omega_\ell$ and $\omega_{\rm s}$ are the angular frequencies of the
+longer-period and shorter-period signals. By not including a phase
+parameter in the shorter-period model, we have implicitly defined the
+zero point of the phase scale. The relative phase of the longer-period
+model is specified by the phase parameter $\phi_\ell$.  Since we did
+not know in advance how many harmonics would be appropriate to include
+in the model, we considered a number of different choices for $N_{\rm
+s}$ and $N_{\rm \ell}$, and used the Bayesian information criterion to
+select the final model (Table~1).
 
 As an example, one possible model consists of a transit, $N_{\rm s}=2$
-sines and cosines at the shorter period, plus $N_{\rm \ell}=1$
-sine and cosine at the longer period.  In this case, the free parameters would
-be as follows.  The transit model parameters are the impact
+sines and cosines at the shorter period, plus $N_{\rm \ell}=1$ sine
+and cosine at the longer period.  In this case, the free parameters
+would be as follows.  The transit model parameters are the impact
 parameter, the planet-to-star radius ratio, two quadratic limb
-darkening parameters, the planet's orbital period (set equal to $P_{\rm s}$)
-the time of a particular transit, and the mean flux.  There
-would be $2N_{\rm s}=4$ additional Fourier amplitudes at the shorter period,
-plus $2N_{\rm \ell}=2$ Fourier amplitudes at the longer period, and well as $P_\ell$
-itself and the relative phase $\phi_\ell$.  The total number
-of parameters is 14.
+darkening parameters, the planet's orbital period (set equal to
+$P_{\rm s}$) the time of a particular transit, and the mean flux.
+There would be $2N_{\rm s}=4$ additional Fourier amplitudes at the
+shorter period, plus $2N_{\rm \ell}=2$ Fourier amplitudes at the
+longer period, and well as $P_\ell$ itself and the relative phase
+$\phi_\ell$.  The total number of parameters is 14.
 
@@
 built on \texttt{theano}
-\citep{salvatier_2016_PyMC3,exoplanet:theano}.
-For the Fourier terms,
-we used the default math operators.
-For the exoplanet transit, we
-used the model and derivatives implemented in the \texttt{exoplanet} code
-\citep{exoplanet:exoplanet}.  Our priors are listed in Table~2.  To
-speed up the fitting process, we averaged the 2-minute lightcurve
+\citep{salvatier_2016_PyMC3,exoplanet:theano}.  For the Fourier terms,
+we used the default math operators.  For the exoplanet transit, we
+used the model and derivatives implemented in the \texttt{exoplanet}
+code \citep{exoplanet:exoplanet}.  Our priors are listed in Table~2.
+To speed up the fitting process, we averaged the 2-minute lightcurve
 to 10-minute samples.  We correspondingly scaled the uncertainties in
 the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
-initialized each model with the parameters of the maximum a posteriori (MAP) model.
-We then sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
-\citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
-diagnostic \citep{gelman_inference_1992}.  We tested our ability to
-successfully recover injected parameters using synthetic data before
-fitting the \ptfo\ lightcurves.
+initialized each model with the parameters of the maximum a posteriori
+(MAP) model.  We then sampled using \texttt{PyMC3}'s gradient-based
+No-U-Turn Sampler \citep{hoffman_no-u-turn_2014}, and used $\hat{R}$
+as our convergence diagnostic \citep{gelman_inference_1992}.  We
+tested our ability to successfully recover injected parameters using
+synthetic data before fitting the \ptfo\ lightcurves.
 
@@
 We considered nine models, with the number of modes per frequency
-$N_{\rm s}$ and $N_{\rm \ell}$ ranging from one to three.  To select our preferred model,
-we used the Bayesian information criterion (Table~1).  The model with
-the lowest BIC had two modes at the shorter 10.74$\,$hr period, and
-two modes at the longer 11.96$\,$hr period.  The next-best model had
-an additional harmonic at the longer period ($N_{\rm \ell}=3$), but was otherwise
-identical.  All nine models have reduced $\chi^2$ values ranging between 1.37
-and 1.51, which suggests a plausible though imperfect agreement
-between the data and the model to within the formal uncertainties.
-Table~2 gives the best-fitting parameters for the preferred model,
-which has the lowest BIC value.
+$N_{\rm s}$ and $N_{\rm \ell}$ ranging from one to three.  To select
+our preferred model, we used the Bayesian information criterion
+(Table~1).  The model with the lowest BIC had two modes at the shorter
+10.74$\,$hr period, and two modes at the longer 11.96$\,$hr period.
+The next-best model had an additional harmonic at the longer period
+($N_{\rm \ell}=3$), but was otherwise identical.  All nine models have
+reduced $\chi^2$ values ranging between 1.37 and 1.51, which suggests
+a plausible though imperfect agreement between the data and the model
+to within the formal uncertainties.  Table~2 gives the best-fitting
+parameters for the preferred model, which has the lowest BIC value.
 
@@
 with peak-to-peak amplitude $\approx$10\%.  The $10.74\,{\rm hr}$
-variability is always more complex.  The overall impression is of
-a distorted sinusoidal function,
-with a peak-to-peak
-amplitude of about 4\%.
-The asymmetric sinusoid rises to a maximum near phase 0.25,
-and reaches minimum brightness between phases $-0.5$ and $-0.25$.
-Between phases $-0.5$ and $0.0$ there appears to be complex shorter-timescale variability, 
-including an abrupt jump at phase $\approx -0.33$,
-and ending with a ``dip'' 
-of depth $\approx$1.2\%, lasting $\approx$0.75 hours.
-
-The periodogram of the final residual (Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}
-bottom row) shows a barely significant and poorly-resolved peak at
-$\approx$8 days, consistent with the visual impression of some slower
-trends in the residuals after subtracting the shorter- and longer-period
-signals.
+variability is always more complex.  The overall impression is of a
+distorted sinusoidal function, with a peak-to-peak amplitude of about
+4\%.  The asymmetric sinusoid rises to a maximum near phase 0.25, and
+reaches minimum brightness between phases $-0.5$ and $-0.25$.  Between
+phases $-0.5$ and $0.0$ there appears to be complex shorter-timescale
+variability, including an abrupt jump at phase $\approx -0.33$, and
+ending with a ``dip'' of depth $\approx$1.2\%, lasting $\approx$0.75
+hours.
+
+The periodogram of the final residual (Figures~\ref{fig:splitsignal}
+and~\ref{fig:splitsignalii} bottom row) shows a barely significant and
+poorly-resolved peak at $\approx$8 days, consistent with the visual
+impression of some slower trends in the residuals after subtracting
+the shorter- and longer-period signals.
 
@@
     logarithmic grayscale. The yellow star is the position of \ptfo.
-		Orange crosses are neighboring stars with $T<17$. The
-		\texttt{X} and \texttt{/} hatches show the apertures
-		used to measure the background and target star flux, respectively.
-		{\it Bottom:} Digitized Sky Survey $R$-band image of the same
-		field, with a linear grayscale. The circles show the apertures of radii
-		1, 1.5, and 2.25 pixels used in our blend analysis. To the northwest
-		of \ptfo\ and between the
-		blue and orange circles is ``Star A'',
-		the only star bright and close enough to be contributing
-		to the signal attributed to \ptfo\. However, the pixel-level TESS data
-		showed that any such contamination is negligible (see
+    Orange crosses are neighboring stars with $T<17$. The \texttt{X}
+    and \texttt{/} hatches show the apertures used to measure the
+    background and target star flux, respectively.  {\it Bottom:}
+    Digitized Sky Survey $R$-band image of the same field, with a
+    linear grayscale. The circles show the apertures of radii 1, 1.5,
+    and 2.25 pixels used in our blend analysis. To the northwest of
+    \ptfo\ and between the blue and orange circles is ``Star A'', the
+    only star bright and close enough to be contributing to the signal
+    attributed to \ptfo\. However, the pixel-level TESS data showed
+    that any such contamination is negligible (see
     Section~\ref{subsec:blend}).
@@
 The TESS pixels are $\approx21$'' per side. Before making any
-interpretations, we needed to consider whether light from
-neighboring stars could have contributed to the photometric signal
-we are attributing to \ptfo. The scene is
-shown in Figure~\ref{fig:scene}.  In the upper panels, the pixels used
-to measure the background level in the SPOC lightcurve are indicated
-with `\texttt{X}' hatching, and the pixels used in the final
-lightcurve aperture are shown with `\texttt{/}' hatching.
+interpretations, we needed to consider whether light from neighboring
+stars could have contributed to the photometric signal we are
+attributing to \ptfo. The scene is shown in Figure~\ref{fig:scene}.
+In the upper panels, the pixels used to measure the background level
+in the SPOC lightcurve are indicated with `\texttt{X}' hatching, and
+the pixels used in the final lightcurve aperture are shown with
+`\texttt{/}' hatching.
 
@@
 contaminate the target star is TIC 264461980, with $T=14.8$, which we
-dub ``Star A''.  Star A is 23.6'' northwest of the target, and based on the
-magnitude difference could contribute flux variations as large as 48\% the flux of our
-target star, \ptfo.
-
-The variability of \ptfo\ with a period consistent with $P_{\rm s}$ had already been observed
-based on images with arcsecond resolution.
-Thus, our main concern regarding blending
-was whether 
-the longer-period signal with period $P_{\rm \ell}$ originated from \ptfo,
-or from Star A. We took two approaches to investigate the source of the long-period
-signal.
+dub ``Star A''.  Star A is 23.6'' northwest of the target, and based
+on the magnitude difference could contribute flux variations as large
+as 48\% the flux of our target star, \ptfo.
+
+The variability of \ptfo\ with a period consistent with $P_{\rm s}$
+had already been observed based on images with arcsecond resolution.
+Thus, our main concern regarding blending was whether the
+longer-period signal with period $P_{\rm \ell}$ originated from \ptfo,
+or from Star A. We took two approaches to investigate the source of
+the long-period signal.
 
 First, we examined the CDIPS FFI lightcurves of the target, which are
-available on MAST \citep{bouma_cluster_2019}. Three lightcurves are available,
-based on photometric apertures with a radius of 1, 1.5, or 2.5 pixels. The maximal
-peak-to-peak beat amplitude was the same to within a percent,
-regardless of the size of the photometric
-aperture that was used to create the lightcurve.  If Star A
-were the source of the long-period variability, we would expect the
-peak variability amplitude to be smallest in the 1 pixel aperture,
-based on the separation of the sources (Figure~\ref{fig:scene},
-bottom).  From this test alone, it seems unlikely that Star A is the
-source of the long-period signal.
-
-Second, we examined the 2-minute lightcurve of each individual pixel in the scene,
-using the interactive tools implemented in
+available on MAST \citep{bouma_cluster_2019}. Three lightcurves are
+available, based on photometric apertures with a radius of 1, 1.5, or
+2.5 pixels. The maximal peak-to-peak beat amplitude was the same to
+within a percent, regardless of the size of the photometric aperture
+that was used to create the lightcurve.  If Star A were the source of
+the long-period variability, we would expect the peak variability
+amplitude to be smallest in the 1 pixel aperture, based on the
+separation of the sources (Figure~\ref{fig:scene}, bottom).  From this
+test alone, it seems unlikely that Star A is the source of the
+long-period signal.
+
+Second, we examined the 2-minute lightcurve of each individual pixel
+in the scene, using the interactive tools implemented in
 \texttt{lightkurve} \citep{lightkurve_2018}.  If Star A were the
@@
 nearest to Star A to show a sinusoidal signal with amplitude exceeding
-$10\%$.  The data do not show this pattern.  The pixel directly
-below Star A does not clearly show any sinusoidal variability; the
+$10\%$.  The data do not show this pattern.  The pixel directly below
+Star A does not clearly show any sinusoidal variability; the
 peak-to-peak variability in that pixel is $\lesssim 8\%$.  In
@@
 pixel furthest from Star A that was used in the optimal aperture)
-shows the longer-period sinusoidal variability signal with an amplitude of
-14\%.  We conclude that within the resolution of the Gaia
+shows the longer-period sinusoidal variability signal with an
+amplitude of 14\%.  We conclude that within the resolution of the Gaia
 DR2 source catalog, the $P_{\rm s}$ and $P_{\rm \ell}$ signals
-originate from \ptfo.  Based on the work of \citet{ziegler_measuring_2018}, we can
-surmise that stellar companions with separations wider than $\approx$1'' (349~AU) and
-within $\Delta G \approx 3$ magnitudes of \ptfo\ would have likely
-been detected through this approach.
+originate from \ptfo.  Based on the work of
+\citet{ziegler_measuring_2018}, we can surmise that stellar companions
+with separations wider than $\approx$1'' (349~AU) and within $\Delta G
+\approx 3$ magnitudes of \ptfo\ would have likely been detected
+through this approach.
 
 \citet{van_eyken_ptf_2012} obtained stronger constraints on possible
-stellar companions based on high-resolution imaging with the NIRC2 camera on Keck II.  They reported
-3-$\sigma$ $H$-band magnitude difference limits of 4.3, 6.4, and 8.9 at
-angular separations of 0.25, 0.5, and 1.0 arcseconds (87, 175, and
-349~AU).  They also detected a point source, not present in Gaia DR2,
-7.0 magnitudes fainter than the target, and 1.8$''$ to the north-east.  Due
-to the brightness difference, this point source\footnote{This
-point source was claimed to be a potential planetary-mass object
-\citep{schmidt_direct_2016}. Subsequent analysis of its colors showed
-that it is a background star \citep{lee_evidence_2018}.} cannot be the
-source of our signals.
+stellar companions based on high-resolution imaging with the NIRC2
+camera on Keck II.  They reported 3-$\sigma$ $H$-band magnitude
+difference limits of 4.3, 6.4, and 8.9 at angular separations of 0.25,
+0.5, and 1.0 arcseconds (87, 175, and 349~AU).  They also detected a
+point source, not present in Gaia DR2, 7.0 magnitudes fainter than the
+target, and 1.8$''$ to the north-east.  Due to the brightness
+difference, this point source\footnote{This point source was claimed
+to be a potential planetary-mass object \citep{schmidt_direct_2016}.
+Subsequent analysis of its colors showed that it is a background star
+\citep{lee_evidence_2018}.} cannot be the source of our signals.
 
 While long-baseline radial velocity (RV) measurements could provide
-complementary lower limits on companion masses and separations, the available RV
-data for \ptfo\ have large uncertainties due to the faintness of the star.  The
-RV dataset with the longest timespan in the literature appears to be the 5
-Keck/HIRES measurements acquired over 10 days in April 2011 by
-\citet{van_eyken_ptf_2012}.  The root-mean-squared RV over that 10 day span was
-$1.6\,{\rm km}\,{\rm s}^{-1}$, consistent with the measurement
-precision.  Although \citet{yu_tests_2015} acquired 22 further
-Keck/HIRES spectra on one night in December 2013, those points
-were not reduced to velocities. Further Keck/HIRES
-measurements of this target could confirm or refute the existence of
-possible binary companions.
+complementary lower limits on companion masses and separations, the
+available RV data for \ptfo\ have large uncertainties due to the
+faintness of the star.  The RV dataset with the longest timespan in
+the literature appears to be the 5 Keck/HIRES measurements acquired
+over 10 days in April 2011 by \citet{van_eyken_ptf_2012}.  The
+root-mean-squared RV over that 10 day span was $1.6\,{\rm km}\,{\rm
+s}^{-1}$, consistent with the measurement precision.  Although
+\citet{yu_tests_2015} acquired 22 further Keck/HIRES spectra on one
+night in December 2013, those points were not reduced to velocities.
+Further Keck/HIRES measurements of this target could confirm or refute
+the existence of possible binary companions.
 
@@
     circles) were identified by \citet{kounkel_apogee2_2018}.  The
-    gray circles are stars in the ``neighborhood'', i.e.,
-    non-member stars within 5
-    standard deviations of the mean 25$\,$Ori-1 right ascension,
-    declination, and parallax.  It contains members of the Orion
-    complex with its full spread of ages, in addition to field
+    gray circles are stars in the ``neighborhood'', i.e., non-member
+    stars within 5 standard deviations of the mean 25$\,$Ori-1 right
+    ascension, declination, and parallax.  It contains members of the
+    Orion complex with its full spread of ages, in addition to field
     interlopers.  $G$ denotes Gaia broadband magnitudes, $Bp$ Gaia
     blue, $Rp$ Gaia red, and $\omega_{\rm as}$ the parallax in
-    arcseconds.  The $x$-axis limits are set to show only K and
-    M dwarfs, to accentuate \ptfo's separation from the
-    single-star sequence.  {\it Bottom:} Astrometric goodness-of-fit
-    versus $Rp$ magnitude for 25$\,$Ori-1 members.  The single-source
-    astrometric model for \ptfo\ provides a poor fit, which could be due to
-    stellar variability or binarity.  Cluster members that
-    are at least as variable as \ptfo\ show lower astrometric excesses
-    (black squares), suggesting binarity as the root cause.
+    arcseconds.  The $x$-axis limits are set to show only K and M
+    dwarfs, to accentuate \ptfo's separation from the single-star
+    sequence.  {\it Bottom:} Astrometric goodness-of-fit versus $Rp$
+    magnitude for 25$\,$Ori-1 members.  The single-source astrometric
+    model for \ptfo\ provides a poor fit, which could be due to
+    stellar variability or binarity.  Cluster members that are at
+    least as variable as \ptfo\ show lower astrometric excesses (black
+    squares), suggesting binarity as the root cause.
 		\label{fig:gaia}
@@
 is too bright to be a single star, i.e., if it is a ``photometric
-binary.''
-To assemble a set of stars coeval with \ptfo,
-we used the 25$\,$Ori-1 members identified by
-\citet{kounkel_apogee2_2018}, and discussed in
-Section~\ref{subsec:hierarchical}.
-
-To define a set of non-member stars that nonetheless are subject
-to similar selection criteria,
-we defined the reference ``neighborhood'' as the
-group of at most $10^4$ randomly selected non-member stars within 5
-standard deviations of the mean 25$\,$Ori-1 right ascension,
+binary.'' To assemble a set of stars coeval with \ptfo, we used the
+25$\,$Ori-1 members identified by \citet{kounkel_apogee2_2018}, and
+discussed in Section~\ref{subsec:hierarchical}.
+
+To define a set of non-member stars that nonetheless are subject to
+similar selection criteria, we defined the reference ``neighborhood''
+as the group of at most $10^4$ randomly selected non-member stars
+within 5 standard deviations of the mean 25$\,$Ori-1 right ascension,
 declination, and parallax.  We queried Gaia DR2 for these stars using
@@
 
-We examined the resulting five-dimensional distribution of right ascension,
-declination, proper motion in both directions, and parallax.  The first point we noted
-was that 25$\,$Ori-1 is a clearly defined over-density in each
-dimension: the cluster was confirmed to exist, and to be distinct from the neighborhood. \ptfo\ was also within the cluster in each of these projected
-dimensions.
-
-Figure~\ref{fig:gaia} shows
-the HR diagram we constructed from the data.  The diagram shows
-that \ptfo\ is $\approx$0.75 magnitudes brighter than the average
-25$\,$Ori-1 star of the same color.  In other words, it is about twice
-as bright as expected for a single star in the cluster.
-It also seems to be part of a ``photometric binary''
+We examined the resulting five-dimensional distribution of right
+ascension, declination, proper motion in both directions, and
+parallax.  The first point we noted was that 25$\,$Ori-1 is a clearly
+defined over-density in each dimension: the cluster was confirmed to
+exist, and to be distinct from the neighborhood. \ptfo\ was also
+within the cluster in each of these projected dimensions.
+
+Figure~\ref{fig:gaia} shows the HR diagram we constructed from the
+data.  The diagram shows that \ptfo\ is $\approx$0.75 magnitudes
+brighter than the average 25$\,$Ori-1 star of the same color.  In
+other words, it is about twice as bright as expected for a single star
+in the cluster.  It also seems to be part of a ``photometric binary''
 track that runs parallel to the main track.
@@
 \ptfo\ is a binary with two components of nearly equal brightness.
-Since there is no other reason to
-suspect an age difference, and because the source showed two separate
-photometric signals with similar but distinct periods,
-the binary interpretation seems more probable.
+Since there is no other reason to suspect an age difference, and
+because the source showed two separate photometric signals with
+similar but distinct periods, the binary interpretation seems more
+probable.
 
@@
 astrometry.  As noted in Section~\ref{sec:observations}, the Gaia DR2
-astrometric solution for \ptfo\ shows a 10.3$\sigma$ ``astrometric excess'',
-a parameter that quantifies the degree to which a single-star model
-fails to fit the astrometric measurements.  Specifically,
-the single-source astrometric model yielded $\chi^2=325.2$.  There are
-121 astrometric measurements, and 5 free parameters, and therefore 116
-degrees of freedom. The reduced $\chi^2$ is 2.80.  The majority of
-stars with comparable brightness in Gaia do not show such poor
-goodness-of-fit \citep[see][Appendix A]{lindegren_gaiasoln_2018}.
+astrometric solution for \ptfo\ shows a 10.3$\sigma$ ``astrometric
+excess'', a parameter that quantifies the degree to which a
+single-star model fails to fit the astrometric measurements.
+Specifically, the single-source astrometric model yielded
+$\chi^2=325.2$.  There are 121 astrometric measurements, and 5 free
+parameters, and therefore 116 degrees of freedom. The reduced $\chi^2$
+is 2.80.  The majority of stars with comparable brightness in Gaia do
+not show such poor goodness-of-fit \citep[see][Appendix
+A]{lindegren_gaiasoln_2018}.
 
@@
 CDIPS lightcurve database at MAST \citep{bouma_cluster_2019} to find
-the subset of members that were at least as variable as \ptfo.
-We measured the variability amplitude by taking the difference between the
-$95^{\rm th}$ and $5^{\rm th}$ percentiles of the flux measurements.
-This yielded 30 stars of equal or greater variability.
-The lower panel of Figure~\ref{fig:gaia} shows the
-reduced $\chi^2$ as a function of stellar brightness.  \ptfo\ is in
-the upper 90$^{\rm th}$ percentile of stars showing astrometric
-excesses within the 25$\,$Ori-1 group.  Relative to other M-dwarf
-group members with comparable brightnesses and variability
-characteristics, \ptfo\ still stands out by virtue of its failure
-to conform to a single-star astrometric model. This supports
-the interpretation that \ptfo\ is a binary star.
-
-We will have to wait for the full release of the
-nominal Gaia mission to determine definitively whether the astrometric
-excess is caused by stellar binarity or photometric variability.
-Nonetheless the fact that comparably variable stars do not show
-comparably large astrometric excesses suggests that stellar binarity
-is indeed the root cause.
+the subset of members that were at least as variable as \ptfo.  We
+measured the variability amplitude by taking the difference between
+the $95^{\rm th}$ and $5^{\rm th}$ percentiles of the flux
+measurements.  This yielded 30 stars of equal or greater variability.
+The lower panel of Figure~\ref{fig:gaia} shows the reduced $\chi^2$ as
+a function of stellar brightness.  \ptfo\ is in the upper 90$^{\rm
+th}$ percentile of stars showing astrometric excesses within the
+25$\,$Ori-1 group.  Relative to other M-dwarf group members with
+comparable brightnesses and variability characteristics, \ptfo\ still
+stands out by virtue of its failure to conform to a single-star
+astrometric model. This supports the interpretation that \ptfo\ is a
+binary star.
+
+We will have to wait for the full release of the nominal Gaia mission
+to determine definitively whether the astrometric excess is caused by
+stellar binarity or photometric variability.  Nonetheless the fact
+that comparably variable stars do not show comparably large
+astrometric excesses suggests that stellar binarity is indeed the root
+cause.
 
@@
 
-The standard interpretation for 11.96$\,$hour nearly sinusoidal modulations
-of a pre-main-sequence M dwarf is stellar rotation.  This is the
-dominant signal in the system with 10\% amplitude, and there is no
-evidence to suggest that this signal has any other origin.
-
-In their report on the discovery of the unusual
-photometric variability, \citet{van_eyken_ptf_2012} saw an alias of the
-longer-period signal ({\it e.g.}, their Figure~7),
-and identified it as a
-periodogram peak at $0.9985 \pm 0.0061\,$days. They ascribed it to
-their observing cadence, because of its close correspondence to the
-sidereal day.
-Our
-pixel-level analysis showed that the signal is specific to only pixels
-near \ptfo, and no other pixels.  We therefore conclude that the
-signal is not an artifact of systematic errors.
+The standard interpretation for 11.96$\,$hour nearly sinusoidal
+modulations of a pre-main-sequence M dwarf is stellar rotation.  This
+is the dominant signal in the system with 10\% amplitude, and there is
+no evidence to suggest that this signal has any other origin.
+
+In their report on the discovery of the unusual photometric
+variability, \citet{van_eyken_ptf_2012} saw an alias of the
+longer-period signal ({\it e.g.}, their Figure~7), and identified it
+as a periodogram peak at $0.9985 \pm 0.0061\,$days. They ascribed it
+to their observing cadence, because of its close correspondence to the
+sidereal day.  Our pixel-level analysis showed that the signal is
+specific to only pixels near \ptfo, and no other pixels.  We therefore
+conclude that the signal is not an artifact of systematic errors.
 
@@
 	\caption{
-		{\bf Timing residuals for \ptfob\ based on a decade of monitoring.}
-		Black points are times of dips, minus the indicated linear
-		ephemeris.  The phase of the shorter-period signal
-		is plotted on the $y$-axis. The star symbol represents
-		the TESS ephemeris.
-		Dips were observed by \citet{van_eyken_ptf_2012},
+    {\bf Timing residuals for \ptfob\ based on a decade of
+    monitoring.} Black points are times of dips, minus the indicated
+    linear ephemeris.  The phase of the shorter-period signal is
+    plotted on the $y$-axis. The star symbol represents the TESS
+    ephemeris.  Dips were observed by \citet{van_eyken_ptf_2012},
     \citet{ciardi_followup_2015}, \citet{yu_tests_2015},
@@
     at phase 0 in mid-2019) are consistent with noise, and were likely
-		reported because dips were expected, rather than
-		convincingly observed.  Horizontal dashed lines are drawn at
-		$\pm (P_{\rm \ell} - P_{\rm s})/P_{\rm s}$, highlighting either a
-		numerical coincidence or an observational bias.  The orbital phase
-		observed by TESS (lower-right) is consistent with that of
+    reported because dips were expected, rather than convincingly
+    observed.  Horizontal dashed lines are drawn at $\pm (P_{\rm \ell}
+    - P_{\rm s})/P_{\rm s}$, highlighting either a numerical
+    coincidence or an observational bias.  The orbital phase observed
+    by TESS (lower-right) is consistent with that of
     \citet{tanimoto_evidence_2020}.
@@
 
-One particularly interesting feature of the dip is its epoch.  The dips do
-not occur with strict periodicity
-\citep{yu_tests_2015}.  In fact, \citet{tanimoto_evidence_2020}
-provided stark evidence for different behavior altogether:
-over a time-span of years, the dip ``split'' into distinct groups at
-particular repeating phases.  See for instance their Figures~2
-through~4.  Fitting a decade of observations, they provided the
-following linear ephemeris, which we did not find any need to update.
+One particularly interesting feature of the dip is its epoch.  The
+dips do not occur with strict periodicity \citep{yu_tests_2015}.  In
+fact, \citet{tanimoto_evidence_2020} provided stark evidence for
+different behavior altogether: over a time-span of years, the dip
+``split'' into distinct groups at particular repeating phases.  See
+for instance their Figures~2 through~4.  Fitting a decade of
+observations, they provided the following linear ephemeris, which we
+did not find any need to update.
 \begin{align}
@@
 
-Figure~\ref{fig:o_minus_c} shows the phase of the dip we detected
-in the TESS data, relative to their ephemeris.  It agrees with the
+Figure~\ref{fig:o_minus_c} shows the phase of the dip we detected in
+the TESS data, relative to their ephemeris.  It agrees with the
 independent December 2018 measurements by
@@
 of multiple dips per cycle in December 2018 at phase 0 and $-0{.}47$
-seems less plausible---the phase $\text{-}0{.}47$ dips are strongly detected,
-while the suggested phase 0 dip is not clearly present in the data.
+seems less plausible---the phase $\text{-}0{.}47$ dips are strongly
+detected, while the suggested phase 0 dip is not clearly present in
+the data.
 
@@
 that also peak at quadrature \citep[see][]{shporer_astrophysics_2017}.
-Interpreting the second cosine harmonic as planet-induced tidal distortion,
-it would imply a minimum planet mass $M_{\rm p} \sin i$ of
+Interpreting the second cosine harmonic as planet-induced tidal
+distortion, it would imply a minimum planet mass $M_{\rm p} \sin i$ of
 $3.8\,M_{\rm Jup}$.  For this estimate, we assumed $R_\star = 1.39
 R_\odot$, and $M_\star = 0.39 M_\odot$ \citep{van_eyken_ptf_2012}.
-This ellipsoidal amplitude is larger than the typical
-modulations induced by close-in giant planets because the host star is
-puffy, and still on the pre-main-sequence.
+This ellipsoidal amplitude is larger than the typical modulations
+induced by close-in giant planets because the host star is puffy, and
+still on the pre-main-sequence.
 
@@
 ($0.86\,M_\odot$).  Interpreting the cosine component as reflected or
-emitted light from the planet's surface is nonsensical because the sign is wrong---the
-planet would need to be {\it absorbing} light.
+emitted light from the planet's surface is nonsensical because the
+sign is wrong---the planet would need to be {\it absorbing} light.
 
@@
     EPIC 205483258 (RIK-210).
-    %and EPIC 204787516.
-    RIK-210 has the longest period of any of these objects. 
-    All the analogs displayed are either in Taurus or Upper~Sco, and
-    meet the characteristics of Section~\ref{subsec:dipstars}.
-    % EPIC 204787516 has a less angular dip, and could also be an
-    % eclipsing binary.
-    % All stars have similar properties as \ptfo, but are in Upper~Sco.
-    We found these objects through studies by
-    \citet{stauffer_orbiting_2017}, \citet{david_transient_2017}, and
-    \citet{rebull_usco_2018}.
+    RIK-210 has the longest period of any of these objects.  All the
+    analogs displayed are either in Taurus or Upper~Sco, and meet the
+    characteristics of Section~\ref{subsec:dipstars}.  We found these
+    objects through studies by \citet{stauffer_orbiting_2017},
+    \citet{david_transient_2017}, and \citet{rebull_usco_2018}.
 		\label{fig:brethren}
@@
 
-When physical explanations are not forthcoming, taxonomy is a
-useful exercise.  By searching the literature, we have found about a dozen lightcurves with
-similar morphologies to \ptfo, drawn from surveys of low-mass weak-lined
-T Tauri stars in regions including $\rho$ Oph, Upper~Sco, Taurus, and
-perhaps the Pleiades
+When physical explanations are not forthcoming, taxonomy is a useful
+exercise.  By searching the literature, we have found about a dozen
+lightcurves with similar morphologies to \ptfo, drawn from surveys of
+low-mass weak-lined T Tauri stars in regions including $\rho$ Oph,
+Upper~Sco, Taurus, and perhaps the Pleiades
 \citep{rebull_rotation_2016,david_transient_2017,stauffer_orbiting_2017,stauffer_rotevol_2018,rebull_usco_2018,rebull_rotation_2020}.
 These surveys were performed using K2 \citep{howell_k2_2014}.  We
-downloaded some of these lightcurves from MAST, opting for the
-EVEREST reductions \citep{luger_everest_2016,luger_update_2018}.
-They are plotted in Figure~\ref{fig:brethren}.
+downloaded some of these lightcurves from MAST, opting for the EVEREST
+reductions \citep{luger_everest_2016,luger_update_2018}.  They are
+plotted in Figure~\ref{fig:brethren}.
 
@@
 their depths and durations show variability over the 90-day K2
-campaigns \citep{stauffer_orbiting_2017}.
-In the terminology of \citet{stauffer_orbiting_2017},
-these objects are morphologically distinct from ``scallop shell'' lightcurves,
-and are present in stars at more advanced evolutionary disk stages than
-the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
-The persistent and transient flux dip stars all show angular dips
-that are cannot be explained as the effects of starspots.
-These stars typically have the following things in common:
+campaigns \citep{stauffer_orbiting_2017}.  In the terminology of
+\citet{stauffer_orbiting_2017}, these objects are morphologically
+distinct from ``scallop shell'' lightcurves, and are present in stars
+at more advanced evolutionary disk stages than the ``dipper'' stars
+\citep{ansdell_young_2016,cody_manyfaceted_2018}.  The persistent and
+transient flux dip stars all show angular dips that are cannot be
+explained as the effects of starspots.  These stars typically have the
+following things in common:
 \begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
@@
   \item The age is $\lesssim$ 100$\,$Myr.
-  \item The lightcurves show shallow, angular dips, usually
-    superposed on large-amplitude smooth variability. The latter
-    is interpreted as stellar rotation.
+  \item The lightcurves show shallow, angular dips, usually superposed
+    on large-amplitude smooth variability. The latter is interpreted
+    as stellar rotation.
   \item The rotation is rapid---usually between 0.5 and 2.0 days.
@@
   \item They sometimes show multiple dips per cycle.
-  \item The dip depths, durations, and phases can vary over just a few cycles
-    ({\it e.g.}, EPIC 204143627).
+  \item The dip depths, durations, and phases can vary over just a few
+    cycles ({\it e.g.}, EPIC 204143627).
   \item The dip depths can change after flares.
@@
 \end{enumerate}
-The 10.74 hour signal in \ptfo\ meets all of these criteria.
-This is the first connection of \ptfo\ with this class of objects
-because the TESS lightcurve was needed to resolve the different
-rotation signals.
+The 10.74 hour signal in \ptfo\ meets all of these criteria.  This is
+the first connection of \ptfo\ with this class of objects because the
+TESS lightcurve was needed to resolve the different rotation signals.
 
 There are two crucial additional points concerning the transient flux
-dips.  First, the dip durations seem to scale linearly with
-the photometric periods \citep[][Figure~26]{stauffer_orbiting_2017}.  In the usual
-idealized limits, the transit duration $T$ of a point source across
-the stellar disk scales as $T \propto R_\star (P/M_\star)^{1/3}$
-\citep{winn_exoplanet_2010}.  While the shortest period
-$\approx$0.5-day transient flux dip stars have dip durations
+dips.  First, the dip durations seem to scale linearly with the
+photometric periods \citep[][Figure~26]{stauffer_orbiting_2017}.  In
+the usual idealized limits, the transit duration $T$ of a point source
+across the stellar disk scales as $T \propto R_\star
+(P/M_\star)^{1/3}$ \citep{winn_exoplanet_2010}.  While the shortest
+period $\approx$0.5-day transient flux dip stars have dip durations
 consistent with point sources, at longer periods of 1 to 5 days the
@@
 Second, between 40-50\% of the transient flux dip stars discovered in
-$\rho$~Oph and Upper~Sco show two
-Lomb-Scargle periods, and so are apparently binaries
-\citep[][Table~1]{stauffer_orbiting_2017}.
-This is higher than the main-sequence companion
-fraction of ${\rm CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$
-\citep{henry_solar_2006,duchene_stellar_2013}.
-Low-mass pre-main-sequence stars however have been shown to 
-companion fractions up to twice as high in dispersed clusters such as
-Upper~Sco and Taurus \citep{kraus_mapping_2008,kraus_mapping_2011}.
-A detailed high-resolution imaging survey would be necessary to determine
-whether the transient flux dip stars truly have any distinct
-population-level binarity properties relative to other young low-mass
-stars.
+$\rho$~Oph and Upper~Sco show two Lomb-Scargle periods, and so are
+apparently binaries \citep[][Table~1]{stauffer_orbiting_2017}.  This
+is higher than the main-sequence companion fraction of ${\rm
+CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$
+\citep{henry_solar_2006,duchene_stellar_2013}.  Low-mass
+pre-main-sequence stars however have been shown to companion fractions
+up to twice as high in dispersed clusters such as Upper~Sco and Taurus
+\citep{kraus_mapping_2008,kraus_mapping_2011}.  A detailed
+high-resolution imaging survey would be necessary to determine whether
+the transient flux dip stars truly have any distinct population-level
+binarity properties relative to other young low-mass stars.
 
@@
 photometrically twice as bright as stars of the same color in its
-kinematic group (Figure~\ref{fig:gaia}).  Second, it shows two distinct
-photometric signals.  These points alone suggest binarity
+kinematic group (Figure~\ref{fig:gaia}).  Second, it shows two
+distinct photometric signals.  These points alone suggest binarity
 \citep{stauffer_rotevol_2018}.  For the case of \ptfo, there is a
 third line of evidence: the Gaia DR2 entry for \ptfo\ reports a poor
-fit of the single-star model to the astrometric data.  While this could be caused
-by stellar variability, other cluster members
-that are just as variable do not typically show the same level of
-excess astrometric motion.
-Therefore the astrometric excess is a suggestive third line
-of evidence for binarity in \ptfo. To us, the evidence leads to
-the conclusion that \ptfo\ is a nearly equal-mass binary consisting
-of two rapidly rotating stars.
+fit of the single-star model to the astrometric data.  While this
+could be caused by stellar variability, other cluster members that are
+just as variable do not typically show the same level of excess
+astrometric motion.  Therefore the astrometric excess is a suggestive
+third line of evidence for binarity in \ptfo. To us, the evidence
+leads to the conclusion that \ptfo\ is a nearly equal-mass binary
+consisting of two rapidly rotating stars.
 
@@
 
-The main physical question is what causes the
-transient dips. This is an unsolved problem not only for \ptfo\ but
-also for an emerging class of similar young rapidly rotating M-dwarfs.
-Many possible explanations been detailed by \citet{rebull_rotation_2016}, \citet{david_transient_2017},
-\citet{stauffer_orbiting_2017}, and \citet{zhan_complex_2019}.
-Their disfavored explanations include that the dips are caused by
+The main physical question is what causes the transient dips. This is
+an unsolved problem not only for \ptfo\ but also for an emerging class
+of similar young rapidly rotating M-dwarfs.  Many possible
+explanations been detailed by \citet{rebull_rotation_2016},
+\citet{david_transient_2017}, \citet{stauffer_orbiting_2017}, and
+\citet{zhan_complex_2019}.  Their disfavored explanations include that
+the dips are caused by
 {\it (i)} eclipsing binaries;
@@
 {\it (vi)} dust clouds of plausible composition.
-We also view the possibility of {\it (vii)} tidally disrupted planetary
-or cometary material to be implausible, given the synchronicity
-between dip and rotation periods seen across many systems.
-
-The explanations that are not yet ruled out include
-{\it (i)} transiting clumps of gas at the Keplerian corotation radius;
-{\it (ii)} transits of enshrouded protoplanets;
-{\it (iii)} occultations of starspots by an optically thick disk.
-The first and last explanations have added appeal because they are
-flexible enough to explain not only the transient and persistent-dip
-M-dwarfs, but also the ``scallop shell'' M-dwarfs
-\citep{stauffer_orbiting_2017}.
-Despite this appeal, the possibility of distinct mechanisms explaining
-these distinct variability classes remains open.
-
-The evolution of \ptfo\
-over the past decade (Figure~\ref{fig:o_minus_c}) could offer important
-hints.  Specifically, \ptfo's transition between having none, one, and
+We also view the possibility of {\it (vii)} tidally disrupted
+planetary or cometary material to be implausible, given the
+synchronicity between dip and rotation periods seen across many
+systems.
+
+The explanations that are not yet ruled out include {\it (i)}
+transiting clumps of gas at the Keplerian corotation radius; {\it
+(ii)} transits of enshrouded protoplanets; {\it (iii)} occultations of
+starspots by an optically thick disk.  The first and last explanations
+have added appeal because they are flexible enough to explain not only
+the transient and persistent-dip M-dwarfs, but also the ``scallop
+shell'' M-dwarfs \citep{stauffer_orbiting_2017}.  Despite this appeal,
+the possibility of distinct mechanisms explaining these distinct
+variability classes remains open.
+
+The evolution of \ptfo\ over the past decade
+(Figure~\ref{fig:o_minus_c}) could offer important hints.
+Specifically, \ptfo's transition between having none, one, and
 multiple dips per cycle seems important.  It strains the ``enshrouded
@@
     beat, and therefore cannot be an artifact linked to data
-    processing.
-    Within the angular resolution of the Gaia source
+    processing.  Within the angular resolution of the Gaia source
     catalog, both signals need to originate from \ptfo.
@@
     (Figure~\ref{fig:gaia}, top).  Relative to stars in its group that
-    are at least as photometrically variable, \ptfo\ also shows
-    signs of astrometric binarity (Figure~\ref{fig:gaia}, bottom).
+    are at least as photometrically variable, \ptfo\ also shows signs
+    of astrometric binarity (Figure~\ref{fig:gaia}, bottom).
   \item {\it The orbital phase of the dip has changed since the
@@
 The physical mechanism that explains the transient and persistent flux
-dips is unresolved. Our preferred explanations include
-transiting clumps of gas at the Keplerian corotation radius, and
-occultations of starspots by an optically thick disk \citep[{\it
+dips is unresolved. Our preferred explanations include transiting
+clumps of gas at the Keplerian corotation radius, and occultations of
+starspots by an optically thick disk \citep[{\it
 e.g.},][]{stauffer_orbiting_2017,david_transient_2017,zhan_complex_2019}.
-The jumping orbital phase disfavors the explanation of an
-enshrouded, transiting protoplanet.  Though \ptfob\ may not be a planet,
-as we and others had hoped, understanding \ptfo\ and its analogs is a worthy
+The jumping orbital phase disfavors the explanation of an enshrouded,
+transiting protoplanet.  Though \ptfob\ may not be a planet, as we and
+others had hoped, understanding \ptfo\ and its analogs is a worthy
 problem that might even teach us about the natal environments of the

### 2020-04-27 billy/plotting.py
look into RUWE

@@ -861,24 +861,32 @@ def plot_hr(outdir):
     savefig(f, outpath)
 
 
-def plot_astrometric_excess(outdir):
+def plot_astrometric_excess(outdir, ruwe=0):
 
     varamppath = '/Users/luke/Dropbox/proj/billy/data/25ori-1/var_amps.csv'
     va_df = pd.read_csv(varamppath)
 
-    pklpath = '/Users/luke/Dropbox/proj/billy/results/cluster_membership/nbhd_info_3222255959210123904.pkl'
-    d = pickle.load(open(pklpath, 'rb'))
-    g_df = d[2]
+    if not ruwe:
+        pklpath = '/Users/luke/Dropbox/proj/billy/results/cluster_membership/nbhd_info_3222255959210123904.pkl'
+        d = pickle.load(open(pklpath, 'rb'))
+        g_df = d[2]
+    else:
+        csvpath = '/Users/luke/Dropbox/proj/billy/results/cluster_membership/25ori-1_group_ruwe.csv'
+        g_df = pd.read_csv(csvpath)
 
     g_df = g_df.merge(va_df, on='source_id', how='left')
 
     chisq_red = g_df.astrometric_chi2_al / (g_df.astrometric_n_obs_al - 5)
+    if not ruwe:
+        yval = chisq_red
+    else:
+        yval = np.array(g_df.ruwe)
 
     ######
 
     f,ax = plt.subplots(figsize=(4,3))
 
-    ax.scatter(g_df.phot_rp_mean_mag, chisq_red, c='gray', zorder=3, s=3,
+    ax.scatter(g_df.phot_rp_mean_mag, yval, c='gray', zorder=3, s=3,
                marker='s', rasterized=True, linewidths=0, label='All members',
                alpha=1)
 
@@ -886,27 +894,39 @@ def plot_astrometric_excess(outdir):
     ptfo_amp = float(g_df[ptfosel].var_amp)
     sel = (g_df.var_amp > ptfo_amp)
 
-    ax.scatter(g_df[sel].phot_rp_mean_mag, chisq_red[sel], c='k', alpha=1.,
+    ax.scatter(g_df[sel].phot_rp_mean_mag, yval[sel], c='k', alpha=1.,
                zorder=4, s=12,
                rasterized=True, linewidths=0, marker='s',
                label='Amplitude > {:.1f}%'.format(ptfo_amp*100))
 
     ax.plot(
         g_df[ptfosel].phot_rp_mean_mag,
-        chisq_red[ptfosel],
[…]

### 2020-04-27 drivers/plot_astrometric_excess.py

@@ -5,4 +5,5 @@ from billy import __path__
 RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
 PLOTDIR = os.path.join(RESULTSDIR, 'cluster_membership')
 
-bp.plot_astrometric_excess(PLOTDIR)
+bp.plot_astrometric_excess(PLOTDIR, ruwe=0)
+bp.plot_astrometric_excess(PLOTDIR, ruwe=1)

### 2020-04-27 paper/ms.tex
add words to describe RUWE test

@@
 binary star.
+
+Performing the same analysis using the renormalized unit weight error
+(RUWE\footnote{ See the Gaia DPAC technical note
+GAIA-C3-TN-LU-LL-124-01,
+\url{http://www.rssd.esa.int/doc_fetch.php?id=3757412}, accessed
+2020-04-27. }) rather than the reduced $\chi^2$ yielded similar
+results.  \ptfo\ has a RUWE of 1{.}22, which corresponds to the
+93$^{\rm rd}$ percentile of 25$\,$Ori-1 members.  Two of thirty stars
+with variability amplitudes greater than 9.7\% showed higher RUWE.
+One was CVSO~35, which has a TESS light-curve that varies by 2
+magnitudes, and shows a strong IR excess and a 10$\mu$m silicate emission
+feature \citep{mauco_herschel_2018}.  The other is GAIA DR2
+3222210363837122048.
 

### 2020-04-27 billy/plotting.py

@@ -914,7 +914,14 @@ def plot_astrometric_excess(outdir, ruwe=0):
         ax.set_ylabel('Renormalized Unit Weight Error', fontsize='large')
         ax.set_ylim([0.7,1.8])
         print('WRN! ylimits omit')
-        print( g_df[yval>2].source_id )
+        print( g_df[(yval>1.2) & (g_df.var_amp > 0.097)].source_id )
+        print(42*'-')
+        print(yval[ptfosel])
+        _df = pd.DataFrame({'ruwe':yval})
+        print(_df.describe())
+        for p in np.arange(75, 100, 2.5):
+           print('{}: {}'.format(p, float(_df.quantile(q=p/100))))
+
         # NOTE: this ylimit omits one variable member, 
         #  Gaia DR2 3222267297922229248 = CVSO 35
         # which has an IR excess and 10 micron silicate emission (Macuo+2018)

### 2020-04-27 billy/convenience.py
minor analysis tweaks

@@ -175,7 +175,8 @@ def initialize_ptfo_prior_d(x_obs, modelcomponents):
     t0_orb = 2458468.63809577 - 2457000 - 1468.2  # non-zero means phase offset.
 
     P_rot = 0.49845 # +/- 1e-3
-    t0_rot = 0.14 # np.nanmin(x_obs) + 0.1 # initial guess
+    t0_rot = 0.14
+    # t0_rot = 0.14 + 0.49845/8 # 0.14 # np.nanmin(x_obs) + 0.1 # initial guess
 
     prior_d = OrderedDict()
 

### 2020-04-27 billy/modelfitter.py

@@ -45,9 +45,9 @@ class ModelParser:
 
 class ModelFitter(ModelParser):
     """
-    Given a modelid of the form "transit_NsincosPorb_NsincosProt", and observed
-    x and y values (typically time and flux), construct and fit the model. In
-    other words, run the inference.
+    Given a modelid of the form "transit_NsincosPorb_NsincosProt", and
+    observed x and y values (typically time and flux), construct and
+    fit the model. In other words, run the inference.
 
     The model implemented is of the form
 
@@ -58,8 +58,8 @@ class ModelFitter(ModelParser):
     σ^2).
     """
 
-    def __init__(self, modelid, x_obs, y_obs, y_err, prior_d,
-                 mstar=1, rstar=1, N_samples=1000, N_cores=16, N_chains=4,
+    def __init__(self, modelid, x_obs, y_obs, y_err, prior_d, mstar=1,
+                 rstar=1, N_samples=1000, N_cores=16, N_chains=3,
                  plotdir=None, pklpath=None, overwrite=1):
 
         self.N_samples = N_samples
@@ -330,8 +330,9 @@ class ModelFitter(ModelParser):
 
             # sample from the posterior defined by this model.
             trace = pm.sample(
-                tune=self.N_samples, draws=self.N_samples, start=map_estimate,
-                cores=self.N_cores, chains=self.N_chains,
+                tune=self.N_samples, draws=self.N_samples,
+                start=map_estimate, cores=self.N_cores,
+                chains=self.N_chains,
                 step=xo.get_dense_nuts_step(target_accept=0.9),
             )
 

### 2020-04-27 billy/plotting.py

@@ -284,7 +284,7 @@ def plot_splitsignal_map_periodogram(ydict, outpath):
         if not k == 'y_resid':
             power = ls.power(frequency)
         else:
-            _period_min, _period_max, _N_freqs = 1, 20, int(2e4)
+            _period_min, _period_max, _N_freqs = 0.3, 20, int(4e4)
             _frequency = np.linspace(1/_period_max, 1/_period_min, _N_freqs)
             _period = 1/_frequency
             power = ls.power(_frequency)

### 2020-04-28 paper/ms.tex
paper updates

@@
 \submitjournal{AAS journals.}
-\shorttitle{Seeing Double: PTFO$\,$8-8695}
+\shorttitle{Two Stars, Two Signals, No Planet}
 
@@
 The authors thank D.~Fabrycky for helpful tidal evolution calculations
-that were eventually omitted.
+that were eventually omitted, and A.~Vanderburg for the suggestion to
+consider RUWE.
 %

### 2020-04-28 billy/modelfitter.py
hacky fix for matplotlib not being threadsafe
see https://github.com/matplotlib/matplotlib/issues/15410

matplotlib's scheme for cacheing fonts in memory leads apparently random
child processes to crash the PyMC3 sampler. (on OSX catalina).

the fix is simple (but also bad): *do not use matplotlib and
multiprocessing in the same program*.

@@ -59,7 +59,7 @@ class ModelFitter(ModelParser):
     """
 
     def __init__(self, modelid, x_obs, y_obs, y_err, prior_d, mstar=1,
-                 rstar=1, N_samples=1000, N_cores=16, N_chains=3,
+                 rstar=1, N_samples=1000, N_cores=16, N_chains=4,
                  plotdir=None, pklpath=None, overwrite=1):
 
         self.N_samples = N_samples
@@ -76,7 +76,7 @@ class ModelFitter(ModelParser):
 
         self.initialize_model(modelid)
         self.verify_inputdata()
-        self.run_inference(prior_d, pklpath)
+        self.run_inference(prior_d, pklpath, make_threadsafe=True)
 
 
     def verify_inputdata(self):
@@ -85,7 +85,7 @@ class ModelFitter(ModelParser):
         assert isinstance(self.y_obs, np.ndarray)
 
 
-    def run_inference(self, prior_d, pklpath):
+    def run_inference(self, prior_d, pklpath, make_threadsafe=True):
 
         # if the model has already been run, pull the result from the
         # pickle. otherwise, run it.
@@ -322,11 +322,19 @@ class ModelFitter(ModelParser):
             # make sure that our initialization looks ok.
             self.y_MAP = map_estimate['mu_model'].flatten()
 
-            if self.PLOTDIR is None:
-                raise NotImplementedError
-            outpath = os.path.join(self.PLOTDIR,
-                                   'test_{}_MAP.png'.format(self.modelid))
-            plot_MAP_data(self.x_obs, self.y_obs, self.y_MAP, outpath)
+            if make_threadsafe:
+                pass
+            else:
+                # as described in
+                # https://github.com/matplotlib/matplotlib/issues/15410
+                # matplotlib is not threadsafe. so do not make plots before
+                # sampling, because some child processes tries to close a
+                # cached file, and crashes the sampler.
+                if self.PLOTDIR is None:
+                    raise NotImplementedError
+                outpath = os.path.join(self.PLOTDIR,
+                                       'test_{}_MAP.png'.format(self.modelid))
+                plot_MAP_data(self.x_obs, self.y_obs, self.y_MAP, outpath)
 
             # sample from the posterior defined by this model.
             trace = pm.sample(

### 2020-04-28 drivers/real.py

@@ -15,6 +15,8 @@ from billy import __path__
 
 def main(modelid):
 
+    make_threadsafe = 0
+
     traceplot = 0
     sampleplot = 1
     cornerplot = 1
@@ -24,9 +26,13 @@ def main(modelid):
     REALID = 'PTFO_8-8695'
     RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
     PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID),
-                           '20200413_v0')
+                           '20200427_v2')
 
     ##########
+
+    print(42*'#')
+    print(modelid)
+    print(42*'#')
 
     if not os.path.exists(PLOTDIR):
         os.mkdir(PLOTDIR)
@@ -45,44 +51,48 @@ def main(modelid):
 
     print(pm.summary(m.trace, varnames=list(prior_d.keys())))
 
-    if traceplot:
-        outpath = join(PLOTDIR, '{}_{}_traceplot.png'.format(REALID, modelid))
-        bp.plot_traceplot(m, outpath)
+    if make_threadsafe:
+        pass
 
-    if sampleplot:
-        outpath = join(PLOTDIR, '{}_{}_sampleplot.png'.format(REALID, modelid))
-        bp.plot_sampleplot(m, outpath, N_samples=100)
+    else:
+        if traceplot:
+            outpath = join(PLOTDIR, '{}_{}_traceplot.png'.format(REALID, modelid))
+            bp.plot_traceplot(m, outpath)
 
-    if splitsignalplot:
-        do_post = 0
-        do_map = 1
-        if do_post:
-            outpath = join(PLOTDIR, '{}_{}_splitsignalpost.png'.format(REALID, modelid))
-            ydict = bp.plot_splitsignal_post(m, outpath)
-            outpath = join(PLOTDIR, '{}_{}_phasefoldpost.png'.format(REALID, modelid))
-            bp.plot_phasefold_post(m, ydict, outpath)
-        if do_map:
-            outpath = join(PLOTDIR, '{}_{}_splitsignalmap_i.png'.format(REALID, modelid))
-            ydict = bp.plot_splitsignal_map(m, outpath, part='i')
-            outpath = join(PLOTDIR, '{}_{}_splitsignalmap_ii.png'.format(REALID, modelid))
-            ydict = bp.plot_splitsignal_map(m, outpath, part='ii')
-            outpath = join(PLOTDIR, '{}_{}_splitsignalmap_periodogram.png'.format(REALID, modelid))
[…]

### 2020-04-28 paper/ms.tex
add subset of coauthors

@@
 % 
-\author[0000-0002-4265-047X]{TESS collaborators}
-\affiliation{ Department of Astrophysical Sciences, Princeton
-	University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+%-------------------------------------
+% TESS Mission Architects:
+% These authors should be listed in this order
+% see https://spacebook.mit.edu/pages/viewpage.action?pageId=24543276
+%-------------------------------------
+%
+%PENDING
+\author{G. R. Ricker} % grr@space.mit.edu
+\affiliation{Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA}
+%
+%PENDING
+\author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu
+\affiliation{Department of Physics and Kavli Institute for Astrophysics
+and Space Research, Massachusetts Institute of Technology, Cambridge, MA
+02139, USA}
+%
+%PENDING
+\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu
+\affiliation{Center for Astrophysics \textbar \ Harvard \&
+Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
+%
+%PENDING
+\author{S. Seager} % seager@mit.edu
+\affiliation{Department of Earth, Atmospheric, and Planetary Sciences,
+Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
+%
+%PENDING
+\author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov
+\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+%
+%-------------------------------------
+% 3 representatives of each of SPOC, POC, TSO, for a total of 9. 
+%These 9 authors should be listed in alphabetical order
+%-------------------------------------
+% 3 TSO COAUTHORS
+% 3 SPOC COAUTHORS
+%     Mark E. Rose mark.rose@nasa.gov
+%     Jeffrey C. Smith  jeffrey.c.smith-1@nasa.gov
+%     Bill Wohler  bill.wohler@nasa.gov
+% 3 POC COAUTHORS: 
+%     John Doty (jpd@noqsi.com)
+%     Joel Villaseñor (jsvilla@space.mit.edu)
+%     Tom Barclay (barclay.astro@gmail.com)
+
+%PENDING
+\author[0000-0001-7139-2724]{T. Barclay} % barclay.astro@gmail.com
+\affiliation{NASA Goddard Space Flight Center, 8800 Greenbelt Road,
+Greenbelt, MD 20771, USA}
+\affiliation{University of Maryland, Baltimore County, 1000 Hilltop
+Circle, Baltimore, MD 21250, USA}
+
+%PENDING
+\author{J. P. Doty} % jpd@noqsi.com
+\affiliation{Noqsi Aerospace Ltd., 15 Blanchard Avenue, Billerica, MA,
+01821, USA}
+
+%PENDING
+\author[0000-0003-4724-745X]{M. E. Rose} %  mark.rose@nasa.gov
+\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+
+%PENDING
+\author[0000-0002-6148-7903]{J. C. Smith} % jeffrey.c.smith-1@nasa.gov
+\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+\affiliation{SETI Institute, Mountain View, CA 94043, USA}
+
+%PENDING
+\author{J. Villase\~nor} % jsvilla@space.mit.edu
+\affiliation{Department of Physics and Kavli Institute for
+Astrophysics and Space Research, Massachusetts Institute of
+Technology, Cambridge, MA 02139, USA}
+
+%PENDING
+\author[0000-0002-5402-9613]{B. Wohler} % bill.wohler@nasa.gov
+\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+\affiliation{SETI Institute, Mountain View, CA 94043, USA}
 

### 2020-04-28 billy/modelfitter.py
reran analysis with weaker rp/rstar prior

@@ -59,7 +59,7 @@ class ModelFitter(ModelParser):
     """
 
     def __init__(self, modelid, x_obs, y_obs, y_err, prior_d, mstar=1,
-                 rstar=1, N_samples=1000, N_cores=16, N_chains=4,
+                 rstar=1, N_samples=2000, N_cores=16, N_chains=4,
                  plotdir=None, pklpath=None, overwrite=1):
 
         self.N_samples = N_samples
@@ -136,7 +136,7 @@ class ModelFitter(ModelParser):
                     )
 
                     r = pm.Normal(
-                        "r", mu=prior_d['r'], sd=0.03*prior_d['r'],
+                        "r", mu=prior_d['r'], sd=0.10*prior_d['r'],
                         testval=prior_d['r']
                     )
                     # r = pm.Uniform(

### 2020-04-28 drivers/make_model_comparison_table.py

@@ -2,8 +2,11 @@ import pickle, os
 import numpy as np, pandas as pd
 from glob import glob
 
+run_id = '20200428_v0'
+
 pklpaths = glob(
-    '/Users/luke/Dropbox/proj/billy/results/PTFO_8-8695_results/20200413_v0/*bicdict*'
+    '/Users/luke/Dropbox/proj/billy/results/PTFO_8-8695_results/{}/*bicdict*'
+    .format(run_id)
 )
 ds = []
 for p in pklpaths:
@@ -23,6 +26,7 @@ df['redchisq'] = np.round(df.redchisq, 3)
 df['BIC'] = np.round(df.BIC, 1)
 df['D_BIC'] = np.round(df.D_BIC, 1)
 
-outpath = '../results/PTFO_8-8695_results/20200413_v0/bic_table_data.tex'
+outpath = '../results/PTFO_8-8695_results/{}/bic_table_data.tex'.format(run_id)
+
 df.to_latex(outpath, index=False)
 print('wrote {}'.format(outpath))

### 2020-04-28 drivers/make_posterior_table.py

@@ -11,31 +11,22 @@ from billy.convenience import (
 )
 from billy import __path__
 
-def normal_str(mu, sd, fmtstr=None):
-    # fmtstr: e.g., "({:.5f}; {:.2f})"
-    if fmtstr is None:
-        return '$\\mathcal{N}'+'({}; {})$'.format(mu, sd)
-    else:
-        return '$\\mathcal{N}'+'{}$'.format(fmtstr).format(mu, sd)
-
-
-def uniform_str(lower, upper, fmtstr=None):
-    # fmtstr: e.g., "({:.5f}; {:.2f})"
-    if fmtstr is None:
-        return '$\\mathcal{U}'+'({}; {})$'.format(lower, upper)
-    else:
-        return '$\\mathcal{U}'+'{}$'.format(fmtstr).format(lower, upper)
-
-
 def main(modelid):
 
     OVERWRITE = 0
     REALID = 'PTFO_8-8695'
     RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
     PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID),
-                           '20200413_v0')
+                           '20200428_v0')
 
-    summarypath = os.path.join(PLOTDIR, 'posterior_table_raw.csv')
+    assert modelid in [
+        'transit_2sincosPorb_2sincosProt',
+        'transit_2sincosPorb_3sincosProt'
+    ]
+
+    summarypath = os.path.join(
+        PLOTDIR, 'posterior_table_raw_{}.csv'.format(modelid)
+    )
 
     pklpath = os.path.join(
         os.path.expanduser('~'), 'local', 'billy',
@@ -66,6 +57,11 @@ def main(modelid):
     srows = ['period', 't0', 'r', 'b', 'u[0]', 'u[1]', 'mean',
              'omegaorb', 'Aorb0', 'Borb0', 'Aorb1', 'Borb1',
              'phirot', 'omegarot', 'Arot0', 'Brot0', 'Arot1', 'Brot1']
+    if modelid == 'transit_2sincosPorb_3sincosProt':
+        srows = ['period', 't0', 'r', 'b', 'u[0]', 'u[1]', 'mean',
+                 'omegaorb', 'Aorb0', 'Borb0', 'Aorb1', 'Borb1',
+                 'phirot', 'omegarot', 'Arot0', 'Brot0', 'Arot1', 'Brot1',
+                 'Arot2', 'Brot2']
 
     df = df.loc[srows]
 
[…]

### 2020-04-28 drivers/real.py

@@ -26,7 +26,7 @@ def main(modelid):
     REALID = 'PTFO_8-8695'
     RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
     PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID),
-                           '20200427_v2')
+                           '20200428_v0')
 
     ##########
 
@@ -61,7 +61,7 @@ def main(modelid):
 
         if sampleplot:
             outpath = join(PLOTDIR, '{}_{}_sampleplot.png'.format(REALID, modelid))
-            bp.plot_sampleplot(m, outpath, N_samples=100)
+            bp.plot_sampleplot(m, outpath, N_samples=10)
 
         if splitsignalplot:
             do_post = 0

### 2020-04-28 paper/ms.tex
update coauthors
update text per broader prior

@@
 \accepted{---}
-\submitjournal{AAS journals.}
+% \submitjournal{AAS journals.}
 \shorttitle{Two Stars, Two Signals, No Planet}
@@
 \author[0000-0002-0514-5538]{L.\,G.\,Bouma}
-\affiliation{ Department of Astrophysical Sciences, Princeton
+\affiliation{Department of Astrophysical Sciences, Princeton
 University, 4 Ivy Lane, Princeton, NJ 08540, USA}
@@
 \author[0000-0002-4265-047X]{J.\,N.\,Winn}
-\affiliation{ Department of Astrophysical Sciences, Princeton
+\affiliation{Department of Astrophysical Sciences, Princeton
 University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+%
+% PENDING
+%\author[0000-0001-7246-5438]{A. Vanderburg}
+%\affiliation{Department of Astronomy, The University of Texas at
+%Austin, Austin, TX 78712, USA}
+%\affiliation{NASA Sagan Fellow}
+
 % 
@@
 % 3 TSO COAUTHORS
+%     Karen Collins karenacollins@outlook.com
+%     Sam Quinn squinn@cfa.harvard.edu
+%     Dana Louie danalouie@astro.umd.edu
 % 3 SPOC COAUTHORS
@@
 
+% PENDING
+\author[0000-0001-6588-9574]{K. A. Collins} % karenacollins@outlook.com
+\affiliation{Center for Astrophysics \textbar \ Harvard \&
+Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
+
 %PENDING
@@
 01821, USA}
+
+% PENDING
+\author{D. Louie} % danalouie@astro.umd.edu
+\affiliation{University of Maryland, Baltimore County, 1000 Hilltop
+Circle, Baltimore, MD 21250, USA}
+
+% PENDING
+\author[0000-0002-8964-8377]{S. N. Quinn} % squinn@cfa.harvard.edu
+\affiliation{Center for Astrophysics \textbar \ Harvard \&
+Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
 
@@
 		\leavevmode
-		\includegraphics[width=0.99\textwidth]{f3.pdf}
+		\includegraphics[width=0.93\textwidth]{f3.pdf}
 	\end{center}
@@
 10.74$\,$hr period, and two modes at the longer 11.96$\,$hr period.
-The next-best model had an additional harmonic at the longer period
-($N_{\rm \ell}=3$), but was otherwise identical.  All nine models have
-reduced $\chi^2$ values ranging between 1.37 and 1.51, which suggests
+Two other models had BIC values that provided comparable evidence
+\citep{burnham_multimodel_2016}: one with ($N_{\rm s}=3$, $N_{\rm
+\ell}=3$), and the other with ($N_{\rm s}=3$, $N_{\rm \ell}=2$).
+All nine models have
+reduced $\chi^2$ values ranging between 1.25 and 1.68, which suggests
 a plausible though imperfect agreement between the data and the model
@@
 ending with a ``dip'' of depth $\approx$1.2\%, lasting $\approx$0.75
-hours.
+hours.  The fact that the higher-order models lead to roughly
+comparable BIC values suggests that these sharper features may indeed
+be real, but that there is only weak statistical support for their
+existence.
 
@@
 interprets the ``out of dip'' signal to potentially show multiple
-dips, while our BIC-based model comparison did not indicate the need
-for additional harmonics after fitting for the largest dip and two
-harmonics at the shorter period (Table~1). Our analysis emphasized the
-agreement between the TESS dip ephemeris and that from
-\citet{tanimoto_evidence_2020}.  \citet{koen_2020} however made a
-clearer case for testing the binarity hypothesis using interferometric
-imaging and high-precision radial velocities.
+dips, while our BIC-based model comparison did not indicate that
+including additional harmonics past the second order was strictly
+necessary (Table~1). Our analysis emphasized the agreement between the
+TESS dip ephemeris and that from \citet{tanimoto_evidence_2020}.
+\citet{koen_2020} however made a clearer case for testing the binarity
+hypothesis using interferometric imaging and high-precision radial
+velocities.
 \\
@@
 that were eventually omitted, and A.~Vanderburg for the suggestion to
-consider RUWE.
-%
-We thank the Heising-Simons Foundation for their generous support of this work.
-%
-\ptfo\ was included on the TESS ``short-cadence'' target list thanks to the
-Guest Investigator programs of S.\ Czesla and C.\ Huang (G011128 and
-G011132 respectively).
+look into RUWE.
+%
+We also thank the Heising-Simons Foundation for
+their generous support of this work.
+%
+\ptfo\ was included on the TESS ``short-cadence'' target list thanks
+to the Guest Investigator programs of S.\ Czesla and C.\ Huang
+(G011128 and G011132 respectively).
 %

### 2020-04-28 paper/collect.sh

@@ -1,6 +1,6 @@
 # Usage: copy in figures for the paper
 
-pdir=../results/PTFO_8-8695_results/20200413_v0/
+pdir=../results/PTFO_8-8695_results/20200428_v0/
 
 cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_i.pdf f1.pdf
 cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_ii.pdf f2.pdf

### 2020-04-29 paper/ms.tex
D Latham's typo catches

@@
 %
-%PENDING
+%CONFIRMED
 \author{G. R. Ricker} % grr@space.mit.edu
@@
 %
-%PENDING
+%CONFIRMED
 \author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu
@@
   photometric periods. The variability is dominated by a sinusoidal
-  signal with a period of 11.96$\,$hr, presumably caused by stellar
-  rotation.  Also present is a 10.74-day signal consisting of a
+  signal with a period of 11.98$\,$hr, presumably caused by stellar
+  rotation.  Also present is a 10.76-day signal consisting of a
   not-quite sinusoid interrupted by hour-long dips, the type of signal
@@
 down-weighted in the astrometric solution.  \ptfo\ was assigned the
-Gaia DR2 identifier 3222255959210123904.  It photometric brightness
+Gaia DR2 identifier 3222255959210123904.  Its photometric brightness
 was measured using selected bands ($G$, $Rp$, and $Bp$) of the Gaia
@@
 refer to these two periods as the ``shorter period'' $P_{\rm s}$ and
-the ``longer period'' $P_{\rm ell}$.  The peak at $P_{\rm \ell}$ is
+the ``longer period'' $P_{\rm \ell}$.  The peak at $P_{\rm \ell}$ is
 the taller of the two peaks.  Lower-power harmonics of both signals
@@
 We fitted a model to the lightcurve consisting of a linear combination
-of Fourier modes with periods $P_{\rm s}$ and $P_{\rm ell}$, as well
+of Fourier modes with periods $P_{\rm s}$ and $P_{\rm \ell}$, as well
 as a number of harmonics chosen as described below. To try accounting
@@
 shorter period, plus $2N_{\rm \ell}=2$ Fourier amplitudes at the
-longer period, and well as $P_\ell$ itself and the relative phase
+longer period, as well as $P_\ell$ itself and the relative phase
 $\phi_\ell$.  The total number of parameters is 14.
@@
 (Table~1).  The model with the lowest BIC had two modes at the shorter
-10.74$\,$hr period, and two modes at the longer 11.96$\,$hr period.
+10.76$\,$hr period, and two modes at the longer 11.98$\,$hr period.
 Two other models had BIC values that provided comparable evidence
@@
 
-In every model, the $11.96\,{\rm hr}$ variability is a simple sinusoid
-with peak-to-peak amplitude $\approx$10\%.  The $10.74\,{\rm hr}$
+In every model, the $11.98\,{\rm hr}$ variability is a simple sinusoid
+with peak-to-peak amplitude $\approx$10\%.  The $10.76\,{\rm hr}$
 variability is always more complex.  The overall impression is of a
@@
     only star bright and close enough to be contributing to the signal
-    attributed to \ptfo\. However, the pixel-level TESS data showed
+    attributed to \ptfo. However, the pixel-level TESS data showed
     that any such contamination is negligible (see
@@
 
-The standard interpretation for 11.96$\,$hour nearly sinusoidal
+The standard interpretation for 11.98$\,$hour nearly sinusoidal
 modulations of a pre-main-sequence M dwarf is stellar rotation.  This
@@
 The TESS lightcurve shows a dip that lasts about 45 minutes, and seems
-to recur every 10.74 hours
+to recur every 10.76 hours
 (Figures~\ref{fig:splitsignal},~\ref{fig:splitsignalii},~\ref{fig:phasefold}).
@@
 
-Visually, the out-of-dip modulation at the 10.74$\,$hour period
+Visually, the out-of-dip modulation at the 10.76$\,$hour period
 resembles a slightly asymmetric sinusoid (Figure~\ref{fig:phasefold}).
@@
 \end{enumerate}
-The 10.74 hour signal in \ptfo\ meets all of these criteria.  This is
+The 10.76 hour signal in \ptfo\ meets all of these criteria.  This is
 the first connection of \ptfo\ with this class of objects because the
@@
 of similar young rapidly rotating M-dwarfs.  Many possible
-explanations been detailed by \citet{rebull_rotation_2016},
+explanations have been detailed by \citet{rebull_rotation_2016},
 \citet{david_transient_2017}, \citet{stauffer_orbiting_2017}, and
@@
     ``long'' signal is a 10\% peak-to-peak sinusoid that repeats every
-    11.96 hours.  The ``short'' signal is a 4\% peak-to-peak ``dip +
-    asymmetric sinusoid'' that repeats every 10.74 hours. The signals
+    11.98 hours.  The ``short'' signal is a 4\% peak-to-peak ``dip +
+    asymmetric sinusoid'' that repeats every 10.76 hours. The signals
     beat, and therefore cannot be an artifact linked to data

### 2020-05-01 paper/ms.tex
fix typo

@@
 
-% PENDING
+%CONFIRMED
 \author[0000-0001-6588-9574]{K. A. Collins} % karenacollins@outlook.com
@@
 
-We also used the Gaia data to see if the observed luminosity of \ptfo
+We also used the Gaia data to see if the observed luminosity of \ptfo\ 
 is too bright to be a single star, i.e., if it is a ``photometric

### 2020-05-08 billy/models.py
plot kwarg tweaking

@@ -64,5 +64,3 @@ def quadratic_model(params, x, x_occ=None):
         return (t0 + period*x + half_dP_dE*x**2,
                 t0 + period/2 + period*x_occ + half_dP_dE*x_occ**2
                )
-
-

### 2020-05-08 billy/plotting.py

@@ -103,10 +103,10 @@ def plot_test_data(x_obs, y_obs, y_mod, modelid, outdir):
     savefig(fig, outpath, writepdf=0, dpi=300)
 
 
-def plot_MAP_data(x_obs, y_obs, y_MAP, outpath):
+def plot_MAP_data(x_obs, y_obs, y_MAP, outpath, ms=4):
     plt.close('all')
     plt.figure(figsize=(14, 4))
-    plt.plot(x_obs, y_obs, ".k", ms=4, label="data")
+    plt.plot(x_obs, y_obs, ".k", ms=ms, label="data")
     plt.plot(x_obs, y_MAP, lw=1)
     plt.ylabel("relative flux")
     plt.xlabel("time [days]")
@@ -115,14 +115,15 @@ def plot_MAP_data(x_obs, y_obs, y_MAP, outpath):
     savefig(fig, outpath, writepdf=0, dpi=300)
 
 
-def plot_sampleplot(m, outpath, N_samples=100):
+def plot_sampleplot(m, outpath, N_samples=100, ms=4, malpha=1):
 
     if os.path.exists(outpath) and not m.OVERWRITE:
         return
 
     plt.close('all')
     fig, ax = plt.subplots(figsize=(14, 4))
-    ax.plot(m.x_obs, m.y_obs, ".k", ms=4, label="data", zorder=N_samples+1)
+    ax.plot(m.x_obs, m.y_obs, ".k", ms=ms, label="data", zorder=N_samples+1,
+            alpha=malpha)
     ax.plot(m.x_obs, m.map_estimate['mu_model'], lw=0.5, label='MAP',
             zorder=N_samples+2, color='C1', alpha=1)
 

### 2020-05-08 paper/ms.tex
update text

@@
 
-%PENDING
+%CONFIRMED
 \author[0000-0001-7139-2724]{T. Barclay} % barclay.astro@gmail.com
@@
 
-%PENDING
+%CONFIRMED
 \author[0000-0002-5402-9613]{B. Wohler} % bill.wohler@nasa.gov
@@
   signal with a period of 11.98$\,$hr, presumably caused by stellar
-  rotation.  Also present is a 10.76-day signal consisting of a
+  rotation.  Also present is a 10.76$\,$hr signal consisting of a
   not-quite sinusoid interrupted by hour-long dips, the type of signal
@@
   is nearly 180$^\circ$ away from the phase of the originally reported
-  dips, which, as noted previously, makes them difficult to explain as
+  dips. As noted previously, this makes them difficult to explain as
   planetary transits.  Instead, we believe that PTFO$\,$8-8695 is a
@@
 initialized each model with the parameters of the maximum a posteriori
-(MAP) model.  We then sampled using \texttt{PyMC3}'s gradient-based
-No-U-Turn Sampler \citep{hoffman_no-u-turn_2014}, and used $\hat{R}$
-as our convergence diagnostic \citep{gelman_inference_1992}.  We
-tested our ability to successfully recover injected parameters using
-synthetic data before fitting the \ptfo\ lightcurves.
+(MAP) model.  We then assumed a Gaussian likelihood, and sampled using
+\texttt{PyMC3}'s gradient-based No-U-Turn Sampler
+\citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
+diagnostic \citep{gelman_inference_1992}.  We tested our ability to
+successfully recover injected parameters using synthetic data before
+fitting the \ptfo\ lightcurves.
 
@@
 
-\citet{van_eyken_ptf_2012} obtained stronger constraints on possible
-stellar companions based on high-resolution imaging with the NIRC2
+Stronger constraints on possible
+stellar companions were obtained by
+\citet{van_eyken_ptf_2012} through high-resolution imaging with the NIRC2
 camera on Keck II.  They reported 3-$\sigma$ $H$-band magnitude
@@
 \citep{lee_evidence_2018}.} cannot be the source of our signals.
-
-While long-baseline radial velocity (RV) measurements could provide
-complementary lower limits on companion masses and separations, the
-available RV data for \ptfo\ have large uncertainties due to the
-faintness of the star.  The RV dataset with the longest timespan in
-the literature appears to be the 5 Keck/HIRES measurements acquired
-over 10 days in April 2011 by \citet{van_eyken_ptf_2012}.  The
-root-mean-squared RV over that 10 day span was $1.6\,{\rm km}\,{\rm
-s}^{-1}$, consistent with the measurement precision.  Although
-\citet{yu_tests_2015} acquired 22 further Keck/HIRES spectra on one
-night in December 2013, those points were not reduced to velocities.
-Further Keck/HIRES measurements of this target could confirm or refute
-the existence of possible binary companions.
 
@@
 
+\subsection{Radial Velocity Binarity}
+
+Long-baseline radial velocity (RV) measurements could also reveal
+the presence of multiple stars in this system.
+Unfortunately,
+the available RV data for \ptfo\ is rather sparse, presumably due to the
+stellar faintness and equatorial velocity.
+The RV dataset with the longest baseline we could find
+in the literature was the 5 Keck/HIRES measurements acquired
+over 10 days in April 2011 by \citet{van_eyken_ptf_2012}.  The
+root-mean-squared RV over that 10 day span was $1.6\,{\rm km}\,{\rm
+	s}^{-1}$, consistent with the measurement precision. 
+Although \citet{van_eyken_ptf_2012} tried a CCF-based RV reduction
+technique, they eventually found that manually selecting
+absorption lines and measuring line centroids was more effective.
+While
+\citet{yu_tests_2015} acquired 22 further Keck/HIRES spectra over one
+night in December 2013, those points were not reduced to velocities, and
+it is not clear whether the appropriate calibration data exist.
+Further Keck/HIRES measurements of \ptfo, if appropriately calibrated,
+could potentially confirm or refute the presence of binary companions.
+
 \section{Discussion}
@@
 
-The standard interpretation for 11.98$\,$hour nearly sinusoidal
+The standard interpretation for 11.98$\,$hr nearly sinusoidal
 modulations of a pre-main-sequence M dwarf is stellar rotation.  This
@@
 
-Visually, the out-of-dip modulation at the 10.76$\,$hour period
+Visually, the out-of-dip modulation at the 10.76$\,$hr period
 resembles a slightly asymmetric sinusoid (Figure~\ref{fig:phasefold}).
@@
     \citealt{rebull_usco_2018},~Figure~20).
-  \item The age is $\lesssim$ 100$\,$Myr.
+  \item The age is typically\footnote{
+    At present, the oldest observed ``scallops'' are in the Pleaides
+    \citep{rebull_rotation_2016}. One of these, EPIC 211013604, might
+    meet the ``persistent dip'' classification.  If so, it is also the
+    oldest known.
+    } $\lesssim$ 100$\,$Myr.
   \item The lightcurves show shallow, angular dips, usually superposed
@@
 \end{enumerate}
-The 10.76 hour signal in \ptfo\ meets all of these criteria.  This is
+The 10.76$\,$hr signal in \ptfo\ meets all of these criteria.  This is
 the first connection of \ptfo\ with this class of objects because the
@@
 (ii)} transits of enshrouded protoplanets; {\it (iii)} occultations of
-starspots by an optically thick disk.  The first and last explanations
+starspots by an optically thick disk.  
+The first and last explanations
 have added appeal because they are flexible enough to explain not only

### 2020-05-08 billy/plotting.py
tweak plot

@@ -828,7 +828,7 @@ def plot_hr(outdir):
         group_df_dr2['phot_bp_mean_mag']-group_df_dr2['phot_rp_mean_mag'],
         yval,
         c='k', alpha=1., zorder=3, s=9, rasterized=True, linewidths=0,
-        label='K+18 members'
+        label='Members'
     )
 
     target_yval = np.array([target_df['phot_g_mean_mag'] +

### 2020-05-08 paper/ms.tex
update paper

@@
 %
-% PENDING
-%\author[0000-0001-7246-5438]{A. Vanderburg}
-%\affiliation{Department of Astronomy, The University of Texas at
-%Austin, Austin, TX 78712, USA}
-%\affiliation{NASA Sagan Fellow}
-
 % 
@@
 \author{G. R. Ricker} % grr@space.mit.edu
-\affiliation{Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA}
-%
-%PENDING
+\affiliation{Department of Physics and Kavli Institute for
+Astrophysics and Space Research, Massachusetts Institute of
+Technology, Cambridge, MA 02139, USA}
+%
+%PENDING yy
 \author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu
-\affiliation{Department of Physics and Kavli Institute for Astrophysics
-and Space Research, Massachusetts Institute of Technology, Cambridge, MA
-02139, USA}
+\affiliation{Department of Physics and Kavli Institute for
+Astrophysics and Space Research, Massachusetts Institute of
+Technology, Cambridge, MA 02139, USA}
 %
@@
 %
-%PENDING
-\author{S. Seager} % seager@mit.edu
-\affiliation{Department of Earth, Atmospheric, and Planetary Sciences,
+%CONFIRMED
+\author[0000-0002-6892-6948]{S.~Seager} % seager@mit.edu
+\affiliation{Department of Physics and Kavli Institute for
+Astrophysics and Space Research, Massachusetts Institute of
+Technology, Cambridge, MA 02139, USA}
+\affiliation{Department of Earth, Atmospheric and Planetary Sciences,
 Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
-%
-%PENDING
+\affiliation{Department of Aeronautics and Astronautics, MIT, 77
+Massachusetts Avenue, Cambridge, MA 02139, USA}
+%
+%PENDING yy
 \author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov
@@
 
-%PENDING
+%CONFIRMED
 \author{J. P. Doty} % jpd@noqsi.com
@@
 
-% PENDING
+% PENDING yy
 \author{D. Louie} % danalouie@astro.umd.edu
@@
 
-% PENDING
+%CONFIRMED
 \author[0000-0002-8964-8377]{S. N. Quinn} % squinn@cfa.harvard.edu
@@
 
-%PENDING
+%PENDING y
 \author[0000-0003-4724-745X]{M. E. Rose} %  mark.rose@nasa.gov
@@
 
-%PENDING
+%PENDING y
 \author[0000-0002-6148-7903]{J. C. Smith} % jeffrey.c.smith-1@nasa.gov
@@
 
-%PENDING
+%CONFIRMED
 \author{J. Villase\~nor} % jsvilla@space.mit.edu
@@
 pixels chosen to maximize the SNR of the total flux of the target
-\citep{smith_kepler_apertures_2017}.  Non-astrophysical variability
+\citep{smith_finding_2016}.  Non-astrophysical variability
 was removed by fitting out trends common to many stars
-\citep{smith_kepler_PDC_2017}.
+\citep{smith_kepler_2012,stumpe_multiscale_2014}.
+% \citep{smith_kepler_PDC_2017}
+% \citep{smith_kepler_apertures_2017}
 

### 2020-05-10 paper/ms.tex
update text

@@
 
-%PENDING y
+%CONFIRMED
 \author[0000-0002-6148-7903]{J. C. Smith} % jeffrey.c.smith-1@nasa.gov
@@
 \cite{barnes_measurement_2013} as the natural effects of gravity
-darkening However, \cite{howarth_reappraisal_2016} argued that the
+darkening.  However, \cite{howarth_reappraisal_2016} argued that the
 necessary amplitude of gravity darkening is too large to be realistic,
@@
 Section~\ref{sec:gaia}, show that \ptfo\ is a photometric binary, and
-suggest it also an astrometric binary.  We discuss the pieces of the
-puzzle in Section~\ref{sec:discussion}, and summarize the situation in
-Section~\ref{sec:conclusions}.  We comment on the study by
-\citet{koen_2020} in postscript.
+suggest it is also an astrometric binary.  We discuss the pieces of
+the puzzle in Section~\ref{sec:discussion}, and summarize the
+situation in Section~\ref{sec:conclusions}.  We comment on the study
+by \citet{koen_2020} in postscript.
 
@@
 largest subgroup of 25$\,$Ori, with 149 identified members.  Based on
-the CMD, the mean age was found to be 6.9$\,$Myr, while the HR diagram
-led to an estimated age of 8.5$\,$Myr \citep[see][Section
-2.3]{kounkel_apogee2_2018}.  \citet{kounkel_apogee2_2018} identified
-seven other smaller groups in the Orion complex near the Be star
-25$\,$Ori. These groups received higher numbers, {\it e.g.},
-25$\,$Ori-2 (${\rm Age}_{\rm CMD} =15.1\,$Myr; ${\rm Age}_{\rm CMD}
-=12.9\,$Myr; see also \citealt{briceno_cidaII_2019}).
+the color-magnitude diagram (CMD), the mean age was found to be
+6.9$\,$Myr, while the HR diagram led to an estimated age of 8.5$\,$Myr
+\citep[see][Section 2.3]{kounkel_apogee2_2018}.
+\citet{kounkel_apogee2_2018} identified seven other smaller groups in
+the Orion complex near the Be star 25$\,$Ori. These groups received
+higher numbers, {\it e.g.}, 25$\,$Ori-2 (${\rm Age}_{\rm CMD}
+=15.1\,$Myr; ${\rm Age}_{\rm CMD} =12.9\,$Myr; see also
+\citealt{briceno_cidaII_2019}).
 
@@
     attributed to \ptfo. However, the pixel-level TESS data showed
-    that any such contamination is negligible (see
+    that Star A is not the source of the observed variability (see
     Section~\ref{subsec:blend}).
@@
 stellar faintness and equatorial velocity.
-The RV dataset with the longest baseline we could find
-in the literature was the 5 Keck/HIRES measurements acquired
-over 10 days in April 2011 by \citet{van_eyken_ptf_2012}.  The
-root-mean-squared RV over that 10 day span was $1.6\,{\rm km}\,{\rm
-	s}^{-1}$, consistent with the measurement precision. 
-Although \citet{van_eyken_ptf_2012} tried a CCF-based RV reduction
-technique, they eventually found that manually selecting
-absorption lines and measuring line centroids was more effective.
-While
-\citet{yu_tests_2015} acquired 22 further Keck/HIRES spectra over one
-night in December 2013, those points were not reduced to velocities, and
-it is not clear whether the appropriate calibration data exist.
-Further Keck/HIRES measurements of \ptfo, if appropriately calibrated,
-could potentially confirm or refute the presence of binary companions.
+The RV datasets with the longest baselines we could find
+in the literature were those reported by \citet{van_eyken_ptf_2012}.
+These included 5 Keck/HIRES measurements acquired
+over 10 days in April 2011, and 4 HET/HRS measurements acquired over
+10 days in February 2011.  The root-mean-squared RV over each 10 day
+span was $\approx$$2\,{\rm km}\,{\rm s}^{-1}$, consistent with the
+measurement precision.  Although \citet{van_eyken_ptf_2012} tried a
+CCF-based RV reduction technique, they eventually found that manually
+selecting absorption lines and measuring line centroids was more
+effective.  While \citet{yu_tests_2015} acquired 22 further Keck/HIRES
+spectra over one night in December 2013, those points were not reduced
+to velocities, and it is not clear whether the appropriate calibration
+data exist.  Further Keck/HIRES measurements of \ptfo, if
+appropriately calibrated, could potentially confirm or refute the
+presence of binary companions.
 
@@
 maximum of $\approx$5\% \citep[{\it
-e.g.},]{koen_multicolour_2015,yu_tests_2015,tanimoto_evidence_2020}.
+e.g.},][]{koen_multicolour_2015,yu_tests_2015,tanimoto_evidence_2020}.
 
@@
     Five transient and persistent flux dip stars selected based on
-    their visual similarity to the short-period signal in \ptfo are as
+    their visual similarity to the short-period signal in \ptfo\ are as
     shown.
@@
 CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$
-\citep{henry_solar_2006,duchene_stellar_2013}.  Low-mass
-pre-main-sequence stars however have been shown to companion fractions
-up to twice as high in dispersed clusters such as Upper~Sco and Taurus
-\citep{kraus_mapping_2008,kraus_mapping_2011}.  A detailed
+\citep{henry_solar_2006,duchene_stellar_2013,winters_solar_2019}.
+Low-mass pre-main-sequence stars however have been shown to companion
+fractions up to twice as high in dispersed clusters such as Upper~Sco
+and Taurus \citep{kraus_mapping_2008,kraus_mapping_2011}.  A detailed
 high-resolution imaging survey would be necessary to determine whether
@@
 %
-The authors thank D.~Fabrycky for helpful tidal evolution calculations
-that were eventually omitted, and A.~Vanderburg for the suggestion to
-look into RUWE.
+The authors thank D.~Fabrycky, S.~Mahadevan, G.~Stef\'ansson, and
+A.~Vanderburg for helpful calculations, observations, and suggestions.
 %

### 2020-05-10 billy/plotting.py
add model underplot to phase-fold

@@ -254,8 +254,8 @@ def plot_splitsignal_map(m, outpath, part='i'):
         'y_rot': m.y_obs-y_orb,
         'y_resid': m.y_obs-m.map_estimate['mu_model'],
         'y_mod_tra': y_tra,
-        'y_mod_rot': y_orb,
-        'y_mod_orb': y_rot,
+        'y_mod_rot': y_rot,
+        'y_mod_orb': y_orb,
         'y_mod': m.map_estimate['mu_model'],
         'y_err': m.y_err
     }
@@ -348,11 +348,18 @@ def plot_phasefold_map(m, d, outpath):
     orb_bd = phase_bin_magseries(
         orb_d['phase'], orb_d['mags'], binsize=0.01
     )
+    morb_d = phase_magseries(
+        d['x_obs'], d['y_mod_orb'], P_orb, t0_orb, wrap=True, sort=True
+    )
+
     rot_d = phase_magseries(
         d['x_obs'], d['y_rot'], P_rot, t0_rot, wrap=True, sort=True
     )
     rot_bd = phase_bin_magseries(
         rot_d['phase'], rot_d['mags'], binsize=0.01
+    )
+    mrot_d = phase_magseries(
+        d['x_obs'], d['y_mod_rot'], P_rot, t0_rot, wrap=True, sort=True
     )
 
     # make tha plot
@@ -363,6 +370,8 @@ def plot_phasefold_map(m, d, outpath):
                    zorder=4, linewidths=0, rasterized=True)
     axs[0].scatter(rot_bd['binnedphases'], rot_bd['binnedmags'], color='black',
                    s=8, alpha=1, zorder=5, linewidths=0)
+    axs[0].plot(mrot_d['phase'], mrot_d['mags'], lw=0.5, color='C0', alpha=1,
+                zorder=3)
 
     txt0 = '$P_{{\mathrm{{\ell}}}}$ = {:.5f}$\,$d'.format(P_rot)
     props = dict(boxstyle='square', facecolor='white', alpha=0.9, pad=0.15,
@@ -378,6 +387,8 @@ def plot_phasefold_map(m, d, outpath):
                    zorder=4, linewidths=0, rasterized=True)
     axs[1].scatter(orb_bd['binnedphases'], orb_bd['binnedmags'], color='black',
                    s=8, alpha=1, zorder=5, linewidths=0)
[…]

### 2020-05-10 drivers/real.py

@@ -92,7 +92,7 @@ def main(modelid):
 
 if __name__ == "__main__":
 
-    DEBUG = 0
+    DEBUG = 1
 
     if DEBUG:
         main('transit_2sincosPorb_2sincosProt')

### 2020-05-12 paper/ms.tex
update paper
merge Jon Jenkins comments

@@
 %
-%PENDING yy
+%CONFIRMED
 \author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov
@@
 
-%PENDING y
+%CONFIRMED
 \author[0000-0003-4724-745X]{M. E. Rose} %  mark.rose@nasa.gov
@@
   its kinematic group, and independently suggest that it is an
-  astrometric binary.  The TESS lightcurve shows two different
+  astrometric binary.  The TESS light curve shows two different
   photometric periods. The variability is dominated by a sinusoidal
@@
 available observations from TESS \citep{ricker_transiting_2015} and
-Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
+Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS light curve shows
 two different periodic signals, which we analyze in
@@
 	\caption{
-    {\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 19).} {\it Top}:
-    The ``raw'' (\texttt{PDCSAP} mean-subtracted) relative flux.
+    {\bf TESS light curve of \ptfo\ (Sector 6, Orbit 19).} {\it Top}:
+    The original (\texttt{PDCSAP} mean-subtracted) relative flux.
     The beat period of 4.48 days is visible by eye.  The blue curve is
@@
     plus 2 harmonics and a transit at the shorter period $P_{\rm s}$.
-    {\it Upper middle}: Longer-period signal, equal to the raw signal
+    {\it Upper middle}: Longer-period signal, equal to the original signal
     minus the shorter-period signal.  {\it Lower middle}:
-    Shorter-period signal, equal to the raw signal minus the
+    Shorter-period signal, equal to the original signal minus the
     longer-period signal.  {\it Bottom}: residual relative flux.  The
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 20).} Same
+  \caption{ {\bf TESS light curve of \ptfo\ (Sector 6, Orbit 20).} Same
   format as Figure~\ref{fig:splitsignal}.
@@
 
-\ptfo\ [also known as CVSO\,30; \citealt{briceno_cida_2005}] was
+\ptfo\ (also known as CVSO\,30; \citealt{briceno_cida_2005}) was
 observed by TESS with Camera 1, CCD 1, from December 15, 2018 until
@@
 
-The 2-minute stacks for \ptfo\ were reduced to lightcurves by the
+The 2-minute stacks for \ptfo\ were reduced to light curves by the
 Science Processing Operations Center (SPOC) at NASA
 Ames~\citep{jenkins_tess_2016}.  We mainly used the Presearch Data
-Conditioning (PDC) lightcurve.  The PDC lightcurve aperture used
+Conditioning (PDC) light curve.  The PDC light curve aperture used
 pixels chosen to maximize the SNR of the total flux of the target
-\citep{smith_finding_2016}.  Non-astrophysical variability
+\citep{bryson_2020_target_aperture}.  Non-astrophysical variability
 was removed by fitting out trends common to many stars
@@
 
-As an independent check on the 2-minute SPOC lightcurve, we examined
-the lightcurve based upon 30-minute image stacks which was produced as
+As an independent check on the 2-minute SPOC light curve, we examined
+the light curve based upon 30-minute image stacks which was produced as
 part of the Cluster Difference Imaging Photometric Survey (CDIPS;
-\citealt{bouma_cluster_2019}).  Our CDIPS lightcurve of choice used a
+\citealt{bouma_cluster_2019}).  Our CDIPS light curve of choice used a
 circular aperture with radius 1 pixel.
@@
 often systematic red noise in the photometry during those times.  Both
-the CDIPS and PDC lightcurves showed a clear discontinuous jump in the
+the CDIPS and PDC light curves showed a clear discontinuous jump in the
 last few days of orbit 20, which seemed likely to be an instrumental
@@
 timestamps ranging from BJD\,2458488.3 until the end of the orbit.
-The PDC lightcurve initially had 15{,}678 points.  The quality-flag
+The PDC light curve initially had 15{,}678 points.  The quality-flag
 cut removed 854 points; masking the orbit edges removed an additional
@@
 
-We normalized the lightcurve by dividing out the median flux, and then
+We normalized the light curve by dividing out the median flux, and then
 opted to subtract 1{.}0 to set the median value to zero, which
@@
     points are binned to 100 points per period.
+    The model (blue line)
+    includes 2 harmonics at the longer period,
+    plus 2 harmonics and a transit at the shorter period.
 		\label{fig:phasefold}
@@
 
-Our initial inspection of the TESS lightcurve, in both its 2-minute
+Our initial inspection of the TESS light curve, in both its 2-minute
 PDCSAP and 30-minute FFI forms, showed a strong sinusoidal beat signal
@@
 
-The peak-to-peak lightcurve amplitude at maximum, when the two signals
+The peak-to-peak light curve amplitude at maximum, when the two signals
 interfere constructively, is about 14\%.  During the times of
@@
 
-\subsection{Lightcurve Model}
-
-We fitted a model to the lightcurve consisting of a linear combination
+\subsection{Light Curve Model}
+
+We fitted a model to the light curve consisting of a linear combination
 of Fourier modes with periods $P_{\rm s}$ and $P_{\rm \ell}$, as well
@@
 code \citep{exoplanet:exoplanet}.  Our priors are listed in Table~2.
-To speed up the fitting process, we averaged the 2-minute lightcurve
+To speed up the fitting process, we averaged the 2-minute light curve
 to 10-minute samples.  We correspondingly scaled the uncertainties in
@@
 successfully recover injected parameters using synthetic data before
-fitting the \ptfo\ lightcurves.
+fitting the \ptfo\ light curves.
 
@@
 
-To explore where each model succeeded and failed, we split the raw
-signal into its respective components (Figures~\ref{fig:splitsignal}
-and~\ref{fig:splitsignalii}).  We also examined the phase-folded
-signals (Figure~\ref{fig:phasefold}).
+To explore where each model succeeded and failed, we split the
+original signal into its respective components
+(Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}).  We also
+examined the phase-folded signals (Figure~\ref{fig:phasefold}).
 
@@
 In the upper panels, the pixels used to measure the background level
-in the SPOC lightcurve are indicated with `\texttt{X}' hatching, and
-the pixels used in the final lightcurve aperture are shown with
+in the SPOC light curve are indicated with `\texttt{X}' hatching, and
+the pixels used in the final light curve aperture are shown with
 `\texttt{/}' hatching.
@@
 
-First, we examined the CDIPS FFI lightcurves of the target, which are
-available on MAST \citep{bouma_cluster_2019}. Three lightcurves are
+First, we examined the CDIPS FFI light curves of the target, which are
+available on MAST \citep{bouma_cluster_2019}. Three light curves are
 available, based on photometric apertures with a radius of 1, 1.5, or
@@
 within a percent, regardless of the size of the photometric aperture
-that was used to create the lightcurve.  If Star A were the source of
+that was used to create the light curve.  If Star A were the source of
 the long-period variability, we would expect the peak variability
@@
 
-Second, we examined the 2-minute lightcurve of each individual pixel
+Second, we examined the 2-minute light curve of each individual pixel
 in the scene, using the interactive tools implemented in
@@
 
-% NB: CDIPS-I only made 131 light-curves for 25Ori-1, because I wasn't
+% NB: CDIPS-I only made 131 light curves for 25Ori-1, because I wasn't
 % aware of the Kounkel+18 member list at the time.
@@
 the astrometric reduced $\chi^2$ for each member.  We then queried the
-CDIPS lightcurve database at MAST \citep{bouma_cluster_2019} to find
+CDIPS light curve database at MAST \citep{bouma_cluster_2019} to find
 the subset of members that were at least as variable as \ptfo.  We
@@
 with variability amplitudes greater than 9.7\% showed higher RUWE.
-One was CVSO~35, which has a TESS light-curve that varies by 2
+One was CVSO~35, which has a TESS light curve that varies by 2
 magnitudes, and shows a strong IR excess and a 10$\mu$m silicate emission
@@
 
-The TESS lightcurve shows a dip that lasts about 45 minutes, and seems
+The TESS light curve shows a dip that lasts about 45 minutes, and seems
 to recur every 10.76 hours
@@
 \citet{tanimoto_evidence_2020}.  It therefore seems credible.
-Inspecting the \citet{tanimoto_evidence_2020} lightcurves, the claim
+Inspecting the \citet{tanimoto_evidence_2020} light curves, the claim
 of multiple dips per cycle in December 2018 at phase 0 and $-0{.}47$
@@
 
-\subsubsection{Similar Lightcurves}
+\subsubsection{Similar Light Curves}
 \label{subsec:dipstars}
@@
 exercise.  By searching the literature, we have found about a dozen
-lightcurves with similar morphologies to \ptfo, drawn from surveys of
+light curves with similar morphologies to \ptfo, drawn from surveys of
 low-mass weak-lined T Tauri stars in regions including $\rho$ Oph,
@@
 These surveys were performed using K2 \citep{howell_k2_2014}.  We
-downloaded some of these lightcurves from MAST, opting for the EVEREST
+downloaded some of these light curves from MAST, opting for the EVEREST
 reductions \citep{luger_everest_2016,luger_update_2018}.  They are
@@
 
-These lightcurves have been phenomenologically classified as
+These light curves have been phenomenologically classified as
 ``persistent flux dips'' or ``transient flux dips'', based on whether
@@
 \citet{stauffer_orbiting_2017}, these objects are morphologically
-distinct from ``scallop shell'' lightcurves, and are present in stars
+distinct from ``scallop shell'' light curves, and are present in stars
 at more advanced evolutionary disk stages than the ``dipper'' stars
@@
     } $\lesssim$ 100$\,$Myr.
-  \item The lightcurves show shallow, angular dips, usually superposed
+  \item The light curves show shallow, angular dips, usually superposed
     on large-amplitude smooth variability. The latter is interpreted
@@
 the first connection of \ptfo\ with this class of objects because the
-TESS lightcurve was needed to resolve the different rotation signals.
+TESS light curve was needed to resolve the different rotation signals.
 
@@
 \begin{itemize}
-  \item {\it The TESS light-curve shows two periodic signals.} The
+  \item {\it The TESS light curve shows two periodic signals.} The
     ``long'' signal is a 10\% peak-to-peak sinusoid that repeats every
@@
     class of transient and persistent flux dip stars.} Analogous
-    lightcurves are shown in Figure~\ref{fig:brethren}.  Properties of
+    light curves are shown in Figure~\ref{fig:brethren}.  Properties of
     this variability class are enumerated in
@@
 studies independently reached the key conclusions that the TESS
-lightcurve shows two periodic signals, and that the properties of
+light curve shows two periodic signals, and that the properties of
 \ptfo\ are consistent with the emerging class of transient and
 persistent flux dip stars.  \citet{koen_2020} reached these
-conclusions by modeling the TESS lightcurve as a truncated sum of
+conclusions by modeling the TESS light curve as a truncated sum of
 Fourier terms, and concluded that the two signals are most simply
@@
 (G011128 and G011132 respectively).
+%
+Resources supporting this work were provided by the NASA High-End Computing (HEC) Program through the NASA Advanced Supercomputing (NAS) Division at Ames Research Center for the production of the SPOC data products.
 %
@@
   %\texttt{BATMAN} \citep{kreidberg_batman_2015},
-  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019}
+  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
   \texttt{corner} \citep{corner_2016},
   %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
-  \texttt{exoplanet} \citep{exoplanet:agol19}
+  \texttt{exoplanet} \citep{exoplanet:agol19},
   \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
   dependencies \citep{exoplanet:agol19, exoplanet:kipping13, exoplanet:luger18,
-  	exoplanet:theano}.
+  	exoplanet:theano},
   \texttt{IPython} \citep{perez_2007},
@@
   \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
-  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
+  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015},
   \texttt{numpy} \citep{walt_numpy_2011}, 
@@
   %\texttt{scikit-learn} \citep{scikit-learn},
-  \texttt{scipy} \citep{jones_scipy_2001}.
+  \texttt{scipy} \citep{jones_scipy_2001},
+  \texttt{SPOC R4.0} \citep{jenkins_tess_2016},
   \texttt{tesscut} \citep{brasseur_astrocut_2019},

### 2020-05-12 billy/plotting.py

@@ -167,7 +167,7 @@ def plot_splitsignal_map(m, outpath, part='i'):
     # 8.5x11 is letter paper. x10 allows space for caption.
     fig, axs = plt.subplots(nrows=4, figsize=(8.5, 10), sharex=True)
 
-    axs[0].set_ylabel('Raw flux', fontsize='x-large')
+    axs[0].set_ylabel('Original flux', fontsize='x-large')
     # axs[0].set_ylabel('f', fontsize='x-large')
     axs[0].plot(m.x_obs[g], m.y_obs[g], ".k", ms=4, label="data", zorder=2,
                 rasterized=True)

### 2020-05-12 paper/ms.tex
merge Dana Louie comments

@@
 % PENDING yy
-\author{D. Louie} % danalouie@astro.umd.edu
-\affiliation{University of Maryland, Baltimore County, 1000 Hilltop
-Circle, Baltimore, MD 21250, USA}
+\author[0000-0002-2457-272X]{D.~R.~Louie} % danalouie@astro.umd.edu
+\affiliation{Department of Astronomy, University of Maryland, College
+Park, MD 20742, USA}
 
@@
 
-The periodogram of the final residual (Figures~\ref{fig:splitsignal}
-and~\ref{fig:splitsignalii} bottom row) shows a barely significant and
+The periodogram of the final residual shows a barely significant and
 poorly-resolved peak at $\approx$8 days, consistent with the visual
-impression of some slower trends in the residuals after subtracting
-the shorter- and longer-period signals.
+impression of some slower trends in the bottom rows of
+Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}.
 
@@
 \begin{align}
-t_0\ {\rm BJD}_{\rm TDB} &= 2455543.943 \pm 0.002 \\
-P &= 0.4483993 \pm 0.0000006 d.
+t_0\ {\rm BJD}_{\rm TDB} &= 2455543.943 \pm 0.002 \nonumber \\
+P &= 0.4483993 \pm 0.0000006\,{\rm d}.
+\label{eq:ephem}
 \end{align}
 
-Figure~\ref{fig:o_minus_c} shows the phase of the dip we detected in
-the TESS data, relative to their ephemeris.  It agrees with the
-independent December 2018 measurements by
+Figure~\ref{fig:o_minus_c} shows the observed mid-transit times of
+dips minus the times predicted from Equation~\ref{eq:ephem}.  The
+phase of the dip we detected in the TESS data (yellow star) agrees
+with the independent December 2018 measurements by
 \citet{tanimoto_evidence_2020}: either the dip abruptly shifted phase

### 2020-05-13 billy/convenience.py
analysis: sample over Mstar and Rstar, too

@@ -6,6 +6,12 @@ from astropy.io import fits
 from astrobase.lcmath import time_bin_magseries_with_errs
 from cdips.lcproc.mask_orbit_edges import mask_orbit_start_and_end
 
+MSTAR_VANEYKEN = 0.39 # 0.34 or 0.39, from Briceno+2005 Baraffe/Siess
+MSTAR_STDEV = 0.25
+RSTAR_VANEYKEN = 1.23 # 1.39 (Brinceno+05) or 1.07 (Table 3 van Eyken+12)
+RSTAR_STDEV = 0.40
+
+
 def chisq(y_mod, y_obs, y_err):
     return np.sum( (y_mod - y_obs )**2 / y_err**2 )
 
@@ -33,7 +39,8 @@ def get_bic(m, ydict, outdir):
     k_tra = 7
     k_Porb = 2*_N # amplitudes
     k_Prot = 2*_M + 2 # amplitudes, plus period and phase
-    k = k_tra + k_Porb + k_Prot
+    k_stparam = 2
+    k = k_tra + k_Porb + k_Prot + k_stparam
 
     n = len(y_obs)
     BIC = bic( χ2, k, n )
@@ -189,6 +196,8 @@ def initialize_ptfo_prior_d(x_obs, modelcomponents):
             prior_d['b'] = 0.5  # initialize for broad prior
             prior_d['u'] = [0.30,0.35] # 0.39Msun, 1.39Rsun -> logg=3.74. Teff 3500K -> Claret18
             prior_d['mean'] = 0
+            prior_d['m_star'] = MSTAR_VANEYKEN
+            prior_d['r_star'] = RSTAR_VANEYKEN
 
         if 'sincos' in modelcomponent:
 

### 2020-05-13 billy/modelfitter.py

@@ -2,6 +2,7 @@ import numpy as np, matplotlib.pyplot as plt, pandas as pd, pymc3 as pm
 import pickle, os
 from copy import deepcopy
 from collections import OrderedDict
+from astropy import units as units, constants as const
 
 import exoplanet as xo
 
@@ -9,6 +10,10 @@ from billy import __path__
 from billy.models import sin_model, cos_model, transit_model
 from billy.plotting import plot_test_data, savefig, plot_MAP_data
 from billy.convenience import flatten as bflatten
+
+from billy.convenience import (
+    MSTAR_VANEYKEN, MSTAR_STDEV, RSTAR_VANEYKEN, RSTAR_STDEV
+)
 
 RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
 
@@ -58,8 +63,8 @@ class ModelFitter(ModelParser):
     σ^2).
     """
 
-    def __init__(self, modelid, x_obs, y_obs, y_err, prior_d, mstar=1,
-                 rstar=1, N_samples=2000, N_cores=16, N_chains=4,
+    def __init__(self, modelid, x_obs, y_obs, y_err, prior_d,
+                 N_samples=2000, N_cores=16, N_chains=4,
                  plotdir=None, pklpath=None, overwrite=1):
 
         self.N_samples = N_samples
@@ -70,8 +75,6 @@ class ModelFitter(ModelParser):
         self.x_obs = x_obs
         self.y_obs = y_obs
         self.y_err = y_err
-        self.mstar = mstar
-        self.rstar = rstar
         self.t_exp = np.nanmedian(np.diff(x_obs))
 
         self.initialize_model(modelid)
@@ -107,6 +110,12 @@ class ModelFitter(ModelParser):
             for modelcomponent in self.modelcomponents:
 
                 if 'transit' in modelcomponent:
+
+                    BoundedNormal = pm.Bound(pm.Normal, lower=0, upper=3)
+                    m_star = BoundedNormal("m_star", mu=MSTAR_VANEYKEN,
+                                           sd=MSTAR_STDEV)
+                    r_star = BoundedNormal("r_star", mu=RSTAR_VANEYKEN,
+                                           sd=RSTAR_STDEV)
 
                     # mean = pm.Normal(
                     #     "mean", mu=prior_d['mean'], sd=0.02, testval=prior_d['mean']
@@ -150,7 +159,7 @@ class ModelFitter(ModelParser):
 
                     orbit = xo.orbits.KeplerianOrbit(
[…]

### 2020-05-13 drivers/real.py
drivers: sample over Mstar/Rstar

@@ -26,7 +26,7 @@ def main(modelid):
     REALID = 'PTFO_8-8695'
     RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
     PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID),
-                           '20200428_v0')
+                           '20200513_v0')
 
     ##########
 
@@ -92,7 +92,7 @@ def main(modelid):
 
 if __name__ == "__main__":
 
-    DEBUG = 1
+    DEBUG = 0
 
     if DEBUG:
         main('transit_2sincosPorb_2sincosProt')

### 2020-05-13 drivers/make_model_comparison_table.py
tables: update for new samples + modelcomp results

@@ -2,7 +2,7 @@ import pickle, os
 import numpy as np, pandas as pd
 from glob import glob
 
-run_id = '20200428_v0'
+run_id = '20200513_v0'
 
 pklpaths = glob(
     '/Users/luke/Dropbox/proj/billy/results/PTFO_8-8695_results/{}/*bicdict*'

### 2020-05-13 drivers/make_posterior_table.py

@@ -17,11 +17,12 @@ def main(modelid):
     REALID = 'PTFO_8-8695'
     RESULTSDIR = os.path.join(os.path.dirname(__path__[0]), 'results')
     PLOTDIR = os.path.join(RESULTSDIR, '{}_results'.format(REALID),
-                           '20200428_v0')
+                           '20200513_v0')
 
     assert modelid in [
         'transit_2sincosPorb_2sincosProt',
-        'transit_2sincosPorb_3sincosProt'
+        'transit_2sincosPorb_3sincosProt',
+        'transit_3sincosPorb_2sincosProt'
     ]
 
     summarypath = os.path.join(
@@ -44,8 +45,12 @@ def main(modelid):
         m = ModelFitter(modelid, x_obs, y_obs, y_err, prior_d, plotdir=PLOTDIR,
                         pklpath=pklpath, overwrite=OVERWRITE)
 
+        # NOTE: you could pass "varnames = varnames", but don't to enable
+        # derived parameter collection
+        # varnames = list(prior_d.keys())
+
         df = pm.summary(
-            m.trace, varnames=list(prior_d.keys()),
+            m.trace,
             round_to=10, kind='stats'
         )
 
@@ -54,16 +59,34 @@ def main(modelid):
     else:
         df = pd.read_csv(summarypath, index_col=0)
 
-    srows = ['period', 't0', 'r', 'b', 'u[0]', 'u[1]', 'mean',
-             'omegaorb', 'Aorb0', 'Borb0', 'Aorb1', 'Borb1',
-             'phirot', 'omegarot', 'Arot0', 'Brot0', 'Arot1', 'Brot1']
+    srows = [
+        # fitted params
+        'period', 't0', 'r', 'b', 'u[0]', 'u[1]', 'mean', 'r_star', 'm_star',
+        'Aorb0', 'Borb0', 'Aorb1', 'Borb1', 'phirot', 'omegarot', 'Arot0',
+        'Brot0', 'Arot1', 'Brot1',
+        # derived params
+        'omegaorb', 'rhostar', 'r_planet', 'a_Rs'
+    ]
     if modelid == 'transit_2sincosPorb_3sincosProt':
-        srows = ['period', 't0', 'r', 'b', 'u[0]', 'u[1]', 'mean',
-                 'omegaorb', 'Aorb0', 'Borb0', 'Aorb1', 'Borb1',
[…]

### 2020-05-13 paper/collect.sh
paper: update figures from 20200513 modelcomp

@@ -1,11 +1,11 @@
 # Usage: copy in figures for the paper
 
-pdir=../results/PTFO_8-8695_results/20200428_v0/
+pdir=../results/PTFO_8-8695_results/20200513_v0/
 
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_i.pdf f1.pdf
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_splitsignalmap_ii.pdf f2.pdf
+cp ${pdir}PTFO_8-8695_transit_3sincosPorb_2sincosProt_splitsignalmap_i.pdf f1.pdf
+cp ${pdir}PTFO_8-8695_transit_3sincosPorb_2sincosProt_splitsignalmap_ii.pdf f2.pdf
 
-cp ${pdir}PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap.pdf f3.pdf
+cp ${pdir}PTFO_8-8695_transit_3sincosPorb_2sincosProt_phasefoldmap.pdf f3.pdf
 
 cp ../results/PTFO_8-8695_results/scene.pdf f4.pdf
 

### 2020-05-13 billy/plotting.py
update brethren plot per 3orb2rot

@@ -1256,7 +1256,7 @@ def plot_brethren(outdir):
 
     lc_dict = pickle.load(open(k2pkl, 'rb'))
 
-    pklpath = '/Users/luke/Dropbox/proj/billy/results/PTFO_8-8695_results/20200413_v0/PTFO_8-8695_transit_2sincosPorb_2sincosProt_phasefoldmap_points.pkl'
+    pklpath = '/Users/luke/Dropbox/proj/billy/results/PTFO_8-8695_results/20200513_v0/PTFO_8-8695_transit_3sincosPorb_2sincosProt_phasefoldmap_points.pkl'
     ptfo_d = pickle.load(open(pklpath, 'rb'))
 
     ##########

### 2020-05-13 paper/ms.tex
paper: update tables with new modelcomp results
paper: cleanup tex
paper: update text per new modelcomp results

@@
 \accepted{---}
-% \submitjournal{AAS journals.}
+\submitjournal{AAS journals.}
 \shorttitle{Two Stars, Two Signals, No Planet}
 
-%\objectname{PTFO 8-8695}
-%\objectname{PTFO 8-8695b}
-%\objectname{CVSO 30}
-
 \begin{document}
@@
 
-%\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
-% \title{Seeing Double: Still Against the Planetary Interpretation of
-% PTFO$\,$8-8695b}
-% \title{Seeing Double: TESS and Gaia Show That PTFO$\,$8-8695b is Unlikely a
-% Planet}
-%\title{Seeing Double: Indications From {\it TESS} and {\it Gaia} That
-%PTFO$\,$8-8695b is Not a Planet}
-%\title{Double Trouble: Transient Dips and Photometric Binarity in
-%PTFO$\,$8-8695}
 % \title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
 % PTFO 8-8695}
-% \title{PTFO\,8-8695: A binary star, not a hot Jupiter host}   
-%\title{Two Stars and Two Signals in PTFO 8-8695}
 \title{PTFO 8-8695: Two Stars, Two Signals, No Planet}
@@
 
-% PENDING yy
+%CONFIRMED
 \author[0000-0002-2457-272X]{D.~R.~Louie} % danalouie@astro.umd.edu
@@
     a model including 2 harmonics at the longer period $P_{\rm \ell}$,
-    plus 2 harmonics and a transit at the shorter period $P_{\rm s}$.
+    plus 3 harmonics and a transit at the shorter period $P_{\rm s}$.
     {\it Upper middle}: Longer-period signal, equal to the original signal
@@
 \begin{equation}
-  %{\rm PTFO\,8\text{-}8695}
-  % \in
   {\rm 25\,Ori\text{-}1}
@@
 		\leavevmode
-		\includegraphics[width=0.93\textwidth]{f3.pdf}
+		\includegraphics[width=1\textwidth]{f3.pdf}
 	\end{center}
@@
     includes 2 harmonics at the longer period,
-    plus 2 harmonics and a transit at the shorter period.
+    plus 3 harmonics and a transit at the shorter period.
 		\label{fig:phasefold}
@@
 darkening parameters, the planet's orbital period (set equal to
-$P_{\rm s}$) the time of a particular transit, and the mean flux.
-There would be $2N_{\rm s}=4$ additional Fourier amplitudes at the
-shorter period, plus $2N_{\rm \ell}=2$ Fourier amplitudes at the
-longer period, as well as $P_\ell$ itself and the relative phase
-$\phi_\ell$.  The total number of parameters is 14.
+$P_{\rm s}$) the time of a particular transit, and the mean flux.  We
+also independently sampled the stellar radius and mass, implicitly
+defining the stellar density.  Additionally, there would be $2N_{\rm
+s}=4$ additional Fourier amplitudes at the shorter period, plus
+$2N_{\rm \ell}=2$ Fourier amplitudes at the longer period, as well as
+$P_\ell$ itself and the relative phase $\phi_\ell$.  The total number
+of parameters would be 17.
 
@@
 our preferred model, we used the Bayesian information criterion
-(Table~1).  The model with the lowest BIC had two modes at the shorter
-10.76$\,$hr period, and two modes at the longer 11.98$\,$hr period.
-Two other models had BIC values that provided comparable evidence
-\citep{burnham_multimodel_2016}: one with ($N_{\rm s}=3$, $N_{\rm
-\ell}=3$), and the other with ($N_{\rm s}=3$, $N_{\rm \ell}=2$).
+(Table~1).  The model with the lowest BIC had three modes at the
+shorter 10.76$\,$hr period, and two modes at the longer 11.98$\,$hr
+period.
+The other models had BIC values that implied significantly less
+support \citep{burnham_multimodel_2016}.
 All nine models have
-reduced $\chi^2$ values ranging between 1.25 and 1.68, which suggests
+reduced $\chi^2$ values ranging between 1.21 and 1.68, which suggests
 a plausible though imperfect agreement between the data and the model
@@
 phases $-0.5$ and $0.0$ there appears to be complex shorter-timescale
-variability, including an abrupt jump at phase $\approx -0.33$, and
-ending with a ``dip'' of depth $\approx$1.2\%, lasting $\approx$0.75
-hours.  The fact that the higher-order models lead to roughly
-comparable BIC values suggests that these sharper features may indeed
-be real, but that there is only weak statistical support for their
-existence.
+variability, ending with a ``dip'' of depth $\approx$1.2\%, lasting
+$\approx$0.75 hours.  The fact that our preferred model has three
+rather than two ``short period'' harmonics is linked to the degree of
+curvature required between phases $-0.5$ and $-0.05$: the analogous
+$(N_{\rm \ell},N_{\rm s})=(2,2)$ model prefers a longer transit duration,
+but does not fit the out-of-transit curvature as well, particularly
+immediately before ingress.
 
@@
 detected (Table~2).  The first sine and cosine harmonic both have
-amplitudes of roughly $0.9\pm0.1\%$.  The second sine harmonic has
-amplitude $0.16 \pm 0.07\%$, so is non-zero at a significance of only
-2.3$\sigma$.  The second cosine harmonic has negative amplitude $0.53
-\pm 0.06\%$.  In our sign convention, the fact that it is negative
+amplitudes of roughly $0.90\pm0.04\%$.  The second sine harmonic has
+amplitude $0.16 \pm 0.04\%$, so is non-zero at a significance of
+$\approx$4$\sigma$.  The second cosine harmonic has negative amplitude $0.55
+\pm 0.03\%$.  In our sign convention, the fact that it is negative
 means that this component peaks at phase 0.25 and 0.75.
@@
 The planetary interpretation however does not readily explain the
-first sine and cosine harmonics.  Interpreting the sine component as
+large first sine and cosine harmonics.  Interpreting the sine component as
 Doppler beaming would imply a secondary mass greater than the primary
@@
     processing.  Within the angular resolution of the Gaia source
-    catalog, both signals need to originate from \ptfo.
+    catalog, both signals originate from \ptfo.
   \item {\it The Gaia data imply binarity.} Relative to stars in its
@@
 clumps of gas at the Keplerian corotation radius, and occultations of
-starspots by an optically thick disk \citep[{\it
+starspots by a tenuous gas disk \citep[{\it
 e.g.},][]{stauffer_orbiting_2017,david_transient_2017,zhan_complex_2019}.
@@
 others had hoped, understanding \ptfo\ and its analogs is a worthy
-problem that might even teach us about the natal environments of the
-majority of habitable-zone Earth-sized planets in the Milky Way
+problem.  It might even teach us about the birth environments of the
+majority of habitable-zone Earth-sized planets
 \citep{dressing_occurrence_2013}.
@@
 provide independent support for the conclusion that \ptfo\ is a
-binary.
-
-Minor differences in our approaches are that \citet{koen_2020}
-interprets the ``out of dip'' signal to potentially show multiple
-dips, while our BIC-based model comparison did not indicate that
-including additional harmonics past the second order was strictly
-necessary (Table~1). Our analysis emphasized the agreement between the
+binary, and emphasized the agreement between the
 TESS dip ephemeris and that from \citet{tanimoto_evidence_2020}.
-\citet{koen_2020} however made a clearer case for testing the binarity
-hypothesis using interferometric imaging and high-precision radial
-velocities.
 \\
@@
 
-%
-% The following are entries from Table 1 that are not otherwise cited
-% in the text
-%
-% \nocite{wilson_wasp-4b_2008}
-% \nocite{gillon_improved_2009}
-% \nocite{winn_transit_2009}
-% \nocite{hoyer_tramos_2013}
-% \nocite{dragomir_terms_2011}
-% \nocite{sanchis-ojeda_starspots_2011}
-% \nocite{nikolov_wasp-4b_2012}
-% \nocite{ranjan_atmospheric_2014}
-% \nocite{huitson_gemini_2017}
-
-% \input{WASP-4b_transit_time_table.tex}
-% \input{WASP-4b_rv_table.tex}
-% \input{model_fit_table.tex}
-% \input{rv_model_posterior_table.tex}
-% \input{pdot_table.tex}
-
 % \clearpage

### 2020-05-13 paper/ms.tex
merge Roland Vanderspek comments

@@
 %
-%PENDING yy
+%CONFIRMED
 \author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu
@@
 a T Tauri star in the Orion-OB1a cluster.  It would have the shortest
-orbital period (10.7 hours) of any hot Jupiter.  With such a short
+orbital period of any hot Jupiter.  With such a short
 period, it would probably be filling its Roche lobe, and actively
@@
 
-While the planetary interpetation clearly faces challenges, there is
+While the planetary interpretation clearly faces challenges, there is
 no completely satisfactory alternate explanation.  High-latitude
@@
 Section~\ref{sec:tess}.  The Gaia data, analyzed in
-Section~\ref{sec:gaia}, show that \ptfo\ is a photometric binary, and
+Section~\ref{sec:gaia}, show that \ptfo\ is a photometric binary and
 suggest it is also an astrometric binary.  We discuss the pieces of
@@
 \citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
-The two tallest peaks occur at 0.448\,days and 0.499\,days.  We will
-refer to these two periods as the ``shorter period'' $P_{\rm s}$ and
-the ``longer period'' $P_{\rm \ell}$.  The peak at $P_{\rm \ell}$ is
-the taller of the two peaks.  Lower-power harmonics of both signals
-are also present.
+The two tallest peaks occur at 0.448\,d (10.76$\,$hr) and 0.499\,d
+(11.98$\,$hr).  We will refer to these two periods as the ``shorter
+period'' $P_{\rm s}$ and the ``longer period'' $P_{\rm \ell}$.  The
+peak at $P_{\rm \ell}$ is the taller of the two peaks.  Lower-power
+harmonics of both signals are also present.
 
@@
 the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
-initialized each model with the parameters of the maximum a posteriori
-(MAP) model.  We then assumed a Gaussian likelihood, and sampled using
-\texttt{PyMC3}'s gradient-based No-U-Turn Sampler
+initialized each model with the parameters of the maximum {\it a
+posteriori} (MAP) model.  We then assumed a Gaussian likelihood, and
+sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
 \citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
@@
 
-In every model, the $11.98\,{\rm hr}$ variability is a simple sinusoid
-with peak-to-peak amplitude $\approx$10\%.  The $10.76\,{\rm hr}$
+In every model, the 11.98$\,$hr variability is a simple sinusoid
+with peak-to-peak amplitude $\approx$10\%.  The 10.76$\,$hr
 variability is always more complex.  The overall impression is of a
@@
     ``long'' signal is a 10\% peak-to-peak sinusoid that repeats every
-    11.98 hours.  The ``short'' signal is a 4\% peak-to-peak ``dip +
-    asymmetric sinusoid'' that repeats every 10.76 hours. The signals
+    11.98$\,$hr.  The ``short'' signal is a 4\% peak-to-peak ``dip +
+    asymmetric sinusoid'' that repeats every 10.76$\,$hr. The signals
     beat, and therefore cannot be an artifact linked to data

### 2020-05-13 paper/ms.tex
manuscript: fix column width

@@
 a T Tauri star in the Orion-OB1a cluster.  It would have the shortest
-orbital period of any hot Jupiter.  With such a short
-period, it would probably be filling its Roche lobe, and actively
-losing mass to its host star.  Not only that, but the rapidly-rotating
-host star is probably oblate enough to torque the planet's orbit into
-and out of the transiting configuration on a time scale of years
+orbital period of any hot Jupiter.  With such a short period, it would
+probably be filling its Roche lobe, and actively losing mass to its
+host star.  Not only that, but the rapidly-rotating host star is
+probably oblate enough to torque the planet's orbit into and out of
+the transiting configuration on a time scale of years
 \citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.
@@
     {\bf TESS light curve of \ptfo\ (Sector 6, Orbit 19).} {\it Top}:
-    The original (\texttt{PDCSAP} mean-subtracted) relative flux.
-    The beat period of 4.48 days is visible by eye.  The blue curve is
-    a model including 2 harmonics at the longer period $P_{\rm \ell}$,
+    The original (\texttt{PDCSAP} mean-subtracted) relative flux.  The
+    beat period of 4.48 days is visible by eye.  The blue curve is a
+    model including 2 harmonics at the longer period $P_{\rm \ell}$,
     plus 3 harmonics and a transit at the shorter period $P_{\rm s}$.
-    {\it Upper middle}: Longer-period signal, equal to the original signal
-    minus the shorter-period signal.  {\it Lower middle}:
+    {\it Upper middle}: Longer-period signal, equal to the original
+    signal minus the shorter-period signal.  {\it Lower middle}:
     Shorter-period signal, equal to the original signal minus the
@@
 	\vspace{-0.7cm}
-  \caption{ {\bf TESS light curve of \ptfo\ (Sector 6, Orbit 20).} Same
-  format as Figure~\ref{fig:splitsignal}.
+  \caption{ {\bf TESS light curve of \ptfo\ (Sector 6, Orbit 20).}
+  Same format as Figure~\ref{fig:splitsignal}.
   \label{fig:splitsignalii}
@@
 Ames~\citep{jenkins_tess_2016}.  We mainly used the Presearch Data
-Conditioning (PDC) light curve.  The PDC light curve aperture used
-pixels chosen to maximize the SNR of the total flux of the target
-\citep{bryson_2020_target_aperture}.  Non-astrophysical variability
-was removed by fitting out trends common to many stars
-\citep{smith_kepler_2012,stumpe_multiscale_2014}.
-% \citep{smith_kepler_PDC_2017}
-% \citep{smith_kepler_apertures_2017}
+Conditioning Simple Aperture Photometry (PDCSAP) light curve.  The PDC
+light curve aperture used pixels chosen to maximize the SNR of the
+total flux of the target \citep{bryson_2020_target_aperture}.
+Non-astrophysical variability was removed by fitting out trends common
+to many stars \citep{smith_kepler_2012,stumpe_multiscale_2014}.
 
 As an independent check on the 2-minute SPOC light curve, we examined
-the light curve based upon 30-minute image stacks which was produced as
-part of the Cluster Difference Imaging Photometric Survey (CDIPS;
+the light curve based upon 30-minute image stacks which was produced
+as part of the Cluster Difference Imaging Photometric Survey (CDIPS;
 \citealt{bouma_cluster_2019}).  Our CDIPS light curve of choice used a
@@
 often systematic red noise in the photometry during those times.  Both
-the CDIPS and PDC light curves showed a clear discontinuous jump in the
-last few days of orbit 20, which seemed likely to be an instrumental
-systematic effect.  We correspondingly masked out the data with
-timestamps ranging from BJD\,2458488.3 until the end of the orbit.
-The PDC light curve initially had 15{,}678 points.  The quality-flag
-cut removed 854 points; masking the orbit edges removed an additional
-716 points; and removing the data from the final few days of orbit 20
-removed an additional 1079 points.  After cleaning, 83\% of the
-initial flux measurements remained.
-
-We normalized the light curve by dividing out the median flux, and then
-opted to subtract 1{.}0 to set the median value to zero, which
+the CDIPS and PDC light curves showed a clear discontinuous jump in
+the last few days of orbit 20, which seemed likely to be an
+instrumental systematic effect.  We correspondingly masked out the
+data with timestamps ranging from BJD\,2458488.3 until the end of the
+orbit.  The PDC light curve initially had 15{,}678 points.  The
+quality-flag cut removed 854 points; masking the orbit edges removed
+an additional 716 points; and removing the data from the final few
+days of orbit 20 removed an additional 1079 points.  After cleaning,
+83\% of the initial flux measurements remained.
+
+We normalized the light curve by dividing out the median flux, and
+then opted to subtract 1{.}0 to set the median value to zero, which
 simplified subsequent interpretation.  Many of these and subsequent
@@
     points are the 10 minute cadence \texttt{PDCSAP} flux.  Black
-    points are binned to 100 points per period.
-    The model (blue line)
-    includes 2 harmonics at the longer period,
-    plus 3 harmonics and a transit at the shorter period.
+    points are binned to 100 points per period.  The model (blue line)
+    includes 2 harmonics at the longer period, plus 3 harmonics and a
+    transit at the shorter period.
 		\label{fig:phasefold}
@@
 
-The peak-to-peak light curve amplitude at maximum, when the two signals
-interfere constructively, is about 14\%.  During the times of
+The peak-to-peak light curve amplitude at maximum, when the two
+signals interfere constructively, is about 14\%.  During the times of
 destructive interference, the peak-to-peak amplitude is about 6\%.
@@
 shorter 10.76$\,$hr period, and two modes at the longer 11.98$\,$hr
-period.
-The other models had BIC values that implied significantly less
-support \citep{burnham_multimodel_2016}.
-All nine models have
+period.  The other models had BIC values that implied significantly
+less support \citep{burnham_multimodel_2016}.  All nine models have
 reduced $\chi^2$ values ranging between 1.21 and 1.68, which suggests
@@
 
-In every model, the 11.98$\,$hr variability is a simple sinusoid
-with peak-to-peak amplitude $\approx$10\%.  The 10.76$\,$hr
-variability is always more complex.  The overall impression is of a
-distorted sinusoidal function, with a peak-to-peak amplitude of about
-4\%.  The asymmetric sinusoid rises to a maximum near phase 0.25, and
-reaches minimum brightness between phases $-0.5$ and $-0.25$.  Between
-phases $-0.5$ and $0.0$ there appears to be complex shorter-timescale
+In every model, the 11.98$\,$hr variability is a simple sinusoid with
+peak-to-peak amplitude $\approx$10\%.  The 10.76$\,$hr variability is
+always more complex.  The overall impression is of a distorted
+sinusoidal function, with a peak-to-peak amplitude of about 4\%.  The
+asymmetric sinusoid rises to a maximum near phase 0.25, and reaches
+minimum brightness between phases $-0.5$ and $-0.25$.  Between phases
+$-0.5$ and $0.0$ there appears to be complex shorter-timescale
 variability, ending with a ``dip'' of depth $\approx$1.2\%, lasting
@@
 curvature required between phases $-0.5$ and $-0.05$: the analogous
-$(N_{\rm \ell},N_{\rm s})=(2,2)$ model prefers a longer transit duration,
-but does not fit the out-of-transit curvature as well, particularly
-immediately before ingress.
+$(N_{\rm \ell},N_{\rm s})=(2,2)$ model prefers a longer transit
+duration, but does not fit the out-of-transit curvature as well,
+particularly immediately before ingress.
 
@@
 
-Stronger constraints on possible
-stellar companions were obtained by
-\citet{van_eyken_ptf_2012} through high-resolution imaging with the NIRC2
-camera on Keck II.  They reported 3-$\sigma$ $H$-band magnitude
+Stronger constraints on possible stellar companions were obtained by
+\citet{van_eyken_ptf_2012} through high-resolution imaging with the
+NIRC2 camera on Keck II.  They reported 3-$\sigma$ $H$-band magnitude
 difference limits of 4.3, 6.4, and 8.9 at angular separations of 0.25,
@@
 One was CVSO~35, which has a TESS light curve that varies by 2
-magnitudes, and shows a strong IR excess and a 10$\mu$m silicate emission
-feature \citep{mauco_herschel_2018}.  The other is GAIA DR2
+magnitudes, and shows a strong IR excess and a 10$\mu$m silicate
+emission feature \citep{mauco_herschel_2018}.  The other is GAIA DR2
 3222210363837122048.
@@
 
-Long-baseline radial velocity (RV) measurements could also reveal
-the presence of multiple stars in this system.
-Unfortunately,
-the available RV data for \ptfo\ is rather sparse, presumably due to the
-stellar faintness and equatorial velocity.
-The RV datasets with the longest baselines we could find
-in the literature were those reported by \citet{van_eyken_ptf_2012}.
-These included 5 Keck/HIRES measurements acquired
-over 10 days in April 2011, and 4 HET/HRS measurements acquired over
-10 days in February 2011.  The root-mean-squared RV over each 10 day
-span was $\approx$$2\,{\rm km}\,{\rm s}^{-1}$, consistent with the
-measurement precision.  Although \citet{van_eyken_ptf_2012} tried a
-CCF-based RV reduction technique, they eventually found that manually
-selecting absorption lines and measuring line centroids was more
-effective.  While \citet{yu_tests_2015} acquired 22 further Keck/HIRES
-spectra over one night in December 2013, those points were not reduced
-to velocities, and it is not clear whether the appropriate calibration
-data exist.  Further Keck/HIRES measurements of \ptfo, if
-appropriately calibrated, could potentially confirm or refute the
-presence of binary companions.
+Long-baseline radial velocity (RV) measurements could also reveal the
+presence of multiple stars in this system.  Unfortunately, the
+available RV data for \ptfo\ is rather sparse, presumably due to the
+stellar faintness and equatorial velocity.  The RV datasets with the
+longest baselines we could find in the literature were those reported
+by \citet{van_eyken_ptf_2012}.  These included 5 Keck/HIRES
+measurements acquired over 10 days in April 2011, and 4 HET/HRS
+measurements acquired over 10 days in February 2011.  The
+root-mean-squared RV over each 10 day span was $\approx$$2\,{\rm
+km}\,{\rm s}^{-1}$, consistent with the measurement precision.
+Although \citet{van_eyken_ptf_2012} tried a CCF-based RV reduction
+technique, they eventually found that manually selecting absorption
+lines and measuring line centroids was more effective.  While
+\citet{yu_tests_2015} acquired 22 further Keck/HIRES spectra over one
+night in December 2013, those points were not reduced to velocities,
+and it is not clear whether the appropriate calibration data exist.
+Further Keck/HIRES measurements of \ptfo, if appropriately calibrated,
+could potentially confirm or refute the presence of binary companions.
 
@@
 
-The TESS light curve shows a dip that lasts about 45 minutes, and seems
-to recur every 10.76 hours
+The TESS light curve shows a dip that lasts about 45 minutes, and
+seems to recur every 10.76 hours
 (Figures~\ref{fig:splitsignal},~\ref{fig:splitsignalii},~\ref{fig:phasefold}).
@@
 amplitude $0.16 \pm 0.04\%$, so is non-zero at a significance of
-$\approx$4$\sigma$.  The second cosine harmonic has negative amplitude $0.55
-\pm 0.03\%$.  In our sign convention, the fact that it is negative
-means that this component peaks at phase 0.25 and 0.75.
+$\approx$4$\sigma$.  The second cosine harmonic has negative amplitude
+$0.55 \pm 0.03\%$.  In our sign convention, the fact that it is
+negative means that this component peaks at phase 0.25 and 0.75.
 
@@
 The planetary interpretation however does not readily explain the
-large first sine and cosine harmonics.  Interpreting the sine component as
-Doppler beaming would imply a secondary mass greater than the primary
-($0.86\,M_\odot$).  Interpreting the cosine component as reflected or
-emitted light from the planet's surface is nonsensical because the
-sign is wrong---the planet would need to be {\it absorbing} light.
+large first sine and cosine harmonics.  Interpreting the sine
+component as Doppler beaming would imply a secondary mass greater than
+the primary ($0.86\,M_\odot$).  Interpreting the cosine component as
+reflected or emitted light from the planet's surface is nonsensical
+because the sign is wrong---the planet would need to be {\it
+absorbing} light.
 
@@
     Five transient and persistent flux dip stars selected based on
-    their visual similarity to the short-period signal in \ptfo\ are as
-    shown.
-    They include
-    EPIC 204143627,
-    EPIC 204270520,
-    EPIC 204321142,
-    EPIC 246938594,
-    and
-    EPIC 205483258 (RIK-210).
-    RIK-210 has the longest period of any of these objects.  All the
-    analogs displayed are either in Taurus or Upper~Sco, and meet the
+    their visual similarity to the short-period signal in \ptfo\ are
+    as shown.  They include EPIC 204143627, EPIC 204270520, EPIC
+    204321142, EPIC 246938594, and EPIC 205483258 (RIK-210).  RIK-210
+    has the longest period of any of these objects.  All the analogs
+    displayed are either in Taurus or Upper~Sco, and meet the
     characteristics of Section~\ref{subsec:dipstars}.  We found these
@@
 These surveys were performed using K2 \citep{howell_k2_2014}.  We
-downloaded some of these light curves from MAST, opting for the EVEREST
-reductions \citep{luger_everest_2016,luger_update_2018}.  They are
-plotted in Figure~\ref{fig:brethren}.
+downloaded some of these light curves from MAST, opting for the
+EVEREST reductions \citep{luger_everest_2016,luger_update_2018}.  They
+are plotted in Figure~\ref{fig:brethren}.
 
@@
 (ii)} transits of enshrouded protoplanets; {\it (iii)} occultations of
-starspots by an optically thick disk.  
-The first and last explanations
+starspots by an optically thick disk.  The first and last explanations
 have added appeal because they are flexible enough to explain not only

### 2020-05-14 paper/ms.tex
minor
Update ms.tex

@@
   Orion-OB1a cluster that shows brightness dips that resemble
-  planetary transits.  While strong evidence against the planet
-  hypothesis has already been provided, the possibility remains
+  planetary transits.  Although strong evidence against the planet
+  hypothesis has been presented, the possibility remains
   debated in the literature.  To obtain further clues, we inspected
   data from the NASA {\it Transiting Exoplanet Survey Satellite}
-  (TESS) and the ESA Gaia mission.  The Gaia data show that
-  PTFO$\,$8-8695 is a photometric binary with respect to members of
-  its kinematic group, and independently suggest that it is an
-  astrometric binary.  The TESS light curve shows two different
+  (TESS) and the ESA Gaia mission.  The Gaia data suggest that
+  PTFO$\,$8-8695 is a binary: the photometric data show
+  it to be overluminous with respect to members of
+  its kinematic group, and the astrometric data are inconsistent
+  with a single star.  The TESS light curve shows two different
   photometric periods. The variability is dominated by a sinusoidal
@@
   pair of young and rapidly rotating M dwarfs, one of which shows the
-  same ``transient-dipper'' behavior that been seen in at least 5
+  same ``transient-dipper'' behavior that has been seen in at least 5
   other cases.  The origin of these transient dips is still unknown
@@
 but require fine-tuning to produce dips of the appropriate duration.
-Furthermore, \ptfo\ does not have an infrared (IR) excess associated
-with the presence of dust at the inner edge of a primordial disk
+Transits by dust clumps or other dusty features are questionable
+because \ptfo\ does not have a detectable infrared (IR) excess associated
+with the presence of warm dust
 \citep[{\it e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude
-starspots, hot or cold, struggle to produce photometric features as
+starspots, hot or cold, would struggle to produce photometric features as
 short as some of the observed dips.
@@
 A relevant fact is that between 0.1\% and 1\% of rapidly rotating
-low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations show
+low-mass stars in $\mathcal{O}$\,(10)$\,$Myr old associations show
 short-duration dips as part of their overall periodic variability
@@
 Section~\ref{sec:tess}.  The Gaia data, analyzed in
-Section~\ref{sec:gaia}, show that \ptfo\ is a photometric binary and
-suggest it is also an astrometric binary.  We discuss the pieces of
+Section~\ref{sec:gaia}, show that \ptfo\ is too bright to be a single star and
+also suggest it is an astrometric binary.  We discuss the pieces of
 the puzzle in Section~\ref{sec:discussion}, and summarize the
-situation in Section~\ref{sec:conclusions}.  We comment on the study
-by \citet{koen_2020} in postscript.
+situation in Section~\ref{sec:conclusions}.  In a postscript, we
+comment on a recent study by \citet{koen_2020} which reached similar
+conclusions.
 
@@
     {\bf TESS light curve of \ptfo\ (Sector 6, Orbit 19).} {\it Top}:
-    The original (\texttt{PDCSAP} mean-subtracted) relative flux.  The
+    The original (\texttt{PDCSAP} median-subtracted) relative flux.  The
     beat period of 4.48 days is visible by eye.  The blue curve is a
@@
     longer-period signal.  {\it Bottom}: residual relative flux.  The
-    data are binned from 2 to 10 minute cadence as a convenience for
+    data are binned from 2 to 10 minute cadence for convenience in
     plotting and fitting.
@@
 January 6, 2019, during the sixth sector of science operations
-\citep{ricker_transiting_2015}.  The star was designated
+\citep{ricker_transiting_2015}.  The star is designated
 TIC\,264461976 in the TESS Input Catalog
@@
 The 2-minute stacks for \ptfo\ were reduced to light curves by the
-Science Processing Operations Center (SPOC) at NASA
-Ames~\citep{jenkins_tess_2016}.  We mainly used the Presearch Data
+Science Processing Operations Center (SPOC) at the NASA
+Ames Research Center~\citep{jenkins_tess_2016}.  We mainly used the Presearch Data
 Conditioning Simple Aperture Photometry (PDCSAP) light curve.  The PDC
-light curve aperture used pixels chosen to maximize the SNR of the
+light curve is based on pixels chosen to maximize the SNR of the
 total flux of the target \citep{bryson_2020_target_aperture}.
@@
 data from the first and last 6 hours of each orbit, since there are
-often systematic red noise in the photometry during those times.  Both
-the CDIPS and PDC light curves showed a clear discontinuous jump in
+often systematic effects in the photometry during those times.  Both
+the CDIPS and PDC light curves showed a discontinuous jump in
 the last few days of orbit 20, which seemed likely to be an
-instrumental systematic effect.  We correspondingly masked out the
+instrumental systematic effect, and led us to mask out the
 data with timestamps ranging from BJD\,2458488.3 until the end of the
@@
 an additional 716 points; and removing the data from the final few
-days of orbit 20 removed an additional 1079 points.  After cleaning,
+days of orbit 20 removed an additional 1{,}079 points.  After cleaning,
 83\% of the initial flux measurements remained.
@@
 \citep{gaia_collaboration_gaia_2016,lindegren_gaiasoln_2018,gaia_collaboration_gaia_2018}.
-In the Gaia second data release (DR2), these CCD observations were
+For the Gaia second data release (DR2), these CCD observations were
 used to determine positions, proper motions, and parallaxes of the
-brighest 1{.}3 billion stars, including \ptfo\
+brighest 1{.}3 billion stars
 \citep{lindegren_gaiasoln_2018}.  For \ptfo, there were 121 ``good''
@@
 down-weighted in the astrometric solution.  \ptfo\ was assigned the
-Gaia DR2 identifier 3222255959210123904.  Its photometric brightness
+Gaia DR2 identifier 3222255959210123904.  Its brightness
 was measured using selected bands ($G$, $Rp$, and $Bp$) of the Gaia
@@
 
-The majority of Gaia's derived parameters for \ptfo\ agreed with
-expectations from former studies
+The majority of Gaia's derived parameters for \ptfo\ agree with
+expectations based on previous studies
 \citep{briceno_cida_2005,van_eyken_ptf_2012}.  The main novelty is
@@
 complex in particular has numerous subgroups, with ages ranging from
-0.5 to 15$\,$Myr. For an incomplete sampling, see for instance
+0.5 to 15$\,$Myr. See, for instance,
 \citet{briceno_cida_2005,jeffries_kinematic_2006,briceno_25_2007,kounkel_apogee2_2018}
@@
 
-\ptfo\ was initially identified as a member of the Orion$\,$OB1a
-sub-association and designated CVSO\,30 by \citet{briceno_cida_2005}
-through combined photometry and spectroscopy.  Later work by
-\citet{briceno_25_2007} clarified that \ptfo\ was in a kinematically
+\ptfo\ was initially designated CVSO\,30 and identified as a member of the Orion$\,$OB1a
+sub-association by \citet{briceno_cida_2005},
+based on photometry and spectroscopy.  Later work by
+\citet{briceno_25_2007} clarified that \ptfo\ is in a kinematically
 distinct subgroup of Orion~OB1a, named the ``25$\,$Ori'' group after
@@
 \end{equation}
-where `$\subset$' means ``is a proper subset of''.  25$\,$Ori-1 is the
+where `$\subset$' means `is a proper subset of'.  25$\,$Ori-1 is the
 largest subgroup of 25$\,$Ori, with 149 identified members.  Based on
@@
 6.9$\,$Myr, while the HR diagram led to an estimated age of 8.5$\,$Myr
+% JNW: What is the difference between HR and CMD?
 \citep[see][Section 2.3]{kounkel_apogee2_2018}.
@@
 
-These details concerning the group membership for one object are
-cumbersome to those accustomed to the simple distinction between
+These details concerning the group membership for one object may
+seem excessive to those accustomed to the simple distinction between
 ``young cluster members'' and ``old field stars''.  Although all
@@
 	\caption{ {\bf Phase-folded longer and shorter-period signals.}
-    {\it Top}: Relative flux at the longer period, as in
-    Figure~\ref{fig:splitsignal}.  {\it Bottom}: Relative flux at the
-    shorter period. The reference phase is set to the dip.  Gray
-    points are the 10 minute cadence \texttt{PDCSAP} flux.  Black
+    {\it Top}: The longer-period signal.  {\it Bottom}: The
+    shorter-period signal. The phase is defined such that the dip occurs at zero phase.  Gray
+    points are the \texttt{PDCSAP} data binned to 10-minute cadence.  Black
     points are binned to 100 points per period.  The model (blue line)
@@
 \citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
-The two tallest peaks occur at 0.448\,d (10.76$\,$hr) and 0.499\,d
-(11.98$\,$hr).  We will refer to these two periods as the ``shorter
-period'' $P_{\rm s}$ and the ``longer period'' $P_{\rm \ell}$.  The
-peak at $P_{\rm \ell}$ is the taller of the two peaks.  Lower-power
+The tallest peak occurs at 0.499\,d (11.98$\,$hr) and a second strong
+peak occurs at 0.448\,d (10.76$\,$hr). We will refer to these two periods as the ``longer
+period'' $P_{\rm \ell}$ and the ``shorter period'' $P_{\rm s}$.  Lower-power
 harmonics of both signals are also present.
@@
 
-The peak-to-peak light curve amplitude at maximum, when the two
+The peak-to-peak maximum amplitude of the light curve, when the two
 signals interfere constructively, is about 14\%.  During the times of
@@
 the phase-folded light curves of the longer-period signal, we
-subtracted the best-fitting sinusoid at the shorter period; the
+subtracted the best-fitting sinusoid with a period equal to $P_{\rm s}$. The
 resulting light curve appears smooth and nearly sinusoidal.  But after
-subtracting the best-fitting sinusoid at the longer period, visual
-inspection of the phase-folded light curve of the shorter-period
-signal revealed substructure resembling the ``dips'' seen in previous
+subtracting the best-fitting sinusoid with a period equal to $P_{\ell}$, visual
+inspection of the phase-folded light curve revealed substructure resembling the ``dips'' seen in previous
 observations. In particular, there was a $\approx$1\% dip lasting
@@
 
-As an example, one possible model consists of a transit, $N_{\rm s}=2$
-sines and cosines at the shorter period, plus $N_{\rm \ell}=1$ sine
-and cosine at the longer period.  In this case, the free parameters
-would be as follows.  The transit model parameters are the impact
+The free parameters are as follows.  The transit model parameters are the impact
 parameter, the planet-to-star radius ratio, two quadratic limb
@@
 $P_{\rm s}$) the time of a particular transit, and the mean flux.  We
-also independently sampled the stellar radius and mass, implicitly
-defining the stellar density.  Additionally, there would be $2N_{\rm
+sampled the stellar radius and mass from prior probability distributions, implicitly
+defining the stellar density which (together with the orbital period) sets the transit timescale. 
+There are also the parameters defining the Fourier modes.
+As an example, one possible model consists of a transit, $N_{\rm s}=2$
+sines and cosines at the shorter period, plus $N_{\rm \ell}=1$ sine
+and cosine at the longer period.  There are $2N_{\rm
 s}=4$ additional Fourier amplitudes at the shorter period, plus
@@
 $P_\ell$ itself and the relative phase $\phi_\ell$.  The total number
-of parameters would be 17.
+of parameters is 17 for this case.
 
@@
 To speed up the fitting process, we averaged the 2-minute light curve
-to 10-minute samples.  We correspondingly scaled the uncertainties in
+into 10-minute samples.  We correspondingly scaled down the uncertainties in
 the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
@@
 We considered nine models, with the number of modes per frequency
-$N_{\rm s}$ and $N_{\rm \ell}$ ranging from one to three.  To select
+($N_{\rm s}$ and $N_{\rm \ell}$) ranging from one to three.  To select
 our preferred model, we used the Bayesian information criterion
@@
 
-The periodogram of the final residual shows a barely significant and
+The periodogram of the residuals between the data and the preferred model shows a barely significant and
 poorly-resolved peak at $\approx$8 days, consistent with the visual
@@
 
-The TESS pixels are $\approx21$'' per side. Before making any
+The portion of the sky subtended by each TESS pixel is about 21'' on a side. Before making any
 interpretations, we needed to consider whether light from neighboring
@@
 cannot contribute more than about 10\% to the total signal.  The only
-neighbor that is sufficiently close and bright that its light might
-contaminate the target star is TIC 264461980, with $T=14.8$, which we
-dub ``Star A''.  Star A is 23.6'' northwest of the target, and based
-on the magnitude difference could contribute flux variations as large
-as 48\% the flux of our target star, \ptfo.
+other known star that is sufficiently close and bright that its light might
+contaminate the signal from the target star is TIC 264461980, with $T=14.8$.
+This star, we we dub ``Star A'', is 23.6'' northwest of the target. Based
+on the magnitude difference, Star A could contribute flux variations as large
+as 48\% of the flux of the target star.
 
@@
 nearest to Star A to show a sinusoidal signal with amplitude exceeding
-$10\%$.  The data do not show this pattern.  The pixel directly below
-Star A does not clearly show any sinusoidal variability; the
-peak-to-peak variability in that pixel is $\lesssim 8\%$.  In
-contrast, the south-easternmost pixel within the \ptfo\ aperture (the
+$10\%$.  The data do not show this pattern.  The data from the pixel directly below
+Star A does not show any sinusoidal variability; the
+peak-to-peak variability seen in that pixel is $\lesssim 8\%$.  In
+contrast, the southeastern-most pixel within the \ptfo\ aperture (the
 pixel furthest from Star A that was used in the optimal aperture)
@@
 \citet{van_eyken_ptf_2012} through high-resolution imaging with the
-NIRC2 camera on Keck II.  They reported 3-$\sigma$ $H$-band magnitude
+NIRC2 camera on the Keck II 10m telescope.  They reported 3-$\sigma$ $H$-band magnitude
 difference limits of 4.3, 6.4, and 8.9 at angular separations of 0.25,
 0.5, and 1.0 arcseconds (87, 175, and 349~AU).  They also detected a
-point source, not present in Gaia DR2, 7.0 magnitudes fainter than the
-target, and 1.8$''$ to the north-east.  Due to the brightness
-difference, this point source\footnote{This point source was claimed
+point source 7.0 magnitudes fainter than the
+target, and 1.8$''$ to the north-east (which is not included in the Gaia DR2 catalog). Due to its relative
+faintness, this object cannot be the source of the shorter and longer-period TESS signals.\footnote{This point source was claimed
 to be a potential planetary-mass object \citep{schmidt_direct_2016}.
 Subsequent analysis of its colors showed that it is a background star
-\citep{lee_evidence_2018}.} cannot be the source of our signals.
+\citep{lee_evidence_2018}.}
 
@@
     {\it Top:} Hertzsprung-Russell diagram of \ptfo\ and late-type
-    members of 25$\,$Ori-1. Members of the 25$\,$Ori-1 group (black
-    circles) were identified by \citet{kounkel_apogee2_2018}.  The
-    gray circles are stars in the ``neighborhood'', i.e., non-member
-    stars within 5 standard deviations of the mean 25$\,$Ori-1 right
-    ascension, declination, and parallax.  It contains members of the
+    members of 25$\,$Ori-1. Black circles are members of the 25$\,$Ori-1 group
+    identified by \citet{kounkel_apogee2_2018}.  
+    Gray circles are stars in the ``neighborhood'', i.e., non-member
+    stars for which the right
+    ascension, declination, and parallax are within 5 standard deviations of the mean values for 25$\,$Ori-1.
+    The neighborhood contains members of the
     Orion complex with its full spread of ages, in addition to field
-    interlopers.  $G$ denotes Gaia broadband magnitudes, $Bp$ Gaia
+    interlopers.  $G$ denotes the Gaia broadband magnitude, $Bp$ Gaia
     blue, $Rp$ Gaia red, and $\omega_{\rm as}$ the parallax in
-    arcseconds.  The $x$-axis limits are set to show only K and M
-    dwarfs, to accentuate \ptfo's separation from the single-star
+    arcseconds.  The $x$-axis limits are chosen to display only the K and M
+    dwarfs, accentuating \ptfo's separation from the single-star
     sequence.  {\it Bottom:} Astrometric goodness-of-fit versus $Rp$
     magnitude for 25$\,$Ori-1 members.  The single-source astrometric
-    model for \ptfo\ provides a poor fit, which could be due to
-    stellar variability or binarity.  Cluster members that are at
+    model for \ptfo\ provides a poor fit to the data, which could be due to
+    either stellar variability or binarity.  But since cluster members that are at
     least as variable as \ptfo\ show lower astrometric excesses (black
-    squares), suggesting binarity as the root cause.
+    squares), binarity is the likely reason.
 		\label{fig:gaia}
@@
 We also used the Gaia data to see if the observed luminosity of \ptfo\
-is too bright to be a single star, i.e., if it is a ``photometric
+is too high to be from a single star, i.e., if the object is a ``photometric
 binary.'' To assemble a set of stars coeval with \ptfo, we used the
@@
 discussed in Section~\ref{subsec:hierarchical}.
-
 To define a set of non-member stars that nonetheless are subject to
@@
 as the group of at most $10^4$ randomly selected non-member stars
-within 5 standard deviations of the mean 25$\,$Ori-1 right ascension,
-declination, and parallax.  We queried Gaia DR2 for these stars using
+within 5 standard deviations of the mean values of the right ascension,
+declination, and parallax of 25$\,$Ori-1.  We queried Gaia DR2 for these stars using
 \texttt{astroquery} \citep{astroquery_2018}.  This yielded 1{,}819
@@
 defined over-density in each dimension: the cluster was confirmed to
-exist, and to be distinct from the neighborhood. \ptfo\ was also
-within the cluster in each of these projected dimensions.
+exist, and to be distinct from the neighborhood.  The second point we noted is that \ptfo\ belongs
+to the cluster, based on its properties in each of these dimensions.
 
@@
 in the cluster.  It also seems to be part of a ``photometric binary''
-track that runs parallel to the main track.
+track that runs above and parallel to the main track.
 
@@
 
-A separate possible line of evidence for binarity is the Gaia DR2
+A separate line of evidence for binarity is the Gaia DR2
 astrometry.  As noted in Section~\ref{sec:observations}, the Gaia DR2
@@
 e.g.},][]{rizzuto_ZEIT8_2018,belokurov_unresolved_2020}.  If
-photometric variability were the cause, we would expect comparably
-faint stars in the same kinematic group of Orion to show similar
-astrometric excesses, as the majority of young stars are highly
+photometric variability were the cause, we would expect stars of similar
+brightness in the same kinematic group of Orion to show similar
+astrometric excesses, because the majority of young stars are highly
 variable.
@@
 
-We will have to wait for the full release of the nominal Gaia mission
-to determine definitively whether the astrometric excess is caused by
+We will have to wait for the next data release of the Gaia mission
+for a more definitive determination of whether the astrometric excess is caused by
 stellar binarity or photometric variability.  Nonetheless the fact
@@
 
-Long-baseline radial velocity (RV) measurements could also reveal the
+Radial velocity (RV) measurements over sufficiently long timescales could also reveal the
 presence of multiple stars in this system.  Unfortunately, the
-available RV data for \ptfo\ is rather sparse, presumably due to the
-stellar faintness and equatorial velocity.  The RV datasets with the
-longest baselines we could find in the literature were those reported
+available RV data for \ptfo\ is sparse, presumably due to the difficulties
+of performing RV observations of such a faint and rapidly rotating
+star.  The RV datasets with the
+longest time baselines we could find in the literature were those reported
 by \citet{van_eyken_ptf_2012}.  These included 5 Keck/HIRES
@@
 measurements acquired over 10 days in February 2011.  The
-root-mean-squared RV over each 10 day span was $\approx$$2\,{\rm
+root-mean-squared RV over each 10-day span was $\approx$$2\,{\rm
 km}\,{\rm s}^{-1}$, consistent with the measurement precision.
@@
 \citet{yu_tests_2015} acquired 22 further Keck/HIRES spectra over one
-night in December 2013, those points were not reduced to velocities,
-and it is not clear whether the appropriate calibration data exist.
-Further Keck/HIRES measurements of \ptfo, if appropriately calibrated,
+night in December 2013, those points were not reduced to velocities
+Further radial-velocity observations
 could potentially confirm or refute the presence of binary companions.
@@
 variability, \citet{van_eyken_ptf_2012} saw an alias of the
-longer-period signal ({\it e.g.}, their Figure~7), and identified it
-as a periodogram peak at $0.9985 \pm 0.0061\,$days. They ascribed it
+longer-period signal ({\it e.g.}, their Figure~7), in the form of a
+peak in the periodogram at $0.9985 \pm 0.0061\,$days. They ascribed it
 to their observing cadence, because of its close correspondence to the
@@
 We are not the first to reach the conclusion that the long period
-sinusoidal modulation is astrophysical.  A study by
+sinusoidal modulation is astrophysical. 
 \citet{koen_multicolour_2015} identified the same modes and aliases as
-\citet{van_eyken_ptf_2012}, but argued that the signal was
-astrophysical (however they were still unsure of the exact period).
+\citet{van_eyken_ptf_2012}, and argued that the signal was
+astrophysical, even if the exact period was still unclear.
 Using photometry from the YETI global telescope network,
-\citet{raetz_yeti_2016} eventually came to the conclusion that the
+\citet{raetz_yeti_2016} came to the conclusion that the
 $0.50\,{\rm d}$ signal was indeed from stellar rotation.  The TESS
@@
     observed.  Horizontal dashed lines are drawn at $\pm (P_{\rm \ell}
-    - P_{\rm s})/P_{\rm s}$, highlighting either a numerical
-    coincidence or an observational bias.  The orbital phase observed
+    - P_{\rm s})/P_{\rm s}$, highlighting a possible numerical
+    coincidence.  The orbital phase observed
     by TESS (lower-right) is consistent with that of
@@
 The TESS light curve shows a dip that lasts about 45 minutes, and
-seems to recur every 10.76 hours
+recurs every 10.76 hours
 (Figures~\ref{fig:splitsignal},~\ref{fig:splitsignalii},~\ref{fig:phasefold}).
@@
 
-One particularly interesting feature of the dip is its epoch.  The
-dips do not occur with strict periodicity \citep{yu_tests_2015}.  In
+An interesting feature of the sequence of dips is that the phase of the dips has been observed to change with time \citep{yu_tests_2015}.  In
 fact, \citet{tanimoto_evidence_2020} provided stark evidence for
-different behavior altogether: over a time-span of years, the dip
-``split'' into distinct groups at particular repeating phases.  See
-for instance their Figures~2 through~4.  Fitting a decade of
-observations, they provided the following linear ephemeris, which we
-did not find any need to update.
+different behavior altogether: over a timespan of years, the dip
+``split'' into distinct groups at particular phases.  See,
+for instance, their Figures~2 through~4.  Fitting a decade of
+observations, they provided the following constant-period ephemeris, which we
+did not find any need to update:
 \begin{align}
@@
 
-Figure~\ref{fig:o_minus_c} shows the observed mid-transit times of
-dips minus the times predicted from Equation~\ref{eq:ephem}.  The
-phase of the dip we detected in the TESS data (yellow star) agrees
+Figure~\ref{fig:o_minus_c} shows the differences bewteen the observed ``mid-transit'' times of
+the dips and the times calculated using Equation~\ref{eq:ephem}.  The
+phase of the dips seen by TESS (yellow star) agrees
 with the independent December 2018 measurements by
@@
 (i)}  multiple dips per cycle, and {\it (ii)} a set of dips
-numerically coincident with phase $(P_{\rm \ell} - P_{\rm s}) / P_{\rm
+at a phase that is numerically coincident with $(P_{\rm \ell} - P_{\rm s}) / P_{\rm
 s}$.  The observation of multiple dips per cycle in 2015 was seen
@@
 of multiple dips per cycle in December 2018 at phase 0 and $-0{.}47$
-seems less plausible---the phase $\text{-}0{.}47$ dips are strongly
-detected, while the suggested phase 0 dip is not clearly present in
-the data.
+seems less plausible. The dips at phase $\text{-}0{.}47$ are strongly
+detected, while the suggested dip at phase 0 is not clearly deteceted.
 
@@
 of long to short periods is roughly 10:9.  It is not clear that this
-would obviously translate into an observational bias unless by some
-fluke three season's worth of observations managed to only observe
+would obviously translate into an observational bias unless, by some
+fluke, three season's worth of observations managed to only observe
 every ninth dip.  This is of course not the case, and we therefore
@@
 % energy to ponder it now, but it deserves pondering.
-
-\subsection{Short Period Out-of-Dip Modulation}
+% JNW: Could this mean that both stars are dippers? A change in period
+% is degenerate with a change in period, over a short interval, right?
+% so maybe one set of dips belongs to one star, the other set to the 2nd star?
+% Features with a different period would appear to lie along a line
+% of constant slope in this diagram...what would the slope be, in this case,
+% if the feature period were P_long?
+
+\subsection{Short Period Modulation Outside of Dips}
 
@@
 resembles a slightly asymmetric sinusoid (Figure~\ref{fig:phasefold}).
-Non-zero contributions in both the first and second harmonics are
-detected (Table~2).  The first sine and cosine harmonic both have
+The best model has non-zero amplitudes for both the first and second harmonics (Table~2).
+The first sine and cosine harmonic both have
 amplitudes of roughly $0.90\pm0.04\%$.  The second sine harmonic has
 amplitude $0.16 \pm 0.04\%$, so is non-zero at a significance of
-$\approx$4$\sigma$.  The second cosine harmonic has negative amplitude
-$0.55 \pm 0.03\%$.  In our sign convention, the fact that it is
-negative means that this component peaks at phase 0.25 and 0.75.
+$\approx$4$\sigma$.  The second cosine harmonic has an amplitude of
+$-0.55 \pm 0.03\%$.  In our sign convention, the fact that it is
+negative means that this component peaks at phase 0.25 and 0.75,
+i.e., the quadratures of the orbit.
 
@@
 distort the host star, and cause ellipsoidal photometric modulations
-that also peak at quadrature \citep[see][]{shporer_astrophysics_2017}.
+that peak at the quadratures \citep[see][]{shporer_astrophysics_2017}.
 Interpreting the second cosine harmonic as planet-induced tidal
@@
   \caption{ {\bf \ptfo\ and its brethren.}
-    Five transient and persistent flux dip stars selected based on
-    their visual similarity to the short-period signal in \ptfo\ are
-    as shown.  They include EPIC 204143627, EPIC 204270520, EPIC
+    Shown are the light cuerves of five transient and persistent flux dip stars, selected based on
+    their similarity to the short-period signal of \ptfo.  The stars are EPIC 204143627, EPIC 204270520, EPIC
     204321142, EPIC 246938594, and EPIC 205483258 (RIK-210).  RIK-210
@@
     displayed are either in Taurus or Upper~Sco, and meet the
-    characteristics of Section~\ref{subsec:dipstars}.  We found these
-    objects through studies by \citet{stauffer_orbiting_2017},
+    characteristics of Section~\ref{subsec:dipstars}.  These objects
+    were originally reported by \citet{stauffer_orbiting_2017},
     \citet{david_transient_2017}, and \citet{rebull_usco_2018}.
@@
 
-When physical explanations are not forthcoming, taxonomy is a useful
-exercise.  By searching the literature, we have found about a dozen
+When physical explanations are not forthcoming, we often resort to taxonomy.
+By searching the literature, we have found about a dozen
 light curves with similar morphologies to \ptfo, drawn from surveys of
@@
     \citealt{rebull_usco_2018},~Figure~20).
-  \item The age is typically\footnote{
+  \item The age is typically $\lesssim$ 100$\,$Myr.\footnote{
     At present, the oldest observed ``scallops'' are in the Pleaides
@@
     meet the ``persistent dip'' classification.  If so, it is also the
-    oldest known.
-    } $\lesssim$ 100$\,$Myr.
+    oldest known.} 
   \item The light curves show shallow, angular dips, usually superposed
@@
     as stellar rotation.
-  \item The rotation is rapid---usually between 0.5 and 2.0 days.
-  \item There is rarely any detectable infrared excess in WISE data
+  \item The rotation is rapid, with a period that is usually between 0.5 and 2.0 days.
+  \item There is rarely any infrared excess that is detectable in the WISE data
   (never any W4 detection; only a few W3 detections).
+  % JNW: Spell out acronym and give reference?
   \item They sometimes show multiple dips per cycle.
@@
   \item The dip depths can change after flares.
-  \item They are rare at a population level: $\lesssim 1\%$ of relevant stars
-    \citep{rebull_usco_2018}.
+  \item They are rare at a population level, with an occurrence rate of $\lesssim 1\%$ of
+  the relevant stars \citep{rebull_usco_2018}.
+% JNW: what is 'relevant stars'? Can this be replaced by 'young M2-M5 stars'?
 \end{enumerate}
@@
 TESS light curve was needed to resolve the different rotation signals.
+% JNW: Despite the preceding sentence, I don't understand why this connection was impossible before TESS.
 
@@
 photometric periods \citep[][Figure~26]{stauffer_orbiting_2017}.  In
-the usual idealized limits, the transit duration $T$ of a point source
+contrast, the transit duration $T$ of a small obstructing object
 across the stellar disk scales as $T \propto R_\star
@@
 consistent with point sources, at longer periods of 1 to 5 days the
-dip durations become many hours, far too long to be caused by
+dip durations become many hours, which is too long to be caused by
 planetary transits.
 
-Second, between 40-50\% of the transient flux dip stars discovered in
+Second, approximately 40-50\% of the transient flux dip stars discovered in
 $\rho$~Oph and Upper~Sco show two Lomb-Scargle periods, and so are
@@
 fractions up to twice as high in dispersed clusters such as Upper~Sco
-and Taurus \citep{kraus_mapping_2008,kraus_mapping_2011}.  A detailed
-high-resolution imaging survey would be necessary to determine whether
-the transient flux dip stars truly have any distinct population-level
+and Taurus \citep{kraus_mapping_2008,kraus_mapping_2011}.  A 
+high-resolution imaging survey would be interesting, to determine whether
+the transient flux dip stars truly have distinct population-level
 binarity properties relative to other young low-mass stars.
@@
 The evidence for binarity in \ptfo\ is as follows.  First, the star is 
-photometrically twice as bright as stars of the same color in its
+twice as bright as stars of the same color in its
 kinematic group (Figure~\ref{fig:gaia}).  Second, it shows two
@@
 \cite{hernandez_spitzer_2006,hernandez_spitzer_ob1_2007,hernandez_spitzer_sig_2007,hernandez_spitzer_2009}
-show that it is.} around both stars in \ptfo. The stars are therefore
+show that it is.}
+% JNW: I don't understand why we can be so definitive...
+% an infrared excess is not either 'present' or 'absent'. The upper limit in Jy
+% corresponds to a constraint on the mass/radius/etc of any dusty disk.
+around both stars in \ptfo. The stars are therefore
 no longer magnetically locked to their disks.  This is consistent with
@@
 evolution.
-
-The main physical question is what causes the transient dips. This is
-an unsolved problem not only for \ptfo\ but also for an emerging class
-of similar young rapidly rotating M-dwarfs.  Many possible
-explanations have been detailed by \citet{rebull_rotation_2016},
+% JNW: I don't understand the last sentence.
+
+The key question is what causes the transient dips. This is
+an unsolved problem not only for \ptfo\ but also for the emerging class
+of similar young and rapidly rotating M-dwarfs.  Many possible
+explanations were discussed by \citet{rebull_rotation_2016},
 \citet{david_transient_2017}, \citet{stauffer_orbiting_2017}, and
-\citet{zhan_complex_2019}.  Their disfavored explanations include that
+\citet{zhan_complex_2019}.  Among the {\it disfavored} explanations are that
 the dips are caused by
@@
 {\it (iii)} eclipses of prominences;
+% JNW: why is this one ruled out? Sounds similar to the accretion-hotspot hypothesis.
 {\it (iv)} high-latitude accretion hotspots;
@@
 systems.
+% JNW Could the material get magnetically locked at the corotation radius?
 
@@
 \acknowledgements
-When this manuscript was at an advanced stage, we learned of the
-complementary work by \citet{koen_2020}, which was in press at the
+When this manuscript was at an advanced stage, we received notice of a
+paper by \citet{koen_2020} that was in press at the
 {\it Monthly Notices} before submission of our manuscript.  Our
-studies independently reached the key conclusions that the TESS
-light curve shows two periodic signals, and that the properties of
+studies independently reached the same conclusions: the TESS
+light curve shows two periodic signals, and the properties of
 \ptfo\ are consistent with the emerging class of transient and
@@
 provide independent support for the conclusion that \ptfo\ is a
-binary, and emphasized the agreement between the
+binary. We also note the agreement between the
 TESS dip ephemeris and that from \citet{tanimoto_evidence_2020}.

### 2020-05-14 paper/ms.tex
fixes to 21055a6

@@
 A relevant fact is that between 0.1\% and 1\% of rapidly rotating
-low-mass stars in $\mathcal{O}$\,(10)$\,$Myr old associations show
+low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations show
 short-duration dips as part of their overall periodic variability
@@
 where `$\subset$' means `is a proper subset of'.  25$\,$Ori-1 is the
-largest subgroup of 25$\,$Ori, with 149 identified members.  Based on
-the color-magnitude diagram (CMD), the mean age was found to be
-6.9$\,$Myr, while the HR diagram led to an estimated age of 8.5$\,$Myr
-% JNW: What is the difference between HR and CMD?
+largest subgroup of 25$\,$Ori, with 149 identified members.  The mean
+age of the 25$\,$Ori-1 subgroup, determined by fitting isochrones to
+group members that had effective temperatures from APOGEE and Gaia
+parallaxes, was determined to be $8.5\pm1.2\,{\rm Myr}$
 \citep[see][Section 2.3]{kounkel_apogee2_2018}.
@@
 the Orion complex near the Be star 25$\,$Ori. These groups received
-higher numbers, {\it e.g.}, 25$\,$Ori-2 (${\rm Age}_{\rm CMD}
-=15.1\,$Myr; ${\rm Age}_{\rm CMD} =12.9\,$Myr; see also
-\citealt{briceno_cidaII_2019}).
+sequential identifiers, {\it e.g.}, 25$\,$Ori-2 (${\rm Age}
+=12.9\pm2.8\,$Myr; see also \citealt{briceno_cidaII_2019}).
 
@@
 
-We will have to wait for the next data release of the Gaia mission
-for a more definitive determination of whether the astrometric excess is caused by
-stellar binarity or photometric variability.  Nonetheless the fact
-that comparably variable stars do not show comparably large
+We will have to wait for the next data release of the Gaia mission for
+a more definitive determination of whether the astrometric excess is
+caused by stellar binarity or photometric variability.  Nonetheless
+the fact that comparably variable stars do not show comparably large
 astrometric excesses suggests that stellar binarity is indeed the root
@@
 resembles a slightly asymmetric sinusoid (Figure~\ref{fig:phasefold}).
-The best model has non-zero amplitudes for both the first and second harmonics (Table~2).
-The first sine and cosine harmonic both have
-amplitudes of roughly $0.90\pm0.04\%$.  The second sine harmonic has
-amplitude $0.16 \pm 0.04\%$, so is non-zero at a significance of
-$\approx$4$\sigma$.  The second cosine harmonic has an amplitude of
-$-0.55 \pm 0.03\%$.  In our sign convention, the fact that it is
-negative means that this component peaks at phase 0.25 and 0.75,
-i.e., the quadratures of the orbit.
+The best model has non-zero amplitudes for both the first and second
+harmonics (Table~2).  The third harmonic is formally present with
+marginal ($\approx$2$\sigma$) significance.  The first sine and cosine
+harmonic both have amplitudes of roughly $0.90\pm0.04\%$.  The second
+sine harmonic has amplitude $0.16 \pm 0.04\%$, so is non-zero at a
+significance of $\approx$4$\sigma$.  The second cosine harmonic has an
+amplitude of $-0.55 \pm 0.03\%$.  In our sign convention, the fact
+that it is negative means that this component peaks at phase 0.25 and
+0.75, i.e., the quadratures of the orbit.
 
@@
   \caption{ {\bf \ptfo\ and its brethren.}
-    Shown are the light cuerves of five transient and persistent flux dip stars, selected based on
+    Shown are the light curves of five transient and persistent flux dip stars, selected based on
     their similarity to the short-period signal of \ptfo.  The stars are EPIC 204143627, EPIC 204270520, EPIC

### 2020-05-14 paper/ms.tex
ditto

@@
 One was CVSO~35, which has a TESS light curve that varies by 2
-magnitudes, and shows a strong IR excess and a 10$\mu$m silicate
+magnitudes.
+CVSO~35 also shows a strong Wide-field Infrared Survey Exoplorer
+(WISE, \citealt{wright_WISE_2010}) IR excess as well as a 10$\mu$m silicate
 emission feature \citep{mauco_herschel_2018}.  The other is GAIA DR2
@@
 % if the feature period were P_long?
+% LGB: Seems unlikely, because we're seeing a something with constant
+% 10.74 hour periodicity, but the phase is offset by 1/9th.
+% If both stars were dippers, we would be seeing a 11.98 hour
+% periodicity instead.
 
@@
   (never any W4 detection; only a few W3 detections).
-  % JNW: Spell out acronym and give reference?
   \item They sometimes show multiple dips per cycle.
@@
   \item They are rare at a population level, with an occurrence rate of $\lesssim 1\%$ of
-  the relevant stars \citep{rebull_usco_2018}.
-% JNW: what is 'relevant stars'? Can this be replaced by 'young M2-M5 stars'?
+  young M2-M5 stars \citep{rebull_usco_2018}.
 \end{enumerate}
 The 10.76$\,$hr signal in \ptfo\ meets all of these criteria.  This is
-the first connection of \ptfo\ with this class of objects because the
-TESS light curve was needed to resolve the different rotation signals.
-% JNW: Despite the preceding sentence, I don't understand why this connection was impossible before TESS.
+the first connection of \ptfo\ with this class of objects, 
+likely because the TESS data enabled us to put the dips in 
+the context of the asymmetric out-of-dip modulation.
 
@@
 % corresponds to a constraint on the mass/radius/etc of any dusty disk.
-around both stars in \ptfo. The stars are therefore
-no longer magnetically locked to their disks.  This is consistent with
-the $\approx$half-day periodicities of both rotation signals: young
-disked M dwarfs typically rotate with periods of two days or more due
-to magnetic locking \citep[{\it e.g.},][]{rebull_rotation_2020}.  If
-the two stars are within $\approx$50$\,$AU of each other, as required
-by the NIRC2 adaptive optics imaging, then it would also be expected
-that the stars would truncate the outer edges of their respective
+around both stars in \ptfo. The stars are therefore no longer
+magnetically locked to their disks.  This is consistent with the
+$\approx$half-day periodicities of both rotation signals: young disked
+M dwarfs typically rotate with periods of two days or more due to
+magnetic locking \citep[{\it e.g.},][]{rebull_rotation_2020}.  If the
+two stars are within $\approx$50$\,$AU of each other, as required by
+the NIRC2 adaptive optics imaging, then it would also be expected that
+the stars would have truncated the outer edges of their respective
 disks, in a manner seen at the population level in exoplanetary
-systems \citep{kraus_impact_2016,moe_impact_2019}.  This hard outer
-boundary condition could propagate to the inner disk and affect its
-evolution.
-% JNW: I don't understand the last sentence.
-
-The key question is what causes the transient dips. This is
-an unsolved problem not only for \ptfo\ but also for the emerging class
-of similar young and rapidly rotating M-dwarfs.  Many possible
+systems \citep{kraus_impact_2016,moe_impact_2019}.
+
+The key remaining question is what causes the transient dips. This is
+an unsolved problem not only for \ptfo\ but also for the emerging
+class of similar young and rapidly rotating M-dwarfs.  Many possible
 explanations were discussed by \citet{rebull_rotation_2016},
 \citet{david_transient_2017}, \citet{stauffer_orbiting_2017}, and
-\citet{zhan_complex_2019}.  Among the {\it disfavored} explanations are that
-the dips are caused by
+\citet{zhan_complex_2019}.  Among the {\it disfavored} explanations
+are that the dips are caused by
 {\it (i)} eclipsing binaries;

### 2020-05-14 paper/ms.tex
fix tex column width

@@
   planetary transits.  Although strong evidence against the planet
-  hypothesis has been presented, the possibility remains
-  debated in the literature.  To obtain further clues, we inspected
-  data from the NASA {\it Transiting Exoplanet Survey Satellite}
-  (TESS) and the ESA Gaia mission.  The Gaia data suggest that
-  PTFO$\,$8-8695 is a binary: the photometric data show
-  it to be overluminous with respect to members of
-  its kinematic group, and the astrometric data are inconsistent
-  with a single star.  The TESS light curve shows two different
-  photometric periods. The variability is dominated by a sinusoidal
-  signal with a period of 11.98$\,$hr, presumably caused by stellar
-  rotation.  Also present is a 10.76$\,$hr signal consisting of a
-  not-quite sinusoid interrupted by hour-long dips, the type of signal
-  previously intepreted as planetary transits.  The phase of the dips
-  is nearly 180$^\circ$ away from the phase of the originally reported
-  dips. As noted previously, this makes them difficult to explain as
-  planetary transits.  Instead, we believe that PTFO$\,$8-8695 is a
-  pair of young and rapidly rotating M dwarfs, one of which shows the
-  same ``transient-dipper'' behavior that has been seen in at least 5
-  other cases.  The origin of these transient dips is still unknown
-  but likely involves circumstellar material.
+  hypothesis has been presented, the possibility remains debated in
+  the literature.  To obtain further clues, we inspected data from the
+  NASA {\it Transiting Exoplanet Survey Satellite} (TESS) and the ESA
+  Gaia mission.  The Gaia data suggest that PTFO$\,$8-8695 is a
+  binary: the photometric data show it to be overluminous with respect
+  to members of its kinematic group, and the astrometric data are
+  inconsistent with a single star.  The TESS light curve shows two
+  different photometric periods. The variability is dominated by a
+  sinusoidal signal with a period of 11.98$\,$hr, presumably caused by
+  stellar rotation.  Also present is a 10.76$\,$hr signal consisting
+  of a not-quite sinusoid interrupted by hour-long dips, the type of
+  signal previously intepreted as planetary transits.  The phase of
+  the dips is nearly 180$^\circ$ away from the phase of the originally
+  reported dips. As noted previously, this makes them difficult to
+  explain as planetary transits.  Instead, we believe that
+  PTFO$\,$8-8695 is a pair of young and rapidly rotating M dwarfs, one
+  of which shows the same ``transient-dipper'' behavior that has been
+  seen in at least 5 other cases.  The origin of these transient dips
+  is still unknown but likely involves circumstellar material.
 \end{abstract}
@@
 Transits by dust clumps or other dusty features are questionable
-because \ptfo\ does not have a detectable infrared (IR) excess associated
-with the presence of warm dust
-\citep[{\it e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude
-starspots, hot or cold, would struggle to produce photometric features as
-short as some of the observed dips.
+because \ptfo\ does not have a detectable infrared (IR) excess
+associated with the presence of warm dust \citep[{\it
+e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude starspots, hot or
+cold, would struggle to produce photometric features as short as some
+of the observed dips.
 
@@
 Section~\ref{sec:tess}.  The Gaia data, analyzed in
-Section~\ref{sec:gaia}, show that \ptfo\ is too bright to be a single star and
-also suggest it is an astrometric binary.  We discuss the pieces of
-the puzzle in Section~\ref{sec:discussion}, and summarize the
-situation in Section~\ref{sec:conclusions}.  In a postscript, we
+Section~\ref{sec:gaia}, show that \ptfo\ is too bright to be a single
+star and also suggest it is an astrometric binary.  We discuss the
+pieces of the puzzle in Section~\ref{sec:discussion}, and summarize
+the situation in Section~\ref{sec:conclusions}.  In a postscript, we
 comment on a recent study by \citet{koen_2020} which reached similar
@@
     {\bf TESS light curve of \ptfo\ (Sector 6, Orbit 19).} {\it Top}:
-    The original (\texttt{PDCSAP} median-subtracted) relative flux.  The
-    beat period of 4.48 days is visible by eye.  The blue curve is a
-    model including 2 harmonics at the longer period $P_{\rm \ell}$,
+    The original (\texttt{PDCSAP} median-subtracted) relative flux.
+    The beat period of 4.48 days is visible by eye.  The blue curve is
+    a model including 2 harmonics at the longer period $P_{\rm \ell}$,
     plus 3 harmonics and a transit at the shorter period $P_{\rm s}$.
@@
 January 6, 2019, during the sixth sector of science operations
-\citep{ricker_transiting_2015}.  The star is designated
-TIC\,264461976 in the TESS Input Catalog
-\citep{stassun_TIC_2018,stassun_TIC8_2019}.  The pixel data for an
-$11\times11$ array surrounding \ptfo\ were averaged into 2-minute
-stacks by the onboard computer.  Each 2048$\times$2048 image from the
-CCD was also averaged into 30-minute stacks, and saved as a ``full
-frame image'' (FFI).
+\citep{ricker_transiting_2015}.  The star is designated TIC\,264461976
+in the TESS Input Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.
+The pixel data for an $11\times11$ array surrounding \ptfo\ were
+averaged into 2-minute stacks by the onboard computer.  Each
+2048$\times$2048 image from the CCD was also averaged into 30-minute
+stacks, and saved as a ``full frame image'' (FFI).
 
 The 2-minute stacks for \ptfo\ were reduced to light curves by the
-Science Processing Operations Center (SPOC) at the NASA
-Ames Research Center~\citep{jenkins_tess_2016}.  We mainly used the Presearch Data
+Science Processing Operations Center (SPOC) at the NASA Ames Research
+Center~\citep{jenkins_tess_2016}.  We mainly used the Presearch Data
 Conditioning Simple Aperture Photometry (PDCSAP) light curve.  The PDC
-light curve is based on pixels chosen to maximize the SNR of the
-total flux of the target \citep{bryson_2020_target_aperture}.
+light curve is based on pixels chosen to maximize the SNR of the total
+flux of the target \citep{bryson_2020_target_aperture}.
 Non-astrophysical variability was removed by fitting out trends common
@@
 often systematic effects in the photometry during those times.  Both
-the CDIPS and PDC light curves showed a discontinuous jump in
-the last few days of orbit 20, which seemed likely to be an
-instrumental systematic effect, and led us to mask out the
-data with timestamps ranging from BJD\,2458488.3 until the end of the
-orbit.  The PDC light curve initially had 15{,}678 points.  The
-quality-flag cut removed 854 points; masking the orbit edges removed
-an additional 716 points; and removing the data from the final few
-days of orbit 20 removed an additional 1{,}079 points.  After cleaning,
-83\% of the initial flux measurements remained.
+the CDIPS and PDC light curves showed a discontinuous jump in the last
+few days of orbit 20, which seemed likely to be an instrumental
+systematic effect, and led us to mask out the data with timestamps
+ranging from BJD\,2458488.3 until the end of the orbit.  The PDC light
+curve initially had 15{,}678 points.  The quality-flag cut removed 854
+points; masking the orbit edges removed an additional 716 points; and
+removing the data from the final few days of orbit 20 removed an
+additional 1{,}079 points.  After cleaning, 83\% of the initial flux
+measurements remained.
 
@@
 used to determine positions, proper motions, and parallaxes of the
-brighest 1{.}3 billion stars
-\citep{lindegren_gaiasoln_2018}.  For \ptfo, there were 121 ``good''
-observations, {\it i.e.}, observations that were not strongly
-down-weighted in the astrometric solution.  \ptfo\ was assigned the
-Gaia DR2 identifier 3222255959210123904.  Its brightness
-was measured using selected bands ($G$, $Rp$, and $Bp$) of the Gaia
-Radial Velocity Spectrometer
+brighest 1{.}3 billion stars \citep{lindegren_gaiasoln_2018}.  For
+\ptfo, there were 121 ``good'' observations, {\it i.e.}, observations
+that were not strongly down-weighted in the astrometric solution.
+\ptfo\ was assigned the Gaia DR2 identifier 3222255959210123904.  Its
+brightness was measured using selected bands ($G$, $Rp$, and $Bp$) of
+the Gaia Radial Velocity Spectrometer
 \citep{cropper_gaia_2018,evans_gaia_2018}.  We accessed the pipeline
@@
 
-\ptfo\ was initially designated CVSO\,30 and identified as a member of the Orion$\,$OB1a
-sub-association by \citet{briceno_cida_2005},
-based on photometry and spectroscopy.  Later work by
-\citet{briceno_25_2007} clarified that \ptfo\ is in a kinematically
-distinct subgroup of Orion~OB1a, named the ``25$\,$Ori'' group after
-its brightest member. They reported that the 25$\,$Ori group has an
-isochrone age of 7--10$\,$Myr, and a smaller fraction of stars with
-disks than younger nearby sub-associations
-\citep{hernandez_spitzer_ob1_2007}.
+\ptfo\ was initially designated CVSO\,30 and identified as a member of
+the Orion$\,$OB1a sub-association by \citet{briceno_cida_2005}, based
+on photometry and spectroscopy.  Later work by \citet{briceno_25_2007}
+clarified that \ptfo\ is in a kinematically distinct subgroup of
+Orion~OB1a, named the ``25$\,$Ori'' group after its brightest member.
+They reported that the 25$\,$Ori group has an isochrone age of
+7--10$\,$Myr, and a smaller fraction of stars with disks than younger
+nearby sub-associations \citep{hernandez_spitzer_ob1_2007}.
 
@@
 
-These details concerning the group membership for one object may
-seem excessive to those accustomed to the simple distinction between
+These details concerning the group membership for one object may seem
+excessive to those accustomed to the simple distinction between
 ``young cluster members'' and ``old field stars''.  Although all
@@
     {\it Top}: The longer-period signal.  {\it Bottom}: The
-    shorter-period signal. The phase is defined such that the dip occurs at zero phase.  Gray
-    points are the \texttt{PDCSAP} data binned to 10-minute cadence.  Black
-    points are binned to 100 points per period.  The model (blue line)
-    includes 2 harmonics at the longer period, plus 3 harmonics and a
-    transit at the shorter period.
+    shorter-period signal. The phase is defined such that the dip
+    occurs at zero phase.  Gray points are the \texttt{PDCSAP} data
+    binned to 10-minute cadence.  Black points are binned to 100
+    points per period.  The model (blue line) includes 2 harmonics at
+    the longer period, plus 3 harmonics and a transit at the shorter
+    period.
 		\label{fig:phasefold}
@@
 The tallest peak occurs at 0.499\,d (11.98$\,$hr) and a second strong
-peak occurs at 0.448\,d (10.76$\,$hr). We will refer to these two periods as the ``longer
-period'' $P_{\rm \ell}$ and the ``shorter period'' $P_{\rm s}$.  Lower-power
-harmonics of both signals are also present.
-
-% 0.14 = A1+A2
-% 0.06 = A1-A2
-% 2A1 = 0.20
-% -> A1 = 0.1
-% 2A2 = 0.08
-% -> A2 = 0.04
+peak occurs at 0.448\,d (10.76$\,$hr). We will refer to these two
+periods as the ``longer period'' $P_{\rm \ell}$ and the ``shorter
+period'' $P_{\rm s}$.  Lower-power harmonics of both signals are also
+present.
 
@@
 the phase-folded light curves of the longer-period signal, we
-subtracted the best-fitting sinusoid with a period equal to $P_{\rm s}$. The
-resulting light curve appears smooth and nearly sinusoidal.  But after
-subtracting the best-fitting sinusoid with a period equal to $P_{\ell}$, visual
-inspection of the phase-folded light curve revealed substructure resembling the ``dips'' seen in previous
-observations. In particular, there was a $\approx$1\% dip lasting
-about an hour.  These initial impressions turned out to be consistent
-with the results of our more complicated analysis, described below.
+subtracted the best-fitting sinusoid with a period equal to $P_{\rm
+s}$. The resulting light curve appears smooth and nearly sinusoidal.
+But after subtracting the best-fitting sinusoid with a period equal to
+$P_{\ell}$, visual inspection of the phase-folded light curve revealed
+substructure resembling the ``dips'' seen in previous observations. In
+particular, there was a $\approx$1\% dip lasting about an hour.  These
+initial impressions turned out to be consistent with the results of
+our more complicated analysis, described below.
 
@@
 
-The free parameters are as follows.  The transit model parameters are the impact
-parameter, the planet-to-star radius ratio, two quadratic limb
-darkening parameters, the planet's orbital period (set equal to
+The free parameters are as follows.  The transit model parameters are
+the impact parameter, the planet-to-star radius ratio, two quadratic
+limb darkening parameters, the planet's orbital period (set equal to
 $P_{\rm s}$) the time of a particular transit, and the mean flux.  We
-sampled the stellar radius and mass from prior probability distributions, implicitly
-defining the stellar density which (together with the orbital period) sets the transit timescale. 
-There are also the parameters defining the Fourier modes.
-As an example, one possible model consists of a transit, $N_{\rm s}=2$
-sines and cosines at the shorter period, plus $N_{\rm \ell}=1$ sine
-and cosine at the longer period.  There are $2N_{\rm
-s}=4$ additional Fourier amplitudes at the shorter period, plus
-$2N_{\rm \ell}=2$ Fourier amplitudes at the longer period, as well as
-$P_\ell$ itself and the relative phase $\phi_\ell$.  The total number
-of parameters is 17 for this case.
+sampled the stellar radius and mass from prior probability
+distributions, implicitly defining the stellar density which (together
+with the orbital period) sets the transit timescale.  There are also
+the parameters defining the Fourier modes.  As an example, one
+possible model consists of a transit, $N_{\rm s}=2$ sines and cosines
+at the shorter period, plus $N_{\rm \ell}=1$ sine and cosine at the
+longer period.  There are $2N_{\rm s}=4$ additional Fourier amplitudes
+at the shorter period, plus $2N_{\rm \ell}=2$ Fourier amplitudes at
+the longer period, as well as $P_\ell$ itself and the relative phase
+$\phi_\ell$.  The total number of parameters is 17 for this case.
 
@@
 To speed up the fitting process, we averaged the 2-minute light curve
-into 10-minute samples.  We correspondingly scaled down the uncertainties in
-the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
-initialized each model with the parameters of the maximum {\it a
-posteriori} (MAP) model.  We then assumed a Gaussian likelihood, and
-sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
-\citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
-diagnostic \citep{gelman_inference_1992}.  We tested our ability to
-successfully recover injected parameters using synthetic data before
-fitting the \ptfo\ light curves.
+into 10-minute samples.  We correspondingly scaled down the
+uncertainties in the flux measurements by a factor of $\sqrt{5}$.
+Before sampling, we initialized each model with the parameters of the
+maximum {\it a posteriori} (MAP) model.  We then assumed a Gaussian
+likelihood, and sampled using \texttt{PyMC3}'s gradient-based
+No-U-Turn Sampler \citep{hoffman_no-u-turn_2014}, and used $\hat{R}$
+as our convergence diagnostic \citep{gelman_inference_1992}.  We
+tested our ability to successfully recover injected parameters using
+synthetic data before fitting the \ptfo\ light curves.
 
@@
 
-The periodogram of the residuals between the data and the preferred model shows a barely significant and
-poorly-resolved peak at $\approx$8 days, consistent with the visual
-impression of some slower trends in the bottom rows of
-Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}.
+The periodogram of the residuals between the data and the preferred
+model shows a barely significant and poorly-resolved peak at
+$\approx$8 days, consistent with the visual impression of some slower
+trends in the bottom rows of Figures~\ref{fig:splitsignal}
+and~\ref{fig:splitsignalii}.
 
@@
 
-The portion of the sky subtended by each TESS pixel is about 21'' on a side. Before making any
-interpretations, we needed to consider whether light from neighboring
-stars could have contributed to the photometric signal we are
-attributing to \ptfo. The scene is shown in Figure~\ref{fig:scene}.
-In the upper panels, the pixels used to measure the background level
-in the SPOC light curve are indicated with `\texttt{X}' hatching, and
-the pixels used in the final light curve aperture are shown with
-`\texttt{/}' hatching.
+The portion of the sky subtended by each TESS pixel is about 21'' on a
+side. Before making any interpretations, we needed to consider whether
+light from neighboring stars could have contributed to the photometric
+signal we are attributing to \ptfo. The scene is shown in
+Figure~\ref{fig:scene}.  In the upper panels, the pixels used to
+measure the background level in the SPOC light curve are indicated
+with `\texttt{X}' hatching, and the pixels used in the final light
+curve aperture are shown with `\texttt{/}' hatching.
 
@@
 cannot contribute more than about 10\% to the total signal.  The only
-other known star that is sufficiently close and bright that its light might
-contaminate the signal from the target star is TIC 264461980, with $T=14.8$.
-This star, we we dub ``Star A'', is 23.6'' northwest of the target. Based
-on the magnitude difference, Star A could contribute flux variations as large
-as 48\% of the flux of the target star.
+other known star that is sufficiently close and bright that its light
+might contaminate the signal from the target star is TIC 264461980,
+with $T=14.8$.  This star, we we dub ``Star A'', is 23.6'' northwest
+of the target. Based on the magnitude difference, Star A could
+contribute flux variations as large as 48\% of the flux of the target
+star.
 
@@
 nearest to Star A to show a sinusoidal signal with amplitude exceeding
-$10\%$.  The data do not show this pattern.  The data from the pixel directly below
-Star A does not show any sinusoidal variability; the
+$10\%$.  The data do not show this pattern.  The data from the pixel
+directly below Star A does not show any sinusoidal variability; the
 peak-to-peak variability seen in that pixel is $\lesssim 8\%$.  In
@@
 \citet{van_eyken_ptf_2012} through high-resolution imaging with the
-NIRC2 camera on the Keck II 10m telescope.  They reported 3-$\sigma$ $H$-band magnitude
-difference limits of 4.3, 6.4, and 8.9 at angular separations of 0.25,
-0.5, and 1.0 arcseconds (87, 175, and 349~AU).  They also detected a
-point source 7.0 magnitudes fainter than the
-target, and 1.8$''$ to the north-east (which is not included in the Gaia DR2 catalog). Due to its relative
-faintness, this object cannot be the source of the shorter and longer-period TESS signals.\footnote{This point source was claimed
-to be a potential planetary-mass object \citep{schmidt_direct_2016}.
-Subsequent analysis of its colors showed that it is a background star
+NIRC2 camera on the Keck II 10m telescope.  They reported 3-$\sigma$
+$H$-band magnitude difference limits of 4.3, 6.4, and 8.9 at angular
+separations of 0.25, 0.5, and 1.0 arcseconds (87, 175, and 349~AU).
+They also detected a point source 7.0 magnitudes fainter than the
+target, and 1.8$''$ to the north-east (which is not included in the
+Gaia DR2 catalog). Due to its relative faintness, this object cannot
+be the source of the shorter and longer-period TESS
+signals.\footnote{This point source was claimed to be a potential
+planetary-mass object \citep{schmidt_direct_2016}.  Subsequent
+analysis of its colors showed that it is a background star
 \citep{lee_evidence_2018}.}
@@
     {\it Top:} Hertzsprung-Russell diagram of \ptfo\ and late-type
-    members of 25$\,$Ori-1. Black circles are members of the 25$\,$Ori-1 group
-    identified by \citet{kounkel_apogee2_2018}.  
+    members of 25$\,$Ori-1. Black circles are members of the
+    25$\,$Ori-1 group identified by \citet{kounkel_apogee2_2018}.
     Gray circles are stars in the ``neighborhood'', i.e., non-member
-    stars for which the right
-    ascension, declination, and parallax are within 5 standard deviations of the mean values for 25$\,$Ori-1.
-    The neighborhood contains members of the
-    Orion complex with its full spread of ages, in addition to field
-    interlopers.  $G$ denotes the Gaia broadband magnitude, $Bp$ Gaia
-    blue, $Rp$ Gaia red, and $\omega_{\rm as}$ the parallax in
-    arcseconds.  The $x$-axis limits are chosen to display only the K and M
-    dwarfs, accentuating \ptfo's separation from the single-star
-    sequence.  {\it Bottom:} Astrometric goodness-of-fit versus $Rp$
-    magnitude for 25$\,$Ori-1 members.  The single-source astrometric
-    model for \ptfo\ provides a poor fit to the data, which could be due to
-    either stellar variability or binarity.  But since cluster members that are at
-    least as variable as \ptfo\ show lower astrometric excesses (black
-    squares), binarity is the likely reason.
+    stars for which the right ascension, declination, and parallax are
+    within 5 standard deviations of the mean values for 25$\,$Ori-1.
+    The neighborhood contains members of the Orion complex with its
+    full spread of ages, in addition to field interlopers.  $G$
+    denotes the Gaia broadband magnitude, $Bp$ Gaia blue, $Rp$ Gaia
+    red, and $\omega_{\rm as}$ the parallax in arcseconds.  The
+    $x$-axis limits are chosen to display only the K and M dwarfs,
+    accentuating \ptfo's separation from the single-star sequence.
+    {\it Bottom:} Astrometric goodness-of-fit versus $Rp$ magnitude
+    for 25$\,$Ori-1 members.  The single-source astrometric model for
+    \ptfo\ provides a poor fit to the data, which could be due to
+    either stellar variability or binarity.  But since cluster members
+    that are at least as variable as \ptfo\ show lower astrometric
+    excesses (black squares), binarity is the likely reason.
 		\label{fig:gaia}
@@
 We also used the Gaia data to see if the observed luminosity of \ptfo\
-is too high to be from a single star, i.e., if the object is a ``photometric
-binary.'' To assemble a set of stars coeval with \ptfo, we used the
-25$\,$Ori-1 members identified by \citet{kounkel_apogee2_2018}, and
-discussed in Section~\ref{subsec:hierarchical}.
-To define a set of non-member stars that nonetheless are subject to
-similar selection criteria, we defined the reference ``neighborhood''
-as the group of at most $10^4$ randomly selected non-member stars
-within 5 standard deviations of the mean values of the right ascension,
-declination, and parallax of 25$\,$Ori-1.  We queried Gaia DR2 for these stars using
+is too high to be from a single star, i.e., if the object is a
+``photometric binary.'' To assemble a set of stars coeval with \ptfo,
+we used the 25$\,$Ori-1 members identified by
+\citet{kounkel_apogee2_2018}, and discussed in
+Section~\ref{subsec:hierarchical}.  To define a set of non-member
+stars that nonetheless are subject to similar selection criteria, we
+defined the reference ``neighborhood'' as the group of at most $10^4$
+randomly selected non-member stars within 5 standard deviations of the
+mean values of the right ascension, declination, and parallax of
+25$\,$Ori-1.  We queried Gaia DR2 for these stars using
 \texttt{astroquery} \citep{astroquery_2018}.  This yielded 1{,}819
@@
 defined over-density in each dimension: the cluster was confirmed to
-exist, and to be distinct from the neighborhood.  The second point we noted is that \ptfo\ belongs
-to the cluster, based on its properties in each of these dimensions.
+exist, and to be distinct from the neighborhood.  The second point we
+noted is that \ptfo\ belongs to the cluster, based on its properties
+in each of these dimensions.
 
@@
 
-A separate line of evidence for binarity is the Gaia DR2
-astrometry.  As noted in Section~\ref{sec:observations}, the Gaia DR2
-astrometric solution for \ptfo\ shows a 10.3$\sigma$ ``astrometric
-excess'', a parameter that quantifies the degree to which a
-single-star model fails to fit the astrometric measurements.
-Specifically, the single-source astrometric model yielded
-$\chi^2=325.2$.  There are 121 astrometric measurements, and 5 free
-parameters, and therefore 116 degrees of freedom. The reduced $\chi^2$
-is 2.80.  The majority of stars with comparable brightness in Gaia do
-not show such poor goodness-of-fit \citep[see][Appendix
-A]{lindegren_gaiasoln_2018}.
+A separate line of evidence for binarity is the Gaia DR2 astrometry.
+As noted in Section~\ref{sec:observations}, the Gaia DR2 astrometric
+solution for \ptfo\ shows a 10.3$\sigma$ ``astrometric excess'', a
+parameter that quantifies the degree to which a single-star model
+fails to fit the astrometric measurements.  Specifically, the
+single-source astrometric model yielded $\chi^2=325.2$.  There are 121
+astrometric measurements, and 5 free parameters, and therefore 116
+degrees of freedom. The reduced $\chi^2$ is 2.80.  The majority of
+stars with comparable brightness in Gaia do not show such poor
+goodness-of-fit \citep[see][Appendix A]{lindegren_gaiasoln_2018}.
 
@@
 e.g.},][]{rizzuto_ZEIT8_2018,belokurov_unresolved_2020}.  If
-photometric variability were the cause, we would expect stars of similar
-brightness in the same kinematic group of Orion to show similar
-astrometric excesses, because the majority of young stars are highly
-variable.
-
-% NB: CDIPS-I only made 131 light curves for 25Ori-1, because I wasn't
-% aware of the Kounkel+18 member list at the time.
+photometric variability were the cause, we would expect stars of
+similar brightness in the same kinematic group of Orion to show
+similar astrometric excesses, because the majority of young stars are
+highly variable.
+
 Using the same 149 members in the 25$\,$Ori-1 subgroup, we calculated
@@
 One was CVSO~35, which has a TESS light curve that varies by 2
-magnitudes.
-CVSO~35 also shows a strong Wide-field Infrared Survey Exoplorer
-(WISE, \citealt{wright_WISE_2010}) IR excess as well as a 10$\mu$m silicate
-emission feature \citep{mauco_herschel_2018}.  The other is GAIA DR2
-3222210363837122048.
+magnitudes.  CVSO~35 also shows a strong Wide-field Infrared Survey
+Exoplorer (WISE, \citealt{wright_WISE_2010}) IR excess as well as a
+10$\mu$m silicate emission feature \citep{mauco_herschel_2018}.  The
+other is GAIA DR2 3222210363837122048.
 
@@
 
-Radial velocity (RV) measurements over sufficiently long timescales could also reveal the
-presence of multiple stars in this system.  Unfortunately, the
-available RV data for \ptfo\ is sparse, presumably due to the difficulties
-of performing RV observations of such a faint and rapidly rotating
-star.  The RV datasets with the
-longest time baselines we could find in the literature were those reported
-by \citet{van_eyken_ptf_2012}.  These included 5 Keck/HIRES
-measurements acquired over 10 days in April 2011, and 4 HET/HRS
-measurements acquired over 10 days in February 2011.  The
-root-mean-squared RV over each 10-day span was $\approx$$2\,{\rm
-km}\,{\rm s}^{-1}$, consistent with the measurement precision.
-Although \citet{van_eyken_ptf_2012} tried a CCF-based RV reduction
-technique, they eventually found that manually selecting absorption
-lines and measuring line centroids was more effective.  While
-\citet{yu_tests_2015} acquired 22 further Keck/HIRES spectra over one
-night in December 2013, those points were not reduced to velocities
-Further radial-velocity observations
-could potentially confirm or refute the presence of binary companions.
+Radial velocity (RV) measurements over sufficiently long timescales
+could also reveal the presence of multiple stars in this system.
+Unfortunately, the available RV data for \ptfo\ is sparse, presumably
+due to the difficulties of performing RV observations of such a faint
+and rapidly rotating star.  The RV datasets with the longest time
+baselines we could find in the literature were those reported by
+\citet{van_eyken_ptf_2012}.  These included 5 Keck/HIRES measurements
+acquired over 10 days in April 2011, and 4 HET/HRS measurements
+acquired over 10 days in February 2011.  The root-mean-squared RV over
+each 10-day span was $\approx$$2\,{\rm km}\,{\rm s}^{-1}$, consistent
+with the measurement precision.  Although \citet{van_eyken_ptf_2012}
+tried a CCF-based RV reduction technique, they eventually found that
+manually selecting absorption lines and measuring line centroids was
+more effective.  While \citet{yu_tests_2015} acquired 22 further
+Keck/HIRES spectra over one night in December 2013, those points were
+not reduced to velocities Further radial-velocity observations could
+potentially confirm or refute the presence of binary companions.
 
@@
 We are not the first to reach the conclusion that the long period
-sinusoidal modulation is astrophysical. 
-\citet{koen_multicolour_2015} identified the same modes and aliases as
-\citet{van_eyken_ptf_2012}, and argued that the signal was
-astrophysical, even if the exact period was still unclear.
-Using photometry from the YETI global telescope network,
-\citet{raetz_yeti_2016} came to the conclusion that the
+sinusoidal modulation is astrophysical.  \citet{koen_multicolour_2015}
+identified the same modes and aliases as \citet{van_eyken_ptf_2012},
+and argued that the signal was astrophysical, even if the exact period
+was still unclear.  Using photometry from the YETI global telescope
+network, \citet{raetz_yeti_2016} came to the conclusion that the
 $0.50\,{\rm d}$ signal was indeed from stellar rotation.  The TESS
@@
     - P_{\rm s})/P_{\rm s}$, highlighting a possible numerical
-    coincidence.  The orbital phase observed
-    by TESS (lower-right) is consistent with that of
-    \citet{tanimoto_evidence_2020}.
+    coincidence.  The orbital phase observed by TESS (lower-right) is
+    consistent with that of \citet{tanimoto_evidence_2020}.
 		\label{fig:o_minus_c}
@@
 
-An interesting feature of the sequence of dips is that the phase of the dips has been observed to change with time \citep{yu_tests_2015}.  In
-fact, \citet{tanimoto_evidence_2020} provided stark evidence for
+An interesting feature of the sequence of dips is that the phase of
+the dips has been observed to change with time \citep{yu_tests_2015}.
+In fact, \citet{tanimoto_evidence_2020} provided stark evidence for
 different behavior altogether: over a timespan of years, the dip
-``split'' into distinct groups at particular phases.  See,
-for instance, their Figures~2 through~4.  Fitting a decade of
-observations, they provided the following constant-period ephemeris, which we
-did not find any need to update:
+``split'' into distinct groups at particular phases.  See, for
+instance, their Figures~2 through~4.  Fitting a decade of
+observations, they provided the following constant-period ephemeris,
+which we did not find any need to update:
 \begin{align}
@@
 
-Figure~\ref{fig:o_minus_c} shows the differences bewteen the observed ``mid-transit'' times of
-the dips and the times calculated using Equation~\ref{eq:ephem}.  The
-phase of the dips seen by TESS (yellow star) agrees
-with the independent December 2018 measurements by
+Figure~\ref{fig:o_minus_c} shows the differences bewteen the observed
+``mid-transit'' times of the dips and the times calculated using
+Equation~\ref{eq:ephem}.  The phase of the dips seen by TESS (yellow
+star) agrees with the independent December 2018 measurements by
 \citet{tanimoto_evidence_2020}: either the dip abruptly shifted phase
@@
 Figure~\ref{fig:o_minus_c} shows two additional strange features: {\it
-(i)}  multiple dips per cycle, and {\it (ii)} a set of dips
-at a phase that is numerically coincident with $(P_{\rm \ell} - P_{\rm s}) / P_{\rm
-s}$.  The observation of multiple dips per cycle in 2015 was seen
-independently by both \citet{yu_tests_2015} and
+(i)}  multiple dips per cycle, and {\it (ii)} a set of dips at a phase
+that is numerically coincident with $(P_{\rm \ell} - P_{\rm s}) /
+P_{\rm s}$.  The observation of multiple dips per cycle in 2015 was
+seen independently by both \citet{yu_tests_2015} and
 \citet{tanimoto_evidence_2020}.  It therefore seems credible.
@@
 leave this curiosity as observation {\it sans} interpretation.
-% JNW: This seems like an important clue. I don't have the mental
-% energy to ponder it now, but it deserves pondering.
-% JNW: Could this mean that both stars are dippers? A change in period
-% is degenerate with a change in period, over a short interval, right?
-% so maybe one set of dips belongs to one star, the other set to the 2nd star?
-% Features with a different period would appear to lie along a line
-% of constant slope in this diagram...what would the slope be, in this case,
-% if the feature period were P_long?
-% LGB: Seems unlikely, because we're seeing a something with constant
-% 10.74 hour periodicity, but the phase is offset by 1/9th.
-% If both stars were dippers, we would be seeing a 11.98 hour
-% periodicity instead.
 
 \subsection{Short Period Modulation Outside of Dips}
-
 Visually, the out-of-dip modulation at the 10.76$\,$hr period
@@
   \caption{ {\bf \ptfo\ and its brethren.}
-    Shown are the light curves of five transient and persistent flux dip stars, selected based on
-    their similarity to the short-period signal of \ptfo.  The stars are EPIC 204143627, EPIC 204270520, EPIC
-    204321142, EPIC 246938594, and EPIC 205483258 (RIK-210).  RIK-210
-    has the longest period of any of these objects.  All the analogs
-    displayed are either in Taurus or Upper~Sco, and meet the
+    Shown are the light curves of five transient and persistent flux
+    dip stars, selected based on their similarity to the short-period
+    signal of \ptfo.  The stars are EPIC 204143627, EPIC 204270520,
+    EPIC 204321142, EPIC 246938594, and EPIC 205483258 (RIK-210).
+    RIK-210 has the longest period of any of these objects.  All the
+    analogs displayed are either in Taurus or Upper~Sco, and meet the
     characteristics of Section~\ref{subsec:dipstars}.  These objects
@@
 
-When physical explanations are not forthcoming, we often resort to taxonomy.
-By searching the literature, we have found about a dozen
+When physical explanations are not forthcoming, we often resort to
+taxonomy.  By searching the literature, we have found about a dozen
 light curves with similar morphologies to \ptfo, drawn from surveys of
@@
     \citealt{rebull_usco_2018},~Figure~20).
-  \item The age is typically $\lesssim$ 100$\,$Myr.\footnote{
-    At present, the oldest observed ``scallops'' are in the Pleaides
+  \item The age is typically $\lesssim$ 100$\,$Myr.\footnote{ At
+    present, the oldest observed ``scallops'' are in the Pleaides
     \citep{rebull_rotation_2016}. One of these, EPIC 211013604, might
@@
     oldest known.} 
-  \item The light curves show shallow, angular dips, usually superposed
-    on large-amplitude smooth variability. The latter is interpreted
-    as stellar rotation.
-  \item The rotation is rapid, with a period that is usually between 0.5 and 2.0 days.
-  \item There is rarely any infrared excess that is detectable in the WISE data
-  (never any W4 detection; only a few W3 detections).
+  \item The light curves show shallow, angular dips, usually
+    superposed on large-amplitude smooth variability. The latter is
+    interpreted as stellar rotation.
+  \item The rotation is rapid, with a period that is usually between
+    0.5 and 2.0 days.
+  \item There is rarely any infrared excess that is detectable in the
+    WISE data (never any W4 detection; only a few W3 detections).
   \item They sometimes show multiple dips per cycle.
@@
   \item The dip depths can change after flares.
-  \item They are rare at a population level, with an occurrence rate of $\lesssim 1\%$ of
-  young M2-M5 stars \citep{rebull_usco_2018}.
+  \item They are rare at a population level, with an occurrence rate
+    of $\lesssim 1\%$ of young M2-M5 stars \citep{rebull_usco_2018}.
 \end{enumerate}
 The 10.76$\,$hr signal in \ptfo\ meets all of these criteria.  This is
-the first connection of \ptfo\ with this class of objects, 
-likely because the TESS data enabled us to put the dips in 
-the context of the asymmetric out-of-dip modulation.
+the first connection of \ptfo\ with this class of objects, likely
+because the TESS data enabled us to put the dips in the context of the
+asymmetric out-of-dip modulation.
 
@@
 
-Second, approximately 40-50\% of the transient flux dip stars discovered in
-$\rho$~Oph and Upper~Sco show two Lomb-Scargle periods, and so are
-apparently binaries \citep[][Table~1]{stauffer_orbiting_2017}.  This
-is higher than the main-sequence companion fraction of ${\rm
+Second, approximately 40-50\% of the transient flux dip stars
+discovered in $\rho$~Oph and Upper~Sco show two Lomb-Scargle periods,
+and so are apparently binaries
+\citep[][Table~1]{stauffer_orbiting_2017}.  This is higher than the
+main-sequence companion fraction of ${\rm
 CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$
@@
 and Taurus \citep{kraus_mapping_2008,kraus_mapping_2011}.  A
-high-resolution imaging survey would be interesting, to determine whether
-the transient flux dip stars truly have distinct population-level
-binarity properties relative to other young low-mass stars.
+high-resolution imaging survey would be interesting, to determine
+whether the transient flux dip stars truly have distinct
+population-level binarity properties relative to other young low-mass
+stars.
 
@@
 The evidence for binarity in \ptfo\ is as follows.  First, the star is
-twice as bright as stars of the same color in its
-kinematic group (Figure~\ref{fig:gaia}).  Second, it shows two
-distinct photometric signals.  These points alone suggest binarity
+twice as bright as stars of the same color in its kinematic group
+(Figure~\ref{fig:gaia}).  Second, it shows two distinct photometric
+signals.  These points alone suggest binarity
 \citep{stauffer_rotevol_2018}.  For the case of \ptfo, there is a
@@
 When this manuscript was at an advanced stage, we received notice of a
-paper by \citet{koen_2020} that was in press at the
-{\it Monthly Notices} before submission of our manuscript.  Our
-studies independently reached the same conclusions: the TESS
-light curve shows two periodic signals, and the properties of
-\ptfo\ are consistent with the emerging class of transient and
-persistent flux dip stars.  \citet{koen_2020} reached these
-conclusions by modeling the TESS light curve as a truncated sum of
-Fourier terms, and concluded that the two signals are most simply
-interpreted as coming from two stars.  Our analysis of the Gaia data
-provide independent support for the conclusion that \ptfo\ is a
-binary. We also note the agreement between the
-TESS dip ephemeris and that from \citet{tanimoto_evidence_2020}.
+paper by \citet{koen_2020} that was in press at the {\it Monthly
+Notices} before submission of our manuscript.  Our studies
+independently reached the same conclusions: the TESS light curve shows
+two periodic signals, and the properties of \ptfo\ are consistent with
+the emerging class of transient and persistent flux dip stars.
+\citet{koen_2020} reached these conclusions by modeling the TESS light
+curve as a truncated sum of Fourier terms, and concluded that the two
+signals are most simply interpreted as coming from two stars.  Our
+analysis of the Gaia data provide independent support for the
+conclusion that \ptfo\ is a binary. We also note the agreement between
+the TESS dip ephemeris and that from \citet{tanimoto_evidence_2020}.
 \\

### 2020-05-14 drivers/calc_mmin_given_gammadot.py
add hires spectra driver scripts

@@ -1,0 +1,9 @@
+from astropy import units as u
+
+tau = 2.66*u.year
+gammadot = 50*u.m/u.s/u.day
+mstar = 0.35 * u.Msun
+
+m_min = 6*u.Mjup * (tau/(1*u.year))**(4/3) * (gammadot/(1*u.m/u.s/u.day)) * (mstar/(1*u.Msun))**(2/3)
+
+print(m_min.to(u.Msun))

### 2020-05-14 drivers/plot_hires_spectra.py

@@ -1,0 +1,12 @@
+from astropy.io import fits
+from glob import glob
+import numpy as np, matplotlib.pyplot as plt
+from cdips_followup.spectools import plot_orders
+import os
+
+outdir = '../results/hires_spectra/plot_orders/'
+specpaths = glob('../data/hires_spectra/deblazed/*117*fits')
+
+for s in specpaths:
+    idstring = os.path.basename(s).replace('.fits', '')
+    plot_orders(s, outdir=outdir, idstring=idstring)

### 2020-05-15 paper/ms.tex
update disk discussion

@@
 Based on the lack of an infrared excess seen by \citet{yu_tests_2015},
-the primordial gas disks seem to be have been depleted\footnote{A
-potential confounding factor is completeness. Is Spitzer sensitive to
-infrared excesses at the distance of Orion? The series of studies by
-\cite{hernandez_spitzer_2006,hernandez_spitzer_ob1_2007,hernandez_spitzer_sig_2007,hernandez_spitzer_2009}
-show that it is.}
-% JNW: I don't understand why we can be so definitive...
-% an infrared excess is not either 'present' or 'absent'. The upper limit in Jy
-% corresponds to a constraint on the mass/radius/etc of any dusty disk.
-around both stars in \ptfo. The stars are therefore no longer
-magnetically locked to their disks.  This is consistent with the
-$\approx$half-day periodicities of both rotation signals: young disked
-M dwarfs typically rotate with periods of two days or more due to
-magnetic locking \citep[{\it e.g.},][]{rebull_rotation_2020}.  If the
-two stars are within $\approx$50$\,$AU of each other, as required by
-the NIRC2 adaptive optics imaging, then it would also be expected that
-the stars would have truncated the outer edges of their respective
-disks, in a manner seen at the population level in exoplanetary
-systems \citep{kraus_impact_2016,moe_impact_2019}.
-
-The key remaining question is what causes the transient dips. This is
+the primordial gas disks of both stars in \ptfo\ seem to be have been
+depleted.  This is consistent with the $8.5\pm1.2\,{\rm Myr}$ age of
+the 25$\,$Ori-1 group, and the rapid rate at which stars show lose
+their disks between 1 and 10$\,$Myr \citep[{\it
+e.g.},][]{hernandez_spitzer_ob1_2007}.  The stars are therefore
+presumably no longer magnetically locked to their disks.  This is also
+suggested by the $\approx$half-day periodicities of both rotation
+signals: young disked M dwarfs typically rotate with periods of two
+days or more due to magnetic locking \citep[{\it
+e.g.},][]{rebull_rotation_2020}.  If the two stars are within
+$\approx$50$\,$AU of each other, as required by the NIRC2 adaptive
+optics imaging, then it would also be expected that the stars would
+have truncated the outer edges of their respective disks, in a manner
+seen at the population level in exoplanetary systems
+\citep{kraus_impact_2016,moe_impact_2019}.
+
+The key question is what causes the transient dips. This is
 an unsolved problem not only for \ptfo\ but also for the emerging

### 2020-05-15 billy/plotting.py
O-C: add "P_long" phasing option, just to check
doesn't lead to much insight!

@@ -952,7 +952,8 @@ def plot_astrometric_excess(outdir, ruwe=0):
 def plot_O_minus_C(
     x, y, sigma_y, theta_linear, refs, savpath=None, xlabel='Epoch',
     ylabel='Deviation from constant period [min]', xlim=None, ylim=None,
-    ylim1=None, include_all_points=False, onlytransits=False):
+    ylim1=None, include_all_points=False, onlytransits=False,
+    plongphasing=False):
 
     xfit = np.linspace(10*np.min(x), 10*np.max(x), 10000)
 
@@ -976,14 +977,19 @@ def plot_O_minus_C(
     P_short = 0.44854
     P_long = 0.49914
     phase_diff = np.abs(P_long - P_short) / P_short
-    a0.hlines(
-        phase_diff, min(xlim), max(xlim), colors='gray', alpha=0.5,
-        linestyles='--', zorder=-2, linewidths=0.5
-    )
-    a0.hlines(
-        -phase_diff, min(xlim), max(xlim), colors='gray', alpha=0.5,
-        linestyles='--', zorder=-2, linewidths=0.5
-    )
+    if isinstance(xlim, tuple):
+        pass
+    else:
+        xlim = a0.get_xlim()
+    if not plongphasing:
+        a0.hlines(
+            phase_diff, min(xlim), max(xlim), colors='gray', alpha=0.5,
+            linestyles='--', zorder=-2, linewidths=0.5
+        )
+        a0.hlines(
+            -phase_diff, min(xlim), max(xlim), colors='gray', alpha=0.5,
+            linestyles='--', zorder=-2, linewidths=0.5
+        )
     a0.hlines(
         0, min(xlim), max(xlim), colors='gray', alpha=0.5,
         linestyles='-', zorder=-2, linewidths=0.5
@@ -992,12 +998,20 @@ def plot_O_minus_C(
 
     props = dict(boxstyle='square', facecolor='white', alpha=0.9, pad=0.15,
                  linewidth=0)
-    txt = (
-        '$t_{{\mathrm{{s}}}}$ = {:.6f}\n$P_{{\mathrm{{s}}}}$ = {:.6f}$\,$d'.
-        format(theta_linear[0], theta_linear[1])
-    )
-    a0.text(0.03, 0.03, txt, ha='left', va='bottom', transform=a0.transAxes,
-            bbox=props, zorder=3)
[…]

### 2020-05-15 drivers/plot_O_minus_C.py

@@ -13,7 +13,8 @@ PERIOD = 0.448399
 
 def get_data(
     datacsv='../data/WASP-18b_transits_and_TESS_times_O-C_vs_epoch_selected.csv',
-    is_occultation=False
+    is_occultation=False,
+    plongphasing=False
     ):
 
     df = pd.read_csv(datacsv, sep=';')
@@ -29,14 +30,20 @@ def get_data(
 
     y, sigma_y, refs = y[sel], sigma_y[sel], refs[sel]
 
-    x, _ = get_epochs_given_midtimes_and_period(
-        y, PERIOD, t0_fixed=EPOCH, verbose=True
-    )
+    if not plongphasing:
+        x, _ = get_epochs_given_midtimes_and_period(
+            y, PERIOD, t0_fixed=EPOCH, verbose=True
+        )
+    else:
+        x, _ = get_epochs_given_midtimes_and_period(
+            y, 0.4991110, t0_fixed=EPOCH, verbose=True
+        )
 
     return x, y, sigma_y, refs
 
 
-def main(xlim=None, ylim=None, include_all_points=False, ylim1=None):
+def main(xlim=None, ylim=None, include_all_points=False, ylim1=None,
+         plongphasing=0):
 
     homedir = os.path.expanduser('~')
     basedir = os.path.join(homedir, 'Dropbox/proj/billy/')
@@ -46,22 +53,27 @@ def main(xlim=None, ylim=None, include_all_points=False, ylim1=None):
     occpath = 'foo.csv'
 
     print('getting data from {:s}'.format(transitpath))
-    x, y, sigma_y, refs = get_data(datacsv=transitpath)
+    x, y, sigma_y, refs = get_data(datacsv=transitpath,
+                                   plongphasing=plongphasing)
     print('getting data from {:s}'.format(occpath))
 
     if os.path.exists(occpath):
         raise NotImplementedError
 
     linear_params = nparr([EPOCH,  PERIOD])
+    if plongphasing:
+        linear_params = nparr([EPOCH, 0.4991110])
 
     savpath = '../results/ephemeris/O_minus_C.png'
+    if plongphasing:
+        savpath = '../results/ephemeris/O_minus_C_plongphasing.png'
 
     onlytransits = True
     bp.plot_O_minus_C(
         x, y, sigma_y, linear_params, refs, savpath=savpath,
         xlabel='Cycle number', ylabel=None, xlim=xlim, ylim=ylim,
[…]

### 2020-05-15 paper/ms.tex
update a few words in intro

@@
 While the planetary interpretation clearly faces challenges, there is
-no completely satisfactory alternate explanation.  High-latitude
-accretion hotspots might produce the observed H$\alpha$ variability,
-but require fine-tuning to produce dips of the appropriate duration.
-Transits by dust clumps or other dusty features are questionable
-because \ptfo\ does not have a detectable infrared (IR) excess
-associated with the presence of warm dust \citep[{\it
-e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude starspots, hot or
-cold, would struggle to produce photometric features as short as some
-of the observed dips.
+no completely satisfactory alternate explanation.  Low-latitude
+starspots, hot or cold, would struggle to produce photometric features
+as short as some of the observed dips.  High-latitude accretion
+hotspots might produce the observed H$\alpha$ variability, but require
+fine-tuning to produce dips of the appropriate duration.  Transits by
+dust clumps or other dusty features are questionable because \ptfo\
+does not have a detectable infrared (IR) excess associated with the
+presence of warm dust \citep[{\it e.g.},][Figure~18]{yu_tests_2015}.
+In addition, the sublimation times for dust grains of plausible
+composition are quite short \citep{zhan_complex_2019}.
 
@@
 		\leavevmode
-		\includegraphics[width=0.5\textwidth]{f6.pdf}
+		\includegraphics[width=0.48\textwidth]{f6.pdf}
 	\end{center}
@@
 {\it (iii)} eclipses of prominences;
-% JNW: why is this one ruled out? Sounds similar to the accretion-hotspot hypothesis.
 {\it (iv)} high-latitude accretion hotspots;
@@
 systems.
-% JNW Could the material get magnetically locked at the corotation radius?
 

### 2020-05-15 paper/ms.tex
minor word smithing

@@
   Pre-main sequence stars (1290),
-	Stellar ages (1581),
 	Stellar rotation (1629),
@@
 age of the 25$\,$Ori-1 subgroup, determined by fitting isochrones to
-group members that had effective temperatures from APOGEE and Gaia
-parallaxes, was determined to be $8.5\pm1.2\,{\rm Myr}$
-\citep[see][Section 2.3]{kounkel_apogee2_2018}.
-\citet{kounkel_apogee2_2018} identified seven other smaller groups in
-the Orion complex near the Be star 25$\,$Ori. These groups received
-sequential identifiers, {\it e.g.}, 25$\,$Ori-2 (${\rm Age}
-=12.9\pm2.8\,$Myr; see also \citealt{briceno_cidaII_2019}).
+group members with APOGEE effective temperatures and Gaia parallaxes,
+was determined to be $8.5\pm1.2\,{\rm Myr}$ \citep[see][Section
+2.3]{kounkel_apogee2_2018}.  \citet{kounkel_apogee2_2018} also
+identified seven smaller groups in the Orion complex near the Be star
+25$\,$Ori. These groups received sequential identifiers, {\it e.g.},
+25$\,$Ori-2 (${\rm Age} =12.9\pm2.8\,$Myr; see also
+\citealt{briceno_cidaII_2019}).
 
@@
 		\leavevmode
-		\includegraphics[width=1\textwidth]{f3.pdf}
+		\includegraphics[width=.93\textwidth]{f3.pdf}
 	\end{center}

### 2020-05-15 paper/ms.tex
paper: fix typos

@@
 probably oblate enough to torque the planet's orbit into and out of
-the transiting configuration on a time scale of years
+the transiting configuration on a timescale of years
 \citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.

### 2020-05-20 paper/ms.tex
typo
fix typo

@@
 %% Add "Submitted to " argument.
-\received{\today}
-\revised{---}
+\received{May 15, 2020}
+\revised{\today}
 \accepted{---}
@@
   of a not-quite sinusoid interrupted by hour-long dips, the type of
-  signal previously intepreted as planetary transits.  The phase of
+  signal previously interpreted as planetary transits.  The phase of
   the dips is nearly 180$^\circ$ away from the phase of the originally
@@
 Keck/HIRES spectra over one night in December 2013, those points were
-not reduced to velocities Further radial-velocity observations could
+not reduced to velocities. Further radial-velocity observations could
 potentially confirm or refute the presence of binary companions.
@@
 signals are most simply interpreted as coming from two stars.  Our
-analysis of the Gaia data provide independent support for the
+analysis of the Gaia data provides independent support for the
 conclusion that \ptfo\ is a binary. We also note the agreement between

### 2020-06-12 paper_review/response_20200605.txt
WIP for the referee report
added some nice details to flesh out the torus of gas+dust in favor of
the starspot occultations

Dear Dr. Shafter,

We thank you for organizing the review of our paper, and also
thank the referee for their helpful comments.  We have
considered the referee's remarks carefully and revised our
manuscript accordingly.

Below, we have reproduced the relevant portions of the
referee's report, alongside our responses. A list of changes is
appended to the manuscript using the trackchanges AASTeX
macros.

Sincerely,

Luke Bouma

******************************************
REFEREE COMMENT
----------
Overall: It would help if the authors used hours or days throughout the paper
in a consistent manner. The period is referred to as 0.5d and 11.98h just a
couple paragraphs apart (e.g. in Section 5.1). The abstract uses hours, most
plots use days.

RESPONSE
----------
> Throughout the manuscript, we have shifted the units wherever possible to
> hours.
>
> Exceptions include the ephemeris (which is quoted in days), and the periods
> listed in Figures 3 and 6, which are quoted in days. The former is standard
> convention. The latter was because we felt it important to list the periods
> used to generate the figures at a higher precision, and quoting higher
> precision in units of hours seems atypical in a field in which days are more
> typically seen.

******************************************
REFEREE COMMENT
----------

Introduction: The introduction takes a light/jovial tone, e.g., did they
actually wish it were a planet, or is it more that they aimed to confirm it and
found the opposite? I would normally suggest making more formal. However, I
enjoyed reading it. I'm sorry the authors ruined their own favorite planet.

RESPONSE
----------
> So are we!  Our aim was to clarify what could be happening in the system, and
> hopefully this contribution is valuable for that.  When a planetary door
> closes, a window on young star variability opens...

******************************************
REFEREE COMMENT
----------

Introduction: I do not think this is youngest hot Jupiter. That should be true
just for transiting systems, and it was the youngest at discovery, but the two
RV planets from ESPaDOnS are younger (Donati et al. 2016 and Yu et al. 2017).
Both are hot Jupiters and in Taurus. I believe the age of Tap 26 is more
ambiguous due to age spreads (multiple groups?) in Taurus, but the age of V830
Tau is certainly <3 Myr.

RESPONSE
----------

> Thank you for this correction. We have qualified the relevant sentence as the
> youngest known *transiting* hot Jupiter. We have opted to not cite the V830
> Tau or TAP 26 results, which we hope to see confirmed by independent groups.

******************************************
REFEREE COMMENT
----------

Introduction: I also think it wasn't the first H-alpha detection. That belongs
to LkCa15, which was imaged in H-alpha in 2015 (Sallum et al. 2015). Although
that planet might not exist, so this comment is probably OK. It would still be
the first close-in planet either way. I'd suggest weakening the statement
slightly to err on the side of caution.

RESPONSE
----------
> We have changed the wording to err on the side of caution, and qualify it as
> the first direct detection of Hα emission from a close-in planet.

******************************************
REFEREE COMMENT
----------

Introduction: Barnes et al. (2013) claimed that the planet and star precessed
on a period of 300-600 days, and that this could somewhat fit variable depths
and missing transits. The Ciardi et al. 2015 paper uses this as their model, in
addition to gravity darkening as mentioned in the current introduction. I
suggest mentioning the precession of the planet in the introduction, as many
papers used the precession model or at least discussed it. The line "...the
gravity-darkened star precessed..." seems to allude to this, but it's unclear
from the rest of the text.

RESPONSE
----------
> We have changed the wording of the last sentence of the first paragraph in
> the introduction to explicitly use the word "precess", rather than "torque".

******************************************
REFEREE COMMENT
----------

Section 2: The first paragraph discusses the full frame images, while the next
section discusses the 2m data. Why did the authors use the FFIs when they have
the (higher quality) 2m data?

RESPONSE
----------
> For the essential analysis of the study (i.e., extracting and fitting the
> long and short signals), we used the 2m SPOC lightcurve.
>
> We primarily mention our use of the FFI lightcurves in Section 2 because of
> the systematic jump mentioned at BJD 2458488.3, which was eventually excluded
> from further analysis. We wanted to be sure though that this jump was truly
> an instrumental systematic, and not a software systematic. Checking
> lightcurves from independent pipelines meets this goal.
>
> A secondary reason to introduce the concept of FFIs in Section 2 is that they
> are used to measure the variability amplitudes of the other members of 25
> Ori-1 in Section 4.3.

******************************************
REFEREE COMMENT
----------

Section 2.2.2: "These details concerning the group membership for one object
may seem excessive to those accustomed to the simple distinction between "young
cluster members" and "old field stars"." I suggest removing this. It's not
adding anything and sounds like the authors are making assumptions about the
reader. The membership of this object is essential for an age assessment, and
can/should be updated given the new Gaia data. Since most of the object class
in question are at similar ages, this directly relates to the study. If it had
turned out to be ~100 Myr, we would all be more skeptical (although ~100 Myr
stars like this are known, they are more rare).

RESPONSE
----------
> We've removed it -- thank you for the suggestion.

******************************************
REFEREE COMMENT
----------

Section 3/4: This deals mostly with identification of the second star. However,
the second star appears to be an assumption of section 3.2. If the periods come
from two stars that could be separated (e.g., wide binaries), it would have
been best to do so rather than try to model both by including fs and fl terms.
So it is odd that the binary search is after the light curve model in the
paper. It might make more narrative sense to move the light curve model to
after section 4, since 4 motivates 3.2.

RESPONSE
----------
> While we see the referee's point, we prefer the narrative structure of
> presenting the TESS data first, and getting into considerations of stellar
> binarity, both resolved and unresolved, afterwards.
>
> The reasons are that i) we wrote the paper primarily because of the TESS
> data; ii) the multiple signals come from TESS data. From a story-telling
> perspective, why would we worry about binarity if we didn't already know
> there were multiple signals?

******************************************
REFEREE COMMENT
----------

Section 4.3: My biggest concern with this analysis is reliance on the chi^2
from Gaia to assess binarity. This is not the suggested path from Gaia, as it
is subject to a noise model Gaia now knows has issues. For example, nearly 100%
of mid/late M dwarfs have a high chi^2 (and a large excess astrometric noise),
but clearly they are not all binaries. This is (partially) because of a
color-term in the PSF model that was not initially understood. Similarly,
almost all bright stars have a large RUWE, as do the bluest stars. The authors
correctly use RUWE, the suggested method, but the RUWE analysis is less
convincing and clearly the authors are leaning on the chi^2.

A good reference on this would be Belokurov et al. 2020, which shows what kinds
of binaries RUWE can and cannot detect. This binary might simply be too tight
to create an astrometric signal (in which case the RUWE makes sense). The
typical cutoff for binarity is a RUWE of ~1.4, see Rizzuto et al. 2018, Ziegler
et al. 2019, Pearce et al. 2020, among many others.  I suggest removing the
chi^2 analysis. If the authors want to keep it, they need to argue why it is a
more meaningful metric than RUWE, and attempt to account for color and other
parameters that impact chi^2.

RESPONSE
----------
> We've done as suggested, and replaced this sub-plot with RUWE instead of
> reduced chi^2. The text in Section 4.3 and captions have been updated
> accordingly, and chi^2 is now mostly ignored.

******************************************
REFEREE COMMENT
----------

Section 4.4:
Need a period between velocities and Further.

RESPONSE
----------
> Corrected, thank you.

******************************************
REFEREE COMMENT
----------

Section 5.2: That clear offset at (Pl-Ps)/Ps doesn't seem like a coincidence,
and the paper reads like the authors agree, but couldn't explain it. I'm not
sure I can help much here, although it is interesting. Could this be that both
stars are dippers and one turns off while the other turns on by coincidence?
Spots and hot-spot clusters can shift in phase as the magnetic fields realign,
which creates a new spot cluster at the same period but a different phase. If
it is circumstellar material, a tight binary could be shepherding the material,
but I think the systems would have to be very tight for that, and it could not
last for such a long time.

I think this can be assessed statistically; given the scatter in those points,
the odds of lining up to any given line is small. But without a physical
explanation this is hard to test. Probably this requires looking at many
dippers as an ensemble. It is not required for the primary analysis of this
paper, so I will defer to the authors if they want to dig deeper on this.

RESPONSE
----------
> We're not sure whether or not it is a coincidence, but it seemed interesting
> enough to mention.
>
> ..

******************************************
REFEREE COMMENT
----------

Section 5.3.2: The difference between the fraction of transient flux dip stars
and the binary fraction is more significant than it appears to the authors.
High mass ratio binaries would not be included, as they are too faint to detect
the second period. Further, at the distance of Upper Sco (~150pc) and assuming
Stauffer et al. 2017 would notice/separate two stars wider than a K2 pixel
(4"), the presence of two periods would not include binaries wider than 600pc.
Eyeballing the Winters et al. 2019 distribution suggests ~20% of binaries are
beyond this. Stauffer et al. 2017 also probably checked against existing
imaging data for obvious binaries down to 1-2", so likely even more are getting
cut.

RESPONSE
----------
>

******************************************
REFEREE COMMENT
----------

Section 5.4: The authors conclude this subsection with a reasonable argument
that this isn't a protoplanet. Does this same argument not rule out gas/debris
at the Keplerian coronation radius? What could maintain such clouds for
years/decades and then have them jump phase but at the same period? It's hard
to do this if it is not somehow physically attached to the star in some what
(e.g., spots can do this).

RESPONSE
----------
>

******************************************
ADDITIONAL CHANGES

A few additional changes are noted in the "List of Changes" appended at the end
of the manuscript.

We modified the text throughout to clarify that Stauffer+ 2017 favored the
explanation of a clumpy torus of gas and dust at the Keplerian co-rotation
radius, rather than our initial (incomplete) description of "clouds of gas" at
the same location.

We also modified the labels of the axes in Figure 5a for clarity.

### 2020-06-12 paper/ms.tex
WIP for the referee report
added some nice details to flesh out the torus of gas+dust in favor of
the starspot occultations

@@
 
-\documentclass[12pt,twocolumn,tighten]{aastex62}
-%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
+%\documentclass[12pt,twocolumn,tighten]{aastex62}
+\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
 \usepackage{amsmath,amstext,amssymb}
@@
 We wish \ptfob\ were a planet. It would be quite exceptional. It would
-be the youngest known hot Jupiter \citep{van_eyken_ptf_2012}, orbiting
+be the youngest known \added{transiting} hot Jupiter \citep{van_eyken_ptf_2012}, orbiting
 a T Tauri star in the Orion-OB1a cluster.  It would have the shortest
@@
 host star.  Not only that, but the rapidly-rotating host star is
-probably oblate enough to torque the planet's orbit into and out of
-the transiting configuration on a timescale of years
+probably oblate enough to \replaced{torque the planet's orbit into and out of
+the transiting configuration}{cause the planet's orbit to precess into and out of the 
+transiting configuration} on a timescale of years
 \citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.
@@
 Another first would be the direct detection of H$\alpha$ emission from
-the planet itself \citep{johnskrull_h_2016}.  In addition to the
+\replaced{the planet itself}{a close-in planet} \citep{johnskrull_h_2016}.  In addition to the
 chromospheric H$\alpha$ emisson, it seems that there is an additional
@@
 In addition, the sublimation times for dust grains of plausible
-composition are quite short \citep{zhan_complex_2019}.
+composition \replaced{are}{could be} quite short \citep{zhan_complex_2019}.
 
@@
 and \citet{david_transient_2017} to explain this novel class of
-variable stars is that a circumstellar cloud of gas is orbiting near
-the co-rotation radius.  To this point, though, it has not been clear
+variable stars is that a \replaced{circumstellar cloud of gas is orbiting near
+the co-rotation radius}{clumpy torus of dust and gas orbits near the
+Keplerian co-rotation radius}.  To this point, though, it has not been clear
 if this explanation applies to \ptfo, because the determination of the
@@
 
-These details concerning the group membership for one object may seem
+\replaced{These details concerning the group membership for one object may seem
 excessive to those accustomed to the simple distinction between
 ``young cluster members'' and ``old field stars''.  Although all
-members of the Orion complex are indeed young relative to the field,
-these details are essential for assessing the photometric evidence for
+members of the Orion complex are indeed young relative to the field,}
+{While all members of the Orion complex are young relative to the field,}
+these \replaced{details}{considerations} are essential for assessing \deleted{the} photometric evidence for
 the binarity of \ptfo, because of the degeneracy between stellar
@@
 \citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
-The tallest peak occurs at 0.499\,d (11.98$\,$hr) and a second strong
-peak occurs at 0.448\,d (10.76$\,$hr). We will refer to these two
+The tallest peak occurs at \replaced{0.499\,d (11.98$\,$hr)}{11.98$\,$hr} and a second strong
+peak occurs at \replaced{0.448\,d (10.76$\,$hr)}{10.76$\,$hr}. We will refer to these two
 periods as the ``longer period'' $P_{\rm \ell}$ and the ``shorter
@@
     accentuating \ptfo's separation from the single-star sequence.
-    {\it Bottom:} Astrometric goodness-of-fit versus $Rp$ magnitude
+    {\it Bottom:} Renormalized astrometric
+    unit weight error (RUWE) versus $Rp$ magnitude
     for 25$\,$Ori-1 members.  The single-source astrometric model for
@@
 
+\explain{All paragraphs in this section have been heavily
+	reworded to emphasize RUWE over $\chi_{\rm red}^2$, as requested by the
+	referee.}
 A separate line of evidence for binarity is the Gaia DR2 astrometry.
@@
 astrometric measurements, and 5 free parameters, and therefore 116
-degrees of freedom. The reduced $\chi^2$ is 2.80.  The majority of
+degrees of freedom. The reduced $\chi^2$ is 2.80.  While the majority of
 stars with comparable brightness in Gaia do not show such poor
-goodness-of-fit \citep[see][Appendix A]{lindegren_gaiasoln_2018}.
-
-Potential explanations for the poor astrometric fit include
-photometric variability and unresolved stellar binarity \citep[{\it
-e.g.},][]{rizzuto_ZEIT8_2018,belokurov_unresolved_2020}.  If
-photometric variability were the cause, we would expect stars of
-similar brightness in the same kinematic group of Orion to show
-similar astrometric excesses, because the majority of young stars are
-highly variable.
-
-Using the same 149 members in the 25$\,$Ori-1 subgroup, we calculated
-the astrometric reduced $\chi^2$ for each member.  We then queried the
+goodness-of-fit\added{, many M dwarfs do, because of a color-dependent
+term in the mission's PSF model} \citep[see][Appendix A]{lindegren_gaiasoln_2018}.
+The reduced $\chi^2$ has been
+``renormalized'' to account for this, yielding a modified
+statistic called the renormalized unit weight error
+(RUWE\footnote{ See the Gaia DPAC technical note
+	GAIA-C3-TN-LU-LL-124-01,
+	\url{http://www.rssd.esa.int/doc_fetch.php?id=3757412}, accessed
+	2020-04-27. }).
+
+We acquired the RUWE for each of the
+149 members in 25$\,$Ori-1 identified by
+identified by \citet{kounkel_apogee2_2018}.
+We then queried the
 CDIPS light curve database at MAST \citep{bouma_cluster_2019} to find
@@
 measurements.  This yielded 30 stars of equal or greater variability.
-The lower panel of Figure~\ref{fig:gaia} shows the reduced $\chi^2$ as
-a function of stellar brightness.  \ptfo\ is in the upper 90$^{\rm
-th}$ percentile of stars showing astrometric excesses within the
-25$\,$Ori-1 group.  Relative to other M-dwarf group members with
-comparable brightnesses and variability characteristics, \ptfo\ still
-stands out by virtue of its failure to conform to a single-star
-astrometric model. This supports the interpretation that \ptfo\ is a
-binary star.
-
-Performing the same analysis using the renormalized unit weight error
-(RUWE\footnote{ See the Gaia DPAC technical note
-GAIA-C3-TN-LU-LL-124-01,
-\url{http://www.rssd.esa.int/doc_fetch.php?id=3757412}, accessed
-2020-04-27. }) rather than the reduced $\chi^2$ yielded similar
-results.  \ptfo\ has a RUWE of 1{.}22, which corresponds to the
-93$^{\rm rd}$ percentile of 25$\,$Ori-1 members.  Two of thirty stars
+The lower panel of Figure~\ref{fig:gaia} shows the RUWE of
+these stars as
+a function of stellar brightness.  \ptfo\ is at the 93$^{\rm
+	rd}$ percentile of equally-variable stars within the
+25$\,$Ori-1 group.  
+Two of the thirty stars
 with variability amplitudes greater than 9.7\% showed higher RUWE.
@@
 10$\mu$m silicate emission feature \citep{mauco_herschel_2018}.  The
-other is GAIA DR2 3222210363837122048.
+other star was GAIA DR2 3222210363837122048.
+
+Potential explanations for an elevated RUWE include
+photometric variability and unresolved stellar binarity \citep[{\it
+e.g.},][]{rizzuto_ZEIT8_2018,belokurov_unresolved_2020}.  
+If photometric variability were the root cause, we would expect stars with
+similar brightness and color in the same kinematic group of Orion to show
+similar astrometric excesses, because the majority of young stars are
+highly variable.
+However
+relative to other M-dwarf group members with
+comparable brightnesses and variability characteristics, \ptfo\ still
+stands out by virtue of its failure to conform to a single-star
+astrometric model. This supports the interpretation that \ptfo\ is a
+binary star.
 
@@
 Keck/HIRES spectra over one night in December 2013, those points were
-not reduced to velocities. Further radial-velocity observations could
+not reduced to velocities\added{.} Further radial-velocity observations could
 potentially confirm or refute the presence of binary companions.
@@
 longer-period signal ({\it e.g.}, their Figure~7), in the form of a
-peak in the periodogram at $0.9985 \pm 0.0061\,$days. They ascribed it
+peak in the periodogram at \replaced{$0.9985 \pm 0.0061\,$days}{$23.96\pm0.15\,$hr}. They ascribed it
 to their observing cadence, because of its close correspondence to the
@@
 network, \citet{raetz_yeti_2016} came to the conclusion that the
-$0.50\,{\rm d}$ signal was indeed from stellar rotation.  The TESS
+\replaced{$0.50\,{\rm d}$}{11.98$\,$hr} signal was indeed from stellar rotation.  The TESS
 data strongly support this conclusion.
@@
 
+\explain{Note: Citations to Parks et al 2014 and Stauffer et al 2015 were
+	added to the paragraph below}
 When physical explanations are not forthcoming, we often resort to
-taxonomy.  By searching the literature, we have found about a dozen
-light curves with similar morphologies to \ptfo, drawn from surveys of
-low-mass weak-lined T Tauri stars in regions including $\rho$ Oph,
-Upper~Sco, Taurus, and perhaps the Pleiades
-\citep{rebull_rotation_2016,david_transient_2017,stauffer_orbiting_2017,stauffer_rotevol_2018,rebull_usco_2018,rebull_rotation_2020}.
-These surveys were performed using K2 \citep{howell_k2_2014}.  We
-downloaded some of these light curves from MAST, opting for the
-EVEREST reductions \citep{luger_everest_2016,luger_update_2018}.  They
-are plotted in Figure~\ref{fig:brethren}.
-
+taxonomy.  
+By searching the literature, we found about a dozen light curves with
+similar morphologies to \ptfo, drawn from surveys of low-mass
+weak-lined T Tauri stars in regions including $\rho$ Oph, Upper~Sco,
+Taurus, and perhaps the Pleiades
+\citep{parks_periodic_2014,rebull_rotation_2016,david_transient_2017,stauffer_orbiting_2017,stauffer_rotevol_2018,rebull_usco_2018,rebull_rotation_2020}.
+The clearest matches came from K2 \citep{howell_k2_2014}, but a few
+analogs were potentially also found using 2MASS (YLW 1C and YLW 10C;
+\citealt{parks_periodic_2014}) and CoRoT
+(\citealt{stauffer_csi_2015}).  We downloaded a subset of the K2 light
+curves from MAST, opting for the EVEREST reductions
+\citep{luger_everest_2016,luger_update_2018}.  They are plotted in
+Figure~\ref{fig:brethren}.
+
+\explain{Note: added citation to Bodman et al 2017 in this paragraph.}
 These light curves have been phenomenologically classified as
@@
 at more advanced evolutionary disk stages than the ``dipper'' stars
-\citep{ansdell_young_2016,cody_manyfaceted_2018}.  The persistent and
+\citep{ansdell_young_2016,bodman_dippers_2017,cody_manyfaceted_2018}.  The persistent and
 transient flux dip stars all show angular dips that are cannot be
@@
     \citealt{rebull_usco_2018},~Figure~20).
-  \item The age is typically $\lesssim$ 100$\,$Myr.\footnote{ At
+  \item \replaced{The age is typically $\lesssim$ 100$\,$Myr.}{The
+    ages typically range between 5 and 100$\,$Myr.}\footnote{ At
     present, the oldest observed ``scallops'' are in the Pleaides
     \citep{rebull_rotation_2016}. One of these, EPIC 211013604, might
-    meet the ``persistent dip'' classification.  If so, it is also the
-    oldest known.} 
-  \item The light curves show shallow, angular dips, usually
-    superposed on large-amplitude smooth variability. The latter is
+    meet the ``persistent dip'' classification.  If so, it is \deleted{also} the
+    oldest known. \added{The lower limit comes from the lack of detectable
+   WISE infrared excesses.}} 
+  \item The light curves show shallow, angular dips, \deleted{usually}
+    superposed on large-amplitude smooth variability \added{at or near
+    1:1 synchronicity}. The \replaced{latter}{smooth variability} is
     interpreted as stellar rotation.
@@
   \item They are rare at a population level, with an occurrence rate
-    of $\lesssim 1\%$ of young M2-M5 stars \citep{rebull_usco_2018}.
+  \replaced{of $\lesssim 1\%$}{between 0.1--1\%}
+   of young M2-M5 stars \citep{rebull_usco_2018}.
 \end{enumerate}
@@
 Based on the lack of an infrared excess seen by \citet{yu_tests_2015},
-the primordial gas disks of both stars in \ptfo\ seem to be have been
-depleted.  This is consistent with the $8.5\pm1.2\,{\rm Myr}$ age of
-the 25$\,$Ori-1 group, and the rapid rate at which stars show lose
+the primordial \deleted{gas} disks of both stars in \ptfo\ seem to be have been
+depleted \added{of hot dust within the inner $\sim$1$\,$AU}.  This is consistent with the $8.5\pm1.2\,{\rm Myr}$ age of
+the 25$\,$Ori-1 group, and the rapid rate at which stars \deleted{show} lose
 their disks between 1 and 10$\,$Myr \citep[{\it
@@
 or
-{\it (vi)} dust clouds of plausible composition.
+{\it (vi)} dust clouds of \replaced{plausible composition}{particular compositions \citep[][Figure~10]{zhan_complex_2019}}.
 We also view the possibility of {\it (vii)} tidally disrupted
@@
 The explanations that are not yet ruled out include {\it (i)}
-transiting clumps of gas at the Keplerian corotation radius; {\it
+transiting clumps of \added{dust or }gas at the Keplerian co-rotation radius; {\it
 (ii)} transits of enshrouded protoplanets; {\it (iii)} occultations of
@@
 the transient and persistent-dip M-dwarfs, but also the ``scallop
-shell'' M-dwarfs \citep{stauffer_orbiting_2017}.  Despite this appeal,
+shell'' M-dwarfs \citep{stauffer_orbiting_2017}.  \replaced{Despite this appeal,
 the possibility of distinct mechanisms explaining these distinct
-variability classes remains open.
-
-The evolution of \ptfo\ over the past decade
-(Figure~\ref{fig:o_minus_c}) could offer important hints.
-Specifically, \ptfo's transition between having none, one, and
-multiple dips per cycle seems important.  It strains the ``enshrouded
+variability classes remains open.}{
+The model of starspot occultations by an optically thick disk \citep{zhan_complex_2019}
+is a qualitative inverse of the torus of circumstellar material:
+the difference is whether to place the ``blobs'' that
+provide the needed brightness contrast on the
+surface of the star, or at the co-rotation radius.
+}
+
+\replaced{The evolution of \ptfo\ over}{Observations of \ptfo\ from} the past decade
+could offer important hints.
+\deleted{Specifically, }\ptfo's transition between having none, one, and
+multiple dips per cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.  It strains the ``enshrouded
 protoplanet'' interpretation, because there are no known processes
-that cause a planet's orbital phase to jump.  The dips would then need
+that cause a planet's orbital phase to jump.  \deleted{The dips would then need
 to be caused by material that was somehow disrupted from the planet,
 but somehow remained co-orbital for an extended duration. This seems
-implausible.
+implausible.}
+\added{
+The H$\alpha$
+excess observed by \citet{johnskrull_h_2016} also seems like an
+important piece of evidence.
+The model with circumstellar optically thick hydrogen
+gas near the co-rotation radius easily explains this excess,
+while the starspot occultation model does not.
+}
+
+\added{Out of the available explanations, we therefore consider the
+idea of a clumpy torus of optically thick dust or gas near the co-rotation radius
+to be most plausible.
+There is theoretical support for
+the idea that accretion disks subject to strong magnetic
+fields could enter stalled states and become trapped near co-rotation
+\citep{dangelo_accretion_2012}.
+The concept of an ``evaporating dust wall'' within which all
+dust evaporates \citep{millan-gabet_circumstellar_2007,dullemond_inner_2010}
+will likely be an important first-order consideration in
+physical models of this scenario.
+The additional question of whether the opacity in fact comes from dust 
+or partially-ionized gas ({\it e.g.}, via gas continuum, atomic lines, or perhaps
+TiO molecular opacity) remains a topic for future exploration.
+}
 
@@
 
+\explain{Updated this paragraph to remove spot occultations by a disk as
+a ``preferred explanation''.}
 The physical mechanism that explains the transient and persistent flux
-dips is unresolved. Our preferred explanations include transiting
-clumps of gas at the Keplerian corotation radius, and occultations of
-starspots by a tenuous gas disk \citep[{\it
-e.g.},][]{stauffer_orbiting_2017,david_transient_2017,zhan_complex_2019}.
-The jumping orbital phase disfavors the explanation of an enshrouded,
-transiting protoplanet.  Though \ptfob\ may not be a planet, as we and
-others had hoped, understanding \ptfo\ and its analogs is a worthy
-problem.  It might even teach us about the birth environments of the
+dips is unresolved. Our preferred explanation is that a clumpy torus of dust or gas is near the Keplerian co-rotation radius, and periodically occults the star.
+The \added{chromaticity, out-of-eclipse variability, and} jumping orbital phase \replaced{disfavored}{of \ptfo\ all disfavor} the explanation of an enshrouded,
+transiting protoplanet.  Though \ptfo\ may not be a planet,\deleted{ as we and
+others had hoped,} understanding it and its analogs is a worthy
+problem\replaced{.}\added{that seems likely to provide insight into how gas
+and dust disperse from young protostellar disks.} \replaced{It might even}{Understanding
+this process would, in turn,} teach us about the birth environments of the
 majority of habitable-zone Earth-sized planets

### 2020-06-12 paper/collect.sh

@@ -10,7 +10,7 @@ cp ${pdir}PTFO_8-8695_transit_3sincosPorb_2sincosProt_phasefoldmap.pdf f3.pdf
 cp ../results/PTFO_8-8695_results/scene.pdf f4.pdf
 
 cp ../results/cluster_membership/hr.pdf f5a.pdf
-cp ../results/cluster_membership/astrometric_excess.pdf f5b.pdf
+cp ../results/cluster_membership/astrometric_excess_ruwe_talklabels.pdf f5b.pdf
 
 cp ../results/ephemeris/O_minus_C.pdf f6.pdf
 

### 2020-06-12 paper_review/response_20200605.txt
wrap up the referee response

@@
 
-We thank you for organizing the review of our paper, and also
-thank the referee for their helpful comments.  We have
-considered the referee's remarks carefully and revised our
-manuscript accordingly.
+We thank you for organizing the review of our paper, and also thank the referee
+for their helpful comments.  We have considered the referee's remarks carefully
+and revised our manuscript accordingly.
 
-Below, we have reproduced the relevant portions of the
-referee's report, alongside our responses. A list of changes is
-appended to the manuscript using the trackchanges AASTeX
-macros.
+Below, we have reproduced the relevant portions of the referee's report,
+alongside our responses. A list of changes is appended to the manuscript using
+the trackchanges AASTeX macros.
 
@@
 > Thank you for this correction. We have qualified the relevant sentence as the
-> youngest known *transiting* hot Jupiter. We have opted to not cite the V830
-> Tau or TAP 26 results, which we hope to see confirmed by independent groups.
+> youngest known *transiting* hot Jupiter, and added a footnote to the V830
+> Tau or TAP 26 results.
 
@@
 >
-> The reasons are that i) we wrote the paper primarily because of the TESS
-> data; ii) the multiple signals come from TESS data. From a story-telling
-> perspective, why would we worry about binarity if we didn't already know
-> there were multiple signals?
+> The reason is that the multiple signals come from TESS data processed at the
+> ``default'' level. From a story-telling perspective, why would we worry about
+> binarity if we didn't already know there were multiple photometric signals?
 
@@
 ----------
-> We're not sure whether or not it is a coincidence, but it seemed interesting
-> enough to mention.
+> Indeed, we are not sure whether or not it is a coincidence, but we
+> thought it was interesting enough to mention, even though we did not
+> have an interpretation.
 >
-> ..
+> It seems unlikely in our view that both stars are showing dips.
+> Our interpretation of the two smooth photometric signals is that one
+> star is rotating at 11.98 hr, and the other star is rotating at
+> 10.76 hr.
+>
+> All of the dips occur with a periodicity of 10.76 hr. (If we make
+> Figure 6 but instead phasing by P_long, it yields a mess of points).
+> Therefore all the dips are presumably associated with the star that
+> has that rotation period.
 
@@
 ----------
+> We have updated the text to include this point -- thank you for
+> making it.
 > 
+> We have also updated the text to reflect our hesitation about
+> whether or not the phenomenon is associated with binarity, primarily
+> because multiple photometric periods do not necessarily imply
+> binarity.
 
@@
 ----------
->
+> We have modified the last paragraphs of Section 5.4 to reflect an
+> updated preference that we came to after submission, for the clumpy
+> torus model over the spot occultation model. The reasons
+> for the preference are given in the text: 1) the spot occultation
+> model seems more geometrically contrived than we initially thought,
+> and 2) the Halpha excess seen by Johns-Krull+16 (and also in
+> unpublished observations by Stauffer of scallop shells) seems likely
+> to require hydrogen in orbit near the corotation radius.
 
@@
 We also modified the labels of the axes in Figure 5a for clarity.
+
+We have also added a few citations that we omitted in the original
+manuscript.

### 2020-06-12 paper/ms.tex
wrap up the referee response

@@
 We wish \ptfob\ were a planet. It would be quite exceptional. It would
-be the youngest known \added{transiting} hot Jupiter \citep{van_eyken_ptf_2012}, orbiting
+be the youngest known \added{transiting\footnote{Two younger hot
+Jupiters have been reported around V830 Tau and TAP 26
+\citep{donati_hot_2016,donati_hot_2017,yu_hot_2017}.}} hot Jupiter \citep{van_eyken_ptf_2012}, orbiting
 a T Tauri star in the Orion-OB1a cluster.  It would have the shortest
@@
 One was CVSO~35, which has a TESS light curve that varies by 2
-magnitudes.  CVSO~35 also shows a strong Wide-field Infrared Survey
-Exoplorer (WISE, \citealt{wright_WISE_2010}) IR excess as well as a
+magnitudes, and shows both a strong Wide-field Infrared Survey
+Exoplorer (WISE, \citealt{wright_WISE_2010}) IR excess and also a
 10$\mu$m silicate emission feature \citep{mauco_herschel_2018}.  The
@@
 discovered in $\rho$~Oph and Upper~Sco show two Lomb-Scargle periods,
-and so are apparently binaries
+and so \replaced{are apparently}{could be} binaries
 \citep[][Table~1]{stauffer_orbiting_2017}.  This is higher than the
 main-sequence companion fraction of ${\rm
-CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$
+CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$\added{, and so it
+could be an indication that the transient flux dip phenomenon is
+linked to stellar binarity}
 \citep{henry_solar_2006,duchene_stellar_2013,winters_solar_2019}.
-Low-mass pre-main-sequence stars however have been shown to companion
-fractions up to twice as high in dispersed clusters such as Upper~Sco
-and Taurus \citep{kraus_mapping_2008,kraus_mapping_2011}.  A
-high-resolution imaging survey would be interesting, to determine
-whether the transient flux dip stars truly have distinct
-population-level binarity properties relative to other young low-mass
-stars.
+\added{This fraction might be even more significant given the
+	selection bias against high
+mass ratio binaries, in which
+the second star would be too faint to detect a second period.}
+
+\added{However, the association between transient flux dips and binarity is
+not iron-clad.  Firstly, multiple photometric periods do not necessarily
+imply binarity. For instance, circumstellar material could exist
+around single stars inside, at, or wide of co-rotation.  Additionally, } low-mass pre-main-sequence stars
+\deleted{however} have been shown to \added{have} companion fractions
+up to twice as high in dispersed clusters such as Upper~Sco and Taurus
+\citep{kraus_mapping_2008,kraus_mapping_2011}.  A high-resolution
+imaging survey would be interesting, to determine whether the
+transient flux dip stars truly have distinct population-level binarity
+properties relative to other young low-mass stars.
 
@@
 
-The explanations that are not yet ruled out include {\it (i)}
+The explanations that \replaced{are not yet}{have not yet been} ruled out include {\it (i)}
 transiting clumps of \added{dust or }gas at the Keplerian co-rotation radius; {\it
 (ii)} transits of enshrouded protoplanets; {\it (iii)} occultations of
-starspots by an optically thick disk.  The first and last explanations
+\replaced{starspots}{hot and cold spots on the stellar surface} by an optically thick disk.  The first and last explanations
 have added appeal because they are flexible enough to explain not only
@@
 variability classes remains open.}{
-The model of starspot occultations by an optically thick disk \citep{zhan_complex_2019}
-is a qualitative inverse of the torus of circumstellar material:
-the difference is whether to place the ``blobs'' that
-provide the needed brightness contrast on the
-surface of the star, or at the co-rotation radius.
+The model of hot and coldspot occultations by an optically thick disk
+\citep{zhan_complex_2019} is a qualitative inverse of the torus of
+circumstellar material: the difference is whether to place the
+``blobs'' that provide the needed brightness contrast on the surface
+of the star, or at the co-rotation radius.  In the spot occultation
+model, the observing geometry would require an inner disk to be
+projected into a narrow band across the stellar surface, in order to
+produce variations with sufficiently short timescales.  The outer
+disk would need either to not be present, or to be misaligned enough,
+and narrow enough, that it did not fall along the line of sight. These
+combined requirements seem geometrically contrived.
 }
@@
 \replaced{The evolution of \ptfo\ over}{Observations of \ptfo\ from} the past decade
-could offer important hints.
+could offer additional hints.
 \deleted{Specifically, }\ptfo's transition between having none, one, and
@@
 The model with circumstellar optically thick hydrogen
-gas near the co-rotation radius easily explains this excess,
-while the starspot occultation model does not.
+gas near the co-rotation radius readily explains this excess.
 }

### 2020-06-12 paper_review/response_20200605.txt
fix column widths

--- 2020-06-12 paper_review/response_20200605.txt
@@
 We thank you for organizing the review of our paper, and also
-thank the referee for their helpful comments.  We have
-considered the referee's remarks carefully and revised our
-manuscript accordingly.
+thank the referee for their helpful comments.  We have considered
+the referee's remarks carefully and revised our manuscript
+accordingly.
 
-Below, we have reproduced the relevant portions of the
-referee's report, alongside our responses. A list of changes is
-appended to the manuscript using the trackchanges AASTeX
-macros.
+Below, we have reproduced the relevant portions of the referee's
+report, alongside our responses. A list of changes is appended to
+the manuscript using the trackchanges AASTeX macros.
 
@@
 ----------
-Overall: It would help if the authors used hours or days throughout the paper
-in a consistent manner. The period is referred to as 0.5d and 11.98h just a
-couple paragraphs apart (e.g. in Section 5.1). The abstract uses hours, most
-plots use days.
+Overall: It would help if the authors used hours or days
+throughout the paper in a consistent manner. The period is
+referred to as 0.5d and 11.98h just a couple paragraphs apart
+(e.g. in Section 5.1). The abstract uses hours, most plots use
+days.
 
@@
 ----------
-> Throughout the manuscript, we have shifted the units wherever possible to
-> hours.
+> Throughout the manuscript, we have shifted the units wherever
+> possible to hours.
 >
-> Exceptions include the ephemeris (which is quoted in days), and the periods
-> listed in Figures 3 and 6, which are quoted in days. The former is standard
-> convention. The latter was because we felt it important to list the periods
-> used to generate the figures at a higher precision, and quoting higher
-> precision in units of hours seems atypical in a field in which days are more
+> Exceptions include the ephemeris (which is quoted in days), and
+> the periods listed in Figures 3 and 6, which are quoted in
+> days. The former is standard convention. The latter was because
+> we felt it important to list the periods used to generate the
+> figures at a higher precision, and quoting higher precision in
+> units of hours seems atypical in a field in which days are more
 > typically seen.
@@
 
-Introduction: The introduction takes a light/jovial tone, e.g., did they
-actually wish it were a planet, or is it more that they aimed to confirm it and
-found the opposite? I would normally suggest making more formal. However, I
-enjoyed reading it. I'm sorry the authors ruined their own favorite planet.
+Introduction: The introduction takes a light/jovial tone, e.g.,
+did they actually wish it were a planet, or is it more that they
+aimed to confirm it and found the opposite? I would normally
+suggest making more formal. However, I enjoyed reading it. I'm
+sorry the authors ruined their own favorite planet.
 
@@
 ----------
-> So are we!  Our aim was to clarify what could be happening in the system, and
-> hopefully this contribution is valuable for that.  When a planetary door
-> closes, a window on young star variability opens...
+> So are we!  Our aim was to clarify what could be happening in
+> the system, and hopefully this contribution is valuable for
+> that.  When a planetary door closes, a window on young star
+> variability opens...
 
@@
 
-Introduction: I do not think this is youngest hot Jupiter. That should be true
-just for transiting systems, and it was the youngest at discovery, but the two
-RV planets from ESPaDOnS are younger (Donati et al. 2016 and Yu et al. 2017).
-Both are hot Jupiters and in Taurus. I believe the age of Tap 26 is more
-ambiguous due to age spreads (multiple groups?) in Taurus, but the age of V830
-Tau is certainly <3 Myr.
+Introduction: I do not think this is youngest hot Jupiter. That
+should be true just for transiting systems, and it was the
+youngest at discovery, but the two RV planets from ESPaDOnS are
+younger (Donati et al. 2016 and Yu et al. 2017).  Both are hot
+Jupiters and in Taurus. I believe the age of Tap 26 is more
+ambiguous due to age spreads (multiple groups?) in Taurus, but
+the age of V830 Tau is certainly <3 Myr.
 
@@
 
-> Thank you for this correction. We have qualified the relevant sentence as the
-> youngest known *transiting* hot Jupiter. We have opted to not cite the V830
-> Tau or TAP 26 results, which we hope to see confirmed by independent groups.
+> Thank you for this correction. We have qualified the relevant
+> sentence as the youngest known *transiting* hot Jupiter, and
+> added a footnote to the V830 Tau or TAP 26 results.
 
@@
 
-Introduction: I also think it wasn't the first H-alpha detection. That belongs
-to LkCa15, which was imaged in H-alpha in 2015 (Sallum et al. 2015). Although
-that planet might not exist, so this comment is probably OK. It would still be
-the first close-in planet either way. I'd suggest weakening the statement
-slightly to err on the side of caution.
+Introduction: I also think it wasn't the first H-alpha detection.
+That belongs to LkCa15, which was imaged in H-alpha in 2015
+(Sallum et al. 2015). Although that planet might not exist, so
+this comment is probably OK. It would still be the first close-in
+planet either way. I'd suggest weakening the statement slightly
+to err on the side of caution.
 
@@
 ----------
-> We have changed the wording to err on the side of caution, and qualify it as
-> the first direct detection of Hα emission from a close-in planet.
+> We have changed the wording to err on the side of caution, and
+> qualify it as the first direct detection of Hα emission from a
+> close-in planet.
 
@@
 
-Introduction: Barnes et al. (2013) claimed that the planet and star precessed
-on a period of 300-600 days, and that this could somewhat fit variable depths
-and missing transits. The Ciardi et al. 2015 paper uses this as their model, in
-addition to gravity darkening as mentioned in the current introduction. I
-suggest mentioning the precession of the planet in the introduction, as many
-papers used the precession model or at least discussed it. The line "...the
-gravity-darkened star precessed..." seems to allude to this, but it's unclear
-from the rest of the text.
+Introduction: Barnes et al. (2013) claimed that the planet and
+star precessed on a period of 300-600 days, and that this could
+somewhat fit variable depths and missing transits. The Ciardi et
+al. 2015 paper uses this as their model, in addition to gravity
+darkening as mentioned in the current introduction. I suggest
+mentioning the precession of the planet in the introduction, as
+many papers used the precession model or at least discussed it.
+The line "...the gravity-darkened star precessed..." seems to
+allude to this, but it's unclear from the rest of the text.
 
@@
 ----------
-> We have changed the wording of the last sentence of the first paragraph in
-> the introduction to explicitly use the word "precess", rather than "torque".
+> We have changed the wording of the last sentence of the first
+> paragraph in the introduction to explicitly use the word
+> "precess", rather than "torque".
 
@@
 
-Section 2: The first paragraph discusses the full frame images, while the next
-section discusses the 2m data. Why did the authors use the FFIs when they have
-the (higher quality) 2m data?
+Section 2: The first paragraph discusses the full frame images,
+while the next section discusses the 2m data. Why did the authors
+use the FFIs when they have the (higher quality) 2m data?
 
@@
 ----------
-> For the essential analysis of the study (i.e., extracting and fitting the
-> long and short signals), we used the 2m SPOC lightcurve.
+> For the essential analysis of the study (i.e., extracting and
+> fitting the long and short signals), we used the 2m SPOC
+> lightcurve.
 >
-> We primarily mention our use of the FFI lightcurves in Section 2 because of
-> the systematic jump mentioned at BJD 2458488.3, which was eventually excluded
-> from further analysis. We wanted to be sure though that this jump was truly
-> an instrumental systematic, and not a software systematic. Checking
-> lightcurves from independent pipelines meets this goal.
+> We primarily mention our use of the FFI lightcurves in Section
+> 2 because of the systematic jump mentioned at BJD 2458488.3,
+> which was eventually excluded from further analysis. We wanted
+> to be sure though that this jump was truly an instrumental
+> systematic, and not a software systematic. Checking lightcurves
+> from independent pipelines meets this goal.
 >
-> A secondary reason to introduce the concept of FFIs in Section 2 is that they
-> are used to measure the variability amplitudes of the other members of 25
-> Ori-1 in Section 4.3.
+> A secondary reason to introduce the concept of FFIs in Section
+> 2 is that they are used to measure the variability amplitudes
+> of the other members of 25 Ori-1 in Section 4.3.
 
@@
 
-Section 2.2.2: "These details concerning the group membership for one object
-may seem excessive to those accustomed to the simple distinction between "young
-cluster members" and "old field stars"." I suggest removing this. It's not
-adding anything and sounds like the authors are making assumptions about the
-reader. The membership of this object is essential for an age assessment, and
-can/should be updated given the new Gaia data. Since most of the object class
-in question are at similar ages, this directly relates to the study. If it had
-turned out to be ~100 Myr, we would all be more skeptical (although ~100 Myr
-stars like this are known, they are more rare).
+Section 2.2.2: "These details concerning the group membership for
+one object may seem excessive to those accustomed to the simple
+distinction between "young cluster members" and "old field
+stars"." I suggest removing this. It's not adding anything and
+sounds like the authors are making assumptions about the reader.
+The membership of this object is essential for an age assessment,
+and can/should be updated given the new Gaia data. Since most of
+the object class in question are at similar ages, this directly
+relates to the study. If it had turned out to be ~100 Myr, we
+would all be more skeptical (although ~100 Myr stars like this
+are known, they are more rare).
 
@@
 
-Section 3/4: This deals mostly with identification of the second star. However,
-the second star appears to be an assumption of section 3.2. If the periods come
-from two stars that could be separated (e.g., wide binaries), it would have
-been best to do so rather than try to model both by including fs and fl terms.
-So it is odd that the binary search is after the light curve model in the
-paper. It might make more narrative sense to move the light curve model to
-after section 4, since 4 motivates 3.2.
+Section 3/4: This deals mostly with identification of the second
+star. However, the second star appears to be an assumption of
+section 3.2. If the periods come from two stars that could be
+separated (e.g., wide binaries), it would have been best to do so
+rather than try to model both by including fs and fl terms.  So
+it is odd that the binary search is after the light curve model
+in the paper. It might make more narrative sense to move the
+light curve model to after section 4, since 4 motivates 3.2.
 
@@
 ----------
-> While we see the referee's point, we prefer the narrative structure of
-> presenting the TESS data first, and getting into considerations of stellar
-> binarity, both resolved and unresolved, afterwards.
+> While we see the referee's point, we prefer the narrative
+> structure of presenting the TESS data first, and getting into
+> considerations of stellar binarity, both resolved and
+> unresolved, afterwards.
 >
-> The reasons are that i) we wrote the paper primarily because of the TESS
-> data; ii) the multiple signals come from TESS data. From a story-telling
-> perspective, why would we worry about binarity if we didn't already know
-> there were multiple signals?
+> The reason is that the multiple signals come from TESS data
+> processed at the ``default'' level. From a story-telling
+> perspective, why would we worry about binarity if we didn't
+> already know there were multiple photometric signals?
 
@@
 
-Section 4.3: My biggest concern with this analysis is reliance on the chi^2
-from Gaia to assess binarity. This is not the suggested path from Gaia, as it
-is subject to a noise model Gaia now knows has issues. For example, nearly 100%
-of mid/late M dwarfs have a high chi^2 (and a large excess astrometric noise),
-but clearly they are not all binaries. This is (partially) because of a
-color-term in the PSF model that was not initially understood. Similarly,
-almost all bright stars have a large RUWE, as do the bluest stars. The authors
-correctly use RUWE, the suggested method, but the RUWE analysis is less
-convincing and clearly the authors are leaning on the chi^2.
+Section 4.3: My biggest concern with this analysis is reliance on
+the chi^2 from Gaia to assess binarity. This is not the suggested
+path from Gaia, as it is subject to a noise model Gaia now knows
+has issues. For example, nearly 100% of mid/late M dwarfs have a
+high chi^2 (and a large excess astrometric noise), but clearly
+they are not all binaries. This is (partially) because of a
+color-term in the PSF model that was not initially understood.
+Similarly, almost all bright stars have a large RUWE, as do the
+bluest stars. The authors correctly use RUWE, the suggested
+method, but the RUWE analysis is less convincing and clearly the
+authors are leaning on the chi^2.
 
-A good reference on this would be Belokurov et al. 2020, which shows what kinds
-of binaries RUWE can and cannot detect. This binary might simply be too tight
-to create an astrometric signal (in which case the RUWE makes sense). The
-typical cutoff for binarity is a RUWE of ~1.4, see Rizzuto et al. 2018, Ziegler
-et al. 2019, Pearce et al. 2020, among many others.  I suggest removing the
-chi^2 analysis. If the authors want to keep it, they need to argue why it is a
-more meaningful metric than RUWE, and attempt to account for color and other
-parameters that impact chi^2.
+A good reference on this would be Belokurov et al. 2020, which
+shows what kinds of binaries RUWE can and cannot detect. This
+binary might simply be too tight to create an astrometric signal
+(in which case the RUWE makes sense). The typical cutoff for
+binarity is a RUWE of ~1.4, see Rizzuto et al. 2018, Ziegler et
+al. 2019, Pearce et al. 2020, among many others.  I suggest
+removing the chi^2 analysis. If the authors want to keep it, they
+need to argue why it is a more meaningful metric than RUWE, and
+attempt to account for color and other parameters that impact
+chi^2.
 
@@
 ----------
-> We've done as suggested, and replaced this sub-plot with RUWE instead of
-> reduced chi^2. The text in Section 4.3 and captions have been updated
-> accordingly, and chi^2 is now mostly ignored.
+> We've done as suggested, and replaced this sub-plot with RUWE
+> instead of reduced chi^2. The text in Section 4.3 and captions
+> have been updated accordingly, and chi^2 is now mostly ignored.
 
@@
 
-Section 4.4:
-Need a period between velocities and Further.
+Section 4.4: Need a period between velocities and Further.
 
@@
 
-Section 5.2: That clear offset at (Pl-Ps)/Ps doesn't seem like a coincidence,
-and the paper reads like the authors agree, but couldn't explain it. I'm not
-sure I can help much here, although it is interesting. Could this be that both
-stars are dippers and one turns off while the other turns on by coincidence?
-Spots and hot-spot clusters can shift in phase as the magnetic fields realign,
-which creates a new spot cluster at the same period but a different phase. If
-it is circumstellar material, a tight binary could be shepherding the material,
-but I think the systems would have to be very tight for that, and it could not
-last for such a long time.
+Section 5.2: That clear offset at (Pl-Ps)/Ps doesn't seem like a
+coincidence, and the paper reads like the authors agree, but
+couldn't explain it. I'm not sure I can help much here, although
+it is interesting. Could this be that both stars are dippers and
+one turns off while the other turns on by coincidence?  Spots and
+hot-spot clusters can shift in phase as the magnetic fields
+realign, which creates a new spot cluster at the same period but
+a different phase. If it is circumstellar material, a tight
+binary could be shepherding the material, but I think the systems
+would have to be very tight for that, and it could not last for
+such a long time.
 
-I think this can be assessed statistically; given the scatter in those points,
-the odds of lining up to any given line is small. But without a physical
-explanation this is hard to test. Probably this requires looking at many
-dippers as an ensemble. It is not required for the primary analysis of this
-paper, so I will defer to the authors if they want to dig deeper on this.
+I think this can be assessed statistically; given the scatter in
+those points, the odds of lining up to any given line is small.
+But without a physical explanation this is hard to test. Probably
+this requires looking at many dippers as an ensemble. It is not
+required for the primary analysis of this paper, so I will defer
+to the authors if they want to dig deeper on this.
 
@@
 ----------
-> We're not sure whether or not it is a coincidence, but it seemed interesting
-> enough to mention.
+> Indeed, we are not sure whether or not it is a coincidence, but
+> we thought it was interesting enough to mention, even though we
+> do not have an interpretation.
 >
-> ..
+> It seems unlikely in our view that both stars are showing dips.
+> Our interpretation of the two smooth photometric signals is
+> that one star is rotating at 11.98 hr, and the other star is
+> rotating at 10.76 hr.
+>
+> All of the dips occur with a periodicity of 10.76 hr. (If we
+> make Figure 6 but instead phasing by P_long, it yields a mess
+> of points).  Therefore all the dips are presumably associated
+> with the star that has that rotation period.
 
@@
 
-Section 5.3.2: The difference between the fraction of transient flux dip stars
-and the binary fraction is more significant than it appears to the authors.
-High mass ratio binaries would not be included, as they are too faint to detect
-the second period. Further, at the distance of Upper Sco (~150pc) and assuming
-Stauffer et al. 2017 would notice/separate two stars wider than a K2 pixel
-(4"), the presence of two periods would not include binaries wider than 600pc.
-Eyeballing the Winters et al. 2019 distribution suggests ~20% of binaries are
-beyond this. Stauffer et al. 2017 also probably checked against existing
-imaging data for obvious binaries down to 1-2", so likely even more are getting
-cut.
+Section 5.3.2: The difference between the fraction of transient
+flux dip stars and the binary fraction is more significant than
+it appears to the authors.  High mass ratio binaries would not be
+included, as they are too faint to detect the second period.
+Further, at the distance of Upper Sco (~150pc) and assuming
+Stauffer et al. 2017 would notice/separate two stars wider than a
+K2 pixel (4"), the presence of two periods would not include
+binaries wider than 600pc.  Eyeballing the Winters et al. 2019
+distribution suggests ~20% of binaries are beyond this. Stauffer
+et al. 2017 also probably checked against existing imaging data
+for obvious binaries down to 1-2", so likely even more are
+getting cut.
 
@@
 ----------
+> We have updated the text to include this point -- thank you for
+> making it.
 > 
+> We have also updated the text to reflect our hesitation about
+> whether or not the phenomenon is associated with binarity,
+> primarily because multiple photometric periods do not
+> necessarily imply binarity.
 
@@
 
-Section 5.4: The authors conclude this subsection with a reasonable argument
-that this isn't a protoplanet. Does this same argument not rule out gas/debris
-at the Keplerian coronation radius? What could maintain such clouds for
-years/decades and then have them jump phase but at the same period? It's hard
-to do this if it is not somehow physically attached to the star in some what
-(e.g., spots can do this).
+Section 5.4: The authors conclude this subsection with a
+reasonable argument that this isn't a protoplanet. Does this same
+argument not rule out gas/debris at the Keplerian coronation
+radius? What could maintain such clouds for years/decades and
+then have them jump phase but at the same period? It's hard to do
+this if it is not somehow physically attached to the star in some
+what (e.g., spots can do this).
 
@@
 ----------
+> We have modified the last paragraphs of Section 5.4 to reflect
+> an updated preference that we came to after submission, for the
+> clumpy torus model over the spot occultation model. The reasons
+> for the preference are given in the text: 1) the spot
+> occultation model seems more geometrically contrived than we
+> initially thought, and 2) the Halpha excess seen by
+> Johns-Krull+16 (and also in unpublished observations by
+> Stauffer of scallop shells) seems likely to require hydrogen in
+> orbit near the corotation radius.
 >
+> In response to the question of: "why can the shifting phase be
+> explained more easily by dust/gas clumps than the
+> protoplanet?", we now refer D'Angelo and Spruit 2012 (see in
+> particular their Section 1), which discusses the possibility of
+> cyclic accretion, with periods of "stalled" mass accumulation
+> outside the corotation radius alternating with accreting
+> phases.
 
@@
 
-A few additional changes are noted in the "List of Changes" appended at the end
-of the manuscript.
+A few additional changes are noted in the "List of Changes"
+appended at the end of the manuscript.
 
-We modified the text throughout to clarify that Stauffer+ 2017 favored the
-explanation of a clumpy torus of gas and dust at the Keplerian co-rotation
-radius, rather than our initial (incomplete) description of "clouds of gas" at
-the same location.
+We modified the text throughout to clarify that Stauffer+ 2017
+favored the explanation of a clumpy torus of gas and dust at the
+Keplerian co-rotation radius, rather than our initial
+(incomplete) description of "clouds of gas" at the same location.
 
 We also modified the labels of the axes in Figure 5a for clarity.
+
+We have also added a few citations that we omitted in the
+original manuscript.

### 2020-06-12 paper/ms.tex
fix column widths

@@
 Jupiters have been reported around V830 Tau and TAP 26
-\citep{donati_hot_2016,donati_hot_2017,yu_hot_2017}.}} hot Jupiter \citep{van_eyken_ptf_2012}, orbiting
-a T Tauri star in the Orion-OB1a cluster.  It would have the shortest
-orbital period of any hot Jupiter.  With such a short period, it would
-probably be filling its Roche lobe, and actively losing mass to its
-host star.  Not only that, but the rapidly-rotating host star is
-probably oblate enough to \replaced{torque the planet's orbit into and out of
-the transiting configuration}{cause the planet's orbit to precess into and out of the 
+\citep{donati_hot_2016,donati_hot_2017,yu_hot_2017}.}} hot Jupiter
+\citep{van_eyken_ptf_2012}, orbiting a T Tauri star in the Orion-OB1a
+cluster.  It would have the shortest orbital period of any hot
+Jupiter.  With such a short period, it would probably be filling its
+Roche lobe, and actively losing mass to its host star.  Not only that,
+but the rapidly-rotating host star is probably oblate enough to
+\replaced{torque the planet's orbit into and out of the transiting
+configuration}{cause the planet's orbit to precess into and out of the
 transiting configuration} on a timescale of years
@@
 Another first would be the direct detection of H$\alpha$ emission from
-\replaced{the planet itself}{a close-in planet} \citep{johnskrull_h_2016}.  In addition to the
-chromospheric H$\alpha$ emisson, it seems that there is an additional
-H$\alpha$ emission with radial velocity variations in phase with the
-planetary orbit.  The average velocity width of the excess H$\alpha$
-emission is 87$\,$km$\,$s$^{-1}$, and its equivalent width is 70-80\%
-that of the stellar chromosphere \citep{johnskrull_h_2016}.  The
-proposed explanation is that the emission is from hot material flowing
-away from the planet \citep{johnskrull_h_2016}.
+\replaced{the planet itself}{a close-in planet}
+\citep{johnskrull_h_2016}.  In addition to the chromospheric H$\alpha$
+emisson, it seems that there is an additional H$\alpha$ emission with
+radial velocity variations in phase with the planetary orbit.  The
+average velocity width of the excess H$\alpha$ emission is
+87$\,$km$\,$s$^{-1}$, and its equivalent width is 70-80\% that of the
+stellar chromosphere \citep{johnskrull_h_2016}.  The proposed
+explanation is that the emission is from hot material flowing away
+from the planet \citep{johnskrull_h_2016}.
 
@@
 In addition, the sublimation times for dust grains of plausible
-composition \replaced{are}{could be} quite short \citep{zhan_complex_2019}.
+composition \replaced{are}{could be} quite short
+\citep{zhan_complex_2019}.
 
@@
 and \citet{david_transient_2017} to explain this novel class of
-variable stars is that a \replaced{circumstellar cloud of gas is orbiting near
-the co-rotation radius}{clumpy torus of dust and gas orbits near the
-Keplerian co-rotation radius}.  To this point, though, it has not been clear
-if this explanation applies to \ptfo, because the determination of the
-stellar rotation period has been somewhat ambiguous
+variable stars is that a \replaced{circumstellar cloud of gas is
+orbiting near the co-rotation radius}{clumpy torus of dust and gas
+orbits near the Keplerian co-rotation radius}.  To this point, though,
+it has not been clear if this explanation applies to \ptfo, because
+the determination of the stellar rotation period has been somewhat
+ambiguous
 \citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.
@@
 
-\replaced{These details concerning the group membership for one object may seem
-excessive to those accustomed to the simple distinction between
-``young cluster members'' and ``old field stars''.  Although all
-members of the Orion complex are indeed young relative to the field,}
-{While all members of the Orion complex are young relative to the field,}
-these \replaced{details}{considerations} are essential for assessing \deleted{the} photometric evidence for
-the binarity of \ptfo, because of the degeneracy between stellar
-luminosity and age for pre-main-sequence stars.  Having a clean sample
-of reference stars that are tightly associated with \ptfo\ --- both
-spatially and kinematically --- minimizes contamination not only from
-field stars, but also from older and younger members of the Orion
-complex.
+\replaced{These details concerning the group membership for one object
+may seem excessive to those accustomed to the simple distinction
+between ``young cluster members'' and ``old field stars''.  Although
+all members of the Orion complex are indeed young relative to the
+field,} {While all members of the Orion complex are young relative to
+the field,} these \replaced{details}{considerations} are essential for
+assessing \deleted{the} photometric evidence for the binarity of
+\ptfo, because of the degeneracy between stellar luminosity and age
+for pre-main-sequence stars.  Having a clean sample of reference stars
+that are tightly associated with \ptfo\ --- both spatially and
+kinematically --- minimizes contamination not only from field stars,
+but also from older and younger members of the Orion complex.
 
@@
 \citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
-The tallest peak occurs at \replaced{0.499\,d (11.98$\,$hr)}{11.98$\,$hr} and a second strong
-peak occurs at \replaced{0.448\,d (10.76$\,$hr)}{10.76$\,$hr}. We will refer to these two
-periods as the ``longer period'' $P_{\rm \ell}$ and the ``shorter
+The tallest peak occurs at \replaced{0.499\,d
+(11.98$\,$hr)}{11.98$\,$hr} and a second strong peak occurs at
+\replaced{0.448\,d (10.76$\,$hr)}{10.76$\,$hr}. We will refer to these
+two periods as the ``longer period'' $P_{\rm \ell}$ and the ``shorter
 period'' $P_{\rm s}$.  Lower-power harmonics of both signals are also
@@
 
-We fitted a model to the light curve consisting of a linear combination
-of Fourier modes with periods $P_{\rm s}$ and $P_{\rm \ell}$, as well
-as a number of harmonics chosen as described below. To try accounting
-for the dips, we also added an analytic transit model with period
-$P_{\rm s}$.  Symbolically, the total flux $f$ is given as
+We fitted a model to the light curve consisting of a linear
+combination of Fourier modes with periods $P_{\rm s}$ and $P_{\rm
+\ell}$, as well as a number of harmonics chosen as described below. To
+try accounting for the dips, we also added an analytic transit model
+with period $P_{\rm s}$.  Symbolically, the total flux $f$ is given as
 \begin{equation}
@@
 
-\explain{All paragraphs in this section have been heavily
-	reworded to emphasize RUWE over $\chi_{\rm red}^2$, as requested by the
-	referee.}
+\explain{All paragraphs in this section have been heavily reworded to
+emphasize RUWE over $\chi_{\rm red}^2$, as requested by the referee.}
 A separate line of evidence for binarity is the Gaia DR2 astrometry.
@@
 astrometric measurements, and 5 free parameters, and therefore 116
-degrees of freedom. The reduced $\chi^2$ is 2.80.  While the majority of
-stars with comparable brightness in Gaia do not show such poor
+degrees of freedom. The reduced $\chi^2$ is 2.80.  While the majority
+of stars with comparable brightness in Gaia do not show such poor
 goodness-of-fit\added{, many M dwarfs do, because of a color-dependent
-term in the mission's PSF model} \citep[see][Appendix A]{lindegren_gaiasoln_2018}.
-The reduced $\chi^2$ has been
-``renormalized'' to account for this, yielding a modified
-statistic called the renormalized unit weight error
-(RUWE\footnote{ See the Gaia DPAC technical note
-	GAIA-C3-TN-LU-LL-124-01,
+term in the mission's PSF model} \citep[see][Appendix
+A]{lindegren_gaiasoln_2018}.  The reduced $\chi^2$ has been
+``renormalized'' to account for this, yielding a modified statistic
+called the renormalized unit weight error (RUWE\footnote{ See the Gaia
+DPAC technical note GAIA-C3-TN-LU-LL-124-01,
 \url{http://www.rssd.esa.int/doc_fetch.php?id=3757412}, accessed
@@
 
-We acquired the RUWE for each of the
-149 members in 25$\,$Ori-1 identified by
-identified by \citet{kounkel_apogee2_2018}.
-We then queried the
-CDIPS light curve database at MAST \citep{bouma_cluster_2019} to find
-the subset of members that were at least as variable as \ptfo.  We
-measured the variability amplitude by taking the difference between
-the $95^{\rm th}$ and $5^{\rm th}$ percentiles of the flux
-measurements.  This yielded 30 stars of equal or greater variability.
-The lower panel of Figure~\ref{fig:gaia} shows the RUWE of
-these stars as
-a function of stellar brightness.  \ptfo\ is at the 93$^{\rm
-	rd}$ percentile of equally-variable stars within the
-25$\,$Ori-1 group.  
-Two of the thirty stars
-with variability amplitudes greater than 9.7\% showed higher RUWE.
-One was CVSO~35, which has a TESS light curve that varies by 2
-magnitudes, and shows both a strong Wide-field Infrared Survey
-Exoplorer (WISE, \citealt{wright_WISE_2010}) IR excess and also a
-10$\mu$m silicate emission feature \citep{mauco_herschel_2018}.  The
-other star was GAIA DR2 3222210363837122048.
-
-Potential explanations for an elevated RUWE include
-photometric variability and unresolved stellar binarity \citep[{\it
-e.g.},][]{rizzuto_ZEIT8_2018,belokurov_unresolved_2020}.  
-If photometric variability were the root cause, we would expect stars with
-similar brightness and color in the same kinematic group of Orion to show
-similar astrometric excesses, because the majority of young stars are
-highly variable.
-However
-relative to other M-dwarf group members with
-comparable brightnesses and variability characteristics, \ptfo\ still
-stands out by virtue of its failure to conform to a single-star
-astrometric model. This supports the interpretation that \ptfo\ is a
-binary star.
+We acquired the RUWE for each of the 149 members in 25$\,$Ori-1
+identified by identified by \citet{kounkel_apogee2_2018}.  We then
+queried the CDIPS light curve database at MAST
+\citep{bouma_cluster_2019} to find the subset of members that were at
+least as variable as \ptfo.  We measured the variability amplitude by
+taking the difference between the $95^{\rm th}$ and $5^{\rm th}$
+percentiles of the flux measurements.  This yielded 30 stars of equal
+or greater variability.  The lower panel of Figure~\ref{fig:gaia}
+shows the RUWE of these stars as a function of stellar brightness.
+\ptfo\ is at the 93$^{\rm rd}$ percentile of equally-variable stars
+within the 25$\,$Ori-1 group.  Two of the thirty stars with
+variability amplitudes greater than 9.7\% showed higher RUWE.  One was
+CVSO~35, which has a TESS light curve that varies by 2 magnitudes, and
+shows both a strong Wide-field Infrared Survey Exoplorer (WISE,
+\citealt{wright_WISE_2010}) IR excess and also a 10$\mu$m silicate
+emission feature \citep{mauco_herschel_2018}.  The other star was GAIA
+DR2 3222210363837122048.
+
+Potential explanations for an elevated RUWE include photometric
+variability and unresolved stellar binarity \citep[{\it
+e.g.},][]{rizzuto_ZEIT8_2018,belokurov_unresolved_2020}.  If
+photometric variability were the root cause, we would expect stars
+with similar brightness and color in the same kinematic group of Orion
+to show similar astrometric excesses, because the majority of young
+stars are highly variable.  However relative to other M-dwarf group
+members with comparable brightnesses and variability characteristics,
+\ptfo\ still stands out by virtue of its failure to conform to a
+single-star astrometric model. This supports the interpretation that
+\ptfo\ is a binary star.
 
@@
 longer-period signal ({\it e.g.}, their Figure~7), in the form of a
-peak in the periodogram at \replaced{$0.9985 \pm 0.0061\,$days}{$23.96\pm0.15\,$hr}. They ascribed it
-to their observing cadence, because of its close correspondence to the
-sidereal day.  Our pixel-level analysis showed that the signal is
-specific to only pixels near \ptfo, and no other pixels.  We therefore
-conclude that the signal is not an artifact of systematic errors.
+peak in the periodogram at \replaced{$0.9985 \pm
+0.0061\,$days}{$23.96\pm0.15\,$hr}. They ascribed it to their
+observing cadence, because of its close correspondence to the sidereal
+day.  Our pixel-level analysis showed that the signal is specific to
+only pixels near \ptfo, and no other pixels.  We therefore conclude
+that the signal is not an artifact of systematic errors.
 
@@
 network, \citet{raetz_yeti_2016} came to the conclusion that the
-\replaced{$0.50\,{\rm d}$}{11.98$\,$hr} signal was indeed from stellar rotation.  The TESS
-data strongly support this conclusion.
+\replaced{$0.50\,{\rm d}$}{11.98$\,$hr} signal was indeed from stellar
+rotation.  The TESS data strongly support this conclusion.
 
@@
 
-\explain{Note: Citations to Parks et al 2014 and Stauffer et al 2015 were
-	added to the paragraph below}
-When physical explanations are not forthcoming, we often resort to
-taxonomy.  
-By searching the literature, we found about a dozen light curves with
-similar morphologies to \ptfo, drawn from surveys of low-mass
-weak-lined T Tauri stars in regions including $\rho$ Oph, Upper~Sco,
-Taurus, and perhaps the Pleiades
+\explain{Note: Citations to Parks et al 2014 and Stauffer et al 2015
+were added to the paragraph below} When physical explanations are not
+forthcoming, we often resort to taxonomy.  By searching the
+literature, we found about a dozen light curves with similar
+morphologies to \ptfo, drawn from surveys of low-mass weak-lined T
+Tauri stars in regions including $\rho$ Oph, Upper~Sco, Taurus, and
+perhaps the Pleiades
 \citep{parks_periodic_2014,rebull_rotation_2016,david_transient_2017,stauffer_orbiting_2017,stauffer_rotevol_2018,rebull_usco_2018,rebull_rotation_2020}.
@@
 at more advanced evolutionary disk stages than the ``dipper'' stars
-\citep{ansdell_young_2016,bodman_dippers_2017,cody_manyfaceted_2018}.  The persistent and
-transient flux dip stars all show angular dips that are cannot be
-explained as the effects of starspots.  These stars typically have the
-following things in common:
+\citep{ansdell_young_2016,bodman_dippers_2017,cody_manyfaceted_2018}.
+The persistent and transient flux dip stars all show angular dips that
+are cannot be explained as the effects of starspots.  These stars
+typically have the following things in common:
 \begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
@@
 \added{This fraction might be even more significant given the
-	selection bias against high
-mass ratio binaries, in which
-the second star would be too faint to detect a second period.}
-
-\added{However, the association between transient flux dips and binarity is
-not iron-clad.  Firstly, multiple photometric periods do not necessarily
-imply binarity. For instance, circumstellar material could exist
-around single stars inside, at, or wide of co-rotation.  Additionally, } low-mass pre-main-sequence stars
-\deleted{however} have been shown to \added{have} companion fractions
-up to twice as high in dispersed clusters such as Upper~Sco and Taurus
+selection bias against high mass ratio binaries, in which the second
+star would be too faint to detect a second period.}
+
+\added{However, the association between transient flux dips and
+binarity is not iron-clad.  Firstly, multiple photometric periods do
+not necessarily imply binarity. For instance, circumstellar material
+could exist around single stars inside, at, or wide of co-rotation.
+Additionally, } low-mass pre-main-sequence stars \deleted{however}
+have been shown to \added{have} companion fractions up to twice as
+high in dispersed clusters such as Upper~Sco and Taurus
 \citep{kraus_mapping_2008,kraus_mapping_2011}.  A high-resolution
@@
 Based on the lack of an infrared excess seen by \citet{yu_tests_2015},
-the primordial \deleted{gas} disks of both stars in \ptfo\ seem to be have been
-depleted \added{of hot dust within the inner $\sim$1$\,$AU}.  This is consistent with the $8.5\pm1.2\,{\rm Myr}$ age of
-the 25$\,$Ori-1 group, and the rapid rate at which stars \deleted{show} lose
-their disks between 1 and 10$\,$Myr \citep[{\it
+the primordial \deleted{gas} disks of both stars in \ptfo\ seem to be
+have been depleted \added{of hot dust within the inner $\sim$1$\,$AU}.
+This is consistent with the $8.5\pm1.2\,{\rm Myr}$ age of the
+25$\,$Ori-1 group, and the rapid rate at which stars \deleted{show}
+lose their disks between 1 and 10$\,$Myr \citep[{\it
 e.g.},][]{hernandez_spitzer_ob1_2007}.  The stars are therefore
@@
 
-The key question is what causes the transient dips. This is
-an unsolved problem not only for \ptfo\ but also for the emerging
-class of similar young and rapidly rotating M-dwarfs.  Many possible
+The key question is what causes the transient dips. This is an
+unsolved problem not only for \ptfo\ but also for the emerging class
+of similar young and rapidly rotating M-dwarfs.  Many possible
 explanations were discussed by \citet{rebull_rotation_2016},
@@
 {\it (v)} high-latitude starspots;
-or
-{\it (vi)} dust clouds of \replaced{plausible composition}{particular compositions \citep[][Figure~10]{zhan_complex_2019}}.
-We also view the possibility of {\it (vii)} tidally disrupted
-planetary or cometary material to be implausible, given the
-synchronicity between dip and rotation periods seen across many
-systems.
-
-The explanations that \replaced{are not yet}{have not yet been} ruled out include {\it (i)}
-transiting clumps of \added{dust or }gas at the Keplerian co-rotation radius; {\it
-(ii)} transits of enshrouded protoplanets; {\it (iii)} occultations of
-\replaced{starspots}{hot and cold spots on the stellar surface} by an optically thick disk.  The first and last explanations
-have added appeal because they are flexible enough to explain not only
-the transient and persistent-dip M-dwarfs, but also the ``scallop
-shell'' M-dwarfs \citep{stauffer_orbiting_2017}.  \replaced{Despite this appeal,
-the possibility of distinct mechanisms explaining these distinct
-variability classes remains open.}{
-The model of hot and coldspot occultations by an optically thick disk
-\citep{zhan_complex_2019} is a qualitative inverse of the torus of
-circumstellar material: the difference is whether to place the
-``blobs'' that provide the needed brightness contrast on the surface
-of the star, or at the co-rotation radius.  In the spot occultation
-model, the observing geometry would require an inner disk to be
-projected into a narrow band across the stellar surface, in order to
-produce variations with sufficiently short timescales.  The outer
-disk would need either to not be present, or to be misaligned enough,
-and narrow enough, that it did not fall along the line of sight. These
-combined requirements seem geometrically contrived.
-}
-
-\replaced{The evolution of \ptfo\ over}{Observations of \ptfo\ from} the past decade
-could offer additional hints.
-\deleted{Specifically, }\ptfo's transition between having none, one, and
-multiple dips per cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.  It strains the ``enshrouded
-protoplanet'' interpretation, because there are no known processes
-that cause a planet's orbital phase to jump.  \deleted{The dips would then need
-to be caused by material that was somehow disrupted from the planet,
-but somehow remained co-orbital for an extended duration. This seems
-implausible.}
-\added{
-The H$\alpha$
+or {\it (vi)} dust clouds of \replaced{plausible
+composition}{particular compositions
+\citep[][Figure~10]{zhan_complex_2019}}.  We also view the possibility
+of {\it (vii)} tidally disrupted planetary or cometary material to be
+implausible, given the synchronicity between dip and rotation periods
+seen across many systems.
+
+The explanations that \replaced{are not yet}{have not yet been} ruled
+out include {\it (i)} transiting clumps of \added{dust or }gas at the
+Keplerian co-rotation radius; {\it (ii)} transits of enshrouded
+protoplanets; {\it (iii)} occultations of \replaced{starspots}{hot and
+cold spots on the stellar surface} by an optically thick disk.  The
+first and last explanations have added appeal because they are
+flexible enough to explain not only the transient and persistent-dip
+M-dwarfs, but also the ``scallop shell'' M-dwarfs
+\citep{stauffer_orbiting_2017}.  \replaced{Despite this appeal, the
+possibility of distinct mechanisms explaining these distinct
+variability classes remains open.}{ The model of hot and coldspot
+occultations by an optically thick disk \citep{zhan_complex_2019} is a
+qualitative inverse of the torus of circumstellar material: the
+difference is whether to place the ``blobs'' that provide the needed
+brightness contrast on the surface of the star, or at the co-rotation
+radius.  In the spot occultation model, the observing geometry would
+require an inner disk to be projected into a narrow band across the
+stellar surface, in order to produce variations with sufficiently
+short timescales.  The outer disk would need either to not be present,
+or to be misaligned enough, and narrow enough, that it did not fall
+along the line of sight. These combined requirements seem
+geometrically contrived.  }
+
+\replaced{The evolution of \ptfo\ over}{Observations of \ptfo\ from}
+the past decade could offer additional hints.  \deleted{Specifically,
+}\ptfo's transition between having none, one, and multiple dips per
+cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.  It
+strains the ``enshrouded protoplanet'' interpretation, because there
+are no known processes that cause a planet's orbital phase to jump.
+\deleted{The dips would then need to be caused by material that was
+somehow disrupted from the planet, but somehow remained co-orbital for
+an extended duration. This seems implausible.} \added{ The H$\alpha$
 excess observed by \citet{johnskrull_h_2016} also seems like an
-important piece of evidence.
-The model with circumstellar optically thick hydrogen
-gas near the co-rotation radius readily explains this excess.
-}
+important piece of evidence.  The model with circumstellar optically
+thick hydrogen gas near the co-rotation radius readily explains this
+excess.  }
 
 \added{Out of the available explanations, we therefore consider the
-idea of a clumpy torus of optically thick dust or gas near the co-rotation radius
-to be most plausible.
-There is theoretical support for
-the idea that accretion disks subject to strong magnetic
-fields could enter stalled states and become trapped near co-rotation
-\citep{dangelo_accretion_2012}.
-The concept of an ``evaporating dust wall'' within which all
-dust evaporates \citep{millan-gabet_circumstellar_2007,dullemond_inner_2010}
-will likely be an important first-order consideration in
-physical models of this scenario.
-The additional question of whether the opacity in fact comes from dust 
-or partially-ionized gas ({\it e.g.}, via gas continuum, atomic lines, or perhaps
-TiO molecular opacity) remains a topic for future exploration.
-}
+idea of a clumpy torus of optically thick dust or gas near the
+co-rotation radius to be most plausible.  There is theoretical support
+for the idea that accretion disks subject to strong magnetic fields
+could enter stalled states and become trapped near co-rotation
+\citep{dangelo_accretion_2012}.  The concept of an ``evaporating dust
+wall'' within which all dust evaporates
+\citep{millan-gabet_circumstellar_2007,dullemond_inner_2010} will
+likely be an important first-order consideration in physical models of
+this scenario.  The additional question of whether the opacity in fact
+comes from dust or partially-ionized gas ({\it e.g.}, via gas
+continuum, atomic lines, or perhaps TiO molecular opacity) remains a
+topic for future exploration.  }
 

### 2020-06-12 paper/ms.tex
wordsmith

@@
 \citep{dangelo_accretion_2012}.  The concept of an ``evaporating dust
-wall'' within which all dust evaporates
+rim'' within which all dust evaporates
 \citep{millan-gabet_circumstellar_2007,dullemond_inner_2010} will
@@
 comes from dust or partially-ionized gas ({\it e.g.}, via gas
-continuum, atomic lines, or perhaps TiO molecular opacity) remains a
-topic for future exploration.  }
+continuum, atomic lines, or perhaps TiO molecular opacity) also
+remains a topic for future exploration.  }
 

### 2020-06-12 paper/ms.tex
wordsmith

@@
 
-%\documentclass[12pt,twocolumn,tighten]{aastex62}
-\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
+\documentclass[12pt,twocolumn,tighten]{aastex62}
+%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
 \usepackage{amsmath,amstext,amssymb}
@@
 stellar surface, in order to produce variations with sufficiently
-short timescales.  The outer disk would need either to not be present,
+short timescales.  The outer disk would need to either not be present,
 or to be misaligned enough, and narrow enough, that it did not fall
@@
 others had hoped,} understanding it and its analogs is a worthy
-problem\replaced{.}\added{that seems likely to provide insight into how gas
+problem\replaced{.}{ that seems likely to provide insight into how gas
 and dust disperse from young protostellar disks.} \replaced{It might even}{Understanding

### 2020-06-13 paper_review/response_20200605.txt
Update response_20200605.txt

@@
 We thank you for organizing the review of our paper, and also
-thank the referee for their helpful comments.  We have considered
+thank the referee for constructive criticism.  We have considered
 the referee's remarks carefully and revised our manuscript
@@
 > figures at a higher precision, and quoting higher precision in
-> units of hours seems atypical in a field in which days are more
-> typically seen.
+> units of hours seems atypical.
 
@@
 > sentence as the youngest known *transiting* hot Jupiter, and
-> added a footnote to the V830 Tau or TAP 26 results.
+> added a footnote to cite the V830 Tau or TAP 26 results.
 
@@
 ----------
-> We have changed the wording to err on the side of caution, and
-> qualify it as the first direct detection of Hα emission from a
-> close-in planet.
+> We followed the referee's suggestion by stating that PTFO 8-8695
+> would have been the first close-in planet for which Hα emission
+> was detected.
 
@@
 > We have changed the wording of the last sentence of the first
-> paragraph in the introduction to explicitly use the word
+> paragraph in the introduction to use the word
 > "precess", rather than "torque".
@@
 >
-> We primarily mention our use of the FFI lightcurves in Section
-> 2 because of the systematic jump mentioned at BJD 2458488.3,
+> Our discussion of the FFI lightcurves in Section 2 was motivated
+> by the discontinuity in the light curve at BJD 2458488.3,
 > which was eventually excluded from further analysis. We wanted
-> to be sure though that this jump was truly an instrumental
-> systematic, and not a software systematic. Checking lightcurves
-> from independent pipelines meets this goal.
+> to compare the FFI and 2m light ccurves to see whether this jump
+> was an instrumental systematic or a software systematic.
 >
@@
 distinction between "young cluster members" and "old field
-stars"." I suggest removing this. It's not adding anything and
+stars". I suggest removing this. It's not adding anything and
 sounds like the authors are making assumptions about the reader.
@@
 ----------
-> We've removed it -- thank you for the suggestion.
+> We appreciate this suggestion and have omitted this sentence.
 
@@
 ----------
-> While we see the referee's point, we prefer the narrative
-> structure of presenting the TESS data first, and getting into
-> considerations of stellar binarity, both resolved and
-> unresolved, afterwards.
->
-> The reason is that the multiple signals come from TESS data
-> processed at the ``default'' level. From a story-telling
-> perspective, why would we worry about binarity if we didn't
-> already know there were multiple photometric signals?
+> We considered re-organizing the paper as the referee suggested,
+> but ultimately decided against it.  We chose to present the
+> TESS data, and show the 2 signals, which raises the possibility
+> of binarity - which is then investigated further with Gaia.
+> The referee suggests showing the Gaia evidence for binarity, and then
+> showing the TESS data with 2 signals. To us, there does not seem
+> to be any advantage to inverting the order.
 
@@
 ----------
-> We've done as suggested, and replaced this sub-plot with RUWE
-> instead of reduced chi^2. The text in Section 4.3 and captions
-> have been updated accordingly, and chi^2 is now mostly ignored.
+> We appreciate this criticism. We've followed the referee's advice
+> and replaced this sub-plot with RUWE instead of reduced chi^2.
+> The text in Section 4.3 and captions have been updated accordingly.
 
@@
 ----------
-> We have modified the last paragraphs of Section 5.4 to reflect
-> an updated preference that we came to after submission, for the
+> This comment caused us to consider more carefully the merits
+> and weaknesses of the two hypotheses we discussed. 
+> We ended up with a stronger preference for
 > clumpy torus model over the spot occultation model. The reasons

### 2020-06-13 paper/ms.tex
Update ms.tex

@@
 possibility of distinct mechanisms explaining these distinct
-variability classes remains open.}{ The model of hot and coldspot
-occultations by an optically thick disk \citep{zhan_complex_2019} is a
-qualitative inverse of the torus of circumstellar material: the
-difference is whether to place the ``blobs'' that provide the needed
-brightness contrast on the surface of the star, or at the co-rotation
-radius.  In the spot occultation model, the observing geometry would
-require an inner disk to be projected into a narrow band across the
-stellar surface, in order to produce variations with sufficiently
-short timescales.  The outer disk would need to either not be present,
-or to be misaligned enough, and narrow enough, that it did not fall
-along the line of sight. These combined requirements seem
-geometrically contrived.  }
+variability classes remains open.}{ In the spot occultation scenario
+\citep{zhan_complex_2019},
+the ``blobs'' that provide the necessary
+brightness contrast are on the surface of the star, rather than at the co-rotation
+radius. The observing geometry would
+require an inner ring to be projected into a narrow band in front of the
+stellar disk, in order to produce variations with sufficiently
+short timescales. 
+% The outer disk would need to either not be present,
+% or to be misaligned enough, and narrow enough, that it did not fall
+% along the line of sight.
+% JNW Doesn't this same problem affect the "dusty torus" model?
+This geometric requirement seems too contrived to occur in even
+$\lesssim$1\% of young M dwarf systems. }
 
@@
 }\ptfo's transition between having none, one, and multiple dips per
-cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.  It
-strains the ``enshrouded protoplanet'' interpretation, because there
-are no known processes that cause a planet's orbital phase to jump.
+cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.
+%It
+%strains the ``enshrouded protoplanet'' interpretation, because there
+%are no known processes that cause a planet's orbital phase to jump.
+% JNW Doesn't seem worth beating on a dead horse.
 \deleted{The dips would then need to be caused by material that was

### 2020-06-14 paper_review/response_20200605.txt
paper & report cleanup

@@
 > sentence as the youngest known *transiting* hot Jupiter, and
-> added a footnote to cite the V830 Tau or TAP 26 results.
+> added a footnote to cite the V830 Tau and TAP 26 results.
 
@@
 > which was eventually excluded from further analysis. We wanted
-> to compare the FFI and 2m light ccurves to see whether this jump
+> to compare the FFI and 2m light curves to see whether this jump
 > was an instrumental systematic or a software systematic.
@@
 > explained more easily by dust/gas clumps than the
-> protoplanet?", we now refer D'Angelo and Spruit 2012 (see in
-> particular their Section 1), which discusses the possibility of
-> cyclic accretion, with periods of "stalled" mass accumulation
-> outside the corotation radius alternating with accreting
-> phases.
+> protoplanet?", we now reference D'Angelo and Spruit 2012 (see
+> in particular their Section 1), which discusses the possibility
+> of cyclic accretion, with periods of "stalled" mass
+> accumulation outside the corotation radius alternating with
+> accreting phases.
 

### 2020-06-14 paper/ms.tex
paper & report cleanup

@@
 variability classes remains open.}{ In the spot occultation scenario
-\citep{zhan_complex_2019},
-the ``blobs'' that provide the necessary
-brightness contrast are on the surface of the star, rather than at the co-rotation
-radius. The observing geometry would
-require an inner ring to be projected into a narrow band in front of the
-stellar disk, in order to produce variations with sufficiently
-short timescales. 
-% The outer disk would need to either not be present,
-% or to be misaligned enough, and narrow enough, that it did not fall
-% along the line of sight.
-% JNW Doesn't this same problem affect the "dusty torus" model?
-This geometric requirement seems too contrived to occur in even
-$\lesssim$1\% of young M dwarf systems. }
+\citep{zhan_complex_2019}, the ``blobs'' that provide the necessary
+brightness contrast are on the surface of the star, rather than at the
+co-rotation radius. The observing geometry would require an inner ring
+to be projected into a narrow band in front of the stellar disk, in
+order to produce variations with sufficiently short timescales.  The
+outer disk would also need to either not be present, or to be
+misaligned enough and insufficiently flared, that it did not fall
+along the line of sight.  This geometric requirement seems too
+contrived to occur in even $\lesssim$1\% of young M dwarf systems. }
 
@@
 cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.
-%It
-%strains the ``enshrouded protoplanet'' interpretation, because there
-%are no known processes that cause a planet's orbital phase to jump.
-% JNW Doesn't seem worth beating on a dead horse.
 \deleted{The dips would then need to be caused by material that was
@@
 \citep{millan-gabet_circumstellar_2007,dullemond_inner_2010} will
-likely be an important first-order consideration in physical models of
+likely also be an important first-order consideration in physical models of
 this scenario.  The additional question of whether the opacity in fact
@@
 The physical mechanism that explains the transient and persistent flux
-dips is unresolved. Our preferred explanation is that a clumpy torus of dust or gas is near the Keplerian co-rotation radius, and periodically occults the star.
-The \added{chromaticity, out-of-eclipse variability, and} jumping orbital phase \replaced{disfavored}{of \ptfo\ all disfavor} the explanation of an enshrouded,
-transiting protoplanet.  Though \ptfo\ may not be a planet,\deleted{ as we and
-others had hoped,} understanding it and its analogs is a worthy
-problem\replaced{.}{ that seems likely to provide insight into how gas
-and dust disperse from young protostellar disks.} \replaced{It might even}{Understanding
-this process would, in turn,} teach us about the birth environments of the
-majority of habitable-zone Earth-sized planets
-\citep{dressing_occurrence_2013}.
+dips is unresolved. Our preferred explanation is that a clumpy torus
+of dust or gas is near the Keplerian co-rotation radius, and
+periodically occults the star.  The \added{chromaticity,
+out-of-eclipse variability, and} jumping orbital phase
+\replaced{disfavored}{of \ptfo\ all disfavor} the explanation of an
+enshrouded, transiting protoplanet.  Though \ptfo\ may not be a
+planet,\deleted{ as we and others had hoped,} understanding it and its
+analogs is a worthy problem\replaced{.}{ that seems likely to provide
+insight into how gas and dust disperse from young protostellar disks.}
+\replaced{It might even}{Understanding this process would, in turn,}
+teach us about the birth environments of the majority of
+habitable-zone Earth-sized planets \citep{dressing_occurrence_2013}.
 

### 2020-06-15 paper/ms.tex
minor

@@
 \end{enumerate}
-The 10.76$\,$hr signal in \ptfo\ meets all of these criteria.  This is
-the first connection of \ptfo\ with this class of objects, likely
-because the TESS data enabled us to put the dips in the context of the
-asymmetric out-of-dip modulation.
+The 10.76$\,$hr signal in \ptfo\ meets all of these criteria.
+\replaced{This is the first connection of \ptfo\ with this class of
+objects, likely because the TESS data enabled us to put the dips in
+the context of the asymmetric out-of-dip modulation.} {While this was
+in fact already noted by \citet{stauffer_orbiting_2017}, the
+asymmetric out-of-dip modulation shown by the TESS data puts the
+connection on stronger footing.}
 

### 2020-06-16 paper/ms.tex
minor
made discussion better

--- 2020-06-14 paper/ms.tex
@@
 \begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
-  \item They are weak-lined T Tauri stars.
+  \item They are weak-lined T Tauri stars\added{ (under the
+    definition of {\it e.g.}, \citealt{fang_star_2009})}.
   \item The spectral type is M2 to M5 ({\it e.g.},
     \citealt{rebull_usco_2018},~Figure~20).
-  \item \replaced{The age is typically $\lesssim$ 100$\,$Myr.}{The
-    ages typically range between 5 and 100$\,$Myr.}\footnote{ At
+  \item The age is typically $\lesssim$ 100$\,$Myr.\footnote{ At
     present, the oldest observed ``scallops'' are in the Pleaides
@@
     meet the ``persistent dip'' classification.  If so, it is \deleted{also} the
-    oldest known. \added{The lower limit comes from the lack of detectable
-   WISE infrared excesses.}} 
+    oldest known. } 
   \item The light curves show shallow, angular dips, \deleted{usually}
@@
 \end{enumerate}
-The 10.76$\,$hr signal in \ptfo\ meets all of these criteria.  This is
-the first connection of \ptfo\ with this class of objects, likely
-because the TESS data enabled us to put the dips in the context of the
-asymmetric out-of-dip modulation.
+The 10.76$\,$hr signal in \ptfo\ meets all of these criteria.
+\replaced{This is the first connection of \ptfo\ with this class of
+objects, likely because the TESS data enabled us to put the dips in
+the context of the asymmetric out-of-dip modulation.} {While this was
+in fact mentioned by \citet{stauffer_orbiting_2017}, the
+asymmetric out-of-dip modulation shown by the TESS data solidifies the
+connection.}
 
@@
 could exist around single stars inside, at, or wide of co-rotation.
+If it did, it could eclipse with a period different from the stellar
+rotation.
 Additionally, } low-mass pre-main-sequence stars \deleted{however}
@@
 
+\explain{Note :Removed ``high-latitude accretion hotspots'' from the
+list in this paragraph}
 The key question is what causes the transient dips. This is an
@@
 {\it (iii)} eclipses of prominences;
-{\it (iv)} high-latitude accretion hotspots;
-{\it (v)} high-latitude starspots;
-or {\it (vi)} dust clouds of \replaced{plausible
+{\it (iv)} high-latitude starspots;
+or {\it (v)} dust clouds of \replaced{plausible
 composition}{particular compositions
 \citep[][Figure~10]{zhan_complex_2019}}.  We also view the possibility
-of {\it (vii)} tidally disrupted planetary or cometary material to be
+of {\it (vi)} tidally disrupted planetary or cometary material to be
 implausible, given the synchronicity between dip and rotation periods
-seen across many systems.
-
-The explanations that \replaced{are not yet}{have not yet been} ruled
-out include {\it (i)} transiting clumps of \added{dust or }gas at the
-Keplerian co-rotation radius; {\it (ii)} transits of enshrouded
-protoplanets; {\it (iii)} occultations of \replaced{starspots}{hot and
-cold spots on the stellar surface} by an optically thick disk.  The
-first and last explanations have added appeal because they are
-flexible enough to explain not only the transient and persistent-dip
-M-dwarfs, but also the ``scallop shell'' M-dwarfs
+seen across many systems.  \added{Finally, {\it (vii)} transits of an
+enshrouded protoplanet can be added to the list of disfavored
+explanations, based on the chromaticity, asymmetric phase curve, and
+light-curve changes seen in \ptfo.}
+
+\explain{Note: ``High-latitude accretion hotspots'' was moved to this
+paragraph.}
+The explanations that we view as not yet ruled out include {\it A)}
+transiting clumps of \added{dust or }gas at the Keplerian co-rotation
+radius; {\it B)} occultations of \replaced{starspots}{hotspots on the
+stellar surface} by an optically thick disk \citep{zhan_complex_2019};
+and {\it C)} high-latitude accretion hotspots passing behind the star
+\citep{yu_tests_2015}.  \replaced{The evolution of \ptfo\
+over}{Observations of \ptfo\ from} the past decade may offer some
+hints.  \deleted{Specifically, }\ptfo's transition between having
+none, one, and multiple dips per cycle seems important
+\added{(Figure~\ref{fig:o_minus_c})}.  \deleted{The dips would then
+need to be caused by material that was somehow disrupted from the
+planet, but somehow remained co-orbital for an extended duration. This
+seems implausible.} \added{ The H$\alpha$ excess observed by
+\citet{johnskrull_h_2016} also seems like an important piece of
+evidence.  }
+
+\explain{Note: This paragraph is newly added, to support our view of the
+high-latitude accretion hotspot being plausible for \ptfo.}
+The high-latitude accretion hotspot scenario seems qualitatively
+capable of explaining most of the characteristics of \ptfo, including
+the fact that its H$\alpha$ equivalent width of 11.4$\,$\AA\
+\citep{van_eyken_ptf_2012} places it near the ``classical/weak'' T
+Tauri division \citep{briceno_cidaII_2019}, and the fact that its
+excess H$\alpha$ component varies in velocity between $-200$ and
+$+200$ km$\,$s$^{-1}$ \citep{johnskrull_h_2016}.  The lack of a
+NIR-excess could be problematic, but does not eliminate the
+possibility that a tenuous amount of dust or gas remains near the
+star, and is able to accrete onto it.  We view the main problem with
+this explanation to be the likelihood that it explains the population
+of transient flux dips.  RIK-210 in particular has dips too deep
+($\sim$20\%) and too asymmetric to be explained by occultations of a
+hot spot \citep{david_transient_2017}.
+
+\explain{Note: This paragraph has been expanded to clarify the
+possibility of a hotspot being occulted by a disk.}
+Explanations {\it A} and {\it B} (circumstellar material at the
+co-rotation radius, and occultations of spots by a disk) have the
+appeal that they are flexible enough to explain not only the transient
+and persistent-dip stars, but also the ``scallop shells''
 \citep{stauffer_orbiting_2017}.  \replaced{Despite this appeal, the
@@
 
-\replaced{The evolution of \ptfo\ over}{Observations of \ptfo\ from}
-the past decade could offer additional hints.  \deleted{Specifically,
-}\ptfo's transition between having none, one, and multiple dips per
-cycle seems important \added{(Figure~\ref{fig:o_minus_c})}.
-\deleted{The dips would then need to be caused by material that was
-somehow disrupted from the planet, but somehow remained co-orbital for
-an extended duration. This seems implausible.} \added{ The H$\alpha$
-excess observed by \citet{johnskrull_h_2016} also seems like an
-important piece of evidence.  The model with circumstellar optically
-thick hydrogen gas near the co-rotation radius readily explains this
-excess.  }
-
 \added{Out of the available explanations, we therefore consider the
-idea of a clumpy torus of optically thick dust or gas near the
-co-rotation radius to be most plausible.  There is theoretical support
-for the idea that accretion disks subject to strong magnetic fields
-could enter stalled states and become trapped near co-rotation
-\citep{dangelo_accretion_2012}.  The concept of an ``evaporating dust
-rim'' within which all dust evaporates
-\citep{millan-gabet_circumstellar_2007,dullemond_inner_2010} will
-likely also be an important first-order consideration in physical models of
-this scenario.  The additional question of whether the opacity in fact
-comes from dust or partially-ionized gas ({\it e.g.}, via gas
-continuum, atomic lines, or perhaps TiO molecular opacity) also
-remains a topic for future exploration.  }
+clumpy torus of optically thick dust or gas near the co-rotation
+radius to be most plausible.  The partially ionized circumstellar
+hydrogen would readily explain the excess H$\alpha$ emission, and
+changes in the clumping geometry would lead to changes in the observed
+eclipse shapes.  The long-term stability, punctuated by changes in dip
+phases, might be understood as an accretion disk subject to strong
+magnetic fields going through ``stalled'' states in which material is
+trapped near co-rotation \citep{dangelo_accretion_2012}.  Physical
+models of this scenario may need to consider such states, in addition
+to dissociation of the dust population
+\citep{millan-gabet_circumstellar_2007,dullemond_inner_2010}.
+Finally, the question of whether the opacity in fact comes from dust
+or partially-ionized gas ({\it e.g.}, via gas continuum, atomic lines,
+or perhaps TiO molecular opacity) remains an additional issue for
+future exploration.  }
 

### 2020-06-16 paper/ms.tex
update \explain environment calls

@@
 
-\explain{All paragraphs in this section have been heavily reworded to
+\explain{NOTE: All paragraphs in this section have been heavily reworded to
 emphasize RUWE over $\chi_{\rm red}^2$, as requested by the referee.}
@@
 
-\explain{Note: Citations to Parks et al 2014 and Stauffer et al 2015
+\explain{NOTE: Citations to Parks et al 2014 and Stauffer et al 2015
 were added to the paragraph below} When physical explanations are not
@@
 
-\explain{Note: added citation to Bodman et al 2017 in this paragraph.}
+\explain{NOTE: added citation to Bodman et al 2017 in this paragraph.}
 These light curves have been phenomenologically classified as
@@
 
-\explain{Note :Removed ``high-latitude accretion hotspots'' from the
+\explain{NOTE: Removed ``high-latitude accretion hotspots'' from the
 list in this paragraph}
@@
 
-\explain{Note: ``High-latitude accretion hotspots'' was moved to this
+\explain{NOTE: ``High-latitude accretion hotspots'' was moved to this
 paragraph.}
@@
 
-\explain{Note: This paragraph is newly added, to support our view of the
+\explain{NOTE: This paragraph is newly added, to support our view of the
 high-latitude accretion hotspot being plausible for \ptfo.}
@@
 
-\explain{Note: This paragraph has been expanded to clarify the
+\explain{NOTE: This paragraph has been expanded to clarify the
 possibility of a hotspot being occulted by a disk.}
@@
 
-\explain{Updated this paragraph to remove spot occultations by a disk as
+\explain{NOTE: Updated this paragraph to remove spot occultations by a disk as
 a ``preferred explanation''.}

### 2020-06-16 paper/ms.tex
Update ms.tex

@@
 the context of the asymmetric out-of-dip modulation.} {While this was
-in fact mentioned by \citet{stauffer_orbiting_2017}, the
+mentioned by \citet{stauffer_orbiting_2017}, the
 asymmetric out-of-dip modulation shown by the TESS data solidifies the
@@
 discovered in $\rho$~Oph and Upper~Sco show two Lomb-Scargle periods,
-and so \replaced{are apparently}{could be} binaries
+\replaced{and so are apparently}{a possible sign of binarity} \removed{binaries}
 \citep[][Table~1]{stauffer_orbiting_2017}.  This is higher than the
 main-sequence companion fraction of ${\rm
-CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$\added{, and so it
+CF}_{0.1\text{-}0.5\,M_\odot}^{\rm MS} = 33 \pm 5\%$\added{, which
 could be an indication that the transient flux dip phenomenon is
@@
 not necessarily imply binarity. For instance, circumstellar material
-could exist around single stars inside, at, or wide of co-rotation.
-If it did, it could eclipse with a period different from the stellar
-rotation.
+could exist around single stars at multiple orbital distances, and not
+just at the co-rotation radius.
+This could cause eclipses with periods different from the stellar
+rotation period.
 Additionally, } low-mass pre-main-sequence stars \deleted{however}
@@
 \citep{yu_tests_2015}.  \replaced{The evolution of \ptfo\
-over}{Observations of \ptfo\ from} the past decade may offer some
+over}{Observations of \ptfo\ over} the past decade may offer some
 hints.  \deleted{Specifically, }\ptfo's transition between having
-none, one, and multiple dips per cycle seems important
+no dips, one dip, and multiple dips per cycle seems important
 \added{(Figure~\ref{fig:o_minus_c})}.  \deleted{The dips would then
@@
 high-latitude accretion hotspot being plausible for \ptfo.}
-The high-latitude accretion hotspot scenario seems qualitatively
-capable of explaining most of the characteristics of \ptfo, including
-the fact that its H$\alpha$ equivalent width of 11.4$\,$\AA\
-\citep{van_eyken_ptf_2012} places it near the ``classical/weak'' T
-Tauri division \citep{briceno_cidaII_2019}, and the fact that its
-excess H$\alpha$ component varies in velocity between $-200$ and
-$+200$ km$\,$s$^{-1}$ \citep{johnskrull_h_2016}.  The lack of a
-NIR-excess could be problematic, but does not eliminate the
-possibility that a tenuous amount of dust or gas remains near the
+The high-latitude accretion hotspot scenario seems 
+capable of explaining most of the qualitative characteristics of \ptfo.
+Some degree of accretion onto the star seems likely, given 
+the H$\alpha$ equivalent width of 11.4$\,$\AA\
+\citep{van_eyken_ptf_2012}, which is near the boundary separating ``classical'' and ''weak-lined'' T
+Tauri stars \citep{briceno_cidaII_2019}, and the high-velocity
+components of the observed H$\alpha$ line profile ($\pm\,200$\,km$\,$s$^{-1}$ \citep{johnskrull_h_2016}.
+The non-detection of any NIR excess could be problematic, but does not eliminate the
+possibility that a tenuous amount of dust and gas remains near the
 star, and is able to accrete onto it.  We view the main problem with
-this explanation to be the likelihood that it explains the population
-of transient flux dips.  RIK-210 in particular has dips too deep
+this explanation seems to require rare circumstances,
+making it unlikely to be correct for the entire population
+of transient flux dips.  RIK-210 in particular has dips that seem too deep
 ($\sim$20\%) and too asymmetric to be explained by occultations of a
@@
 outer disk would also need to either not be present, or to be
-misaligned enough and insufficiently flared, that it did not fall
-along the line of sight.  This geometric requirement seems too
+misaligned enough and insufficiently flared, that it did not obstruct
+the line of sight to the star.  This geometric requirement seems too
 contrived to occur in even $\lesssim$1\% of young M dwarf systems. }
 
-\added{Out of the available explanations, we therefore consider the
+\added{Thus, while several of the proferred explanations deserve further
+scrutiny, we consider the
 clumpy torus of optically thick dust or gas near the co-rotation

### 2020-06-17 paper/ms.tex
minor edits

@@
 \received{May 15, 2020}
-\revised{\today}
+\revised{June 16, 2020}
 \accepted{---}
@@
 discovered in $\rho$~Oph and Upper~Sco show two Lomb-Scargle periods,
-\replaced{and so are apparently}{a possible sign of binarity} \removed{binaries}
+\replaced{and so are apparently binaries}{a possible sign of binarity}
 \citep[][Table~1]{stauffer_orbiting_2017}.  This is higher than the
@@
 could exist around single stars at multiple orbital distances, and not
-just at the co-rotation radius.
-This could cause eclipses with periods different from the stellar
-rotation period.
-Additionally, } low-mass pre-main-sequence stars \deleted{however}
-have been shown to \added{have} companion fractions up to twice as
-high in dispersed clusters such as Upper~Sco and Taurus
+just at the co-rotation radius.  This could cause eclipses with
+periods different from the stellar rotation period.  Additionally, }
+low-mass pre-main-sequence stars \deleted{however} have been shown to
+\added{have} companion fractions up to twice as high in dispersed
+clusters such as Upper~Sco and Taurus
 \citep{kraus_mapping_2008,kraus_mapping_2011}.  A high-resolution
@@
 high-latitude accretion hotspot being plausible for \ptfo.}
-The high-latitude accretion hotspot scenario seems 
-capable of explaining most of the qualitative characteristics of \ptfo.
-Some degree of accretion onto the star seems likely, given 
-the H$\alpha$ equivalent width of 11.4$\,$\AA\
-\citep{van_eyken_ptf_2012}, which is near the boundary separating ``classical'' and ''weak-lined'' T
-Tauri stars \citep{briceno_cidaII_2019}, and the high-velocity
-components of the observed H$\alpha$ line profile ($\pm\,200$\,km$\,$s$^{-1}$ \citep{johnskrull_h_2016}.
-The non-detection of any NIR excess could be problematic, but does not eliminate the
-possibility that a tenuous amount of dust and gas remains near the
-star, and is able to accrete onto it.  We view the main problem with
-this explanation seems to require rare circumstances,
-making it unlikely to be correct for the entire population
-of transient flux dips.  RIK-210 in particular has dips that seem too deep
-($\sim$20\%) and too asymmetric to be explained by occultations of a
-hot spot \citep{david_transient_2017}.
+The high-latitude accretion hotspot scenario seems capable of
+explaining most of the qualitative characteristics of \ptfo.  Some
+degree of accretion onto the star seems likely, given 
+the high-velocity ($\pm\,200$\,km$\,$s$^{-1}$) components of
+the observed H$\alpha$ line profile
+\citep{johnskrull_h_2016},
+as well as the H$\alpha$
+equivalent width of 11.4$\,$\AA\ \citep{van_eyken_ptf_2012}, which is
+near the boundary separating ``classical'' and ''weak-lined'' T Tauri
+stars \citep{briceno_cidaII_2019}.  The non-detection of any NIR excess could
+be problematic, but does not eliminate the possibility that a tenuous
+amount of dust and gas remains near the star, and is able to accrete
+onto it.  The main problem with this explanation is that it seems to
+require rare circumstances, making it unlikely to be correct for the
+entire population of transient flux dips.  RIK-210 in particular has
+dips that seem too deep ($\sim$20\%) and too asymmetric to be
+explained by an accretion hot spot \citep{david_transient_2017}.
 
@@
 
-\added{Thus, while several of the proferred explanations deserve further
-scrutiny, we consider the
-clumpy torus of optically thick dust or gas near the co-rotation
-radius to be most plausible.  The partially ionized circumstellar
-hydrogen would readily explain the excess H$\alpha$ emission, and
-changes in the clumping geometry would lead to changes in the observed
-eclipse shapes.  The long-term stability, punctuated by changes in dip
-phases, might be understood as an accretion disk subject to strong
-magnetic fields going through ``stalled'' states in which material is
-trapped near co-rotation \citep{dangelo_accretion_2012}.  Physical
-models of this scenario may need to consider such states, in addition
-to dissociation of the dust population
+\added{Thus, while several of the proferred explanations deserve
+further scrutiny, we consider the clumpy torus of optically thick dust
+or gas near the co-rotation radius to be most plausible.  The
+partially ionized circumstellar hydrogen would readily explain the
+excess H$\alpha$ emission, and changes in the clumping geometry would
+lead to changes in the observed eclipse shapes.  The long-term
+stability, punctuated by changes in dip phases, might be understood as
+an accretion disk subject to strong magnetic fields going through
+``stalled'' states in which material is trapped near co-rotation
+\citep{dangelo_accretion_2012}.  Physical models of this scenario may
+need to consider such states, in addition to dissociation of the dust
+population
 \citep{millan-gabet_circumstellar_2007,dullemond_inner_2010}.

### 2020-06-19 paper/ms.tex
minor correction

@@
 
-We wish \ptfob\ were a planet. It would be quite exceptional. It would
-be the youngest known \added{transiting\footnote{Two younger hot
-Jupiters have been reported around V830 Tau and TAP 26
-\citep{donati_hot_2016,donati_hot_2017,yu_hot_2017}.}} hot Jupiter
-\citep{van_eyken_ptf_2012}, orbiting a T Tauri star in the Orion-OB1a
-cluster.  It would have the shortest orbital period of any hot
-Jupiter.  With such a short period, it would probably be filling its
-Roche lobe, and actively losing mass to its host star.  Not only that,
-but the rapidly-rotating host star is probably oblate enough to
-\replaced{torque the planet's orbit into and out of the transiting
-configuration}{cause the planet's orbit to precess into and out of the
-transiting configuration} on a timescale of years
+We wish \ptfob\ were a planet. It would be quite exceptional. It would be the
+youngest known \added{transiting\footnote{Younger hot Jupiters have been
+reported around V830 Tau, TAP 26, and CI Tau
+\citep{johns-krull_CI_Tau_candidate_2016,donati_hot_2016,donati_hot_2017,yu_hot_2017,flagg_co_2019}.}}
+hot Jupiter \citep{van_eyken_ptf_2012}, orbiting a T Tauri star in the
+Orion-OB1a cluster.  It would have the shortest orbital period of any hot
+Jupiter.  With such a short period, it would probably be filling its Roche
+lobe, and actively losing mass to its host star.  Not only that, but the
+rapidly-rotating host star is probably oblate enough to \replaced{torque the
+planet's orbit into and out of the transiting configuration}{cause the planet's
+orbit to precess into and out of the transiting configuration} on a timescale
+of years
 \citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.

### 2020-06-27 paper/ms.tex
minor proof correction

@@
 reported around V830 Tau, TAP 26, and CI Tau
-\citep{johns-krull_CI_Tau_candidate_2016,donati_hot_2016,donati_hot_2017,yu_hot_2017,flagg_co_2019}.}}
+\citep{johns-krull_CI_Tau_candidate_2016,donati_hot_2016,donati_hot_2017,yu_hot_2017,biddle_k2_2018,flagg_co_2019}.}}
 hot Jupiter \citep{van_eyken_ptf_2012}, orbiting a T Tauri star in the
